### 2018-10-04 ehp_draft.tex
Initial commit

\documentclass[fleqn,10pt,lineno]{olplainarticle}
% Use option lineno for line numbers

\graphicspath{{./figures/}}

\usepackage[implicit=false]{hyperref}

\title{Assessing United States county-level exposure for tropical cyclone impact
research}

\author[a,1]{G. Brooke Anderson} 
\author[a,b]{Joshua Ferreri}
\author[c]{Mohammad Al-Hamdan} 
\author[c]{William Crosson} 
\author[d]{Andrea Schumacher} 
\author[e]{Seth Guikema} 
\author[f]{Steven Quiring} 
\author[g]{Dirk Eddelbuettel} 
\author[a]{Meilin Yan} 
\author[h]{Roger D. Peng}

\affil[a]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[b]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
\affil[c]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[d]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[e]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[f]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[g]{Debian and R Projects; Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[h]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}

\keywords{Hurricanes, Tropical cyclones, Disaster impacts, Exposure assessment}

\begin{abstract} Hurricanes and other tropical cyclones bring severe impacts to
U.S. communities. These impacts can result from a variety of storm-related
hazards, including extreme wind, rain, flooding, and tornadoes. Impact studies
vary widely in how they classify exposure to tropical cyclones, using various
hazard-based metrics and, in some cases, using distance from the storm as a
surrogate for exposure to storm-related hazards. Here we measure county-level
exposure to tropical cyclones in the United States based on distance from the
storm, maximum sustained wind, rainfall, flooding, and tornadoes for all
land-falling or near-land Atlantic basin storms for 1988--2015. We show that the
locations identified as storm-exposed varied substantially when switching among
these metrics. For example, most wind-based storm exposures were limited to
southern counties near the coast, while flood- and rain-based exposures often
extended to inland and northern counties. We also show that distance to the
storm served as, at best, a moderate, and often a poor, surrogate in identifying
exposure to storm-related wind, rain, floods, or tornadoes. Therefore, when
impact studies use distance as a surrogate for tropical cyclone exposures or use
one hazard-based metric (e.g., wind-based) when the impact is partly or fully
caused by a different storm hazard, the analysis will be prone to exposure
misclassification, which can mask true associations, even strong associations.
To facilitate future research, we make this multi-hazard storm exposure data
available through open-source software.  \end{abstract}

\begin{document}

\flushbottom 
\maketitle 
\thispagestyle{empty}

\section*{Introduction}

Tropical cyclones, including hurricanes, can have disasterous impacts in the
United States (U.S.). A tropical cyclone's high winds can bring health risks
and property damage caused by structural damage of houses and other buildings,
falling trees, and wind-borne debris \citep{rappaport2000} and 
cause power outages \citep{liu2005, han2009}, introducing a number of
threats to human and ecological health, including water quality
risks if the outage affects wastewater treatment plants \citep{mallin2006}.
Other risks can exist without severe wind; for example, one study found that
most of the direct hurricane-related deaths in the U.S. between 1970 and 1999
occurred in cases when wind was below hurricane strength, including for
Tropical Storms Charley in 1998 and Alberto in 1994 \citep{rappaport2000}.
Tropical cyclones can produce excessive rain, especially in certain
topographies (e.g., near mountains), so counties well inland sometimes
experience more extreme rain than coastal counties. Flood risks from tropical
cyclones can result from this rain, although the two risks are not perfectly
correlated \citep{chen2015}. In the U.S., over half of hurricane-related direct
deaths from 1970 to 1999 from Atlantic basin storms were a result of freshwater
flooding \citep{rappaport2000}. Flooding can also degrade water quality
\citep{mallin2006}, which can threaten both human and ecological health.

To measure tropical cyclone exposure for either investigations of exposure
patterns or for estimation of storm impacts, including health impacts, studies
have varied widely in how they determine storm exposure (some examples shown in
Fig. S1). While many studies have measured exposure based on specific hazards
of the storm (e.g., wind, rain), many studies have used distance from the
storm's central track as a surrogate in identifying exposure to the hazards of
a storm (Fig. S1) (e.g., \citep{czajkowski2011, tansel2010, kinney2008,
caillouet2008increase}). Distance from the storm's central track can be easily
measured using widely-available hurricane tracking data. There are, however, a
number of limitations to using distance to assess exposure to tropical
cyclones, and there are reasons to suspect that exposure classifications based
on other storm hazards may disagree with distance-based classifications.
Tropical cyclones vary dramatically in size: U.S. storms have been observed
with radii to maximum winds as small as 20 kilometers and as large as 200
kilometers \citep{mallin2006, quiring2011variations}. While a number of storm
hazards are strongly associated with distance from the storm's center (e.g.,
wind and, at the coast, storm surge and waves \citep{rappaport2000, kruk2010}),
other hazards like dangerous rain, floods, and tornadoes can occur well away
from the storm's central track \citep{rappaport2000, atallah2007, moore2012}.
For example, fatal tropical cyclone tornadoes, which were linked to over 300
deaths in the U.S.  between 1995 and 2009, most often occur 200--500 kilometers
from the storm's center \citep{moore2012}. Further, distance-based exposure
metrics that use buffers tend to use an equal buffer distance on each side of
the storm track (e.g., \citep{czajkowski2011, grabich2015, grabich2016,
zandbergen2009, tansel2010}), but the forces of a storm tend to be distributed
around the center in a non-symmetrical way. For example, extreme winds are more
common to the right of the track where counter-clockwise cyclonic winds move in
concert with the storm's forward motion \citep{halverson2015}, and the fatal
tornadoes associated with U.S. tropical cyclones between 1995 and 2009 occurred
almost exclusively to the right of the storm's track, mostly in the right-front
quandrant (northeast) of the storm \citep{moore2012}. Rain, conversely, is
often heaviest to the left of the storm's track, especially when the storm
interacts with other weather systems \citep{atallah2003, atallah2007,
zhu2013variations} or undergoes an extratropical transition
\citep{elsberry2002}.

The use of different exposure metrics across studies hampers the development of
a scientific consensus on storm-related risks, as differences observed between
studies could result from differences in exposure assessment.  Further, if a
study misclassifies exposure to the storm hazard or hazards that cause an
impact, estimates of storm risks or impacts will be biased, making it hard to
identify true associations. These concerns are particularly important if there
are large differences in the locations that are determined to be storm-exposed
based on different metrics.  To investigate this, we measured county-level
exposure to tropical cyclones in all eastern United States counties using five
metrics (Table \ref{tab:exposuremetrics}): (1) distance to storm track; (2)
maximum sustained wind speed; (3) cumulative rainfall; (4) flood events; and
(5) tornado events.  We assessed exposure at the county level since data on
many potential impacts are available at county-level aggregations (e.g., direct
hurricane-related deaths \citep{czajkowski2011}; birth outcomes
\citep{grabich2015, grabich2016}; autism prevalence \citep{kinney2008}) and
since decisions and policies to prepare for, and respond to, storms are often
undertaken at the county level \citep{zandbergen2009, rappaport2000}. We
measured how well exposure classification agreed across these five metrics in
terms of classifying individual counties as exposed to a storm.  Finally, to
make this hazard-specific tropical cyclone exposure data available to other
researchers, we published open source software for county-based hurricane
exposure assessment in U.S. counties \citep{hurricaneexposure}.

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{0.9cm}p{2.5cm}p{9cm}} 
Metric & Available & Criteria for exposure \\ \midrule 
Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\ 
Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
Tornado & 1996--2015 & Tornado event listed NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}

\section*{Methods and Materials}
,
All counties in the eastern half of the U.S. (e.g., Fig.
\ref{fig:averageexposure}), as of the 2010 U.S. Decennial Census, were included
in the study. The study covered all tracked storms in the revised Atlantic
hurricane database (HURDAT2 \citep{landsea2013}) and that came within 250
kilometers of at least one study county between 1988 and 2015 (Fig. S2). HURDAT2
is a post-storm assessment of each storm conducted by the U.S. National
Hurricane Center (NHC) and incorporates data from a variety of sources,
including satellite data and, when available, aircraft reconnaisance data
\citep{landsea2013, jarvinen1988}. These data typically give measurements of
storm center location at 6-hour intervals, at synoptic times (i.e., 6:00 am,
12:00 pm, 6:00 pm, and 12:00 am UTC); some landfalling storms have an additional
observation at the time of landfall \citep{landsea2013}.

\subsection*{Distance-based exposure metric.}

For the rain-based exposure metric, cumulative rainfall was estimated for the
period from two days before to one day after the date of a storm's closest
approach to a county. Daily county-level rainfall was determined by aggregating
hourly, 1/8 degree gridded reanalysis data from the North American Land Data
Assimilation System Phase 2 (NLDAS-2) precipitation data files
\citep{rui2013nldas}. The NLDAS-2 data integrate satellite-based and land-based
monitoring and applies a land-surface model to create a reanalysis dataset that
is spatially and temporally complete across the continental U.S.
\citep{rui2013nldas, alhamdan2014environmental}. These data have been used in
previous research to investigate tropical cyclones (e.g., \citep{chen2015}). To
aggregate to county-level, hourly data at each grid point were summed to create
a daily sum and these grid-point specific rainfall totals were then averaged
across all grid points within a county's 1990 U.S. Census boundaries to create a
county-level daily total \citep{alhamdan2014environmental, cdcwonder}. This
county-level precipitation data are publicly available through the U.S. Centers
for Disease Control's Wide-ranging Online Data for Epidemiological Research
(WONDER) database \citep{cdcwonder, alhamdan2014environmental}. The
rainfall-based exposure metric was only available for storms through 2011, the
period for which the reanalysis rainfall data aggregated to county-level daily
totals was available \citep{cdcwonder, alhamdan2014environmental}. Validation
analysis of this and other hazard-based exposure data is given in the Supporting
Information.

\subsection*{Wind-based exposure metric}

Ground-based observations of wind speed are problematic during storms, as
instruments often fail at high wind speed, while reanalysis data, often
available at hourly or higher resolution, can be too smooth to capture wind
extremes associated with a storm. Therefore, to create a dataset of county-level
sustained winds during historical storms, we modeled maximum sustained wind
speeds at each county's population mean center \citep{countycenters}. HURDAT2
Best Track data were interpolated from 6-hourly reported values to 15-minute
increments using natural cubic splines \citep{stormwindmodel}, and then
sustained wind speed was modeled for each county center at each 15-minute
increment using a double exponential wind speed model
\citep{willoughby2006parametric} to estimate maximum ground-level sustained wind
speed at each county center \citep{stormwindmodel}. Asymmetry in wind speeds
around the storm center was incorporated into the wind speed model
\citep{phadke2003modeling, stormwindmodel}. The maximum value of this modeled
wind speed was determined for each county as the maximum windspeed across the
15-minute incremented modeled values throughout the storm.

\subsection*{Flood- and tornado-based exposure metrics}

To identify flood- and tornado-based exposures to tropical cyclones in U.S.
counties, we used event listings from the National Oceanic and Atmospheric
Administration (NOAA)'s Storm Event Database \citep{stormevents}. For each
storm, we identified all events with event types related to flooding (``Flood",
``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and that occurred
in a county within 500 kilometers of the storm's track and within a five-day
window centered on the date of the storm's closest approach to the county
\citep{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado" events in
this database were reported by county Federal Information Processing Standard
(FIPS) code. ``Coastal Flood" events were reported by forecast zone; for these,
the event was matched to the appropriate county if possible
\citep{noaastormevents}. While this database has recorded storm data since 1950,
its coverage changed substantially in 1996 to cover a wider variety of storm
events \citep{stormevents}. We therefore only considered these metrics of
hurricane exposure for storms in 1996 and later.

\subsection*{Assessing agreement in exposure classification across exposure
metrics}

Two of the exposure metrics (flood- and tornado-based) were inherently binary;
for the other exposure metrics, each county was classified as exposed to a storm
based on whether the continuously-measured exposure metric exceeded a certain
threshold (Table \ref{tab:exposuremetrics}). We assessed county-level tropical
cyclone exposure in the eastern U.S. based on each exposure metrics (Table
\ref{tab:exposuremetrics}) for all land-falling or near-land Atlantic basin
storms. Depending on available exposure data, this assessment included some or
all of the period from 1988 to 2015, a period with 136 tropical cyclones that
made U.S. landfall or passed within 250 kilometers (155 miles) of at least one
U.S. county (Fig. S2).

We then measured agreement between exposure metrics in the classification of a
county as exposed or unexposed to a specific storm. For this measurement, we
calculated the within-storm Jaccard index \citep{jaccard1901distribution,
jaccard1908nouvelles} between every pair-wise combination of hazard metrics.
The Jaccard index ($J_s$) measures similarity between two metrics ($X_{1,s}$
and $X_{2,s}$) for storm $s$ as the proportion of counties in which both of the
metrics classify the county as exposed out of all counties classified as
exposed by at least one of the metrics:

\begin{equation} 
J_s = \frac{X_{1,s} \cap X_{2,s}}{X_{1,s} \cup X_{2,s}}
\end{equation}

This metric can range from 0, in the case of no overlap between the counties
classified as ``exposed" based on the two metrics, to 1, in the case that the
two metrics classify exactly the same counties as ``exposed" to the storm.

\section*{Results}

\begin{figure*}%[tbhp] 
\centering 
\includegraphics[width=16cm]{ivanonly}
\caption{Counties classified as exposed to Hurricane Ivan (2004) under each
exposure metric (Table 1). The red line shows the track of Hurricane Ivan 
based on the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
Similar maps for other large-extent storms are given in Fig. S5.}
\label{fig:ivanexposure} 
\end{figure*}

\subsection*{Patterns in tropical cyclone exposures in eastern U.S. counties}

Table [x] provides summary statistics on ...  Table [x] includes some measures
of the number of counties exposed to the storm based on each exposure metric
(Tab. [x]) Median storm extent was largest when exposure was classified based
on distance (median storm extent: 62 counties exposed). The measured extent of
storms also tended to be higher when using exposure metrics based on rain
(median: 21 counties exposed) or wind (median: 26 counties exposed) versus
based on floods (median: 9 counties exposed) or tornadoes (median: 1 county
exposed).  For every metric except the tornado-based metric, we identified at
least one storm with over 300 counties exposed. However, the largest-extent
storm varied by metric (e.g., Beryl [1994] exposed the most counties based on
distance, Frances [2004] based on rain, and Ike [2008] based on wind).

\subsection*{County-specific exposure classification}

\begin{figure}%[tbhp] 
\centering 
\includegraphics[width = 0.7\linewidth]{jaccard_heatmap} 
\caption{Heatmap of Jaccard index values for
specific exposure metric pairs within storms. Only storms between 1996 and 2011,
and for which at least 250 counties were exposed based on at least one metric,
are included. The color of each cell within the main heatmap indicates the value
of the Jaccard index (proportion of counties classified as exposed by both
metrics out of storms classified as exposed by either metric) for a given pair
of metrics for a given storm. Storms are displayed within clusters that have
similar patterns in county-level exposure agreement for metric pairs, based on
hierarchical clustering using the complete link method
\citep{murtagh2012algorithms} (i.e., storms in the same cluster tend to have
similar patterns for the pairwise strength of agreement among metrics); columns
are also ordered based on hierarchical clustering. The colors to the right of
the main heatmap for each storm indicate the total number of counties classified
as exposed to the storm by any of the five metrics, providing an estimate of
storm extent. Maps are available showing the counties identified as exposed
under each of five metrics for the widest-extent storm in each cluster:
Hurricane Ivan (2004) (Fig. \ref{fig:ivanexposure}) and Hurricanes Floyd (1999),
Lee (2011), Cindy (2005), and Katrina (2005) (Fig. S5).} 
\label{fig:jaccard}
\end{figure}

We found that switching among these metrics typically resulted in large
differences in which counties were identified as exposed to a particular storm.
Fig. \ref{fig:ivanexposure} gives an example for Hurricane Ivan (2004), a
record-breaking storm in terms of its duration as a major hurricane
\citep{franklin2006atlantic}. When county-level exposure was determined based on
the wind metric, only counties near two of the storm's landfalls were assessed
as exposed. For rain- and flood-based metrics, however, exposure extended to the
left of the track, including counties as far north as New York and Connecticut,
while for the tornado metric, exposed counties tended to be to the right of the
track and included several counties in central North Carolina, South Carolina,
and Georgia that were not identified as exposed to Ivan based on any other
metric.

We drew similar conclusions when we investigated all 46 storms between 1996 and
2011 (the period for which all five metrics were available) for which 250 or
more counties were exposed based on at least one metric (Fig.
\ref{fig:jaccard}). The tornado-based metric showed universally poor agreement
with other metrics in county-level classification across all storms considered.
For other pairs of metrics, there were also generally large differences in
which counties were determined to be exposed to the storm, with the Jaccard
index values (see Methods) below 0.4 for most metric pairs in most storms.

There were, however, a few exceptions---storms in which similar counties were
determined to be exposed to the storm for two or more of the metrics
considered.  For Floyd (1999) and Irene (2011), for example, county-level
classification agreed moderately to well (Jaccard index of approximately
0.5--0.8) for all pairs of exposure metrics except those including the
tornado-based metric. These storms both made their first U.S. landfall in North
Carolina at minor hurricane windspeeds (Category 2 and 1, respectively) and
then skimmed the eastern coast of the U.S. north through New England, bringing
large rainfall to much of the eastern coast from North Carolina north and
causing extensive inland flooding in North Carolina (Floyd) and New England
(Irene) \citep{avila2013atlantic, lawrence2000atlantic}. For another set of
storms (e.g., Lee [2011], Ernesto [2006], and Bertha [1996]), there was
moderate to good agreement for all pairwise combinations of distance, rain, and
wind, but poor agreement for all other combinations of metrics. Most of these
storms either skimmed along the eastern coast north of North Carolina (Bertha
[1996], Charley [2004], Gaston [2004], Barry [2007], Hanna [2008]), in a
pattern similar to Hurricanes Floyd (1999) and Irene (2011), or made landfall
in or near the Florida panhandle and cut across to exit into the Atlantic
around North Carolina (Josephine [1996], Earl [1998], Alberto [2006]). However,
these two sets of storms represented the minority of all storms considered; for
most of the storms, there was low overlap in the counties determined to be
exposed to the storm for most pairings of metrics.

\begin{figure*}%[tbhp] 
\centering
\includegraphics[width=16cm]{averageexposureonly} 
\caption{Average number of storm exposures per decade in U.S. counties for 
each exposure metric. The criteria behind each of the five metrics is given 
in Table \ref{tab:exposuremetrics}. The years used to estimate these averages 
are based on years of available exposure data (distance and wind: 1988--2015; 
rain: 1988--2011; flood and tornado: 1996--2015). Similar patterns persist when
analysis is restricted to years with all exposure data available (1996--2011;
Fig. S3).} 
\label{fig:averageexposure} 
\end{figure*}

\subsection*{Geographic patterns in tropical cyclone exposures}

When different metrics were used to assess average tropical cyclone exposure in
U.S. counties, geographic patterns varied substantially; in fact, some areas of
the U.S. that experienced regular exposures to storm-related excessive rainfall
and floods were not identified as areas of regular risk based on more commonly
used assessment metrics like wind or distance (Figs. \ref{fig:averageexposure},
S3). Wind-based exposure had a strong coastal pattern, with almost all
exposures in counties within about 200 kilometers (124 miles) of the coastline,
mostly in eastern North and South Carolina, southern Florida, and the area near
the outlet of the Mississippi River. While exposures by rain- and
distance-based metrics were also more common in coastal areas compared to
inland areas, inland exposures were much more common based on these two metrics
compared to the wind-based metric. Rain and, to some extent, flood exposures
were characterized by a pattern defined by the Appalachian Mountains, with much
lower exposures west of the mountain range than to the east.  Flood- and,
especially, tornado-based exposures were much less common than exposures based
on other metrics. Almost all tornado-based exposures were in coastal states,
with most in Florida, Alabama, South Carolina, and North Carolina and almost
none north of Maryland. Flood-based exposures were highest in North Carolina
and along the South Carolina coast, as well as further north in New Jersey,
southeastern Pennsylvania, and southeastern New York.

We have included a case study in the Supporting Information showing how these
differences can strongly influence estimates of which U.S. counties have the
largest expected physical exposure to tropical cyclones among a susceptible
subpopulation, Medicare beneficiaries who are reliant on electricity for
medical equipment \citep{empower}.

\subsection*{Software}

To assist with future tropical cyclone impact studies, we created and published
open source software that includes the hazard-specific, county-level tropical
cyclone data in this paper, as well as tools to explore and map that data
\citep{hurricaneexposure, hurricaneexposuredata}.  We investigated how well
these data correspond with data from other available sources (Figs. S6--8).
While generally in agreement with data from other sources, there are a few
limitations. The wind data are based on modeled, rather than observed, values,
and while the modeled wind data generally agree well with post-analyis maximum
wind radii from the revised Atlantic hurricane database (HURDAT2
\citep{landsea2013}) (Fig. S6), for a few storms it did not (e.g., Fig.  S9).
The rainfall data are from re-analysis data, which are generally
well-correlated with observed ground-based station data but may oversmooth
extreme measurements (Fig. S7).

\section*{Discussion}

Tropical cyclones impact public health in many ways, and epidemiological studies
of the health impacts of tropical cyclone exposures could help improve
preparedness for and response to future storms. However, tropical storms are
multi-hazard events, making it more complicated to measure exposure to tropical
cyclones compared to other weather exposures like temperature and precipitation.
... Our results can inform exposure assessment for future county-level studies
of the health impacts of tropical cyclones exposures, and the data and software
presented provides multi-hazard, county-level exposure data for multiple
tropical cyclone hazards for such studies.

Wind-based exposure had a strong coastal pattern, with almost all exposures in
counties within about 200 kilometers (124 miles) of the coastline, mostly in
eastern North and South Carolina, southern Florida, and the area near the
outlet of the Mississippi River. This is consistent with the dramatic decrease
in wind intensity that typically characterizes the landfall of tropical
cyclones. While exposures by rain- and distance-based metrics were also more
common in coastal areas compared to inland areas, inland exposures were much
more common based on these two metrics compared to the wind-based metric. Rain
and, to some extent, flood exposures were characterized by a pattern defined by
the Appalachian Mountains, with much lower exposures west of the mountain range
than to the east. This agrees with previous research indicating that the
Appalachian mountains' topography both enhances precipitation during tropical
cyclones and provides hydrological conditions for severe flooding
\citep{rees2001}.  Almost all tornado-based exposures were in coastal states,
with most in Florida, Alabama, South Carolina, and North Carolina and almost
none north of Maryland. This pattern may be linked with previous findings that
most noteworthy storm-related tornadoes occur on the right-side of the storm
track in Atlantic Basin U.S. storms \citep{moore2012}.

The tropical cyclone exposure averages shown in Fig. \ref{fig:averageexposure}
are limited as estimates of long-term frequencies, as tropical cyclones follow
decadal patterns \citep{kossin2007more} likely not adequately captured in the
available data. However, these frequency maps do illustrate the potential for
strong differences in spatial patterns in tropical cyclone exposures, depending
on which storm hazards are considered. A few previous studies have sought to
determine county-level exposure to tropical cyclones over multi-year periods,
including \citep{zandbergen2009}, which estimated exposure in U.S. counties to
all U.S. landfalling Atlantic basin tropical cyclones between 1851 and 2003,
using both a distance-based metric and a metric that combined distance and
windspeed, and \citep{kruk2010}, which explored exposure to hurricane-related
winds in the U.S., including inland areas, for 1900--2008. Our results suggest
that such exposure assessments may perform well in capturing some storm hazards
(e.g., wind), but likely miss other potentially dangerous tropical cyclone
exposures, especially for hazards that repeatedly threaten northern or inland
counties (e.g., rain, flooding). We have included a case study in the Supporting
Information showing how these differences can strongly influence estimates of
which U.S. counties have the largest expected physical exposure to tropical
cyclones among a susceptible subpopulation, Medicare beneficiaries who are
reliant on electricity for medical equipment \citep{empower}.

Median storm extent was largest when exposure was classified based on distance,
likely because this metric includes counties along or near the paths of less
severe storms, as well as inland counties near a storm that has weakened
substantially.

Previous research supports the idea that some storms may differ dramatically in
extent based on which hazards of the storm are of interest. While storm rainfall
and windspeed may be well-correlated when the storm is over water
\citep{cerveny2000}, this relationship does not remain as strong once the
hurricane has made landfall \citep{jiang2008}. Fast-moving storms bring higher
risks of dangerous winds inland \citep{kruk2010}, while slow-moving storms are
likely to bring more rain \citep{rappaport2000} and cause more damage because of
sustained hazardous conditions \citep{rezapour2014}. Further, while the
likelihood and extent of flooding during a tropical cyclone is related to the
storm's rainfall, it is also driven by factors like top soil saturation and the
structure of the water basin's drainage network \citep{chen2015, rees2001}.

The tornado-based metric showed universally poor agreement
with other metrics in county-level classification across all storms considered.
This likely results from tornado-based exposure being very rare, even during
large storms (Fig. \ref{fig:severityagreement}).

For both sets of storms for which agreement between metrics were unusually high
(e.g., Floyd (1999), Irene (2011), ...), the storms' persistent proximity to
water may have helped maintain wind speeds in similar patterns to rain and
distance exposures, resulting in this moderate to good agreement among exposure
assessments based on different metrics.

Distance from a tropical cyclone's track is relatively easy to measure and has
been used as an operational metric of exposure to tropical cyclones in previous
large-scale studies (Fig. S1). Since distance itself does not constitute a
hazard, distance is meant in these cases as a surrogate to capture exposure to
hazards from the storm. However, here we found that in assessing U.S.
county-level exposure to tropical cyclones, distance is, at best, a moderate,
and often a very poor, surrogate for exposure to the specific storm hazards of
high wind, extreme rainfall, flooding, and tornadoes (Fig. \ref{fig:jaccard}).
Therefore, use of distance to assess tropical cyclone exposure for impact
studies could result in problematic exposure misclassification, which could mask
true associations, even strong associations, between storm exposure and outcomes
of interest in impact studies \citep{savitz2016interpreting,
armstrong1998effect}. In some cases, this exposure misclassification may be
differential (i.e., associated with the outcome of interest or with factors
associated with risk of the outcome of interest). For example, tropical cyclone
wind exposures tend to be concentrated in counties near the coast, since most
storms rapidly decrease in sustained windspeed following landfall. However,
tropical cyclone exposures based on distance can extend well inland, following
the storm's tracks, but may not adequately capture all wind-exposed counties
near the coast. In this case, if the etiologically-relevant exposure is high
wind but exposure is classified based on distance, the probabilty of being
misclassified as unexposed would be higher in coastal counties, while the
probability of being misclassified as exposed would be higher in inland
counties. If coastal counties differ from inland counties in either the outcome
of interest or in factors associated with risk of that outcome, differential
exposure misclassification would exist \citep{savitz2016interpreting}. Such
differential exposure misclassification can bias estimates of storm effects
either towards the null (estimating a lower or null association compared to the
true association that exists) or away from the null (estimating a larger
association than actually exists) \citep{savitz2016interpreting,
armstrong1998effect}. The use of a single hazard-based metric (e.g., wind) could
cause similar problems if the impact is driven, at least in part, by a different
hazard or by multiple hazards of the storm.

\subsection*{Software}

To assist with future tropical cyclone impact studies, we created and published
open source software that includes the hazard-specific, county-level tropical
cyclone data in this paper, as well as tools to explore and map that data
\citep{hurricaneexposure, hurricaneexposuredata}. Many previous studies have
used geographical information system software (e.g., ArcGIS) to assess exposure
to tropical cyclones in the U.S. \citep{grabich2016, zandbergen2009,
czajkowski2011, kruk2010}. Here, we offer methods to map and output historic
exposure to tropical cyclones that does not require the use of proprietary
software but instead uses a package written in the R statistical programming
language \citep{R}, which is free and open-source.

In addition to providing tropical cyclone exposure metrics for individual
hazards, this software and data also allow users to create storm exposure
profiles based on multiple hazards or craft exposure indices that combine hazard
metrics \citep{chakraborty2005population, peduzzi2009assessing}. This ability
can be critical, as different hazards of extreme storms often act
synergistically in their impacts \citep{smith2009}. Further, by including
measurements of different hazard exposures in each county for each storm, this
software allows for the development of more complex models that incoroporate
multiple potential predictors of an impact (e.g., random forests, multivariable
generalized linear models).

In creating this hurricane exposure dataset, we aimed for data that are
reliable, available for all eastern U.S. counties, and have no missing data
within available dates. Further, we have investigated how well these data
correspond with data from other available sources (Figs. S6--8). However, there
are still some important limitations. The wind data are based on modeled, rather
than observed, values, and while the modeled wind data generally agree well with
post-analyis maximum wind radii from the revised Atlantic hurricane database
(HURDAT2 \citep{landsea2013}) (Fig. S6), for a few storms it did not (e.g., Fig.
S9). The rainfall data are from re-analysis data, which are generally
well-correlated with observed ground-based station data but may oversmooth
extreme measurements (Fig. S7). The flood and tornado data came from the
National Oceanic and Atmospheric Administration's (NOAA's) Storm Events
database, which, while a widely-used database of events maintained by NOAA, is
based on reports, and so may be prone to underreporting \citep{Ashley2008flood,
Curran2000}, especially in less populated areas \citep{Witt1998, Ashley2007}, as
well as to other reporting errors. While these are important limitations, we
selected these data sources as among the best currently available for measuring
each of these hazards at a multi-county, multi-year scale.

\subsection*{Conclusions}

Previous research has highlighted the range of impacts that tropical cyclones
can have in U.S. communities (e.g., \cite{pielke1997, rappaport2000, liu2005,
han2009, mallin2006}). Determining the metrics of storm exposure that are most
associated with loss of life, property damage, and other impacts may help
identify and quantify the important threats that remain from tropical cyclones
in the U.S., which could allow future success in reducing these threats through
community planning, warning systems, and other measures. Here we found large
differences in which counties are exposed to different hazards of tropical
cyclones and that distance is, at best, a moderate, and often a very poor,
surrogate for exposure to the specific storm hazards of high wind, extreme
rainfall, flooding, and tornadoes. Use of distance as a surrogate for any of
these hazards could lead to exposure misclassification and, in the case of storm
impact studies, result in biased storm risk estimates. Similarly, if studies
disagree on their findings of tropical cyclone impacts, it could result from the
often poor agreement between exposure classifications based on these different
metrics, making such studies that use different metrics hard to compare or
combine in meta-analyses.

\section*{Acknowledgments}

This work was supported in part by grants from the National Institute of
Environmental Health Sciences (R00ES022631), the National Science Foundation
(1331399), the Department of Energy (Grant No. DE-FG02-08ER64644), and a
National Aeronautics and Space Administration Applied Sciences Program/Public
Health Program Grant (NNX09AV81G). Rainfall data are based on data acquired as
part of the mission of the National Aeronautics and Space Administration's Earth
Science Division and archived and distributed by the Goddard Earth Sciences
(GES) Data and Information Services Center (DISC).

\bibliography{hurr_exposure}

\end{document}

### 2018-10-04 poster/hurr_exposure_poster.tex
Initial commit

\documentclass[25pt, a0paper, portrait, margin=0mm, innermargin=15mm,
     blockverticalspace=15mm, colspace=15mm, subcolspace=8mm]{tikzposter}
\usetheme{Wave}
\useblockstyle{TornOut}
\usenotestyle{Sticky}

\makeatletter
\def\title#1{\gdef\@title{\scalebox{\TP@titletextscale}{%
\begin{minipage}[t]{\linewidth}
\centering
#1
\par
\vspace{0.5em}
\end{minipage}%
}}}
\makeatother

\makeatletter
\newcounter{tablecounter}
\newenvironment{tikztable}[1][]{
  \def \rememberparameter{#1}
  \vspace{10pt}
  \refstepcounter{tablecounter}
  \begin{center}
  }{
    \ifx\rememberparameter\@empty
    \else
    \\[10pt]
    {\small Table~\thetablecounter: \rememberparameter}
    \fi
  \end{center}
}
\makeatother

\usepackage{booktabs}
\usepackage{array}
\usepackage{graphicx}  % package for inserting images, including .pdf
\usepackage{adjustbox} % package for cropping images
\usepackage[colorlinks=true, urlcolor=orange]{hyperref} % package for url and hyperlinks
\usepackage{wrapfig}

\title{Assessing United States county-level exposure for 
  hurricane and tropical storm  epidemiological research}
\author{Brooke Anderson*, Joshua Ferreri, Mohammad Al-Hamdan, William Crosson, Andrea Schumacher, Seth Guikema, Steven Quiring, Dirk Eddelbuettel, Meilin Yan, Roger Peng}
\institute{*Colorado State University, Fort Collins, CO, USA \\[0.5cm] 
E-mail: \texttt{\href{mailto:brooke.anderson@colostate.edu}{brooke.anderson@colostate.edu}}}

\begin{document}

\maketitle

\begin{columns}

%% Left column
\column{0.33}

% Summary block
\block{1. Motivation}{
Hurricanes and other tropical storms bring severe impacts to U.S. communities. These impacts can result from a variety of storm-related hazards, including extreme wind, rain, flooding, and tornadoes. Epidemiological studies vary widely in how they classify \textbf{exposure to tropical cyclones}, using various hazard-based metrics and, in some cases, using distance from the storm as a surrogate for exposure to storm-related hazards.

} % Closes block 1, Summary

\block{2. Identifying storm exposures}{

Here we measure county-level exposure to tropical cyclones in the United States based on several of these metrics for Atlantic basin storms for 1996--2011 (Table 1). \\[0.01cm]

\coloredbox{
% Exposure definition table
\small
\begin{tikztable}[Criteria for tropical cyclone exposures]
\begin{tabular}{p{3cm}p{19cm}} 
\toprule
Metric &  Criteria for exposure \\ \midrule 
\textbf{Distance} &  County population mean center within 100 kilometers of storm track \\ 
\textbf{Rain} &  County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\ 
\textbf{Wind} &  Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
\textbf{Flood} &  Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
\textbf{Tornado} &  Tornado event listed NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
\bottomrule 
\end{tabular} 
\end{tikztable}
}

\medskip

We considered all tracked storms in the revised Atlantic
hurricane database (HURDAT2).
For the \textbf{rain}-based exposure metric, cumulative rainfall was estimated from two days before to one day after the storm's closest
approach to the county, using North American Land Data
Assimilation System Phase 2 (NLDAS-2) data. 
For the \textbf{wind}-based exposure metric, we modeled maximum sustained wind
speeds at each county's population mean center.
To identify \textbf{flood}- and \textbf{tornado}-based exposures to tropical cyclones in U.S.
counties, we used event listings from the National Oceanic and Atmospheric
Administration's Storm Event Database. \\[0.01cm]

  % Sub-block with R package information
  \innerblock{R package \texttt{hurricaneepxosure}}{

To assist with future tropical cyclone epidemiological studies, we created 
the R package \textbf{hurricaneexposure}, which is
open source software with hazard-specific, county-level tropical
cyclone exposure data, as well as tools to explore and map the data.
  }

} % Closes block 2, identifying tropical cyclone exposures

\block{3. Measuring agreement}{

We
measured agreement between exposure metrics by calculating
the within-storm \textbf{Jaccard index} between every pair-wise combination of metrics. The
Jaccard index ($J_s$) measures similarity between two metrics ($X_{1,s}$ and
$X_{2,s}$) for storm $s$ as the proportion of counties in which both of the
metrics classify the county as exposed, out of all counties classified as exposed
by at least one of the metrics:

\begin{equation} 
J_s = \frac{X_{1,s} \cap X_{2,s}}{X_{1,s} \cup X_{2,s}}
\end{equation}

} % Closes block 3., measuring agreement between metrics

%% Right column
\column{0.67}

\block{4. Agreement among exposure classifications}{

  % Sub-block with Ivan example
  \innerblock{Agreement for one example storm: Hurricane Ivan (2004)}{
  \begin{wrapfigure}[20]{r}{0.68\colwidth}
    \begin{tikzfigure}[Counties classified as exposed to Hurricane Ivan (2004) 
                       (orange) based on each exposure metric. The red line shows 
                       the track of Hurricane Ivan based on the revised Atlantic
                       hurricane database (HURDAT2).] 
    \includegraphics[width = 0.68\colwidth]{../figures/ivanonly}
    \end{tikzfigure} 
  \end{wrapfigure}
  We found that switching among these metrics can dramatically change which counties are identified as exposed to a particular storm.
Fig. 1 gives an example for Hurricane Ivan (2004). When county-level exposure was determined based on
the \textbf{wind} metric, only counties near two of the storm's landfalls were assessed
as exposed. For \textbf{rain}- and \textbf{flood}-based metrics, however, exposure extended to the
left of the track, including counties as far north as New York and Connecticut,
while for the \textbf{tornado} metric, exposed counties tended to be to the right of the
track and included several counties not identified as exposed by any other
metric.
  }

  \vspace{0.5cm}

  % Sub-block with Jaccard agreement
  \innerblock{Agreement for all major storms}{

  \begin{wrapfigure}[18]{l}{0.45\colwidth}
  \begin{tikzfigure}[Jaccard index values for
agreement between exposure metric pairs for major storms, 1996--2011.]
  \includegraphics[width=0.45\colwidth]{../figures/jaccard_heatmap_presentation.pdf}
  \end{tikzfigure}
  \end{wrapfigure}

   We drew similar conclusions when we investigated all 24 storms for which 250 or
more counties were exposed based on at least one metric (Fig.
2). The color of each cell within the main heatmap indicates the Jaccard index for a given pair
of metrics for a given storm. Storms are displayed within clusters that have
similar patterns in county-level exposure agreement for metric pairs. The colors to the right of
the main heatmap for each storm indicate the total number of counties classified
as exposed to the storm by any of the five metrics.

The tornado-based metric showed universally poor agreement
with other metrics in county-level classification across all storms considered.
For other pairs of metrics, there were also \textbf{generally large differences} in
which counties were determined to be exposed to the storm, with the Jaccard
index values below 0.4 for most metric pairs in most storms.

There were, however, \textbf{a few exceptions}---storms in which similar counties were
determined to be exposed to the storm for two or more of the metrics
considered.  For Floyd (1999) and Irene (2011), for example, county-level
classification agreed moderately to well (Jaccard index of approximately
0.5--0.8) for all pairs of exposure metrics except those including the
tornado-based metric. These storms both made their first U.S. landfall in North
Carolina at minor hurricane windspeeds (Category 2 and 1, respectively) and
then skimmed the eastern coast of the U.S. north through New England, bringing
large rainfall to much of the eastern coast from North Carolina north and
causing extensive inland flooding in North Carolina (Floyd) and New England
(Irene).
  }

} % Closes block 4, Agreement among exposure classification

\block{5. Conclusions}{

Previous research has highlighted the range of impacts that tropical cyclones
can have in US communities. 
Determining the metrics of storm exposure that are most
associated with loss of life, property damage, and other impacts may help
identify and quantify the important threats that remain from tropical cyclones
in the US, which could allow future success in reducing these threats through
community planning, warning systems, and other measures. Locations identified as storm-exposed varied substantially when switching among metrics based on different storm hazards, and distance to the storm was at best a moderate, and often a poor, surrogate for storm hazard exposures. Therefore, when impact studies use distance as a surrogate of exposure to tropical cyclone exposures or use one hazard-based metric (e.g., wind-based) when the impact is partly or fully caused by a different storm hazard (e.g., flooding), the analysis will be prone to exposure misclassification, which can mask true associations. To facilitate future research, we make multi-hazard storm exposure data for the US available through open-source software.

}

\note[angle=270, radius=9cm, width=50cm, innersep=1cm]{
\small  
\textbf{Acknowledgements.} 
This work was supported in part by grants from the National Institute of
Environmental Health Sciences (R00ES022631), the National Science Foundation
(1331399), the Department of Energy (Grant No. DE-FG02-08ER64644), and a
National Aeronautics and Space Administration Applied Sciences Program/Public
Health Program Grant (NNX09AV81G). Rainfall data are based on data acquired as
part of the mission of the National Aeronautics and Space Administration's Earth
Science Division and archived and distributed by the Goddard Earth Sciences
(GES) Data and Information Services Center (DISC).

} % Closes Acknowledgements block

\end{columns}

\end{document}

### 2018-10-04 tables/stormseverity.tex
Initial commit

% latex table generated in R 3.3.2 by xtable 1.8-2 package
% Mon Mar 27 10:38:20 2017
Metric & Median number of exposed counties (Interquartile range) & Storm with most counties exposed (\# exposed counties) \\ 
  \midrule
Distance & 62 (11, 150) & Beryl, 1994 (330) \\ 
  Wind & 26 (2, 66) & Ike, 2008 (355) \\ 
  Rain & 21 (2, 113) & Frances, 2004 (464) \\ 
  Flood & 9 (0, 39) & Ivan, 2004 (317) \\ 
  Tornado & 1 (0, 9) & Ivan, 2004 (91) \\ 
   \bottomrule

### 2018-10-04 R/average_elec_depend_exposure_figure.R

@@ -1,0 +1,162 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+library(rgdal)
+library(maptools)
+library(RColorBrewer)
+library(choroplethr)
+data(closest_dist)
+load("clean_empower.RData") # Loads dataframe named `medi`
+
+my_fips <- unique(closest_dist$fips)
+dist <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2015, dist_limit = 100) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 28) %>%
+        mutate(metric = "distance")
+
+rain <- county_rain(counties = my_fips, start_year = 1988,
+                             end_year = 2011, dist_limit = 500,
+                             rain_limit = 75) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 24) %>%
+        mutate(metric = "rain")
+
+wind <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2015, wind_limit = 15) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 28) %>%
+        mutate(metric = "wind")
+
+flood <- county_events(counties = my_fips,
+                                start_year = 1996, end_year = 2015,
+                                event_type = "flood") %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 20) %>%
+        mutate(metric = "flood")
+
+tornado <- county_events(counties = my_fips,
+                                start_year = 1996, end_year = 2015,
+                                event_type = "tornado") %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 20) %>%
+        mutate(metric = "tornado")
+
+our_fips <- hurricaneexposure:::get_eastern_map() %>% dplyr::select(fips) %>% dplyr::distinct() %>%
+        dplyr::mutate(fips = as.numeric(fips))
+all_exposures <- bind_rows(dist, rain, wind, flood, tornado) %>%
+        mutate(n_exposures = n_exposures * 10,
+               fips = as.numeric(fips)) %>%
+        right_join(our_fips, by = "fips") %>%
+        tidyr::spread(metric, n_exposures, fill = 0) %>%
[…]

### 2018-10-04 R/average_exposure_figure.R

@@ -1,0 +1,330 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+library(rgdal)
+library(maptools)
+library(RColorBrewer)
+library(choroplethr)
+data(closest_dist)
+
+my_fips <- unique(closest_dist$fips)
+dist <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2015, dist_limit = 100) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 28) %>%
+        mutate(metric = "distance")
+
+rain <- county_rain(counties = my_fips, start_year = 1988,
+                             end_year = 2011, dist_limit = 500,
+                             rain_limit = 75) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 24) %>%
+        mutate(metric = "rain")
+
+wind <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2015, wind_limit = 15) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 28) %>%
+        mutate(metric = "wind")
+
+flood <- county_events(counties = my_fips,
+                                start_year = 1996, end_year = 2015,
+                                event_type = "flood") %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 20) %>%
+        mutate(metric = "flood")
+
+tornado <- county_events(counties = my_fips,
+                                start_year = 1996, end_year = 2015,
+                                event_type = "tornado") %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / 20) %>%
+        mutate(metric = "tornado")
+
+our_fips <- hurricaneexposure:::get_eastern_map() %>% dplyr::select(fips) %>% dplyr::distinct() %>%
+        dplyr::mutate(fips = as.numeric(fips))
+all_exposures <- bind_rows(dist, rain, wind, flood, tornado) %>%
+        mutate(n_exposures = n_exposures * 10,
+               fips = as.numeric(fips)) %>%
+        right_join(our_fips, by = "fips") %>%
+        tidyr::spread(metric, n_exposures, fill = 0) %>%
+        tidyr::gather(metric, n_exposures, -fips)
+
+out_data <- hurricaneexposure:::get_eastern_map() %>%
[…]

### 2018-10-04 R/clean_empower_data.R

@@ -1,0 +1,132 @@
+# Script to clean up EmPower data, so we can match it up by FIPS with the hurricane
+# exposure data.
+
+# setwd("writing/pnas_draft/")
+library(readxl)
+library(readr)
+library(dplyr)
+library(maps)
+library(tidyr)
+library(stringr)
+library(choroplethr)
+
+
+county_fips <- read_csv("https://www2.census.gov/geo/docs/reference/codes/files/national_county.txt",
+                        col_names = c("abb", "state_fips", "county_fips",
+                                      "county_name", "h")) %>%
+        select(-h) %>%
+        unite(fips, state_fips, county_fips, sep = "") %>%
+        mutate(county_name = str_replace(str_to_lower(county_name), " parish| county", ""),
+               abb = str_to_lower(abb),
+               fips = as.integer(fips)) %>%
+        unite(county_name, county_name, abb, sep = ",")
+
+extra_fips <- data_frame(county_name = c("de soto,fl",
+                                         "saint johns,fl",
+                                         "saint lucie,fl",
+                                         "radford,va",
+                                         "salem,va",
+                                         "prince georges,md",
+                                         "queen annes,md",
+                                         "saint marys,md",
+                                         "saint lawrence,ny",
+                                         "saint clair,mi",
+                                         "saint joseph,mi",
+                                         "saint croix,wi",
+                                         "dewitt,il",
+                                         "la salle,il",
+                                         "saint clair,il",
+                                         "de kalb,in",
+                                         "la porte,in",
+                                         "st joseph,in",
+                                         "saint charles,mo",
+                                         "saint clair,mo",
+                                         "saint francois,mo",
+                                         "saint louis,mo",
+                                         "saint louis city,mo",
+                                         "sainte genevieve,mo",
+                                         "obrien,ia",
+                                         "de witt,tx",
+                                         "saint francis,ar",
+                                         "saint bernard,la",
+                                         "saint charles,la",
+                                         "saint helena,la",
+                                         "saint james,la",
+                                         "saint landry,la",
+                                         "saint martin,la",
+                                         "saint mary,la",
+                                         "saint tammany,la",
+                                         "st john the baptist,la",
+                                         "de kalb,al",
+                                         "saint clair,al",
+                                         "bedford,va"),
+                         fips = c("12027",
+                                  "12109",
+                                  "12111",
+                                  "51750",
+                                  "51775",
[…]

### 2018-10-04 R/extent_disagreement_figure.R

@@ -1,0 +1,266 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+
+storm <- "Frances-2004"
+
+rain_data <- filter_storm_data(storm = storm,
+                               days_included = c(-2, -1, 0, 1),
+                               include_rain = TRUE,
+                               output_vars = c("fips", "tot_precip",
+                                               "storm_dist")) %>%
+        dplyr::mutate_(rain = ~ tot_precip >= 75 &
+                               storm_dist <= 500) %>%
+        dplyr::select(fips, rain) %>%
+        dplyr::tbl_df()
+
+wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
+                              output_vars = c("fips", "vmax_sust")) %>%
+        `colnames<-`(c("fips", "wind_value")) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::select(fips, wind) %>%
+        dplyr::tbl_df()
+
+
+storm_data <- rain_data %>%
+        full_join(wind_data, by = "fips") %>%
+        gather(key = metric, value = Exposed, -fips) %>%
+        mutate(metric = factor(metric,
+                               levels = c("rain", "wind"),
+                               labels = c("Rain-based metric", "Wind-based metric")))
+
+out_data <- hurricaneexposure:::get_eastern_map() %>%
+        dplyr::left_join(storm_data, by = "fips")
+out <- ggplot2::ggplot() +
+        ggplot2::geom_polygon(data = out_data,
+                              ggplot2::aes_(x = ~ long, y = ~ lat, group = ~ group,
+                                            fill = ~ Exposed, color = ~ Exposed),
+                              size = 0.2) +
+        ggplot2::borders("state", regions = c("virginia", "north carolina", "south carolina",
+                                              "georgia", "florida", "alabama", "kentucky",
+                                              "tennessee", "maryland", "west virginia",
+                                              "district of columbia", "pennsylvania",
+                                              "new jersey", "delaware", "mississippi",
+                                              "louisiana", "texas", "oklahoma", "arkansas",
+                                              "new york", "connecticut", "rhode island",
+                                              "massachusetts", "new hampshire", "vermont",
+                                              "maine", "kansas", "missouri", "iowa", "michigan",
[…]

### 2018-10-04 R/flood_comparison_figure.R

@@ -1,0 +1,121 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(countyfloods)
+library(lubridate)
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(ggplot2)
+library(viridis)
+library(scales)
+
+## Fairfield County, Connecticut 09001
+# Baltimore County, MD 24005
+## Bergen County, New Jersey 34003
+## Montgomery County, Pennsylvania 42091
+## Mobile County, Alabama 01097
+## DC 11001
+## Wake County, North Carolina 37183
+## Harris County, TX, 48201
+## Fulton County, GA 13121
+## Escambia County, FL 12033
+# Spartanburg, SC 45083
+# Prince William County, VA 51153
+# Forrest County, MS 28035
+
+county_fips <- c("37183", "42091", "01097", "13121", "12033", "34003", "48201",
+                 "24005", "09001")
+county_floods <- county_events(county_fips, start_year = 1996, end_year = 2015, event_type = "flood") %>%
+        select(storm_id, fips) %>%
+        mutate(flood = TRUE)
+county_storms <- county_distance(county_fips, start_year = 1996, end_year = 2015, dist_limit = 500) %>%
+        left_join(county_floods, by = c("fips", "storm_id")) %>%
+        mutate(flood = ifelse(is.na(flood), FALSE, flood),
+               closest_date = ymd(closest_date)) %>%
+        dplyr::select(storm_id, fips, closest_date, flood) %>%
+        rename(date_0 = closest_date) %>%
+        mutate(date_m2 = date_0 - 2,
+               date_m1 = date_0 - 1,
+               date_1 = date_0 + 1,
+               date_2 = date_0 + 2,
+               date_3 = date_0 + 3,
+               date_4 = date_0 + 4,
+               date_5 = date_0 + 5,
+               date_6 = date_0 + 6) %>%
+        gather(key = lag, value = date, -storm_id, -fips, -flood) %>%
+        mutate(lag = str_replace(lag, "date_", ""),
+               lag = str_replace(lag, "m", "-"),
+               lag = as.numeric(lag))
+
+# county_gages <- get_gages(county_fips, start_date = "1996-01-01", end_date = "2015-12-31")
+# gage_q2s <- find_q2(county_gages$site_no)
+# county_gages <- full_join(county_gages, gage_q2s, by = "site_no")
[…]

### 2018-10-04 R/hurr_tracks_figure.R

@@ -1,0 +1,35 @@
+## Map of hurricane tracks
+library(ggplot2)
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+
+data(closest_dist)
+storms <- unique(closest_dist$storm_id)
+storms <- storms[gsub("*.+-", "", storms) <= 2015 &
+                         gsub("*.+-", "", storms) >= 1988]
+a <- map_tracks(storms, plot_points = FALSE, alpha = 0.3, color = "darkcyan")
+# Add storms whose names were retired
+# Source: http://www.nhc.noaa.gov/aboutnames_history.shtml
+fig <- map_tracks(c("Hugo-1989", "Bob-1991", "Andrew-1992",
+             "Opal-1995", "Fran-1996", "Georges-1998", "Mitch-1998",
+             "Floyd-1999", "Allison-2001", "Michelle-2001",
+             "Isidore-2002", "Lili-2002", "Isabel-2003",
+             "Charley-2004", "Frances-2004",
+             "Ivan-2004", "Jeanne-2004",
+             "Dennis-2005", "Katrina-2005", "Rita-2005",
+             "Wilma-2005", "Noel-2007",
+             "Gustav-2008", "Ike-2008", "Paloma-2008",
+             "Irene-2011", "Sandy-2012"),
+           plot_object = a, plot_points = FALSE, color = "darkcyan",
+           padding = 0) +
+        theme_void() +
+        theme(panel.background = element_rect(fill = "white")) +
+        coord_map()
+
+pdf(file = "figures/hurrtracks.pdf", width = 4.8, height = 4.0)
+print(fig)
+dev.off()
+
+tiff(filename = "figures/hurrtracks.tiff", width = 480, height = 400)
+print(fig)
+dev.off()

### 2018-10-04 R/ivan_figure.R

@@ -1,0 +1,116 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+
+storm <- "Ivan-2004"
+
+distance_data <- filter_storm_data(storm = storm,
+                              output_vars = c("fips", "storm_dist")) %>%
+        dplyr::mutate_(distance = ~ storm_dist <= 100) %>%
+        dplyr::select(fips, distance) %>%
+        dplyr::tbl_df()
+
+rain_data <- filter_storm_data(storm = storm,
+                              days_included = c(-2, -1, 0, 1),
+                              include_rain = TRUE,
+                              output_vars = c("fips", "tot_precip",
+                                              "storm_dist")) %>%
+        dplyr::mutate_(rain = ~ tot_precip >= 75 &
+                               storm_dist <= 500) %>%
+        dplyr::select(fips, rain) %>%
+        dplyr::tbl_df()
+
+wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
+                             output_vars = c("fips", "vmax_sust")) %>%
+        `colnames<-`(c("fips", "wind_value")) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::select(fips, wind) %>%
+        dplyr::tbl_df()
+
+storm_year <- gsub("*.+-", "", storm)
+counties <- hurricaneexposuredata::closest_dist %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips), quote(storm_dist))
+
+flood_data <- county_events(counties = counties$fips,
+                          start_year = storm_year,
+                          end_year = storm_year,
+                          event_type = "flood") %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips)) %>%
+        dplyr::mutate_(event = ~ 1) %>%
+        dplyr::right_join(counties, by = "fips") %>%
+        dplyr::mutate_(event = ~ !is.na(event)) %>%
+        dplyr::rename_(flood = ~ event) %>%
+        dplyr::select_(quote(-storm_dist))
+
+tornado_data <- county_events(counties = counties$fips,
+                            start_year = storm_year,
+                            end_year = storm_year,
+                            event_type = "tornado") %>%
[…]

### 2018-10-04 R/ivan_figure_presentation.R

@@ -1,0 +1,112 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+
+storm <- "Ivan-2004"
+
+distance_data <- filter_storm_data(storm = storm,
+                              output_vars = c("fips", "storm_dist")) %>%
+        dplyr::mutate_(distance = ~ storm_dist <= 100) %>%
+        dplyr::select(fips, distance) %>%
+        dplyr::tbl_df()
+
+rain_data <- filter_storm_data(storm = storm,
+                              days_included = c(-2, -1, 0, 1),
+                              include_rain = TRUE,
+                              output_vars = c("fips", "tot_precip",
+                                              "storm_dist")) %>%
+        dplyr::mutate_(rain = ~ tot_precip >= 75 &
+                               storm_dist <= 500) %>%
+        dplyr::select(fips, rain) %>%
+        dplyr::tbl_df()
+
+wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
+                             output_vars = c("fips", "vmax_sust")) %>%
+        `colnames<-`(c("fips", "wind_value")) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::select(fips, wind) %>%
+        dplyr::tbl_df()
+
+storm_year <- gsub("*.+-", "", storm)
+counties <- hurricaneexposuredata::closest_dist %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips), quote(storm_dist))
+
+flood_data <- county_events(counties = counties$fips,
+                          start_year = storm_year,
+                          end_year = storm_year,
+                          event_type = "flood") %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips)) %>%
+        dplyr::mutate_(event = ~ 1) %>%
+        dplyr::right_join(counties, by = "fips") %>%
+        dplyr::mutate_(event = ~ !is.na(event)) %>%
+        dplyr::rename_(flood = ~ event) %>%
+        dplyr::select_(quote(-storm_dist))
+
+tornado_data <- county_events(counties = counties$fips,
+                            start_year = storm_year,
+                            end_year = storm_year,
+                            event_type = "tornado") %>%
[…]

### 2018-10-04 R/jaccard_heatmap.R

@@ -1,0 +1,274 @@
+library(hurricaneexposure)
+library(hurricaneexposuredata)
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(purrr)
+library(viridis)
+library(dendextend)
+library(ComplexHeatmap)
+library(circlize)
+
+data(closest_dist)
+
+calc_jaccard <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        jaccard <- proxy::simil(t(x_mat), method = "jaccard")
+        out <- combn(colnames(x_mat), 2) %>%
+                t() %>%
+                tbl_df() %>%
+                dplyr::rename(metric_1 = V1,
+                              metric_2 = V2) %>%
+                dplyr::mutate(jaccard = as.vector(jaccard))
+        return(out)
+}
+
+calc_metric_1_not_2 <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        met_1_not_2 <- function(i, j){
+                x <- x_mat[ , i]
+                y <- x_mat[ , j]
+                if(sum(x | y) == 0) return(NA)
+                a <- sum(x & y)
+                b <- sum(x & !y)
+                c <- sum(!x & y)
+                metric <- b / (a + b + c)
+                return(metric)
+        }
+        out_metric <- outer(1:5, 1:5, FUN = Vectorize(met_1_not_2))
+        out <- combn(colnames(x_mat), 2) %>%
+                t() %>%
+                tbl_df() %>%
+                dplyr::rename(metric_1 = V1,
+                              metric_2 = V2) %>%
+                dplyr::mutate(met_1_not_2 = as.vector(t(out_metric))[t(upper.tri(out_metric))])
+        return(out)
+}
+
+
+calc_metric_2_not_1 <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        met_2_not_1 <- function(i, j){
+                x <- x_mat[ , i]
+                y <- x_mat[ , j]
+                if(sum(x | y) == 0) return(NA)
+                a <- sum(x & y)
+                b <- sum(x & !y)
+                c <- sum(!x & y)
+                metric <- c / (a + b + c)
[…]

### 2018-10-04 R/jaccard_heatmap_presentation.R

@@ -1,0 +1,272 @@
+library(hurricaneexposure)
+library(hurricaneexposuredata)
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(purrr)
+library(viridis)
+library(dendextend)
+library(ComplexHeatmap)
+library(circlize)
+
+data(closest_dist)
+
+calc_jaccard <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        jaccard <- proxy::simil(t(x_mat), method = "jaccard")
+        out <- combn(colnames(x_mat), 2) %>%
+                t() %>%
+                tbl_df() %>%
+                dplyr::rename(metric_1 = V1,
+                              metric_2 = V2) %>%
+                dplyr::mutate(jaccard = as.vector(jaccard))
+        return(out)
+}
+
+calc_metric_1_not_2 <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        met_1_not_2 <- function(i, j){
+                x <- x_mat[ , i]
+                y <- x_mat[ , j]
+                if(sum(x | y) == 0) return(NA)
+                a <- sum(x & y)
+                b <- sum(x & !y)
+                c <- sum(!x & y)
+                metric <- b / (a + b + c)
+                return(metric)
+        }
+        out_metric <- outer(1:5, 1:5, FUN = Vectorize(met_1_not_2))
+        out <- combn(colnames(x_mat), 2) %>%
+                t() %>%
+                tbl_df() %>%
+                dplyr::rename(metric_1 = V1,
+                              metric_2 = V2) %>%
+                dplyr::mutate(met_1_not_2 = as.vector(t(out_metric))[t(upper.tri(out_metric))])
+        return(out)
+}
+
+
+calc_metric_2_not_1 <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        met_2_not_1 <- function(i, j){
+                x <- x_mat[ , i]
+                y <- x_mat[ , j]
+                if(sum(x | y) == 0) return(NA)
+                a <- sum(x & y)
+                b <- sum(x & !y)
+                c <- sum(!x & y)
+                metric <- c / (a + b + c)
[…]

### 2018-10-04 R/other_top_storm_exposures.R

@@ -1,0 +1,419 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+
+storm <- "Floyd-1999"
+
+distance_data <- filter_storm_data(storm = storm,
+                              output_vars = c("fips", "storm_dist")) %>%
+        dplyr::mutate_(distance = ~ storm_dist <= 100) %>%
+        dplyr::select(fips, distance) %>%
+        dplyr::tbl_df()
+
+rain_data <- filter_storm_data(storm = storm,
+                              days_included = c(-2, -1, 0, 1),
+                              include_rain = TRUE,
+                              output_vars = c("fips", "tot_precip",
+                                              "storm_dist")) %>%
+        dplyr::mutate_(rain = ~ tot_precip >= 75 &
+                               storm_dist <= 500) %>%
+        dplyr::select(fips, rain) %>%
+        dplyr::tbl_df()
+
+wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
+                             output_vars = c("fips", "vmax_sust")) %>%
+        `colnames<-`(c("fips", "wind_value")) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::select(fips, wind) %>%
+        dplyr::tbl_df()
+
+storm_year <- gsub("*.+-", "", storm)
+counties <- hurricaneexposuredata::closest_dist %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips), quote(storm_dist))
+
+flood_data <- county_events(counties = counties$fips,
+                          start_year = storm_year,
+                          end_year = storm_year,
+                          event_type = "flood") %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips)) %>%
+        dplyr::mutate_(event = ~ 1) %>%
+        dplyr::right_join(counties, by = "fips") %>%
+        dplyr::mutate_(event = ~ !is.na(event)) %>%
+        dplyr::rename_(flood = ~ event) %>%
+        dplyr::select_(quote(-storm_dist))
+
+tornado_data <- county_events(counties = counties$fips,
+                            start_year = storm_year,
+                            end_year = storm_year,
+                            event_type = "tornado") %>%
[…]

### 2018-10-04 R/previous_exposure_metrics.R

@@ -1,0 +1,93 @@
+library(tibble)
+library(stringr)
+library(dplyr)
+library(tidyr)
+library(forcats)
+library(purrr)
+library(ggplot2)
+library(viridis)
+library(ggthemes)
+
+exp_metrics <- tribble(
+        ~ study, ~ area, ~ metric, ~ n,
+        "Grabich et al. 2015", "birth outcomes", "wind, distance", "1, 1",
+        "Grabich et al. 2016a", "birth outcomes", "damage, distance, wind", "1, 3, 2",
+        "Currie and Rossin-Slater 2013", "birth outcomes", "distance, wind", "4, 1",
+        "Kinney et al. 2008", "other health", "distance", "1",
+        "Czajkowski et al. 2011", "mortality", "distance", "3",
+        "Swerdel et al. 2014", "mortality", "damage", "1",
+        "McKinney et al. 2011", "mortality", "distance, other", "1, 1",
+        "Zandbergen 2009", "tropical cyclone exposure", "distance, wind", "1, 1",
+        "Amstadter et al. 2010", "mental health", "flooding, wind, damage, other", "1, 1, 1, 1",
+        "Escobedo et al. 2009", "physical damage", "damage", "1",
+        "Acierno et al. 2006", "mental health", "distance, damage, wind", "1, 1, 1",
+        "Tansel and Sizirici 2010", "physical damage", "distance", "3",
+        "Baggerly and Ferretti 2008", "educational outcomes", "other", "1",
+        "Zahran et al. 2010", "birth outcomes", "damage, other", "1, 1",
+        "Grabich et al. 2016b", "birth outcomes", "wind", "2",
+        "Esnard et al. 2011", "tropical cyclone exposure", "wind, distance", "1, 1",
+        "Fuller 2014", "educational outcomes", "damage", "1",
+        "Lieberman-Cribbin et al. 2017", "mental health", "flooding", "2",
+        "Le et al. 2013", "mental health", "other", "1",
+        "Mongin et al. 2017", "other health", "damage, flooding", "2, 2",
+        "Logan and Xu 2015", "tropical cyclone exposure", "wind, damage, other", "1, 1, 1",
+        "Kim et al. 2016", "other health", "other", "1",
+        "Kim and Marcouiller 2015", "physical damage", "damage, other", "1, 1",
+        "Horney et al. 2016", "tropical cyclone exposure", "damage", "1",
[…]

### 2018-10-04 R/rain_comparison_figure.R

@@ -1,0 +1,285 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(lubridate)
+library(ggplot2)
+library(gridExtra)
+
+data("closest_dist")
+# library(countyweather) # If pulling data rather than using saved data
+
+
+plot_county_rain_compare <- function(ex_fips, ex_dir, ex_title,
+                                     get_data = FALSE){
+        if(get_data){
+                write_daily_timeseries(ex_fips, coverage = 0,
+                                       date_min = "1988-01-01",
+                                       date_max = "2011-12-31",
+                                       var = "PRCP",
+                                       out_directory = ex_dir,
+                                       keep_map = FALSE)
+        }
+
+        check_dates <- closest_dist %>%
+                dplyr::filter(fips == ex_fips) %>%
+                dplyr::select(-storm_dist) %>%
+                dplyr::mutate(closest_date = ymd(closest_date)) %>%
+                dplyr::rename(day_0 = closest_date) %>%
+                dplyr::mutate(fips = as.integer(fips),
+                              day_0 = day_0 + days(0),
+                              day_b1 = day_0 - days(1),
+                              day_b2 = day_0 - days(2),
+                              day_a1 = day_0 + days(1)) %>%
+                dplyr::select(storm_id, day_b2, day_b1, day_0, day_a1) %>%
+                tidyr::gather(key = lag, value = day, -storm_id) %>%
+                dplyr::rename(date = day)
+        all_dates <- unique(check_dates$date)
+
+        ex_weather <- readRDS(paste0(ex_dir, ex_fips, ".rds"))
+
+        ex_weather <- ex_weather %>%
+                dplyr::filter(date %in% all_dates) %>%
+                dplyr::right_join(check_dates, by = "date") %>%
+                dplyr::group_by(storm_id) %>%
+                dplyr::summarize(prcp = sum(prcp),
+                                 ave_n = mean(prcp_reporting))
+
+        ex_rain <- county_rain(counties = ex_fips,
+                               start_year = 1988, end_year = 2011,
+                               rain_limit = 0, dist_limit = 500) %>%
+                full_join(ex_weather, by = "storm_id") %>%
[…]

### 2018-10-04 R/storm_severity_table.R

@@ -1,0 +1,71 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(xtable)
+library(stringr)
+
+my_fips <- unique(closest_dist$fips)
+distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2015, dist_limit = 100) %>%
+        select(storm_id, fips) %>%
+        mutate(distance_exposed = TRUE,
+               storm_id = as.character(storm_id))
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500) %>%
+        select(storm_id, fips) %>%
+        mutate(rain_exposed = TRUE,
+               storm_id = as.character(storm_id))
+wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2015, wind_limit = 15) %>%
+        select(storm_id, fips) %>%
+        mutate(wind_exposed = TRUE,
+               storm_id = as.character(storm_id))
+flood_exposure <- county_events(counties = my_fips,
+                                start_year = 1996, end_year = 2015,
+                                event_type = "flood") %>%
+        select(storm_id, fips) %>%
+        mutate(flood_exposed = TRUE,
+               storm_id = as.character(storm_id))
+tornado_exposure <- county_events(counties = my_fips,
+                                  start_year = 1996, end_year = 2015,
+                                  event_type = "tornado") %>%
+        select(storm_id, fips) %>%
+        mutate(tornado_exposed = TRUE,
+               storm_id = as.character(storm_id))
+total_exposure <- distance_exposure %>%
+        full_join(rain_exposure, by = c("storm_id", "fips")) %>%
+        full_join(wind_exposure, by = c("storm_id", "fips")) %>%
+        full_join(flood_exposure, by = c("storm_id", "fips")) %>%
+        full_join(tornado_exposure, by = c("storm_id", "fips")) %>%
+        gather(metric, exposed,
+               -storm_id, -fips, na.rm = FALSE) %>%
+        mutate(exposed = ifelse(is.na(exposed), FALSE, exposed))
+
+total_exposure %>%
+        group_by(storm_id, metric) %>%
+        dplyr::summarize(n_exposed = sum(exposed)) %>%
+        dplyr::ungroup() %>%
+        dplyr::group_by(metric) %>%
+        dplyr::summarize(median_exposure2 = median(n_exposed),
[…]

### 2018-10-04 R/wind_comparison_figure.R

@@ -1,0 +1,65 @@
+library(hurricaneexposuredata)
+library(dplyr)
+library(ggplot2)
+library(polycor)
+data(storm_winds)
+data(ext_tracks_wind)
+
+wind_comp <- storm_winds %>%
+        select(fips, vmax_sust, storm_id) %>%
+        rename(vmax_sust_model = vmax_sust) %>%
+        inner_join(select(ext_tracks_wind, fips, storm_id, vmax_sust),
+                  by = c("fips", "storm_id")) %>%
+        rename(vmax_sust_ext = vmax_sust) %>%
+        mutate(vmax_sust_model_cat = ifelse(vmax_sust_model < 17.4896, 0, NA),
+               vmax_sust_model_cat = ifelse(vmax_sust_model >= 17.4896 & vmax_sust_model < 25.7200,
+                                            17.4896, vmax_sust_model_cat),
+               vmax_sust_model_cat = ifelse(vmax_sust_model >= 25.7200 & vmax_sust_model < 32.9216,
+                                            25.7200, vmax_sust_model_cat),
+               vmax_sust_model_cat = ifelse(vmax_sust_model >= 32.9216, 32.9216, vmax_sust_model_cat),
+               cats_equal = vmax_sust_model_cat == vmax_sust_ext)
+
+fig <- wind_comp %>%
+        group_by(storm_id) %>%
+        summarize(perc_same = sum(cats_equal) / n(),
+                  non_zero = sum(vmax_sust_ext != 0)) %>%
+        filter(non_zero > 0) %>%
+        ggplot(aes(x = perc_same, y = reorder(storm_id, perc_same), color = non_zero)) +
+        geom_point(size = 2) +
+        scale_x_continuous(labels = scales::percent) +
+        theme_classic() +
+        labs(x = "% of counties classified in same wind category\nby modeled storm wind and Extended Best Tracks",
+             y = "",
+             color = "Number of counties with modeled\nsustained winds of \u2265 34 knots",
+             parse = TRUE
+             ) +
+        viridis::scale_color_viridis(direction = -1, option = "B") +
+        theme(legend.position = "bottom")
+
+quartz.save(file = "figures/windcomparison.pdf", type = "pdf", width = 5.25, height = 7)
+print(fig)
+dev.off()
+
+tiff(filename = "figures/windcomparison.tiff", width = 525, height = 700)
+print(fig)
+dev.off()
+
+wind_comp_ordinal <- wind_comp %>%
[…]

### 2018-10-04 R/wind_examples_figure.R

@@ -1,0 +1,156 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(ggplot2)
+library(viridis)
+
+modeled_winds <- data.table::data.table(hurricaneexposuredata::storm_winds) %>%
+        filter(storm_id %in% c("Sandy-2012", "Ike-2008")) %>%
+        select(fips, vmax_sust, storm_id) %>%
+        mutate(wind = cut(vmax_sust, breaks = c(0, 17.4896, 25.7200, 32.9216, 100),
+                             include.lowest = TRUE, right = TRUE,
+                             labels = c("0-33.9 knots", "34-49.9 knots", "50-63.9 knots", "64+ knots")),
+               wind_source = "Modeled")
+
+ext_winds <- data.table::data.table(hurricaneexposuredata::ext_tracks_wind) %>%
+        filter(storm_id %in% c("Sandy-2012", "Ike-2008")) %>%
+        select(fips, vmax_sust, storm_id) %>%
+        mutate(wind = cut(vmax_sust, breaks = c(0, 17.4896, 25.7200, 32.9216, 100),
+                             include.lowest = TRUE, right = TRUE,
+                             labels = c("0-33.9 knots", "34-49.9 knots", "50-63.9 knots", "64+ knots")),
+               wind_source = "Extended Best Tracks")
+
+to_plot <- bind_rows(modeled_winds, ext_winds) %>%
+        mutate(wind_source = factor(wind_source, levels = c("Modeled", "Extended Best Tracks")),
+               storm_id = factor(storm_id, levels = c("Ike-2008", "Sandy-2012"),
+                                 labels = c("Hurricane Ike\n(2008)", "Hurricane Sandy\n(2012)")))
+
+
+map_dim <- apply(matrix(c(-106.65037, 25.12993, -67.00742, 47.48101),
+                        byrow = TRUE, ncol = 2),
+                 MARGIN = 2,
+                 function(x) range(x) + c(-1, 1) * 2)
+tracks <- hurricaneexposuredata::hurr_tracks %>%
+        dplyr::select_(~ latitude, ~ longitude, ~ storm_id,
+                       ~ date) %>%
+        dplyr::filter_(~ as.character(storm_id) %in% c("Sandy-2012", "Ike-2008") &
+                               longitude > map_dim[1, 1] &
+                               longitude < map_dim[2, 1] &
+                               latitude > map_dim[1, 2] &
[…]

### 2019-01-26 tables/metric_definitions.tex
Add a bunch of stuff

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{0.9cm}p{2.5cm}p{9cm}} 
Metric & Available & Criteria for exposure \\ \midrule 
Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\ 
Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
Tornado & 1996--2015 & Tornado event listed in the NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}

### 2019-01-26 tables/stormseverity.tex
Add a bunch of stuff

@@
-% latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Mon Mar 27 10:38:20 2017
-Metric & Median number of exposed counties (Interquartile range) & Storm with most counties exposed (\# exposed counties) \\ 
+% latex table generated in R 3.4.3 by xtable 1.8-2 package
+% Wed Oct 10 12:11:35 2018
+\begin{table}[ht]
+\centering
+\caption{Summary statistics for exposure metrics considered to assess county-level exposure to
+         tropical cyclones. A county exposure represents the exposure of a specific eastern U.S.  
+         county to a specific tropical cyclone. A county can be exposed to multiple tropical cyclones within
+         a year, and a tropical cyclone can create multiple county exposures under a given metric.
+         The median and interquartile range of county exposures per tropical cyclone is based on
+         the tropical cyclones for which at least one U.S. county was exposed. The years for which
+         data are available for each metric are given in Table 1.} 
+\label{tab:exposuresummaries}
+\begin{tabular}{p{1.75cm}p{3cm}p{4cm}p{4cm}}
+  \toprule
+Metric & Mean (interquartile range) of county exposures per year & Median (interquartile range) of county exposures per tropical cyclone & Tropical cyclone with most counties exposures (\# exposed counties) \\
   \midrule
-Distance & 62 (11, 150) & Beryl, 1994 (330) \\ 
-  Wind & 26 (2, 66) & Ike, 2008 (355) \\ 
-  Rain & 21 (2, 113) & Frances, 2004 (464) \\ 
-  Flood & 9 (0, 39) & Ivan, 2004 (317) \\ 
-  Tornado & 1 (0, 9) & Ivan, 2004 (91) \\ 
+Distance & 433 (224, 503) & 84 (23, 173) & Beryl, 1994 (330) \\ 
+  Rain & 401 (152, 553) & 68 (14, 146) & Frances, 2004 (464) \\ 
+  Wind & 224 (83, 353) & 34 (8, 80) & Ike, 2008 (355) \\ 
+  Flood & 213 (84, 249) & 26 (6, 57) & Ivan, 2004 (317) \\ 
+  Tornado & 53 (16, 43) & 7 (2, 16) & Ivan, 2004 (91) \\ 
    \bottomrule
+\end{tabular}
+\end{table}

### 2019-01-26 writing/anderson_ehp_supp.tex
Add a bunch of stuff

\documentclass[11pt, titlepage]{article}
\usepackage{microtype}
\usepackage{charter}

\input{ehp_preamble}

\begin{document}

\title{Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"}

\author[1,*]{G. Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1]{Meilin Yan} 
\author[8]{Roger D. Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Debian and R Projects; Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}

\date{}

\maketitle

\beginsupplement

\section*{Supplemental Text}

\subsection*{Case study: Physical exposure of electricity-dependent Medicare beneficiaries to tropical cyclones}

The outcomes of disasters depend both on the geophysical forces of the disaster as well as on the vulnerability of the society living in the geographical areas affected by those forces. \citep{chakraborty2005population, anderson2003community, cutter1996vulnerability} As a case study of how differences in tropical cyclone exposure assessments across different hazard metrics influence estimates of physical exposures of susceptible populations to tropical cyclones, we investigated how the use of different metrics influenced estimates of which eastern U.S. counties have the highest expected average exposures of electricity-dependent Medicare beneficiaries to tropical cyclones. This subpopulation was selected for this analysis since it may be particularly susceptible to health impacts from tropical cyclone exposure, especially through the pathway of storm-associated power outages and evacuations. We collected data on the number of electricity-dependent Medicare beneficiaries in each study county from the U.S. Department of Health \& Human Service's emPower Map 2.0 \citep{empower}. We then calculated the physical exposure of the electricity-dependent Medicare population in each county, based on tropical cyclone assessments using each hazard metric, following \citep{peduzzi2009assessing}:

\begin{equation}
E_c = F_c * P_c
\end{equation}

\noindent where $E_c$ is the average yearly physical exposure among electricity-dependent Medicare beneficiaries in county $c$ to tropical cyclone exposures based on a given metric, $F_c$ is the estimated yearly expected frequency of tropical cyclone exposures in county $c$ based on that metric, and $P_c$ is the size of the electricity-dependent Medicare population in county $c$, as of July 2017. When combined with estimates of vulnerability of a population to a natural hazard, such measurements of physical exposure can be used to calculate risk of human losses from the hazard \citep{peduzzi2009assessing}.

We found that a few counties (Miami-Dade County, FL; Harris County, TX) were ranked in the top three counties of physical exposure among electricity-dependent Medicare beneficiaries for almost all exposure metrics (Figure \ref{fig:topelecdependexposure}). However, there were key differences when using different exposure metrics. For example, Florida counties dominated the lists of top exposed counties based on wind- and tornado-based metrics, while over half of the counties for the flood-based metric were in Mid-Atlantic states (PA, NY).

Dissimilarity in exposure to specific hazards of tropical cyclones can result in differing patterns in estimates of county-level exposure (Fig. 1 of main text), which can lead to substantial differences in estimates of population physical exposure (Fig. \ref{fig:topelecdependexposure}), based on which storm hazard is considered. The measure of a hurricane's intensity, the Saffir-Simpson scale, expresses intensity as a function of wind speed, and so fails to capture potential risks from other hazards of the storm \citep{smith2009}, and previous assessments of physical exposure to tropical cyclones have often focused on winds or distance from the storm. For example, a global study of exposure and vulnerability to natural hazards used a combination of windspeed above a certain threshold and distance from the storm's track to assess exposure \citep{peduzzi2009assessing}. While many storm impacts might be most strongly linked to wind hazards, there is a growing recognition of the potential risks of adverse health outcomes and property damage from rain- and flood-related tropical cyclone hazards in the U.S. \citep{smith2009}. Such assessments may underemphasize potentially dangerous exposures to certain regions of the U.S., particularly inland locations. In fact, based on our assessment of physical exposures to tropical cyclones among electricity-dependent Medicare beneficiaries, the highest expected physical exposure under any metric was measured for Philadelphia County, PA, under the flood metric (Fig. \ref{fig:topelecdependexposure}). While this county has a similar-sized population of electricity-dependent Medicare beneficiaries as Miami-Dade County, FL (Fig. \ref{fig:topelecdependexposure}), the frequency of storm-related flood events in this inland county results in a very high expected physical exposure among this subpopulation, while this county is not included in the top ten counties for physical exposure based on most of the other exposure metrics.

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{previous_exposure_metrics}
\caption{Metrics used in a sample of previous studies to assign exposure to tropical cyclones in U.S.-based tropical cyclone exposure or impact studies. Citations for each study are shown on the left and the broad topic of the study on the left. Presence of a bar for a study-metric combination indicates that the given metric was used in the study in assessing exposure to tropical cyclones for that study, either alone or in combination with other metrics for which that study has bars. The color transparency of each bar indicates the number of different thresholds considered for defining storm exposure in the study under that hazard (e.g., less transparent bars indicate that a study considered multiple thresholds of the metric in defining tropical cyclone exposure as a method of sensitivity analysis). Studies for which there are two or more bars (e.g., bars for both distance and wind) indicate studies that either used separate exposure assessments, one for each metric, (e.g., as a sensitivity analysis) or used a single exposure index that incorporated multiple metrics in its definition. ``Damage" often indicate use of Federal Emergency Management Agency (FEMA) reports, although occasionally used damage reports from other sources, including newspaper reports. Examples of ``Other" include evacuation orders, excessive school closures, length of time a storm was near a location, and detailed, holistic analyses of a specific storm or storms. References for each study are included in the
bibliography of this Supplemental Material.}
\label{fig:previousmetrics}
\end{figure}

\nocite{logan2015, gares2014, esnard2011, zandbergen2009, horney2016, grabich2016, 
antipova2015post, zahran2010, currie2013, grabich2015, grabich2016hurricane, amstadter2010,
acierno2006, le2013, lieberman2017, mckinney2011, dosa2012evacuate, swerdel2014, 
czajkowski2011, mongin2017, kessler2007hurricane, caillouet2008increase, 
rosenheim2018disaster, domino2003disasters, kim2016, kinney2008, kim2015, tansel2010,
escobedo2009, fuller2014, baggerly2008impact, strobl2011economic, mukherjee2017}

\newpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.7\linewidth]{hurrtracks}
\caption{Tracks of all tropical cyclones included in the study. The study included all tracked storms in the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}) that came within 250 kilometers of at least on U.S. county, 1988--2015. Thicker lines show the tracks of storms whose names were retired, indicating that the storm was particularly severe or had notable impacts \citep{retirednames}.}
\label{fig:hurrtracks}
\end{figure}

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{othertopstorms}
\caption{Counties exposed, based on different exposure metrics, to the storms with largest extent (as measured by the number of counties exposed based on any metric) from each of the clusters shown in the Jaccard heatmap in the main text (Figure 2).}
\label{fig:othertopstorms}
\end{figure}

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{extentdisagreement}
\caption{Examples of storms for which exposure assessments differed substantially between rain-based and wind-based metrics. Counties shown in orange were classified as exposed to a given storm by a given metric; those in yellow were classified as not exposed. The storm's track is shown in red. For both Frances (2004) and Isidore (2002) (top panel), substantially more counties were assessed as exposed when measuring exposure based on rain rather than wind. Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel), substantially more counties were assessed as exposed based on wind compared to rain.}
\label{fig:extentdisagreement}
\end{figure}

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{averageexposureonlysupp}
\caption{Average number of storm exposure per decade in U.S. counties for each exposure metric, based on years of for which data for all five exposure metrics are available (1996--2011). The criteria behind each of the five metrics is given in Table 1 of the main text.}
\label{fig:averageexposuresupp}
\end{figure}

\clearpage

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=15.5cm]{topelecdependexposure}
\caption{Counties with the highest estimated physical exposure among electricity-dependent Medicare beneficiaries exposed to storms per year in U.S. counties for each exposure metric. The criteria behind each of the five metrics is given in Table 1 of the main text and details of the physical exposure calculation are given in the Methods. The color of each bar indicates the number of Medicare beneficiaries in the county reliant on electricity-dependent medical and assistive equipment as of July 2017 \citep{empower}. The length of each bar shows the average expected number of these electricity-dependent Medicare beneficiaries exposed to tropical storms per year based on a given exposure metric.}
\label{fig:topelecdependexposure}
\end{figure*}

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.6\linewidth]{windcomparison}
\caption{Comparison of estimated wind categories from modeled sustained wind speed (as used for the wind exposure metric in the main text; \citep{stormwindmodel}) versus sustained wind speed determined from the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}). In HURDAT2, wind radii were available for study storms from 2004--2015 and gives estimates of the maximum radii to which storm winds of 34 knots, 50 knots, and 64 knots extended in each four quadrant of the storm at each of the six-hour storm track observations. We interpolated this data to 15-minute increments and assessed a county as exposed to winds at or above one of these cutoffs if its population mean center was within 85\% of the maximum radius for that wind speed in its quadrant of the storm. For each county in each of the available storms, we also calculated the modeled wind speed \citep{stormwindmodel} and then categorized this value into the same categories available for the HURDAT2 wind radii ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots). Each point in this graph shows, for a storm, the percent of counties in which the wind category ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots) was the same based on both modeled winds and the HURDAT2 wind radii. Only storms with at least one county with sustained wind of $\ge$34 knots, based on the HURDAT2 wind radii, are shown. Hurricanes Sandy (2012) and Ike (2008) were he two storms with the worst agreement between these two sets of wind data; maps with further comparisons between the two data sources for these two storms are given in Fig. \ref{fig:windexamples}.}
\label{fig:windcomparison}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.9\linewidth]{raincomparison}
\caption{Comparison of county-level rain exposure based on the reanalysis data used for exposure assessment in this study (NLDAS-2) \citep{rui2013nldas, alhamdan2014environmental} and ground-based weather stations \citep{menne2012overview, rnoaa, countyweather}. Each small graph shows a sample county, and each point on the graph represents a tropical storm that came within 500 kilometers of the population mean center of the county. The x-axis shows the estimated cumulative rainfall (in millimeters) from two days before to one day after the storm's closest approach to the county based on the county-aggregated NLDAS-2 precipitation data used to measure rain exposure in our main analysis. The ground-based observations (y-axis) are based on collecting precipitation data from available weather stations in the county, 1988--2011, summed over the same period from two days before to one day after the storm's closest approach for each available station. These cumulative station-based precipitation totals were then averaged across all available stations in the county to create a county-wide estimate of cumulative storm-related precipitation for each storm in the county. Horizontal and vertical lines show the threshold used to classify a storm as exposed based on cumulative rainfall; storms in the lower left and upper right quadrants are classified the same regardless of whether reanalysis or station-based precipitation data is considered, while storms in the upper left and lower right quadrants would be classified differently. The number of storms within each county and the average number of stations reporting rainfall during the county's storms are given above each plot. Note that x- and y-axis ranges differ by county. An estimate of the rank correlation (Spearman's $\rho$ \citep{spearman1904proof}) between measurements from the two rain data sources is also shown on each graph.}
\label{fig:raincomparison}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.8\linewidth]{floodcomparison}
\caption{Comparison of for a sample of study counties of flood status, based on NOAA Storm Events listings \citep{stormevents, noaastormevents}, with total streamflows at county streamflow gages during tropical storms \citep{usgsgages, countyfloods, dataRetrieval}. Each panel represents a separate county. For each county, we first identified all streamflow gages in the county that had complete data for Jan. 1 1996--Dec. 31, 2015. The number of streamflow gages meeting this criterion are given in parentheses beside the county's name in the panel title. This analysis covered tropical storm that came within 500 kilometers of each sample county between 1996 and 2015. In the graphs, each point represents a single tropical storm, and the point's position along the x-axis shows the highest daily total streamflow (cubic feet / second), summed across all identified streamgages in the county, for a five-day window centered on the day of the storm's closest approach to the county. The y-axis separates storms for which a flood event was reported in NOAA's Storm Events database for the county with a start date within the five-day window of the storm's closest approach to the county \citep{stormevents, noaastormevents}. Boxplots are added to show the distribution of the total streamflow measurements across storms in the county for storms grouped by whether or not they were linked with a flood event based on the NOAA Storm Events data. The color of each point gives the percent of streamflow gages in the county with a daily streamflow that exceeded a threshold of flooding (the streamgage's median value for annual peak flow \citep{countyfloods}) on any day during the five-day window.  Note that the x-axis scales differ by county, depending on the number of streamflow gages and typical flow rates for each gage, and are on a log-10 scale.}
\label{fig:floodcomparison}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{windexamples}
\caption{Wind classifications for the two storms, Hurricanes Sandy (2012) and Ike (2008) for which exposure classifications from modeled wind data \citep{stormwindmodel} and HURDAT2 wind radii \citep{demuth2006improvement} showed the highest disagreement for storms between 2004 and 2015 (Fig. \ref{fig:windcomparison}). The top two maps show modeled (left) and HURDAT2 (right) windspeed categorizations for Hurricane Ike (2008), while the bottom two maps show the same windspeed characterizations for Hurricane Sandy (2012). In both cases, the storms were exceptionally large systems for which high winds persisted well inland from landfall and for which, based on the HURDAT2 data, a large number of counties were exposed to winds of $\ge$34 knots (Figure \ref{fig:windcomparison}).}
\label{fig:windexamples}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.9\linewidth]{rainexamples}
\caption{Examples of storms where the rain exposure classification method fails to capture full storm-related rainfall because of the distance constraint used in that assessment metric. The red line on each map shows the track of the storm, based on the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}). The color of each county in the map gives the cumulative precipitation (in millimeters) for that county during the window from two days before to one day after the storm's closest approach to the county. The blue outline identifies the collection of counties that were classified as ``exposed" based on the criteria that the storm passed within 500 kilometers of the county's population mean center and the cumulative rainfall was 75 millimeters or more (Table 1 of main text). The map on the left shows rain-based exposure for Hurricane Ike. For this exceptionally large storm, there are some counties in northern Texas that received high rainfall from the storm but are too far from the storm center's track to be classified as exposed by this metric. The map on the right shows rain-based exposure for Tropical Storm Lee. For this storm, rain from the storm affected counties in the Mid-Atlantic and New England as the remnants of the storm moved over those regions, but storm tracking stopped while the storm was still over the southeastern part of the U.S., and so these counties are not classified as exposed to the storm based on this metric.}
\label{fig:rainexamples}
\end{figure}

\clearpage

\bibliography{hurr_exposure}

\end{document}

### 2019-01-26 writing/ehp_abstract.tex
Add a bunch of stuff

\begin{abstract}

\noindent \textbf{Background:} Tropical cyclones bring severe impacts to U.S.
communities, including substantial human health risks, from hazards that
include wind, rain, flooding, and tornadoes. However, studies quantifying
tropical cyclone risks and impacts vary widely in how they classify
exposure to tropical cyclones, using various hazard-based metrics and, in some
cases, using distance from the storm track as a surrogate. \\
\textbf{Objectives:} (1) Investigate how well county-level assessments of
tropical cyclone exposure agree when based on distance versus hazard-based
measurements, as well as among assessments based on different hazard
measurements, and (2) provide software to improve tropical cyclone exposure
assessment for health-related research. \\ 
\textbf{Methods:} We measured county-level exposure to tropical cyclones in the
United States based on distance from the storm track, maximum sustained wind,
rainfall, flooding, and tornadoes for all land-falling or near-land Atlantic
basin storms, for 1996--2011 for all metrics and up to 1988--2015 for specific
metrics. We quantified how well county-level exposure assessment agreed for all
pairwise combinations of metrics.\\ 
\textbf{Results:} In most of the storms investigated, agreement in county-level
exposure assessment was at best moderate, and often poor, when using distance
compared with four hazard-based measurements. There was also substantial
disagreement in exposure assessment based on all pairwise combinations of the
four hazard-based measurements. There were geographic patterns in exposures
based on these metrics, with most wind-based exposures near the coast, flood-
and rain-based exposures extending to inland and northern counties, and
tornado-based exposures in southern counties. \\ 
\textbf{Discussion:} Our results suggest that when impact studies use distance
as a surrogate for tropical cyclone exposure, or use one hazard-based metric
when the impact is partly or fully caused by a different hazard, the study will
be prone to exposure misclassification.  To facilitate future research, we make
this multi-hazard storm exposure data available through open-source software. 
\end{abstract}

### 2019-01-26 writing/ehp_acknowledgements.tex
Add a bunch of stuff

--- 2018-10-04 ehp_draft.tex
@@
-\documentclass[fleqn,10pt,lineno]{olplainarticle}
-% Use option lineno for line numbers
-
-\graphicspath{{./figures/}}
-
-\usepackage[implicit=false]{hyperref}
-
-\title{Assessing United States county-level exposure for tropical cyclone impact
-research}
-
-\author[a,1]{G. Brooke Anderson} 
-\author[a,b]{Joshua Ferreri}
-\author[c]{Mohammad Al-Hamdan} 
-\author[c]{William Crosson} 
-\author[d]{Andrea Schumacher} 
-\author[e]{Seth Guikema} 
-\author[f]{Steven Quiring} 
-\author[g]{Dirk Eddelbuettel} 
-\author[a]{Meilin Yan} 
-\author[h]{Roger D. Peng}
-
-\affil[a]{Department of Environmental \& Radiological Health Sciences, Colorado 
-  State University, Fort Collins, CO, 80523} 
-\affil[b]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
-\affil[c]{Universities Space Research Association, NASA Marshall Space Flight 
-  Center, Huntsville, AL, 35805}
-\affil[d]{Cooperative Institute for Research in the Atmosphere, Colorado State
-  University, Fort Collins, CO, 80523} 
-\affil[e]{Department of Industrial and Operations Engineering, University of 
-  Michigan, Ann Arbor, MI, 48109}
-\affil[f]{Department of Geography, Ohio State University, Columbus, OH, 43210}
-\affil[g]{Debian and R Projects; Department of Statistics, University of
-  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
-\affil[h]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
-  Health, Baltimore, MD, 21205}
-
-\keywords{Hurricanes, Tropical cyclones, Disaster impacts, Exposure assessment}
-
-\begin{abstract} Hurricanes and other tropical cyclones bring severe impacts to
-U.S. communities. These impacts can result from a variety of storm-related
-hazards, including extreme wind, rain, flooding, and tornadoes. Impact studies
-vary widely in how they classify exposure to tropical cyclones, using various
-hazard-based metrics and, in some cases, using distance from the storm as a
-surrogate for exposure to storm-related hazards. Here we measure county-level
-exposure to tropical cyclones in the United States based on distance from the
-storm, maximum sustained wind, rainfall, flooding, and tornadoes for all
-land-falling or near-land Atlantic basin storms for 1988--2015. We show that the
-locations identified as storm-exposed varied substantially when switching among
-these metrics. For example, most wind-based storm exposures were limited to
-southern counties near the coast, while flood- and rain-based exposures often
-extended to inland and northern counties. We also show that distance to the
-storm served as, at best, a moderate, and often a poor, surrogate in identifying
-exposure to storm-related wind, rain, floods, or tornadoes. Therefore, when
-impact studies use distance as a surrogate for tropical cyclone exposures or use
-one hazard-based metric (e.g., wind-based) when the impact is partly or fully
-caused by a different storm hazard, the analysis will be prone to exposure
-misclassification, which can mask true associations, even strong associations.
-To facilitate future research, we make this multi-hazard storm exposure data
-available through open-source software.  \end{abstract}
-
-\begin{document}
-
-\flushbottom 
-\maketitle 
-\thispagestyle{empty}
-
-\section*{Introduction}
-
-Tropical cyclones, including hurricanes, can have disasterous impacts in the
-United States (U.S.). A tropical cyclone's high winds can bring health risks
-and property damage caused by structural damage of houses and other buildings,
-falling trees, and wind-borne debris \citep{rappaport2000} and 
-cause power outages \citep{liu2005, han2009}, introducing a number of
-threats to human and ecological health, including water quality
-risks if the outage affects wastewater treatment plants \citep{mallin2006}.
-Other risks can exist without severe wind; for example, one study found that
-most of the direct hurricane-related deaths in the U.S. between 1970 and 1999
-occurred in cases when wind was below hurricane strength, including for
-Tropical Storms Charley in 1998 and Alberto in 1994 \citep{rappaport2000}.
-Tropical cyclones can produce excessive rain, especially in certain
-topographies (e.g., near mountains), so counties well inland sometimes
-experience more extreme rain than coastal counties. Flood risks from tropical
-cyclones can result from this rain, although the two risks are not perfectly
-correlated \citep{chen2015}. In the U.S., over half of hurricane-related direct
-deaths from 1970 to 1999 from Atlantic basin storms were a result of freshwater
-flooding \citep{rappaport2000}. Flooding can also degrade water quality
-\citep{mallin2006}, which can threaten both human and ecological health.
-
-To measure tropical cyclone exposure for either investigations of exposure
-patterns or for estimation of storm impacts, including health impacts, studies
-have varied widely in how they determine storm exposure (some examples shown in
-Fig. S1). While many studies have measured exposure based on specific hazards
-of the storm (e.g., wind, rain), many studies have used distance from the
-storm's central track as a surrogate in identifying exposure to the hazards of
-a storm (Fig. S1) (e.g., \citep{czajkowski2011, tansel2010, kinney2008,
-caillouet2008increase}). Distance from the storm's central track can be easily
-measured using widely-available hurricane tracking data. There are, however, a
-number of limitations to using distance to assess exposure to tropical
-cyclones, and there are reasons to suspect that exposure classifications based
-on other storm hazards may disagree with distance-based classifications.
-Tropical cyclones vary dramatically in size: U.S. storms have been observed
-with radii to maximum winds as small as 20 kilometers and as large as 200
-kilometers \citep{mallin2006, quiring2011variations}. While a number of storm
-hazards are strongly associated with distance from the storm's center (e.g.,
-wind and, at the coast, storm surge and waves \citep{rappaport2000, kruk2010}),
-other hazards like dangerous rain, floods, and tornadoes can occur well away
-from the storm's central track \citep{rappaport2000, atallah2007, moore2012}.
-For example, fatal tropical cyclone tornadoes, which were linked to over 300
-deaths in the U.S.  between 1995 and 2009, most often occur 200--500 kilometers
-from the storm's center \citep{moore2012}. Further, distance-based exposure
-metrics that use buffers tend to use an equal buffer distance on each side of
-the storm track (e.g., \citep{czajkowski2011, grabich2015, grabich2016,
-zandbergen2009, tansel2010}), but the forces of a storm tend to be distributed
-around the center in a non-symmetrical way. For example, extreme winds are more
-common to the right of the track where counter-clockwise cyclonic winds move in
-concert with the storm's forward motion \citep{halverson2015}, and the fatal
-tornadoes associated with U.S. tropical cyclones between 1995 and 2009 occurred
-almost exclusively to the right of the storm's track, mostly in the right-front
-quandrant (northeast) of the storm \citep{moore2012}. Rain, conversely, is
-often heaviest to the left of the storm's track, especially when the storm
-interacts with other weather systems \citep{atallah2003, atallah2007,
-zhu2013variations} or undergoes an extratropical transition
-\citep{elsberry2002}.
-
-The use of different exposure metrics across studies hampers the development of
-a scientific consensus on storm-related risks, as differences observed between
-studies could result from differences in exposure assessment.  Further, if a
-study misclassifies exposure to the storm hazard or hazards that cause an
-impact, estimates of storm risks or impacts will be biased, making it hard to
-identify true associations. These concerns are particularly important if there
-are large differences in the locations that are determined to be storm-exposed
-based on different metrics.  To investigate this, we measured county-level
-exposure to tropical cyclones in all eastern United States counties using five
-metrics (Table \ref{tab:exposuremetrics}): (1) distance to storm track; (2)
-maximum sustained wind speed; (3) cumulative rainfall; (4) flood events; and
-(5) tornado events.  We assessed exposure at the county level since data on
-many potential impacts are available at county-level aggregations (e.g., direct
-hurricane-related deaths \citep{czajkowski2011}; birth outcomes
-\citep{grabich2015, grabich2016}; autism prevalence \citep{kinney2008}) and
-since decisions and policies to prepare for, and respond to, storms are often
-undertaken at the county level \citep{zandbergen2009, rappaport2000}. We
-measured how well exposure classification agreed across these five metrics in
-terms of classifying individual counties as exposed to a storm.  Finally, to
-make this hazard-specific tropical cyclone exposure data available to other
-researchers, we published open source software for county-based hurricane
-exposure assessment in U.S. counties \citep{hurricaneexposure}.
-
-\begin{table}%[tbhp] 
-\centering 
-\caption{Exposure metrics considered to assess county-level exposure to 
-tropical cyclones}
-\begin{tabular}{p{0.9cm}p{2.5cm}p{9cm}} 
-Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Tornado & 1996--2015 & Tornado event listed NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
-\bottomrule 
-\end{tabular} 
-\label{tab:exposuremetrics} 
-\end{table}
-
-\section*{Methods and Materials}
-,
-All counties in the eastern half of the U.S. (e.g., Fig.
-\ref{fig:averageexposure}), as of the 2010 U.S. Decennial Census, were included
-in the study. The study covered all tracked storms in the revised Atlantic
-hurricane database (HURDAT2 \citep{landsea2013}) and that came within 250
-kilometers of at least one study county between 1988 and 2015 (Fig. S2). HURDAT2
-is a post-storm assessment of each storm conducted by the U.S. National
-Hurricane Center (NHC) and incorporates data from a variety of sources,
-including satellite data and, when available, aircraft reconnaisance data
-\citep{landsea2013, jarvinen1988}. These data typically give measurements of
-storm center location at 6-hour intervals, at synoptic times (i.e., 6:00 am,
-12:00 pm, 6:00 pm, and 12:00 am UTC); some landfalling storms have an additional
-observation at the time of landfall \citep{landsea2013}.
-
-\subsection*{Distance-based exposure metric.}
-
-For the rain-based exposure metric, cumulative rainfall was estimated for the
-period from two days before to one day after the date of a storm's closest
-approach to a county. Daily county-level rainfall was determined by aggregating
-hourly, 1/8 degree gridded reanalysis data from the North American Land Data
-Assimilation System Phase 2 (NLDAS-2) precipitation data files
-\citep{rui2013nldas}. The NLDAS-2 data integrate satellite-based and land-based
-monitoring and applies a land-surface model to create a reanalysis dataset that
-is spatially and temporally complete across the continental U.S.
-\citep{rui2013nldas, alhamdan2014environmental}. These data have been used in
-previous research to investigate tropical cyclones (e.g., \citep{chen2015}). To
-aggregate to county-level, hourly data at each grid point were summed to create
-a daily sum and these grid-point specific rainfall totals were then averaged
-across all grid points within a county's 1990 U.S. Census boundaries to create a
-county-level daily total \citep{alhamdan2014environmental, cdcwonder}. This
-county-level precipitation data are publicly available through the U.S. Centers
-for Disease Control's Wide-ranging Online Data for Epidemiological Research
-(WONDER) database \citep{cdcwonder, alhamdan2014environmental}. The
-rainfall-based exposure metric was only available for storms through 2011, the
-period for which the reanalysis rainfall data aggregated to county-level daily
-totals was available \citep{cdcwonder, alhamdan2014environmental}. Validation
-analysis of this and other hazard-based exposure data is given in the Supporting
-Information.
-
-\subsection*{Wind-based exposure metric}
-
-Ground-based observations of wind speed are problematic during storms, as
-instruments often fail at high wind speed, while reanalysis data, often
-available at hourly or higher resolution, can be too smooth to capture wind
-extremes associated with a storm. Therefore, to create a dataset of county-level
-sustained winds during historical storms, we modeled maximum sustained wind
-speeds at each county's population mean center \citep{countycenters}. HURDAT2
-Best Track data were interpolated from 6-hourly reported values to 15-minute
-increments using natural cubic splines \citep{stormwindmodel}, and then
-sustained wind speed was modeled for each county center at each 15-minute
-increment using a double exponential wind speed model
-\citep{willoughby2006parametric} to estimate maximum ground-level sustained wind
-speed at each county center \citep{stormwindmodel}. Asymmetry in wind speeds
-around the storm center was incorporated into the wind speed model
-\citep{phadke2003modeling, stormwindmodel}. The maximum value of this modeled
-wind speed was determined for each county as the maximum windspeed across the
-15-minute incremented modeled values throughout the storm.
-
-\subsection*{Flood- and tornado-based exposure metrics}
-
-To identify flood- and tornado-based exposures to tropical cyclones in U.S.
-counties, we used event listings from the National Oceanic and Atmospheric
-Administration (NOAA)'s Storm Event Database \citep{stormevents}. For each
-storm, we identified all events with event types related to flooding (``Flood",
-``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and that occurred
-in a county within 500 kilometers of the storm's track and within a five-day
-window centered on the date of the storm's closest approach to the county
-\citep{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado" events in
-this database were reported by county Federal Information Processing Standard
-(FIPS) code. ``Coastal Flood" events were reported by forecast zone; for these,
-the event was matched to the appropriate county if possible
-\citep{noaastormevents}. While this database has recorded storm data since 1950,
-its coverage changed substantially in 1996 to cover a wider variety of storm
-events \citep{stormevents}. We therefore only considered these metrics of
-hurricane exposure for storms in 1996 and later.
-
-\subsection*{Assessing agreement in exposure classification across exposure
-metrics}
-
-Two of the exposure metrics (flood- and tornado-based) were inherently binary;
-for the other exposure metrics, each county was classified as exposed to a storm
-based on whether the continuously-measured exposure metric exceeded a certain
-threshold (Table \ref{tab:exposuremetrics}). We assessed county-level tropical
-cyclone exposure in the eastern U.S. based on each exposure metrics (Table
-\ref{tab:exposuremetrics}) for all land-falling or near-land Atlantic basin
-storms. Depending on available exposure data, this assessment included some or
-all of the period from 1988 to 2015, a period with 136 tropical cyclones that
-made U.S. landfall or passed within 250 kilometers (155 miles) of at least one
-U.S. county (Fig. S2).
-
-We then measured agreement between exposure metrics in the classification of a
-county as exposed or unexposed to a specific storm. For this measurement, we
-calculated the within-storm Jaccard index \citep{jaccard1901distribution,
-jaccard1908nouvelles} between every pair-wise combination of hazard metrics.
-The Jaccard index ($J_s$) measures similarity between two metrics ($X_{1,s}$
-and $X_{2,s}$) for storm $s$ as the proportion of counties in which both of the
-metrics classify the county as exposed out of all counties classified as
-exposed by at least one of the metrics:
-
-\begin{equation} 
-J_s = \frac{X_{1,s} \cap X_{2,s}}{X_{1,s} \cup X_{2,s}}
-\end{equation}
-
-This metric can range from 0, in the case of no overlap between the counties
-classified as ``exposed" based on the two metrics, to 1, in the case that the
-two metrics classify exactly the same counties as ``exposed" to the storm.
-
-\section*{Results}
-
-\begin{figure*}%[tbhp] 
-\centering 
-\includegraphics[width=16cm]{ivanonly}
-\caption{Counties classified as exposed to Hurricane Ivan (2004) under each
-exposure metric (Table 1). The red line shows the track of Hurricane Ivan 
-based on the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
-Similar maps for other large-extent storms are given in Fig. S5.}
-\label{fig:ivanexposure} 
-\end{figure*}
-
-\subsection*{Patterns in tropical cyclone exposures in eastern U.S. counties}
-
-Table [x] provides summary statistics on ...  Table [x] includes some measures
-of the number of counties exposed to the storm based on each exposure metric
-(Tab. [x]) Median storm extent was largest when exposure was classified based
-on distance (median storm extent: 62 counties exposed). The measured extent of
-storms also tended to be higher when using exposure metrics based on rain
-(median: 21 counties exposed) or wind (median: 26 counties exposed) versus
-based on floods (median: 9 counties exposed) or tornadoes (median: 1 county
-exposed).  For every metric except the tornado-based metric, we identified at
-least one storm with over 300 counties exposed. However, the largest-extent
-storm varied by metric (e.g., Beryl [1994] exposed the most counties based on
-distance, Frances [2004] based on rain, and Ike [2008] based on wind).
-
-\subsection*{County-specific exposure classification}
-
-\begin{figure}%[tbhp] 
-\centering 
-\includegraphics[width = 0.7\linewidth]{jaccard_heatmap} 
-\caption{Heatmap of Jaccard index values for
-specific exposure metric pairs within storms. Only storms between 1996 and 2011,
-and for which at least 250 counties were exposed based on at least one metric,
-are included. The color of each cell within the main heatmap indicates the value
-of the Jaccard index (proportion of counties classified as exposed by both
-metrics out of storms classified as exposed by either metric) for a given pair
-of metrics for a given storm. Storms are displayed within clusters that have
-similar patterns in county-level exposure agreement for metric pairs, based on
-hierarchical clustering using the complete link method
-\citep{murtagh2012algorithms} (i.e., storms in the same cluster tend to have
-similar patterns for the pairwise strength of agreement among metrics); columns
-are also ordered based on hierarchical clustering. The colors to the right of
-the main heatmap for each storm indicate the total number of counties classified
-as exposed to the storm by any of the five metrics, providing an estimate of
-storm extent. Maps are available showing the counties identified as exposed
-under each of five metrics for the widest-extent storm in each cluster:
-Hurricane Ivan (2004) (Fig. \ref{fig:ivanexposure}) and Hurricanes Floyd (1999),
-Lee (2011), Cindy (2005), and Katrina (2005) (Fig. S5).} 
-\label{fig:jaccard}
-\end{figure}
-
-We found that switching among these metrics typically resulted in large
-differences in which counties were identified as exposed to a particular storm.
-Fig. \ref{fig:ivanexposure} gives an example for Hurricane Ivan (2004), a
-record-breaking storm in terms of its duration as a major hurricane
-\citep{franklin2006atlantic}. When county-level exposure was determined based on
-the wind metric, only counties near two of the storm's landfalls were assessed
-as exposed. For rain- and flood-based metrics, however, exposure extended to the
-left of the track, including counties as far north as New York and Connecticut,
-while for the tornado metric, exposed counties tended to be to the right of the
-track and included several counties in central North Carolina, South Carolina,
-and Georgia that were not identified as exposed to Ivan based on any other
-metric.
-
-We drew similar conclusions when we investigated all 46 storms between 1996 and
-2011 (the period for which all five metrics were available) for which 250 or
-more counties were exposed based on at least one metric (Fig.
-\ref{fig:jaccard}). The tornado-based metric showed universally poor agreement
-with other metrics in county-level classification across all storms considered.
-For other pairs of metrics, there were also generally large differences in
-which counties were determined to be exposed to the storm, with the Jaccard
-index values (see Methods) below 0.4 for most metric pairs in most storms.
-
-There were, however, a few exceptions---storms in which similar counties were
-determined to be exposed to the storm for two or more of the metrics
-considered.  For Floyd (1999) and Irene (2011), for example, county-level
-classification agreed moderately to well (Jaccard index of approximately
-0.5--0.8) for all pairs of exposure metrics except those including the
-tornado-based metric. These storms both made their first U.S. landfall in North
-Carolina at minor hurricane windspeeds (Category 2 and 1, respectively) and
-then skimmed the eastern coast of the U.S. north through New England, bringing
-large rainfall to much of the eastern coast from North Carolina north and
-causing extensive inland flooding in North Carolina (Floyd) and New England
-(Irene) \citep{avila2013atlantic, lawrence2000atlantic}. For another set of
-storms (e.g., Lee [2011], Ernesto [2006], and Bertha [1996]), there was
-moderate to good agreement for all pairwise combinations of distance, rain, and
-wind, but poor agreement for all other combinations of metrics. Most of these
-storms either skimmed along the eastern coast north of North Carolina (Bertha
-[1996], Charley [2004], Gaston [2004], Barry [2007], Hanna [2008]), in a
-pattern similar to Hurricanes Floyd (1999) and Irene (2011), or made landfall
-in or near the Florida panhandle and cut across to exit into the Atlantic
-around North Carolina (Josephine [1996], Earl [1998], Alberto [2006]). However,
-these two sets of storms represented the minority of all storms considered; for
-most of the storms, there was low overlap in the counties determined to be
-exposed to the storm for most pairings of metrics.
-
-\begin{figure*}%[tbhp] 
-\centering
-\includegraphics[width=16cm]{averageexposureonly} 
-\caption{Average number of storm exposures per decade in U.S. counties for 
-each exposure metric. The criteria behind each of the five metrics is given 
-in Table \ref{tab:exposuremetrics}. The years used to estimate these averages 
-are based on years of available exposure data (distance and wind: 1988--2015; 
-rain: 1988--2011; flood and tornado: 1996--2015). Similar patterns persist when
-analysis is restricted to years with all exposure data available (1996--2011;
-Fig. S3).} 
-\label{fig:averageexposure} 
-\end{figure*}
-
-\subsection*{Geographic patterns in tropical cyclone exposures}
-
-When different metrics were used to assess average tropical cyclone exposure in
-U.S. counties, geographic patterns varied substantially; in fact, some areas of
-the U.S. that experienced regular exposures to storm-related excessive rainfall
-and floods were not identified as areas of regular risk based on more commonly
-used assessment metrics like wind or distance (Figs. \ref{fig:averageexposure},
-S3). Wind-based exposure had a strong coastal pattern, with almost all
-exposures in counties within about 200 kilometers (124 miles) of the coastline,
-mostly in eastern North and South Carolina, southern Florida, and the area near
-the outlet of the Mississippi River. While exposures by rain- and
-distance-based metrics were also more common in coastal areas compared to
-inland areas, inland exposures were much more common based on these two metrics
-compared to the wind-based metric. Rain and, to some extent, flood exposures
-were characterized by a pattern defined by the Appalachian Mountains, with much
-lower exposures west of the mountain range than to the east.  Flood- and,
-especially, tornado-based exposures were much less common than exposures based
-on other metrics. Almost all tornado-based exposures were in coastal states,
-with most in Florida, Alabama, South Carolina, and North Carolina and almost
-none north of Maryland. Flood-based exposures were highest in North Carolina
-and along the South Carolina coast, as well as further north in New Jersey,
-southeastern Pennsylvania, and southeastern New York.
-
-We have included a case study in the Supporting Information showing how these
-differences can strongly influence estimates of which U.S. counties have the
-largest expected physical exposure to tropical cyclones among a susceptible
-subpopulation, Medicare beneficiaries who are reliant on electricity for
-medical equipment \citep{empower}.
-
-\subsection*{Software}
-
-To assist with future tropical cyclone impact studies, we created and published
-open source software that includes the hazard-specific, county-level tropical
-cyclone data in this paper, as well as tools to explore and map that data
-\citep{hurricaneexposure, hurricaneexposuredata}.  We investigated how well
-these data correspond with data from other available sources (Figs. S6--8).
-While generally in agreement with data from other sources, there are a few
-limitations. The wind data are based on modeled, rather than observed, values,
-and while the modeled wind data generally agree well with post-analyis maximum
-wind radii from the revised Atlantic hurricane database (HURDAT2
-\citep{landsea2013}) (Fig. S6), for a few storms it did not (e.g., Fig.  S9).
-The rainfall data are from re-analysis data, which are generally
-well-correlated with observed ground-based station data but may oversmooth
-extreme measurements (Fig. S7).
-
-\section*{Discussion}
-
-Tropical cyclones impact public health in many ways, and epidemiological studies
-of the health impacts of tropical cyclone exposures could help improve
-preparedness for and response to future storms. However, tropical storms are
-multi-hazard events, making it more complicated to measure exposure to tropical
-cyclones compared to other weather exposures like temperature and precipitation.
-... Our results can inform exposure assessment for future county-level studies
-of the health impacts of tropical cyclones exposures, and the data and software
-presented provides multi-hazard, county-level exposure data for multiple
-tropical cyclone hazards for such studies.
-
-Wind-based exposure had a strong coastal pattern, with almost all exposures in
-counties within about 200 kilometers (124 miles) of the coastline, mostly in
-eastern North and South Carolina, southern Florida, and the area near the
-outlet of the Mississippi River. This is consistent with the dramatic decrease
-in wind intensity that typically characterizes the landfall of tropical
-cyclones. While exposures by rain- and distance-based metrics were also more
-common in coastal areas compared to inland areas, inland exposures were much
-more common based on these two metrics compared to the wind-based metric. Rain
-and, to some extent, flood exposures were characterized by a pattern defined by
-the Appalachian Mountains, with much lower exposures west of the mountain range
-than to the east. This agrees with previous research indicating that the
-Appalachian mountains' topography both enhances precipitation during tropical
-cyclones and provides hydrological conditions for severe flooding
-\citep{rees2001}.  Almost all tornado-based exposures were in coastal states,
-with most in Florida, Alabama, South Carolina, and North Carolina and almost
-none north of Maryland. This pattern may be linked with previous findings that
-most noteworthy storm-related tornadoes occur on the right-side of the storm
-track in Atlantic Basin U.S. storms \citep{moore2012}.
-
-The tropical cyclone exposure averages shown in Fig. \ref{fig:averageexposure}
-are limited as estimates of long-term frequencies, as tropical cyclones follow
-decadal patterns \citep{kossin2007more} likely not adequately captured in the
-available data. However, these frequency maps do illustrate the potential for
-strong differences in spatial patterns in tropical cyclone exposures, depending
-on which storm hazards are considered. A few previous studies have sought to
-determine county-level exposure to tropical cyclones over multi-year periods,
-including \citep{zandbergen2009}, which estimated exposure in U.S. counties to
-all U.S. landfalling Atlantic basin tropical cyclones between 1851 and 2003,
-using both a distance-based metric and a metric that combined distance and
-windspeed, and \citep{kruk2010}, which explored exposure to hurricane-related
-winds in the U.S., including inland areas, for 1900--2008. Our results suggest
-that such exposure assessments may perform well in capturing some storm hazards
-(e.g., wind), but likely miss other potentially dangerous tropical cyclone
-exposures, especially for hazards that repeatedly threaten northern or inland
-counties (e.g., rain, flooding). We have included a case study in the Supporting
-Information showing how these differences can strongly influence estimates of
-which U.S. counties have the largest expected physical exposure to tropical
-cyclones among a susceptible subpopulation, Medicare beneficiaries who are
-reliant on electricity for medical equipment \citep{empower}.
-
-Median storm extent was largest when exposure was classified based on distance,
-likely because this metric includes counties along or near the paths of less
-severe storms, as well as inland counties near a storm that has weakened
-substantially.
-
-Previous research supports the idea that some storms may differ dramatically in
-extent based on which hazards of the storm are of interest. While storm rainfall
-and windspeed may be well-correlated when the storm is over water
-\citep{cerveny2000}, this relationship does not remain as strong once the
-hurricane has made landfall \citep{jiang2008}. Fast-moving storms bring higher
-risks of dangerous winds inland \citep{kruk2010}, while slow-moving storms are
-likely to bring more rain \citep{rappaport2000} and cause more damage because of
-sustained hazardous conditions \citep{rezapour2014}. Further, while the
-likelihood and extent of flooding during a tropical cyclone is related to the
-storm's rainfall, it is also driven by factors like top soil saturation and the
-structure of the water basin's drainage network \citep{chen2015, rees2001}.
-
-The tornado-based metric showed universally poor agreement
-with other metrics in county-level classification across all storms considered.
-This likely results from tornado-based exposure being very rare, even during
-large storms (Fig. \ref{fig:severityagreement}).
-
-For both sets of storms for which agreement between metrics were unusually high
-(e.g., Floyd (1999), Irene (2011), ...), the storms' persistent proximity to
-water may have helped maintain wind speeds in similar patterns to rain and
-distance exposures, resulting in this moderate to good agreement among exposure
-assessments based on different metrics.
-
-Distance from a tropical cyclone's track is relatively easy to measure and has
-been used as an operational metric of exposure to tropical cyclones in previous
-large-scale studies (Fig. S1). Since distance itself does not constitute a
-hazard, distance is meant in these cases as a surrogate to capture exposure to
-hazards from the storm. However, here we found that in assessing U.S.
-county-level exposure to tropical cyclones, distance is, at best, a moderate,
-and often a very poor, surrogate for exposure to the specific storm hazards of
-high wind, extreme rainfall, flooding, and tornadoes (Fig. \ref{fig:jaccard}).
-Therefore, use of distance to assess tropical cyclone exposure for impact
-studies could result in problematic exposure misclassification, which could mask
-true associations, even strong associations, between storm exposure and outcomes
-of interest in impact studies \citep{savitz2016interpreting,
-armstrong1998effect}. In some cases, this exposure misclassification may be
-differential (i.e., associated with the outcome of interest or with factors
-associated with risk of the outcome of interest). For example, tropical cyclone
-wind exposures tend to be concentrated in counties near the coast, since most
-storms rapidly decrease in sustained windspeed following landfall. However,
-tropical cyclone exposures based on distance can extend well inland, following
-the storm's tracks, but may not adequately capture all wind-exposed counties
-near the coast. In this case, if the etiologically-relevant exposure is high
-wind but exposure is classified based on distance, the probabilty of being
-misclassified as unexposed would be higher in coastal counties, while the
-probability of being misclassified as exposed would be higher in inland
-counties. If coastal counties differ from inland counties in either the outcome
-of interest or in factors associated with risk of that outcome, differential
-exposure misclassification would exist \citep{savitz2016interpreting}. Such
-differential exposure misclassification can bias estimates of storm effects
-either towards the null (estimating a lower or null association compared to the
-true association that exists) or away from the null (estimating a larger
-association than actually exists) \citep{savitz2016interpreting,
-armstrong1998effect}. The use of a single hazard-based metric (e.g., wind) could
-cause similar problems if the impact is driven, at least in part, by a different
-hazard or by multiple hazards of the storm.
-
-\subsection*{Software}
-
-To assist with future tropical cyclone impact studies, we created and published
-open source software that includes the hazard-specific, county-level tropical
-cyclone data in this paper, as well as tools to explore and map that data
-\citep{hurricaneexposure, hurricaneexposuredata}. Many previous studies have
-used geographical information system software (e.g., ArcGIS) to assess exposure
-to tropical cyclones in the U.S. \citep{grabich2016, zandbergen2009,
-czajkowski2011, kruk2010}. Here, we offer methods to map and output historic
-exposure to tropical cyclones that does not require the use of proprietary
-software but instead uses a package written in the R statistical programming
-language \citep{R}, which is free and open-source.
-
-In addition to providing tropical cyclone exposure metrics for individual
-hazards, this software and data also allow users to create storm exposure
-profiles based on multiple hazards or craft exposure indices that combine hazard
-metrics \citep{chakraborty2005population, peduzzi2009assessing}. This ability
-can be critical, as different hazards of extreme storms often act
-synergistically in their impacts \citep{smith2009}. Further, by including
-measurements of different hazard exposures in each county for each storm, this
-software allows for the development of more complex models that incoroporate
-multiple potential predictors of an impact (e.g., random forests, multivariable
-generalized linear models).
-
-In creating this hurricane exposure dataset, we aimed for data that are
-reliable, available for all eastern U.S. counties, and have no missing data
-within available dates. Further, we have investigated how well these data
-correspond with data from other available sources (Figs. S6--8). However, there
-are still some important limitations. The wind data are based on modeled, rather
-than observed, values, and while the modeled wind data generally agree well with
-post-analyis maximum wind radii from the revised Atlantic hurricane database
-(HURDAT2 \citep{landsea2013}) (Fig. S6), for a few storms it did not (e.g., Fig.
-S9). The rainfall data are from re-analysis data, which are generally
-well-correlated with observed ground-based station data but may oversmooth
-extreme measurements (Fig. S7). The flood and tornado data came from the
-National Oceanic and Atmospheric Administration's (NOAA's) Storm Events
-database, which, while a widely-used database of events maintained by NOAA, is
-based on reports, and so may be prone to underreporting \citep{Ashley2008flood,
-Curran2000}, especially in less populated areas \citep{Witt1998, Ashley2007}, as
-well as to other reporting errors. While these are important limitations, we
-selected these data sources as among the best currently available for measuring
-each of these hazards at a multi-county, multi-year scale.
-
-\subsection*{Conclusions}
-
-Previous research has highlighted the range of impacts that tropical cyclones
-can have in U.S. communities (e.g., \cite{pielke1997, rappaport2000, liu2005,
-han2009, mallin2006}). Determining the metrics of storm exposure that are most
-associated with loss of life, property damage, and other impacts may help
-identify and quantify the important threats that remain from tropical cyclones
-in the U.S., which could allow future success in reducing these threats through
-community planning, warning systems, and other measures. Here we found large
-differences in which counties are exposed to different hazards of tropical
-cyclones and that distance is, at best, a moderate, and often a very poor,
-surrogate for exposure to the specific storm hazards of high wind, extreme
-rainfall, flooding, and tornadoes. Use of distance as a surrogate for any of
-these hazards could lead to exposure misclassification and, in the case of storm
-impact studies, result in biased storm risk estimates. Similarly, if studies
-disagree on their findings of tropical cyclone impacts, it could result from the
-often poor agreement between exposure classifications based on these different
-metrics, making such studies that use different metrics hard to compare or
-combine in meta-analyses.
-
 \section*{Acknowledgments}
@@
 (GES) Data and Information Services Center (DISC).
-
-\bibliography{hurr_exposure}
-
-\end{document}

### 2019-01-26 writing/ehp_cover.tex
Add a bunch of stuff

--- 2019-01-26 writing/anderson_ehp_supp.tex
@@
-\documentclass[11pt, titlepage]{article}
-\usepackage{microtype}
-\usepackage{charter}
-
-\input{ehp_preamble}
-
-\begin{document}
-
-\title{Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"}
+\title{Assessing United States county-level exposure for research on tropical cyclones and human health}
 
@@
   Health, Baltimore, MD, 21205}
-\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}
+\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu,
+  1681 Campus Delivery, Fort Collins, CO, 80523}
 
 \date{}
-
-\maketitle
-
-\beginsupplement
-
-\section*{Supplemental Text}
-
-\subsection*{Case study: Physical exposure of electricity-dependent Medicare beneficiaries to tropical cyclones}
-
-The outcomes of disasters depend both on the geophysical forces of the disaster as well as on the vulnerability of the society living in the geographical areas affected by those forces.
-\citep{chakraborty2005population, anderson2003community, cutter1996vulnerability} As a case study of how differences in tropical cyclone exposure assessments across different hazard metrics influence estimates of physical exposures of susceptible populations to tropical cyclones, we investigated how the use of different metrics influenced estimates of which eastern U.S.
-counties have the highest expected average exposures of electricity-dependent Medicare beneficiaries to tropical cyclones.
-This subpopulation was selected for this analysis since it may be particularly susceptible to health impacts from tropical cyclone exposure, especially through the pathway of storm-associated power outages and evacuations.
-We collected data on the number of electricity-dependent Medicare beneficiaries in each study county from the U.S.
-Department of Health \& Human Service's emPower Map 2.0 \citep{empower}.
-We then calculated the physical exposure of the electricity-dependent Medicare population in each county, based on tropical cyclone assessments using each hazard metric, following \citep{peduzzi2009assessing}:
-
-\begin{equation}
-E_c = F_c * P_c
-\end{equation}
-
-\noindent where $E_c$ is the average yearly physical exposure among electricity-dependent Medicare beneficiaries in county $c$ to tropical cyclone exposures based on a given metric, $F_c$ is the estimated yearly expected frequency of tropical cyclone exposures in county $c$ based on that metric, and $P_c$ is the size of the electricity-dependent Medicare population in county $c$, as of July 2017.
-When combined with estimates of vulnerability of a population to a natural hazard, such measurements of physical exposure can be used to calculate risk of human losses from the hazard \citep{peduzzi2009assessing}.
-
-We found that a few counties (Miami-Dade County, FL; Harris County, TX) were ranked in the top three counties of physical exposure among electricity-dependent Medicare beneficiaries for almost all exposure metrics (Figure \ref{fig:topelecdependexposure}).
-However, there were key differences when using different exposure metrics.
-For example, Florida counties dominated the lists of top exposed counties based on wind- and tornado-based metrics, while over half of the counties for the flood-based metric were in Mid-Atlantic states (PA, NY).
-
-Dissimilarity in exposure to specific hazards of tropical cyclones can result in differing patterns in estimates of county-level exposure (Fig.
-1 of main text), which can lead to substantial differences in estimates of population physical exposure (Fig.
-\ref{fig:topelecdependexposure}), based on which storm hazard is considered.
-The measure of a hurricane's intensity, the Saffir-Simpson scale, expresses intensity as a function of wind speed, and so fails to capture potential risks from other hazards of the storm \citep{smith2009}, and previous assessments of physical exposure to tropical cyclones have often focused on winds or distance from the storm.
-For example, a global study of exposure and vulnerability to natural hazards used a combination of windspeed above a certain threshold and distance from the storm's track to assess exposure \citep{peduzzi2009assessing}.
-While many storm impacts might be most strongly linked to wind hazards, there is a growing recognition of the potential risks of adverse health outcomes and property damage from rain- and flood-related tropical cyclone hazards in the U.S.
-\citep{smith2009}.
-Such assessments may underemphasize potentially dangerous exposures to certain regions of the U.S., particularly inland locations.
-In fact, based on our assessment of physical exposures to tropical cyclones among electricity-dependent Medicare beneficiaries, the highest expected physical exposure under any metric was measured for Philadelphia County, PA, under the flood metric (Fig.
-\ref{fig:topelecdependexposure}).
-While this county has a similar-sized population of electricity-dependent Medicare beneficiaries as Miami-Dade County, FL (Fig.
-\ref{fig:topelecdependexposure}), the frequency of storm-related flood events in this inland county results in a very high expected physical exposure among this subpopulation, while this county is not included in the top ten counties for physical exposure based on most of the other exposure metrics.
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{previous_exposure_metrics}
-\caption{Metrics used in a sample of previous studies to assign exposure to tropical cyclones in U.S.-based tropical cyclone exposure or impact studies.
-Citations for each study are shown on the left and the broad topic of the study on the left.
-Presence of a bar for a study-metric combination indicates that the given metric was used in the study in assessing exposure to tropical cyclones for that study, either alone or in combination with other metrics for which that study has bars.
-The color transparency of each bar indicates the number of different thresholds considered for defining storm exposure in the study under that hazard (e.g., less transparent bars indicate that a study considered multiple thresholds of the metric in defining tropical cyclone exposure as a method of sensitivity analysis).
-Studies for which there are two or more bars (e.g., bars for both distance and wind) indicate studies that either used separate exposure assessments, one for each metric, (e.g., as a sensitivity analysis) or used a single exposure index that incorporated multiple metrics in its definition.
-``Damage" often indicate use of Federal Emergency Management Agency (FEMA) reports, although occasionally used damage reports from other sources, including newspaper reports.
-Examples of ``Other" include evacuation orders, excessive school closures, length of time a storm was near a location, and detailed, holistic analyses of a specific storm or storms.
-References for each study are included in the
-bibliography of this Supplemental Material.}
-\label{fig:previousmetrics}
-\end{figure}
-
-\nocite{logan2015, gares2014, esnard2011, zandbergen2009, horney2016, grabich2016, 
-antipova2015post, zahran2010, currie2013, grabich2015, grabich2016hurricane, amstadter2010,
-acierno2006, le2013, lieberman2017, mckinney2011, dosa2012evacuate, swerdel2014, 
-czajkowski2011, mongin2017, kessler2007hurricane, caillouet2008increase, 
-rosenheim2018disaster, domino2003disasters, kim2016, kinney2008, kim2015, tansel2010,
-escobedo2009, fuller2014, baggerly2008impact, strobl2011economic, mukherjee2017}
-
-\newpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=0.7\linewidth]{hurrtracks}
-\caption{Tracks of all tropical cyclones included in the study.
-The study included all tracked storms in the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}) that came within 250 kilometers of at least on U.S.
-county, 1988--2015.
-Thicker lines show the tracks of storms whose names were retired, indicating that the storm was particularly severe or had notable impacts \citep{retirednames}.}
-\label{fig:hurrtracks}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{othertopstorms}
-\caption{Counties exposed, based on different exposure metrics, to the storms with largest extent (as measured by the number of counties exposed based on any metric) from each of the clusters shown in the Jaccard heatmap in the main text (Figure 2).}
-\label{fig:othertopstorms}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{extentdisagreement}
-\caption{Examples of storms for which exposure assessments differed substantially between rain-based and wind-based metrics.
-Counties shown in orange were classified as exposed to a given storm by a given metric; those in yellow were classified as not exposed.
-The storm's track is shown in red.
-For both Frances (2004) and Isidore (2002) (top panel), substantially more counties were assessed as exposed when measuring exposure based on rain rather than wind.
-Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel), substantially more counties were assessed as exposed based on wind compared to rain.}
-\label{fig:extentdisagreement}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{averageexposureonlysupp}
-\caption{Average number of storm exposure per decade in U.S.
-counties for each exposure metric, based on years of for which data for all five exposure metrics are available (1996--2011).
-The criteria behind each of the five metrics is given in Table 1 of the main text.}
-\label{fig:averageexposuresupp}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=15.5cm]{topelecdependexposure}
-\caption{Counties with the highest estimated physical exposure among electricity-dependent Medicare beneficiaries exposed to storms per year in U.S.
-counties for each exposure metric.
-The criteria behind each of the five metrics is given in Table 1 of the main text and details of the physical exposure calculation are given in the Methods.
-The color of each bar indicates the number of Medicare beneficiaries in the county reliant on electricity-dependent medical and assistive equipment as of July 2017 \citep{empower}.
-The length of each bar shows the average expected number of these electricity-dependent Medicare beneficiaries exposed to tropical storms per year based on a given exposure metric.}
-\label{fig:topelecdependexposure}
-\end{figure*}
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=0.6\linewidth]{windcomparison}
-\caption{Comparison of estimated wind categories from modeled sustained wind speed (as used for the wind exposure metric in the main text; \citep{stormwindmodel}) versus sustained wind speed determined from the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
-In HURDAT2, wind radii were available for study storms from 2004--2015 and gives estimates of the maximum radii to which storm winds of 34 knots, 50 knots, and 64 knots extended in each four quadrant of the storm at each of the six-hour storm track observations.
-We interpolated this data to 15-minute increments and assessed a county as exposed to winds at or above one of these cutoffs if its population mean center was within 85\% of the maximum radius for that wind speed in its quadrant of the storm.
-For each county in each of the available storms, we also calculated the modeled wind speed \citep{stormwindmodel} and then categorized this value into the same categories available for the HURDAT2 wind radii ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots).
-Each point in this graph shows, for a storm, the percent of counties in which the wind category ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots) was the same based on both modeled winds and the HURDAT2 wind radii.
-Only storms with at least one county with sustained wind of $\ge$34 knots, based on the HURDAT2 wind radii, are shown.
-Hurricanes Sandy (2012) and Ike (2008) were he two storms with the worst agreement between these two sets of wind data; maps with further comparisons between the two data sources for these two storms are given in Fig.
-\ref{fig:windexamples}.}
-\label{fig:windcomparison}
-\end{figure}
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=0.9\linewidth]{raincomparison}
-\caption{Comparison of county-level rain exposure based on the reanalysis data used for exposure assessment in this study (NLDAS-2) \citep{rui2013nldas, alhamdan2014environmental} and ground-based weather stations \citep{menne2012overview, rnoaa, countyweather}.
-Each small graph shows a sample county, and each point on the graph represents a tropical storm that came within 500 kilometers of the population mean center of the county.
-The x-axis shows the estimated cumulative rainfall (in millimeters) from two days before to one day after the storm's closest approach to the county based on the county-aggregated NLDAS-2 precipitation data used to measure rain exposure in our main analysis.
-The ground-based observations (y-axis) are based on collecting precipitation data from available weather stations in the county, 1988--2011, summed over the same period from two days before to one day after the storm's closest approach for each available station.
-These cumulative station-based precipitation totals were then averaged across all available stations in the county to create a county-wide estimate of cumulative storm-related precipitation for each storm in the county.
-Horizontal and vertical lines show the threshold used to classify a storm as exposed based on cumulative rainfall; storms in the lower left and upper right quadrants are classified the same regardless of whether reanalysis or station-based precipitation data is considered, while storms in the upper left and lower right quadrants would be classified differently.
-The number of storms within each county and the average number of stations reporting rainfall during the county's storms are given above each plot.
-Note that x- and y-axis ranges differ by county.
-An estimate of the rank correlation (Spearman's $\rho$ \citep{spearman1904proof}) between measurements from the two rain data sources is also shown on each graph.}
-\label{fig:raincomparison}
-\end{figure}
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=0.8\linewidth]{floodcomparison}
-\caption{Comparison of for a sample of study counties of flood status, based on NOAA Storm Events listings \citep{stormevents, noaastormevents}, with total streamflows at county streamflow gages during tropical storms \citep{usgsgages, countyfloods, dataRetrieval}.
-Each panel represents a separate county.
-For each county, we first identified all streamflow gages in the county that had complete data for Jan.
-1 1996--Dec.
-31, 2015.
-The number of streamflow gages meeting this criterion are given in parentheses beside the county's name in the panel title.
-This analysis covered tropical storm that came within 500 kilometers of each sample county between 1996 and 2015.
-In the graphs, each point represents a single tropical storm, and the point's position along the x-axis shows the highest daily total streamflow (cubic feet / second), summed across all identified streamgages in the county, for a five-day window centered on the day of the storm's closest approach to the county.
-The y-axis separates storms for which a flood event was reported in NOAA's Storm Events database for the county with a start date within the five-day window of the storm's closest approach to the county \citep{stormevents, noaastormevents}.
-Boxplots are added to show the distribution of the total streamflow measurements across storms in the county for storms grouped by whether or not they were linked with a flood event based on the NOAA Storm Events data.
-The color of each point gives the percent of streamflow gages in the county with a daily streamflow that exceeded a threshold of flooding (the streamgage's median value for annual peak flow \citep{countyfloods}) on any day during the five-day window.
-Note that the x-axis scales differ by county, depending on the number of streamflow gages and typical flow rates for each gage, and are on a log-10 scale.}
-\label{fig:floodcomparison}
-\end{figure}
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{windexamples}
-\caption{Wind classifications for the two storms, Hurricanes Sandy (2012) and Ike (2008) for which exposure classifications from modeled wind data \citep{stormwindmodel} and HURDAT2 wind radii \citep{demuth2006improvement} showed the highest disagreement for storms between 2004 and 2015 (Fig.
-\ref{fig:windcomparison}).
-The top two maps show modeled (left) and HURDAT2 (right) windspeed categorizations for Hurricane Ike (2008), while the bottom two maps show the same windspeed characterizations for Hurricane Sandy (2012).
-In both cases, the storms were exceptionally large systems for which high winds persisted well inland from landfall and for which, based on the HURDAT2 data, a large number of counties were exposed to winds of $\ge$34 knots (Figure \ref{fig:windcomparison}).}
-\label{fig:windexamples}
-\end{figure}
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=0.9\linewidth]{rainexamples}
-\caption{Examples of storms where the rain exposure classification method fails to capture full storm-related rainfall because of the distance constraint used in that assessment metric.
-The red line on each map shows the track of the storm, based on the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
-The color of each county in the map gives the cumulative precipitation (in millimeters) for that county during the window from two days before to one day after the storm's closest approach to the county.
-The blue outline identifies the collection of counties that were classified as ``exposed" based on the criteria that the storm passed within 500 kilometers of the county's population mean center and the cumulative rainfall was 75 millimeters or more (Table 1 of main text).
-The map on the left shows rain-based exposure for Hurricane Ike.
-For this exceptionally large storm, there are some counties in northern Texas that received high rainfall from the storm but are too far from the storm center's track to be classified as exposed by this metric.
-The map on the right shows rain-based exposure for Tropical Storm Lee.
-For this storm, rain from the storm affected counties in the Mid-Atlantic and New England as the remnants of the storm moved over those regions, but storm tracking stopped while the storm was still over the southeastern part of the U.S., and so these counties are not classified as exposed to the storm based on this metric.}
-\label{fig:rainexamples}
-\end{figure}
-
-\clearpage
-
-\bibliography{hurr_exposure}
-
-\end{document}

### 2019-01-26 writing/ehp_discussion.tex
Add a bunch of stuff

\section*{Discussion}

Tropical cyclones impact public health in many ways, and epidemiological
studies of the health risk and impacts associated with tropical cyclone
exposures could help improve preparedness for and response to future tropical
cyclones.  However, tropical cyclones are multi-hazard events, making it more
complicated to measure exposure compared to other weather
exposures like temperature and precipitation. Here, we find the county-level
tropical cyclone exposure assessments vary substantially when using a
distance-based metric versus four hazard-based metrics, as well as among
different hazard-based metrics. Our results can inform exposure assessment for
future county-level studies of the health risks and impacts associated with tropical cyclones
exposures, and the data and software we have created to accompany this article
provides multi-hazard, county-level exposure data for multiple tropical cyclone
hazards for such studies.

Distance from a tropical cyclone's track is relatively easy to measure and has
been used as an operational metric of exposure to tropical cyclones in previous
large-scale studies (examples included in Figure S1). Since distance itself
does not constitute a hazard, distance is meant in these cases as a surrogate
to capture exposure to hazards from the tropical cyclone. However, here we
found that in assessing U.S.  county-level exposure to tropical cyclones,
distance is, at best, a moderate, and often a very poor, surrogate for exposure
to the specific tropical cyclone hazards of high wind, extreme rainfall,
flooding, and tornadoes (Figure \ref{fig:jaccard}).  Therefore, use of distance
to assess tropical cyclone exposure for impact studies could result in
problematic exposure misclassification, which could mask true associations,
even strong associations, between tropical cyclone exposure and outcomes of
interest in impact studies \citep{savitz2016interpreting, armstrong1998effect}.

In some cases, this exposure misclassification may be differential (i.e.,
associated with the outcome of interest or with factors associated with risk of
the outcome of interest). For example, tropical cyclone wind exposures tend to
be concentrated in counties near the coast, since most tropical cyclones
rapidly decrease in sustained windspeed following landfall. However, tropical
cyclone exposures based on distance can extend well inland, following the
tropical cyclone's tracks, but may not adequately capture all wind-exposed
counties near the coast. In this case, if the etiologically-relevant exposure
is high wind but exposure is classified based on distance, the probability of
being misclassified as unexposed would be higher in coastal counties, while the
probability of being misclassified as exposed would be higher in inland
counties. If coastal counties differ from inland counties in either the outcome
of interest or in factors associated with risk of that outcome, differential
exposure misclassification would exist \citep{savitz2016interpreting}. Such
differential exposure misclassification can bias estimates of tropical cyclone
effects either towards the null (estimating a lower or null association
compared to the true association that exists) or away from the null (estimating
a larger association than actually exists) \citep{savitz2016interpreting,
armstrong1998effect}. The use of a single hazard-based metric (e.g., wind)
could cause similar problems if the impact is driven, at least in part, by a
different hazard or by multiple hazards of the tropical cyclone.

Our findings---that county-level tropical cyclone exposure assessment varies
substantially when different exposure measurements are considered---aligns with
previous research from atmospheric science and related fields on the
characteristics of tropical cyclones. While tropical cyclone rainfall and
windspeed are often well-correlated when the tropical cyclone is over water
\citep{cerveny2000}, this relationship often does not remain as strong once the
hurricane has made landfall \citep{jiang2008}.  Fast-moving tropical cyclones
bring higher risks of dangerous winds inland \citep{kruk2010}, while
slow-moving tropical cyclone are likely to bring more rain
\citep{rappaport2000} and cause more damage because of sustained hazardous
conditions \citep{rezapour2014}. Further, while the likelihood and extent of
flooding during a tropical cyclone is related to the tropical cyclone's
rainfall, it is also driven by factors like top soil saturation and the
structure of the water basin's drainage network \citep{chen2015, rees2001}.
However, we did find a small set of tropical cyclone for which for which
agreement between metrics were unusually high (e.g., Floyd in 1999, Irene in
2011, Hannah in 2008, Bertha in 1996) For this set of tropical cyclones, the
tropical cyclones' persistent proximity to water may have helped maintain wind
speeds in similar patterns to rain and distance exposures, resulting in this
moderate to good agreement among exposure assessments based on different
metrics.

In our results, wind-based exposure had a strong coastal pattern, with almost
all exposures in counties within about 200 kilometers (124 miles) of the
coastline. This is consistent with the
dramatic decrease in wind intensity that typically characterizes the landfall
of tropical cyclones. Exposures by rain- and distance-based metrics often extended further inland compared to the wind-based
metric, up to the Appalachian mountains. This agrees with previous research
indicating that the Appalachian mountains' topography both enhances
precipitation during tropical cyclones and provides hydrological conditions for
severe flooding \citep{rees2001}.  Almost all tornado-based exposures were in
southern coastal states. This pattern echoes
previous findings that most noteworthy tropical cyclone-related tornadoes occur
on the right-side of the tropical cyclone track in Atlantic Basin U.S. storms
\citep{moore2012}.

The tropical cyclone exposure averages shown in Figure
\ref{fig:averageexposure} are limited as estimates of long-term frequencies, as
tropical cyclones follow decadal patterns \citep{kossin2007more} likely not
adequately captured in the available data. However, these frequency maps,
together with evidence from specific tropical cyclones (Figures
\ref{fig:ivanexposure} and \ref{fig:jaccard}),  do illustrate the potential for
strong differences in spatial patterns in tropical cyclone exposures, depending
on which tropical cyclone hazards are considered. A few previous studies have
sought to determine county-level exposure to tropical cyclones over multi-year
periods, including \citet{zandbergen2009}, which estimated exposure in U.S.
counties to all U.S.  landfalling Atlantic basin tropical cyclones between 1851
and 2003, using both a distance-based metric and a metric that combined
distance and windspeed, and \citet{kruk2010}, which explored exposure to
hurricane-related winds in the U.S., including inland areas, for 1900--2008.
Our results suggest that such exposure assessments may perform well in
capturing some tropical cyclone hazards (e.g., wind), but likely miss other
potentially dangerous tropical cyclone exposures, especially for hazards that
repeatedly threaten northern or inland counties (e.g., rain, flooding).

\subsection*{Software}

To assist with future tropical cyclone studies, we created and published
open source software to accompany this article \citep{hurricaneexposure,
hurricaneexposuredata}. Many previous studies have used geographical
information system software (e.g., ArcGIS) to assess exposure to tropical
cyclones in the U.S. \citep{grabich2016, zandbergen2009, czajkowski2011,
kruk2010}. Here, we offer methods to map and output historic exposure to
tropical cyclones that does not require the use of proprietary software but
instead uses a package written in the R statistical programming language
\citep{R}, which is free and open-source.

In addition to providing tropical cyclone exposure metrics for individual
hazards, this software and data also allow users to create tropical cyclone
exposure profiles based on multiple hazards or craft exposure indices that
combine hazard metrics \citep{chakraborty2005population, peduzzi2009assessing}.
This ability can be critical, as different hazards of tropical cyclones
often act synergistically in causing impacts \citep{smith2009}. Further, by
including measurements of different hazard exposures in each county for each
tropical cyclone, this software allows for the development of more complex
exposure indices or models (e.g., random forests, multivariable generalized
linear models) that incorporate multiple tropical cyclone hazard measurements.

In creating this hurricane exposure dataset, we aimed for data that are
reliable, available for all eastern U.S. counties, and have no missing data
within available dates. Further, we have investigated how well these data
correspond with data from other available sources (Figures S7--11). However,
there are still some important limitations. For example, for a few tropical cyclones, 
the model-based wind data might differ from post-storm analysis of wind radii
(e.g., Figure S10), and the re-analysis rainfall data seems to sometime oversmooth
very extreme precipitation (Figure S8). 
The flood and tornado data
came from the National Oceanic and Atmospheric Administration's (NOAA's) Storm
Events database, which, while a widely-used database of events maintained by
NOAA, is based on reports, and so may be prone to underreporting
\citep{Ashley2008flood, Curran2000}, especially in less populated areas
\citep{Witt1998, Ashley2007}, as well as to other reporting errors. While these
are important limitations, we selected these data sources as among the best
currently available for measuring each of these hazards at a multi-county,
multi-year scale.

\subsection*{Conclusions}

Previous research has highlighted the range of impacts that tropical cyclones
can have in U.S. communities. However, studies have varied widely in how they assess
exposure to tropical cyclones for exposure, health impact, and other impact
studies, in some cases using distance to the tropical cyclone's track as a
surrogate metric of exposure to a tropical cyclone's hazards. Here we found
large differences in which counties are exposed to different hazards of
tropical cyclones and that distance is, at best, a moderate, and often a very
poor, surrogate for exposure to the specific tropical cyclone hazards of high
wind, extreme rainfall, flooding, and tornadoes. Use of distance as a surrogate
for any of these hazards could lead to exposure misclassification and, in the
case of tropical cyclone risk and impact studies, including epidemiological
studies, result in biased estimates.  Similarly, if such studies disagree on
their findings, it could result from the often poor agreement between exposure
classifications based on these different metrics, making studies that use
different metrics hard to compare or combine in meta-analyses. Our findings
highlight the importance of clarifying the potential pathway from tropical
cyclone hazards to health impacts when conducting tropical cyclone
epidemiological studies and basing exposure assessment on measurements of these
hazards. To help with this, we provide extensive tropical cyclone
hazard-related measurements for a large collection of historical Atlantic-basin
cyclones, aggregated at the county level to align with the spatial scale at
which much health-related data is collected.

### 2019-01-26 writing/ehp_draft.tex
Add a bunch of stuff

--- 2018-10-04 ehp_draft.tex
@@
 
-\graphicspath{{./figures/}}
+\graphicspath{{../figures/}}
 
 \usepackage[implicit=false]{hyperref}
+\usepackage{acronym}
 
-\title{Assessing United States county-level exposure for tropical cyclone impact
-research}
+\title{Assessing United States county-level exposure for research on tropical cyclones and human health}
 
@@
 
-\begin{abstract} Hurricanes and other tropical cyclones bring severe impacts to
-U.S. communities. These impacts can result from a variety of storm-related
-hazards, including extreme wind, rain, flooding, and tornadoes. Impact studies
-vary widely in how they classify exposure to tropical cyclones, using various
-hazard-based metrics and, in some cases, using distance from the storm as a
-surrogate for exposure to storm-related hazards. Here we measure county-level
-exposure to tropical cyclones in the United States based on distance from the
-storm, maximum sustained wind, rainfall, flooding, and tornadoes for all
-land-falling or near-land Atlantic basin storms for 1988--2015. We show that the
-locations identified as storm-exposed varied substantially when switching among
-these metrics. For example, most wind-based storm exposures were limited to
-southern counties near the coast, while flood- and rain-based exposures often
-extended to inland and northern counties. We also show that distance to the
-storm served as, at best, a moderate, and often a poor, surrogate in identifying
-exposure to storm-related wind, rain, floods, or tornadoes. Therefore, when
-impact studies use distance as a surrogate for tropical cyclone exposures or use
-one hazard-based metric (e.g., wind-based) when the impact is partly or fully
-caused by a different storm hazard, the analysis will be prone to exposure
-misclassification, which can mask true associations, even strong associations.
-To facilitate future research, we make this multi-hazard storm exposure data
-available through open-source software.  \end{abstract}
+\begin{abstract} 
+  \input{ehp_abstract}
+\end{abstract}
 
@@
 
-\section*{Introduction}
-
-Tropical cyclones, including hurricanes, can have disasterous impacts in the
-United States (U.S.). A tropical cyclone's high winds can bring health risks
-and property damage caused by structural damage of houses and other buildings,
-falling trees, and wind-borne debris \citep{rappaport2000} and 
-cause power outages \citep{liu2005, han2009}, introducing a number of
-threats to human and ecological health, including water quality
-risks if the outage affects wastewater treatment plants \citep{mallin2006}.
-Other risks can exist without severe wind; for example, one study found that
-most of the direct hurricane-related deaths in the U.S. between 1970 and 1999
-occurred in cases when wind was below hurricane strength, including for
-Tropical Storms Charley in 1998 and Alberto in 1994 \citep{rappaport2000}.
-Tropical cyclones can produce excessive rain, especially in certain
-topographies (e.g., near mountains), so counties well inland sometimes
-experience more extreme rain than coastal counties. Flood risks from tropical
-cyclones can result from this rain, although the two risks are not perfectly
-correlated \citep{chen2015}. In the U.S., over half of hurricane-related direct
-deaths from 1970 to 1999 from Atlantic basin storms were a result of freshwater
-flooding \citep{rappaport2000}. Flooding can also degrade water quality
-\citep{mallin2006}, which can threaten both human and ecological health.
-
-To measure tropical cyclone exposure for either investigations of exposure
-patterns or for estimation of storm impacts, including health impacts, studies
-have varied widely in how they determine storm exposure (some examples shown in
-Fig. S1). While many studies have measured exposure based on specific hazards
-of the storm (e.g., wind, rain), many studies have used distance from the
-storm's central track as a surrogate in identifying exposure to the hazards of
-a storm (Fig. S1) (e.g., \citep{czajkowski2011, tansel2010, kinney2008,
-caillouet2008increase}). Distance from the storm's central track can be easily
-measured using widely-available hurricane tracking data. There are, however, a
-number of limitations to using distance to assess exposure to tropical
-cyclones, and there are reasons to suspect that exposure classifications based
-on other storm hazards may disagree with distance-based classifications.
-Tropical cyclones vary dramatically in size: U.S. storms have been observed
-with radii to maximum winds as small as 20 kilometers and as large as 200
-kilometers \citep{mallin2006, quiring2011variations}. While a number of storm
-hazards are strongly associated with distance from the storm's center (e.g.,
-wind and, at the coast, storm surge and waves \citep{rappaport2000, kruk2010}),
-other hazards like dangerous rain, floods, and tornadoes can occur well away
-from the storm's central track \citep{rappaport2000, atallah2007, moore2012}.
-For example, fatal tropical cyclone tornadoes, which were linked to over 300
-deaths in the U.S.  between 1995 and 2009, most often occur 200--500 kilometers
-from the storm's center \citep{moore2012}. Further, distance-based exposure
-metrics that use buffers tend to use an equal buffer distance on each side of
-the storm track (e.g., \citep{czajkowski2011, grabich2015, grabich2016,
-zandbergen2009, tansel2010}), but the forces of a storm tend to be distributed
-around the center in a non-symmetrical way. For example, extreme winds are more
-common to the right of the track where counter-clockwise cyclonic winds move in
-concert with the storm's forward motion \citep{halverson2015}, and the fatal
-tornadoes associated with U.S. tropical cyclones between 1995 and 2009 occurred
-almost exclusively to the right of the storm's track, mostly in the right-front
-quandrant (northeast) of the storm \citep{moore2012}. Rain, conversely, is
-often heaviest to the left of the storm's track, especially when the storm
-interacts with other weather systems \citep{atallah2003, atallah2007,
-zhu2013variations} or undergoes an extratropical transition
-\citep{elsberry2002}.
-
-The use of different exposure metrics across studies hampers the development of
-a scientific consensus on storm-related risks, as differences observed between
-studies could result from differences in exposure assessment.  Further, if a
-study misclassifies exposure to the storm hazard or hazards that cause an
-impact, estimates of storm risks or impacts will be biased, making it hard to
-identify true associations. These concerns are particularly important if there
-are large differences in the locations that are determined to be storm-exposed
-based on different metrics.  To investigate this, we measured county-level
-exposure to tropical cyclones in all eastern United States counties using five
-metrics (Table \ref{tab:exposuremetrics}): (1) distance to storm track; (2)
-maximum sustained wind speed; (3) cumulative rainfall; (4) flood events; and
-(5) tornado events.  We assessed exposure at the county level since data on
-many potential impacts are available at county-level aggregations (e.g., direct
-hurricane-related deaths \citep{czajkowski2011}; birth outcomes
-\citep{grabich2015, grabich2016}; autism prevalence \citep{kinney2008}) and
-since decisions and policies to prepare for, and respond to, storms are often
-undertaken at the county level \citep{zandbergen2009, rappaport2000}. We
-measured how well exposure classification agreed across these five metrics in
-terms of classifying individual counties as exposed to a storm.  Finally, to
-make this hazard-specific tropical cyclone exposure data available to other
-researchers, we published open source software for county-based hurricane
-exposure assessment in U.S. counties \citep{hurricaneexposure}.
+\input{ehp_introduction}
 
@@
 
-\section*{Methods and Materials}
-,
-All counties in the eastern half of the U.S. (e.g., Fig.
-\ref{fig:averageexposure}), as of the 2010 U.S. Decennial Census, were included
-in the study. The study covered all tracked storms in the revised Atlantic
-hurricane database (HURDAT2 \citep{landsea2013}) and that came within 250
-kilometers of at least one study county between 1988 and 2015 (Fig. S2). HURDAT2
-is a post-storm assessment of each storm conducted by the U.S. National
-Hurricane Center (NHC) and incorporates data from a variety of sources,
-including satellite data and, when available, aircraft reconnaisance data
-\citep{landsea2013, jarvinen1988}. These data typically give measurements of
-storm center location at 6-hour intervals, at synoptic times (i.e., 6:00 am,
-12:00 pm, 6:00 pm, and 12:00 am UTC); some landfalling storms have an additional
-observation at the time of landfall \citep{landsea2013}.
-
-\subsection*{Distance-based exposure metric.}
-
-For the rain-based exposure metric, cumulative rainfall was estimated for the
-period from two days before to one day after the date of a storm's closest
-approach to a county. Daily county-level rainfall was determined by aggregating
-hourly, 1/8 degree gridded reanalysis data from the North American Land Data
-Assimilation System Phase 2 (NLDAS-2) precipitation data files
-\citep{rui2013nldas}. The NLDAS-2 data integrate satellite-based and land-based
-monitoring and applies a land-surface model to create a reanalysis dataset that
-is spatially and temporally complete across the continental U.S.
-\citep{rui2013nldas, alhamdan2014environmental}. These data have been used in
-previous research to investigate tropical cyclones (e.g., \citep{chen2015}). To
-aggregate to county-level, hourly data at each grid point were summed to create
-a daily sum and these grid-point specific rainfall totals were then averaged
-across all grid points within a county's 1990 U.S. Census boundaries to create a
-county-level daily total \citep{alhamdan2014environmental, cdcwonder}. This
-county-level precipitation data are publicly available through the U.S. Centers
-for Disease Control's Wide-ranging Online Data for Epidemiological Research
-(WONDER) database \citep{cdcwonder, alhamdan2014environmental}. The
-rainfall-based exposure metric was only available for storms through 2011, the
-period for which the reanalysis rainfall data aggregated to county-level daily
-totals was available \citep{cdcwonder, alhamdan2014environmental}. Validation
-analysis of this and other hazard-based exposure data is given in the Supporting
-Information.
-
-\subsection*{Wind-based exposure metric}
-
-Ground-based observations of wind speed are problematic during storms, as
-instruments often fail at high wind speed, while reanalysis data, often
-available at hourly or higher resolution, can be too smooth to capture wind
-extremes associated with a storm. Therefore, to create a dataset of county-level
-sustained winds during historical storms, we modeled maximum sustained wind
-speeds at each county's population mean center \citep{countycenters}. HURDAT2
-Best Track data were interpolated from 6-hourly reported values to 15-minute
-increments using natural cubic splines \citep{stormwindmodel}, and then
-sustained wind speed was modeled for each county center at each 15-minute
-increment using a double exponential wind speed model
-\citep{willoughby2006parametric} to estimate maximum ground-level sustained wind
-speed at each county center \citep{stormwindmodel}. Asymmetry in wind speeds
-around the storm center was incorporated into the wind speed model
-\citep{phadke2003modeling, stormwindmodel}. The maximum value of this modeled
-wind speed was determined for each county as the maximum windspeed across the
-15-minute incremented modeled values throughout the storm.
-
-\subsection*{Flood- and tornado-based exposure metrics}
-
-To identify flood- and tornado-based exposures to tropical cyclones in U.S.
-counties, we used event listings from the National Oceanic and Atmospheric
-Administration (NOAA)'s Storm Event Database \citep{stormevents}. For each
-storm, we identified all events with event types related to flooding (``Flood",
-``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and that occurred
-in a county within 500 kilometers of the storm's track and within a five-day
-window centered on the date of the storm's closest approach to the county
-\citep{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado" events in
-this database were reported by county Federal Information Processing Standard
-(FIPS) code. ``Coastal Flood" events were reported by forecast zone; for these,
-the event was matched to the appropriate county if possible
-\citep{noaastormevents}. While this database has recorded storm data since 1950,
-its coverage changed substantially in 1996 to cover a wider variety of storm
-events \citep{stormevents}. We therefore only considered these metrics of
-hurricane exposure for storms in 1996 and later.
-
-\subsection*{Assessing agreement in exposure classification across exposure
-metrics}
-
-Two of the exposure metrics (flood- and tornado-based) were inherently binary;
-for the other exposure metrics, each county was classified as exposed to a storm
-based on whether the continuously-measured exposure metric exceeded a certain
-threshold (Table \ref{tab:exposuremetrics}). We assessed county-level tropical
-cyclone exposure in the eastern U.S. based on each exposure metrics (Table
-\ref{tab:exposuremetrics}) for all land-falling or near-land Atlantic basin
-storms. Depending on available exposure data, this assessment included some or
-all of the period from 1988 to 2015, a period with 136 tropical cyclones that
-made U.S. landfall or passed within 250 kilometers (155 miles) of at least one
-U.S. county (Fig. S2).
-
-We then measured agreement between exposure metrics in the classification of a
-county as exposed or unexposed to a specific storm. For this measurement, we
-calculated the within-storm Jaccard index \citep{jaccard1901distribution,
-jaccard1908nouvelles} between every pair-wise combination of hazard metrics.
-The Jaccard index ($J_s$) measures similarity between two metrics ($X_{1,s}$
-and $X_{2,s}$) for storm $s$ as the proportion of counties in which both of the
-metrics classify the county as exposed out of all counties classified as
-exposed by at least one of the metrics:
-
-\begin{equation} 
-J_s = \frac{X_{1,s} \cap X_{2,s}}{X_{1,s} \cup X_{2,s}}
-\end{equation}
-
-This metric can range from 0, in the case of no overlap between the counties
-classified as ``exposed" based on the two metrics, to 1, in the case that the
-two metrics classify exactly the same counties as ``exposed" to the storm.
-
-\section*{Results}
+\input{ehp_methods}
 
@@
 
-\subsection*{Patterns in tropical cyclone exposures in eastern U.S. counties}
-
-Table [x] provides summary statistics on ...  Table [x] includes some measures
-of the number of counties exposed to the storm based on each exposure metric
-(Tab. [x]) Median storm extent was largest when exposure was classified based
-on distance (median storm extent: 62 counties exposed). The measured extent of
-storms also tended to be higher when using exposure metrics based on rain
-(median: 21 counties exposed) or wind (median: 26 counties exposed) versus
-based on floods (median: 9 counties exposed) or tornadoes (median: 1 county
-exposed).  For every metric except the tornado-based metric, we identified at
-least one storm with over 300 counties exposed. However, the largest-extent
-storm varied by metric (e.g., Beryl [1994] exposed the most counties based on
-distance, Frances [2004] based on rain, and Ike [2008] based on wind).
-
-\subsection*{County-specific exposure classification}
+\input{../tables/stormseverity.tex}
 
@@
 
-We found that switching among these metrics typically resulted in large
-differences in which counties were identified as exposed to a particular storm.
-Fig. \ref{fig:ivanexposure} gives an example for Hurricane Ivan (2004), a
-record-breaking storm in terms of its duration as a major hurricane
-\citep{franklin2006atlantic}. When county-level exposure was determined based on
-the wind metric, only counties near two of the storm's landfalls were assessed
-as exposed. For rain- and flood-based metrics, however, exposure extended to the
-left of the track, including counties as far north as New York and Connecticut,
-while for the tornado metric, exposed counties tended to be to the right of the
-track and included several counties in central North Carolina, South Carolina,
-and Georgia that were not identified as exposed to Ivan based on any other
-metric.
-
-We drew similar conclusions when we investigated all 46 storms between 1996 and
-2011 (the period for which all five metrics were available) for which 250 or
-more counties were exposed based on at least one metric (Fig.
-\ref{fig:jaccard}). The tornado-based metric showed universally poor agreement
-with other metrics in county-level classification across all storms considered.
-For other pairs of metrics, there were also generally large differences in
-which counties were determined to be exposed to the storm, with the Jaccard
-index values (see Methods) below 0.4 for most metric pairs in most storms.
-
-There were, however, a few exceptions---storms in which similar counties were
-determined to be exposed to the storm for two or more of the metrics
-considered.  For Floyd (1999) and Irene (2011), for example, county-level
-classification agreed moderately to well (Jaccard index of approximately
-0.5--0.8) for all pairs of exposure metrics except those including the
-tornado-based metric. These storms both made their first U.S. landfall in North
-Carolina at minor hurricane windspeeds (Category 2 and 1, respectively) and
-then skimmed the eastern coast of the U.S. north through New England, bringing
-large rainfall to much of the eastern coast from North Carolina north and
-causing extensive inland flooding in North Carolina (Floyd) and New England
-(Irene) \citep{avila2013atlantic, lawrence2000atlantic}. For another set of
-storms (e.g., Lee [2011], Ernesto [2006], and Bertha [1996]), there was
-moderate to good agreement for all pairwise combinations of distance, rain, and
-wind, but poor agreement for all other combinations of metrics. Most of these
-storms either skimmed along the eastern coast north of North Carolina (Bertha
-[1996], Charley [2004], Gaston [2004], Barry [2007], Hanna [2008]), in a
-pattern similar to Hurricanes Floyd (1999) and Irene (2011), or made landfall
-in or near the Florida panhandle and cut across to exit into the Atlantic
-around North Carolina (Josephine [1996], Earl [1998], Alberto [2006]). However,
-these two sets of storms represented the minority of all storms considered; for
-most of the storms, there was low overlap in the counties determined to be
-exposed to the storm for most pairings of metrics.
-
 \begin{figure*}%[tbhp] 
@@
 
-\subsection*{Geographic patterns in tropical cyclone exposures}
+\input{ehp_results}
 
-When different metrics were used to assess average tropical cyclone exposure in
-U.S. counties, geographic patterns varied substantially; in fact, some areas of
-the U.S. that experienced regular exposures to storm-related excessive rainfall
-and floods were not identified as areas of regular risk based on more commonly
-used assessment metrics like wind or distance (Figs. \ref{fig:averageexposure},
-S3). Wind-based exposure had a strong coastal pattern, with almost all
-exposures in counties within about 200 kilometers (124 miles) of the coastline,
-mostly in eastern North and South Carolina, southern Florida, and the area near
-the outlet of the Mississippi River. While exposures by rain- and
-distance-based metrics were also more common in coastal areas compared to
-inland areas, inland exposures were much more common based on these two metrics
-compared to the wind-based metric. Rain and, to some extent, flood exposures
-were characterized by a pattern defined by the Appalachian Mountains, with much
-lower exposures west of the mountain range than to the east.  Flood- and,
-especially, tornado-based exposures were much less common than exposures based
-on other metrics. Almost all tornado-based exposures were in coastal states,
-with most in Florida, Alabama, South Carolina, and North Carolina and almost
-none north of Maryland. Flood-based exposures were highest in North Carolina
-and along the South Carolina coast, as well as further north in New Jersey,
-southeastern Pennsylvania, and southeastern New York.
+\input{ehp_discussion}
 
-We have included a case study in the Supporting Information showing how these
-differences can strongly influence estimates of which U.S. counties have the
-largest expected physical exposure to tropical cyclones among a susceptible
-subpopulation, Medicare beneficiaries who are reliant on electricity for
-medical equipment \citep{empower}.
-
-\subsection*{Software}
-
-To assist with future tropical cyclone impact studies, we created and published
-open source software that includes the hazard-specific, county-level tropical
-cyclone data in this paper, as well as tools to explore and map that data
-\citep{hurricaneexposure, hurricaneexposuredata}.  We investigated how well
-these data correspond with data from other available sources (Figs. S6--8).
-While generally in agreement with data from other sources, there are a few
-limitations. The wind data are based on modeled, rather than observed, values,
-and while the modeled wind data generally agree well with post-analyis maximum
-wind radii from the revised Atlantic hurricane database (HURDAT2
-\citep{landsea2013}) (Fig. S6), for a few storms it did not (e.g., Fig.  S9).
-The rainfall data are from re-analysis data, which are generally
-well-correlated with observed ground-based station data but may oversmooth
-extreme measurements (Fig. S7).
-
-\section*{Discussion}
-
-Tropical cyclones impact public health in many ways, and epidemiological studies
-of the health impacts of tropical cyclone exposures could help improve
-preparedness for and response to future storms. However, tropical storms are
-multi-hazard events, making it more complicated to measure exposure to tropical
-cyclones compared to other weather exposures like temperature and precipitation.
-... Our results can inform exposure assessment for future county-level studies
-of the health impacts of tropical cyclones exposures, and the data and software
-presented provides multi-hazard, county-level exposure data for multiple
-tropical cyclone hazards for such studies.
-
-Wind-based exposure had a strong coastal pattern, with almost all exposures in
-counties within about 200 kilometers (124 miles) of the coastline, mostly in
-eastern North and South Carolina, southern Florida, and the area near the
-outlet of the Mississippi River. This is consistent with the dramatic decrease
-in wind intensity that typically characterizes the landfall of tropical
-cyclones. While exposures by rain- and distance-based metrics were also more
-common in coastal areas compared to inland areas, inland exposures were much
-more common based on these two metrics compared to the wind-based metric. Rain
-and, to some extent, flood exposures were characterized by a pattern defined by
-the Appalachian Mountains, with much lower exposures west of the mountain range
-than to the east. This agrees with previous research indicating that the
-Appalachian mountains' topography both enhances precipitation during tropical
-cyclones and provides hydrological conditions for severe flooding
-\citep{rees2001}.  Almost all tornado-based exposures were in coastal states,
-with most in Florida, Alabama, South Carolina, and North Carolina and almost
-none north of Maryland. This pattern may be linked with previous findings that
-most noteworthy storm-related tornadoes occur on the right-side of the storm
-track in Atlantic Basin U.S. storms \citep{moore2012}.
-
-The tropical cyclone exposure averages shown in Fig. \ref{fig:averageexposure}
-are limited as estimates of long-term frequencies, as tropical cyclones follow
-decadal patterns \citep{kossin2007more} likely not adequately captured in the
-available data. However, these frequency maps do illustrate the potential for
-strong differences in spatial patterns in tropical cyclone exposures, depending
-on which storm hazards are considered. A few previous studies have sought to
-determine county-level exposure to tropical cyclones over multi-year periods,
-including \citep{zandbergen2009}, which estimated exposure in U.S. counties to
-all U.S. landfalling Atlantic basin tropical cyclones between 1851 and 2003,
-using both a distance-based metric and a metric that combined distance and
-windspeed, and \citep{kruk2010}, which explored exposure to hurricane-related
-winds in the U.S., including inland areas, for 1900--2008. Our results suggest
-that such exposure assessments may perform well in capturing some storm hazards
-(e.g., wind), but likely miss other potentially dangerous tropical cyclone
-exposures, especially for hazards that repeatedly threaten northern or inland
-counties (e.g., rain, flooding). We have included a case study in the Supporting
-Information showing how these differences can strongly influence estimates of
-which U.S. counties have the largest expected physical exposure to tropical
-cyclones among a susceptible subpopulation, Medicare beneficiaries who are
-reliant on electricity for medical equipment \citep{empower}.
-
-Median storm extent was largest when exposure was classified based on distance,
-likely because this metric includes counties along or near the paths of less
-severe storms, as well as inland counties near a storm that has weakened
-substantially.
-
-Previous research supports the idea that some storms may differ dramatically in
-extent based on which hazards of the storm are of interest. While storm rainfall
-and windspeed may be well-correlated when the storm is over water
-\citep{cerveny2000}, this relationship does not remain as strong once the
-hurricane has made landfall \citep{jiang2008}. Fast-moving storms bring higher
-risks of dangerous winds inland \citep{kruk2010}, while slow-moving storms are
-likely to bring more rain \citep{rappaport2000} and cause more damage because of
-sustained hazardous conditions \citep{rezapour2014}. Further, while the
-likelihood and extent of flooding during a tropical cyclone is related to the
-storm's rainfall, it is also driven by factors like top soil saturation and the
-structure of the water basin's drainage network \citep{chen2015, rees2001}.
-
-The tornado-based metric showed universally poor agreement
-with other metrics in county-level classification across all storms considered.
-This likely results from tornado-based exposure being very rare, even during
-large storms (Fig. \ref{fig:severityagreement}).
-
-For both sets of storms for which agreement between metrics were unusually high
-(e.g., Floyd (1999), Irene (2011), ...), the storms' persistent proximity to
-water may have helped maintain wind speeds in similar patterns to rain and
-distance exposures, resulting in this moderate to good agreement among exposure
-assessments based on different metrics.
-
-Distance from a tropical cyclone's track is relatively easy to measure and has
-been used as an operational metric of exposure to tropical cyclones in previous
-large-scale studies (Fig. S1). Since distance itself does not constitute a
-hazard, distance is meant in these cases as a surrogate to capture exposure to
-hazards from the storm. However, here we found that in assessing U.S.
-county-level exposure to tropical cyclones, distance is, at best, a moderate,
-and often a very poor, surrogate for exposure to the specific storm hazards of
-high wind, extreme rainfall, flooding, and tornadoes (Fig. \ref{fig:jaccard}).
-Therefore, use of distance to assess tropical cyclone exposure for impact
-studies could result in problematic exposure misclassification, which could mask
-true associations, even strong associations, between storm exposure and outcomes
-of interest in impact studies \citep{savitz2016interpreting,
-armstrong1998effect}. In some cases, this exposure misclassification may be
-differential (i.e., associated with the outcome of interest or with factors
-associated with risk of the outcome of interest). For example, tropical cyclone
-wind exposures tend to be concentrated in counties near the coast, since most
-storms rapidly decrease in sustained windspeed following landfall. However,
-tropical cyclone exposures based on distance can extend well inland, following
-the storm's tracks, but may not adequately capture all wind-exposed counties
-near the coast. In this case, if the etiologically-relevant exposure is high
-wind but exposure is classified based on distance, the probabilty of being
-misclassified as unexposed would be higher in coastal counties, while the
-probability of being misclassified as exposed would be higher in inland
-counties. If coastal counties differ from inland counties in either the outcome
-of interest or in factors associated with risk of that outcome, differential
-exposure misclassification would exist \citep{savitz2016interpreting}. Such
-differential exposure misclassification can bias estimates of storm effects
-either towards the null (estimating a lower or null association compared to the
-true association that exists) or away from the null (estimating a larger
-association than actually exists) \citep{savitz2016interpreting,
-armstrong1998effect}. The use of a single hazard-based metric (e.g., wind) could
-cause similar problems if the impact is driven, at least in part, by a different
-hazard or by multiple hazards of the storm.
-
-\subsection*{Software}
-
-To assist with future tropical cyclone impact studies, we created and published
-open source software that includes the hazard-specific, county-level tropical
-cyclone data in this paper, as well as tools to explore and map that data
-\citep{hurricaneexposure, hurricaneexposuredata}. Many previous studies have
-used geographical information system software (e.g., ArcGIS) to assess exposure
-to tropical cyclones in the U.S. \citep{grabich2016, zandbergen2009,
-czajkowski2011, kruk2010}. Here, we offer methods to map and output historic
-exposure to tropical cyclones that does not require the use of proprietary
-software but instead uses a package written in the R statistical programming
-language \citep{R}, which is free and open-source.
-
-In addition to providing tropical cyclone exposure metrics for individual
-hazards, this software and data also allow users to create storm exposure
-profiles based on multiple hazards or craft exposure indices that combine hazard
-metrics \citep{chakraborty2005population, peduzzi2009assessing}. This ability
-can be critical, as different hazards of extreme storms often act
-synergistically in their impacts \citep{smith2009}. Further, by including
-measurements of different hazard exposures in each county for each storm, this
-software allows for the development of more complex models that incoroporate
-multiple potential predictors of an impact (e.g., random forests, multivariable
-generalized linear models).
-
-In creating this hurricane exposure dataset, we aimed for data that are
-reliable, available for all eastern U.S. counties, and have no missing data
-within available dates. Further, we have investigated how well these data
-correspond with data from other available sources (Figs. S6--8). However, there
-are still some important limitations. The wind data are based on modeled, rather
-than observed, values, and while the modeled wind data generally agree well with
-post-analyis maximum wind radii from the revised Atlantic hurricane database
-(HURDAT2 \citep{landsea2013}) (Fig. S6), for a few storms it did not (e.g., Fig.
-S9). The rainfall data are from re-analysis data, which are generally
-well-correlated with observed ground-based station data but may oversmooth
-extreme measurements (Fig. S7). The flood and tornado data came from the
-National Oceanic and Atmospheric Administration's (NOAA's) Storm Events
-database, which, while a widely-used database of events maintained by NOAA, is
-based on reports, and so may be prone to underreporting \citep{Ashley2008flood,
-Curran2000}, especially in less populated areas \citep{Witt1998, Ashley2007}, as
-well as to other reporting errors. While these are important limitations, we
-selected these data sources as among the best currently available for measuring
-each of these hazards at a multi-county, multi-year scale.
-
-\subsection*{Conclusions}
-
-Previous research has highlighted the range of impacts that tropical cyclones
-can have in U.S. communities (e.g., \cite{pielke1997, rappaport2000, liu2005,
-han2009, mallin2006}). Determining the metrics of storm exposure that are most
-associated with loss of life, property damage, and other impacts may help
-identify and quantify the important threats that remain from tropical cyclones
-in the U.S., which could allow future success in reducing these threats through
-community planning, warning systems, and other measures. Here we found large
-differences in which counties are exposed to different hazards of tropical
-cyclones and that distance is, at best, a moderate, and often a very poor,
-surrogate for exposure to the specific storm hazards of high wind, extreme
-rainfall, flooding, and tornadoes. Use of distance as a surrogate for any of
-these hazards could lead to exposure misclassification and, in the case of storm
-impact studies, result in biased storm risk estimates. Similarly, if studies
-disagree on their findings of tropical cyclone impacts, it could result from the
-often poor agreement between exposure classifications based on these different
-metrics, making such studies that use different metrics hard to compare or
-combine in meta-analyses.
-
-\section*{Acknowledgments}
-
-This work was supported in part by grants from the National Institute of
-Environmental Health Sciences (R00ES022631), the National Science Foundation
-(1331399), the Department of Energy (Grant No. DE-FG02-08ER64644), and a
-National Aeronautics and Space Administration Applied Sciences Program/Public
-Health Program Grant (NNX09AV81G). Rainfall data are based on data acquired as
-part of the mission of the National Aeronautics and Space Administration's Earth
-Science Division and archived and distributed by the Goddard Earth Sciences
-(GES) Data and Information Services Center (DISC).
+\input{ehp_acknowledgements}
 

### 2019-01-26 writing/ehp_figures.tex
Add a bunch of stuff

--- 2019-01-26 writing/ehp_draft.tex
@@
-\documentclass[fleqn,10pt,lineno]{olplainarticle}
-% Use option lineno for line numbers
+\section*{Figures}
 
-\graphicspath{{../figures/}}
+\listoffigures
 
-\usepackage[implicit=false]{hyperref}
-\usepackage{acronym}
+\clearpage
 
-\title{Assessing United States county-level exposure for research on tropical cyclones and human health}
-
-\author[a,1]{G. Brooke Anderson} 
-\author[a,b]{Joshua Ferreri}
-\author[c]{Mohammad Al-Hamdan} 
-\author[c]{William Crosson} 
-\author[d]{Andrea Schumacher} 
-\author[e]{Seth Guikema} 
-\author[f]{Steven Quiring} 
-\author[g]{Dirk Eddelbuettel} 
-\author[a]{Meilin Yan} 
-\author[h]{Roger D. Peng}
-
-\affil[a]{Department of Environmental \& Radiological Health Sciences, Colorado 
-  State University, Fort Collins, CO, 80523} 
-\affil[b]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
-\affil[c]{Universities Space Research Association, NASA Marshall Space Flight 
-  Center, Huntsville, AL, 35805}
-\affil[d]{Cooperative Institute for Research in the Atmosphere, Colorado State
-  University, Fort Collins, CO, 80523} 
-\affil[e]{Department of Industrial and Operations Engineering, University of 
-  Michigan, Ann Arbor, MI, 48109}
-\affil[f]{Department of Geography, Ohio State University, Columbus, OH, 43210}
-\affil[g]{Debian and R Projects; Department of Statistics, University of
-  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
-\affil[h]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
-  Health, Baltimore, MD, 21205}
-
-\keywords{Hurricanes, Tropical cyclones, Disaster impacts, Exposure assessment}
-
-\begin{abstract} 
-  \input{ehp_abstract}
-\end{abstract}
-
-\begin{document}
-
-\flushbottom 
-\maketitle 
-\thispagestyle{empty}
-
-\input{ehp_introduction}
-
-\begin{table}%[tbhp] 
-\centering 
-\caption{Exposure metrics considered to assess county-level exposure to 
-tropical cyclones}
-\begin{tabular}{p{0.9cm}p{2.5cm}p{9cm}} 
-Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Tornado & 1996--2015 & Tornado event listed NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
-\bottomrule 
-\end{tabular} 
-\label{tab:exposuremetrics} 
-\end{table}
-
-\input{ehp_methods}
-
-\begin{figure*}%[tbhp] 
+\begin{figure}%[tbhp]
 \centering
@@
 exposure metric (Table 1). The red line shows the track of Hurricane Ivan 
-based on the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
-Similar maps for other large-extent storms are given in Fig. S5.}
+based on the revised Atlantic hurricane database (HURDAT2 \cite{landsea2013}).
+Similar maps for other large-extent storms are given in Fig. S3.}
 \label{fig:ivanexposure} 
-\end{figure*}
+\end{figure}
 
-\input{../tables/stormseverity.tex}
+\clearpage
 
@@
 \centering 
-\includegraphics[width = 0.7\linewidth]{jaccard_heatmap} 
+\includegraphics[width = 0.6\linewidth]{jaccard_heatmap} 
 \caption{Heatmap of Jaccard index values for
@@
 hierarchical clustering using the complete link method
-\citep{murtagh2012algorithms} (i.e., storms in the same cluster tend to have
+\cite{murtagh2012algorithms} (i.e., storms in the same cluster tend to have
 similar patterns for the pairwise strength of agreement among metrics); columns
@@
 Hurricane Ivan (2004) (Fig. \ref{fig:ivanexposure}) and Hurricanes Floyd (1999),
-Lee (2011), Cindy (2005), and Katrina (2005) (Fig. S5).} 
+Lee (2011), Cindy (2005), and Katrina (2005) (Fig. S3).} 
 \label{fig:jaccard}
@@
 
-\begin{figure*}%[tbhp] 
+\clearpage
+
+\begin{figure}%[tbhp] 
 \centering
-\includegraphics[width=16cm]{averageexposureonly} 
+\includegraphics[width=15cm]{averageexposureonly} 
 \caption{Average number of storm exposures per decade in U.S. counties for 
@@
 analysis is restricted to years with all exposure data available (1996--2011;
-Fig. S3).} 
+Fig. S5).} 
 \label{fig:averageexposure} 
-\end{figure*}
-
-\input{ehp_results}
-
-\input{ehp_discussion}
-
-\input{ehp_acknowledgements}
-
-\bibliography{hurr_exposure}
-
-\end{document}
+\end{figure}

### 2019-01-26 writing/ehp_introduction.tex
Add a bunch of stuff

\section*{Introduction}

Tropical cyclones, including hurricanes, can have disastrous impacts in the
\ac{US}. A tropical cyclone's high winds can bring health risks
and property damage through structural damage of houses and other buildings,
falling trees, and wind-borne debris~\citep{rappaport2000} and cause power
outages~\citep{liu2005, han2009}, which introduce a number of threats to human
and ecological health, including water quality risks if the outage affects
wastewater treatment plants~\citep{mallin2006}.  Other risks can exist without
severe wind; for example, one study found that most of the direct
hurricane-related deaths in the \ac{US} between 1970 and 1999 occurred in cases
when wind was below hurricane strength, including for Tropical Storms Charley
in 1998 and Alberto in 1994~\citep{rappaport2000}.  Tropical cyclones can
produce excessive rain, especially in certain topographies (e.g., near
mountains), so counties well inland sometimes experience more extreme rain than
coastal counties. Flood risks from tropical cyclones can result from this rain,
although the two risks are not perfectly correlated~\citep{chen2015}. In the
\ac{US}, over half of hurricane-related direct deaths from 1970 to 1999 from
Atlantic basin storms were a result of freshwater flooding~\citep{rappaport2000}. 
Flooding can also degrade water quality~\citep{mallin2006}, which can threaten both 
human and ecological health.

To measure tropical cyclone exposure for either investigations of exposure
patterns or for estimation of tropical cyclone risks and impacts, including
those related to human health, studies have varied widely in how they determine
tropical cyclone exposure (some examples shown in Figure~S1). While some
studies have measured exposure based on specific hazards of the tropical
cyclone (e.g., wind, rain), many have used distance from the tropical
cyclone's central track as a surrogate in identifying exposure to the hazards
of a tropical cyclone (e.g.,~\citet{czajkowski2011, tansel2010, kinney2008,
caillouet2008increase}).

Distance from the tropical cyclone's central track can be easily measured using
widely-available hurricane tracking data. There are, however, a number of
limitations to using distance to assess exposure to tropical cyclones, and
there are reasons to suspect that exposure classifications based on other
tropical cyclone hazards may disagree with distance-based classifications.
Tropical cyclones vary dramatically in size: \ac{US} tropical cyclones have been
observed with radii to maximum winds as small as 20 kilometers and as large as
200 kilometers~\citep{mallin2006, quiring2011variations}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and 
waves~\citep{rappaport2000, kruk2010}), other hazards like dangerous rain, 
floods, and tornadoes can occur well away from the tropical cyclone's central 
track~\citep{rappaport2000, atallah2007, moore2012}.  For example, fatal tropical
cyclone tornadoes, which were linked to over 300 deaths in the \ac{US}  between
1995 and 2009, most often occur 200\,--\,500 kilometers from the tropical cyclone's
center~\citep{moore2012}.

Further, distance-based exposure metrics that use
buffers tend to use an equal buffer distance on each side of the tropical
cyclone track (e.g.,~\citep{czajkowski2011, grabich2015, grabich2016,
zandbergen2009, tansel2010}), but the forces of a tropical cyclone tend to be
distributed around the center in a non-symmetrical way. For example, extreme
winds are more common to the right of the track, where counter-clockwise
cyclonic winds move in concert with the tropical cyclone's forward 
motion~\citep{halverson2015}, and the fatal tornadoes associated with \ac{US} tropical
cyclones between 1995 and 2009 occurred almost exclusively to the right of the
tropical cyclone's track, mostly in the right front quadrant of the tropical
cyclone~\citep{moore2012}. Rain, conversely, is often heaviest to the left of
the tropical cyclone's track, especially when the tropical cyclone interacts
with other weather systems~\citep{atallah2003, atallah2007, zhu2013variations}
or undergoes an extratropical transition~\citep{elsberry2002}.

[Something about differences in other metrics]

The use of different exposure metrics across studies hampers the development of
a scientific consensus on tropical cyclone-related risks, as differences
observed between studies could result from differences in exposure assessment.
Further, if a study misclassifies exposure to the hazard or hazards that cause
an impact, estimates of tropical cyclone risks or impacts will be biased,
making it hard to identify true associations. These concerns are particularly
important if there are large differences in the locations that are determined
to be exposed based on different metrics.

To investigate this, we measured county-level exposure to tropical cyclones in
all eastern \ac{US} counties using five metrics (
Table~\ref{tab:exposuremetrics}): (1) distance to tropical cyclone track; (2) maximum
sustained wind speed; (3) cumulative rainfall; (4) flood events; and (5)
tornado events.  We assessed exposure at the county level since data on many
potential impacts are available at county-level aggregations (e.g., direct
hurricane-related deaths~\citep{czajkowski2011}; birth outcomes~\citep{grabich2015,
grabich2016}; autism prevalence~\citep{kinney2008}) and
since decisions and policies to prepare for, and respond to, tropical cyclones
are often undertaken at the county level~\citep{zandbergen2009, rappaport2000}.
We explored patterns in these exposure assessments and measured how well
exposure classification agreed across these five metrics in terms of
classifying specific counties as exposed to a tropical cyclone.  Finally, to
make this hazard-specific tropical cyclone exposure data available to other
researchers, we published open source software for county-based hurricane
exposure assessment in \ac{US} counties~\citep{hurricaneexposure}.

### 2019-01-26 writing/ehp_methods.tex
Add a bunch of stuff

\section*{Methods and Materials}

All counties in the eastern half of the \ac{US} (e.g., Figure~\ref{fig:ivanexposure}), 
as of the 2010 \ac{US} Decennial Census, were included in
the study. The study covered all tracked storms in the \ac{HURDAT2} \citep{landsea2013} 
that came within 250
kilometers of at least one study county between 1988 and 2015 (Figure~S2).
HURDAT2 is a post-storm assessment conducted by the \ac{US} National Hurricane
Center (NHC) and incorporates data from a variety of sources, including
satellite data and, when available, aircraft reconnaissance 
data~\citep{landsea2013, jarvinen1988}. These data typically give measurements of
the tropical cyclone center's location at 6-hour intervals at synoptic times
(i.e., 6:00 am, 12:00 pm, 6:00 pm, and 12:00 am UTC); some landfalling tropical
cyclones have an additional observation at the time of landfall~\citep{landsea2013}.

\subsection*{Distance-based exposure metric}

For each storm, the distance of closest approach was measured for every study county. 
Storm center track location data from \ac{HURDAT2} was 
interpolated using natural cubic splines to 15-minute intervals~\cite{hurricaneexposure}. 
The distance between each county's population mean center, as assessed by the 2010 Decennial
Census~\cite{countycenters}, and each 15-minute interpolated storm center location was
measured using Great Circle (WGS84 ellipsoid) distance~\cite{bivand2013applied}. The 
distance-based exposure metric was based on the minimum value of this storm 
center-to-county distance throughout the period the storm was tracked. The time of 
closest approach for each county-storm pairing was also recorded for use assessing 
rain-, flood-, and tornado-based metrics. To allow matching with data based on local 
time, times were converted from Coordinated Universal Time (UTC) to local 
time~\cite{countytimezones}.

\subsection*{Rain-based exposure metric}

For the rain-based exposure metric, cumulative rainfall was estimated for the
period from two days before to one day after the date of a tropical cyclone's
closest approach to a county. Daily county-level rainfall was determined by
aggregating hourly, 1/8 degree gridded reanalysis data from the North American
Land Data Assimilation System Phase 2 (NLDAS-2) precipitation data 
files~\citep{rui2013nldas}. The NLDAS-2 data integrate satellite-based and land-based
monitoring, applying a land-surface model to create a reanalysis dataset that
is spatially and temporally complete across the continental 
\ac{US}~\citep{rui2013nldas, alhamdan2014environmental}. To aggregate to county-level,
hourly data at each grid point were summed to create a daily rainfall total,
and these grid point rainfall totals were then averaged across all grid points
within a county's 1990 \ac{US} Census boundaries to create a county-level daily
total~\citep{alhamdan2014environmental, cdcwonder}. This county-level
precipitation data are publicly available through the \ac{US} Centers for Disease
Control's Wide-ranging Online Data for Epidemiological Research (WONDER)
database~\citep{cdcwonder, alhamdan2014environmental}. The rainfall-based
exposure metric was only available for tropical cyclones through 2011, the
period for which the county-aggregated reanalysis rainfall data were 
available~\citep{cdcwonder, alhamdan2014environmental}. Validation analysis of this and
other hazard-based exposure data is given in the Supplemental Material.

\subsection*{Wind-based exposure metric}

Ground-based observations of wind speed are problematic during tropical
cyclones, as instruments often fail at high wind speed, while reanalysis data,
often available at hourly or higher resolution, can be too smooth to capture
wind extremes associated with a tropical cyclone. Therefore, to create a
dataset of county-level sustained winds during historical tropical cyclones, we
modeled maximum sustained wind speeds at each county's population mean 
center~\citep{countycenters}. HURDAT2 Best Track data were interpolated from 6-hourly
reported values to 15-minute increments using natural cubic 
splines~\citep{stormwindmodel}, and then sustained wind speed was modeled for each
county center at each 15-minute increment using a double exponential wind speed
model~\citep{willoughby2006parametric} to estimate maximum ground-level
sustained wind speed at each county center~\citep{stormwindmodel}. Asymmetry in
wind speeds around the tropical cyclone center was incorporated into the wind
speed model~\citep{phadke2003modeling, stormwindmodel}. The maximum value of
this modeled wind speed was determined for each county as the maximum windspeed
across the 15-minute incremented modeled values throughout the tropical
cyclone.

\subsection*{Flood- and tornado-based exposure metrics}

To identify flood- and tornado-based exposures to tropical cyclones in \ac{US}
counties, we used event listings from the National Oceanic and Atmospheric
Administration (NOAA)'s Storm Event Database~\citep{stormevents}. For each
tropical cyclone, we identified all events with event types related to flooding
(``Flood", ``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and
that occurred in a county within 500 kilometers of the tropical cyclone's track
and within a five-day window centered on the date of the tropical cyclone's
closest approach to the county~\citep{hurricaneexposuredata}. ``Flood", ``Flash
Flood" and ``Tornado" events in this database were reported by county Federal
Information Processing Standard (FIPS) code. ``Coastal Flood" events were
reported by forecast zone; for these, the event was matched to the appropriate
county if possible~\citep{noaastormevents}. While this database has recorded
storm data since 1950, its coverage changed substantially in 1996 to cover a
wider variety of storm events~\citep{stormevents}. We therefore only considered
these metrics of hurricane exposure for tropical cyclones in 1996 and later.

\subsection*{Assessment of tropical cyclone exposure data}

To assess the rainfall estimates from the reanalysis data used in analysis in the the 
main text, we compared cumulative rainfall estimates based on this reanalysis rainfall 
data to rainfall observations from ground-based weather stations in nine sample study 
counties, selected as counties with data available from multiple stations. In these nine
counties, observed rainfall was collected for all storms, 1988\,--\,2011, for the same period 
of two days before to one day after each storm's closest approach, using all stations with
available daily data in the Global Historical Climatology Network throughout 
1988\,--\,2011~\citep{menne2012overview, rnoaa, countyweather}. Agreement between 
measured storm-related rainfall between these two data sources was assessed by 
Spearman's rank correlation.

To validate the metric of county-level wind speed used in the main text, we compared this 
wind metric to estimated wind radii from the revised Atlantic hurricane database (HURDAT2),
available for storms in our study from 2004~\citep{landsea2013}. HURDAT2 gives estimates of
the maximum radii to winds of 64, 50, and 34 knots in each four quadrants of a storm (i.e.,
distance from the storm's center to the furthest point in a quadrant with winds at or above
that speed) at each 6-hour storm observation.

As validation for the flood data, we compared streamflow data from United States (\ac{US})
Geological Survey gages~\citep{usgsgages, countyfloods, dataRetrieval} in nine counties with
tropical cyclone flood classifications based on the NOAA Storm Events Database 
(Figure~\ref{fig:floodcomparison}).

\subsection*{Assessing agreement in exposure classification between
metrics}

Two of the exposure metrics (flood- and tornado-based) were inherently binary,
since these metrics were based on whether an event was listed in the NOAA Storm
Events database or not.  For the other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
assessed county-level tropical cyclone exposure in the eastern \ac{US} based on
each exposure metric for all land-falling or near-land Atlantic basin tropical
cyclones. Depending on available exposure data, this assessment included some
or all of the period from 1988 to 2015, a period with 136 tropical cyclones
that made \ac{US} landfall or passed within 250 kilometers (155 miles) of at least
one \ac{US} county (Figure~S2).

We measured agreement between exposure metrics in the classification of a
county as exposed or unexposed to a specific tropical cyclone. For this
measurement, we calculated the within-storm Jaccard 
index~\citep{jaccard1901distribution, jaccard1908nouvelles} between every pair-wise
combination of hazard metrics.  The Jaccard index ($J_s$) measures similarity
between two metrics ($X_{1,s}$ and $X_{2,s}$) for tropical cyclone $s$ as the
proportion of counties in which both of the metrics classify the county as
exposed out of all counties classified as exposed by at least one of the
metrics:

\begin{equation} 
J_s = \frac{X_{1,s} \cap X_{2,s}}{X_{1,s} \cup X_{2,s}}
\end{equation}

\noindent This metric can range from 0, in the case of no overlap between the counties
classified as exposed based on the two metrics, to 1, in the case that the
two metrics classify exactly the same counties as exposed to the tropical cyclone.

### 2019-01-26 writing/ehp_results.tex
Add a bunch of stuff

\section*{Results}

\subsection*{Assessment of tropical cyclone exposure data}

Within
the nine sample counties, storm-related rainfall from the reanalysis precipitation dataset 
was well-correlated with station-based rainfall measurements, although the reanalysis data 
did tend to somewhat underestimate rainfall totals in storms with extremely high rainfall 
(Figure~\ref{fig:raincomparison}). A distance constraint (county population mean center within
500 kilometers of the storm's closest approach) was necessary for assessing storm exposure
based on this metric to ensure that rains unrelated and far from the storm track were not
misattributed to storm exposures. We found that this distance constraint was typically
adequately large to capture storm-related rain, although in rare examples of exceptionally
large storms (e.g., Hurricane Ike [2008]) or storms for which tracking was stopped at
extratropical transition (e.g., Tropical Storm Lee [2011]), some storm-related rains were
missed because of this distance constraint (Figure~\ref{fig:rainexamples}).

The modeled wind speeds generally agreed well
with the HURDAT2 wind radii (Figure~\ref{fig:windcomparison}), with disagreement mostly 
limited to large storms that were characterized by high winds well inland (e.g., Hurricanes
Ike [2008] and Sandy [2012]; Figure~\ref{fig:windexamples}).

Stream gage data generally indicated higher discharge in these
counties on dates identified as flood events, particularly in counties with more reporting
stream gages (e.g., Harris County, TX, with 10 reporting stream gages; 
Figure~\ref{fig:floodcomparison}), although these two flooding data sources were 
not completely
consistent.

\subsection*{Patterns in tropical cyclone exposures in eastern \ac{US} counties}

We first calculated summary statistics for the exposure assessments (Table
\ref{tab:exposuresummaries}). There was wide variation in the average number of
county exposures per year under the five different tropical cyclone metrics
considered. For the tornado metric, there were approximately 50 county
exposures identified per year, on average, within our study.  For flood and
wind metrics, there were substantially more county exposures identified (over
200 per year on average), and even more when assessing exposure using the
distance and rain metrics (over 400 per year on average).

We also measured the average number of county exposures per tropical cyclone,
within those tropical cyclones in which at least one county was exposed to the
hazard (Table~\ref{tab:exposuresummaries}). The typical number of county
exposures per tropical cyclone was lowest when assessing exposure using the
tornado metric (median: 7 county exposures per tropical cyclone), higher under
the flood and wind metrics (median: 26 and 34, respectively), and even higher for the metrics of rain (median: 68)
and distance (median: 84).  For every metric except the tornado-based metric,
we identified at least one tropical cyclone with over 300 counties exposed.
However, the largest-extent tropical cyclone varied by metric (e.g., Beryl in
1994 exposed the most counties based on distance, Frances in 2004 based on
rain, and Ike in 2008 based on wind).

\subsection*{Agreement in county-specific exposure classification to specific
tropical cyclones}

The counties that were identified as exposed to a tropical cyclone differed
substantially depending on which metric was used in exposure assessment.  This
was true both when comparing exposure assessment based on distance versus the
four hazard-specific metrics considered (wind, rain, floods, and tornadoes)
and also when comparing exposure assessments among the four hazard-specific
metrics.

Figure~\ref{fig:ivanexposure} gives an example of how sensitive county-level
tropical cyclone exposure assessment is to the choice of metric, using as an
example Hurricane Ivan (2004), a record-breaking tropical cyclone for its
duration as a major hurricane~\citep{franklin2006atlantic}. When county-level
exposure was determined based on distance, the counties assessed as exposed
followed the tropical cyclone's track, including counties well inland. For the
wind metric, only counties near two of the tropical cyclone's landfalls were
assessed as exposed. For rain- and flood-based metrics, however, exposure
extended to the left of the track, including counties as far north as New York
and Connecticut, while for the tornado metric, exposed counties tended to be to
the right of the track and included several counties in central North Carolina,
South Carolina, and Georgia that were not identified as exposed to Ivan based
on any other metric.  Figure~S3 provides similar maps for four other example
tropical cyclones (selected because they exposed many \ac{US}  counties based on
at least one metric), and Figure~S4 provides four examples of tropical cyclones
in which the disagreement in exposure assessment between rain- and wind-based
metrics was particularly notable.

We drew similar conclusions when we investigated all 46 tropical cyclones
between 1996 and 2011 (the period for which all five metrics were available)
for which 250 or more counties were exposed based on at least one metric
(Figure~\ref{fig:jaccard}). In this figure, each row provides results for a
specific tropical cyclone, and each box in that row shows a measure of
agreement between a specific pair of metrics in terms of the counties assessed
as exposed to that tropical cyclone, with darker colors representing greater
agreement in county-level exposure assessments.

For most tropical cyclone, distance-based county-level exposure assessment was
at best in moderate agreement with exposure assessments based on the four
hazard-based metrics. For almost all tropical cyclones, Jaccard index values
between distance and any of the four hazard-based metrics were 0.6 or lower
(i.e., out of the counties assessed as exposed to the tropical cyclone by at
least one of the two metrics considered, 60\% or fewer of the counties had the
same exposure assessment under the two metrics). For many tropical cyclone, the
Jaccard index values were 0.2 or lower, suggesting very poor agreement in
exposure assessment between the metrics.  Among the four hazard-based metrics,
the tornado-based metric showed universally poor agreement with other metrics
in county-level classification across all tropical cyclones considered.  For
other pairs of metrics, there were also generally large differences in which
counties were determined to be exposed to the tropical cyclone, with the
Jaccard index values below 0.4 for most metric pairs in most tropical cyclones.

There were, however, a few exceptions\,---\,tropical cyclones in which similar
counties were determined to be exposed to the tropical cyclone for two or more
of the metrics considered.  For Floyd in 1999 (Figure~S3) and Irene in 2011,
for example, county-level classification agreed moderately to well (Jaccard
index of approximately 0.5\,--\,0.8) for all pairs of exposure metrics except those
including the tornado-based metric. These tropical cyclones both made their
first \ac{US}  landfall in North Carolina at minor hurricane windspeeds (Category
2 and 1, respectively) and then skimmed the eastern coast of the \ac{US} north
through New England, bringing large rainfall to much of the eastern coast from
North Carolina north and causing extensive inland flooding in North Carolina
(Floyd) and New England (Irene)~\citep{avila2013atlantic,
lawrence2000atlantic}. For another set of tropical cyclones (e.g., Lee in 2011
[Figure~S3], Ernesto in 2006, and Bertha in 1996), there was moderate to good
agreement for all pairwise combinations of distance, rain, and wind, but poor
agreement for all other combinations of metrics. Most of these tropical
cyclones either skimmed along the eastern coast north of North Carolina (Bertha
in 1996, Charley in 2004, Gaston in 2004, Barry in 2007, and Hanna in 2008), in
a pattern similar to Hurricanes Floyd in 1999 and Irene in 2011, or made
landfall in or near the Florida panhandle and cut across to exit into the
Atlantic around North Carolina (Josephine in 1996, Earl in 1998, Alberto in
2006). However, these two sets of tropical cyclones represented the minority of
all tropical cyclones considered; for most tropical cyclones, there was low
overlap in the counties determined to be exposed to the tropical cyclones
between most pairings of metrics.

\subsection*{Geographic patterns in tropical cyclone exposures}

Strong geographical patterns were clear across tropical cyclone exposure
assessments based on these different metrics when the average number of
exposures per decade in each county were calculated 
(Figure~\ref{fig:averageexposure}). Wind-based exposure had a strong coastal pattern,
with almost all exposures in counties within about 200 kilometers (124 miles)
of the coastline, mostly in eastern North and South Carolina, southern Florida,
and the area near the outlet of the Mississippi River. While exposures by rain-
and distance-based metrics were also more common in coastal areas compared to
inland areas, inland exposures were much more common based on these two metrics
compared to the wind-based metric. Rain and, to some extent, flood exposures
were characterized by a pattern defined by the Appalachian Mountains, with much
lower exposures west of the mountain range than to the east. Almost all
tornado-based exposures were in coastal states, with most in Florida, Alabama,
South Carolina, and North Carolina and almost none north of Maryland.
Flood-based exposures were highest in North Carolina and along the South
Carolina coast, as well as further north in New Jersey, southeastern
Pennsylvania, and southeastern New York. Notably, some areas of the \ac{US} that
experienced regular exposures to tropical cyclone-related excessive rainfall
and floods were not identified as areas of regular risk based on the more
commonly used tropical cyclone exposure assessment metrics of wind or distance
(Figure~\ref{fig:averageexposure}).

We included a case study in the Supporting Information showing how these
differences in exposure assessment across the five metrics considered could
strongly influence health-related studies based on physical exposure to
tropical cyclones. For this example, we compared estimates of which \ac{US}
counties have the largest expected physical exposure to tropical cyclones among
a susceptible subpopulation, Medicare beneficiaries who are reliant on
electricity for medical equipment (Supplemental Text, Figure~S6)~\citep{empower}.

\subsection*{Software}

To assist with future tropical cyclone impact studies, we created and published
open source software that includes the hazard-specific, county-level tropical
cyclone data in this paper, as well as tools to explore and map that 
data~\citep{hurricaneexposure, hurricaneexposuredata}. This software includes continuous
data for the distance-, rain-, and wind-based metrics, allowing users to also explore
different thresholds of these metrics when assessing exposure.

We investigated how well these data correspond with data from other available
sources (Figures~S7\,--\,11).  While generally in agreement with data from other
sources, there are a few limitations. The wind data are based on modeled,
rather than observed, values, and while the modeled wind data generally agree
well with post-analysis maximum wind radii from the revised Atlantic hurricane
database (\ac{HURDAT2}~\citep{landsea2013}) (Figure~S7), for a few tropical cyclones
it did not (e.g., Figure~S10). The rainfall data are from re-analysis data,
which are generally well-correlated with observed ground-based station data but
sometimes oversmooth extreme measurements (Figure~S8). More details on the assessment
of this data are provided in the Supplemental Material.

### 2019-01-26 writing/ehp_supp.tex
Add a bunch of stuff

--- 2019-01-26 writing/anderson_ehp_supp.tex
@@
-\documentclass[11pt, titlepage]{article}
-\usepackage{microtype}
-\usepackage{charter}
+\documentclass[fleqn,10pt,lineno]{olplainarticle}
+% Remove the twocolumn option to create a single column SI file if required.
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
 
-\input{ehp_preamble}
+\title{Supporting Information for ``Assessing United States county-level exposure for research on tropical cyclones and human health"}
+\author{Anderson et al.}
 
@@
 
-\title{Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"}
+\flushbottom 
+\maketitle 
+\thispagestyle{empty}
 
-\author[1,*]{G. Brooke Anderson} 
-\author[1,2]{Joshua Ferreri}
-\author[3]{Mohammad Al-Hamdan} 
-\author[3]{William Crosson} 
-\author[4]{Andrea Schumacher} 
-\author[5]{Seth Guikema} 
-\author[6]{Steven Quiring} 
-\author[7]{Dirk Eddelbuettel} 
-\author[1]{Meilin Yan} 
-\author[8]{Roger D. Peng}
+\section*{Supporting Information (SI)}
 
-\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
-  State University, Fort Collins, CO, 80523} 
-\affil[2]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
-\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
-  Center, Huntsville, AL, 35805}
-\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
-  University, Fort Collins, CO, 80523} 
-\affil[5]{Department of Industrial and Operations Engineering, University of 
-  Michigan, Ann Arbor, MI, 48109}
-\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
-\affil[7]{Debian and R Projects; Department of Statistics, University of
-  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
-\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
-  Health, Baltimore, MD, 21205}
-\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}
+\subsection*{Assessment of tropical cyclone exposure data}
 
-\date{}
+To assess the rainfall estimates from the reanalysis data used in analysis in the the main text, we compared cumulative rainfall estimates based on this reanalysis rainfall data to rainfall observations from ground-based weather stations in nine sample study counties, selected as counties with data available from multiple stations.
+In these nine counties, observed rainfall was collected for all storms, 1988--2011, for the same period of two days before to one day after each storm's closest approach, using all stations with available daily data in the Global Historical Climatology Network throughout 1988--2011 \cite{menne2012overview, rnoaa, countyweather}.
+Agreement between measured storm-related rainfall between these two data sources was assessed by Spearman's rank correlation.
+Within the nine sample counties, storm-related rainfall from the reanalysis precipitation dataset was well-correlated with station-based rainfall measurements, although the reanalysis data did tend to somewhat underestimate rainfall totals in storms with extremely high rainfall (Fig.
+\ref{fig:raincomparison}).
+A distance constraint (county population mean center within 500 kilometers of the storm's closest approach) was necessary for assessing storm exposure based on this metric to ensure that rains unrelated and far from the storm track were not misattributed to storm exposures.
+We found that this distance constraint was typically adequately large to capture storm-related rain, although in rare examples of exceptionally large storms (e.g., Hurricane Ike [2008]) or storms for which tracking was stopped at extratropical transition (e.g., Tropical Storm Lee [2011]), some storm-related rains were missed because of this distance constraint (Fig.
+\ref{fig:rainexamples}).
 
-\maketitle
+To validate the metric of county-level wind speed used in the main text, we compared this wind metric to estimated wind radii from the revised Atlantic hurricane database (HURDAT2), available for storms in our study from 2004 \cite{landsea2013}.
+HURDAT2 gives estimates of the maximum radii to winds of 64, 50, and 34 knots in each four quadrants of a storm (i.e., distance from the storm's center to the furtherest point in a quadrant with winds at or above that speed) at each 6-hour storm observation.
+The modeled wind speeds generally agreed well with the HURDAT2 wind radii (Fig.
+\ref{fig:windcomparison}), with disagreement mostly limited to large storms that were characterized by high winds well inland (e.g., Hurricanes Ike [2008] and Sandy [2012]; Fig.
+\ref{fig:windexamples}).
 
-\beginsupplement
-
-\section*{Supplemental Text}
+As validation for the flood data, we compared streamflow data from United States (U.S.) Geological Survey gages \cite{usgsgages, countyfloods, dataRetrieval} in nine counties with tropical cyclone flood classifications based on the NOAA Storm Events Database (Fig.
+\ref{fig:floodcomparison}).
+Stream gage data generally indicated higher discharge in these counties on dates identified as flood events, particularly in counties with more reporting stream gages (e.g., Harris County, TX, with 10 reporting stream gages; Fig.
+\ref{fig:floodcomparison}), although these two flooding data sources were not completely consistent.
 
@@
 
-The outcomes of disasters depend both on the geophysical forces of the disaster as well as on the vulnerability of the society living in the geographical areas affected by those forces.
-\citep{chakraborty2005population, anderson2003community, cutter1996vulnerability} As a case study of how differences in tropical cyclone exposure assessments across different hazard metrics influence estimates of physical exposures of susceptible populations to tropical cyclones, we investigated how the use of different metrics influenced estimates of which eastern U.S.
+The outcomes of disasters depend both on the geophysical forces of the disaster as well as on the vulnerability of the society living in the geographical areas affected by those forces.\cite{chakraborty2005population, anderson2003community, cutter1996vulnerability} As a case study of how differences in tropical cyclone exposure assessments across different hazard metrics influence estimates of physical exposures of susceptible populations to tropical cyclones, we investigated how the use of different metrics influenced estimates of which eastern U.S.
 counties have the highest expected average exposures of electricity-dependent Medicare beneficiaries to tropical cyclones.
@@
 We collected data on the number of electricity-dependent Medicare beneficiaries in each study county from the U.S.
-Department of Health \& Human Service's emPower Map 2.0 \citep{empower}.
-We then calculated the physical exposure of the electricity-dependent Medicare population in each county, based on tropical cyclone assessments using each hazard metric, following \citep{peduzzi2009assessing}:
+Department of Health \& Human Service's emPower Map 2.0 \cite{empower}.
+We then calculated the physical exposure of the electricity-dependent Medicare population in each county, based on tropical cyclone assessments using each hazard metric, following \cite{peduzzi2009assessing}:
 
@@
 \noindent where $E_c$ is the average yearly physical exposure among electricity-dependent Medicare beneficiaries in county $c$ to tropical cyclone exposures based on a given metric, $F_c$ is the estimated yearly expected frequency of tropical cyclone exposures in county $c$ based on that metric, and $P_c$ is the size of the electricity-dependent Medicare population in county $c$, as of July 2017.
-When combined with estimates of vulnerability of a population to a natural hazard, such measurements of physical exposure can be used to calculate risk of human losses from the hazard \citep{peduzzi2009assessing}.
+When combined with estimates of vulnerability of a population to a natural hazard, such measurements of physical exposure can be used to calculate risk of human losses from the hazard \cite{peduzzi2009assessing}.
 
@@
 \ref{fig:topelecdependexposure}), based on which storm hazard is considered.
-The measure of a hurricane's intensity, the Saffir-Simpson scale, expresses intensity as a function of wind speed, and so fails to capture potential risks from other hazards of the storm \citep{smith2009}, and previous assessments of physical exposure to tropical cyclones have often focused on winds or distance from the storm.
-For example, a global study of exposure and vulnerability to natural hazards used a combination of windspeed above a certain threshold and distance from the storm's track to assess exposure \citep{peduzzi2009assessing}.
+The measure of a hurricane's intensity, the Saffir-Simpson scale, expresses intensity as a function of wind speed, and so fails to capture potential risks from other hazards of the storm \cite{smith2009}, and previous assessments of physical exposure to tropical cyclones have often focused on winds or distance from the storm.
+For example, a global study of exposure and vulnerability to natural hazards used a combination of windspeed above a certain threshold and distance from the storm's track to assess exposure \cite{peduzzi2009assessing}.
 While many storm impacts might be most strongly linked to wind hazards, there is a growing recognition of the potential risks of adverse health outcomes and property damage from rain- and flood-related tropical cyclone hazards in the U.S.
-\citep{smith2009}.
+\cite{smith2009}.
 Such assessments may underemphasize potentially dangerous exposures to certain regions of the U.S., particularly inland locations.
-In fact, based on our assessment of physical exposures to tropical cyclones among electricity-dependent Medicare beneficiaries, the highest expected physical exposure under any metric was measured for Philadelphia County, PA, under the flood metric (Fig.
+In fact, based on our assessment of physical exposures to tropical cyclones among electicity-dependent Medicare beneficiaries, the highest expected physical exposure under any metric was measured for Philadelphia County, PA, under the flood metric (Fig.
 \ref{fig:topelecdependexposure}).
@@
 \centering
-\includegraphics[width=\linewidth]{previous_exposure_metrics}
+\includegraphics[width=\linewidth]{figures/previous_exposure_metrics}
 \caption{Metrics used in a sample of previous studies to assign exposure to tropical cyclones in U.S.-based tropical cyclone exposure or impact studies.
@@
 Studies for which there are two or more bars (e.g., bars for both distance and wind) indicate studies that either used separate exposure assessments, one for each metric, (e.g., as a sensitivity analysis) or used a single exposure index that incorporated multiple metrics in its definition.
-``Damage" often indicate use of Federal Emergency Management Agency (FEMA) reports, although occasionally used damage reports from other sources, including newspaper reports.
+``Damage" often indicate use of Federal Emergency Management Agency (FEMA) reports, although occassionally used damage reports from other sources, including newspaper reports.
 Examples of ``Other" include evacuation orders, excessive school closures, length of time a storm was near a location, and detailed, holistic analyses of a specific storm or storms.
-References for each study are included in the
-bibliography of this Supplemental Material.}
+References for each study are: Logan and Xu 2015 \cite{logan2015}; Gares and Montz 2014 \cite{gares2014}; Esnard et al.
+2011 \cite{esnard2011}; Zandbergen 2009 \cite{zandbergen2009}; Horney et al.
+2016 \cite{horney2016}; Grabich et al.
+2016a \cite{grabich2016}; Antipova and Curtis 2015 \cite{antipova2015post}; Zahran et al.
+2010 \cite{zahran2010}; Currie and Rossin-Slater 2013 \cite{currie2013}; Grabich et al.
+2015 \cite{grabich2015}; Grabich et al.
+2016b \cite{grabich2016hurricane}; Amstadter et al.
+2010 \cite{amstadter2010}; Acierno et al.
+2006 \cite{acierno2006}; Le et al.
+2013 \cite{le2013}; Lieberman-Cribben et al.
+2017 \cite{lieberman2017}; McKinney et al.
+2011 \cite{mckinney2011}; Dosa et al.
+2012 \cite{dosa2012evacuate}; Swerdel et al.
+2014 \cite{swerdel2014}; Czajkowski et al.
+2011 \cite{czajkowski2011}; Mongin et al.
+2017 \cite{mongin2017}; Kessler et al.
+2007 \cite{kessler2007hurricane}; Caillou{\"e}t et al.
+2008 \cite{caillouet2008increase}; Rosenheim et al.
+2018 \cite{rosenheim2018disaster}; Domino et al.
+2003 \cite{domino2003disasters}; Kim et al.
+2016 \cite{kim2016}; Kinney et al.
+2008 \cite{kinney2008}; Kim and Marcouiller 2015 \cite{kim2015}; Tansel and Sizirici 2010 \cite{tansel2010}; Escobedo et al.
+2009 \cite{escobedo2009}; Fuller 2014 \cite{fuller2014}; Baggerly and Ferretti 2008 \cite{baggerly2008impact}; Strobl 2011 \cite{strobl2011economic}; Mukherjee et al.
+2017 \cite{mukherjee2017}.}
 \label{fig:previousmetrics}
 \end{figure}
-
-\nocite{logan2015, gares2014, esnard2011, zandbergen2009, horney2016, grabich2016, 
-antipova2015post, zahran2010, currie2013, grabich2015, grabich2016hurricane, amstadter2010,
-acierno2006, le2013, lieberman2017, mckinney2011, dosa2012evacuate, swerdel2014, 
-czajkowski2011, mongin2017, kessler2007hurricane, caillouet2008increase, 
-rosenheim2018disaster, domino2003disasters, kim2016, kinney2008, kim2015, tansel2010,
-escobedo2009, fuller2014, baggerly2008impact, strobl2011economic, mukherjee2017}
 
@@
 \centering
-\includegraphics[width=0.7\linewidth]{hurrtracks}
+\includegraphics[width=0.7\linewidth]{figures/hurrtracks}
 \caption{Tracks of all tropical cyclones included in the study.
-The study included all tracked storms in the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}) that came within 250 kilometers of at least on U.S.
+The study included all tracked storms in the revised Atlantic hurricane database (HURDAT2 \cite{landsea2013}) that came within 250 kilometers of at least on U.S.
 county, 1988--2015.
-Thicker lines show the tracks of storms whose names were retired, indicating that the storm was particularly severe or had notable impacts \citep{retirednames}.}
+Thicker lines show the tracks of storms whose names were retired, indicating that the storm was particularly severe or had notable impacts \cite{retirednames}.}
 \label{fig:hurrtracks}
@@
 \centering
-\includegraphics[width=\linewidth]{othertopstorms}
-\caption{Counties exposed, based on different exposure metrics, to the storms with largest extent (as measured by the number of counties exposed based on any metric) from each of the clusters shown in the Jaccard heatmap in the main text (Figure 2).}
-\label{fig:othertopstorms}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{extentdisagreement}
-\caption{Examples of storms for which exposure assessments differed substantially between rain-based and wind-based metrics.
-Counties shown in orange were classified as exposed to a given storm by a given metric; those in yellow were classified as not exposed.
-The storm's track is shown in red.
-For both Frances (2004) and Isidore (2002) (top panel), substantially more counties were assessed as exposed when measuring exposure based on rain rather than wind.
-Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel), substantially more counties were assessed as exposed based on wind compared to rain.}
-\label{fig:extentdisagreement}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{averageexposureonlysupp}
+\includegraphics[width=\linewidth]{figures/averageexposureonlysupp}
 \caption{Average number of storm exposure per decade in U.S.
@@
 
-\begin{figure*}%[tbhp]
+\begin{figure}[tbhp!]
 \centering
-\includegraphics[width=15.5cm]{topelecdependexposure}
-\caption{Counties with the highest estimated physical exposure among electricity-dependent Medicare beneficiaries exposed to storms per year in U.S.
-counties for each exposure metric.
-The criteria behind each of the five metrics is given in Table 1 of the main text and details of the physical exposure calculation are given in the Methods.
-The color of each bar indicates the number of Medicare beneficiaries in the county reliant on electricity-dependent medical and assistive equipment as of July 2017 \citep{empower}.
-The length of each bar shows the average expected number of these electricity-dependent Medicare beneficiaries exposed to tropical storms per year based on a given exposure metric.}
-\label{fig:topelecdependexposure}
-\end{figure*}
-
-\clearpage
+\includegraphics[width=\linewidth]{figures/extentdisagreement}
+\caption{Examples of storms for which storm extent, as measured by the number of counties exposed, differed substantially between rain-based and wind-based metrics.
+Counties shown in orange were classified as exposed to a given storm by a given metric; those in yellow were classified as not exposed.
+The storm's track is shown in red.
+For both Frances (2004) and Isidore (2002) (top panel), storm extent was much larger when measuring exposure based on rain rather than wind.
+Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel), storm extent was large when exposure was measured based on wind compared to rain.}
+\label{fig:extentdisagreement}
+\end{figure}
 
@@
 \centering
-\includegraphics[width=0.6\linewidth]{windcomparison}
-\caption{Comparison of estimated wind categories from modeled sustained wind speed (as used for the wind exposure metric in the main text; \citep{stormwindmodel}) versus sustained wind speed determined from the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
+\includegraphics[width=\linewidth]{figures/othertopstorms}
+\caption{Counties exposed, based on different exposure metrics, to the storms with largest extent (as measured by the number of counties exposed based on any metric) from each of the clusters shown in the Jaccard heatmap in the main text (Fig.
+4).}
+\label{fig:othertopstorms}
+\end{figure}
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=0.6\linewidth]{figures/windcomparison}
+\caption{Comparison of estimated wind categories from modeled sustained wind speed (as used for the wind exposure metric in the main text; \cite{stormwindmodel}) versus sustained wind speed determined from the revised Atlantic hurricane database (HURDAT2 \cite{landsea2013}).
 In HURDAT2, wind radii were available for study storms from 2004--2015 and gives estimates of the maximum radii to which storm winds of 34 knots, 50 knots, and 64 knots extended in each four quadrant of the storm at each of the six-hour storm track observations.
 We interpolated this data to 15-minute increments and assessed a county as exposed to winds at or above one of these cutoffs if its population mean center was within 85\% of the maximum radius for that wind speed in its quadrant of the storm.
-For each county in each of the available storms, we also calculated the modeled wind speed \citep{stormwindmodel} and then categorized this value into the same categories available for the HURDAT2 wind radii ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots).
+For each county in each of the available storms, we also calculated the modeled wind speed \cite{stormwindmodel} and then categorized this value into the same categories available for the HURDAT2 wind radii ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots).
 Each point in this graph shows, for a storm, the percent of counties in which the wind category ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots) was the same based on both modeled winds and the HURDAT2 wind radii.
@@
 \centering
-\includegraphics[width=0.9\linewidth]{raincomparison}
-\caption{Comparison of county-level rain exposure based on the reanalysis data used for exposure assessment in this study (NLDAS-2) \citep{rui2013nldas, alhamdan2014environmental} and ground-based weather stations \citep{menne2012overview, rnoaa, countyweather}.
+\includegraphics[width=0.9\linewidth]{figures/raincomparison}
+\caption{Comparison of county-level rain exposure based on the reanalysis data used for exposure assessment in this study (NLDAS-2) \cite{rui2013nldas, alhamdan2014environmental} and ground-based weather stations \cite{menne2012overview, rnoaa, countyweather}.
 Each small graph shows a sample county, and each point on the graph represents a tropical storm that came within 500 kilometers of the population mean center of the county.
@@
 Note that x- and y-axis ranges differ by county.
-An estimate of the rank correlation (Spearman's $\rho$ \citep{spearman1904proof}) between measurements from the two rain data sources is also shown on each graph.}
+An estimate of the rank correlation (Spearman's $\rho$ \cite{spearman1904proof}) between measurements from the two rain data sources is also shown on each graph.}
 \label{fig:raincomparison}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{floodcomparison}
-\caption{Comparison of for a sample of study counties of flood status, based on NOAA Storm Events listings \citep{stormevents, noaastormevents}, with total streamflows at county streamflow gages during tropical storms \citep{usgsgages, countyfloods, dataRetrieval}.
+\includegraphics[width=0.8\linewidth]{figures/floodcomparison}
+\caption{Comparison of for a sample of study counties of flood status, based on NOAA Storm Events listings \cite{stormevents, noaastormevents}, with total streamflows at county streamflow gages during tropical storms \cite{usgsgages, countyfloods, dataRetrieval}.
 Each panel represents a separate county.
@@
 In the graphs, each point represents a single tropical storm, and the point's position along the x-axis shows the highest daily total streamflow (cubic feet / second), summed across all identified streamgages in the county, for a five-day window centered on the day of the storm's closest approach to the county.
-The y-axis separates storms for which a flood event was reported in NOAA's Storm Events database for the county with a start date within the five-day window of the storm's closest approach to the county \citep{stormevents, noaastormevents}.
+The y-axis separates storms for which a flood event was reported in NOAA's Storm Events database for the county with a start date within the five-day window of the storm's closest approach to the county \cite{stormevents, noaastormevents}.
 Boxplots are added to show the distribution of the total streamflow measurements across storms in the county for storms grouped by whether or not they were linked with a flood event based on the NOAA Storm Events data.
-The color of each point gives the percent of streamflow gages in the county with a daily streamflow that exceeded a threshold of flooding (the streamgage's median value for annual peak flow \citep{countyfloods}) on any day during the five-day window.
+The color of each point gives the percent of streamflow gages in the county with a daily streamflow that exceeded a threshold of flooding (the streamgage's median value for annual peak flow \cite{countyfloods}) on any day during the five-day window.
 Note that the x-axis scales differ by county, depending on the number of streamflow gages and typical flow rates for each gage, and are on a log-10 scale.}
@@
 \centering
-\includegraphics[width=\linewidth]{windexamples}
-\caption{Wind classifications for the two storms, Hurricanes Sandy (2012) and Ike (2008) for which exposure classifications from modeled wind data \citep{stormwindmodel} and HURDAT2 wind radii \citep{demuth2006improvement} showed the highest disagreement for storms between 2004 and 2015 (Fig.
+\includegraphics[width=\linewidth]{figures/windexamples}
+\caption{Wind classifications for the two storms, Hurricanes Sandy (2012) and Ike (2008) for which exposure classifications from modeled wind data \cite{stormwindmodel} and HURDAT2 wind radii \cite{demuth2006improvement} showed the highest disagreement for storms between 2004 and 2015 (Fig.
 \ref{fig:windcomparison}).
@@
 \centering
-\includegraphics[width=0.9\linewidth]{rainexamples}
+\includegraphics[width=0.9\linewidth]{figures/rainexamples}
 \caption{Examples of storms where the rain exposure classification method fails to capture full storm-related rainfall because of the distance constraint used in that assessment metric.
-The red line on each map shows the track of the storm, based on the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
+The red line on each map shows the track of the storm, based on the revised Atlantic hurricane database (HURDAT2 \cite{landsea2013}).
 The color of each county in the map gives the cumulative precipitation (in millimeters) for that county during the window from two days before to one day after the storm's closest approach to the county.
@@
 
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=15.5cm]{figures/topelecdependexposure}
+\caption{Counties with the highest estimated physical exposure among electricity-dependent Medicare beneficiaries exposed to storms per year in U.S.
+counties for each exposure metric.
+The criteria behind each of the five metrics is given in Table 1 of the main text and details of the physical exposure calculation are given in the Methods.
+The color of each bar indicates the number of Medicare beneficiaries in the county reliant on electricity-dependent medical and assistive equipment as of July 2017 \cite{empower}.
+The length of each bar shows the average expected number of these electricity-dependent Medicare beneficiaries exposed to tropical storms per year based on a given exposure metric.}
+\label{fig:topelecdependexposure}
+\end{figure*}
+
+\clearpage
+
+\subsection*{SI Bibliography}
+
 \bibliography{hurr_exposure}

### 2019-01-26 writing/ehp_tables.tex
Add a bunch of stuff

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{1.5cm}p{2.5cm}p{10cm}} 
\toprule
Metric & Available & Criteria for exposure \\ \midrule 
Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\ 
Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
Tornado & 1996--2015 & Tornado event listed NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}

\clearpage

\input{../tables/stormseverity.tex}

### 2019-01-26 R/storm_severity_table.R

@@ -6,66 +6,177 @@ library(xtable)
 library(stringr)
 
 my_fips <- unique(closest_dist$fips)
+
 distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
                                      end_year = 2015, dist_limit = 100) %>%
         select(storm_id, fips) %>%
-        mutate(distance_exposed = TRUE,
+        mutate(exposed = TRUE,
                storm_id = as.character(storm_id))
+
+dist_sum <- distance_exposure %>%  
+  group_by(storm_id) %>% 
+  count() %>% 
+  ungroup() %>% 
+  summarize(n_storms = n() / (2015 - 1988 + 1),
+            n_exposures = sum(n) / (2015 - 1988 + 1),
+            med = median(n),
+            p25 = quantile(n, 0.25),
+            p75 = quantile(n, 0.75),
+            max = max(n),
+            which_max = storm_id[which.max(n)])
+
 rain_exposure <- county_rain(counties = my_fips,
                              start_year = 1988, end_year = 2011,
                              rain_limit = 75, dist_limit = 500) %>%
         select(storm_id, fips) %>%
-        mutate(rain_exposed = TRUE,
+        mutate(exposed = TRUE,
                storm_id = as.character(storm_id))
+rain_sum <- rain_exposure %>%  
+  group_by(storm_id) %>% 
+  count() %>% 
+  ungroup() %>% 
+  summarize(n_storms = n() / (2011 - 1988 + 1),
+            n_exposures = sum(n) / (2011 - 1988 + 1),
+            med = median(n),
+            p25 = quantile(n, 0.25),
+            p75 = quantile(n, 0.75),
+            max = max(n),
+            which_max = storm_id[which.max(n)])
+
 wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
                              end_year = 2015, wind_limit = 15) %>%
         select(storm_id, fips) %>%
-        mutate(wind_exposed = TRUE,
+        mutate(exposed = TRUE,
                storm_id = as.character(storm_id))
+wind_sum <- wind_exposure %>%  
+  group_by(storm_id) %>% 
+  count() %>% 
+  ungroup() %>% 
+  summarize(n_storms = n() / (2015 - 1988 + 1),
+            n_exposures = sum(n) / (2015 - 1988 + 1),
+            med = median(n),
+            p25 = quantile(n, 0.25),
+            p75 = quantile(n, 0.75),
+            max = max(n),
+            which_max = storm_id[which.max(n)])
+
[…]

### 2019-01-27 writing/anderson_ehp_supp.tex
Rearrange content

@@
-\documentclass[11pt, titlepage]{article}
-\usepackage{microtype}
-\usepackage{charter}
+\documentclass[11pt, titlepage, twoside]{article}
 
@@
 
-\title{Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"}
+\title{Supplemental Material for ``Assessing United States county-level
+exposure for research on tropical cyclones and human health"}
 
@@
 
-\section*{Supplemental Text}
-
-\subsection*{Case study: Physical exposure of electricity-dependent Medicare beneficiaries to tropical cyclones}
-
-The outcomes of disasters depend both on the geophysical forces of the disaster as well as on the vulnerability of the society living in the geographical areas affected by those forces.
-\citep{chakraborty2005population, anderson2003community, cutter1996vulnerability} As a case study of how differences in tropical cyclone exposure assessments across different hazard metrics influence estimates of physical exposures of susceptible populations to tropical cyclones, we investigated how the use of different metrics influenced estimates of which eastern U.S.
-counties have the highest expected average exposures of electricity-dependent Medicare beneficiaries to tropical cyclones.
-This subpopulation was selected for this analysis since it may be particularly susceptible to health impacts from tropical cyclone exposure, especially through the pathway of storm-associated power outages and evacuations.
-We collected data on the number of electricity-dependent Medicare beneficiaries in each study county from the U.S.
-Department of Health \& Human Service's emPower Map 2.0 \citep{empower}.
-We then calculated the physical exposure of the electricity-dependent Medicare population in each county, based on tropical cyclone assessments using each hazard metric, following \citep{peduzzi2009assessing}:
-
-\begin{equation}
-E_c = F_c * P_c
-\end{equation}
-
-\noindent where $E_c$ is the average yearly physical exposure among electricity-dependent Medicare beneficiaries in county $c$ to tropical cyclone exposures based on a given metric, $F_c$ is the estimated yearly expected frequency of tropical cyclone exposures in county $c$ based on that metric, and $P_c$ is the size of the electricity-dependent Medicare population in county $c$, as of July 2017.
-When combined with estimates of vulnerability of a population to a natural hazard, such measurements of physical exposure can be used to calculate risk of human losses from the hazard \citep{peduzzi2009assessing}.
-
-We found that a few counties (Miami-Dade County, FL; Harris County, TX) were ranked in the top three counties of physical exposure among electricity-dependent Medicare beneficiaries for almost all exposure metrics (Figure \ref{fig:topelecdependexposure}).
-However, there were key differences when using different exposure metrics.
-For example, Florida counties dominated the lists of top exposed counties based on wind- and tornado-based metrics, while over half of the counties for the flood-based metric were in Mid-Atlantic states (PA, NY).
-
-Dissimilarity in exposure to specific hazards of tropical cyclones can result in differing patterns in estimates of county-level exposure (Fig.
-1 of main text), which can lead to substantial differences in estimates of population physical exposure (Fig.
-\ref{fig:topelecdependexposure}), based on which storm hazard is considered.
-The measure of a hurricane's intensity, the Saffir-Simpson scale, expresses intensity as a function of wind speed, and so fails to capture potential risks from other hazards of the storm \citep{smith2009}, and previous assessments of physical exposure to tropical cyclones have often focused on winds or distance from the storm.
-For example, a global study of exposure and vulnerability to natural hazards used a combination of windspeed above a certain threshold and distance from the storm's track to assess exposure \citep{peduzzi2009assessing}.
-While many storm impacts might be most strongly linked to wind hazards, there is a growing recognition of the potential risks of adverse health outcomes and property damage from rain- and flood-related tropical cyclone hazards in the U.S.
-\citep{smith2009}.
-Such assessments may underemphasize potentially dangerous exposures to certain regions of the U.S., particularly inland locations.
-In fact, based on our assessment of physical exposures to tropical cyclones among electricity-dependent Medicare beneficiaries, the highest expected physical exposure under any metric was measured for Philadelphia County, PA, under the flood metric (Fig.
-\ref{fig:topelecdependexposure}).
-While this county has a similar-sized population of electricity-dependent Medicare beneficiaries as Miami-Dade County, FL (Fig.
-\ref{fig:topelecdependexposure}), the frequency of storm-related flood events in this inland county results in a very high expected physical exposure among this subpopulation, while this county is not included in the top ten counties for physical exposure based on most of the other exposure metrics.
-
-\clearpage
-
 \begin{figure}[tbhp!]
@@
 \includegraphics[width=\linewidth]{previous_exposure_metrics}
-\caption{Metrics used in a sample of previous studies to assign exposure to tropical cyclones in U.S.-based tropical cyclone exposure or impact studies.
-Citations for each study are shown on the left and the broad topic of the study on the left.
-Presence of a bar for a study-metric combination indicates that the given metric was used in the study in assessing exposure to tropical cyclones for that study, either alone or in combination with other metrics for which that study has bars.
-The color transparency of each bar indicates the number of different thresholds considered for defining storm exposure in the study under that hazard (e.g., less transparent bars indicate that a study considered multiple thresholds of the metric in defining tropical cyclone exposure as a method of sensitivity analysis).
-Studies for which there are two or more bars (e.g., bars for both distance and wind) indicate studies that either used separate exposure assessments, one for each metric, (e.g., as a sensitivity analysis) or used a single exposure index that incorporated multiple metrics in its definition.
-``Damage" often indicate use of Federal Emergency Management Agency (FEMA) reports, although occasionally used damage reports from other sources, including newspaper reports.
-Examples of ``Other" include evacuation orders, excessive school closures, length of time a storm was near a location, and detailed, holistic analyses of a specific storm or storms.
-References for each study are included in the
-bibliography of this Supplemental Material.}
+\caption{Metrics used in a sample of previous studies to assign exposure to
+tropical cyclones in \ac{US}-based tropical cyclone exposure or impact studies.
+Citations for each study are shown on the left and the broad topic of the study
+on the left. Presence of a bar for a study-metric combination indicates that
+the given metric was used in the study in assessing exposure to tropical
+cyclones for that study, either alone or in combination with other metrics for
+which that study has bars. The color transparency of each bar indicates the
+number of different thresholds considered for defining storm exposure in the
+study under that hazard (e.g., less transparent bars indicate that a study
+considered multiple thresholds of the metric in defining tropical cyclone
+exposure as a method of sensitivity analysis). Studies for which there are two
+or more bars (e.g., bars for both distance and wind) indicate studies that
+either used separate exposure assessments, one for each metric, (e.g., as a
+sensitivity analysis) or used a single exposure index that incorporated
+multiple metrics in its definition. ``Damage" often indicate use of Federal
+Emergency Management Agency (FEMA) reports, although occasionally used damage
+reports from other sources, including newspaper reports. Examples of ``Other"
+include evacuation orders, excessive school closures, length of time a storm
+was near a location, and detailed, holistic analyses of a specific storm or
+storms. References for each study are included in the bibliography of this
+Supplemental Material.}
 \label{fig:previousmetrics}
@@
 
-\nocite{logan2015, gares2014, esnard2011, zandbergen2009, horney2016, grabich2016, 
-antipova2015post, zahran2010, currie2013, grabich2015, grabich2016hurricane, amstadter2010,
-acierno2006, le2013, lieberman2017, mckinney2011, dosa2012evacuate, swerdel2014, 
-czajkowski2011, mongin2017, kessler2007hurricane, caillouet2008increase, 
-rosenheim2018disaster, domino2003disasters, kim2016, kinney2008, kim2015, tansel2010,
-escobedo2009, fuller2014, baggerly2008impact, strobl2011economic, mukherjee2017}
+\nocite{logan2015, gares2014, esnard2011, zandbergen2009, horney2016,
+grabich2016, antipova2015post, zahran2010, currie2013, grabich2015,
+grabich2016hurricane, amstadter2010, acierno2006, le2013, lieberman2017,
+mckinney2011, dosa2012evacuate, swerdel2014, czajkowski2011, mongin2017,
+kessler2007hurricane, caillouet2008increase, rosenheim2018disaster,
+domino2003disasters, kim2016, kinney2008, kim2015, tansel2010, escobedo2009,
+fuller2014, baggerly2008impact, strobl2011economic, mukherjee2017}
 
@@
 \includegraphics[width=0.7\linewidth]{hurrtracks}
-\caption{Tracks of all tropical cyclones included in the study.
-The study included all tracked storms in the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}) that came within 250 kilometers of at least on U.S.
-county, 1988--2015.
-Thicker lines show the tracks of storms whose names were retired, indicating that the storm was particularly severe or had notable impacts \citep{retirednames}.}
+\caption{Tracks of all tropical cyclones included in the study. The study
+included all tracked storms in the revised Atlantic hurricane database
+\ac{HURDAT2}~\citep{landsea2013} that came within 250~\si{\kilo\metre} of at least on
+\ac{US} county, 1988\,--\,2015. Thicker lines show the tracks of storms whose
+names were retired, indicating that the storm was particularly severe or had
+notable impacts~\citep{retirednames}.}
 \label{fig:hurrtracks}
@@
 \includegraphics[width=\linewidth]{othertopstorms}
-\caption{Counties exposed, based on different exposure metrics, to the storms with largest extent (as measured by the number of counties exposed based on any metric) from each of the clusters shown in the Jaccard heatmap in the main text (Figure 2).}
+\caption{Counties exposed, based on different exposure metrics, to the storms
+with largest extent (as measured by the number of counties exposed based on any
+metric) from each of the clusters shown in the Jaccard heatmap in the main text
+(Figure 2).}
 \label{fig:othertopstorms}
@@
 \includegraphics[width=\linewidth]{extentdisagreement}
-\caption{Examples of storms for which exposure assessments differed substantially between rain-based and wind-based metrics.
-Counties shown in orange were classified as exposed to a given storm by a given metric; those in yellow were classified as not exposed.
-The storm's track is shown in red.
-For both Frances (2004) and Isidore (2002) (top panel), substantially more counties were assessed as exposed when measuring exposure based on rain rather than wind.
-Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel), substantially more counties were assessed as exposed based on wind compared to rain.}
+\caption{Examples of storms for which exposure assessments differed
+substantially between rain-based and wind-based metrics. Counties shown in
+orange were classified as exposed to a given storm by a given metric; those in
+yellow were classified as not exposed. The storm's track is shown in red. For
+both Frances (2004) and Isidore (2002) (top panel), substantially more counties
+were assessed as exposed when measuring exposure based on rain rather than
+wind. Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel),
+substantially more counties were assessed as exposed based on wind compared to
+rain.}
 \label{fig:extentdisagreement}
@@
 \includegraphics[width=\linewidth]{averageexposureonlysupp}
-\caption{Average number of storm exposure per decade in U.S.
-counties for each exposure metric, based on years of for which data for all five exposure metrics are available (1996--2011).
-The criteria behind each of the five metrics is given in Table 1 of the main text.}
-\label{fig:averageexposuresupp}
+\caption{Average number of storm exposure per decade in \ac{US} counties for
+each exposure metric, based on years of for which data for all five exposure
+metrics are available (1996\,--\,2011). The criteria behind each of the five
+metrics is given in Table 1 of the main text.} \label{fig:averageexposuresupp}
 \end{figure}
-
-\clearpage
-
-\begin{figure*}%[tbhp]
-\centering
-\includegraphics[width=15.5cm]{topelecdependexposure}
-\caption{Counties with the highest estimated physical exposure among electricity-dependent Medicare beneficiaries exposed to storms per year in U.S.
-counties for each exposure metric.
-The criteria behind each of the five metrics is given in Table 1 of the main text and details of the physical exposure calculation are given in the Methods.
-The color of each bar indicates the number of Medicare beneficiaries in the county reliant on electricity-dependent medical and assistive equipment as of July 2017 \citep{empower}.
-The length of each bar shows the average expected number of these electricity-dependent Medicare beneficiaries exposed to tropical storms per year based on a given exposure metric.}
-\label{fig:topelecdependexposure}
-\end{figure*}
 
@@
 \centering
-\includegraphics[width=0.6\linewidth]{windcomparison}
-\caption{Comparison of estimated wind categories from modeled sustained wind speed (as used for the wind exposure metric in the main text; \citep{stormwindmodel}) versus sustained wind speed determined from the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
-In HURDAT2, wind radii were available for study storms from 2004--2015 and gives estimates of the maximum radii to which storm winds of 34 knots, 50 knots, and 64 knots extended in each four quadrant of the storm at each of the six-hour storm track observations.
-We interpolated this data to 15-minute increments and assessed a county as exposed to winds at or above one of these cutoffs if its population mean center was within 85\% of the maximum radius for that wind speed in its quadrant of the storm.
-For each county in each of the available storms, we also calculated the modeled wind speed \citep{stormwindmodel} and then categorized this value into the same categories available for the HURDAT2 wind radii ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots).
-Each point in this graph shows, for a storm, the percent of counties in which the wind category ($<$34 knots; 34--49.9 knots; 50--63.9 knots; $\ge$64 knots) was the same based on both modeled winds and the HURDAT2 wind radii.
-Only storms with at least one county with sustained wind of $\ge$34 knots, based on the HURDAT2 wind radii, are shown.
-Hurricanes Sandy (2012) and Ike (2008) were he two storms with the worst agreement between these two sets of wind data; maps with further comparisons between the two data sources for these two storms are given in Fig.
-\ref{fig:windexamples}.}
-\label{fig:windcomparison}
-\end{figure}
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=0.9\linewidth]{raincomparison}
-\caption{Comparison of county-level rain exposure based on the reanalysis data used for exposure assessment in this study (NLDAS-2) \citep{rui2013nldas, alhamdan2014environmental} and ground-based weather stations \citep{menne2012overview, rnoaa, countyweather}.
-Each small graph shows a sample county, and each point on the graph represents a tropical storm that came within 500 kilometers of the population mean center of the county.
-The x-axis shows the estimated cumulative rainfall (in millimeters) from two days before to one day after the storm's closest approach to the county based on the county-aggregated NLDAS-2 precipitation data used to measure rain exposure in our main analysis.
-The ground-based observations (y-axis) are based on collecting precipitation data from available weather stations in the county, 1988--2011, summed over the same period from two days before to one day after the storm's closest approach for each available station.
-These cumulative station-based precipitation totals were then averaged across all available stations in the county to create a county-wide estimate of cumulative storm-related precipitation for each storm in the county.
-Horizontal and vertical lines show the threshold used to classify a storm as exposed based on cumulative rainfall; storms in the lower left and upper right quadrants are classified the same regardless of whether reanalysis or station-based precipitation data is considered, while storms in the upper left and lower right quadrants would be classified differently.
-The number of storms within each county and the average number of stations reporting rainfall during the county's storms are given above each plot.
-Note that x- and y-axis ranges differ by county.
-An estimate of the rank correlation (Spearman's $\rho$ \citep{spearman1904proof}) between measurements from the two rain data sources is also shown on each graph.}
-\label{fig:raincomparison}
-\end{figure}
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=0.8\linewidth]{floodcomparison}
-\caption{Comparison of for a sample of study counties of flood status, based on NOAA Storm Events listings \citep{stormevents, noaastormevents}, with total streamflows at county streamflow gages during tropical storms \citep{usgsgages, countyfloods, dataRetrieval}.
-Each panel represents a separate county.
-For each county, we first identified all streamflow gages in the county that had complete data for Jan.
-1 1996--Dec.
-31, 2015.
-The number of streamflow gages meeting this criterion are given in parentheses beside the county's name in the panel title.
-This analysis covered tropical storm that came within 500 kilometers of each sample county between 1996 and 2015.
-In the graphs, each point represents a single tropical storm, and the point's position along the x-axis shows the highest daily total streamflow (cubic feet / second), summed across all identified streamgages in the county, for a five-day window centered on the day of the storm's closest approach to the county.
-The y-axis separates storms for which a flood event was reported in NOAA's Storm Events database for the county with a start date within the five-day window of the storm's closest approach to the county \citep{stormevents, noaastormevents}.
-Boxplots are added to show the distribution of the total streamflow measurements across storms in the county for storms grouped by whether or not they were linked with a flood event based on the NOAA Storm Events data.
-The color of each point gives the percent of streamflow gages in the county with a daily streamflow that exceeded a threshold of flooding (the streamgage's median value for annual peak flow \citep{countyfloods}) on any day during the five-day window.
-Note that the x-axis scales differ by county, depending on the number of streamflow gages and typical flow rates for each gage, and are on a log-10 scale.}
-\label{fig:floodcomparison}
-\end{figure}
-
-\begin{figure}[tbhp!]
-\centering
 \includegraphics[width=\linewidth]{windexamples}
-\caption{Wind classifications for the two storms, Hurricanes Sandy (2012) and Ike (2008) for which exposure classifications from modeled wind data \citep{stormwindmodel} and HURDAT2 wind radii \citep{demuth2006improvement} showed the highest disagreement for storms between 2004 and 2015 (Fig.
-\ref{fig:windcomparison}).
-The top two maps show modeled (left) and HURDAT2 (right) windspeed categorizations for Hurricane Ike (2008), while the bottom two maps show the same windspeed characterizations for Hurricane Sandy (2012).
-In both cases, the storms were exceptionally large systems for which high winds persisted well inland from landfall and for which, based on the HURDAT2 data, a large number of counties were exposed to winds of $\ge$34 knots (Figure \ref{fig:windcomparison}).}
+\caption{Wind classifications for the two storms, Hurricanes Sandy (2012) and
+Ike (2008) for which exposure classifications from modeled wind
+data~\citep{stormwindmodel} and \ac{HURDAT2} wind
+radii~\citep{demuth2006improvement} showed the highest disagreement for storms
+between 2004 and 2015 (Figure~\ref{fig:windcomparison}). The top two maps show
+modeled (left) and \ac{HURDAT2} (right) windspeed categorizations for Hurricane
+Ike (2008), while the bottom two maps show the same windspeed characterizations
+for Hurricane Sandy (2012).  In both cases, the storms were exceptionally large
+systems for which high winds persisted well inland from landfall and for which,
+based on the \ac{HURDAT2} data, a large number of counties were exposed to
+winds of $\ge$34 knots (Figure~\ref{fig:windcomparison}).}
 \label{fig:windexamples}
@@
 \includegraphics[width=0.9\linewidth]{rainexamples}
-\caption{Examples of storms where the rain exposure classification method fails to capture full storm-related rainfall because of the distance constraint used in that assessment metric.
-The red line on each map shows the track of the storm, based on the revised Atlantic hurricane database (HURDAT2 \citep{landsea2013}).
-The color of each county in the map gives the cumulative precipitation (in millimeters) for that county during the window from two days before to one day after the storm's closest approach to the county.
-The blue outline identifies the collection of counties that were classified as ``exposed" based on the criteria that the storm passed within 500 kilometers of the county's population mean center and the cumulative rainfall was 75 millimeters or more (Table 1 of main text).
-The map on the left shows rain-based exposure for Hurricane Ike.
-For this exceptionally large storm, there are some counties in northern Texas that received high rainfall from the storm but are too far from the storm center's track to be classified as exposed by this metric.
-The map on the right shows rain-based exposure for Tropical Storm Lee.
-For this storm, rain from the storm affected counties in the Mid-Atlantic and New England as the remnants of the storm moved over those regions, but storm tracking stopped while the storm was still over the southeastern part of the U.S., and so these counties are not classified as exposed to the storm based on this metric.}
+\caption{Examples of storms where the rain exposure classification method fails
+to capture full storm-related rainfall because of the distance constraint used
+in that assessment metric. The red line on each map shows the track of the
+storm, based on the revised Atlantic hurricane database
+\ac{HURDAT2}~\citep{landsea2013}. The color of each county in the map gives the
+cumulative precipitation (in millimeters) for that county during the window
+from two days before to one day after the storm's closest approach to the
+county. The blue outline identifies the collection of counties that were
+classified as ``exposed" based on the criteria that the storm passed within
+500~\si{\kilo\metre} of the county's population mean center and the cumulative
+rainfall was 75 millimeters or more (Table 1 of main text). The map on the left
+shows rain-based exposure for Hurricane Ike. For this exceptionally large
+storm, there are some counties in northern Texas that received high rainfall
+from the storm but are too far from the storm center's track to be classified
+as exposed by this metric. The map on the right shows rain-based exposure for
+Tropical Storm Lee. For this storm, rain from the storm affected counties in
+the Mid-Atlantic and New England as the remnants of the storm moved over those
+regions, but storm tracking stopped while the storm was still over the
+southeastern part of the \ac{US}, and so these counties are not classified as
+exposed to the storm based on this metric.}
 \label{fig:rainexamples}

### 2019-01-27 writing/ehp_abstract.tex
Rearrange content

@@
 
-\noindent \textbf{Background:} Tropical cyclones bring severe impacts to U.S.
-communities, including substantial human health risks, from hazards that
-include wind, rain, flooding, and tornadoes. However, studies quantifying
-tropical cyclone risks and impacts vary widely in how they classify
-exposure to tropical cyclones, using various hazard-based metrics and, in some
-cases, using distance from the storm track as a surrogate. \\
-\textbf{Objectives:} (1) Investigate how well county-level assessments of
+\noindent \textbf{Background:} Tropical cyclones bring severe impacts to
+\ac{US} communities, including substantial human health risks, from hazards
+that include wind, rain, flooding, and tornadoes. However, studies quantifying
+tropical cyclone risks and impacts vary widely in how they classify exposure to
+tropical cyclones, using various hazard-based metrics and, in some cases, using
+distance from the storm track as a surrogate. \\ 
+\textbf{Objectives:} (1)~Investigate how well county-level assessments of
 tropical cyclone exposure agree when based on distance versus hazard-based
 measurements, as well as among assessments based on different hazard
-measurements, and (2) provide software to improve tropical cyclone exposure
+measurements, and (2)~provide software to improve tropical cyclone exposure
 assessment for health-related research. \\ 
@@
 rainfall, flooding, and tornadoes for all land-falling or near-land Atlantic
-basin storms, for 1996--2011 for all metrics and up to 1988--2015 for specific
-metrics. We quantified how well county-level exposure assessment agreed for all
-pairwise combinations of metrics.\\ 
+basin storms, for~1996\,--\,2011 for all metrics and up to~1988\,--\,2015 for
+specific metrics. We quantified how well county-level exposure assessment
+agreed for all pairwise combinations of metrics.\\ 
 \textbf{Results:} In most of the storms investigated, agreement in county-level

### 2019-01-27 writing/ehp_discussion.tex
Rearrange content

@@
 cyclones.  However, tropical cyclones are multi-hazard events, making it more
-complicated to measure exposure compared to other weather
-exposures like temperature and precipitation. Here, we find the county-level
-tropical cyclone exposure assessments vary substantially when using a
-distance-based metric versus four hazard-based metrics, as well as among
-different hazard-based metrics. Our results can inform exposure assessment for
-future county-level studies of the health risks and impacts associated with tropical cyclones
+complicated to measure exposure compared to other weather exposures like
+temperature and precipitation. Here, we find the county-level tropical cyclone
+exposure assessments vary substantially when using a distance-based metric
+versus four hazard-based metrics, as well as among different hazard-based
+metrics. Our results can inform exposure assessment for future county-level
+studies of the health risks and impacts associated with tropical cyclones
 exposures, and the data and software we have created to accompany this article
@@
 been used as an operational metric of exposure to tropical cyclones in previous
-large-scale studies (examples included in Figure S1). Since distance itself
+large-scale studies (examples include ...). Since distance itself
 does not constitute a hazard, distance is meant in these cases as a surrogate
 to capture exposure to hazards from the tropical cyclone. However, here we
-found that in assessing U.S.  county-level exposure to tropical cyclones,
+found that in assessing \ac{US} county-level exposure to tropical cyclones,
 distance is, at best, a moderate, and often a very poor, surrogate for exposure
 to the specific tropical cyclone hazards of high wind, extreme rainfall,
-flooding, and tornadoes (Figure \ref{fig:jaccard}).  Therefore, use of distance
+flooding, and tornadoes (Figure~\ref{fig:jaccard}).  Therefore, use of distance
 to assess tropical cyclone exposure for impact studies could result in
@@
 even strong associations, between tropical cyclone exposure and outcomes of
-interest in impact studies \citep{savitz2016interpreting, armstrong1998effect}.
+interest in impact studies~\citep{savitz2016interpreting, armstrong1998effect}.
 
@@
 of interest or in factors associated with risk of that outcome, differential
-exposure misclassification would exist \citep{savitz2016interpreting}. Such
+exposure misclassification would exist~\citep{savitz2016interpreting}. Such
 differential exposure misclassification can bias estimates of tropical cyclone
@@
 compared to the true association that exists) or away from the null (estimating
-a larger association than actually exists) \citep{savitz2016interpreting,
+a larger association than actually exists)~\citep{savitz2016interpreting,
 armstrong1998effect}. The use of a single hazard-based metric (e.g., wind)
@@
 
-Our findings---that county-level tropical cyclone exposure assessment varies
-substantially when different exposure measurements are considered---aligns with
+Our findings\,---\,that county-level tropical cyclone exposure assessment varies
+substantially when different exposure measurements are considered\,---\,aligns with
 previous research from atmospheric science and related fields on the
 characteristics of tropical cyclones. While tropical cyclone rainfall and
-windspeed are often well-correlated when the tropical cyclone is over water
-\citep{cerveny2000}, this relationship often does not remain as strong once the
-hurricane has made landfall \citep{jiang2008}.  Fast-moving tropical cyclones
-bring higher risks of dangerous winds inland \citep{kruk2010}, while
-slow-moving tropical cyclone are likely to bring more rain
-\citep{rappaport2000} and cause more damage because of sustained hazardous
-conditions \citep{rezapour2014}. Further, while the likelihood and extent of
+windspeed are often well-correlated when the tropical cyclone is over
+water~\citep{cerveny2000}, this relationship often does not remain as strong
+once the hurricane has made landfall~\citep{jiang2008}.  Fast-moving tropical
+cyclones bring higher risks of dangerous winds inland~\citep{kruk2010}, while
+slow-moving tropical cyclone are likely to bring more
+rain~\citep{rappaport2000} and cause more damage because of sustained hazardous
+conditions~\citep{rezapour2014}. Further, while the likelihood and extent of
 flooding during a tropical cyclone is related to the tropical cyclone's
 rainfall, it is also driven by factors like top soil saturation and the
-structure of the water basin's drainage network \citep{chen2015, rees2001}.
-However, we did find a small set of tropical cyclone for which for which
-agreement between metrics were unusually high (e.g., Floyd in 1999, Irene in
-2011, Hannah in 2008, Bertha in 1996) For this set of tropical cyclones, the
+structure of the water basin's drainage network~\citep{chen2015, rees2001}.
+However, we did find a small set of tropical cyclones for which for which
+agreement between metrics were unusually high (e.g., Floyd in~1999, Irene in
+2011, Hannah in~2008, Bertha in~1996) For this set of tropical cyclones, the
 tropical cyclones' persistent proximity to water may have helped maintain wind
@@
 In our results, wind-based exposure had a strong coastal pattern, with almost
-all exposures in counties within about 200 kilometers (124 miles) of the
-coastline. This is consistent with the
-dramatic decrease in wind intensity that typically characterizes the landfall
-of tropical cyclones. Exposures by rain- and distance-based metrics often extended further inland compared to the wind-based
-metric, up to the Appalachian mountains. This agrees with previous research
-indicating that the Appalachian mountains' topography both enhances
+all exposures in counties within about~200~\si{\kilo\metre} (124~mi) of the
+coastline. This is consistent with the dramatic decrease in wind intensity that
+typically characterizes the landfall of tropical cyclones. Exposures by rain-
+and distance-based metrics often extended further inland compared to the
+wind-based metric, up to the Appalachian mountains. This agrees with previous
+research indicating that the Appalachian mountains' topography both enhances
 precipitation during tropical cyclones and provides hydrological conditions for
-severe flooding \citep{rees2001}.  Almost all tornado-based exposures were in
-southern coastal states. This pattern echoes
-previous findings that most noteworthy tropical cyclone-related tornadoes occur
-on the right-side of the tropical cyclone track in Atlantic Basin U.S. storms
-\citep{moore2012}.
+severe flooding~\citep{rees2001}.  Almost all tornado-based exposures were in
+southern coastal states. This pattern echoes previous findings that most
+noteworthy tropical cyclone-related tornadoes occur on the right-side of the
+tropical cyclone track in Atlantic-basin \ac{US} storms~\citep{moore2012}.
 
-The tropical cyclone exposure averages shown in Figure
-\ref{fig:averageexposure} are limited as estimates of long-term frequencies, as
-tropical cyclones follow decadal patterns \citep{kossin2007more} likely not
-adequately captured in the available data. However, these frequency maps,
-together with evidence from specific tropical cyclones (Figures
-\ref{fig:ivanexposure} and \ref{fig:jaccard}),  do illustrate the potential for
-strong differences in spatial patterns in tropical cyclone exposures, depending
-on which tropical cyclone hazards are considered. A few previous studies have
-sought to determine county-level exposure to tropical cyclones over multi-year
-periods, including \citet{zandbergen2009}, which estimated exposure in U.S.
-counties to all U.S.  landfalling Atlantic basin tropical cyclones between 1851
-and 2003, using both a distance-based metric and a metric that combined
-distance and windspeed, and \citet{kruk2010}, which explored exposure to
-hurricane-related winds in the U.S., including inland areas, for 1900--2008.
-Our results suggest that such exposure assessments may perform well in
-capturing some tropical cyclone hazards (e.g., wind), but likely miss other
-potentially dangerous tropical cyclone exposures, especially for hazards that
-repeatedly threaten northern or inland counties (e.g., rain, flooding).
+The tropical cyclone exposure averages shown in
+Figure~\ref{fig:averageexposure} are limited as estimates of long-term
+frequencies, as tropical cyclones follow decadal
+patterns~\citep{kossin2007more} likely not adequately captured in the available
+data.  However, these frequency maps, together with evidence from specific
+tropical cyclones (Figures~\ref{fig:ivanexposure} and~\ref{fig:jaccard}),  do
+illustrate the potential for strong differences in spatial patterns in tropical
+cyclone exposures, depending on which tropical cyclone hazards are considered.
+A few previous studies have sought to determine county-level exposure to
+tropical cyclones over multi-year periods, including~\citet{zandbergen2009},
+which estimated exposure in \ac{US} counties to all \ac{US} landfalling
+Atlantic-basin tropical cyclones between~1851 and~2003, using both a
+distance-based metric and a metric that combined distance and windspeed,
+and~\citet{kruk2010}, which explored exposure to hurricane-related winds in the
+\ac{US}, including inland areas, for~1900\,--\,2008.  Our results suggest that
+such exposure assessments may perform well in capturing some tropical cyclone
+hazards (e.g., wind), but likely miss other potentially dangerous tropical
+cyclone exposures, especially for hazards that repeatedly threaten northern or
+inland counties (e.g., rain, flooding).
+
+Dissimilarity in exposure to specific hazards of tropical cyclones can result
+in differing patterns in estimates of county-level exposure (Figure~1 [change
+to fig. ref]), which can lead to substantial differences in estimates of
+population physical exposure (Figure~\ref{fig:topelecdependexposure}), based on
+which storm hazard is considered. The measure of a hurricane's intensity, the
+Saffir-Simpson scale, expresses intensity as a function of wind speed, and so
+fails to capture potential risks from other hazards of the
+storm~\citep{smith2009}, and previous assessments of physical exposure to
+tropical cyclones have often focused on winds or distance from the storm. For
+example, a global study of exposure and vulnerability to natural hazards used a
+combination of windspeed above a certain threshold and distance from the
+storm's track to assess exposure~\citep{peduzzi2009assessing}. While many storm
+impacts might be most strongly linked to wind hazards, there is a growing
+recognition of the potential risks of adverse health outcomes and property
+damage from rain- and flood-related tropical cyclone hazards in the
+\ac{US}~\citep{smith2009}. Such assessments may underemphasize potentially
+dangerous exposures to certain regions of the \ac{US}, particularly inland
+locations. In fact, based on our assessment of physical exposures to tropical
+cyclones among electricity-dependent Medicare beneficiaries, the highest
+expected physical exposure under any metric was measured for Philadelphia
+County, PA, under the flood metric (Figure~\ref{fig:topelecdependexposure}).
+While this county has a similar-sized population of electricity-dependent
+Medicare beneficiaries as Miami-Dade County, FL
+(Figure~\ref{fig:topelecdependexposure}), the frequency of storm-related flood
+events in this inland county results in a very high expected physical exposure
+among this subpopulation, while this county is not included in the top ten
+counties for physical exposure based on most of the other exposure metrics.
 
@@
 
-To assist with future tropical cyclone studies, we created and published
-open source software to accompany this article \citep{hurricaneexposure,
-hurricaneexposuredata}. Many previous studies have used geographical
+To assist with future tropical cyclone studies, we created and published open
+source software to accompany this article~\citep{hurricaneexposuredata,
+hurricaneexposure}. Many previous studies have used geographical
 information system software (e.g., ArcGIS) to assess exposure to tropical
-cyclones in the U.S. \citep{grabich2016, zandbergen2009, czajkowski2011,
+cyclones in the \ac{US}~\citep{grabich2016, zandbergen2009, czajkowski2011,
 kruk2010}. Here, we offer methods to map and output historic exposure to
 tropical cyclones that does not require the use of proprietary software but
-instead uses a package written in the R statistical programming language
-\citep{R}, which is free and open-source.
+instead uses a package written in the R statistical programming
+language~\citep{R}, which is free and open-source.
 
@@
 exposure profiles based on multiple hazards or craft exposure indices that
-combine hazard metrics \citep{chakraborty2005population, peduzzi2009assessing}.
+combine hazard metrics~\citep{chakraborty2005population, peduzzi2009assessing}.
 This ability can be critical, as different hazards of tropical cyclones
-often act synergistically in causing impacts \citep{smith2009}. Further, by
+often act synergistically in causing impacts~\citep{smith2009}. Further, by
 including measurements of different hazard exposures in each county for each
@@
 In creating this hurricane exposure dataset, we aimed for data that are
-reliable, available for all eastern U.S. counties, and have no missing data
+reliable, available for all eastern \ac{US} counties, and have no missing data
 within available dates. Further, we have investigated how well these data
-correspond with data from other available sources (Figures S7--11). However,
-there are still some important limitations. For example, for a few tropical cyclones, 
-the model-based wind data might differ from post-storm analysis of wind radii
-(e.g., Figure S10), and the re-analysis rainfall data seems to sometime oversmooth
-very extreme precipitation (Figure S8). 
-The flood and tornado data
-came from the National Oceanic and Atmospheric Administration's (NOAA's) Storm
-Events database, which, while a widely-used database of events maintained by
-NOAA, is based on reports, and so may be prone to underreporting
-\citep{Ashley2008flood, Curran2000}, especially in less populated areas
-\citep{Witt1998, Ashley2007}, as well as to other reporting errors. While these
-are important limitations, we selected these data sources as among the best
-currently available for measuring each of these hazards at a multi-county,
-multi-year scale.
+correspond with data from other available sources (Figures~S7\,--\,11). While
+generally in agreement with data from other sources, there are a few
+limitations. The wind data are based on modeled, rather than observed, values,
+and while the modeled wind data generally agree well with post-analysis maximum
+wind radii from \ac{HURDAT2}~\citep{landsea2013} (Figure~S7), for a few
+tropical cyclones it did not (e.g., Figure~S10). The rainfall data are from
+re-analysis data, which are generally well-correlated with observed
+ground-based station data but sometimes oversmooth extreme measurements
+(Figure~S8).  The flood and tornado data came from the \ac{NOAA} Storm Events
+database, which, while a widely-used database of events maintained by
+\ac{NOAA}, is based on reports, and so may be prone to
+underreporting~\citep{Ashley2008flood, Curran2000}, especially in less
+populated areas~\citep{Witt1998, Ashley2007}, as well as to other reporting
+errors. While these are important limitations, we selected these data sources
+as among the best currently available for measuring each of these hazards at a
+multi-county, multi-year scale.
 
@@
 Previous research has highlighted the range of impacts that tropical cyclones
-can have in U.S. communities. However, studies have varied widely in how they assess
-exposure to tropical cyclones for exposure, health impact, and other impact
-studies, in some cases using distance to the tropical cyclone's track as a
-surrogate metric of exposure to a tropical cyclone's hazards. Here we found
+can have in \ac{US} communities. However, studies have varied widely in how they
+assess exposure to tropical cyclones for exposure, health impact, and other
+impact studies, in some cases using distance to the tropical cyclone's track as
+a surrogate metric of exposure to a tropical cyclone's hazards. Here we found
 large differences in which counties are exposed to different hazards of

### 2019-01-27 writing/ehp_figures.tex
Rearrange content

@@
 \centering
-\includegraphics[width=16cm]{ivanonly}
+\includegraphics[width=16cm]{ivanonly.pdf}
 \caption{Counties classified as exposed to Hurricane Ivan (2004) under each
-exposure metric (Table 1). The red line shows the track of Hurricane Ivan 
-based on the revised Atlantic hurricane database (HURDAT2 \cite{landsea2013}).
-Similar maps for other large-extent storms are given in Fig. S3.}
+exposure metric (Table~1). The red line shows the track of Hurricane Ivan 
+based on the revised Atlantic hurricane database \ac{HURDAT2}~\citep{landsea2013}.
+Similar maps for other large-extent storms are given in Figure~S3.}
 \label{fig:ivanexposure} 
@@
 \centering 
-\includegraphics[width = 0.6\linewidth]{jaccard_heatmap} 
-\caption{Heatmap of Jaccard index values for
-specific exposure metric pairs within storms. Only storms between 1996 and 2011,
-and for which at least 250 counties were exposed based on at least one metric,
-are included. The color of each cell within the main heatmap indicates the value
-of the Jaccard index (proportion of counties classified as exposed by both
-metrics out of storms classified as exposed by either metric) for a given pair
-of metrics for a given storm. Storms are displayed within clusters that have
-similar patterns in county-level exposure agreement for metric pairs, based on
-hierarchical clustering using the complete link method
-\cite{murtagh2012algorithms} (i.e., storms in the same cluster tend to have
-similar patterns for the pairwise strength of agreement among metrics); columns
-are also ordered based on hierarchical clustering. The colors to the right of
-the main heatmap for each storm indicate the total number of counties classified
-as exposed to the storm by any of the five metrics, providing an estimate of
-storm extent. Maps are available showing the counties identified as exposed
-under each of five metrics for the widest-extent storm in each cluster:
-Hurricane Ivan (2004) (Fig. \ref{fig:ivanexposure}) and Hurricanes Floyd (1999),
-Lee (2011), Cindy (2005), and Katrina (2005) (Fig. S3).} 
+\includegraphics[width = 0.6\linewidth]{jaccard_heatmap.pdf} 
+\caption{Heatmap of Jaccard index values for specific exposure metric pairs
+within storms. Only storms between~1996 and~2011, and for which at least~250
+counties were exposed based on at least one metric, are included. The color of
+each cell within the main heatmap indicates the value of the Jaccard index
+(proportion of counties classified as exposed by both metrics out of storms
+classified as exposed by either metric) for a given pair of metrics for a given
+storm. Storms are displayed within clusters that have similar patterns in
+county-level exposure agreement for metric pairs, based on hierarchical
+clustering using the complete link method~\citep{murtagh2012algorithms} (i.e.,
+storms in the same cluster tend to have similar patterns for the pairwise
+strength of agreement among metrics); columns are also ordered based on
+hierarchical clustering. The colors to the right of the main heatmap for each
+storm indicate the total number of counties classified as exposed to the storm
+by any of the five metrics, providing an estimate of storm extent. Maps are
+available showing the counties identified as exposed under each of five metrics
+for the widest-extent storm in each cluster: Hurricane Ivan (2004)
+(Figure~\ref{fig:ivanexposure}) and Hurricanes Floyd (1999), Lee (2011), Cindy
+(2005), and Katrina (2005) (Figure~S3).} 
 \label{fig:jaccard}
@@
 \centering
-\includegraphics[width=15cm]{averageexposureonly} 
-\caption{Average number of storm exposures per decade in U.S. counties for 
-each exposure metric. The criteria behind each of the five metrics is given 
-in Table \ref{tab:exposuremetrics}. The years used to estimate these averages 
-are based on years of available exposure data (distance and wind: 1988--2015; 
-rain: 1988--2011; flood and tornado: 1996--2015). Similar patterns persist when
-analysis is restricted to years with all exposure data available (1996--2011;
-Fig. S5).} 
+\includegraphics[width=15cm]{averageexposureonly.pdf} 
+\caption{Average number of storm exposures per decade in U.S. counties for each
+exposure metric. The criteria behind each of the five metrics is given in Table
+\ref{tab:exposuremetrics}. The years used to estimate these averages are based
+on years of available exposure data (distance and wind:~1988\,--\,2015; rain:
+1988\,--\,2011; flood and tornado:~1996\,--\,2015). Similar patterns persist when
+analysis is restricted to years with all exposure data available (1996\,--\,2011;
+Figure~S5).} 
 \label{fig:averageexposure} 
 \end{figure}
+
+\clearpage
+
+\begin{figure*}%[tbhp]
+\centering
+\includegraphics[width=15.5cm]{topelecdependexposure}
+\caption{Counties with the highest estimated physical exposure among
+electricity-dependent Medicare beneficiaries exposed to storms per year in U.S.
+counties for each exposure metric. The criteria behind each of the five metrics
+are given in Table~1 [use ref] and details of the physical exposure
+calculation are given in the Methods. The color of each bar indicates the
+number of Medicare beneficiaries in the county reliant on electricity-dependent
+medical and assistive equipment as of July~2017~\citep{empower}. The length of
+each bar shows the average expected number of these electricity-dependent
+Medicare beneficiaries exposed to tropical storms per year based on a given
+exposure metric.}
+\label{fig:topelecdependexposure}
+\end{figure*}
+
+\clearpage
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=0.6\linewidth]{windcomparison}
+\caption{Comparison of estimated wind categories from modeled sustained wind
+speed (as used for the wind exposure metric in the main
+text;~\citep{stormwindmodel}) versus sustained wind speed determined from the
+revised Atlantic hurricane database \ac{HURDAT2}~\citep{landsea2013}. In
+\ac{HURDAT2}, wind radii were available for study storms from~2004\,--\,2015
+and gives estimates of the maximum radii to which storm winds of 34~\si{\knot},
+50~\si{\knot}, and 64~\si{\knot} extended in each four quadrant of the storm at
+each of the 6-\si{\hour} storm track observations. We interpolated this data
+to~15-\si{\minute} increments and assessed a county as exposed to winds at or above
+one of these cutoffs if its population mean center was within~85\% of the
+maximum radius for that wind speed in its quadrant of the storm. For each
+county in each of the available storms, we also calculated the modeled wind
+speed~\citep{stormwindmodel} and then categorized this value into the same
+categories available for the \ac{HURDAT2} wind radii ($<$34~\si{\knot};
+34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}). Each
+point in this graph shows, for a storm, the percent of counties in which the
+wind category ($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot};
+50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) was the same based on both modeled
+winds and the \ac{HURDAT2} wind radii. Only storms with at least one county
+with sustained wind of~$\ge$34 \si{\knot}, based on the \ac{HURDAT2} wind
+radii, are shown. Hurricanes Sandy (2012) and Ike (2008) were he two storms
+with the worst agreement between these two sets of wind data; maps with further
+comparisons between the two data sources for these two storms are given in
+Figure~\ref{fig:windexamples}.}
+\label{fig:windcomparison}
+\end{figure}
+
+\begin{figure}[tbhp!] \centering
+\includegraphics[width=0.9\linewidth]{raincomparison} \caption{Comparison of
+county-level rain exposure based on the reanalysis data used for exposure
+assessment in this study \ac{NLDAS-2}~\citep{rui2013nldas,
+alhamdan2014environmental} and ground-based weather
+stations~\citep{menne2012overview, rnoaa, countyweather}. Each small graph
+shows a sample county, and each point on the graph represents a tropical storm
+that came within~500 \si{\kilo\metre} of the population mean center of the county.
+The x-axis shows the estimated cumulative rainfall (in millimeters) from two
+days before to one day after the storm's closest approach to the county based
+on the county-aggregated \ac{NLDAS-2} recipitation data used to measure rain
+exposure in our main analysis. The ground-based observations (y-axis) are based
+on collecting precipitation data from available weather stations in the county,
+1988\,--\,2011, summed over the same period from two days before to one day
+after the storm's closest approach for each available station. These cumulative
+station-based precipitation totals were then averaged across all available
+stations in the county to create a county-wide estimate of cumulative
+storm-related precipitation for each storm in the county. Horizontal and
+vertical lines show the threshold used to classify a storm as exposed based on
+cumulative rainfall; storms in the lower left and upper right quadrants are
+classified the same regardless of whether reanalysis or station-based
+precipitation data is considered, while storms in the upper left and lower
+right quadrants would be classified differently. The number of storms within
+each county and the average number of stations reporting rainfall during the
+county's storms are given above each plot. Note that x- and y-axis ranges
+differ by county. An estimate of the rank correlation
+(Spearman's~$\rho$~\citep{spearman1904proof}) between measurements from the two
+rain data sources is also shown on each graph.} \label{fig:raincomparison}
+\end{figure}
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=0.8\linewidth]{floodcomparison}
+\caption{Comparison of for a sample of study counties of flood status, based on
+NOAA Storm Events listings~\citep{stormevents, noaastormevents}, with total
+streamflows at county streamflow gages during tropical storms~\citep{usgsgages,
+countyfloods, dataRetrieval}. Each panel represents a separate county. For each
+county, we first identified all streamflow gages in the county that had
+complete data for Jan.~1~1996\,--\,Dec.~31,~2015. The number of streamflow
+gages meeting this criterion are given in parentheses beside the county's name
+in the panel title. This analysis covered tropical storm that came
+within~500~\si{\kilo\metre} of each sample county between~1996 and~2015. In the
+graphs, each point represents a single tropical storm, and the point's position
+along the x-axis shows the highest daily total streamflow (cubic feet /
+second), summed across all identified streamgages in the county, for a five-day
+window centered on the day of the storm's closest approach to the county. The
+y-axis separates storms for which a flood event was reported in NOAA's Storm
+Events database for the county with a start date within the five-day window of
+the storm's closest approach to the county~\citep{stormevents,
+noaastormevents}. Boxplots are added to show the distribution of the total
+streamflow measurements across storms in the county for storms grouped by
+whether or not they were linked with a flood event based on the NOAA Storm
+Events data. The color of each point gives the percent of streamflow gages in
+the county with a daily streamflow that exceeded a threshold of flooding (the
+streamgage's median value for annual peak flow~\citep{countyfloods}) on any day
+during the five-day window.  Note that the x-axis scales differ by county,
+depending on the number of streamflow gages and typical flow rates for each
+gage, and are on a log-10 scale.}
+\label{fig:floodcomparison}
+\end{figure}

### 2019-01-27 writing/ehp_introduction.tex
Rearrange content

@@
 severe wind; for example, one study found that most of the direct
-hurricane-related deaths in the \ac{US} between 1970 and 1999 occurred in cases
+hurricane-related deaths in the \ac{US} between~1970 and~1999 occurred in cases
 when wind was below hurricane strength, including for Tropical Storms Charley
-in 1998 and Alberto in 1994~\citep{rappaport2000}.  Tropical cyclones can
+in~1998 and Alberto in~1994~\citep{rappaport2000}.  Tropical cyclones can
 produce excessive rain, especially in certain topographies (e.g., near
@@
 although the two risks are not perfectly correlated~\citep{chen2015}. In the
-\ac{US}, over half of hurricane-related direct deaths from 1970 to 1999 from
+\ac{US}, over half of hurricane-related direct deaths from~1970 to~1999 from
 Atlantic basin storms were a result of freshwater flooding~\citep{rappaport2000}. 
@@
 tropical cyclone hazards may disagree with distance-based classifications.
-Tropical cyclones vary dramatically in size: \ac{US} tropical cyclones have been
-observed with radii to maximum winds as small as 20 kilometers and as large as
-200 kilometers~\citep{mallin2006, quiring2011variations}. While a number of
-tropical cyclone hazards are strongly associated with distance from the
-tropical cyclone's center (e.g., wind and, at the coast, storm surge and 
+Tropical cyclones vary dramatically in size: \ac{US} tropical cyclones have
+been observed with radii to maximum winds as small as~20~\si{\kilo\metre} and as
+large as~200~\si{\kilo\metre}~\citep{mallin2006, quiring2011variations}. While a
+number of tropical cyclone hazards are strongly associated with distance from
+the tropical cyclone's center (e.g., wind and, at the coast, storm surge and
 waves~\citep{rappaport2000, kruk2010}), other hazards like dangerous rain,
 floods, and tornadoes can occur well away from the tropical cyclone's central
-track~\citep{rappaport2000, atallah2007, moore2012}.  For example, fatal tropical
-cyclone tornadoes, which were linked to over 300 deaths in the \ac{US}  between
-1995 and 2009, most often occur 200\,--\,500 kilometers from the tropical cyclone's
-center~\citep{moore2012}.
+track~\citep{rappaport2000, atallah2007, moore2012}.  For example, fatal
+tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
+between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
+tropical cyclone's center~\citep{moore2012}.
 
@@
 buffers tend to use an equal buffer distance on each side of the tropical
-cyclone track (e.g.,~\citep{czajkowski2011, grabich2015, grabich2016,
+cyclone track (e.g.,~\cite{czajkowski2011, grabich2015, grabich2016,
 zandbergen2009, tansel2010}), but the forces of a tropical cyclone tend to be
@@
 motion~\citep{halverson2015}, and the fatal tornadoes associated with \ac{US} tropical
-cyclones between 1995 and 2009 occurred almost exclusively to the right of the
+cyclones between~1995 and~2009 occurred almost exclusively to the right of the
 tropical cyclone's track, mostly in the right front quadrant of the tropical
@@
 To investigate this, we measured county-level exposure to tropical cyclones in
-all eastern \ac{US} counties using five metrics (
-Table~\ref{tab:exposuremetrics}): (1) distance to tropical cyclone track; (2) maximum
-sustained wind speed; (3) cumulative rainfall; (4) flood events; and (5)
-tornado events.  We assessed exposure at the county level since data on many
-potential impacts are available at county-level aggregations (e.g., direct
-hurricane-related deaths~\citep{czajkowski2011}; birth outcomes~\citep{grabich2015,
-grabich2016}; autism prevalence~\citep{kinney2008}) and
-since decisions and policies to prepare for, and respond to, tropical cyclones
-are often undertaken at the county level~\citep{zandbergen2009, rappaport2000}.
-We explored patterns in these exposure assessments and measured how well
-exposure classification agreed across these five metrics in terms of
-classifying specific counties as exposed to a tropical cyclone.  Finally, to
-make this hazard-specific tropical cyclone exposure data available to other
-researchers, we published open source software for county-based hurricane
-exposure assessment in \ac{US} counties~\citep{hurricaneexposure}.
+all eastern \ac{US} counties using five metrics
+(Table~\ref{tab:exposuremetrics}): (1)~distance to tropical cyclone track;
+(2)~maximum sustained wind speed; (3)~cumulative rainfall; (4)~flood events;
+and (5)~tornado events.  We assessed exposure at the county level since data on
+many potential impacts are available at county-level aggregations (e.g., direct
+hurricane-related deaths~\citep{czajkowski2011}; birth
+outcomes~\citep{grabich2015, grabich2016}; autism
+prevalence~\citep{kinney2008}) and since decisions and policies to prepare for,
+and respond to, tropical cyclones are often undertaken at the county
+level~\citep{zandbergen2009, rappaport2000}.  We explored patterns in these
+exposure assessments and measured how well exposure classification agreed
+across these five metrics in terms of classifying specific counties as exposed
+to a tropical cyclone.  Finally, to make this hazard-specific tropical cyclone
+exposure data available to other researchers, we published open source software
+for county-based hurricane exposure assessment in \ac{US}
+counties~\citep{hurricaneexposure}.

### 2019-01-27 writing/ehp_methods.tex
Rearrange content

@@
 
-All counties in the eastern half of the \ac{US} (e.g., Figure~\ref{fig:ivanexposure}), 
-as of the 2010 \ac{US} Decennial Census, were included in
-the study. The study covered all tracked storms in the \ac{HURDAT2} \citep{landsea2013} 
-that came within 250
-kilometers of at least one study county between 1988 and 2015 (Figure~S2).
-HURDAT2 is a post-storm assessment conducted by the \ac{US} National Hurricane
-Center (NHC) and incorporates data from a variety of sources, including
-satellite data and, when available, aircraft reconnaissance 
-data~\citep{landsea2013, jarvinen1988}. These data typically give measurements of
-the tropical cyclone center's location at 6-hour intervals at synoptic times
-(i.e., 6:00 am, 12:00 pm, 6:00 pm, and 12:00 am UTC); some landfalling tropical
-cyclones have an additional observation at the time of landfall~\citep{landsea2013}.
+All counties in the eastern half of the \ac{US}
+(Figure~\ref{fig:ivanexposure}), as of the~2010 \ac{US} Decennial Census, were
+included in the study. The study covered all tracked storms in the \ac{HURDAT2}
+\citep{landsea2013} that came within~250 \si{\kilo\metre} of at least one study
+county between~1988 and~2015 (Figure~S2).  \ac{HURDAT2} is a post-storm
+assessment conducted by the \ac{US} \ac{NHC} and incorporates data from a
+variety of sources, including satellite data and, when available, aircraft
+reconnaissance data~\citep{landsea2013, jarvinen1988}. These data typically
+give measurements of the tropical cyclone center's location at~6-\si{\hour} intervals
+at synoptic times (i.e., 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC});
+some landfalling tropical cyclones have an additional observation at the time
+of landfall~\citep{landsea2013}.
 
@@
 
-For each storm, the distance of closest approach was measured for every study county. 
-Storm center track location data from \ac{HURDAT2} was 
-interpolated using natural cubic splines to 15-minute intervals~\cite{hurricaneexposure}. 
-The distance between each county's population mean center, as assessed by the 2010 Decennial
-Census~\cite{countycenters}, and each 15-minute interpolated storm center location was
-measured using Great Circle (WGS84 ellipsoid) distance~\cite{bivand2013applied}. The 
-distance-based exposure metric was based on the minimum value of this storm 
-center-to-county distance throughout the period the storm was tracked. The time of 
-closest approach for each county-storm pairing was also recorded for use assessing 
-rain-, flood-, and tornado-based metrics. To allow matching with data based on local 
-time, times were converted from Coordinated Universal Time (UTC) to local 
-time~\cite{countytimezones}.
+For each storm, the distance of closest approach was measured for every study
+county.  Storm center track location data from \ac{HURDAT2} were interpolated
+using natural cubic splines to~15-\si{\minute} intervals~\citep{hurricaneexposure}.
+The distance between each county's population mean center, as assessed by the
+2010 Decennial Census~\citep{countycenters}, and each~15-\si{\minute} interpolated
+storm center location was measured using Great Circle (WGS84 ellipsoid)
+distance~\citep{bivand2013applied}. The distance-based exposure metric was
+based on the minimum value of this storm center-to-county distance throughout
+the period the storm was tracked. The time of closest approach for each
+county-storm pairing was also recorded for use assessing rain-, flood-, and
+tornado-based metrics. To allow matching with data based on local time, times
+were converted from \ac{UTC} to local time~\citep{countytimezones}.
 
@@
 closest approach to a county. Daily county-level rainfall was determined by
-aggregating hourly, 1/8 degree gridded reanalysis data from the North American
-Land Data Assimilation System Phase 2 (NLDAS-2) precipitation data 
-files~\citep{rui2013nldas}. The NLDAS-2 data integrate satellite-based and land-based
-monitoring, applying a land-surface model to create a reanalysis dataset that
-is spatially and temporally complete across the continental 
-\ac{US}~\citep{rui2013nldas, alhamdan2014environmental}. To aggregate to county-level,
-hourly data at each grid point were summed to create a daily rainfall total,
-and these grid point rainfall totals were then averaged across all grid points
-within a county's 1990 \ac{US} Census boundaries to create a county-level daily
-total~\citep{alhamdan2014environmental, cdcwonder}. This county-level
-precipitation data are publicly available through the \ac{US} Centers for Disease
-Control's Wide-ranging Online Data for Epidemiological Research (WONDER)
-database~\citep{cdcwonder, alhamdan2014environmental}. The rainfall-based
-exposure metric was only available for tropical cyclones through 2011, the
-period for which the county-aggregated reanalysis rainfall data were 
-available~\citep{cdcwonder, alhamdan2014environmental}. Validation analysis of this and
-other hazard-based exposure data is given in the Supplemental Material.
+aggregating hourly~1/8\si{\degree} gridded reanalysis data from the \ac{NLDAS-2}
+precipitation data files~\citep{rui2013nldas}. The \ac{NLDAS-2} data integrate
+satellite-based and land-based monitoring, applying a land-surface model to
+create a reanalysis dataset that is spatially and temporally complete across
+the continental \ac{US}~\citep{rui2013nldas, alhamdan2014environmental}. To
+aggregate to county-level, hourly data at each grid point were summed to create
+a daily rainfall total, and these grid point rainfall totals were then averaged
+across all grid points within a county's~1990 \ac{US} Census boundaries to
+create a county-level daily total~\citep{alhamdan2014environmental, cdcwonder}.
+This county-level precipitation data are publicly available through the \ac{US}
+\ac{CDC}'s \ac{WONDER} database~\citep{cdcwonder, alhamdan2014environmental}.
+The rainfall-based exposure metric was only available for tropical cyclones
+through~2011, the period for which the county-aggregated reanalysis rainfall
+data were available~\citep{cdcwonder, alhamdan2014environmental}.
+
+To assess the rainfall estimates from the reanalysis data used in analysis in
+the the main text, we compared cumulative rainfall estimates based on this
+reanalysis rainfall data to rainfall observations from ground-based weather
+stations in nine sample study counties, selected as counties with data
+available from multiple stations. In these nine counties, observed rainfall was
+collected for all storms,~1988\,--\,2011, for the same period of two days
+before to one day after each storm's closest approach, using all stations with
+available daily data in the Global Historical Climatology Network throughout
+1988\,--\,2011~\citep{menne2012overview, rnoaa, countyweather}. Agreement
+between measured storm-related rainfall between these two data sources was
+assessed by Spearman's rank correlation.
 
@@
 modeled maximum sustained wind speeds at each county's population mean
-center~\citep{countycenters}. HURDAT2 Best Track data were interpolated from 6-hourly
-reported values to 15-minute increments using natural cubic 
-splines~\citep{stormwindmodel}, and then sustained wind speed was modeled for each
-county center at each 15-minute increment using a double exponential wind speed
-model~\citep{willoughby2006parametric} to estimate maximum ground-level
+center~\citep{countycenters}. \ac{HURDAT2} Best Track data were interpolated
+from~6-\si{\hour} reported values to~15-\si{\minute} increments using natural cubic
+splines~\citep{stormwindmodel}, and then sustained wind speed was modeled for
+each county center at each~15-\si{\minute} increment using a double exponential wind
+speed model~\citep{willoughby2006parametric} to estimate maximum ground-level
 sustained wind speed at each county center~\citep{stormwindmodel}. Asymmetry in
@@
 this modeled wind speed was determined for each county as the maximum windspeed
-across the 15-minute incremented modeled values throughout the tropical
+across the~15-\si{\minute} incremented modeled values throughout the tropical
 cyclone.
+
+To validate the metric of county-level wind speed used in the main text, we
+compared this wind metric to estimated wind radii from \ac{HURDAT2}, available
+for storms in our study from~2004~\citep{landsea2013}. \ac{HURDAT2} gives
+estimates of the maximum radii to winds of~64,~50, and~34 \si{\knot} in each four
+quadrants of a storm (i.e., distance from the storm's center to the furthest
+point in a quadrant with winds at or above that speed) at each~6-\si{\hour} storm
+observation.
 
@@
 (``Flood", ``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and
-that occurred in a county within 500 kilometers of the tropical cyclone's track
-and within a five-day window centered on the date of the tropical cyclone's
-closest approach to the county~\citep{hurricaneexposuredata}. ``Flood", ``Flash
-Flood" and ``Tornado" events in this database were reported by county Federal
-Information Processing Standard (FIPS) code. ``Coastal Flood" events were
-reported by forecast zone; for these, the event was matched to the appropriate
-county if possible~\citep{noaastormevents}. While this database has recorded
-storm data since 1950, its coverage changed substantially in 1996 to cover a
-wider variety of storm events~\citep{stormevents}. We therefore only considered
-these metrics of hurricane exposure for tropical cyclones in 1996 and later.
+that occurred in a county within~500 \si{\kilo\metre} of the tropical cyclone's track
+and with a start date within a five-day window centered on the date of the
+tropical cyclone's closest approach to the
+county~\citep{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado"
+events in this database were reported by county \ac{FIPS} code. ``Coastal
+Flood" events were reported by forecast zone; for these, the event was matched
+to the appropriate county if possible~\citep{noaastormevents}. While this
+database has recorded storm data since~1950, its coverage changed substantially
+in~1996 to cover a wider variety of storm events~\citep{stormevents}. We
+therefore only considered these metrics of hurricane exposure for tropical
+cyclones in~1996 and later.
 
-\subsection*{Assessment of tropical cyclone exposure data}
-
-To assess the rainfall estimates from the reanalysis data used in analysis in the the 
-main text, we compared cumulative rainfall estimates based on this reanalysis rainfall 
-data to rainfall observations from ground-based weather stations in nine sample study 
-counties, selected as counties with data available from multiple stations. In these nine
-counties, observed rainfall was collected for all storms, 1988\,--\,2011, for the same period 
-of two days before to one day after each storm's closest approach, using all stations with
-available daily data in the Global Historical Climatology Network throughout 
-1988\,--\,2011~\citep{menne2012overview, rnoaa, countyweather}. Agreement between 
-measured storm-related rainfall between these two data sources was assessed by 
-Spearman's rank correlation.
-
-To validate the metric of county-level wind speed used in the main text, we compared this 
-wind metric to estimated wind radii from the revised Atlantic hurricane database (HURDAT2),
-available for storms in our study from 2004~\citep{landsea2013}. HURDAT2 gives estimates of
-the maximum radii to winds of 64, 50, and 34 knots in each four quadrants of a storm (i.e.,
-distance from the storm's center to the furthest point in a quadrant with winds at or above
-that speed) at each 6-hour storm observation.
-
-As validation for the flood data, we compared streamflow data from United States (\ac{US})
-Geological Survey gages~\citep{usgsgages, countyfloods, dataRetrieval} in nine counties with
-tropical cyclone flood classifications based on the NOAA Storm Events Database 
-(Figure~\ref{fig:floodcomparison}).
+As validation for the flood data, we compared streamflow data from \ac{US}
+Geological Survey gages~\citep{usgsgages, countyfloods, dataRetrieval} in nine
+counties with tropical cyclone flood classifications based on the NOAA Storm
+Events Database (Figure~\ref{fig:floodcomparison}).
 
@@
 cyclones. Depending on available exposure data, this assessment included some
-or all of the period from 1988 to 2015, a period with 136 tropical cyclones
-that made \ac{US} landfall or passed within 250 kilometers (155 miles) of at least
-one \ac{US} county (Figure~S2).
+or all of the period from~1988 to~2015, a period with~136 tropical cyclones
+that made \ac{US} landfall or passed within~250 \si{\kilo\metre} (155 mi) of at
+least one \ac{US} county (Figure~S2).
 
@@
 measurement, we calculated the within-storm Jaccard
-index~\citep{jaccard1901distribution, jaccard1908nouvelles} between every pair-wise
-combination of hazard metrics.  The Jaccard index ($J_s$) measures similarity
-between two metrics ($X_{1,s}$ and $X_{2,s}$) for tropical cyclone $s$ as the
-proportion of counties in which both of the metrics classify the county as
-exposed out of all counties classified as exposed by at least one of the
-metrics:
+index~\citep{jaccard1901distribution, jaccard1908nouvelles} between every
+pair-wise combination of hazard metrics.  The Jaccard index~($J_s$) measures
+similarity between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical
+cyclone~$s$ as the proportion of counties in which both of the metrics classify
+the county as exposed out of all counties classified as exposed by at least one
+of the metrics:
 
@@
 
-\noindent This metric can range from 0, in the case of no overlap between the counties
-classified as exposed based on the two metrics, to 1, in the case that the
-two metrics classify exactly the same counties as exposed to the tropical cyclone.
+\noindent This metric can range from~0, in the case of no overlap between the
+counties classified as exposed based on the two metrics, to~1, in the case that
+the two metrics classify exactly the same counties as exposed to the tropical
+cyclone.
+
+\subsection*{Case study: Physical exposure of electricity-dependent Medicare
+beneficiaries to tropical cyclones}
+
+The outcomes of disasters depend both on the geophysical forces of the disaster
+as well as on the vulnerability of the society living in the geographical areas
+affected by those forces.~\citep{chakraborty2005population,
+anderson2003community, cutter1996vulnerability} As a case study of how
+differences in tropical cyclone exposure assessments across different hazard
+metrics influence estimates of physical exposures of susceptible populations to
+tropical cyclones, we investigated how the use of different metrics influenced
+estimates of which eastern \ac{US} counties have the highest expected average
+exposures of electricity-dependent Medicare beneficiaries to tropical cyclones.
+This subpopulation was selected for this analysis since it may be particularly
+susceptible to health impacts from tropical cyclone exposure, especially
+through the pathway of storm-associated power outages and evacuations. We
+collected data on the number of electricity-dependent Medicare beneficiaries in
+each study county from the U.S. Department of Health \& Human Service's emPower
+Map~2.0~\citep{empower}. We then calculated the physical exposure of the
+electricity-dependent Medicare population in each county, based on tropical
+cyclone assessments using each hazard metric,
+following~\citep{peduzzi2009assessing}:
+
+\begin{equation}
+E_c = F_c * P_c
+\end{equation}
+
+\noindent where~$E_c$ is the average yearly physical exposure among
+electricity-dependent Medicare beneficiaries in county~$c$ to tropical cyclone
+exposures based on a given metric,~$F_c$ is the estimated yearly expected
+frequency of tropical cyclone exposures in county~$c$ based on that metric,
+and~$P_c$ is the size of the electricity-dependent Medicare population in
+county~$c$, as of July~2017. When combined with estimates of vulnerability of a
+population to a natural hazard, such measurements of physical exposure can be
+used to calculate risk of human losses from the
+hazard~\citep{peduzzi2009assessing}.

### 2019-01-27 writing/ehp_results.tex
Rearrange content

@@
 \section*{Results}
-
-\subsection*{Assessment of tropical cyclone exposure data}
-
-Within
-the nine sample counties, storm-related rainfall from the reanalysis precipitation dataset 
-was well-correlated with station-based rainfall measurements, although the reanalysis data 
-did tend to somewhat underestimate rainfall totals in storms with extremely high rainfall 
-(Figure~\ref{fig:raincomparison}). A distance constraint (county population mean center within
-500 kilometers of the storm's closest approach) was necessary for assessing storm exposure
-based on this metric to ensure that rains unrelated and far from the storm track were not
-misattributed to storm exposures. We found that this distance constraint was typically
-adequately large to capture storm-related rain, although in rare examples of exceptionally
-large storms (e.g., Hurricane Ike [2008]) or storms for which tracking was stopped at
-extratropical transition (e.g., Tropical Storm Lee [2011]), some storm-related rains were
-missed because of this distance constraint (Figure~\ref{fig:rainexamples}).
-
-The modeled wind speeds generally agreed well
-with the HURDAT2 wind radii (Figure~\ref{fig:windcomparison}), with disagreement mostly 
-limited to large storms that were characterized by high winds well inland (e.g., Hurricanes
-Ike [2008] and Sandy [2012]; Figure~\ref{fig:windexamples}).
-
-Stream gage data generally indicated higher discharge in these
-counties on dates identified as flood events, particularly in counties with more reporting
-stream gages (e.g., Harris County, TX, with 10 reporting stream gages; 
-Figure~\ref{fig:floodcomparison}), although these two flooding data sources were 
-not completely
-consistent.
 
@@
 
-We first calculated summary statistics for the exposure assessments (Table
-\ref{tab:exposuresummaries}). There was wide variation in the average number of
-county exposures per year under the five different tropical cyclone metrics
-considered. For the tornado metric, there were approximately 50 county
+We first calculated summary statistics for the exposure assessments
+(Table~\ref{tab:exposuresummaries}). There was wide variation in the average
+number of county exposures per year under the five different tropical cyclone
+metrics considered. For the tornado metric, there were approximately~50 county
 exposures identified per year, on average, within our study.  For flood and
-wind metrics, there were substantially more county exposures identified (over
-200 per year on average), and even more when assessing exposure using the
-distance and rain metrics (over 400 per year on average).
+wind metrics, there were substantially more county exposures identified ($>$200
+per year on average), and even more when assessing exposure using the distance
+and rain metrics ($>$400 per year on average).
 
@@
 exposures per tropical cyclone was lowest when assessing exposure using the
-tornado metric (median: 7 county exposures per tropical cyclone), higher under
-the flood and wind metrics (median: 26 and 34, respectively), and even higher for the metrics of rain (median: 68)
-and distance (median: 84).  For every metric except the tornado-based metric,
-we identified at least one tropical cyclone with over 300 counties exposed.
-However, the largest-extent tropical cyclone varied by metric (e.g., Beryl in
-1994 exposed the most counties based on distance, Frances in 2004 based on
-rain, and Ike in 2008 based on wind).
+tornado metric (median:~7 county exposures per tropical cyclone), higher under
+the flood and wind metrics (median:~26 and~34, respectively), and even higher
+for the metrics of rain (median:~68) and distance (median:~84).  For every
+metric except the tornado-based metric, we identified at least one tropical
+cyclone with over~300 counties exposed.  However, the largest-extent tropical
+cyclone varied by metric (e.g., Beryl in~1994 exposed the most counties based
+on distance, Frances in~2004 based on rain, and Ike in~2008 based on wind).
 
@@
 tropical cyclone exposure assessment is to the choice of metric, using as an
-example Hurricane Ivan (2004), a record-breaking tropical cyclone for its
+example Hurricane Ivan in 2004, which was record-breaking for its
 duration as a major hurricane~\citep{franklin2006atlantic}. When county-level
@@
 
-We drew similar conclusions when we investigated all 46 tropical cyclones
-between 1996 and 2011 (the period for which all five metrics were available)
-for which 250 or more counties were exposed based on at least one metric
+We drew similar conclusions when we investigated all~46 tropical cyclones
+between~1996 and~2011 (the period for which all five metrics were available)
+for which~250 or more counties were exposed based on at least one metric
 (Figure~\ref{fig:jaccard}). In this figure, each row provides results for a
@@
 hazard-based metrics. For almost all tropical cyclones, Jaccard index values
-between distance and any of the four hazard-based metrics were 0.6 or lower
+between distance and any of the four hazard-based metrics were~0.6 or lower
 (i.e., out of the counties assessed as exposed to the tropical cyclone by at
-least one of the two metrics considered, 60\% or fewer of the counties had the
-same exposure assessment under the two metrics). For many tropical cyclone, the
-Jaccard index values were 0.2 or lower, suggesting very poor agreement in
-exposure assessment between the metrics.  Among the four hazard-based metrics,
-the tornado-based metric showed universally poor agreement with other metrics
-in county-level classification across all tropical cyclones considered.  For
-other pairs of metrics, there were also generally large differences in which
-counties were determined to be exposed to the tropical cyclone, with the
-Jaccard index values below 0.4 for most metric pairs in most tropical cyclones.
+least one of the two metrics considered,~60\si{\percent} or fewer of the
+counties had the same exposure assessment under the two metrics). For many
+tropical cyclones, the Jaccard index values were~0.2 or lower, suggesting very
+poor agreement in exposure assessment between the metrics.  Among the four
+hazard-based metrics, the tornado-based metric showed universally poor
+agreement with other metrics in county-level classification across all tropical
+cyclones considered.  For other pairs of metrics, there were also generally
+large differences in which counties were determined to be exposed to the
+tropical cyclone, with the Jaccard index values below~0.4 for most metric pairs
+in most tropical cyclones.
 
@@
 counties were determined to be exposed to the tropical cyclone for two or more
-of the metrics considered.  For Floyd in 1999 (Figure~S3) and Irene in 2011,
+of the metrics considered.  For Floyd in~1999 (Figure~S3) and Irene in~2011,
 for example, county-level classification agreed moderately to well (Jaccard
-index of approximately 0.5\,--\,0.8) for all pairs of exposure metrics except those
-including the tornado-based metric. These tropical cyclones both made their
-first \ac{US}  landfall in North Carolina at minor hurricane windspeeds (Category
-2 and 1, respectively) and then skimmed the eastern coast of the \ac{US} north
-through New England, bringing large rainfall to much of the eastern coast from
-North Carolina north and causing extensive inland flooding in North Carolina
-(Floyd) and New England (Irene)~\citep{avila2013atlantic,
-lawrence2000atlantic}. For another set of tropical cyclones (e.g., Lee in 2011
-[Figure~S3], Ernesto in 2006, and Bertha in 1996), there was moderate to good
+index of approximately~0.5\,--\,0.8) for all pairs of exposure metrics except
+those including the tornado-based metric. These tropical cyclones both made
+their first \ac{US}  landfall in North Carolina at minor hurricane windspeeds
+(Category~2 and~1, respectively) and then skimmed the eastern coast of the
+\ac{US} north through New England, bringing large rainfall to much of the
+eastern coast from North Carolina north and causing extensive inland flooding
+in North Carolina (Floyd) and New England (Irene)~\citep{avila2013atlantic,
+lawrence2000atlantic}. For another set of tropical cyclones (e.g., Lee in~2011
+[Figure~S3], Ernesto in~2006, and Bertha in~1996), there was moderate to good
 agreement for all pairwise combinations of distance, rain, and wind, but poor
 agreement for all other combinations of metrics. Most of these tropical
-cyclones either skimmed along the eastern coast north of North Carolina (Bertha
-in 1996, Charley in 2004, Gaston in 2004, Barry in 2007, and Hanna in 2008), in
-a pattern similar to Hurricanes Floyd in 1999 and Irene in 2011, or made
+cyclones either followed the eastern coastline north of North Carolina (Bertha
+in~1996, Charley in~2004, Gaston in~2004, Barry in~2007, and Hanna in~2008), in
+a pattern similar to Hurricanes Floyd in~1999 and Irene in~2011, or made
 landfall in or near the Florida panhandle and cut across to exit into the
-Atlantic around North Carolina (Josephine in 1996, Earl in 1998, Alberto in
-2006). However, these two sets of tropical cyclones represented the minority of
-all tropical cyclones considered; for most tropical cyclones, there was low
+Atlantic around North Carolina (Josephine in~1996, Earl in~1998, Alberto
+in~2006). However, these two sets of tropical cyclones represented the minority
+of all tropical cyclones considered; for most tropical cyclones, there was low
 overlap in the counties determined to be exposed to the tropical cyclones
@@
 
-Strong geographical patterns were clear across tropical cyclone exposure
-assessments based on these different metrics when the average number of
-exposures per decade in each county were calculated 
-(Figure~\ref{fig:averageexposure}). Wind-based exposure had a strong coastal pattern,
-with almost all exposures in counties within about 200 kilometers (124 miles)
-of the coastline, mostly in eastern North and South Carolina, southern Florida,
-and the area near the outlet of the Mississippi River. While exposures by rain-
-and distance-based metrics were also more common in coastal areas compared to
+When the average number of exposures per decade in each county were calculated,
+strong geographical patterns were clear across tropical cyclone exposure
+assessments based on these different metrics
+(Figure~\ref{fig:averageexposure}). Wind-based exposure had a strong coastal
+pattern, with almost all exposures in counties within about~200
+\si{\kilo\metre} (124~mi) of the coastline. While exposures by rain- and
+distance-based metrics were also more common in coastal areas compared to
 inland areas, inland exposures were much more common based on these two metrics
@@
 Carolina coast, as well as further north in New Jersey, southeastern
-Pennsylvania, and southeastern New York. Notably, some areas of the \ac{US} that
-experienced regular exposures to tropical cyclone-related excessive rainfall
-and floods were not identified as areas of regular risk based on the more
-commonly used tropical cyclone exposure assessment metrics of wind or distance
-(Figure~\ref{fig:averageexposure}).
+Pennsylvania, and southeastern New York. Notably, some areas of the \ac{US}
+that experienced regular exposures to tropical cyclone-related excessive
+rainfall and floods were not identified as areas of regular risk based on the
+more commonly used tropical cyclone exposure assessment metrics of wind or
+distance (Figure~\ref{fig:averageexposure}).
 
-We included a case study in the Supporting Information showing how these
-differences in exposure assessment across the five metrics considered could
-strongly influence health-related studies based on physical exposure to
-tropical cyclones. For this example, we compared estimates of which \ac{US}
-counties have the largest expected physical exposure to tropical cyclones among
-a susceptible subpopulation, Medicare beneficiaries who are reliant on
-electricity for medical equipment (Supplemental Text, Figure~S6)~\citep{empower}.
+We conducted a case study to investigate how these differences in exposure
+assessment across the five metrics considered could strongly influence
+health-related studies based on physical exposure to tropical cyclones. For
+this example, we compared estimates of which \ac{US} counties have the largest
+expected physical exposure to tropical cyclones among a susceptible
+subpopulation, Medicare beneficiaries who are reliant on electricity for
+medical equipment (Figure~S6)~\citep{empower}.  We found
+that a few counties (Miami-Dade County, FL; Harris County, TX) were ranked in
+the top three counties of physical exposure among electricity-dependent
+Medicare beneficiaries for almost all exposure metrics
+(Figure~\ref{fig:topelecdependexposure}). However, there were key differences
+when using different exposure metrics. For example, Florida counties dominated
+the lists of top exposed counties based on wind- and tornado-based metrics,
+while over half of the counties for the flood-based metric were in Mid-Atlantic
+states (PA, NY).
 
@@
 We investigated how well these data correspond with data from other available
-sources (Figures~S7\,--\,11).  While generally in agreement with data from other
+sources (Figures~S7\,--\,11). Within the nine sample counties, storm-related
+rainfall from the reanalysis precipitation dataset was well-correlated with
+station-based rainfall measurements, although the reanalysis data did tend to
+somewhat underestimate rainfall totals in storms with extremely high rainfall
+(Figure~\ref{fig:raincomparison}). A distance constraint (county population
+mean center within~500 \si{\kilo\metre} of the storm's closest approach) was
+necessary for assessing storm exposure based on this metric to ensure that
+rains unrelated and far from the storm track were not misattributed to storm
+exposures. We found that this distance constraint was typically adequately
+large to capture storm-related rain, although in rare examples of exceptionally
+large storms (e.g., Hurricane Ike in 2008) or storms for which tracking was
+stopped at extratropical transition (e.g., Tropical Storm Lee in 2011), some
+storm-related rains were missed because of this distance constraint
+(Figure~\ref{fig:rainexamples}).  The modeled wind speeds generally agreed well
+with the \ac{HURDAT2} wind radii (Figure~\ref{fig:windcomparison}), with
+disagreement mostly limited to large storms that were characterized by high
+winds well inland (e.g., Hurricanes Ike in 2008 and Sandy in 2012;
+Figure~\ref{fig:windexamples}).  Stream gage data generally indicated higher
+discharge in these counties on dates identified as flood events, particularly
+in counties with more reporting stream gages (e.g., Harris County, TX, with~10
+reporting stream gages; Figure~\ref{fig:floodcomparison}), although these two
+flooding data sources were not completely consistent.
+
+While generally in agreement with data from other
 sources, there are a few limitations. The wind data are based on modeled,
@@
 well with post-analysis maximum wind radii from the revised Atlantic hurricane
-database (\ac{HURDAT2}~\citep{landsea2013}) (Figure~S7), for a few tropical cyclones
+database \ac{HURDAT2}~\citep{landsea2013} (Figure~S7), for a few tropical cyclones
 it did not (e.g., Figure~S10). The rainfall data are from re-analysis data,
 which are generally well-correlated with observed ground-based station data but
-sometimes oversmooth extreme measurements (Figure~S8). More details on the assessment
-of this data are provided in the Supplemental Material.
+sometimes oversmooth extreme measurements (Figure~S8).

### 2019-01-27 writing/ehp_tables.tex
Rearrange content

@@
 Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
+Distance & 1988--2015 & County population mean center within 100 \si{\kilo\metre} of storm track \\ 
+Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
 Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Tornado & 1996--2015 & Tornado event listed NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
+Flood & 1996--2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
+Tornado & 1996--2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track\\
 \bottomrule 

### 2019-01-28 tables/metric_definitions.tex
Work some on formatting

@@
 Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Tornado & 1996--2015 & Tornado event listed in the NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
+Distance & 1988\,--\,2015 & County population mean center within 100 kilometers of storm track \\ 
+Rain & 1988\,--\,2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
+Wind & 1988\,--\,2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
+Flood & 1996\,--\,2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
+Tornado & 1996\,--\,2015 & Tornado event listed in the NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
 \bottomrule 

### 2019-01-28 tables/stormseverity.tex
Work some on formatting

@@
 \caption{Summary statistics for exposure metrics considered to assess county-level exposure to
-         tropical cyclones. A county exposure represents the exposure of a specific eastern U.S.  
+         tropical cyclones. A county exposure represents the exposure of a specific eastern US
          county to a specific tropical cyclone. A county can be exposed to multiple tropical cyclones within
@@
          The median and interquartile range of county exposures per tropical cyclone is based on
-         the tropical cyclones for which at least one U.S. county was exposed. The years for which
+         the tropical cyclones for which at least one US county was exposed. The years for which
          data are available for each metric are given in Table 1.} 

### 2019-01-28 writing/anderson_ehp_supp.tex
Work some on formatting

@@
 included all tracked storms in the revised Atlantic hurricane database
-\ac{HURDAT2}~\citep{landsea2013} that came within 250~\si{\kilo\metre} of at least on
+\ac{HURDAT2}~\parencite{landsea2013} that came within 250~\si{\kilo\metre} of at least on
 \ac{US} county, 1988\,--\,2015. Thicker lines show the tracks of storms whose
 names were retired, indicating that the storm was particularly severe or had
-notable impacts~\citep{retirednames}.}
+notable impacts~\parencite{retirednames}.}
 \label{fig:hurrtracks}
@@
 metric) from each of the clusters shown in the Jaccard heatmap in the main text
-(Figure 2).}
+(Figure 2\todo[inline]{Check fig ref}).}
 \label{fig:othertopstorms}
@@
 Ike (2008) for which exposure classifications from modeled wind
-data~\citep{stormwindmodel} and \ac{HURDAT2} wind
-radii~\citep{demuth2006improvement} showed the highest disagreement for storms
+data~\parencite{stormwindmodel} and \ac{HURDAT2} wind
+radii~\parencite{demuth2006improvement} showed the highest disagreement for storms
 between 2004 and 2015 (Figure~\ref{fig:windcomparison}). The top two maps show
@@
 based on the \ac{HURDAT2} data, a large number of counties were exposed to
-winds of $\ge$34 knots (Figure~\ref{fig:windcomparison}).}
+winds of $\ge$34 knots (Figure~\ref{fig:windcomparison}\todo[inline]{Fix fig ref}).}
 \label{fig:windexamples}
@@
 storm, based on the revised Atlantic hurricane database
-\ac{HURDAT2}~\citep{landsea2013}. The color of each county in the map gives the
-cumulative precipitation (in millimeters) for that county during the window
+\ac{HURDAT2}~\parencite{landsea2013}. The color of each county in the map gives the
+cumulative precipitation (in \si{\milli\metre}) for that county during the window
 from two days before to one day after the storm's closest approach to the
@@
 500~\si{\kilo\metre} of the county's population mean center and the cumulative
-rainfall was 75 millimeters or more (Table 1 of main text). The map on the left
+rainfall was 75 \si{\milli\metre} or more (Table 1 of main text). The map on the left
 shows rain-based exposure for Hurricane Ike. For this exceptionally large
@@
 
-\bibliography{hurr_exposure}
+\printbibliography
+
+\cleardoublepage
+
+\listoftodos
 

### 2019-01-28 writing/ehp_abstract.tex
Work some on formatting

@@
 \textbf{Methods:} We measured county-level exposure to tropical cyclones in the
-United States based on distance from the storm track, maximum sustained wind,
+\ac{US} based on distance from the storm track, maximum sustained wind,
 rainfall, flooding, and tornadoes for all land-falling or near-land Atlantic

### 2019-01-28 writing/ehp_discussion.tex
Work some on formatting

@@
 even strong associations, between tropical cyclone exposure and outcomes of
-interest in impact studies~\citep{savitz2016interpreting, armstrong1998effect}.
+interest in impact studies~\parencite{savitz2016interpreting, armstrong1998effect}.
 
@@
 of interest or in factors associated with risk of that outcome, differential
-exposure misclassification would exist~\citep{savitz2016interpreting}. Such
+exposure misclassification would exist~\parencite{savitz2016interpreting}. Such
 differential exposure misclassification can bias estimates of tropical cyclone
@@
 compared to the true association that exists) or away from the null (estimating
-a larger association than actually exists)~\citep{savitz2016interpreting,
+a larger association than actually exists)~\parencite{savitz2016interpreting,
 armstrong1998effect}. The use of a single hazard-based metric (e.g., wind)
@@
 windspeed are often well-correlated when the tropical cyclone is over
-water~\citep{cerveny2000}, this relationship often does not remain as strong
-once the hurricane has made landfall~\citep{jiang2008}.  Fast-moving tropical
-cyclones bring higher risks of dangerous winds inland~\citep{kruk2010}, while
+water~\parencite{cerveny2000}, this relationship often does not remain as strong
+once the hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
+cyclones bring higher risks of dangerous winds inland~\parencite{kruk2010}, while
 slow-moving tropical cyclone are likely to bring more
-rain~\citep{rappaport2000} and cause more damage because of sustained hazardous
-conditions~\citep{rezapour2014}. Further, while the likelihood and extent of
+rain~\parencite{rappaport2000} and cause more damage because of sustained hazardous
+conditions~\parencite{rezapour2014}. Further, while the likelihood and extent of
 flooding during a tropical cyclone is related to the tropical cyclone's
 rainfall, it is also driven by factors like top soil saturation and the
-structure of the water basin's drainage network~\citep{chen2015, rees2001}.
+structure of the water basin's drainage network~\parencite{chen2015, rees2001}.
 However, we did find a small set of tropical cyclones for which for which
@@
 precipitation during tropical cyclones and provides hydrological conditions for
-severe flooding~\citep{rees2001}.  Almost all tornado-based exposures were in
+severe flooding~\parencite{rees2001}.  Almost all tornado-based exposures were in
 southern coastal states. This pattern echoes previous findings that most
 noteworthy tropical cyclone-related tornadoes occur on the right-side of the
-tropical cyclone track in Atlantic-basin \ac{US} storms~\citep{moore2012}.
+tropical cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.
 
@@
 frequencies, as tropical cyclones follow decadal
-patterns~\citep{kossin2007more} likely not adequately captured in the available
+patterns~\parencite{kossin2007more} likely not adequately captured in the available
 data.  However, these frequency maps, together with evidence from specific
@@
 A few previous studies have sought to determine county-level exposure to
-tropical cyclones over multi-year periods, including~\citet{zandbergen2009},
+tropical cyclones over multi-year periods, including~\textcite{zandbergen2009},
 which estimated exposure in \ac{US} counties to all \ac{US} landfalling
@@
 distance-based metric and a metric that combined distance and windspeed,
-and~\citet{kruk2010}, which explored exposure to hurricane-related winds in the
+and~\textcite{kruk2010}, which explored exposure to hurricane-related winds in the
 \ac{US}, including inland areas, for~1900\,--\,2008.  Our results suggest that
@@
 Dissimilarity in exposure to specific hazards of tropical cyclones can result
-in differing patterns in estimates of county-level exposure (Figure~1 [change
-to fig. ref]), which can lead to substantial differences in estimates of
-population physical exposure (Figure~\ref{fig:topelecdependexposure}), based on
-which storm hazard is considered. The measure of a hurricane's intensity, the
-Saffir-Simpson scale, expresses intensity as a function of wind speed, and so
-fails to capture potential risks from other hazards of the
-storm~\citep{smith2009}, and previous assessments of physical exposure to
-tropical cyclones have often focused on winds or distance from the storm. For
-example, a global study of exposure and vulnerability to natural hazards used a
-combination of windspeed above a certain threshold and distance from the
-storm's track to assess exposure~\citep{peduzzi2009assessing}. While many storm
-impacts might be most strongly linked to wind hazards, there is a growing
-recognition of the potential risks of adverse health outcomes and property
-damage from rain- and flood-related tropical cyclone hazards in the
-\ac{US}~\citep{smith2009}. Such assessments may underemphasize potentially
-dangerous exposures to certain regions of the \ac{US}, particularly inland
-locations. In fact, based on our assessment of physical exposures to tropical
-cyclones among electricity-dependent Medicare beneficiaries, the highest
-expected physical exposure under any metric was measured for Philadelphia
-County, PA, under the flood metric (Figure~\ref{fig:topelecdependexposure}).
-While this county has a similar-sized population of electricity-dependent
-Medicare beneficiaries as Miami-Dade County, FL
-(Figure~\ref{fig:topelecdependexposure}), the frequency of storm-related flood
-events in this inland county results in a very high expected physical exposure
-among this subpopulation, while this county is not included in the top ten
-counties for physical exposure based on most of the other exposure metrics.
+in differing patterns in estimates of county-level exposure
+(Figure~1\todo{change to fig. ref}), which can lead to substantial differences
+in estimates of population physical exposure
+(Figure~\ref{fig:topelecdependexposure}), based on which storm hazard is
+considered. The measure of a hurricane's intensity, the Saffir-Simpson scale,
+expresses intensity as a function of wind speed, and so fails to capture
+potential risks from other hazards of the storm~\parencite{smith2009}, and
+previous assessments of physical exposure to tropical cyclones have often
+focused on winds or distance from the storm. For example, a global study of
+exposure and vulnerability to natural hazards used a combination of windspeed
+above a certain threshold and distance from the storm's track to assess
+exposure~\parencite{peduzzi2009assessing}. While many storm impacts might be
+most strongly linked to wind hazards, there is a growing recognition of the
+potential risks of adverse health outcomes and property damage from rain- and
+flood-related tropical cyclone hazards in the \ac{US}~\parencite{smith2009}.
+Such assessments may underemphasize potentially dangerous exposures to certain
+regions of the \ac{US}, particularly inland locations. In fact, based on our
+assessment of physical exposures to tropical cyclones among
+electricity-dependent Medicare beneficiaries, the highest expected physical
+exposure under any metric was measured for Philadelphia County, PA, under the
+flood metric (Figure~\ref{fig:topelecdependexposure}).  While this county has a
+similar-sized population of electricity-dependent Medicare beneficiaries as
+Miami-Dade County, FL (Figure~\ref{fig:topelecdependexposure}), the frequency
+of storm-related flood events in this inland county results in a very high
+expected physical exposure among this subpopulation, while this county is not
+included in the top ten counties for physical exposure based on most of the
+other exposure metrics.
 
@@
 To assist with future tropical cyclone studies, we created and published open
-source software to accompany this article~\citep{hurricaneexposuredata,
+source software to accompany this article~\parencite{hurricaneexposuredata,
 hurricaneexposure}. Many previous studies have used geographical
 information system software (e.g., ArcGIS) to assess exposure to tropical
-cyclones in the \ac{US}~\citep{grabich2016, zandbergen2009, czajkowski2011,
+cyclones in the \ac{US}~\parencite{grabich2016, zandbergen2009, czajkowski2011,
 kruk2010}. Here, we offer methods to map and output historic exposure to
@@
 instead uses a package written in the R statistical programming
-language~\citep{R}, which is free and open-source.
+language~\parencite{R}, which is free and open-source.
 
@@
 exposure profiles based on multiple hazards or craft exposure indices that
-combine hazard metrics~\citep{chakraborty2005population, peduzzi2009assessing}.
+combine hazard metrics~\parencite{chakraborty2005population, peduzzi2009assessing}.
 This ability can be critical, as different hazards of tropical cyclones
-often act synergistically in causing impacts~\citep{smith2009}. Further, by
+often act synergistically in causing impacts~\parencite{smith2009}. Further, by
 including measurements of different hazard exposures in each county for each
@@
 within available dates. Further, we have investigated how well these data
-correspond with data from other available sources (Figures~S7\,--\,11). While
-generally in agreement with data from other sources, there are a few
-limitations. The wind data are based on modeled, rather than observed, values,
-and while the modeled wind data generally agree well with post-analysis maximum
-wind radii from \ac{HURDAT2}~\citep{landsea2013} (Figure~S7), for a few
-tropical cyclones it did not (e.g., Figure~S10). The rainfall data are from
-re-analysis data, which are generally well-correlated with observed
-ground-based station data but sometimes oversmooth extreme measurements
-(Figure~S8).  The flood and tornado data came from the \ac{NOAA} Storm Events
-database, which, while a widely-used database of events maintained by
-\ac{NOAA}, is based on reports, and so may be prone to
-underreporting~\citep{Ashley2008flood, Curran2000}, especially in less
-populated areas~\citep{Witt1998, Ashley2007}, as well as to other reporting
+correspond with data from other available sources
+(Figures~S7\,--\,11\todo{Check fig refs}). While generally in agreement with
+data from other sources, there are a few limitations. The wind data are based
+on modeled, rather than observed, values, and while the modeled wind data
+generally agree well with post-analysis maximum wind radii from
+\ac{HURDAT2}~\parencite{landsea2013} (Figure~S7\todo{Check fig refs}), for a
+few tropical cyclones it did not (e.g., Figure~S10\todo{Check fig refs}). The
+rainfall data are from re-analysis data, which are generally well-correlated
+with observed ground-based station data but sometimes oversmooth extreme
+measurements (Figure~S8\todo{Check fig refs}).  The flood and tornado data came
+from the \ac{NOAA} Storm Events database, which, while a widely-used database
+of events maintained by \ac{NOAA}, is based on reports, and so may be prone to
+underreporting~\parencite{Ashley2008flood, Curran2000}, especially in less
+populated areas~\parencite{Witt1998, Ashley2007}, as well as to other reporting
 errors. While these are important limitations, we selected these data sources

### 2019-01-28 writing/ehp_figures.tex
Work some on formatting

@@
 exposure metric (Table~1). The red line shows the track of Hurricane Ivan 
-based on the revised Atlantic hurricane database \ac{HURDAT2}~\citep{landsea2013}.
-Similar maps for other large-extent storms are given in Figure~S3.}
+based on the revised Atlantic hurricane database \ac{HURDAT2}~\parencite{landsea2013}.
+Similar maps for other large-extent storms are given in Figure~S3\todo[inline]{Check fig ref}.}
 \label{fig:ivanexposure} 
@@
 county-level exposure agreement for metric pairs, based on hierarchical
-clustering using the complete link method~\citep{murtagh2012algorithms} (i.e.,
+clustering using the complete link method~\parencite{murtagh2012algorithms} (i.e.,
 storms in the same cluster tend to have similar patterns for the pairwise
@@
 (Figure~\ref{fig:ivanexposure}) and Hurricanes Floyd (1999), Lee (2011), Cindy
-(2005), and Katrina (2005) (Figure~S3).} 
+(2005), and Katrina (2005) (Figure~S3\todo[inline]{Check fig ref}).} 
 \label{fig:jaccard}
@@
 analysis is restricted to years with all exposure data available (1996\,--\,2011;
-Figure~S5).} 
+Figure~S5\todo[inline]{Check fig ref}).} 
 \label{fig:averageexposure} 
@@
 number of Medicare beneficiaries in the county reliant on electricity-dependent
-medical and assistive equipment as of July~2017~\citep{empower}. The length of
+medical and assistive equipment as of July~2017~\parencite{empower}. The length of
 each bar shows the average expected number of these electricity-dependent
@@
 speed (as used for the wind exposure metric in the main
-text;~\citep{stormwindmodel}) versus sustained wind speed determined from the
-revised Atlantic hurricane database \ac{HURDAT2}~\citep{landsea2013}. In
-\ac{HURDAT2}, wind radii were available for study storms from~2004\,--\,2015
-and gives estimates of the maximum radii to which storm winds of 34~\si{\knot},
-50~\si{\knot}, and 64~\si{\knot} extended in each four quadrant of the storm at
-each of the 6-\si{\hour} storm track observations. We interpolated this data
-to~15-\si{\minute} increments and assessed a county as exposed to winds at or above
-one of these cutoffs if its population mean center was within~85\% of the
+text;~\parencite{stormwindmodel}) versus sustained wind speed determined from
+\ac{HURDAT2}~\parencite{landsea2013}. In \ac{HURDAT2}, wind radii were
+available for study storms from~2004\,--\,2015 and gives estimates of the
+maximum radii to which storm winds of 34~\si{\knot}, 50~\si{\knot}, and
+64~\si{\knot} extended in each four quadrant of the storm at each of the
+6-\si{\hour} storm track observations. We interpolated this data
+to~15-\si{\minute} increments and assessed a county as exposed to winds at or
+above one of these cutoffs if its population mean center was within~85\% of the
 maximum radius for that wind speed in its quadrant of the storm. For each
 county in each of the available storms, we also calculated the modeled wind
-speed~\citep{stormwindmodel} and then categorized this value into the same
+speed~\parencite{stormwindmodel} and then categorized this value into the same
 categories available for the \ac{HURDAT2} wind radii ($<$34~\si{\knot};
@@
 comparisons between the two data sources for these two storms are given in
-Figure~\ref{fig:windexamples}.}
+Figure~\ref{fig:windexamples}\todo[inline]{Fix fig ref}.}
 \label{fig:windcomparison}
@@
 county-level rain exposure based on the reanalysis data used for exposure
-assessment in this study \ac{NLDAS-2}~\citep{rui2013nldas,
+assessment in this study \ac{NLDAS-2}~\parencite{rui2013nldas,
 alhamdan2014environmental} and ground-based weather
-stations~\citep{menne2012overview, rnoaa, countyweather}. Each small graph
+stations~\parencite{menne2012overview, rnoaa, countyweather}. Each small graph
 shows a sample county, and each point on the graph represents a tropical storm
@@
 differ by county. An estimate of the rank correlation
-(Spearman's~$\rho$~\citep{spearman1904proof}) between measurements from the two
+(Spearman's~$\rho$~\parencite{spearman1904proof}) between measurements from the two
 rain data sources is also shown on each graph.} \label{fig:raincomparison}
@@
 \caption{Comparison of for a sample of study counties of flood status, based on
-NOAA Storm Events listings~\citep{stormevents, noaastormevents}, with total
-streamflows at county streamflow gages during tropical storms~\citep{usgsgages,
+NOAA Storm Events listings~\parencite{stormevents, noaastormevents}, with total
+streamflows at county streamflow gages during tropical storms~\parencite{usgsgages,
 countyfloods, dataRetrieval}. Each panel represents a separate county. For each
@@
 Events database for the county with a start date within the five-day window of
-the storm's closest approach to the county~\citep{stormevents,
+the storm's closest approach to the county~\parencite{stormevents,
 noaastormevents}. Boxplots are added to show the distribution of the total
@@
 the county with a daily streamflow that exceeded a threshold of flooding (the
-streamgage's median value for annual peak flow~\citep{countyfloods}) on any day
+streamgage's median value for annual peak flow~\parencite{countyfloods}) on any day
 during the five-day window.  Note that the x-axis scales differ by county,

### 2019-01-28 writing/ehp_introduction.tex
Work some on formatting

@@
 Tropical cyclones, including hurricanes, can have disastrous impacts in the
-\ac{US}. A tropical cyclone's high winds can bring health risks
+\acf{US}. A tropical cyclone's high winds can bring health risks
 and property damage through structural damage of houses and other buildings,
-falling trees, and wind-borne debris~\citep{rappaport2000} and cause power
-outages~\citep{liu2005, han2009}, which introduce a number of threats to human
+falling trees, and wind-borne debris~\parencite{rappaport2000} and cause power
+outages~\parencite{liu2005, han2009}, which introduce a number of threats to human
 and ecological health, including water quality risks if the outage affects
-wastewater treatment plants~\citep{mallin2006}.  Other risks can exist without
+wastewater treatment plants~\parencite{mallin2006}.  Other risks can exist without
 severe wind; for example, one study found that most of the direct
@@
 when wind was below hurricane strength, including for Tropical Storms Charley
-in~1998 and Alberto in~1994~\citep{rappaport2000}.  Tropical cyclones can
+in~1998 and Alberto in~1994~\parencite{rappaport2000}.  Tropical cyclones can
 produce excessive rain, especially in certain topographies (e.g., near
@@
 coastal counties. Flood risks from tropical cyclones can result from this rain,
-although the two risks are not perfectly correlated~\citep{chen2015}. In the
+although the two risks are not perfectly correlated~\parencite{chen2015}. In the
 \ac{US}, over half of hurricane-related direct deaths from~1970 to~1999 from
-Atlantic basin storms were a result of freshwater flooding~\citep{rappaport2000}. 
-Flooding can also degrade water quality~\citep{mallin2006}, which can threaten both 
+Atlantic basin storms were a result of freshwater flooding~\parencite{rappaport2000}. 
+Flooding can also degrade water quality~\parencite{mallin2006}, which can threaten both 
 human and ecological health.
@@
 those related to human health, studies have varied widely in how they determine
-tropical cyclone exposure (some examples shown in Figure~S1). While some
-studies have measured exposure based on specific hazards of the tropical
-cyclone (e.g., wind, rain), many have used distance from the tropical
-cyclone's central track as a surrogate in identifying exposure to the hazards
-of a tropical cyclone (e.g.,~\citet{czajkowski2011, tansel2010, kinney2008,
-caillouet2008increase}).
+tropical cyclone exposure (some examples shown in Figure~S1\todo{Check
+fig.ref}). While some studies have measured exposure based on specific hazards
+of the tropical cyclone (e.g., wind, rain), many have used distance from the
+tropical cyclone's central track as a surrogate in identifying exposure to the
+hazards of a tropical cyclone (e.g.,~\textcite{czajkowski2011, tansel2010,
+kinney2008, caillouet2008increase}).
 
@@
 been observed with radii to maximum winds as small as~20~\si{\kilo\metre} and as
-large as~200~\si{\kilo\metre}~\citep{mallin2006, quiring2011variations}. While a
+large as~200~\si{\kilo\metre}~\parencite{mallin2006, quiring2011variations}. While a
 number of tropical cyclone hazards are strongly associated with distance from
 the tropical cyclone's center (e.g., wind and, at the coast, storm surge and
-waves~\citep{rappaport2000, kruk2010}), other hazards like dangerous rain,
+waves~\parencite{rappaport2000, kruk2010}), other hazards like dangerous rain,
 floods, and tornadoes can occur well away from the tropical cyclone's central
-track~\citep{rappaport2000, atallah2007, moore2012}.  For example, fatal
+track~\parencite{rappaport2000, atallah2007, moore2012}.  For example, fatal
 tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
 between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
-tropical cyclone's center~\citep{moore2012}.
+tropical cyclone's center~\parencite{moore2012}.
 
@@
 cyclonic winds move in concert with the tropical cyclone's forward 
-motion~\citep{halverson2015}, and the fatal tornadoes associated with \ac{US} tropical
+motion~\parencite{halverson2015}, and the fatal tornadoes associated with \ac{US} tropical
 cyclones between~1995 and~2009 occurred almost exclusively to the right of the
 tropical cyclone's track, mostly in the right front quadrant of the tropical
-cyclone~\citep{moore2012}. Rain, conversely, is often heaviest to the left of
+cyclone~\parencite{moore2012}. Rain, conversely, is often heaviest to the left of
 the tropical cyclone's track, especially when the tropical cyclone interacts
-with other weather systems~\citep{atallah2003, atallah2007, zhu2013variations}
-or undergoes an extratropical transition~\citep{elsberry2002}.
+with other weather systems~\parencite{atallah2003, atallah2007, zhu2013variations}
+or undergoes an extratropical transition~\parencite{elsberry2002}.
 
-[Something about differences in other metrics]
+\todo[inline]{Something about difference in other metrics}
 
@@
 many potential impacts are available at county-level aggregations (e.g., direct
-hurricane-related deaths~\citep{czajkowski2011}; birth
-outcomes~\citep{grabich2015, grabich2016}; autism
-prevalence~\citep{kinney2008}) and since decisions and policies to prepare for,
+hurricane-related deaths~\parencite{czajkowski2011}; birth
+outcomes~\parencite{grabich2015, grabich2016}; autism
+prevalence~\parencite{kinney2008}) and since decisions and policies to prepare for,
 and respond to, tropical cyclones are often undertaken at the county
-level~\citep{zandbergen2009, rappaport2000}.  We explored patterns in these
+level~\parencite{zandbergen2009, rappaport2000}.  We explored patterns in these
 exposure assessments and measured how well exposure classification agreed
@@
 for county-based hurricane exposure assessment in \ac{US}
-counties~\citep{hurricaneexposure}.
+counties~\parencite{hurricaneexposure}.

### 2019-01-28 writing/ehp_methods.tex
Work some on formatting

@@
 included in the study. The study covered all tracked storms in the \ac{HURDAT2}
-\citep{landsea2013} that came within~250 \si{\kilo\metre} of at least one study
-county between~1988 and~2015 (Figure~S2).  \ac{HURDAT2} is a post-storm
-assessment conducted by the \ac{US} \ac{NHC} and incorporates data from a
-variety of sources, including satellite data and, when available, aircraft
-reconnaissance data~\citep{landsea2013, jarvinen1988}. These data typically
-give measurements of the tropical cyclone center's location at~6-\si{\hour} intervals
-at synoptic times (i.e., 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC});
-some landfalling tropical cyclones have an additional observation at the time
-of landfall~\citep{landsea2013}.
+\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
+study county between~1988 and~2015 (Figure~S2\todo{Check fig ref}).
+\ac{HURDAT2} is a post-storm assessment conducted by the \ac{US} \ac{NHC} and
+incorporates data from a variety of sources, including satellite data and, when
+available, aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.
+These data typically give measurements of the tropical cyclone center's
+location at~6-\si{\hour} intervals at synoptic times (i.e., 6:00~am, 12:00~pm,
+6:00~pm, and 12:00~am \ac{UTC}); some landfalling tropical cyclones have an
+additional observation at the time of landfall~\parencite{landsea2013}.
 
@@
 county.  Storm center track location data from \ac{HURDAT2} were interpolated
-using natural cubic splines to~15-\si{\minute} intervals~\citep{hurricaneexposure}.
+using natural cubic splines to~15-\si{\minute} intervals~\parencite{hurricaneexposure}.
 The distance between each county's population mean center, as assessed by the
-2010 Decennial Census~\citep{countycenters}, and each~15-\si{\minute} interpolated
+2010 Decennial Census~\parencite{countycenters}, and each~15-\si{\minute} interpolated
 storm center location was measured using Great Circle (WGS84 ellipsoid)
-distance~\citep{bivand2013applied}. The distance-based exposure metric was
+distance~\parencite{bivand2013applied}. The distance-based exposure metric was
 based on the minimum value of this storm center-to-county distance throughout
@@
 tornado-based metrics. To allow matching with data based on local time, times
-were converted from \ac{UTC} to local time~\citep{countytimezones}.
+were converted from \ac{UTC} to local time~\parencite{countytimezones}.
 
@@
 aggregating hourly~1/8\si{\degree} gridded reanalysis data from the \ac{NLDAS-2}
-precipitation data files~\citep{rui2013nldas}. The \ac{NLDAS-2} data integrate
+precipitation data files~\parencite{rui2013nldas}. The \ac{NLDAS-2} data integrate
 satellite-based and land-based monitoring, applying a land-surface model to
 create a reanalysis dataset that is spatially and temporally complete across
-the continental \ac{US}~\citep{rui2013nldas, alhamdan2014environmental}. To
+the continental \ac{US}~\parencite{rui2013nldas, alhamdan2014environmental}. To
 aggregate to county-level, hourly data at each grid point were summed to create
@@
 across all grid points within a county's~1990 \ac{US} Census boundaries to
-create a county-level daily total~\citep{alhamdan2014environmental, cdcwonder}.
+create a county-level daily total~\parencite{alhamdan2014environmental, cdcwonder}.
 This county-level precipitation data are publicly available through the \ac{US}
-\ac{CDC}'s \ac{WONDER} database~\citep{cdcwonder, alhamdan2014environmental}.
+\ac{CDC}'s \ac{WONDER} database~\parencite{cdcwonder, alhamdan2014environmental}.
 The rainfall-based exposure metric was only available for tropical cyclones
 through~2011, the period for which the county-aggregated reanalysis rainfall
-data were available~\citep{cdcwonder, alhamdan2014environmental}.
+data were available~\parencite{cdcwonder, alhamdan2014environmental}.
 
@@
 available daily data in the Global Historical Climatology Network throughout
-1988\,--\,2011~\citep{menne2012overview, rnoaa, countyweather}. Agreement
+1988\,--\,2011~\parencite{menne2012overview, rnoaa, countyweather}. Agreement
 between measured storm-related rainfall between these two data sources was
@@
 modeled maximum sustained wind speeds at each county's population mean
-center~\citep{countycenters}. \ac{HURDAT2} Best Track data were interpolated
+center~\parencite{countycenters}. \ac{HURDAT2} Best Track data were interpolated
 from~6-\si{\hour} reported values to~15-\si{\minute} increments using natural cubic
-splines~\citep{stormwindmodel}, and then sustained wind speed was modeled for
+splines~\parencite{stormwindmodel}, and then sustained wind speed was modeled for
 each county center at each~15-\si{\minute} increment using a double exponential wind
-speed model~\citep{willoughby2006parametric} to estimate maximum ground-level
-sustained wind speed at each county center~\citep{stormwindmodel}. Asymmetry in
+speed model~\parencite{willoughby2006parametric} to estimate maximum ground-level
+sustained wind speed at each county center~\parencite{stormwindmodel}. Asymmetry in
 wind speeds around the tropical cyclone center was incorporated into the wind
-speed model~\citep{phadke2003modeling, stormwindmodel}. The maximum value of
+speed model~\parencite{phadke2003modeling, stormwindmodel}. The maximum value of
 this modeled wind speed was determined for each county as the maximum windspeed
@@
 compared this wind metric to estimated wind radii from \ac{HURDAT2}, available
-for storms in our study from~2004~\citep{landsea2013}. \ac{HURDAT2} gives
+for storms in our study from~2004~\parencite{landsea2013}. \ac{HURDAT2} gives
 estimates of the maximum radii to winds of~64,~50, and~34 \si{\knot} in each four
@@
 counties, we used event listings from the National Oceanic and Atmospheric
-Administration (NOAA)'s Storm Event Database~\citep{stormevents}. For each
+Administration (NOAA)'s Storm Event Database~\parencite{stormevents}. For each
 tropical cyclone, we identified all events with event types related to flooding
@@
 tropical cyclone's closest approach to the
-county~\citep{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado"
+county~\parencite{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado"
 events in this database were reported by county \ac{FIPS} code. ``Coastal
 Flood" events were reported by forecast zone; for these, the event was matched
-to the appropriate county if possible~\citep{noaastormevents}. While this
+to the appropriate county if possible~\parencite{noaastormevents}. While this
 database has recorded storm data since~1950, its coverage changed substantially
-in~1996 to cover a wider variety of storm events~\citep{stormevents}. We
+in~1996 to cover a wider variety of storm events~\parencite{stormevents}. We
 therefore only considered these metrics of hurricane exposure for tropical
@@
 As validation for the flood data, we compared streamflow data from \ac{US}
-Geological Survey gages~\citep{usgsgages, countyfloods, dataRetrieval} in nine
+Geological Survey gages~\parencite{usgsgages, countyfloods, dataRetrieval} in nine
 counties with tropical cyclone flood classifications based on the NOAA Storm
@@
 that made \ac{US} landfall or passed within~250 \si{\kilo\metre} (155 mi) of at
-least one \ac{US} county (Figure~S2).
+least one \ac{US} county (Figure~S2\todo{Check fig ref}).
 
@@
 measurement, we calculated the within-storm Jaccard
-index~\citep{jaccard1901distribution, jaccard1908nouvelles} between every
+index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between every
 pair-wise combination of hazard metrics.  The Jaccard index~($J_s$) measures
@@
 as well as on the vulnerability of the society living in the geographical areas
-affected by those forces.~\citep{chakraborty2005population,
+affected by those forces.~\parencite{chakraborty2005population,
 anderson2003community, cutter1996vulnerability} As a case study of how
@@
 each study county from the U.S. Department of Health \& Human Service's emPower
-Map~2.0~\citep{empower}. We then calculated the physical exposure of the
+Map~2.0~\parencite{empower}. We then calculated the physical exposure of the
 electricity-dependent Medicare population in each county, based on tropical
 cyclone assessments using each hazard metric,
-following~\citep{peduzzi2009assessing}:
+following~\parencite{peduzzi2009assessing}:
 
@@
 used to calculate risk of human losses from the
-hazard~\citep{peduzzi2009assessing}.
+hazard~\parencite{peduzzi2009assessing}.

### 2019-01-28 writing/ehp_results.tex
Work some on formatting

@@
 tropical cyclone exposure assessment is to the choice of metric, using as an
-example Hurricane Ivan in 2004, which was record-breaking for its
-duration as a major hurricane~\citep{franklin2006atlantic}. When county-level
-exposure was determined based on distance, the counties assessed as exposed
-followed the tropical cyclone's track, including counties well inland. For the
-wind metric, only counties near two of the tropical cyclone's landfalls were
-assessed as exposed. For rain- and flood-based metrics, however, exposure
-extended to the left of the track, including counties as far north as New York
-and Connecticut, while for the tornado metric, exposed counties tended to be to
-the right of the track and included several counties in central North Carolina,
-South Carolina, and Georgia that were not identified as exposed to Ivan based
-on any other metric.  Figure~S3 provides similar maps for four other example
-tropical cyclones (selected because they exposed many \ac{US}  counties based on
-at least one metric), and Figure~S4 provides four examples of tropical cyclones
-in which the disagreement in exposure assessment between rain- and wind-based
-metrics was particularly notable.
+example Hurricane Ivan in 2004, which was record-breaking for its duration as a
+major hurricane~\parencite{franklin2006atlantic}. When county-level exposure
+was determined based on distance, the counties assessed as exposed followed the
+tropical cyclone's track, including counties well inland. For the wind metric,
+only counties near two of the tropical cyclone's landfalls were assessed as
+exposed. For rain- and flood-based metrics, however, exposure extended to the
+left of the track, including counties as far north as New York and Connecticut,
+while for the tornado metric, exposed counties tended to be to the right of the
+track and included several counties in central North Carolina, South Carolina,
+and Georgia that were not identified as exposed to Ivan based on any other
+metric.  Figure~S3\todo{Check fig ref} provides similar maps for four other
+example tropical cyclones (selected because they exposed many \ac{US}  counties
+based on at least one metric), and Figure~S4\todo{Check fig ref} provides four
+examples of tropical cyclones in which the disagreement in exposure assessment
+between rain- and wind-based metrics was particularly notable.
 
@@
 counties were determined to be exposed to the tropical cyclone for two or more
-of the metrics considered.  For Floyd in~1999 (Figure~S3) and Irene in~2011,
-for example, county-level classification agreed moderately to well (Jaccard
-index of approximately~0.5\,--\,0.8) for all pairs of exposure metrics except
-those including the tornado-based metric. These tropical cyclones both made
-their first \ac{US}  landfall in North Carolina at minor hurricane windspeeds
-(Category~2 and~1, respectively) and then skimmed the eastern coast of the
-\ac{US} north through New England, bringing large rainfall to much of the
-eastern coast from North Carolina north and causing extensive inland flooding
-in North Carolina (Floyd) and New England (Irene)~\citep{avila2013atlantic,
-lawrence2000atlantic}. For another set of tropical cyclones (e.g., Lee in~2011
-[Figure~S3], Ernesto in~2006, and Bertha in~1996), there was moderate to good
-agreement for all pairwise combinations of distance, rain, and wind, but poor
-agreement for all other combinations of metrics. Most of these tropical
-cyclones either followed the eastern coastline north of North Carolina (Bertha
-in~1996, Charley in~2004, Gaston in~2004, Barry in~2007, and Hanna in~2008), in
-a pattern similar to Hurricanes Floyd in~1999 and Irene in~2011, or made
-landfall in or near the Florida panhandle and cut across to exit into the
-Atlantic around North Carolina (Josephine in~1996, Earl in~1998, Alberto
-in~2006). However, these two sets of tropical cyclones represented the minority
-of all tropical cyclones considered; for most tropical cyclones, there was low
-overlap in the counties determined to be exposed to the tropical cyclones
-between most pairings of metrics.
+of the metrics considered.  For Floyd in~1999 (Figure~S3\todo{Check fig ref})
+and Irene in~2011, for example, county-level classification agreed moderately
+to well (Jaccard index of approximately~0.5\,--\,0.8) for all pairs of exposure
+metrics except those including the tornado-based metric. These tropical
+cyclones both made their first \ac{US}  landfall in North Carolina at minor
+hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
+eastern coast of the \ac{US} north through New England, bringing large rainfall
+to much of the eastern coast from North Carolina north and causing extensive
+inland flooding in North Carolina (Floyd) and New England
+(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}. For another set of
+tropical cyclones (e.g., Lee in~2011 [Figure~S3\todo{Check fig ref}], Ernesto
+in~2006, and Bertha in~1996), there was moderate to good agreement for all
+pairwise combinations of distance, rain, and wind, but poor agreement for all
+other combinations of metrics. Most of these tropical cyclones either followed
+the eastern coastline north of North Carolina (Bertha in~1996, Charley in~2004,
+Gaston in~2004, Barry in~2007, and Hanna in~2008), in a pattern similar to
+Hurricanes Floyd in~1999 and Irene in~2011, or made landfall in or near the
+Florida panhandle and cut across to exit into the Atlantic around North
+Carolina (Josephine in~1996, Earl in~1998, Alberto in~2006). However, these two
+sets of tropical cyclones represented the minority of all tropical cyclones
+considered; for most tropical cyclones, there was low overlap in the counties
+determined to be exposed to the tropical cyclones between most pairings of
+metrics.
 
@@
 subpopulation, Medicare beneficiaries who are reliant on electricity for
-medical equipment (Figure~S6)~\citep{empower}.  We found
-that a few counties (Miami-Dade County, FL; Harris County, TX) were ranked in
-the top three counties of physical exposure among electricity-dependent
-Medicare beneficiaries for almost all exposure metrics
+medical equipment (Figure~S6\todo{Check fig ref})~\parencite{empower}.  We
+found that a few counties (Miami-Dade County, FL; Harris County, TX) were
+ranked in the top three counties of physical exposure among
+electricity-dependent Medicare beneficiaries for almost all exposure metrics
 (Figure~\ref{fig:topelecdependexposure}). However, there were key differences
@@
 cyclone data in this paper, as well as tools to explore and map that 
-data~\citep{hurricaneexposure, hurricaneexposuredata}. This software includes continuous
+data~\parencite{hurricaneexposure, hurricaneexposuredata}. This software includes continuous
 data for the distance-, rain-, and wind-based metrics, allowing users to also explore
@@
 We investigated how well these data correspond with data from other available
-sources (Figures~S7\,--\,11). Within the nine sample counties, storm-related
-rainfall from the reanalysis precipitation dataset was well-correlated with
-station-based rainfall measurements, although the reanalysis data did tend to
-somewhat underestimate rainfall totals in storms with extremely high rainfall
-(Figure~\ref{fig:raincomparison}). A distance constraint (county population
-mean center within~500 \si{\kilo\metre} of the storm's closest approach) was
-necessary for assessing storm exposure based on this metric to ensure that
-rains unrelated and far from the storm track were not misattributed to storm
-exposures. We found that this distance constraint was typically adequately
-large to capture storm-related rain, although in rare examples of exceptionally
-large storms (e.g., Hurricane Ike in 2008) or storms for which tracking was
-stopped at extratropical transition (e.g., Tropical Storm Lee in 2011), some
-storm-related rains were missed because of this distance constraint
-(Figure~\ref{fig:rainexamples}).  The modeled wind speeds generally agreed well
-with the \ac{HURDAT2} wind radii (Figure~\ref{fig:windcomparison}), with
-disagreement mostly limited to large storms that were characterized by high
-winds well inland (e.g., Hurricanes Ike in 2008 and Sandy in 2012;
-Figure~\ref{fig:windexamples}).  Stream gage data generally indicated higher
-discharge in these counties on dates identified as flood events, particularly
-in counties with more reporting stream gages (e.g., Harris County, TX, with~10
-reporting stream gages; Figure~\ref{fig:floodcomparison}), although these two
-flooding data sources were not completely consistent.
+sources (Figures~S7\,--\,11\todo{Check fig refs}). Within the nine sample
+counties, storm-related rainfall from the reanalysis precipitation dataset was
+well-correlated with station-based rainfall measurements, although the
+reanalysis data did tend to somewhat underestimate rainfall totals in storms
+with extremely high rainfall (Figure~\ref{fig:raincomparison}). A distance
+constraint (county population mean center within~500 \si{\kilo\metre} of the
+storm's closest approach) was necessary for assessing storm exposure based on
+this metric to ensure that rains unrelated and far from the storm track were
+not misattributed to storm exposures. We found that this distance constraint
+was typically adequately large to capture storm-related rain, although in rare
+examples of exceptionally large storms (e.g., Hurricane Ike in 2008) or storms
+for which tracking was stopped at extratropical transition (e.g., Tropical
+Storm Lee in 2011), some storm-related rains were missed because of this
+distance constraint (Figure~\ref{fig:rainexamples}).  The modeled wind speeds
+generally agreed well with the \ac{HURDAT2} wind radii
+(Figure~\ref{fig:windcomparison}), with disagreement mostly limited to large
+storms that were characterized by high winds well inland (e.g., Hurricanes Ike
+in 2008 and Sandy in 2012; Figure~\ref{fig:windexamples}).  Stream gage data
+generally indicated higher discharge in these counties on dates identified as
+flood events, particularly in counties with more reporting stream gages (e.g.,
+Harris County, TX, with~10 reporting stream gages;
+Figure~\ref{fig:floodcomparison}), although these two flooding data sources
+were not completely consistent.
 
-While generally in agreement with data from other
-sources, there are a few limitations. The wind data are based on modeled,
-rather than observed, values, and while the modeled wind data generally agree
-well with post-analysis maximum wind radii from the revised Atlantic hurricane
-database \ac{HURDAT2}~\citep{landsea2013} (Figure~S7), for a few tropical cyclones
-it did not (e.g., Figure~S10). The rainfall data are from re-analysis data,
-which are generally well-correlated with observed ground-based station data but
-sometimes oversmooth extreme measurements (Figure~S8).
+While generally in agreement with data from other sources, there are a few
+limitations. The wind data are based on modeled, rather than observed, values,
+and while the modeled wind data generally agree well with post-analysis maximum
+wind radii from the revised Atlantic hurricane database
+\ac{HURDAT2}~\parencite{landsea2013} (Figure~S7\todo{Check fig ref}), for a few
+tropical cyclones it did not (e.g., Figure~S10\todo{Check fig ref}). The
+rainfall data are from re-analysis data, which are generally well-correlated
+with observed ground-based station data but sometimes oversmooth extreme
+measurements (Figure~S8\todo{Check fig ref}).

### 2019-01-28 writing/ehp_tables.tex
Work some on formatting

@@
 Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 \si{\kilo\metre} of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
-Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
-Tornado & 1996--2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track\\
+Distance & 1988\,--\,2015 & County population mean center within 100~\si{\kilo\metre} of storm track \\ 
+Rain & 1988\,--\,2011 & County cumulative rainfall of $\ge$75~\si{\milli\metre} over the period from two days before to 
+one day after the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the 
+storm track \\ 
+Wind & 1988\,--\,2015 & Modeled storm-associated maximum sustained winds of $\ge$15~\si{\metre\per\second}
+at the county's population mean center\\ 
+Flood & 1996\,--\,2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the storm track \\
+Tornado & 1996\,--\,2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the storm track\\
 \bottomrule 

### 2019-01-28 R/storm_severity_table.R

@@ -167,11 +167,11 @@ total_sum %>%
          `Tropical cyclone with most counties exposures (# exposed counties)` = max_storm) %>%
   xtable(align = "cp{1.75cm}p{3cm}p{4cm}p{4cm}",
          caption = "Summary statistics for exposure metrics considered to assess county-level exposure to
-         tropical cyclones. A county exposure represents the exposure of a specific eastern U.S.  
+         tropical cyclones. A county exposure represents the exposure of a specific eastern US  
          county to a specific tropical cyclone. A county can be exposed to multiple tropical cyclones within
          a year, and a tropical cyclone can create multiple county exposures under a given metric.
          The median and interquartile range of county exposures per tropical cyclone is based on
-         the tropical cyclones for which at least one U.S. county was exposed. The years for which
+         the tropical cyclones for which at least one US county was exposed. The years for which
          data are available for each metric are given in Table 1.",
          label = "tab:exposuresummaries",
          digits = 0) %>%

### 2019-07-27 writing/anderson_ehp_supp.tex
Work on draft throughout

--- 2019-01-27 writing/anderson_ehp_supp.tex
@@
 \centering
-\includegraphics[width=0.7\linewidth]{hurrtracks}
-\caption{Tracks of all tropical cyclones included in the study. The study
-included all tracked storms in the revised Atlantic hurricane database
-\ac{HURDAT2}~\citep{landsea2013} that came within 250~\si{\kilo\metre} of at least on
-\ac{US} county, 1988\,--\,2015. Thicker lines show the tracks of storms whose
-names were retired, indicating that the storm was particularly severe or had
-notable impacts~\citep{retirednames}.}
-\label{fig:hurrtracks}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
 \includegraphics[width=\linewidth]{othertopstorms}
@@
 each exposure metric, based on years of for which data for all five exposure
-metrics are available (1996\,--\,2011). The criteria behind each of the five
-metrics is given in Table 1 of the main text.} \label{fig:averageexposuresupp}
+metrics are available (1996\,--\,2011).} 
+\label{fig:averageexposuresupp}
 \end{figure}
@@
 \includegraphics[width=\linewidth]{windexamples}
-\caption{Wind classifications for the two storms, Hurricanes Sandy (2012) and
-Ike (2008) for which exposure classifications from modeled wind
-data~\citep{stormwindmodel} and \ac{HURDAT2} wind
-radii~\citep{demuth2006improvement} showed the highest disagreement for storms
-between 2004 and 2015 (Figure~\ref{fig:windcomparison}). The top two maps show
-modeled (left) and \ac{HURDAT2} (right) windspeed categorizations for Hurricane
-Ike (2008), while the bottom two maps show the same windspeed characterizations
-for Hurricane Sandy (2012).  In both cases, the storms were exceptionally large
-systems for which high winds persisted well inland from landfall and for which,
-based on the \ac{HURDAT2} data, a large number of counties were exposed to
-winds of $\ge$34 knots (Figure~\ref{fig:windcomparison}).}
+\caption{Comparison of county-level estimates of maximum sustained surface wind speeds
+	for Hurricane Ike in 2008 (top) and Hurriance Sandy in 2012 (bottom). 
+	Each map shows the estimated maximum sustained surface wind speed 
+	classification ($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot}; 
+	50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) for each study county. 
+	The maps labelled ``Modeled'' (left)
+	shows the classifications based on modeled sustained surface wind
+	speeds, which were included in the open-source data as the main 
+	wind metric and used in further analysis in this research. The maps
+	labelled ``Extended Best Tracks'' (right) show classifications based on the 
+	wind radii given in \ac{HURDAT2} (included as a secondary wind 
+	metric in the open-source data). The red lines show the storms' 
+	tracks.
+	}
 \label{fig:windexamples}
@@
 in that assessment metric. The red line on each map shows the track of the
-storm, based on the revised Atlantic hurricane database
-\ac{HURDAT2}~\citep{landsea2013}. The color of each county in the map gives the
-cumulative precipitation (in millimeters) for that county during the window
-from two days before to one day after the storm's closest approach to the
-county. The blue outline identifies the collection of counties that were
-classified as ``exposed" based on the criteria that the storm passed within
-500~\si{\kilo\metre} of the county's population mean center and the cumulative
-rainfall was 75 millimeters or more (Table 1 of main text). The map on the left
-shows rain-based exposure for Hurricane Ike. For this exceptionally large
-storm, there are some counties in northern Texas that received high rainfall
-from the storm but are too far from the storm center's track to be classified
-as exposed by this metric. The map on the right shows rain-based exposure for
-Tropical Storm Lee. For this storm, rain from the storm affected counties in
-the Mid-Atlantic and New England as the remnants of the storm moved over those
-regions, but storm tracking stopped while the storm was still over the
-southeastern part of the \ac{US}, and so these counties are not classified as
-exposed to the storm based on this metric.}
+storm. The color of each county in the map gives the cumulative precipitation
+(in \si{\milli\metre}). The blue outline identifies the collection of counties
+that were classified as ``exposed" based on our rainfall criteria (Table 1 of
+main text). }
 \label{fig:rainexamples}
@@
 
-\bibliography{hurr_exposure}
+\printbibliography
 

### 2019-07-27 writing/ehp_discussion.tex
Work on draft throughout

@@
 exposures could help improve preparedness for and response to future tropical
-cyclones.  However, tropical cyclones are multi-hazard events, making it more
-complicated to measure exposure compared to other weather exposures like
-temperature and precipitation. Here, we find the county-level tropical cyclone
-exposure assessments vary substantially when using a distance-based metric
-versus four hazard-based metrics, as well as among different hazard-based
-metrics. Our results can inform exposure assessment for future county-level
-studies of the health risks and impacts associated with tropical cyclones
-exposures, and the data and software we have created to accompany this article
-provides multi-hazard, county-level exposure data for multiple tropical cyclone
-hazards for such studies.
+cyclones.  However, tropical cyclones are multi-hazard events, making it
+complicated to measure exposure. Here, we provide open-source county-level data
+on multiple storm exposure metrics, and we find that county-level tropical
+cyclone exposure assessments vary substantially when using a distance-based
+metric versus four hazard-based metrics, as well as among different
+hazard-based metrics. Our results can inform exposure assessment for future
+county-level studies of the health risks and impacts associated with tropical
+cyclones exposures, and the open-source data and software we have created
+provides multi-hazard, county-level exposure data for such studies.
 
-Distance from a tropical cyclone's track is relatively easy to measure and has
-been used as an operational metric of exposure to tropical cyclones in previous
-large-scale studies (examples include ...). Since distance itself
-does not constitute a hazard, distance is meant in these cases as a surrogate
-to capture exposure to hazards from the tropical cyclone. However, here we
-found that in assessing \ac{US} county-level exposure to tropical cyclones,
-distance is, at best, a moderate, and often a very poor, surrogate for exposure
-to the specific tropical cyclone hazards of high wind, extreme rainfall,
-flooding, and tornadoes (Figure~\ref{fig:jaccard}).  Therefore, use of distance
-to assess tropical cyclone exposure for impact studies could result in
+\subsection*{Exposure assessment data and software}
+
+In creating this hurricane exposure dataset, we aimed for data that are
+available for all eastern \ac{US} counties. In addition to providing tropical
+cyclone exposure metrics for individual hazards, this data and its associated
+software allow users not only to access measurements for single hazards, but
+also to create tropical cyclone exposure profiles based on multiple hazards or
+craft exposure indices that combine hazard
+metrics~\parencite{chakraborty2005population, peduzzi2009assessing}. This
+ability can be critical, as different hazards of tropical cyclones often act
+synergistically in causing impacts~\parencite{smith2009}.
+
+\begin{comment}
+To assist with future tropical cyclone studies, we created and published open
+source software with this dataset and accompanying
+tools~\parencite{hurricaneexposuredata, hurricaneexposure}.  Many previous
+studies have used geographical information system software (e.g., ArcGIS) to
+assess exposure to tropical cyclones in the \ac{US}~\parencite{grabich2016,
+zandbergen2009, czajkowski2011, kruk2010}.  Here, we offer methods to map and
+output historic exposure to tropical cyclones that does not require the use of
+proprietary software but instead uses a package written in the R statistical
+programming language~\parencite{R}, which is free and open-source.  Further, by
+including measurements of different hazard exposures in each county for each
+tropical cyclone, this software allows for the development of more complex
+exposure indices or models (e.g., random forests, multivariable generalized
+linear models) that incorporate multiple tropical cyclone hazard measurements.  
+\end{comment}
+
+We investigated how well these data correspond with data from other available
+sources (Figures~\ref{fig:windcomparison}\,--\,\ref{fig:floodcomparison}
+and~S6\,--\,7). While generally in agreement with data from other sources,
+there are a few caveats that should be considered when applying these data in
+future studies. The rainfall data are from re-analysis data are generally
+well-correlated with observed ground-based station data, but may sometimes
+oversmooth very high rainfall values compared to ground-based observations
+(Figure~\ref{fig:raincomparison}). When rainfall data is used to create binary
+exposure classifications, this disagreement is unlikely to influence results,
+as both data sources agree in identifying these as storms with high rainfall,
+but it may be important to consider for cases that include rainfall as a
+continuous measurement.
+
+The wind data are based on modeled, rather than observed, values, and while the
+modeled wind data generally agree well with post-analysis maximum wind radii
+from \ac{HURDAT2}~\parencite{landsea2013} (Figure~\ref{fig:windcomparison}),
+there were a few storms where the two datasets did not agree as well ($<$90\%
+of counties classified into the same wind categories by the two sources of wind
+estimates). These storms were exceptionally large systems for which high winds
+persisted well inland from landfall, like Hurricanes Sandy and Ike
+(Figures~\ref{fig:windcomparison} and\~S6). Since the wind radii are based on
+\ldots, they have [advantages]; their main disadvantages is that they provide
+categorical, rather than continuous, estimates of wind speed, and in many cases
+researchers might prefer more precise estimates of storm wind speeds. However,
+since the wind radii data may be useful in some cases as either a primary
+metric or for sensitivity analysis, we have include county-level wind estimates
+for each storm based on these wind radii, in addition to the continuous modeled
+windspeeds, in the open-source data \cite{hurricaneexposuredata}.
+
+The flood and tornado data came from the \ac{NOAA} Storm Events database,
+which, while a widely-used database of events maintained by \ac{NOAA}, is based
+on reports, and so may be prone to underreporting~\parencite{Ashley2008flood,
+Curran2000}, especially in less populated areas~\parencite{Witt1998,
+Ashley2007}, as well as to other reporting errors. While these are important
+limitations, we selected these data sources as among the best currently
+available for measuring each of these hazards at a multi-county, multi-year
+scale.
+
+\subsection*{Patterns in tropical cyclone exposures}
+
+We found geographic patterns in tropical cyclones based on the different
+exposure metrics that. These patterns were not unexpected based on what is
+known about tropical cyclone hazards, but still highlight variations across
+metrics that are critical to consider in designing studies for tropical cyclone
+epidemiology, and they demonstrate the need for multi-hazard exposure datasets
+for tropical cyclone epidemiology. Wind-based exposures had a strong coastal
+pattern, which is consistent with the dramatic decrease in wind intensity that
+typically characterizes the landfall of tropical cyclones. Exposures by rain-
+and distance-based metrics often extended further inland compared to the
+wind-based metric, up to the Appalachian mountains. This agrees with previous
+research indicating that the Appalachian mountains' topography both enhances
+precipitation during tropical cyclones and provides hydrological conditions for
+severe flooding~\parencite{rees2001}.  Almost all tornado-based exposures were
+in southern coastal states, echoing previous findings that most noteworthy
+tropical cyclone-related tornadoes occur on the right-side of the tropical
+cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.  These
+exposure averages are somewhat limited as estimates of long-term frequencies,
+as tropical cyclones follow decadal patterns~\parencite{kossin2007more} likely
+not adequately captured in the available data. However, exposure assessments
+spanning over a century have found similar patterns for distance- and
+wind-based metrics~\textcite{zandbergen2009, kruk2010} [double-check that refs
+found similar patterns].
+
+\begin{comment}
+However, these frequency maps, together with evidence from
+specific tropical cyclones (Figures~\ref{fig:ivanexposure}
+and~\ref{fig:jaccard}),  do illustrate the potential for strong differences in
+spatial patterns in tropical cyclone exposures, depending on which tropical
+cyclone hazards are considered.  A few previous studies have sought to
+determine county-level exposure to tropical cyclones over multi-year periods,
+including~\textcite{zandbergen2009}, which estimated exposure in \ac{US}
+counties to all \ac{US} landfalling Atlantic-basin tropical cyclones
+between~1851 and~2003, using both a distance-based metric and a metric that
+combined distance and windspeed, and~\textcite{kruk2010}, which explored
+exposure to hurricane-related winds in the \ac{US}, including inland areas,
+for~1900\,--\,2008.  Our results suggest that such exposure assessments may
+perform well in capturing some tropical cyclone hazards (e.g., wind), but
+likely miss other potentially dangerous tropical cyclone exposures, especially
+for hazards that repeatedly threaten northern or inland counties (e.g., rain,
+flooding).
+\end{comment}
+
+\subsection*{Agreement between exposure metrics}
+
+We found that there was typically only moderate or low agreement between
+county-level exposure classifications for a tropical cyclone when comparing
+different exposure metrics. This suggests that the use of one of these metrics
+as a proxy for all storm-based hazards in epidemiological research would often
+be problematic, introducing exposure misclassification.
+
+\begin{comment}
+For example, distance from a tropical cyclone's track is relatively easy to
+measure and has been used as an operational metric of exposure to tropical
+cyclones in previous large-scale studies (examples include ...). Since distance
+itself does not constitute a hazard, distance is meant in these cases as a
+surrogate to capture exposure to hazards from the tropical cyclone. However,
+here we found that in assessing \ac{US} county-level exposure to tropical
+cyclones, distance is, at best, a moderate, and often a very poor, surrogate
+for exposure to the specific tropical cyclone hazards of high wind, extreme
+rainfall, flooding, and tornadoes (Figure~\ref{fig:jaccard}). Therefore, use of
+distance to assess tropical cyclone exposure for impact studies could result in
 problematic exposure misclassification, which could mask true associations,
 even strong associations, between tropical cyclone exposure and outcomes of
-interest in impact studies~\parencite{savitz2016interpreting, armstrong1998effect}.
+interest in impact studies~\parencite{savitz2016interpreting,
+armstrong1998effect}.  
+\end{comment}
 
-In some cases, this exposure misclassification may be differential (i.e.,
-associated with the outcome of interest or with factors associated with risk of
-the outcome of interest). For example, tropical cyclone wind exposures tend to
-be concentrated in counties near the coast, since most tropical cyclones
-rapidly decrease in sustained windspeed following landfall. However, tropical
-cyclone exposures based on distance can extend well inland, following the
-tropical cyclone's tracks, but may not adequately capture all wind-exposed
-counties near the coast. In this case, if the etiologically-relevant exposure
-is high wind but exposure is classified based on distance, the probability of
-being misclassified as unexposed would be higher in coastal counties, while the
-probability of being misclassified as exposed would be higher in inland
-counties. If coastal counties differ from inland counties in either the outcome
-of interest or in factors associated with risk of that outcome, differential
-exposure misclassification would exist~\parencite{savitz2016interpreting}. Such
-differential exposure misclassification can bias estimates of tropical cyclone
-effects either towards the null (estimating a lower or null association
-compared to the true association that exists) or away from the null (estimating
-a larger association than actually exists)~\parencite{savitz2016interpreting,
-armstrong1998effect}. The use of a single hazard-based metric (e.g., wind)
-could cause similar problems if the impact is driven, at least in part, by a
-different hazard or by multiple hazards of the tropical cyclone.
+These findings align with previous results from atmospheric science and related
+fields on the characteristics of tropical cyclones. While tropical cyclone
+rainfall and windspeed are often well-correlated when the tropical cyclone is
+over water~\parencite{cerveny2000}, this relationship often does not remain as
+strong once the hurricane has made landfall~\parencite{jiang2008}.  Fast-moving
+tropical cyclones bring higher risks of dangerous winds
+inland~\parencite{kruk2010}, while slow-moving tropical cyclone are likely to
+bring more rain~\parencite{rappaport2000} and cause more damage because of
+sustained hazardous conditions~\parencite{rezapour2014}. Further, while the
+likelihood and extent of flooding during a tropical cyclone is related to the
+tropical cyclone's rainfall, it is also driven by factors like top soil
+saturation and the structure of the water basin's drainage
+network~\parencite{chen2015, rees2001}.
 
-Our findings\,---\,that county-level tropical cyclone exposure assessment varies
-substantially when different exposure measurements are considered\,---\,aligns with
-previous research from atmospheric science and related fields on the
-characteristics of tropical cyclones. While tropical cyclone rainfall and
-windspeed are often well-correlated when the tropical cyclone is over
-water~\parencite{cerveny2000}, this relationship often does not remain as strong
-once the hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
-cyclones bring higher risks of dangerous winds inland~\parencite{kruk2010}, while
-slow-moving tropical cyclone are likely to bring more
-rain~\parencite{rappaport2000} and cause more damage because of sustained hazardous
-conditions~\parencite{rezapour2014}. Further, while the likelihood and extent of
-flooding during a tropical cyclone is related to the tropical cyclone's
-rainfall, it is also driven by factors like top soil saturation and the
-structure of the water basin's drainage network~\parencite{chen2015, rees2001}.
 However, we did find a small set of tropical cyclones for which for which
 agreement between metrics were unusually high (e.g., Floyd in~1999, Irene in
-2011, Hannah in~2008, Bertha in~1996) For this set of tropical cyclones, the
+2011, Hannah in~2008, Bertha in~1996).  Hurricanes Floyd in~1999 and Irene
+in~2011 both made their first \ac{US} landfall in North Carolina at minor
+hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
+eastern coast of the \ac{US} north through New England, bringing large rainfall
+to much of the eastern coast from North Carolina north and causing extensive
+inland flooding in North Carolina (Floyd) and New England
+(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
+in~2008 and Bertha in~1996 followed the eastern coastline north of North
+Carolina in a pattern similar to Floyd and Irene. For this set of storms, the
 tropical cyclones' persistent proximity to water may have helped maintain wind
@@
 
-In our results, wind-based exposure had a strong coastal pattern, with almost
-all exposures in counties within about~200~\si{\kilo\metre} (124~mi) of the
-coastline. This is consistent with the dramatic decrease in wind intensity that
-typically characterizes the landfall of tropical cyclones. Exposures by rain-
-and distance-based metrics often extended further inland compared to the
-wind-based metric, up to the Appalachian mountains. This agrees with previous
-research indicating that the Appalachian mountains' topography both enhances
-precipitation during tropical cyclones and provides hydrological conditions for
-severe flooding~\parencite{rees2001}.  Almost all tornado-based exposures were in
-southern coastal states. This pattern echoes previous findings that most
-noteworthy tropical cyclone-related tornadoes occur on the right-side of the
-tropical cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.
+In some cases, exposure misclassification from using one of these exposure
+metrics to capture one or more of the other storm hazards may be differential
+(i.e., associated with the outcome of interest or with factors associated with
+risk of the outcome of interest). For example, tropical cyclone wind exposures
+tend to be concentrated in counties near the coast, since most tropical
+cyclones rapidly decrease in their sustained windspeed following landfall.
+However, tropical cyclone exposures based on distance can extend well inland,
+following the tropical cyclone's tracks, but may not adequately capture all
+wind-exposed counties near the coast. In this case, if the
+etiologically-relevant exposure is high wind but exposure is classified based
+on distance, the probability of being misclassified as unexposed would be
+higher in coastal counties, while the probability of being misclassified as
+exposed would be higher in inland counties. If coastal counties differ from
+inland counties in either the outcome of interest or in factors associated with
+risk of that outcome, differential exposure misclassification would
+exist~\parencite{savitz2016interpreting}. Such differential exposure
+misclassification can bias estimates of tropical cyclone effects either towards
+the null (estimating a lower or null association compared to the true
+association that exists) or away from the null (estimating a larger association
+than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.
 
-The tropical cyclone exposure averages shown in
-Figure~\ref{fig:averageexposure} are limited as estimates of long-term
-frequencies, as tropical cyclones follow decadal
-patterns~\parencite{kossin2007more} likely not adequately captured in the available
-data.  However, these frequency maps, together with evidence from specific
-tropical cyclones (Figures~\ref{fig:ivanexposure} and~\ref{fig:jaccard}),  do
-illustrate the potential for strong differences in spatial patterns in tropical
-cyclone exposures, depending on which tropical cyclone hazards are considered.
-A few previous studies have sought to determine county-level exposure to
-tropical cyclones over multi-year periods, including~\textcite{zandbergen2009},
-which estimated exposure in \ac{US} counties to all \ac{US} landfalling
-Atlantic-basin tropical cyclones between~1851 and~2003, using both a
-distance-based metric and a metric that combined distance and windspeed,
-and~\textcite{kruk2010}, which explored exposure to hurricane-related winds in the
-\ac{US}, including inland areas, for~1900\,--\,2008.  Our results suggest that
-such exposure assessments may perform well in capturing some tropical cyclone
-hazards (e.g., wind), but likely miss other potentially dangerous tropical
-cyclone exposures, especially for hazards that repeatedly threaten northern or
-inland counties (e.g., rain, flooding).
+\begin{comment}
+The use of a single hazard-based metric (e.g., wind) could cause similar
+problems if the impact is driven, at least in part, by a different hazard or by
+multiple hazards of the tropical cyclone.
+\end{comment}
+
+\subsection*{Case study}
 
@@
 in differing patterns in estimates of county-level exposure
-(Figure~1\todo{change to fig. ref}), which can lead to substantial differences
+(Figure~\ref{fig:ivanexposure}), which can lead to substantial differences
 in estimates of population physical exposure
@@
 
-\subsection*{Software}
-
-To assist with future tropical cyclone studies, we created and published open
-source software to accompany this article~\parencite{hurricaneexposuredata,
-hurricaneexposure}. Many previous studies have used geographical
-information system software (e.g., ArcGIS) to assess exposure to tropical
-cyclones in the \ac{US}~\parencite{grabich2016, zandbergen2009, czajkowski2011,
-kruk2010}. Here, we offer methods to map and output historic exposure to
-tropical cyclones that does not require the use of proprietary software but
-instead uses a package written in the R statistical programming
-language~\parencite{R}, which is free and open-source.
-
-In addition to providing tropical cyclone exposure metrics for individual
-hazards, this software and data also allow users to create tropical cyclone
-exposure profiles based on multiple hazards or craft exposure indices that
-combine hazard metrics~\parencite{chakraborty2005population, peduzzi2009assessing}.
-This ability can be critical, as different hazards of tropical cyclones
-often act synergistically in causing impacts~\parencite{smith2009}. Further, by
-including measurements of different hazard exposures in each county for each
-tropical cyclone, this software allows for the development of more complex
-exposure indices or models (e.g., random forests, multivariable generalized
-linear models) that incorporate multiple tropical cyclone hazard measurements.
-
-In creating this hurricane exposure dataset, we aimed for data that are
-reliable, available for all eastern \ac{US} counties, and have no missing data
-within available dates. Further, we have investigated how well these data
-correspond with data from other available sources
-(Figures~S7\,--\,11\todo{Check fig refs}). While generally in agreement with
-data from other sources, there are a few limitations. The wind data are based
-on modeled, rather than observed, values, and while the modeled wind data
-generally agree well with post-analysis maximum wind radii from
-\ac{HURDAT2}~\parencite{landsea2013} (Figure~S7\todo{Check fig refs}), for a
-few tropical cyclones it did not (e.g., Figure~S10\todo{Check fig refs}). The
-rainfall data are from re-analysis data, which are generally well-correlated
-with observed ground-based station data but sometimes oversmooth extreme
-measurements (Figure~S8\todo{Check fig refs}).  The flood and tornado data came
-from the \ac{NOAA} Storm Events database, which, while a widely-used database
-of events maintained by \ac{NOAA}, is based on reports, and so may be prone to
-underreporting~\parencite{Ashley2008flood, Curran2000}, especially in less
-populated areas~\parencite{Witt1998, Ashley2007}, as well as to other reporting
-errors. While these are important limitations, we selected these data sources
-as among the best currently available for measuring each of these hazards at a
-multi-county, multi-year scale.
-
 \subsection*{Conclusions}

### 2019-07-27 writing/ehp_figures.tex
Work on draft throughout

--- 2019-01-26 writing/ehp_figures.tex
@@
 
+\begin{figure}[tbhp!] \centering
+\includegraphics[width=0.7\linewidth]{hurrtracks}
+\caption{States and storms considered in this study. All counties in the states
+        shown in this map were investigated. The lines show the paths
+	of the study storms, which included all tracked storms in 1988\,--\,2015
+	\ac{HURDAT2}~\parencite{landsea2013} that came within 250~\si{\kilo\metre} 
+	of at least one \ac{US} county. Thicker lines show the tracks 
+	of storms whose names have been retired, indicating that the storm was particularly 
+	severe or had notable impacts~\parencite{retirednames}.
+	}
+\label{fig:hurrtracks}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[tbhp!] \centering
+\includegraphics[width=0.9\linewidth]{raincomparison} 
+	\caption{Comparison of storm-associated rainfall estimates between 
+	a re-analysis dataset (included as the rainfall exposure metric
+	in the open-source data and included in further analysis) and ground-based
+	observations in nine sample counties. The rainfall estimates  
+	include rainfall from two days before to one day after the storm's closest
+	approach to the county. Each small graph shows data for one sample county, 
+	and each point shows one tropical storm. The number of storms 
+	within each county and the average number of stations reporting rainfall during 
+	the county's storms are given above each plot. Horizontal and vertical lines 
+	in each plot show the threshold of 75~\si{\milli\metre} used to classify a storm 
+	as ``exposed'' in further analysis (Table 1). Note that 
+	x- and y-axis ranges differ by county.
+	} 
+\label{fig:raincomparison}
+\end{figure}
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=0.6\linewidth]{windcomparison}
+\caption{Comparison of modeled maximum sustained surface wind speeds 
+	(included as the primary wind exposure metric in the open-source
+	data and used in further analysis) with estimates from the \ac{HURDAT2}
+	wind speed radii. Each point represents a storm, with the x-axis
+	giving the percent of counties classified in the same category
+	of maximum sustained surface wind speeds ($<$34~\si{\knot}; 
+	34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; 
+	$\ge$64~\si{\knot}) by both sets of wind data for that storm. The 
+	color of each point indicates the extent of the storm, 
+	giving the number of study counties that were exposed to 
+	maximum sustained surface wind speeds of at least 34~\si{\knot}
+	(based on modeled wind speeds). Estimates are shown for [all?]
+	study storms since 2004, the earliest year for which 
+	wind speed radii estimates are available in \ac{HURDAT2} [?].
+	}
+\label{fig:windcomparison}
+\end{figure}
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=0.8\linewidth]{floodcomparison}
+\caption{Comparison for a sample of study counties of NOAA Storm Events
+listings versus five-day total streamflows at county streamflow gages during
+tropical storms. The number of streamflow gages used in analysis are given in
+parentheses beside the county's name in the panel title. Each point represents
+a single tropical storm. Note that the x-axis scales differ by county,
+depending on the number of streamflow gages and typical flow rates for each
+gage, and are on a log-10 scale.}
+\label{fig:floodcomparison}
+\end{figure}
+
+\clearpage
+
 \begin{figure}%[tbhp] 
 \centering
-\includegraphics[width=16cm]{ivanonly}
-\caption{Counties classified as exposed to Hurricane Ivan (2004) under each
-exposure metric (Table 1). The red line shows the track of Hurricane Ivan 
-based on the revised Atlantic hurricane database (HURDAT2 \cite{landsea2013}).
-Similar maps for other large-extent storms are given in Fig. S3.}
+\includegraphics[width=15cm]{averageexposureonly.pdf} 
+\caption{Average number of storm exposures per decade in U.S. counties for each
+exposure metric. The criteria behind each of the five metrics is given in Table
+\ref{tab:exposuremetrics}. The years used to estimate these averages are based
+on years of available exposure data (distance and wind:~1988\,--\,2015; rain:
+1988\,--\,2011; flood and tornado:~1996\,--\,2015). } 
+\label{fig:averageexposure} 
+\end{figure}
+
+\clearpage
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=16cm]{ivanonly.pdf}
+\caption{Counties classified as exposed to Hurricane Ivan in 2004 under each
+exposure metric considered (Table~\ref{tab:exposuremetrics}). The red line
+shows the track of Hurricane Ivan based on the revised Atlantic hurricane
+database \ac{HURDAT2}~\parencite{landsea2013}.  Similar maps for other
+large-extent storms are given in Figure~S3.}
 \label{fig:ivanexposure} 
@@
 \centering 
-\includegraphics[width = 0.6\linewidth]{jaccard_heatmap} 
-\caption{Heatmap of Jaccard index values for
-specific exposure metric pairs within storms. Only storms between 1996 and 2011,
-and for which at least 250 counties were exposed based on at least one metric,
-are included. The color of each cell within the main heatmap indicates the value
-of the Jaccard index (proportion of counties classified as exposed by both
-metrics out of storms classified as exposed by either metric) for a given pair
-of metrics for a given storm. Storms are displayed within clusters that have
-similar patterns in county-level exposure agreement for metric pairs, based on
-hierarchical clustering using the complete link method
-\cite{murtagh2012algorithms} (i.e., storms in the same cluster tend to have
-similar patterns for the pairwise strength of agreement among metrics); columns
-are also ordered based on hierarchical clustering. The colors to the right of
-the main heatmap for each storm indicate the total number of counties classified
-as exposed to the storm by any of the five metrics, providing an estimate of
-storm extent. Maps are available showing the counties identified as exposed
-under each of five metrics for the widest-extent storm in each cluster:
-Hurricane Ivan (2004) (Fig. \ref{fig:ivanexposure}) and Hurricanes Floyd (1999),
-Lee (2011), Cindy (2005), and Katrina (2005) (Fig. S3).} 
+\includegraphics[width = 0.6\linewidth]{jaccard_heatmap.pdf} 
+\caption{Agreement between exposure classifications based on different
+         exposure metrics for all storms between~1996 and~2011 for which 
+	 at least~250 counties were exposed based on at least one metric.
+	 Each row shows one storm, and the color of each cell shows the 
+	 measured Jaccard index for each pair of exposure metrics 
+	 (proportion of counties classified as exposed by both metrics out 
+	 of storms classified as exposed by either metric). The colors to the 
+	 right of the main heatmap for each storm indicate the total number of 
+	 counties classified as exposed to the storm by any of the five metrics, 
+	 providing an indication of each storm's extent. Storms are displayed 
+	 within clusters that have similar patterns in county-level exposure 
+	 agreement for metric pairs, based on hierarchical clustering using the 
+	 complete link method~\parencite{murtagh2012algorithms}; columns are also 
+	 ordered based on hierarchical clustering. Maps are available showing the 
+	 counties identified as exposed under each of five metrics for the widest-extent 
+	 storm in each cluster: Hurricane Ivan in~2004 (Figure~\ref{fig:ivanexposure}) 
+	 and Hurricanes Floyd in~1999, Lee in~2011, Cindy in~2005, and Katrina 
+	 in~2005 (Figure~S3).
+} 
 \label{fig:jaccard}
@@
 
-\begin{figure}%[tbhp] 
+\begin{figure*}%[tbhp]
 \centering
-\includegraphics[width=15cm]{averageexposureonly} 
-\caption{Average number of storm exposures per decade in U.S. counties for 
-each exposure metric. The criteria behind each of the five metrics is given 
-in Table \ref{tab:exposuremetrics}. The years used to estimate these averages 
-are based on years of available exposure data (distance and wind: 1988--2015; 
-rain: 1988--2011; flood and tornado: 1996--2015). Similar patterns persist when
-analysis is restricted to years with all exposure data available (1996--2011;
-Fig. S5).} 
-\label{fig:averageexposure} 
-\end{figure}
+\includegraphics[width=15.5cm]{topelecdependexposure}
+\caption{Counties with the highest estimated physical exposure among
+electricity-dependent Medicare beneficiaries exposed to storms per year in U.S.
+counties for each exposure metric. The color of each bar indicates the number
+of Medicare beneficiaries in the county reliant on electricity-dependent
+medical and assistive equipment as of July~2017. The length
+of each bar shows the average expected number of these electricity-dependent
+Medicare beneficiaries exposed to tropical storms per year based on a given
+exposure metric.}
+\label{fig:topelecdependexposure}
+\end{figure*}

### 2019-07-27 writing/ehp_introduction.tex
Work on draft throughout

@@
 
-Tropical cyclones, including hurricanes, can have disastrous impacts in the
-\acf{US}. A tropical cyclone's high winds can bring health risks
-and property damage through structural damage of houses and other buildings,
-falling trees, and wind-borne debris~\parencite{rappaport2000} and cause power
-outages~\parencite{liu2005, han2009}, which introduce a number of threats to human
-and ecological health, including water quality risks if the outage affects
-wastewater treatment plants~\parencite{mallin2006}.  Other risks can exist without
-severe wind; for example, one study found that most of the direct
-hurricane-related deaths in the \ac{US} between~1970 and~1999 occurred in cases
-when wind was below hurricane strength, including for Tropical Storms Charley
-in~1998 and Alberto in~1994~\parencite{rappaport2000}.  Tropical cyclones can
-produce excessive rain, especially in certain topographies (e.g., near
-mountains), so counties well inland sometimes experience more extreme rain than
-coastal counties. Flood risks from tropical cyclones can result from this rain,
-although the two risks are not perfectly correlated~\parencite{chen2015}. In the
-\ac{US}, over half of hurricane-related direct deaths from~1970 to~1999 from
-Atlantic basin storms were a result of freshwater flooding~\parencite{rappaport2000}. 
-Flooding can also degrade water quality~\parencite{mallin2006}, which can threaten both 
-human and ecological health.
+\acresetall
 
-To measure tropical cyclone exposure for either investigations of exposure
-patterns or for estimation of tropical cyclone risks and impacts, including
-those related to human health, studies have varied widely in how they determine
-tropical cyclone exposure (some examples shown in Figure~S1\todo{Check
-fig.ref}). While some studies have measured exposure based on specific hazards
-of the tropical cyclone (e.g., wind, rain), many have used distance from the
-tropical cyclone's central track as a surrogate in identifying exposure to the
-hazards of a tropical cyclone (e.g.,~\textcite{czajkowski2011, tansel2010,
-kinney2008, caillouet2008increase}).
+Recently, several major tropical cyclones have hit \ac{US} cities, including
+Hurricanes Harvey, Irma, and Maria in 2017 and Hurricanes Florence and Michael
+in 2018. Researchers are currently exploring how these storms impacted human
+health (e.g., [Maria refs, others?]), including through projects funded through
+NIH Rapid Response grants [NIH RePorter ref].  Evidence from this research will
+add to evidence from previous hurricanes, including Sandy in 2012 [refs] and
+Katrina in 2005 [refs], which found these storms can create substantial risk
+not only for accidental health outcomes (e.g., drowning, crushing deaths and
+injuries, electrocution, carbon monoxide poisoning) but also for
+non-accidental, non-communicable diseases [refs]. However, much of the
+epidemiological research to date on the health impacts of tropical cyclones has
+focused on a single storm (e.g., [refs]), or has relied on death and injury
+counts for which a link to the storm was identifiable through mention on a
+death certificate or other information about the death [refs], an approach that
+may miss many of the health impacts of storms, particularly from non-accidental
+causes, as evidenced with Hurricane Maria [ref for this?].
 
-Distance from the tropical cyclone's central track can be easily measured using
-widely-available hurricane tracking data. There are, however, a number of
-limitations to using distance to assess exposure to tropical cyclones, and
-there are reasons to suspect that exposure classifications based on other
-tropical cyclone hazards may disagree with distance-based classifications.
+A recent review highlights the need to move beyond single-storm and descriptive
+analyses to larger-scale studies with consistently [defined?] exposure and
+outcome assessment [ref].  Key to expanding tropical cyclone epidemiology are
+exposure assessment methods that appropriately capture those hazards of the
+storm that are relevant to human health, methods that can be applied
+consistently across multiple storms, locations, and years.
+
+The \ac{NHC}'s ``Best Tracks'' dataset is considered the gold standard in
+recording the location and central intensity of Atlantic-basin tropical
+cyclones. This data is provided through the \ac{HURDAT2} dataset, a post-storm
+assessment conducted by the \ac{US} \ac{NHC} that incorporates data from a
+variety of sources, including satellite data and, when available, aircraft
+reconnaissance data~\parencite{landsea2013, jarvinen1988}. This dataset records
+the central position of a storm every six hours (at the synoptic times of
+6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}) as well as measures of the
+minimum pressure [at the storm's center?] and the storm's central maximum
+sustained winds (typically located in the eye wall [?] if the storm has
+developed an eye).  From the ``Best Tracks'' data, it is straightforward to
+measure whether a study location was in the direct path, or within a certain
+distance of the path, of a storm using geographical software. Some studies have
+used distance to a storm's track either in assessing average exposure patterns
+to tropical storms [refs] or for assigning exposure for health studies [refs].
+
+However, the hazards that tropical cyclones bring that threaten human health
+are not perfectly aligned with distance from the storm's track.  Tropical
+cyclones are multi-hazard events. Recent storms have shown how much storms can
+differ in intensity of these hazards; Hurricanes Michael and Maria demonstrated
+the threats of extreme storm-related winds, for example, while Hurricanes
+Florence and Harvey were more notable for extreme rain and flooding [refs].
+Depending on the health outcome considered, some storm hazards might be more
+plausible as a risk factor than others. Respiratory outcomes in the months
+after a storm, for example, might be increased through a pathway of heightened
+mold exposure from flooding, while heat stroke and other heat-related outcomes
+in the days immediately after a storm might follow from a pathway of extreme
+winds causing power outages, resulting in loss of air conditioning [refs].
+Geographical patterns can differ for these different storm hazards, and so
+exposure assessment based on one hazard could create exposure misclassification
+if the pathway for health risks is through a different storm hazard.
+
+% Tropical cyclones, including hurricanes, can have disastrous impacts in the
+% \ac{US}. A tropical cyclone's high winds can bring health risks
+% and property damage through structural damage of houses and other buildings,
+% falling trees, and wind-borne debris~\parencite{rappaport2000} and cause power
+% outages~\parencite{liu2005, han2009}, which introduce a number of threats to human
+% and ecological health, including water quality risks if the outage affects
+% wastewater treatment plants~\parencite{mallin2006}.  Other risks can exist without
+% severe wind; for example, one study found that most of the direct
+% hurricane-related deaths in the \ac{US} between~1970 and~1999 occurred in cases
+% when wind was below hurricane strength, including for Tropical Storms Charley
+% in~1998 and Alberto in~1994~\parencite{rappaport2000}.  Tropical cyclones can
+% produce excessive rain, especially in certain topographies (e.g., near
+% mountains), so counties well inland sometimes experience more extreme rain than
+% coastal counties. Flood risks from tropical cyclones can result from this rain,
+% although the two risks are not perfectly correlated~\parencite{chen2015}. In the
+% \ac{US}, over half of hurricane-related direct deaths from~1970 to~1999 from
+% Atlantic basin storms were a result of freshwater flooding~\parencite{rappaport2000}. 
+% Flooding can also degrade water quality~\parencite{mallin2006}, which can threaten both 
+% human and ecological health.
+
+% To measure tropical cyclone exposure for either investigations of exposure
+% patterns or for estimation of tropical cyclone risks and impacts, including
+% those related to human health, studies have varied widely in how they determine
+% tropical cyclone exposure (some examples shown in Figure~S1). While some
+% studies have measured exposure based on specific hazards of the tropical
+% cyclone (e.g., wind, rain), many have used distance from the tropical cyclone's
+% central track as a surrogate in identifying exposure to the hazards of a
+% tropical cyclone (e.g.,~\textcite{czajkowski2011, tansel2010, kinney2008,
+% caillouet2008increase}).
+
+% Distance from the tropical cyclone's central track can be easily measured using
+% widely-available hurricane tracking data. There are, however, a number of
+% limitations to using distance to assess exposure to tropical cyclones, and
+
+% Geographical patterns likely differ for these different storm hazards, and so
+% exposure assessment based on one hazard could create exposure misclassification
+% if the pathway for health risks is through a different storm hazard.
+
+There are reasons to suspect that exposure classifications based on tropical
+cyclone hazards like wind and rain may disagree with distance-based
+classifications, based only on distance from the storm's path. Further, an
+assessment based on one storm hazard (e.g., wind) might disagree with assessments based on
+other hazards (e.g., rain or flooding).
 Tropical cyclones vary dramatically in size: \ac{US} tropical cyclones have
-been observed with radii to maximum winds as small as~20~\si{\kilo\metre} and as
-large as~200~\si{\kilo\metre}~\parencite{mallin2006, quiring2011variations}. While a
-number of tropical cyclone hazards are strongly associated with distance from
-the tropical cyclone's center (e.g., wind and, at the coast, storm surge and
-waves~\parencite{rappaport2000, kruk2010}), other hazards like dangerous rain,
-floods, and tornadoes can occur well away from the tropical cyclone's central
-track~\parencite{rappaport2000, atallah2007, moore2012}.  For example, fatal
-tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
-between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
-tropical cyclone's center~\parencite{moore2012}.
+been observed with radii to maximum winds as small as~20~\si{\kilo\metre} and
+as large as~200~\si{\kilo\metre}~\parencite{mallin2006, quiring2011variations}.
+While a number of tropical cyclone hazards are strongly associated with
+distance from the tropical cyclone's center (e.g., wind and, at the coast,
+storm surge and waves~\parencite{rappaport2000, kruk2010}), other hazards like
+dangerous rain, floods, and tornadoes can occur well away from the tropical
+cyclone's central track~\parencite{rappaport2000, atallah2007, moore2012}.  For
+example, fatal tropical cyclone tornadoes, which were linked to over~300 deaths
+in the \ac{US} between~1995 and~2009, most often
+occur~200\,--\,500~\si{\kilo\metre} from the tropical cyclone's
+center~\parencite{moore2012}.
 
-Further, distance-based exposure metrics that use
-buffers tend to use an equal buffer distance on each side of the tropical
-cyclone track (e.g.,~\cite{czajkowski2011, grabich2015, grabich2016,
-zandbergen2009, tansel2010}), but the forces of a tropical cyclone tend to be
-distributed around the center in a non-symmetrical way. For example, extreme
-winds are more common to the right of the track, where counter-clockwise
-cyclonic winds move in concert with the tropical cyclone's forward 
-motion~\parencite{halverson2015}, and the fatal tornadoes associated with \ac{US} tropical
-cyclones between~1995 and~2009 occurred almost exclusively to the right of the
-tropical cyclone's track, mostly in the right front quadrant of the tropical
-cyclone~\parencite{moore2012}. Rain, conversely, is often heaviest to the left of
-the tropical cyclone's track, especially when the tropical cyclone interacts
-with other weather systems~\parencite{atallah2003, atallah2007, zhu2013variations}
-or undergoes an extratropical transition~\parencite{elsberry2002}.
+Further, studies that have used distance-based exposure metrics have tended to
+use an equal buffer distance on each side of the tropical cyclone track
+(e.g.,~\cite{czajkowski2011, grabich2015, grabich2016, zandbergen2009,
+tansel2010}), but the forces of a tropical cyclone tend to be distributed
+around the center in a non-symmetrical way. For example, extreme winds are more
+common to the right of the track, where counter-clockwise cyclonic winds move
+in concert with the tropical cyclone's forward
+motion~\parencite{halverson2015}, and the fatal tornadoes associated with
+\ac{US} tropical cyclones between~1995 and~2009 occurred almost exclusively to
+the right of the tropical cyclone's track, mostly in the right front quadrant
+of the tropical cyclone~\parencite{moore2012}. Rain, conversely, is often
+heaviest to the left of the tropical cyclone's track, especially when the
+tropical cyclone interacts with other weather systems~\parencite{atallah2003,
+atallah2007, zhu2013variations} or undergoes an extratropical
+transition~\parencite{elsberry2002}.
 
-\todo[inline]{Something about difference in other metrics}
-
-The use of different exposure metrics across studies hampers the development of
-a scientific consensus on tropical cyclone-related risks, as differences
-observed between studies could result from differences in exposure assessment.
-Further, if a study misclassifies exposure to the hazard or hazards that cause
-an impact, estimates of tropical cyclone risks or impacts will be biased,
-making it hard to identify true associations. These concerns are particularly
+If a study misclassifies exposure to the hazard or hazards that cause an
+impact, estimates of tropical cyclone risks or impacts will be biased, making
+it hard to identify true associations. The use of different exposure metrics
+across studies hampers the development of a scientific consensus on tropical
+cyclone-related risks, as differences observed between studies could result
+from differences in exposure assessment.  These concerns are particularly
 important if there are large differences in the locations that are determined
@@
 
-To investigate this, we measured county-level exposure to tropical cyclones in
-all eastern \ac{US} counties using five metrics
+To investigate this, and to provide open-source data on tropical cyclone exposures
+to other researchers, we developed a dataset of county-level exposure to
+tropical cyclones in all eastern \ac{US} counties for five different metrics
 (Table~\ref{tab:exposuremetrics}): (1)~distance to tropical cyclone track;
@@
 outcomes~\parencite{grabich2015, grabich2016}; autism
-prevalence~\parencite{kinney2008}) and since decisions and policies to prepare for,
-and respond to, tropical cyclones are often undertaken at the county
+prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
+for, and respond to, tropical cyclones are often undertaken at the county
 level~\parencite{zandbergen2009, rappaport2000}.  We explored patterns in these

### 2019-07-27 writing/ehp_methods.tex
Work on draft throughout

@@
 
-All counties in the eastern half of the \ac{US}
-(Figure~\ref{fig:ivanexposure}), as of the~2010 \ac{US} Decennial Census, were
-included in the study. The study covered all tracked storms in the \ac{HURDAT2}
+This study covered all counties in the eastern half of the \ac{US}
+(Figure~\ref{fig:hurrtracks}), as of the~2010 \ac{US} Decennial Census. The
+study covered all tracked land-falling or near-land Atlantic-basin tropical
+cyclones between~1988 and~2015, considering all 136 storms in \ac{HURDAT2}
 \parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
-study county between~1988 and~2015 (Figure~S2\todo{Check fig ref}).
-\ac{HURDAT2} is a post-storm assessment conducted by the \ac{US} \ac{NHC} and
-incorporates data from a variety of sources, including satellite data and, when
-available, aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.
-These data typically give measurements of the tropical cyclone center's
-location at~6-\si{\hour} intervals at synoptic times (i.e., 6:00~am, 12:00~pm,
-6:00~pm, and 12:00~am \ac{UTC}); some landfalling tropical cyclones have an
-additional observation at the time of landfall~\parencite{landsea2013}.
+eastern \ac{US} county in this period (Figure~\ref{fig:hurrtracks}).
+
+\begin{comment}
+These data typically give measurements of
+the tropical cyclone center's location [and central pressure / maximum
+windspeed?] at~6-\si{\hour} intervals at synoptic times (i.e., 6:00~am,
+12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}); some landfalling tropical cyclones
+have an additional observation at the time of landfall~\parencite{landsea2013}.
+\end{comment}
 
@@
 
-For each storm, the distance of closest approach was measured for every study
-county.  Storm center track location data from \ac{HURDAT2} were interpolated
-using natural cubic splines to~15-\si{\minute} intervals~\parencite{hurricaneexposure}.
-The distance between each county's population mean center, as assessed by the
-2010 Decennial Census~\parencite{countycenters}, and each~15-\si{\minute} interpolated
-storm center location was measured using Great Circle (WGS84 ellipsoid)
-distance~\parencite{bivand2013applied}. The distance-based exposure metric was
-based on the minimum value of this storm center-to-county distance throughout
-the period the storm was tracked. The time of closest approach for each
-county-storm pairing was also recorded for use assessing rain-, flood-, and
-tornado-based metrics. To allow matching with data based on local time, times
-were converted from \ac{UTC} to local time~\parencite{countytimezones}.
+We first measured how close each study storm came to each study county. We
+started with tracking data from \ac{HURDAT2}, which records the storm's
+position every~6 \si{\hour}, and interpolated to~15-\si{\minute} intervals
+using natural cubic splines~\parencite{hurricaneexposure}. We then measured the
+distance between the storm's position at each~15-\si{\minute} interval and each
+county's population mean center, as assessed by the 2010 Decennial
+Census~\parencite{countycenters}, using the Great Circle (WGS84 ellipsoid)
+method~\parencite{bivand2013applied}. These measurements captured the distance
+between the storm's central track and each study county throughout the period
+when the storm was tracked. The closest distance between the storm and each
+county was recorded as the distance of the storm's closest approach to the
+county. We also recorded the time storm came closest to each county, for use in
+assessing rain-, flood-, and tornado-based metrics where observed data on the
+hazards needed to be matched by time to storm events.  To allow matching with
+data based on local time, times were converted from \ac{UTC} to local
+time~\parencite{countytimezones}.
 
@@
 
-For the rain-based exposure metric, cumulative rainfall was estimated for the
-period from two days before to one day after the date of a tropical cyclone's
-closest approach to a county. Daily county-level rainfall was determined by
-aggregating hourly~1/8\si{\degree} gridded reanalysis data from the \ac{NLDAS-2}
-precipitation data files~\parencite{rui2013nldas}. The \ac{NLDAS-2} data integrate
+To estimate storm-associated rainfall, we used precipitation data from a
+re-analysis dataset, as this data was available for every county and every day
+through a continuous period of the study
+(1988\,--\,2011)~\parencite{alhamdan2014environmental, cdcwonder}.  By
+comparison, observations from ground-based monitoring networks had missing
+values spatially (i.e., for some study counties), temporally (for some days),
+or both.
+
+We estimated daily rainfall for each study county by aggregating
+hourly~1/8\si{\degree} gridded reanalysis data from the \ac{NLDAS-2}
+precipitation data files~\parencite{rui2013nldas}. These data integrate
 satellite-based and land-based monitoring, applying a land-surface model to
@@
 the continental \ac{US}~\parencite{rui2013nldas, alhamdan2014environmental}. To
-aggregate to county-level, hourly data at each grid point were summed to create
-a daily rainfall total, and these grid point rainfall totals were then averaged
-across all grid points within a county's~1990 \ac{US} Census boundaries to
-create a county-level daily total~\parencite{alhamdan2014environmental, cdcwonder}.
-This county-level precipitation data are publicly available through the \ac{US}
-\ac{CDC}'s \ac{WONDER} database~\parencite{cdcwonder, alhamdan2014environmental}.
-The rainfall-based exposure metric was only available for tropical cyclones
-through~2011, the period for which the county-aggregated reanalysis rainfall
-data were available~\parencite{cdcwonder, alhamdan2014environmental}.
+aggregate to county-level, the hourly data at each grid point were summed to
+create a daily rainfall total, and these grid point rainfall totals were then
+averaged across all grid points within a county's~1990 \ac{US} Census
+boundaries~\parencite{alhamdan2014environmental, cdcwonder}. These daily
+county-level estimates were then matched temporally with storm tracks, using
+the date when each storm was closest to each county. The cumulative
+storm-associated rainfall was then calculated as the sum of rainfall from two
+days before the storm's closest approach to a county to one day after.
 
-To assess the rainfall estimates from the reanalysis data used in analysis in
-the the main text, we compared cumulative rainfall estimates based on this
-reanalysis rainfall data to rainfall observations from ground-based weather
-stations in nine sample study counties, selected as counties with data
-available from multiple stations. In these nine counties, observed rainfall was
-collected for all storms,~1988\,--\,2011, for the same period of two days
-before to one day after each storm's closest approach, using all stations with
-available daily data in the Global Historical Climatology Network throughout
-1988\,--\,2011~\parencite{menne2012overview, rnoaa, countyweather}. Agreement
-between measured storm-related rainfall between these two data sources was
-assessed by Spearman's rank correlation.
+To validate these rainfall estimates, we investigated a subset of counties that
+also had available ground-based precipitation observations. We selected nine
+sample counties geographically spread across storm-prone regions of the eastern
+\ac{US} and for which precipitation data were available from multiple stations
+throughout the study period: Miami-Dade, FL; Harris County, TX; Mobile County,
+AL; Orleans Parish, LA; Fulton County, GA; Charleston County, SC; Wake County,
+NC; Baltimore County, MD; and Philadelphia County, PA. We collected
+ground-based observations of precipitation data from all stations in the county
+with available daily data in the Global Historical Climatology Network
+throughout 1988\,--\,2011~\parencite{menne2012overview, rnoaa, countyweather}.
+We summed daily station-specific measurments for two days before to one day
+after each storm's closest approach and then averaged these cumulative
+station-based precipitation totals for each county as a county-wide estimate of
+cumulative storm-related precipitation. We measured the rank correlation
+(Spearman's~$\rho$~\parencite{spearman1904proof}) between storm-specific
+cumulative precipitation estimates for the two data sources within each sample
+county.
 
@@
 
-Ground-based observations of wind speed are problematic during tropical
-cyclones, as instruments often fail at high wind speed, while reanalysis data,
-often available at hourly or higher resolution, can be too smooth to capture
-wind extremes associated with a tropical cyclone. Therefore, to create a
-dataset of county-level sustained winds during historical tropical cyclones, we
-modeled maximum sustained wind speeds at each county's population mean
-center~\parencite{countycenters}. \ac{HURDAT2} Best Track data were interpolated
-from~6-\si{\hour} reported values to~15-\si{\minute} increments using natural cubic
-splines~\parencite{stormwindmodel}, and then sustained wind speed was modeled for
-each county center at each~15-\si{\minute} increment using a double exponential wind
-speed model~\parencite{willoughby2006parametric} to estimate maximum ground-level
-sustained wind speed at each county center~\parencite{stormwindmodel}. Asymmetry in
-wind speeds around the tropical cyclone center was incorporated into the wind
-speed model~\parencite{phadke2003modeling, stormwindmodel}. The maximum value of
-this modeled wind speed was determined for each county as the maximum windspeed
-across the~15-\si{\minute} incremented modeled values throughout the tropical
-cyclone.
+Ground-based observations of wind speed are problematic as a metric of exposure
+to tropical cyclones, as instruments often fail at high wind speed, while
+reanalysis data, often available at hourly or higher resolution, can lack a
+fine enough temporal resolution to capture wind extremes associated with a
+tropical cyclone. Therefore, to create a dataset of county-level sustained
+winds during historical tropical cyclones, we modeled maximum sustained wind
+speeds at each county's population mean center~\parencite{countycenters}, using
+a wind speed model based on results from
+Willoughby~\parencite{willoughby2006parametric}. For each storm, we used the
+interpolated storm tracks generated for the distance metric and, for
+each~15-\si{\minute} increment, we modeled maximum ground-level sustained wind
+speed at each county center using a double exponential wind speed model with
+inputs for the storm's forward speed, direction, and central wind
+speed~\parencite{willoughby2006parametric, stormwindmodel}. This model
+incorporated asymmetry in wind speeds around the tropical cyclone center that
+results from the storm's forward movement~\parencite{phadke2003modeling,
+stormwindmodel}. We then determined the maximum value of this estimate over the
+storm tracking period for each study county to determine the maximum sustained
+surface wind speed in each county during each storm.
 
-To validate the metric of county-level wind speed used in the main text, we
-compared this wind metric to estimated wind radii from \ac{HURDAT2}, available
-for storms in our study from~2004~\parencite{landsea2013}. \ac{HURDAT2} gives
-estimates of the maximum radii to winds of~64,~50, and~34 \si{\knot} in each four
-quadrants of a storm (i.e., distance from the storm's center to the furthest
-point in a quadrant with winds at or above that speed) at each~6-\si{\hour} storm
-observation.
+To validate these modeled county-level wind speed estimates, we compared them
+to county-level maximum sustained surface wind speeds based on the wind radii
+estimates in \ac{HURDAT2}, which are based on \ldots. The wind radii values
+have been included in \ac{HURDAT2} since 2004 [?] and estimate the distance
+from the storm's center to the furthest point with winds at or above winds of a
+certain speed in each four quadrants of the storm. They give estimates of these
+maximum radii for three thresholds of maximum surface wind speeds: 64,~50,
+and~34 \si{\knot}.  They therefore allow for the classification of counties
+into four categories of maximum sustained wind speeds: $<$34~\si{\knot};
+34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}. We
+interpolated this data to~15-\si{\minute} increments and classified a county as
+exposed to winds in a given wind speed category if its population mean center
+was within~85\% of the maximum radius for that wind speed in its quadrant of
+the storm [clarify why 85\%]. We compared these wind radii-based estimates with
+the modeled wind speed estimates for the study storms since 2004 for which at
+least one study county had a sustained wind of~$\ge$34 \si{\knot} (based on the
+\ac{HURDAT2} wind radii). For each of these storms, we calculated the percent
+of study counties that were classified in the same wind speed category.
 
@@
 (``Flood", ``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and
-that occurred in a county within~500 \si{\kilo\metre} of the tropical cyclone's track
-and with a start date within a five-day window centered on the date of the
-tropical cyclone's closest approach to the
-county~\parencite{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado"
-events in this database were reported by county \ac{FIPS} code. ``Coastal
-Flood" events were reported by forecast zone; for these, the event was matched
-to the appropriate county if possible~\parencite{noaastormevents}. While this
+that occurred in a county within~500 \si{\kilo\metre} of the tropical cyclone's
+track and with a start date within a five-day window centered on the date of
+the tropical cyclone's closest approach to the
+county~\parencite{hurricaneexposuredata}. ``Flood", ``Flash Flood" and
+``Tornado" events in this database were reported by county \ac{FIPS} code.
+``Coastal Flood" events were reported by forecast zone; for these, the event
+was matched to the appropriate county if possible using regular expression
+matching of listed county names~\parencite{noaastormevents}. While this
 database has recorded storm data since~1950, its coverage changed substantially
@@
 
-As validation for the flood data, we compared streamflow data from \ac{US}
-Geological Survey gages~\parencite{usgsgages, countyfloods, dataRetrieval} in nine
-counties with tropical cyclone flood classifications based on the NOAA Storm
-Events Database (Figure~\ref{fig:floodcomparison}).
+The tornado observations from this dataset form the traditional tornado
+tracking dataset for the \ac{US} [ref for this?], and so we did not conduct
+analyses to validate the tornado event data.  As validation for the
+flood data, we investigated a sample of study counties, comparing the flood
+status, based on NOAA Storm Events listings~\parencite{stormevents,
+noaastormevents}, with streamflow measurements at \ac{US} Geological Survey county
+streamflow gages during tropical storms~\parencite{usgsgages, countyfloods,
+dataRetrieval}.  We considered nine study counties, selecting counties
+geographically spread through storm-prone areas of the eastern \ac{US} and with
+multiple streamflow gages reporting data during events (Baltimore County, MD;
+Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton County,
+GA; Harris County, TX; Mobile County, AL; Montgomery County, PL; and Wake
+County, NC). For each county, we first identified all streamflow gages in the
+county that had complete data for Jan.~1~1996\,--\,Dec.~31,~2015. For each
+tropical cyclone that came within~500~\si{\kilo\metre} of one of these counties
+between~1996 and~2015, we calculated the highest daily total streamflow (cubic
+feet / second), summed across all identified streamgages in the county, for a
+five-day window centered on the day of the storm's closest approach to the
+county.  We also calculated the percent of streamflow gages in the county with
+a daily streamflow that exceeded a threshold of flooding (the streamgage's
+median value for annual peak flow~\parencite{countyfloods}) on any day during
+the five-day window. We compared these measurements to each county's flood
+status based on NOAA Storm Events listings for each tropical cyclone that came
+within~500~\si{\kilo\metre} between~1996 and~2015, to explore if flood event
+listings in the NOAA Storm Events database concurrent with tropical cyclones
+were associated with higher streamflows at gages within the county.
 
-\subsection*{Assessing agreement in exposure classification between
-metrics}
+\subsection*{Binary storm exposure classifications based on different metrics}
+
+In our open-source dataset, we provide the closest distance of each storm to
+each county, as well as county-level storm-related winds and rainfall, as
+continuous metrics. However, epidemiologic studies of tropical cyclones often
+compare ``exposed'' versus ``unexposed'' communities to assess storm-related
+health risks [refs], and so we also used these continuous metrics to classify
+counties as ``exposed'' or ``unexposed'' and explored patterns in storm
+exposure based on these binary clasifications.
 
@@
 since these metrics were based on whether an event was listed in the NOAA Storm
-Events database or not.  For the other exposure metrics, each county was
-classified as exposed to a tropical cyclone based on whether the exposure
-metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
-assessed county-level tropical cyclone exposure in the eastern \ac{US} based on
-each exposure metric for all land-falling or near-land Atlantic basin tropical
-cyclones. Depending on available exposure data, this assessment included some
-or all of the period from~1988 to~2015, a period with~136 tropical cyclones
-that made \ac{US} landfall or passed within~250 \si{\kilo\metre} (155 mi) of at
-least one \ac{US} county (Figure~S2\todo{Check fig ref}).
+Events database.  For the other exposure metrics, each county was classified as
+exposed to a tropical cyclone based on whether the exposure metric exceeded a
+certain threshold (Table~\ref{tab:exposuremetrics}). For the rainfall metric, a
+distance constraint was also necessary, to ensure that rains unrelated and far
+from the storm track were not misattributed to a storm. Through exploratory
+analysis, we set this distance metric at~500 \si{\kilo\metre} (i.e., for a
+county to be classified as exposed based on rainfall, the cumulative rainfall
+had to be over 75 \si{\milli\metre} and the storm must have passed within~500
+\si{\kilo\metre} of the county; Table~\ref{tab:exposuremetrics}). We set this
+distance constraint at a value that was typically large enough to capture
+storm-related rain. However, data users should note that in rare examples of
+exceptionally large storms (e.g., Hurricane Ike in 2008) or storms for which
+storm tracking was stopped at extratropical transition (e.g., Tropical Storm
+Lee in 2011), some storm-related rains may be missed because of this distance
+constraint (Figure~S7).
 
-We measured agreement between exposure metrics in the classification of a
-county as exposed or unexposed to a specific tropical cyclone. For this
-measurement, we calculated the within-storm Jaccard
-index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between every
-pair-wise combination of hazard metrics.  The Jaccard index~($J_s$) measures
-similarity between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical
-cyclone~$s$ as the proportion of counties in which both of the metrics classify
-the county as exposed out of all counties classified as exposed by at least one
-of the metrics:
+We assessed patterns in county-level tropical cyclone exposure in the eastern
+\ac{US} based on each exposure metric. Depending on available exposure data,
+this assessment included some or all of the period from~1988 to~2015 (Table 1).
+For each binary metric of tropical cyclone exposure (Table 1), we first summed
+the total number of county-level exposures over available years for each
+exposure metric and mapped patterns in these exposures. We next investigated
+agreement between storm exposure classifications based on different exposure
+metrics. We calculated the within-storm Jaccard
+index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between each
+pair of exposure metrics. The Jaccard index~($J_s$) measures similarity between
+two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical cyclone~$s$ as the
+proportion of counties in which both of the metrics classify the county as
+exposed out of all counties classified as exposed by at least one of the
+metrics:
 
@@
 counties classified as exposed based on the two metrics, to~1, in the case that
-the two metrics classify exactly the same counties as exposed to the tropical
-cyclone.
+the two metrics classify exactly the same set of counties as exposed to the
+tropical cyclone. We measured these values for all study storms that affected a
+large number of eastern \ac{US} counties (250 or more of the study counties
+classified as exposed by at least one of the exposure metrics) during the years
+when all exposure data were available (1996\,--\,2011).
 
@@
 
-The outcomes of disasters depend both on the geophysical forces of the disaster
-as well as on the vulnerability of the society living in the geographical areas
-affected by those forces.~\parencite{chakraborty2005population,
-anderson2003community, cutter1996vulnerability} As a case study of how
-differences in tropical cyclone exposure assessments across different hazard
-metrics influence estimates of physical exposures of susceptible populations to
-tropical cyclones, we investigated how the use of different metrics influenced
-estimates of which eastern \ac{US} counties have the highest expected average
-exposures of electricity-dependent Medicare beneficiaries to tropical cyclones.
-This subpopulation was selected for this analysis since it may be particularly
-susceptible to health impacts from tropical cyclone exposure, especially
-through the pathway of storm-associated power outages and evacuations. We
-collected data on the number of electricity-dependent Medicare beneficiaries in
-each study county from the U.S. Department of Health \& Human Service's emPower
-Map~2.0~\parencite{empower}. We then calculated the physical exposure of the
-electricity-dependent Medicare population in each county, based on tropical
-cyclone assessments using each hazard metric,
+Disasters' societal impacts depend both on the geophysical forces of the
+disaster and on the vulnerability of those living in the affected geographical
+areas~\parencite{chakraborty2005population, anderson2003community,
+cutter1996vulnerability}. As a case study, we explored how differences in
+tropical cyclone exposure assessments across different exposure metrics might
+influence estimates of physical exposures of a susceptible population to
+tropical cyclones. We calculated expected average physical exposures among
+electricity-dependent Medicare beneficiaries in eastern \ac{US} counties to
+tropical cyclones based on each exposure metric. This subpopulation was
+selected since it may be particularly susceptible to health impacts from
+tropical cyclone exposure, especially through the pathway of storm-associated
+power outages and evacuations.
+
+We collected data on the number of electricity-dependent Medicare beneficiaries
+in each study county from the \ac{US} Department of Health \& Human Service's
+emPower Map~2.0~\parencite{empower}. We then calculated the physical exposure
+of the electricity-dependent Medicare population in each county, based on
+tropical cyclone assessments using each hazard metric,
 following~\parencite{peduzzi2009assessing}:
@@
 frequency of tropical cyclone exposures in county~$c$ based on that metric,
-and~$P_c$ is the size of the electricity-dependent Medicare population in
-county~$c$, as of July~2017. When combined with estimates of vulnerability of a
+and~$P_c$ is the electricity-dependent Medicare population in
+county~$c$, as of July~2017.
+
+\begin{comment}
+When combined with estimates of vulnerability of a
 population to a natural hazard, such measurements of physical exposure can be
@@
 hazard~\parencite{peduzzi2009assessing}.
+\end{comment}

### 2019-07-27 writing/ehp_results.tex
Work on draft throughout

--- 2019-01-27 writing/ehp_results.tex
@@
 \section*{Results}
+
+\subsection*{Exposure assessment data, data validation, and software}
+
+\paragraph{Open-source dataset}
+
+We created and published open source data with multi-year, county-level
+tropical cyclone exposure data for eastern \ac{US}
+counties~\parencite{hurricaneexposuredata}. These data include continuous
+county-level measurements for the closest distance of each storm, cumulative
+rainfall, and maximum sustained surface wind speed. These continuous metrics
+can be used for classifying counties as exposed or unexposed, using thresholds
+selected by the user, or can be used as continuous metrics. The dataset also
+include binary data on flood and tornado events associated with each storm in
+each county. With these data, we published software tools to explore and map
+the data and to integrate it with human health
+datasets~\parencite{hurricaneexposure}.
+
+We explored potential limitations in these data by comparing them with data
+from other available sources.
+
+\paragraph{Rainfall data validation}
+
+For estimates of storm-associated rainfall, we compared estimates from NLDAS-2
+re-analysis precipitation data (included as the rainfall estimates in the
+open-source data) with ground-based observations in nine sample counties
+(Figure~\ref{fig:raincomparison}). Within these counties, storm-related
+rainfall measurements were well-correlated between the two data sources, with
+rank correlations (bottom right of each graph in
+Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some evidence
+that the reanalysis data may tend to underestimate rainfall totals in storms
+with extremely high rainfall, based a few heavy-rainfall storms in Harris
+County, TX, Mobile County, AL, Charleston County, SC, and Wake County, NC
+(Figure~\ref{fig:raincomparison}). However, even in these cases, it was rare
+for a storm to be classified differently (based on the precipitation threshold
+of 75~\si{\milli\metre} we used for exposure classification for later analysis)
+based on the source of precipitation data---horizontal and vertical lines in
+each small plot in Figure~\ref{fig:raincomparison} show the threshold of
+75~\si{\milli\metre}, so storms in the lower left and upper right quadrants
+would be classified the same (``exposed'' or ``unexposed'') regardless of the
+precipitation data source, while storms in the upper left and lower right
+quadrants would be classified differently.
+
+\paragraph{Wind speed data validation}
+
+For maximum sustained surface wind speed estimates, we found that the modeled
+estimates, which we included as the primary wind exposure metric in the open
+source data and used for further analysis in this study, generally agreed well
+with estimates based on the wind radii reported in \ac{HURDAT2}. For most
+storms, $\ge$95\% of counties were assigned the same category of wind speed
+($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot};
+$\ge$64~\si{\knot}) by both data sources (Figure~\ref{fig:windcomparison}).
+Disagreement was limited to large storms that were characterized by high winds
+well inland, like Hurricanes Sandy in 2012 and Ike in 2008. For these two
+storms, the modeled wind speed values somewhat overestimated the extent of the
+storm's most severe winds at landfall but then underestimated the extent of
+34\,--\,49.9 \si{\knot} storm winds further inland (Figure~S6). For researchers
+who would like to conduct sensitivity analysis between these two sets of wind
+data, we have included estimates from the \ac{HURDAT2} wind radii as a
+secondary measure of wind speeds in the open-source dataset
+\parencite{hurricaneexposuredata}.
+
+\paragraph{Flood data validation}
+
+For flood data, we compared the flood status measurements, based on event
+listings in the NOAA Storm Events database, to streamflow measurements at
+\ac{US} Geological Survey gages within nine sample counties
+(Figure~\ref{fig:floodcomparison}). Each small plot in
+Figure~\ref{fig:floodcomparison} shows results for one of the sample counties.
+Each point represents a single tropical storm, and the point's position along
+the x-axis shows the highest daily total streamflow (cubic feet per second),
+summed across all identified streamgages in the county, for a five-day window
+centered on the day of the storm's closest approach to the county. The y-axis
+separates storms for which a flood event was reported in NOAA's Storm Events
+database for the county with a start date within the five-day window of the
+storm's closest approach to the county. The color of each point gives the
+percent of streamflow gages in the county with a daily streamflow that exceeded
+a threshold of flooding on any day during the five-day window.  Stream gage
+data generally indicated higher discharge in these counties on dates identified
+as flood events, particularly in counties with more reporting stream gages
+(e.g., Harris County, TX, with~10 reporting stream gages;
+Figure~\ref{fig:floodcomparison}). There were some cases, however, where the
+two flooding data sources were not completely consistent. 
+For example, there
+were one or two tropical cyclones in several of the counties (Mobile County,
+AL, Escambia County, FL, and Montgomery County, PA) which were classified as
+flood events based on NOAA storm event listings but for which the total
+discharge across county streamflow gages was below the 25\textsuperscript{th}
+percentile of measurements during storms without a flood event listing. 
+For tropical cyclones that did not have an associated flood listing in 
+the NOAA Storm Events database, most had lower total streamflow discharge
+across county gages compareed to storms with flood event listings, and in all
+cases but one none of the gages were over the flooding threshold. The exception
+was \ldots in Fulton County, GA. [More on which storm and why---I think there
+was already a flood event that had started earlier, so it missed the cutoff for
+the startdate being within the range.]
+
+[For discussion?] Based on these results, the two data sources were generally
+in agreement, but the analysis highlights the difficulty of assessing
+storm-related flooding at a multi-county, multi-year scale, as flooding can be
+very localized within a county or can \ldots .
 
@@
 
-We first calculated summary statistics for the exposure assessments
-(Table~\ref{tab:exposuresummaries}). There was wide variation in the average
-number of county exposures per year under the five different tropical cyclone
-metrics considered. For the tornado metric, there were approximately~50 county
-exposures identified per year, on average, within our study.  For flood and
-wind metrics, there were substantially more county exposures identified ($>$200
-per year on average), and even more when assessing exposure using the distance
-and rain metrics ($>$400 per year on average).
+We used the storm exposure metrics to classify counties as exposed or unexposed
+to specific tropical cyclones, and we explored patterns in these exposure
+classifications over years with available data
+(Table~\ref{tab:exposuresummaries}). Across the five exposure metrics, there
+was wide variation in the average number of county exposures per year. For the
+tornado metric, there were approximately~50 county exposures per
+year, on average, within our study.  For flood and wind metrics, there were
+substantially more county exposures ($>$200 / yr on average), and
+even more when assessing exposure using the distance and rain metrics ($>$400 /
+yr on average).
 
-We also measured the average number of county exposures per tropical cyclone,
-within those tropical cyclones in which at least one county was exposed to the
-hazard (Table~\ref{tab:exposuresummaries}). The typical number of county
-exposures per tropical cyclone was lowest when assessing exposure using the
-tornado metric (median:~7 county exposures per tropical cyclone), higher under
-the flood and wind metrics (median:~26 and~34, respectively), and even higher
-for the metrics of rain (median:~68) and distance (median:~84).  For every
-metric except the tornado-based metric, we identified at least one tropical
-cyclone with over~300 counties exposed.  However, the largest-extent tropical
-cyclone varied by metric (e.g., Beryl in~1994 exposed the most counties based
-on distance, Frances in~2004 based on rain, and Ike in~2008 based on wind).
+We also measured the average number of county exposures per storm under each
+exposure metric (limiting this analysis to the tropical cyclones that exposed
+at least one county to that hazard; Table~\ref{tab:exposuresummaries}).
+Tropical cyclones typically exposed only a few counties, if any, to tornadoes
+(median counties exposed per storm:~7), more counties under the flood and wind
+metrics (median:~26 and~34, respectively), and even more counties for the
+metrics of rain (median:~68) and distance (median:~84).  For every metric
+except the tornado-based metric, we identified at least one tropical cyclone
+with over~300 counties exposed.  However, the largest-extent tropical cyclone
+varied by metric (e.g., Beryl in~1994 exposed the most counties based on
+distance, Frances in~2004 based on rain, and Ike in~2008 based on wind).
 
-\subsection*{Agreement in county-specific exposure classification to specific
-tropical cyclones}
-
-The counties that were identified as exposed to a tropical cyclone differed
-substantially depending on which metric was used in exposure assessment.  This
-was true both when comparing exposure assessment based on distance versus the
-four hazard-specific metrics considered (wind, rain, floods, and tornadoes)
-and also when comparing exposure assessments among the four hazard-specific
-metrics.
-
-Figure~\ref{fig:ivanexposure} gives an example of how sensitive county-level
-tropical cyclone exposure assessment is to the choice of metric, using as an
-example Hurricane Ivan in 2004, which was record-breaking for its
-duration as a major hurricane~\citep{franklin2006atlantic}. When county-level
-exposure was determined based on distance, the counties assessed as exposed
-followed the tropical cyclone's track, including counties well inland. For the
-wind metric, only counties near two of the tropical cyclone's landfalls were
-assessed as exposed. For rain- and flood-based metrics, however, exposure
-extended to the left of the track, including counties as far north as New York
-and Connecticut, while for the tornado metric, exposed counties tended to be to
-the right of the track and included several counties in central North Carolina,
-South Carolina, and Georgia that were not identified as exposed to Ivan based
-on any other metric.  Figure~S3 provides similar maps for four other example
-tropical cyclones (selected because they exposed many \ac{US}  counties based on
-at least one metric), and Figure~S4 provides four examples of tropical cyclones
-in which the disagreement in exposure assessment between rain- and wind-based
-metrics was particularly notable.
-
-We drew similar conclusions when we investigated all~46 tropical cyclones
-between~1996 and~2011 (the period for which all five metrics were available)
-for which~250 or more counties were exposed based on at least one metric
-(Figure~\ref{fig:jaccard}). In this figure, each row provides results for a
-specific tropical cyclone, and each box in that row shows a measure of
-agreement between a specific pair of metrics in terms of the counties assessed
-as exposed to that tropical cyclone, with darker colors representing greater
-agreement in county-level exposure assessments.
-
-For most tropical cyclone, distance-based county-level exposure assessment was
-at best in moderate agreement with exposure assessments based on the four
-hazard-based metrics. For almost all tropical cyclones, Jaccard index values
-between distance and any of the four hazard-based metrics were~0.6 or lower
-(i.e., out of the counties assessed as exposed to the tropical cyclone by at
-least one of the two metrics considered,~60\si{\percent} or fewer of the
-counties had the same exposure assessment under the two metrics). For many
-tropical cyclones, the Jaccard index values were~0.2 or lower, suggesting very
-poor agreement in exposure assessment between the metrics.  Among the four
-hazard-based metrics, the tornado-based metric showed universally poor
-agreement with other metrics in county-level classification across all tropical
-cyclones considered.  For other pairs of metrics, there were also generally
-large differences in which counties were determined to be exposed to the
-tropical cyclone, with the Jaccard index values below~0.4 for most metric pairs
-in most tropical cyclones.
-
-There were, however, a few exceptions\,---\,tropical cyclones in which similar
-counties were determined to be exposed to the tropical cyclone for two or more
-of the metrics considered.  For Floyd in~1999 (Figure~S3) and Irene in~2011,
-for example, county-level classification agreed moderately to well (Jaccard
-index of approximately~0.5\,--\,0.8) for all pairs of exposure metrics except
-those including the tornado-based metric. These tropical cyclones both made
-their first \ac{US}  landfall in North Carolina at minor hurricane windspeeds
-(Category~2 and~1, respectively) and then skimmed the eastern coast of the
-\ac{US} north through New England, bringing large rainfall to much of the
-eastern coast from North Carolina north and causing extensive inland flooding
-in North Carolina (Floyd) and New England (Irene)~\citep{avila2013atlantic,
-lawrence2000atlantic}. For another set of tropical cyclones (e.g., Lee in~2011
-[Figure~S3], Ernesto in~2006, and Bertha in~1996), there was moderate to good
-agreement for all pairwise combinations of distance, rain, and wind, but poor
-agreement for all other combinations of metrics. Most of these tropical
-cyclones either followed the eastern coastline north of North Carolina (Bertha
-in~1996, Charley in~2004, Gaston in~2004, Barry in~2007, and Hanna in~2008), in
-a pattern similar to Hurricanes Floyd in~1999 and Irene in~2011, or made
-landfall in or near the Florida panhandle and cut across to exit into the
-Atlantic around North Carolina (Josephine in~1996, Earl in~1998, Alberto
-in~2006). However, these two sets of tropical cyclones represented the minority
-of all tropical cyclones considered; for most tropical cyclones, there was low
-overlap in the counties determined to be exposed to the tropical cyclones
-between most pairings of metrics.
-
-\subsection*{Geographic patterns in tropical cyclone exposures}
-
-When the average number of exposures per decade in each county were calculated,
-strong geographical patterns were clear across tropical cyclone exposure
-assessments based on these different metrics
-(Figure~\ref{fig:averageexposure}). Wind-based exposure had a strong coastal
+When we calculated and mapped the average number of exposures per decade in
+each county for each exposure metric (Figure~\ref{fig:averageexposure}), strong
+geographical patterns were clear. Wind-based exposure had a strong coastal
 pattern, with almost all exposures in counties within about~200
@@
 distance-based metrics were also more common in coastal areas compared to
-inland areas, inland exposures were much more common based on these two metrics
-compared to the wind-based metric. Rain and, to some extent, flood exposures
-were characterized by a pattern defined by the Appalachian Mountains, with much
+inland areas, they also created inland exposures that were not identified by
+the wind-based metric. Rain and, to some extent, flood exposures were
+characterized by a pattern defined by the Appalachian Mountains, with much
 lower exposures west of the mountain range than to the east. Almost all
@@
 more commonly used tropical cyclone exposure assessment metrics of wind or
-distance (Figure~\ref{fig:averageexposure}).
+distance (Figure~\ref{fig:averageexposure}). Patterns were similar when
+analysis was restricted to years with all exposure data available
+(1996\,--\,2011; Figure~S5).
 
-We conducted a case study to investigate how these differences in exposure
-assessment across the five metrics considered could strongly influence
-health-related studies based on physical exposure to tropical cyclones. For
-this example, we compared estimates of which \ac{US} counties have the largest
-expected physical exposure to tropical cyclones among a susceptible
-subpopulation, Medicare beneficiaries who are reliant on electricity for
-medical equipment (Figure~S6)~\citep{empower}.  We found
-that a few counties (Miami-Dade County, FL; Harris County, TX) were ranked in
-the top three counties of physical exposure among electricity-dependent
-Medicare beneficiaries for almost all exposure metrics
-(Figure~\ref{fig:topelecdependexposure}). However, there were key differences
-when using different exposure metrics. For example, Florida counties dominated
-the lists of top exposed counties based on wind- and tornado-based metrics,
-while over half of the counties for the flood-based metric were in Mid-Atlantic
-states (PA, NY).
+\subsection*{Agreement in county-specific exposure classification between exposure metrics}
 
-\subsection*{Software}
+Next, we measured the agreement between exposure classifications for each pair
+of metrics. We found that these exposure classifications typically did not
+strongly agree between pairs of metrics, indicating that the set of counties
+identified as exposed based on one metric can have limited overlap with the
+set of counties identified as exposed to the storm by another metric.
 
-To assist with future tropical cyclone impact studies, we created and published
-open source software that includes the hazard-specific, county-level tropical
-cyclone data in this paper, as well as tools to explore and map that 
-data~\citep{hurricaneexposure, hurricaneexposuredata}. This software includes continuous
-data for the distance-, rain-, and wind-based metrics, allowing users to also explore
-different thresholds of these metrics when assessing exposure.
+Figure~\ref{fig:ivanexposure} shows Hurricane Ivan in 2004 as an example. For
+the distance-based metric, the counties assessed as exposed followed the
+tropical cyclone's track, including counties well inland. For the wind metric,
+only counties near two of the tropical cyclone's landfalls were assessed as
+exposed. For rain- and flood-based metrics, however, exposure extended to the
+left of the track, including counties as far north as New York and Connecticut,
+while for the tornado metric, exposed counties tended to be to the right of the
+track and included several counties in central North Carolina, South Carolina,
+and Georgia that were not identified as exposed to Ivan based on any other
+metric. Figure~S3 provides similar maps for four other example tropical
+cyclones (selected because they exposed many \ac{US}  counties based on at
+least one metric), and Figure~S4 provides four examples of tropical cyclones in
+which the disagreement in exposure assessment between rain- and wind-based
+metrics was particularly notable.
 
-We investigated how well these data correspond with data from other available
-sources (Figures~S7\,--\,11). Within the nine sample counties, storm-related
-rainfall from the reanalysis precipitation dataset was well-correlated with
-station-based rainfall measurements, although the reanalysis data did tend to
-somewhat underestimate rainfall totals in storms with extremely high rainfall
-(Figure~\ref{fig:raincomparison}). A distance constraint (county population
-mean center within~500 \si{\kilo\metre} of the storm's closest approach) was
-necessary for assessing storm exposure based on this metric to ensure that
-rains unrelated and far from the storm track were not misattributed to storm
-exposures. We found that this distance constraint was typically adequately
-large to capture storm-related rain, although in rare examples of exceptionally
-large storms (e.g., Hurricane Ike in 2008) or storms for which tracking was
-stopped at extratropical transition (e.g., Tropical Storm Lee in 2011), some
-storm-related rains were missed because of this distance constraint
-(Figure~\ref{fig:rainexamples}).  The modeled wind speeds generally agreed well
-with the \ac{HURDAT2} wind radii (Figure~\ref{fig:windcomparison}), with
-disagreement mostly limited to large storms that were characterized by high
-winds well inland (e.g., Hurricanes Ike in 2008 and Sandy in 2012;
-Figure~\ref{fig:windexamples}).  Stream gage data generally indicated higher
-discharge in these counties on dates identified as flood events, particularly
-in counties with more reporting stream gages (e.g., Harris County, TX, with~10
-reporting stream gages; Figure~\ref{fig:floodcomparison}), although these two
-flooding data sources were not completely consistent.
+We drew similar conclusions when we investigated all~46 tropical cyclones
+between~1996 and~2011 (the period for which all five metrics were available)
+for which~250 or more counties were exposed based on at least one metric
+(Figure~\ref{fig:jaccard}). In this figure, each row provides results for a
+specific tropical cyclone, and each box in that row shows the measured Jaccard
+coefficient between two of the exposure metrics for that storm.  For most
+tropical cyclones, distance-based county-level exposure assessment was at best
+in moderate agreement with exposure assessments based on the four hazard-based
+metrics. For almost all tropical cyclones, Jaccard index values between
+distance and any of the four hazard-based metrics were~$\le0.6$ (i.e., out of
+the counties assessed as exposed to the tropical cyclone by at least one of the
+two metrics considered,~60\si{\percent} or fewer of the counties had the same
+exposure assessment under the two metrics). For many tropical cyclones, the
+Jaccard index values were~$\le0.2$, suggesting very poor agreement in exposure
+assessment between the metrics.  Among the four hazard-based metrics, the
+tornado-based metric showed universally poor agreement with other metrics in
+county-level classification across all tropical cyclones considered.  For other
+pairs of metrics, there were also generally large differences in which counties
+were determined to be exposed, with the Jaccard index values~$<0.4$ for most
+metric pairs in most tropical cyclones.
 
-While generally in agreement with data from other
-sources, there are a few limitations. The wind data are based on modeled,
-rather than observed, values, and while the modeled wind data generally agree
-well with post-analysis maximum wind radii from the revised Atlantic hurricane
-database \ac{HURDAT2}~\citep{landsea2013} (Figure~S7), for a few tropical cyclones
-it did not (e.g., Figure~S10). The rainfall data are from re-analysis data,
-which are generally well-correlated with observed ground-based station data but
-sometimes oversmooth extreme measurements (Figure~S8).
+\begin{comment}
+Storms are displayed within
+clusters that have similar patterns in county-level exposure agreement for
+metric pairs, based on hierarchical clustering using the complete link
+method~\parencite{murtagh2012algorithms} (i.e., storms in the same cluster tend
+to have similar patterns for the pairwise strength of agreement among metrics);
+columns are also ordered based on hierarchical clustering. 
+Maps are
+available showing the counties identified as exposed under each of five metrics
+for the widest-extent storm in each cluster: Hurricane Ivan in~2004
+(Figure~\ref{fig:ivanexposure}) and Hurricanes Floyd in~1999, Lee in~2011, Cindy
+in~2005, and Katrina in~2005 (Figure~S3).
+\end{comment}
+
+There were, however, a few exceptions\,---\,tropical cyclones in which similar
+counties were determined to be exposed to the tropical cyclone for two or more
+of the metrics considered.  For Floyd in~1999 (Figure~S3) and Irene in~2011,
+for example, county-level classification agreed moderately to well (Jaccard
+index of~$\sim 0.5\mbox{\,--\,}0.8$) for all pairs of exposure metrics except
+those including the tornado-based metric. For another set of tropical cyclones
+(e.g., Lee in~2011 [Figure~S3], Ernesto in~2006, and Bertha in~1996), there was
+moderate to good agreement for all pairwise combinations of distance, rain, and
+wind, but poor agreement for all other combinations of metrics. However, these
+two sets of tropical cyclones represented the minority of all tropical cyclones
+considered; for most tropical cyclones, there was low overlap in the counties
+determined to be exposed to the tropical cyclones between most pairings of
+metrics.
+
+\subsection*{Case study}
+
+Finally, we conducted a case study to investigate how these differences in
+exposure assessment across the exposure metrics could affect estimates of
+expected physical exposure among a vulnerable subpopulation,
+electricity-dependent Medicare beneficiaries.
+Figure~\ref{fig:topelecdependexposure} shows the study counties with the
+highest expected physical exposures in this subpopulation based on each
+exposure metric. A few counties (Miami-Dade County, FL; Harris County, TX) were
+ranked in the top three counties of physical exposure for almost all exposure
+metrics (Figure~\ref{fig:topelecdependexposure}). However, there were key
+differences when using different exposure metrics. For example, Florida
+counties made up most of the counties counties with highest expected physical
+exposure based on wind- and tornado-based metrics, while over half of the
+counties with highest expected physical exposures for the flood-based metric
+were in Mid-Atlantic states (PA, NY).

### 2019-07-27 writing/ehp_tables.tex
Work on draft throughout

--- 2019-01-27 writing/ehp_tables.tex
@@
 Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 \si{\kilo\metre} of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
-Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
-Tornado & 1996--2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track\\
+Distance & 1988\,--\,2015 & Storm track came within 100~\si{\kilo\metre} of county population mean center\\ 
+Rain & 1988\,--\,2011 & County cumulative rainfall of $\ge$75~\si{\milli\metre} over the period from two days before to 
+one day after the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\ 
+Wind & 1988\,--\,2015 & Modeled storm-associated maximum sustained winds of $\ge$15~\si{\metre\per\second}
+at the county's population mean center\\ 
+Flood & 1996\,--\,2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center \\
+Tornado & 1996\,--\,2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\
 \bottomrule 

### 2019-07-28 writing/ehp_discussion.tex
Work more on paper draft

@@
 
-The flood and tornado data came from the \ac{NOAA} Storm Events database,
-which, while a widely-used database of events maintained by \ac{NOAA}, is based
-on reports, and so may be prone to underreporting~\parencite{Ashley2008flood,
-Curran2000}, especially in less populated areas~\parencite{Witt1998,
-Ashley2007}, as well as to other reporting errors. While these are important
-limitations, we selected these data sources as among the best currently
-available for measuring each of these hazards at a multi-county, multi-year
-scale.
+For the flooding events data, we found it often corresponded with measurements
+from \ac{US} Geographical Survey streamgages, with flood event listings more
+likely during storms which elevated stream flows. However, there are
+differences between the two flooding datasets, and these highlight both the
+difficulty of measuring flood exposure at the county level and inherent
+challenges in using the storm event database for epidemiologic exposure
+assessment.  The flood and tornado data came from the \ac{NOAA} Storm Events
+database, which, while a widely-used database of events maintained by
+\ac{NOAA}, is based on reports, and so may be prone to
+underreporting~\parencite{Ashley2008flood, Curran2000}, especially in less
+populated areas~\parencite{Witt1998, Ashley2007}, as well as to other reporting
+errors.
+
+\begin{comment}
+[For discussion?] Based on these results, the two data sources were generally
+in agreement, but the analysis highlights the difficulty of assessing
+storm-related flooding at a multi-county, multi-year scale, as flooding can be
+very localized within a county or can \ldots .   
+\end{comment}
+
+While these are important caveats for the data, we selected these data sources
+as among the best currently available for measuring each of these hazards at a
+multi-county, multi-year scale.
 
@@
 
-Dissimilarity in exposure to specific hazards of tropical cyclones can result
-in differing patterns in estimates of county-level exposure
-(Figure~\ref{fig:ivanexposure}), which can lead to substantial differences
-in estimates of population physical exposure
-(Figure~\ref{fig:topelecdependexposure}), based on which storm hazard is
-considered. The measure of a hurricane's intensity, the Saffir-Simpson scale,
-expresses intensity as a function of wind speed, and so fails to capture
-potential risks from other hazards of the storm~\parencite{smith2009}, and
-previous assessments of physical exposure to tropical cyclones have often
-focused on winds or distance from the storm. For example, a global study of
-exposure and vulnerability to natural hazards used a combination of windspeed
-above a certain threshold and distance from the storm's track to assess
+The conducted case study demonstrates how differences in exposure assessment
+across storm exposure metrics can result in differing estimates of expected
+physical exposure to tropical cyclones among a population
+(Figure~\ref{fig:topelecdependexposure}).  Previous assessments of physical
+exposure to tropical cyclones have often focused on solely storm winds and
+storm proximity. For example, a global study of exposure and vulnerability to
+natural hazards used a combination of windspeed above a certain threshold and
+distance from the storm's track to assess
 exposure~\parencite{peduzzi2009assessing}. While many storm impacts might be

### 2019-07-28 writing/ehp_figures.tex
Work more on paper draft

@@
 \caption{Comparison for a sample of study counties of NOAA Storm Events 
+	listings versus five-day total streamflows at county streamflow 
+	gages during tropical cyclones. Each small plot shows results for 
+	one of the sample counties. Each point represents a single tropical 
+	storm, and the point's position along the x-axis shows the highest 
+	daily total streamflow (cubic feet per second), summed across all 
+	identified streamgages in the county, for a five-day window centered 
+	on the day of the storm's closest approach to the county. The y-axis 
+	separates storms for which a flood event was reported in NOAA's Storm Events 
+	database for the county with a start date within the five-day window 
+	of the storm's closest approach to the county. The color of each point 
+	gives the percent of streamflow gages in the county with a daily streamflow 
+	that exceeded a threshold of flooding on any day during the five-day window.
+	Comparison for a sample of study counties of NOAA Storm Events 
 	listings versus five-day total streamflows at county streamflow gages during 
@@
 \caption{Average number of storm exposures per decade in U.S. counties for each 
-exposure metric. The criteria behind each of the five metrics is given in Table
-\ref{tab:exposuremetrics}. The years used to estimate these averages are based
-on years of available exposure data (distance and wind:~1988\,--\,2015; rain:
-1988\,--\,2011; flood and tornado:~1996\,--\,2015). } 
+	exposure metric. The criteria behind each of the five metrics is given in 
+	Table \ref{tab:exposuremetrics}. The years used to estimate these averages 
+	are based on years of available exposure data 
+	(distance and wind:~1988\,--\,2015; rain: 1988\,--\,2011; 
+	flood and tornado events:~1996\,--\,2015). } 
 \label{fig:averageexposure} 
@@
 \includegraphics[width=15.5cm]{topelecdependexposure}
-\caption{Counties with the highest estimated physical exposure among
-electricity-dependent Medicare beneficiaries exposed to storms per year in U.S.
-counties for each exposure metric. The color of each bar indicates the number
-of Medicare beneficiaries in the county reliant on electricity-dependent
-medical and assistive equipment as of July~2017. The length
-of each bar shows the average expected number of these electricity-dependent
-Medicare beneficiaries exposed to tropical storms per year based on a given
-exposure metric.}
+\caption{Study counties with the highest expected physical exposure per year among
+	 electricity-dependent Medicare beneficiaries for each exposure metric. 
+	 The color of each bar indicates the number of Medicare beneficiaries in the 
+	 county reliant on electricity-dependent medical and assistive equipment as 
+	 of July~2017. The length of each bar shows the average expected physical exposure
+	 to tropical cyclones among this population based on a given exposure metric, i.e., 
+	 the expected number of these electricity-dependent Medicare beneficiaries exposed 
+	 to tropical storms per year based on that metric.}
 \label{fig:topelecdependexposure}

### 2019-07-28 writing/ehp_methods.tex
Work more on paper draft

@@
 
-To identify flood- and tornado-based exposures to tropical cyclones in \ac{US}
+To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
 counties, we used event listings from the National Oceanic and Atmospheric
@@
 county~\parencite{hurricaneexposuredata}. ``Flood", ``Flash Flood" and
-``Tornado" events in this database were reported by county \ac{FIPS} code.
-``Coastal Flood" events were reported by forecast zone; for these, the event
-was matched to the appropriate county if possible using regular expression
-matching of listed county names~\parencite{noaastormevents}. While this
-database has recorded storm data since~1950, its coverage changed substantially
-in~1996 to cover a wider variety of storm events~\parencite{stormevents}. We
-therefore only considered these metrics of hurricane exposure for tropical
-cyclones in~1996 and later.
+``Tornado" events in this database were reported by county \ac{FIPS} code and
+so could be directly linked to counties.  ``Coastal Flood" events were reported
+by forecast zone; for these, the event was matched to the appropriate county if
+possible using regular expression matching of listed county
+names~\parencite{noaastormevents}. While this database has recorded storm data
+since~1950, its coverage changed substantially in~1996 to cover a wider variety
+of storm events~\parencite{stormevents}. We therefore only considered these
+metrics of hurricane exposure for tropical cyclones in~1996 and later.
 
-The tornado observations from this dataset form the traditional tornado
-tracking dataset for the \ac{US} [ref for this?], and so we did not conduct
-analyses to validate the tornado event data.  As validation for the
-flood data, we investigated a sample of study counties, comparing the flood
-status, based on NOAA Storm Events listings~\parencite{stormevents,
-noaastormevents}, with streamflow measurements at \ac{US} Geological Survey county
-streamflow gages during tropical storms~\parencite{usgsgages, countyfloods,
-dataRetrieval}.  We considered nine study counties, selecting counties
-geographically spread through storm-prone areas of the eastern \ac{US} and with
-multiple streamflow gages reporting data during events (Baltimore County, MD;
-Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton County,
-GA; Harris County, TX; Mobile County, AL; Montgomery County, PL; and Wake
-County, NC). For each county, we first identified all streamflow gages in the
-county that had complete data for Jan.~1~1996\,--\,Dec.~31,~2015. For each
-tropical cyclone that came within~500~\si{\kilo\metre} of one of these counties
-between~1996 and~2015, we calculated the highest daily total streamflow (cubic
-feet / second), summed across all identified streamgages in the county, for a
-five-day window centered on the day of the storm's closest approach to the
-county.  We also calculated the percent of streamflow gages in the county with
-a daily streamflow that exceeded a threshold of flooding (the streamgage's
-median value for annual peak flow~\parencite{countyfloods}) on any day during
-the five-day window. We compared these measurements to each county's flood
-status based on NOAA Storm Events listings for each tropical cyclone that came
-within~500~\si{\kilo\metre} between~1996 and~2015, to explore if flood event
-listings in the NOAA Storm Events database concurrent with tropical cyclones
-were associated with higher streamflows at gages within the county.
+The tornado observations from this dataset form the traditional tornado event
+database for the \ac{US}, and so we did not conduct analyses to validate the
+tornado event data. It is difficult to characterize flooding at the county
+level because flooding can be very localized and can be triggered by a variety
+of causes. To investigate the extent to which the NOAA flood event data
+captures extremes that might be identified with other flooding data sources, we
+investigated a sample of study counties, comparing the flood event data during
+tropical storms with streamflow measurements at \ac{US} Geological Survey
+county streamflow gages \parencite{usgsgages, countyfloods, dataRetrieval}.
+
+We considered nine study counties, selecting counties geographically spread
+through storm-prone areas of the eastern \ac{US} and with multiple streamflow
+gages reporting data during events (Baltimore County, MD; Bergen County, NJ;
+Escambia County, FL; Fairfield County, CT; Fulton County, GA; Harris County,
+TX; Mobile County, AL; Montgomery County, PL; and Wake County, NC). For each
+county, we first identified all streamflow gages in the county that had
+complete data for Jan.~1~1996\,--\,Dec.~31,~2015. If a storm did not come
+within~500~\si{\kilo\metre} of a county, it was excluded from this analysis,
+but all other study storms were considered. For each storm and county, we
+summed daily total streamflow across all streamgages for each day in the
+five-day window around the storm's closest approach. We took the maximum daily
+streamflow total during those five days as a measure of the county's maximum
+streamflow during that storm. We also calculated the percent of streamflow
+gages in the county with a daily streamflow that exceeded a threshold of
+flooding (the streamgage's median value for annual peak
+flow~\parencite{countyfloods}) on any day during the five-day window. We
+investigate how these measurements varied between storms with associated flood
+events in the NOAA Storm Events data versus storms without an event listing, to
+explore if storms with flood event listings tended to be  associated with
+higher streamflows at gages within the county.
 

### 2019-07-28 writing/ehp_results.tex
Work more on paper draft

@@
 
-\paragraph{Flood data validation}
+\paragraph{Flood data analysis}
 
-For flood data, we compared the flood status measurements, based on event
-listings in the NOAA Storm Events database, to streamflow measurements at
-\ac{US} Geological Survey gages within nine sample counties
-(Figure~\ref{fig:floodcomparison}). Each small plot in
-Figure~\ref{fig:floodcomparison} shows results for one of the sample counties.
-Each point represents a single tropical storm, and the point's position along
-the x-axis shows the highest daily total streamflow (cubic feet per second),
-summed across all identified streamgages in the county, for a five-day window
-centered on the day of the storm's closest approach to the county. The y-axis
-separates storms for which a flood event was reported in NOAA's Storm Events
-database for the county with a start date within the five-day window of the
-storm's closest approach to the county. The color of each point gives the
-percent of streamflow gages in the county with a daily streamflow that exceeded
-a threshold of flooding on any day during the five-day window.  Stream gage
-data generally indicated higher discharge in these counties on dates identified
-as flood events, particularly in counties with more reporting stream gages
-(e.g., Harris County, TX, with~10 reporting stream gages;
-Figure~\ref{fig:floodcomparison}). There were some cases, however, where the
-two flooding data sources were not completely consistent. 
-For example, there
+For flood data, we compared flood status, based on event listings in the NOAA
+Storm Events database, to streamflow measurements at \ac{US} Geological Survey
+gages within nine sample counties (Figure~\ref{fig:floodcomparison}). Each
+small plot in Figure~\ref{fig:floodcomparison} shows results for one of the
+sample counties.  Each point represents a single tropical storm, and the
+point's position along the x-axis shows the highest daily total streamflow
+(cubic feet per second), summed across all identified streamgages in the
+county, for a five-day window centered on the day of the storm's closest
+approach to the county. The y-axis separates storms for which a flood event was
+reported in NOAA's Storm Events database for the county with a start date
+within the five-day window of the storm's closest approach to the county. The
+color of each point gives the percent of streamflow gages in the county with a
+daily streamflow that exceeded a threshold of flooding on any day during the
+five-day window.
+
+For flood data, we compared flood status, based on event listings in the NOAA
+Storm Events database, to streamflow measurements at \ac{US} Geological Survey
+gages within nine sample counties (Figure~\ref{fig:floodcomparison}).  Across
+the sampled counties, streamgage data generally indicated higher discharge in
+on dates identified as flood events, particularly in counties with more
+reporting stream gages (e.g., Harris County, TX, with~10 reporting stream
+gages; Figure~\ref{fig:floodcomparison}). There were some cases, however, where
+the two flooding data sources were somewhat inconsistent.  For example, there
 were one or two tropical cyclones in several of the counties (Mobile County,
@@
 discharge across county streamflow gages was below the 25\textsuperscript{th}
-percentile of measurements during storms without a flood event listing. 
-For tropical cyclones that did not have an associated flood listing in 
+percentile of measurements taken during storms that did not have a flood event
+listing. For tropical cyclones that did not have an associated flood listing in
 the NOAA Storm Events database, most had lower total streamflow discharge
-across county gages compareed to storms with flood event listings, and in all
-cases but one none of the gages were over the flooding threshold. The exception
-was \ldots in Fulton County, GA. [More on which storm and why---I think there
-was already a flood event that had started earlier, so it missed the cutoff for
-the startdate being within the range.]
-
-[For discussion?] Based on these results, the two data sources were generally
-in agreement, but the analysis highlights the difficulty of assessing
-storm-related flooding at a multi-county, multi-year scale, as flooding can be
-very localized within a county or can \ldots .
+across county gages compared to storms with flood event listings, and in all
+cases but one all gages were below the flooding threshold.  The exception was
+one storm in Fulton County, GA (Ida in 2009). [More on which storm and why---I
+think there was already a flood event that had started earlier, so it missed
+the cutoff for the startdate being within the range.]
 
@@
 exposure metric. A few counties (Miami-Dade County, FL; Harris County, TX) were
-ranked in the top three counties of physical exposure for almost all exposure
-metrics (Figure~\ref{fig:topelecdependexposure}). However, there were key
-differences when using different exposure metrics. For example, Florida
-counties made up most of the counties counties with highest expected physical
-exposure based on wind- and tornado-based metrics, while over half of the
-counties with highest expected physical exposures for the flood-based metric
-were in Mid-Atlantic states (PA, NY).
+ranked in the top three counties for expected physical exposure in this
+subpopulation for almost all exposure metrics
+(Figure~\ref{fig:topelecdependexposure}). However, there were key differences
+across exposure metrics. For example, most of the top counties based on wind-
+and tornado-based metrics were in Florida, while over half of the counties with
+highest expected physical exposures for the flood-based metric were in
+Mid-Atlantic states (PA, NY).

### 2019-07-28 writing/anderson_ehp_supp.tex
Work more on draft

@@
 
+\begin{comment}
+
 \begin{figure}[tbhp!]
@@
 
+\end{comment}
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=0.9\linewidth]{rainexamples}
+\caption{Examples of storms where some storm-related rainfall occured
+         500 \si{\kilo\metre} or further from the storm's track. 
+	 The red line on each map shows the track of the storm. 
+	 The color of each county in the map gives the cumulative 
+	 precipitation in the county from two days before to one day after
+	 the storm's closest approach (in \si{\milli\metre}). The blue 
+	 outline identifies the collection of counties
+	 that were classified as ``exposed" based on the rainfall exposure 
+	 criteria (Table 1 of main text), which includes the constraint
+	 that the storm must have come within 500 \si{\kilo\metre} of the 
+	 county. }
+\label{fig:rainexamples}
+\end{figure}
+
 \begin{figure}[tbhp!]
@@
 \includegraphics[width=\linewidth]{othertopstorms}
-\caption{Counties exposed, based on different exposure metrics, to the storms
-with largest extent (as measured by the number of counties exposed based on any
-metric) from each of the clusters shown in the Jaccard heatmap in the main text
-(Figure 2).}
+\caption{Differences in the counties assessed as ``exposed'', based on different
+	exposure metrics, for a sample of storms. These sample storms were 
+	selected as the  storms with largest extent (as measured by the 
+	number of counties exposed based on any metric) from each of the 
+	clusters shown in the Jaccard heatmap in the main text (Figure 7).}
 \label{fig:othertopstorms}
@@
 \caption{Examples of storms for which exposure assessments differed
-substantially between rain-based and wind-based metrics. Counties shown in
-orange were classified as exposed to a given storm by a given metric; those in
-yellow were classified as not exposed. The storm's track is shown in red. For
-both Frances (2004) and Isidore (2002) (top panel), substantially more counties
-were assessed as exposed when measuring exposure based on rain rather than
-wind. Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel),
-substantially more counties were assessed as exposed based on wind compared to
-rain.}
+	substantially between rain-based and wind-based metrics. Counties 
+	shown in orange were classified as exposed to a given storm by a 
+	given metric; those in yellow were classified as not exposed. 
+	The storm's track is shown in red. For both Frances (2004) and 
+	Isidore (2002) (top panel), substantially more counties were assessed 
+	as exposed when measuring exposure based on rain rather than wind. 
+	Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel), 
+	substantially more counties were assessed as exposed based on wind 
+	compared to rain.}
 \label{fig:extentdisagreement}
@@
 \includegraphics[width=\linewidth]{averageexposureonlysupp}
-\caption{Average number of storm exposure per decade in \ac{US} counties for
-each exposure metric, based on years of for which data for all five exposure
-metrics are available (1996\,--\,2011).} 
+\caption{Average number of storm exposures per decade in U.S. counties for each 
+	exposure metric, limited analysis to years for which data on all five 
+	exposures were available (1996\,--\,2011). The criteria behind each of 
+	the five metrics is given in Table 1 of the main text. 	
+	} 
 \label{fig:averageexposuresupp}
@@
 
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=0.9\linewidth]{rainexamples}
-\caption{Examples of storms where the rain exposure classification method fails
-to capture full storm-related rainfall because of the distance constraint used
-in that assessment metric. The red line on each map shows the track of the
-storm. The color of each county in the map gives the cumulative precipitation
-(in \si{\milli\metre}). The blue outline identifies the collection of counties
-that were classified as ``exposed" based on our rainfall criteria (Table 1 of
-main text). }
-\label{fig:rainexamples}
-\end{figure}
-
 \clearpage

### 2019-07-28 writing/ehp_discussion.tex
Work more on draft

@@
 Previous research has highlighted the range of impacts that tropical cyclones
-can have in \ac{US} communities. However, studies have varied widely in how they
-assess exposure to tropical cyclones for exposure, health impact, and other
-impact studies, in some cases using distance to the tropical cyclone's track as
-a surrogate metric of exposure to a tropical cyclone's hazards. Here we found
+can have in \ac{US} communities. However, studies of tropical cyclone exposures
+and their impacts have varied widely in how they assess exposure. Here we found
 large differences in which counties are exposed to different hazards of
@@
 cyclone hazards to health impacts when conducting tropical cyclone
-epidemiological studies and basing exposure assessment on measurements of these
-hazards. To help with this, we provide extensive tropical cyclone
-hazard-related measurements for a large collection of historical Atlantic-basin
-cyclones, aggregated at the county level to align with the spatial scale at
-which much health-related data is collected.
+epidemiological studies, and then basing exposure assessment on these hazards.
+To help, we provide open-source tropical cyclone hazard-related measurements
+for a large collection of historical Atlantic-basin cyclones, aggregated at the
+county level to align with the spatial scale at which much health-related data
+is collected.

### 2019-07-28 writing/ehp_methods.tex
Work more on draft

@@
 storm-associated rainfall was then calculated as the sum of rainfall from two
-days before the storm's closest approach to a county to one day after.
+days before the storm's closest approach to a county to one day after, as 
+storm-related rains often precede the passing of the storm's center.
 

### 2019-07-28 writing/ehp_results.tex
Work more on draft

@@
 Storm Events database, to streamflow measurements at \ac{US} Geological Survey
-gages within nine sample counties (Figure~\ref{fig:floodcomparison}). Each
-small plot in Figure~\ref{fig:floodcomparison} shows results for one of the
-sample counties.  Each point represents a single tropical storm, and the
-point's position along the x-axis shows the highest daily total streamflow
-(cubic feet per second), summed across all identified streamgages in the
-county, for a five-day window centered on the day of the storm's closest
-approach to the county. The y-axis separates storms for which a flood event was
-reported in NOAA's Storm Events database for the county with a start date
-within the five-day window of the storm's closest approach to the county. The
-color of each point gives the percent of streamflow gages in the county with a
-daily streamflow that exceeded a threshold of flooding on any day during the
-five-day window.
+gages within nine sample counties (Figure~\ref{fig:floodcomparison}).
 

### 2019-07-28 R/wind_comparison_figure.R
Update wind comparison figure

@@ -24,19 +24,19 @@ fig <- wind_comp %>%
         summarize(perc_same = sum(cats_equal) / n(),
                   non_zero = sum(vmax_sust_ext != 0)) %>%
         filter(non_zero > 0) %>%
-        ggplot(aes(x = perc_same, y = reorder(storm_id, perc_same), color = non_zero)) +
-        geom_point(size = 2) +
+        ggplot(aes(x = perc_same, y = reorder(storm_id, perc_same), fill = non_zero)) +
+        geom_point(size = 2, shape = 21, color = "black") +
         scale_x_continuous(labels = scales::percent) +
         theme_classic() +
         labs(x = "% of counties classified in same wind category\nby modeled storm wind and Extended Best Tracks",
              y = "",
-             color = "Number of counties with modeled\nsustained winds of \u2265 34 knots",
+             fill = "Number of counties with modeled\nsustained winds of \u2265 34 knots",
              parse = TRUE
              ) +
-        viridis::scale_color_viridis(direction = -1, option = "B") +
+        viridis::scale_fill_viridis(direction = -1, option = "B") +
         theme(legend.position = "bottom")
 
-quartz.save(file = "figures/windcomparison.pdf", type = "pdf", width = 5.25, height = 7)
+quartz.save(file = "figures/windcomparison.pdf", type = "pdf", width = 5.25, height = 8)
 print(fig)
 dev.off()
 

### 2019-07-28 tables/stormseverity.tex
Update storm severity table

@@
 \centering
-\caption{Summary statistics for exposure metrics considered to assess county-level exposure to
-         tropical cyclones. A county exposure represents the exposure of a specific eastern US
-         county to a specific tropical cyclone. A county can be exposed to multiple tropical cyclones within
-         a year, and a tropical cyclone can create multiple county exposures under a given metric.
-         The median and interquartile range of county exposures per tropical cyclone is based on
+\caption{Summary statistics for the number of county tropical cyclone exposures under each metric.     
+	 The median and interquartile range of number of county exposures per tropical cyclone are based on
          the tropical cyclones for which at least one US county was exposed. The years for which

### 2019-07-28 R/storm_severity_table.R

@@ -166,11 +166,8 @@ total_sum %>%
          `Median (interquartile range) of county exposures per tropical cyclone` = sum,
          `Tropical cyclone with most counties exposures (# exposed counties)` = max_storm) %>%
   xtable(align = "cp{1.75cm}p{3cm}p{4cm}p{4cm}",
-         caption = "Summary statistics for exposure metrics considered to assess county-level exposure to
-         tropical cyclones. A county exposure represents the exposure of a specific eastern US  
-         county to a specific tropical cyclone. A county can be exposed to multiple tropical cyclones within
-         a year, and a tropical cyclone can create multiple county exposures under a given metric.
-         The median and interquartile range of county exposures per tropical cyclone is based on
+         caption = "Summary statistics for the number of county tropical cyclone exposures under each metric.     
+	 The median and interquartile range of number of county exposures per tropical cyclone are based on
          the tropical cyclones for which at least one US county was exposed. The years for which
          data are available for each metric are given in Table 1.",
          label = "tab:exposuresummaries",

### 2019-07-29 writing/ehp_discussion.tex
Update figure references

@@
 We investigated how well these data correspond with data from other available
-sources (Figures~\ref{fig:windcomparison}\,--\,\ref{fig:floodcomparison}
-and~S6\,--\,7). While generally in agreement with data from other sources,
+sources (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison}
+and~S2). While generally in agreement with data from other sources,
 there are a few caveats that should be considered when applying these data in
@@
 persisted well inland from landfall, like Hurricanes Sandy and Ike
-(Figures~\ref{fig:windcomparison} and\~S6). Since the wind radii are based on
+(Figures~\ref{fig:windcomparison} and~S2). Since the wind radii are based on
 \ldots, they have [advantages]; their main disadvantages is that they provide
@@
 for each storm based on these wind radii, in addition to the continuous modeled
-windspeeds, in the open-source data \cite{hurricaneexposuredata}.
+windspeeds, in the open-source data \parencite{hurricaneexposuredata}.
 
@@
 from \ac{US} Geographical Survey streamgages, with flood event listings more
-likely during storms which elevated stream flows. However, there are
-differences between the two flooding datasets, and these highlight both the
-difficulty of measuring flood exposure at the county level and inherent
-challenges in using the storm event database for epidemiologic exposure
-assessment.  The flood and tornado data came from the \ac{NOAA} Storm Events
-database, which, while a widely-used database of events maintained by
-\ac{NOAA}, is based on reports, and so may be prone to
-underreporting~\parencite{Ashley2008flood, Curran2000}, especially in less
-populated areas~\parencite{Witt1998, Ashley2007}, as well as to other reporting
-errors.
+likely during storms which elevated stream flows
+(Figure~\ref{fig:floodcomparison}. However, there are differences between the
+two flooding datasets, and these highlight both the difficulty of measuring
+flood exposure at the county level and inherent challenges in using the storm
+event database for epidemiologic exposure assessment.  The flood and tornado
+data came from the \ac{NOAA} Storm Events database, which, while a widely-used
+database of events maintained by \ac{NOAA}, is based on reports, and so may be
+prone to underreporting~\parencite{Ashley2008flood, Curran2000}, especially in
+less populated areas~\parencite{Witt1998, Ashley2007}, as well as to other
+reporting errors.
 
@@
 We found geographic patterns in tropical cyclones based on the different
-exposure metrics that. These patterns were not unexpected based on what is
-known about tropical cyclone hazards, but still highlight variations across
-metrics that are critical to consider in designing studies for tropical cyclone
-epidemiology, and they demonstrate the need for multi-hazard exposure datasets
-for tropical cyclone epidemiology. Wind-based exposures had a strong coastal
-pattern, which is consistent with the dramatic decrease in wind intensity that
-typically characterizes the landfall of tropical cyclones. Exposures by rain-
-and distance-based metrics often extended further inland compared to the
-wind-based metric, up to the Appalachian mountains. This agrees with previous
-research indicating that the Appalachian mountains' topography both enhances
-precipitation during tropical cyclones and provides hydrological conditions for
-severe flooding~\parencite{rees2001}.  Almost all tornado-based exposures were
-in southern coastal states, echoing previous findings that most noteworthy
-tropical cyclone-related tornadoes occur on the right-side of the tropical
-cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.  These
-exposure averages are somewhat limited as estimates of long-term frequencies,
-as tropical cyclones follow decadal patterns~\parencite{kossin2007more} likely
-not adequately captured in the available data. However, exposure assessments
-spanning over a century have found similar patterns for distance- and
-wind-based metrics~\textcite{zandbergen2009, kruk2010} [double-check that refs
-found similar patterns].
+exposure metrics (Figure~\ref{fig:averageexposures}). These patterns were not
+unexpected based on what is known about tropical cyclone hazards, but still
+highlight variations across metrics that are critical to consider in designing
+studies for tropical cyclone epidemiology, and they demonstrate the need for
+multi-hazard exposure datasets for tropical cyclone epidemiology. Wind-based
+exposures had a strong coastal pattern, which is consistent with the dramatic
+decrease in wind intensity that typically characterizes the landfall of
+tropical cyclones. Exposures by rain- and distance-based metrics often extended
+further inland compared to the wind-based metric, up to the Appalachian
+mountains. This agrees with previous research indicating that the Appalachian
+mountains' topography both enhances precipitation during tropical cyclones and
+provides hydrological conditions for severe flooding~\parencite{rees2001}.
+Almost all tornado-based exposures were in southern coastal states, echoing
+previous findings that most noteworthy tropical cyclone-related tornadoes occur
+on the right-side of the tropical cyclone track in Atlantic-basin \ac{US}
+storms~\parencite{moore2012}.  These exposure averages are somewhat limited as
+estimates of long-term frequencies, as tropical cyclones follow decadal
+patterns~\parencite{kossin2007more} likely not adequately captured in the
+available data. However, exposure assessments spanning over a century have
+found similar patterns for distance- and wind-based
+metrics~\textcite{zandbergen2009, kruk2010} [double-check that refs found
+similar patterns].
 
@@
 county-level exposure classifications for a tropical cyclone when comparing
-different exposure metrics. This suggests that the use of one of these metrics
-as a proxy for all storm-based hazards in epidemiological research would often
-be problematic, introducing exposure misclassification.
+different exposure metrics
+(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard}  and S4\,--\,S5). This
+suggests that the use of one of these metrics as a proxy for all storm-based
+hazards in epidemiological research would often be problematic, introducing
+exposure misclassification.
 
@@
 agreement between metrics were unusually high (e.g., Floyd in~1999, Irene in
-2011, Hannah in~2008, Bertha in~1996).  Hurricanes Floyd in~1999 and Irene
-in~2011 both made their first \ac{US} landfall in North Carolina at minor
-hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
-eastern coast of the \ac{US} north through New England, bringing large rainfall
-to much of the eastern coast from North Carolina north and causing extensive
-inland flooding in North Carolina (Floyd) and New England
-(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
-in~2008 and Bertha in~1996 followed the eastern coastline north of North
-Carolina in a pattern similar to Floyd and Irene. For this set of storms, the
-tropical cyclones' persistent proximity to water may have helped maintain wind
-speeds in similar patterns to rain and distance exposures, resulting in this
-moderate to good agreement among exposure assessments based on different
-metrics.
+2011, Hannah in~2008, Bertha in~1996; Figure~\ref{fig:jaccard}).  Hurricanes
+Floyd in~1999 (Figure~S4) and Irene in~2011 both made their first \ac{US}
+landfall in North Carolina at minor hurricane windspeeds (Category~2 and~1,
+respectively) and then skimmed the eastern coast of the \ac{US} north through
+New England, bringing large rainfall to much of the eastern coast from North
+Carolina north and causing extensive inland flooding in North Carolina (Floyd)
+and New England (Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.
+Hurricanes Hannah in~2008 and Bertha in~1996 followed the eastern coastline
+north of North Carolina in a pattern similar to Floyd and Irene. For this set
+of storms, the tropical cyclones' persistent proximity to water may have helped
+maintain wind speeds in similar patterns to rain and distance exposures,
+resulting in this moderate to good agreement among exposure assessments based
+on different metrics.
 

### 2019-07-29 writing/ehp_methods.tex
Update figure references

@@
 Lee in 2011), some storm-related rains may be missed because of this distance
-constraint (Figure~S7).
+constraint (Figure~S1).
 

### 2019-07-29 writing/ehp_results.tex
Update figure references

@@
 storm's most severe winds at landfall but then underestimated the extent of
-34\,--\,49.9 \si{\knot} storm winds further inland (Figure~S6). For researchers
+34\,--\,49.9 \si{\knot} storm winds further inland (Figure~S2). For researchers
 who would like to conduct sensitivity analysis between these two sets of wind
@@
 analysis was restricted to years with all exposure data available
-(1996\,--\,2011; Figure~S5).
+(1996\,--\,2011; Figure~S3).
 
@@
 and Georgia that were not identified as exposed to Ivan based on any other
-metric. Figure~S3 provides similar maps for four other example tropical
+metric. Figure~S4 provides similar maps for four other example tropical
 cyclones (selected because they exposed many \ac{US}  counties based on at
-least one metric), and Figure~S4 provides four examples of tropical cyclones in
+least one metric), and Figure~S5 provides four examples of tropical cyclones in
 which the disagreement in exposure assessment between rain- and wind-based
@@
 (Figure~\ref{fig:ivanexposure}) and Hurricanes Floyd in~1999, Lee in~2011, Cindy
-in~2005, and Katrina in~2005 (Figure~S3).
+in~2005, and Katrina in~2005 (Figure~S4).
 \end{comment}
@@
 counties were determined to be exposed to the tropical cyclone for two or more
-of the metrics considered.  For Floyd in~1999 (Figure~S3) and Irene in~2011,
+of the metrics considered.  For Floyd in~1999 (Figure~S4) and Irene in~2011,
 for example, county-level classification agreed moderately to well (Jaccard

### 2019-07-29 writing/ehp_methods.tex
Add some on wind radii and Ida

@@
 To validate these modeled county-level wind speed estimates, we compared them
-to county-level maximum sustained surface wind speeds based on the wind radii
-estimates in \ac{HURDAT2}, which are based on \ldots. The wind radii values
-have been included in \ac{HURDAT2} since 2004 [?] and estimate the distance
-from the storm's center to the furthest point with winds at or above winds of a
-certain speed in each four quadrants of the storm. They give estimates of these
-maximum radii for three thresholds of maximum surface wind speeds: 64,~50,
-and~34 \si{\knot}.  They therefore allow for the classification of counties
-into four categories of maximum sustained wind speeds: $<$34~\si{\knot};
+to county-level maximum sustained surface wind speeds estimated based on the
+wind radii estimates in \ac{HURDAT2}, which have been included in \ac{HURDAT2}
+since 2004 [ref]. Reanalysis values for these wind radii estimate the maximum
+extent from the storm's center to winds at or above winds of a certain speed in
+each four quadrants of the storm, based on data from satellites and, where
+available, ground-based and aircraft reconnaissance data [ref].  Wind radii are
+estimated for three thresholds of maximum surface wind speeds: 64,~50, and~34
+\si{\knot}.  They therefore allow for the classification of counties into four
+categories of maximum sustained wind speeds: $<$34~\si{\knot};
 34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}. We
@@
 was within~85\% of the maximum radius for that wind speed in its quadrant of
-the storm [clarify why 85\%]. We compared these wind radii-based estimates with
-the modeled wind speed estimates for the study storms since 2004 for which at
-least one study county had a sustained wind of~$\ge$34 \si{\knot} (based on the
-\ac{HURDAT2} wind radii). For each of these storms, we calculated the percent
-of study counties that were classified in the same wind speed category.
+the storm, as previous research has found  ... [ref]. We compared these wind
+radii-based estimates with the modeled wind speed estimates for the study
+storms since 2004 for which at least one study county had a sustained wind
+of~$\ge$34 \si{\knot} (based on the \ac{HURDAT2} wind radii). For each of these
+storms, we calculated the percent of study counties that were classified in the
+same wind speed category.
 

### 2019-07-29 writing/ehp_results.tex
Add some on wind radii and Ida

@@
 cases but one all gages were below the flooding threshold.  The exception was
-one storm in Fulton County, GA (Ida in 2009). [More on which storm and why---I
-think there was already a flood event that had started earlier, so it missed
-the cutoff for the startdate being within the range.]
+one storm in Fulton County, GA (Ida in 2009). This storm occurred in November
+2009, following a month with historic rainfall and flooding in Georgia [ref].
+Therefore, in this case the flooding associated with Ida was incorporated into
+an ongoing flood event listing, with a start date well before the five-day
+window used to temporally match storm event listings with tropical cyclone
+tracks used here.
 

### 2019-10-09 tables/stormseverity.tex
Work on paper, use 34 knot threshold

@@
-% latex table generated in R 3.4.3 by xtable 1.8-2 package
-% Wed Oct 10 12:11:35 2018
+% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Tue Oct  8 14:03:24 2019
 \begin{table}[ht]
@@
   \midrule
-Distance & 433 (224, 503) & 84 (23, 173) & Beryl, 1994 (330) \\ 
 Rain & 401 (152, 553) & 68 (14, 146) & Frances, 2004 (464) \\ 
-  Wind & 224 (83, 353) & 34 (8, 80) & Ike, 2008 (355) \\ 
-  Flood & 213 (84, 249) & 26 (6, 57) & Ivan, 2004 (317) \\ 
+  Wind & 143 (51, 232) & 25 (14, 58) & Ike, 2008 (251) \\ 
+  Flood & 213 (84, 248) & 26 (6, 58) & Ivan, 2004 (318) \\ 
   Tornado & 53 (16, 43) & 7 (2, 16) & Ivan, 2004 (91) \\ 

### 2019-10-09 writing/darlings_discussion.tex
Work on paper, use 34 knot threshold

--- 2019-07-28 writing/ehp_discussion.tex
@@
-\section*{Discussion}
-
-Tropical cyclones impact public health in many ways, and epidemiological
-studies of the health risk and impacts associated with tropical cyclone
-exposures could help improve preparedness for and response to future tropical
-cyclones.  However, tropical cyclones are multi-hazard events, making it
-complicated to measure exposure. Here, we provide open-source county-level data
-on multiple storm exposure metrics, and we find that county-level tropical
-cyclone exposure assessments vary substantially when using a distance-based
-metric versus four hazard-based metrics, as well as among different
-hazard-based metrics. Our results can inform exposure assessment for future
-county-level studies of the health risks and impacts associated with tropical
-cyclones exposures, and the open-source data and software we have created
-provides multi-hazard, county-level exposure data for such studies.
-
-\subsection*{Exposure assessment data and software}
-
-In creating this hurricane exposure dataset, we aimed for data that are
-available for all eastern \ac{US} counties. In addition to providing tropical
-cyclone exposure metrics for individual hazards, this data and its associated
-software allow users not only to access measurements for single hazards, but
-also to create tropical cyclone exposure profiles based on multiple hazards or
-craft exposure indices that combine hazard
-metrics~\parencite{chakraborty2005population, peduzzi2009assessing}. This
-ability can be critical, as different hazards of tropical cyclones often act
-synergistically in causing impacts~\parencite{smith2009}.
-
-\begin{comment}
 To assist with future tropical cyclone studies, we created and published open
@@
 linear models) that incorporate multiple tropical cyclone hazard measurements.
-\end{comment}
 
-We investigated how well these data correspond with data from other available
-sources (Figures~\ref{fig:windcomparison}\,--\,\ref{fig:floodcomparison}
-and~S6\,--\,7). While generally in agreement with data from other sources,
-there are a few caveats that should be considered when applying these data in
-future studies. The rainfall data are from re-analysis data are generally
-well-correlated with observed ground-based station data, but may sometimes
-oversmooth very high rainfall values compared to ground-based observations
-(Figure~\ref{fig:raincomparison}). When rainfall data is used to create binary
-exposure classifications, this disagreement is unlikely to influence results,
-as both data sources agree in identifying these as storms with high rainfall,
-but it may be important to consider for cases that include rainfall as a
-continuous measurement.
-
-The wind data are based on modeled, rather than observed, values, and while the
-modeled wind data generally agree well with post-analysis maximum wind radii
-from \ac{HURDAT2}~\parencite{landsea2013} (Figure~\ref{fig:windcomparison}),
-there were a few storms where the two datasets did not agree as well ($<$90\%
-of counties classified into the same wind categories by the two sources of wind
-estimates). These storms were exceptionally large systems for which high winds
-persisted well inland from landfall, like Hurricanes Sandy and Ike
-(Figures~\ref{fig:windcomparison} and\~S6). Since the wind radii are based on
-\ldots, they have [advantages]; their main disadvantages is that they provide
-categorical, rather than continuous, estimates of wind speed, and in many cases
-researchers might prefer more precise estimates of storm wind speeds. However,
-since the wind radii data may be useful in some cases as either a primary
-metric or for sensitivity analysis, we have include county-level wind estimates
-for each storm based on these wind radii, in addition to the continuous modeled
-windspeeds, in the open-source data \cite{hurricaneexposuredata}.
-
-For the flooding events data, we found it often corresponded with measurements
-from \ac{US} Geographical Survey streamgages, with flood event listings more
-likely during storms which elevated stream flows. However, there are
-differences between the two flooding datasets, and these highlight both the
-difficulty of measuring flood exposure at the county level and inherent
-challenges in using the storm event database for epidemiologic exposure
-assessment.  The flood and tornado data came from the \ac{NOAA} Storm Events
-database, which, while a widely-used database of events maintained by
-\ac{NOAA}, is based on reports, and so may be prone to
-underreporting~\parencite{Ashley2008flood, Curran2000}, especially in less
-populated areas~\parencite{Witt1998, Ashley2007}, as well as to other reporting
-errors.
-
-\begin{comment}
 [For discussion?] Based on these results, the two data sources were generally
@@
 very localized within a county or can \ldots .
-\end{comment}
 
-While these are important caveats for the data, we selected these data sources
-as among the best currently available for measuring each of these hazards at a
-multi-county, multi-year scale.
-
-\subsection*{Patterns in tropical cyclone exposures}
-
-We found geographic patterns in tropical cyclones based on the different
-exposure metrics that. These patterns were not unexpected based on what is
-known about tropical cyclone hazards, but still highlight variations across
-metrics that are critical to consider in designing studies for tropical cyclone
-epidemiology, and they demonstrate the need for multi-hazard exposure datasets
-for tropical cyclone epidemiology. Wind-based exposures had a strong coastal
-pattern, which is consistent with the dramatic decrease in wind intensity that
-typically characterizes the landfall of tropical cyclones. Exposures by rain-
-and distance-based metrics often extended further inland compared to the
-wind-based metric, up to the Appalachian mountains. This agrees with previous
-research indicating that the Appalachian mountains' topography both enhances
-precipitation during tropical cyclones and provides hydrological conditions for
-severe flooding~\parencite{rees2001}.  Almost all tornado-based exposures were
-in southern coastal states, echoing previous findings that most noteworthy
-tropical cyclone-related tornadoes occur on the right-side of the tropical
-cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.  These
-exposure averages are somewhat limited as estimates of long-term frequencies,
-as tropical cyclones follow decadal patterns~\parencite{kossin2007more} likely
-not adequately captured in the available data. However, exposure assessments
-spanning over a century have found similar patterns for distance- and
-wind-based metrics~\textcite{zandbergen2009, kruk2010} [double-check that refs
-found similar patterns].
-
-\begin{comment}
 However, these frequency maps, together with evidence from
@@
 flooding).
-\end{comment}
 
-\subsection*{Agreement between exposure metrics}
-
-We found that there was typically only moderate or low agreement between
-county-level exposure classifications for a tropical cyclone when comparing
-different exposure metrics. This suggests that the use of one of these metrics
-as a proxy for all storm-based hazards in epidemiological research would often
-be problematic, introducing exposure misclassification.
-
-\begin{comment}
 For example, distance from a tropical cyclone's track is relatively easy to
@@
 armstrong1998effect}.
-\end{comment}
 
-These findings align with previous results from atmospheric science and related
-fields on the characteristics of tropical cyclones. While tropical cyclone
-rainfall and windspeed are often well-correlated when the tropical cyclone is
-over water~\parencite{cerveny2000}, this relationship often does not remain as
-strong once the hurricane has made landfall~\parencite{jiang2008}.  Fast-moving
-tropical cyclones bring higher risks of dangerous winds
-inland~\parencite{kruk2010}, while slow-moving tropical cyclone are likely to
-bring more rain~\parencite{rappaport2000} and cause more damage because of
-sustained hazardous conditions~\parencite{rezapour2014}. Further, while the
-likelihood and extent of flooding during a tropical cyclone is related to the
-tropical cyclone's rainfall, it is also driven by factors like top soil
-saturation and the structure of the water basin's drainage
-network~\parencite{chen2015, rees2001}.
-
-However, we did find a small set of tropical cyclones for which for which
-agreement between metrics were unusually high (e.g., Floyd in~1999, Irene in
-2011, Hannah in~2008, Bertha in~1996).  Hurricanes Floyd in~1999 and Irene
-in~2011 both made their first \ac{US} landfall in North Carolina at minor
-hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
-eastern coast of the \ac{US} north through New England, bringing large rainfall
-to much of the eastern coast from North Carolina north and causing extensive
-inland flooding in North Carolina (Floyd) and New England
-(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
-in~2008 and Bertha in~1996 followed the eastern coastline north of North
-Carolina in a pattern similar to Floyd and Irene. For this set of storms, the
-tropical cyclones' persistent proximity to water may have helped maintain wind
-speeds in similar patterns to rain and distance exposures, resulting in this
-moderate to good agreement among exposure assessments based on different
-metrics.
-
-In some cases, exposure misclassification from using one of these exposure
-metrics to capture one or more of the other storm hazards may be differential
-(i.e., associated with the outcome of interest or with factors associated with
-risk of the outcome of interest). For example, tropical cyclone wind exposures
-tend to be concentrated in counties near the coast, since most tropical
-cyclones rapidly decrease in their sustained windspeed following landfall.
-However, tropical cyclone exposures based on distance can extend well inland,
-following the tropical cyclone's tracks, but may not adequately capture all
-wind-exposed counties near the coast. In this case, if the
-etiologically-relevant exposure is high wind but exposure is classified based
-on distance, the probability of being misclassified as unexposed would be
-higher in coastal counties, while the probability of being misclassified as
-exposed would be higher in inland counties. If coastal counties differ from
-inland counties in either the outcome of interest or in factors associated with
-risk of that outcome, differential exposure misclassification would
-exist~\parencite{savitz2016interpreting}. Such differential exposure
-misclassification can bias estimates of tropical cyclone effects either towards
-the null (estimating a lower or null association compared to the true
-association that exists) or away from the null (estimating a larger association
-than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.
-
-\begin{comment}
 The use of a single hazard-based metric (e.g., wind) could cause similar
@@
 multiple hazards of the tropical cyclone.
-\end{comment}
 
-\subsection*{Case study}
+In creating this hurricane exposure dataset, we aimed for data that are
+available for all eastern \ac{US} counties.
 
-The conducted case study demonstrates how differences in exposure assessment
-across storm exposure metrics can result in differing estimates of expected
-physical exposure to tropical cyclones among a population
-(Figure~\ref{fig:topelecdependexposure}).  Previous assessments of physical
-exposure to tropical cyclones have often focused on solely storm winds and
-storm proximity. For example, a global study of exposure and vulnerability to
-natural hazards used a combination of windspeed above a certain threshold and
-distance from the storm's track to assess
-exposure~\parencite{peduzzi2009assessing}. While many storm impacts might be
-most strongly linked to wind hazards, there is a growing recognition of the
-potential risks of adverse health outcomes and property damage from rain- and
-flood-related tropical cyclone hazards in the \ac{US}~\parencite{smith2009}.
-Such assessments may underemphasize potentially dangerous exposures to certain
-regions of the \ac{US}, particularly inland locations. In fact, based on our
-assessment of physical exposures to tropical cyclones among
-electricity-dependent Medicare beneficiaries, the highest expected physical
-exposure under any metric was measured for Philadelphia County, PA, under the
-flood metric (Figure~\ref{fig:topelecdependexposure}).  While this county has a
-similar-sized population of electricity-dependent Medicare beneficiaries as
-Miami-Dade County, FL (Figure~\ref{fig:topelecdependexposure}), the frequency
-of storm-related flood events in this inland county results in a very high
-expected physical exposure among this subpopulation, while this county is not
-included in the top ten counties for physical exposure based on most of the
-other exposure metrics.
+($<$90\% of counties classified into the same wind categories by the two
+sources of wind estimates)
 
-\subsection*{Conclusions}
+The wind radii, unlike the modeled wind speeds, are based on a storm-specific
+re-analysis of available data. They therefore have the advantage that they can
+capture unusual patterns in specific storms; their main disadvantage is that
+they provide categorical, rather than continuous, estimates of wind speed, and
+in many cases researchers might prefer more precise estimates of storm wind
+speeds. However, since the wind radii data may be useful in some cases as
+either a primary metric or for sensitivity analysis, we have included
+county-level wind estimates for each storm based on these wind radii, in
+addition to the continuous modeled windspeeds, in the open-source data
+\parencite{hurricaneexposuredata}.
 
-Previous research has highlighted the range of impacts that tropical cyclones
-can have in \ac{US} communities. However, studies of tropical cyclone exposures
-and their impacts have varied widely in how they assess exposure. Here we found
-large differences in which counties are exposed to different hazards of
-tropical cyclones and that distance is, at best, a moderate, and often a very
-poor, surrogate for exposure to the specific tropical cyclone hazards of high
-wind, extreme rainfall, flooding, and tornadoes. Use of distance as a surrogate
-for any of these hazards could lead to exposure misclassification and, in the
-case of tropical cyclone risk and impact studies, including epidemiological
-studies, result in biased estimates.  Similarly, if such studies disagree on
-their findings, it could result from the often poor agreement between exposure
-classifications based on these different metrics, making studies that use
-different metrics hard to compare or combine in meta-analyses. Our findings
-highlight the importance of clarifying the potential pathway from tropical
-cyclone hazards to health impacts when conducting tropical cyclone
-epidemiological studies, and then basing exposure assessment on these hazards.
-To help, we provide open-source tropical cyclone hazard-related measurements
-for a large collection of historical Atlantic-basin cyclones, aggregated at the
-county level to align with the spatial scale at which much health-related data
-is collected.
+between county-level classifications of tropical cyclone exposure was typically
+low to moderate across the  five exposure metrics we considered
+
+If measurements of one storm hazard are used as a proxy to capture exposure to
+a different hazard, or a non-hazard proxy like distance to the storm track is
+used to assess exposure to storm hazards, the resulting exposure
+misclassification could be differential (i.e., associated with the outcome of
+interest or with factors associated with risk of the outcome of interest).

### 2019-10-09 writing/darlings_intro.tex
Work on paper, use 34 knot threshold

Much of the
epidemiological research to date on the health impacts of tropical cyclones has, like these studies,
focused on a single storm (e.g., [refs]), or has relied on death and injury
counts for which a link to the storm was identifiable through mention on a
death certificate or other information about the death [refs], an approach that
may miss many of the health impacts of storms, particularly from non-accidental
causes, as evidenced with Hurricane Maria [ref for this?].

To date, epidemiological research has largely focuses on single storms, as with
these studies, or has leveraged fatality counts from diaster surveillance
activities. Single-storm studies clarify the health impacts of a specific
storm, but are limited in extrapolating what risks might be expected under
future storms or which characteristics of a storm modify its likely health
impacts.  Studies based on disaster surveillance have helped in identifying
large-scale patterns in the health impacts of tropical cyclones, especially for
accidental deaths and injuries. However, disaster surveillance can vary from
storm to storm or across agencies estimating fatality and injury tolls, and
this inconsistency makes counts from disaster surveillance difficult to
compare. Further, disaster surveillance activities are typically not 
designed in a way that captures increased risks for non-accidental outcomes. 
For example, details on the circumstances of death are often needed for 
a coroner or medical examiner to note a connection with a disaster on a 
death certificate, but the types of deaths that trigger death scene investigations
are largely accidental causes (e.g., drowning, carbon monoxide poisoning, 
motor vehicle accident, electrocution).

A number of studies have investigated how tropical cyclones affect human
health. These have largely taken two paths. First, many studies focus on a
specific storm, often one that was severe with clear human impacts. For
example, studies have explored how rates of ... changed during and after
Hurricane Sandy, which hit New York and New Jersey in 2012. 
Other studies have explored the health impacts of Hurricane Katrina in 
Mississippi and Louisiana in 2005. These studies show that tropical cyclones
can cause important health impacts from a variety of causes, but they say 
little about how less severe but more frequent tropical cyclone exposures 
might impact human health.

There is a second main path for tropical cyclone epidemiology studies to date.
This has been to leverage data from disaster surveillance activities, in which
[state officials], aid organizations, or [FEMA] estimate the death and injury
tolls from a disaster. These estimated health impacts are recorded across many
tropical cyclones (and other storm or disaster events) across databases like
the NOAA Storm Events database, [EM-DAT], and [SHELDUS]. Studies that use this
data can show large-scale patterns over long time scales. For example, studies
in this vein have shown that average death tolls have decreased dramatically
over the last century for the community where a storm makes landfall, and that
storms can cause numerous drowning deaths in communities well inland from
landfall. However, these studies have two key limitations. First, disaster
surveillance does not have clear standards for counting deaths and injuries
during a disaster---in fact, one recent review found that estimates of death
tolls can vary widely for the same storm across the agencies and organizations
attempting to count deaths. This review found that the estimated death tolls
for Hurricane Ike, for example, included an estimate of [x] deaths from FEMA,
[x] from ..., and [x] as recorded in NOAA's Storm Events database. Similar
inconsistency is likely in estimates made for different storms, particularly
ones that are far apart in time or space. This inconsistency complicates
attempts to link storm characteristics with the type and scale of health
impacts that the storm is likely to cause. The second limitation is that
studies that are based on disaster surveillance counts are likely to miss many
of the impacts from non-accidental causes. For deaths, for example, a medical
examiner or coroner may need to have details on the circumstances and
environment of the death to be able to add a note linking a death to a disaster
on a death certificate, and this death certificate information is a key source
in generating disaster surveillance death toll estimates. However, death scene
investigations are typically only triggered for certain kinds of deaths, mostly
those from accidental causes like drowning, motor vehicle accidents,
electrocution, carbon monoxide poisonings, suicides, and homicides. If this
risk of death from common, non-accidental causes is elevated, this association
is likely not fully captured by disaster surveillance activities.

Given these limitations in the existing literature on tropical cyclone
epidemiology, much can be learned by investigating how the risk of health
outcomes change in the period surrounding a storm's landfall through examining
patterns in administrative data, like all-cause mortality counts across a
community, or insurance claims.  This analysis can help estimate how storms
impact non-accidental mortality and morbidity and, for accidental causes, can
provide an estimate of impacts that may be more consistent across time and
space that estimated tolls from disaster surveillance. Conversely, multi-storm
studies using administrative data will lack the ability to explore the details
of specific deaths, since they can estimate the community-wide change in risk
but not idea which specific deaths were the ones contributing to that increased
risk. Studies based on community-wide administrative data should be considered
a complement to continuing research based on thorough studies of single storms
and on studies that leverage data from disaster surveillance activities. In
fact, one study examined how [heat waves].

For other types of ambient exposures, administrative
data or vital data records have allowed the exploration of changing rates in a
variety of health outcomes, accidental and non-accidental, in association with
changes in ambient exposures.

However, tropical cyclones are multi-hazards events, and a storm's hazards are
not perfectly aligned with distance from the storm's track. Recent storms have
demonstrated the wide variation possible in these hazards; Hurricanes
Michael and Maria brought extreme storm-related winds, for
example, while Hurricanes Florence and Harvey were more notable for extreme
rain and flooding [refs]. Depending on the health outcome considered, some
storm hazards might be more plausible as a risk factor than others. Respiratory
outcomes in the months after a storm, for example, might be increased through a
pathway of heightened mold exposure from flooding, while heat stroke and other
heat-related outcomes in the days immediately after a storm might follow from a
pathway of extreme winds causing power outages, resulting in loss of air
conditioning [refs].

Geographical patterns can differ for these different
storm hazards, and so exposure assessment based on one hazard could create
exposure misclassification if the pathway for health risks is through a
different storm hazard.

Tropical cyclones, including hurricanes, can have disastrous impacts in the
\ac{US}. A tropical cyclone's high winds can bring health risks
and property damage through structural damage of houses and other buildings,
falling trees, and wind-borne debris~\parencite{rappaport2000} and cause power
outages~\parencite{liu2005, han2009}, which introduce a number of threats to human
and ecological health, including water quality risks if the outage affects
wastewater treatment plants~\parencite{mallin2006}.  Other risks can exist without
severe wind; for example, one study found that most of the direct
hurricane-related deaths in the \ac{US} between~1970 and~1999 occurred in cases
when wind was below hurricane strength, including for Tropical Storms Charley
in~1998 and Alberto in~1994~\parencite{rappaport2000}.  Tropical cyclones can
produce excessive rain, especially in certain topographies (e.g., near
mountains), so counties well inland sometimes experience more extreme rain than
coastal counties. Flood risks from tropical cyclones can result from this rain,
although the two risks are not perfectly correlated~\parencite{chen2015}. In the
\ac{US}, over half of hurricane-related direct deaths from~1970 to~1999 from
Atlantic basin storms were a result of freshwater flooding~\parencite{rappaport2000}. 
Flooding can also degrade water quality~\parencite{mallin2006}, which can threaten both 
human and ecological health.

To measure tropical cyclone exposure for either investigations of exposure
patterns or for estimation of tropical cyclone risks and impacts, including
those related to human health, studies have varied widely in how they determine
tropical cyclone exposure (some examples shown in Figure~S1). While some
studies have measured exposure based on specific hazards of the tropical
cyclone (e.g., wind, rain), many have used distance from the tropical cyclone's
central track as a surrogate in identifying exposure to the hazards of a
tropical cyclone (e.g.,~\textcite{czajkowski2011, tansel2010, kinney2008,
caillouet2008increase}).

Distance from the tropical cyclone's central track can be easily measured using
widely-available hurricane tracking data. There are, however, a number of
limitations to using distance to assess exposure to tropical cyclones, and

Geographical patterns likely differ for these different storm hazards, and so
exposure assessment based on one hazard could create exposure misclassification
if the pathway for health risks is through a different storm hazard.

There are reasons to suspect that
exposure classifications based on tropical cyclone hazards like wind and rain
may disagree with distance-based classifications, based only on distance from
the storm's path. Further, an assessment based on one storm hazard (e.g., wind)
might disagree with assessments based on other hazards (e.g., rain or
flooding).

[Previous studies,] which found tropical cyclones can increase risk not only
for accidental health outcomes (e.g., drowning, crushing deaths and injuries,
electrocution, carbon monoxide poisoning [refs]) but also for non-accidental,
non-communicable diseases [refs].

### 2019-10-09 writing/darlings_methods.tex
Work on paper, use 34 knot threshold

These data typically give measurements of
the tropical cyclone center's location [and central pressure / maximum
windspeed?] at~6-\si{\hour} intervals at synoptic times (i.e., 6:00~am,
12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}); some landfalling tropical cyclones
have an additional observation at the time of landfall~\parencite{landsea2013}.

, using the Great Circle %
(WGS84 ellipsoid) method

We selected re-analysis data because of
this completeness.

We selected re-analysis data over other data
sources as it included data for every county and every day through a continuous
period of the study (1988\,--\,2011)~\parencite{alhamdan2014environmental,
cdcwonder}. By comparison, observations from ground-based monitoring networks
had missing values spatially (i.e., for some study counties), temporally (for
some days), or both.

We estimated daily rainfall for each study county by aggregating
hourly~1/8\si{\degree} gridded data from the \ac{NLDAS-2} precipitation data
files~\parencite{rui2013nldas}. These data integrate satellite-based and
land-based monitoring, applying a land-surface model to create a reanalysis
dataset that is spatially and temporally complete across the continental
\ac{US}~\parencite{rui2013nldas, alhamdan2014environmental}. To aggregate to
county-level, the hourly data at each grid point were summed to create a daily
rainfall total, and these grid point rainfall totals were then averaged across
all grid points within a county's~1990 \ac{US} Census
boundaries~\parencite{alhamdan2014environmental, cdcwonder}. These aggregated
county-level daily precipitation data are also available through the \ac{US}
Centers for Disease Control's WONDER database [ref] and are available for all
study counties for 1988\,--\,2011.

When combined with estimates of vulnerability of a
population to a natural hazard, such measurements of physical exposure can be
used to calculate risk of human losses from the
hazard~\parencite{peduzzi2009assessing}.

These measurements
captured the distance between the storm's central track and each study county
throughout the storm tracking period.

To measure storm-associated rainfall in each county, we used re-analysis data.
Re-analysis data integrate data from multiple sources, including
satellite-based and ground-based monitoring, using a land-surface model to
combine data in a dataset that is spatially and temporally complete over a
certain period [ref]. By comparison, observations from ground-based monitoring
networks typically have missing values spatially (i.e., for some study
counties), temporally (for some days), or both.

We included more days prior to the storm because
storm-related rains often precede the passing of the storm's center [ref].

Ground-based observations of wind speed are problematic for measuring exposure
to tropical cyclones, as instruments often fail at high wind speed, while
reanalysis data---often available at hourly or higher resolution---can lack a
fine enough temporal resolution to capture wind extremes associated with a
tropical cyclone. 
A tropical cyclone's rainfall can can begin a day or more before a
storm and last a day or more after, so this functionality makes it possible to
quantify the cumulative precipitation throughout several days surrounding the
storm's approach.

### 2019-10-09 writing/darlings_results.tex
Work on paper, use 34 knot threshold

--- 2019-07-29 writing/ehp_results.tex
@@
-\section*{Results}
-
-\subsection*{Exposure assessment data, data validation, and software}
-
-\paragraph{Open-source dataset}
-
-We created and published open source data with multi-year, county-level
-tropical cyclone exposure data for eastern \ac{US}
-counties~\parencite{hurricaneexposuredata}. These data include continuous
-county-level measurements for the closest distance of each storm, cumulative
-rainfall, and maximum sustained surface wind speed. These continuous metrics
-can be used for classifying counties as exposed or unexposed, using thresholds
-selected by the user, or can be used as continuous metrics. The dataset also
-include binary data on flood and tornado events associated with each storm in
-each county. With these data, we published software tools to explore and map
-the data and to integrate it with human health
-datasets~\parencite{hurricaneexposure}.
-
-We explored potential limitations in these data by comparing them with data
-from other available sources.
-
-\paragraph{Rainfall data validation}
-
-For estimates of storm-associated rainfall, we compared estimates from NLDAS-2
-re-analysis precipitation data (included as the rainfall estimates in the
-open-source data) with ground-based observations in nine sample counties
-(Figure~\ref{fig:raincomparison}). Within these counties, storm-related
-rainfall measurements were well-correlated between the two data sources, with
-rank correlations (bottom right of each graph in
-Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some evidence
-that the reanalysis data may tend to underestimate rainfall totals in storms
-with extremely high rainfall, based a few heavy-rainfall storms in Harris
-County, TX, Mobile County, AL, Charleston County, SC, and Wake County, NC
-(Figure~\ref{fig:raincomparison}). However, even in these cases, it was rare
-for a storm to be classified differently (based on the precipitation threshold
-of 75~\si{\milli\metre} we used for exposure classification for later analysis)
-based on the source of precipitation data---horizontal and vertical lines in
-each small plot in Figure~\ref{fig:raincomparison} show the threshold of
-75~\si{\milli\metre}, so storms in the lower left and upper right quadrants
-would be classified the same (``exposed'' or ``unexposed'') regardless of the
-precipitation data source, while storms in the upper left and lower right
-quadrants would be classified differently.
-
-\paragraph{Wind speed data validation}
-
-For maximum sustained surface wind speed estimates, we found that the modeled
-estimates, which we included as the primary wind exposure metric in the open
-source data and used for further analysis in this study, generally agreed well
-with estimates based on the wind radii reported in \ac{HURDAT2}. For most
-storms, $\ge$95\% of counties were assigned the same category of wind speed
-($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot};
-$\ge$64~\si{\knot}) by both data sources (Figure~\ref{fig:windcomparison}).
-Disagreement was limited to large storms that were characterized by high winds
-well inland, like Hurricanes Sandy in 2012 and Ike in 2008. For these two
-storms, the modeled wind speed values somewhat overestimated the extent of the
-storm's most severe winds at landfall but then underestimated the extent of
-34\,--\,49.9 \si{\knot} storm winds further inland (Figure~S2). For researchers
-who would like to conduct sensitivity analysis between these two sets of wind
-data, we have included estimates from the \ac{HURDAT2} wind radii as a
-secondary measure of wind speeds in the open-source dataset
-\parencite{hurricaneexposuredata}.
-
-\paragraph{Flood data analysis}
-
-For flood data, we compared flood status, based on event listings in the NOAA
-Storm Events database, to streamflow measurements at \ac{US} Geological Survey
-gages within nine sample counties (Figure~\ref{fig:floodcomparison}).
-
-For flood data, we compared flood status, based on event listings in the NOAA
-Storm Events database, to streamflow measurements at \ac{US} Geological Survey
-gages within nine sample counties (Figure~\ref{fig:floodcomparison}).  Across
-the sampled counties, streamgage data generally indicated higher discharge in
-on dates identified as flood events, particularly in counties with more
-reporting stream gages (e.g., Harris County, TX, with~10 reporting stream
-gages; Figure~\ref{fig:floodcomparison}). There were some cases, however, where
-the two flooding data sources were somewhat inconsistent.  For example, there
-were one or two tropical cyclones in several of the counties (Mobile County,
-AL, Escambia County, FL, and Montgomery County, PA) which were classified as
-flood events based on NOAA storm event listings but for which the total
-discharge across county streamflow gages was below the 25\textsuperscript{th}
-percentile of measurements taken during storms that did not have a flood event
-listing. For tropical cyclones that did not have an associated flood listing in
-the NOAA Storm Events database, most had lower total streamflow discharge
-across county gages compared to storms with flood event listings, and in all
-cases but one all gages were below the flooding threshold.  The exception was
-one storm in Fulton County, GA (Ida in 2009). [More on which storm and why---I
-think there was already a flood event that had started earlier, so it missed
-the cutoff for the startdate being within the range.]
-
-\subsection*{Patterns in tropical cyclone exposures in eastern \ac{US} counties}
-
-We used the storm exposure metrics to classify counties as exposed or unexposed
-to specific tropical cyclones, and we explored patterns in these exposure
-classifications over years with available data
-(Table~\ref{tab:exposuresummaries}). Across the five exposure metrics, there
-was wide variation in the average number of county exposures per year. For the
-tornado metric, there were approximately~50 county exposures per
-year, on average, within our study.  For flood and wind metrics, there were
-substantially more county exposures ($>$200 / yr on average), and
-even more when assessing exposure using the distance and rain metrics ($>$400 /
-yr on average).
-
-We also measured the average number of county exposures per storm under each
-exposure metric (limiting this analysis to the tropical cyclones that exposed
-at least one county to that hazard; Table~\ref{tab:exposuresummaries}).
-Tropical cyclones typically exposed only a few counties, if any, to tornadoes
-(median counties exposed per storm:~7), more counties under the flood and wind
-metrics (median:~26 and~34, respectively), and even more counties for the
-metrics of rain (median:~68) and distance (median:~84).  For every metric
-except the tornado-based metric, we identified at least one tropical cyclone
-with over~300 counties exposed.  However, the largest-extent tropical cyclone
-varied by metric (e.g., Beryl in~1994 exposed the most counties based on
-distance, Frances in~2004 based on rain, and Ike in~2008 based on wind).
-
-When we calculated and mapped the average number of exposures per decade in
-each county for each exposure metric (Figure~\ref{fig:averageexposure}), strong
-geographical patterns were clear. Wind-based exposure had a strong coastal
-pattern, with almost all exposures in counties within about~200
-\si{\kilo\metre} (124~mi) of the coastline. While exposures by rain- and
-distance-based metrics were also more common in coastal areas compared to
-inland areas, they also created inland exposures that were not identified by
-the wind-based metric. Rain and, to some extent, flood exposures were
-characterized by a pattern defined by the Appalachian Mountains, with much
-lower exposures west of the mountain range than to the east. Almost all
-tornado-based exposures were in coastal states, with most in Florida, Alabama,
-South Carolina, and North Carolina and almost none north of Maryland.
-Flood-based exposures were highest in North Carolina and along the South
-Carolina coast, as well as further north in New Jersey, southeastern
-Pennsylvania, and southeastern New York. Notably, some areas of the \ac{US}
-that experienced regular exposures to tropical cyclone-related excessive
-rainfall and floods were not identified as areas of regular risk based on the
-more commonly used tropical cyclone exposure assessment metrics of wind or
-distance (Figure~\ref{fig:averageexposure}). Patterns were similar when
-analysis was restricted to years with all exposure data available
-(1996\,--\,2011; Figure~S3).
-
-\subsection*{Agreement in county-specific exposure classification between exposure metrics}
-
-Next, we measured the agreement between exposure classifications for each pair
-of metrics. We found that these exposure classifications typically did not
-strongly agree between pairs of metrics, indicating that the set of counties
-identified as exposed based on one metric can have limited overlap with the
-set of counties identified as exposed to the storm by another metric.
-
-Figure~\ref{fig:ivanexposure} shows Hurricane Ivan in 2004 as an example. For
-the distance-based metric, the counties assessed as exposed followed the
-tropical cyclone's track, including counties well inland. For the wind metric,
-only counties near two of the tropical cyclone's landfalls were assessed as
-exposed. For rain- and flood-based metrics, however, exposure extended to the
-left of the track, including counties as far north as New York and Connecticut,
-while for the tornado metric, exposed counties tended to be to the right of the
-track and included several counties in central North Carolina, South Carolina,
-and Georgia that were not identified as exposed to Ivan based on any other
-metric. Figure~S4 provides similar maps for four other example tropical
-cyclones (selected because they exposed many \ac{US}  counties based on at
-least one metric), and Figure~S5 provides four examples of tropical cyclones in
-which the disagreement in exposure assessment between rain- and wind-based
-metrics was particularly notable.
-
-We drew similar conclusions when we investigated all~46 tropical cyclones
-between~1996 and~2011 (the period for which all five metrics were available)
-for which~250 or more counties were exposed based on at least one metric
-(Figure~\ref{fig:jaccard}). In this figure, each row provides results for a
-specific tropical cyclone, and each box in that row shows the measured Jaccard
-coefficient between two of the exposure metrics for that storm.  For most
-tropical cyclones, distance-based county-level exposure assessment was at best
-in moderate agreement with exposure assessments based on the four hazard-based
-metrics. For almost all tropical cyclones, Jaccard index values between
-distance and any of the four hazard-based metrics were~$\le0.6$ (i.e., out of
-the counties assessed as exposed to the tropical cyclone by at least one of the
-two metrics considered,~60\si{\percent} or fewer of the counties had the same
-exposure assessment under the two metrics). For many tropical cyclones, the
-Jaccard index values were~$\le0.2$, suggesting very poor agreement in exposure
-assessment between the metrics.  Among the four hazard-based metrics, the
-tornado-based metric showed universally poor agreement with other metrics in
-county-level classification across all tropical cyclones considered.  For other
-pairs of metrics, there were also generally large differences in which counties
-were determined to be exposed, with the Jaccard index values~$<0.4$ for most
-metric pairs in most tropical cyclones.
-
-\begin{comment}
 Storms are displayed within
@@
 in~2005, and Katrina in~2005 (Figure~S4).
-\end{comment}
-
-There were, however, a few exceptions\,---\,tropical cyclones in which similar
-counties were determined to be exposed to the tropical cyclone for two or more
-of the metrics considered.  For Floyd in~1999 (Figure~S4) and Irene in~2011,
-for example, county-level classification agreed moderately to well (Jaccard
-index of~$\sim 0.5\mbox{\,--\,}0.8$) for all pairs of exposure metrics except
-those including the tornado-based metric. For another set of tropical cyclones
-(e.g., Lee in~2011 [Figure~S3], Ernesto in~2006, and Bertha in~1996), there was
-moderate to good agreement for all pairwise combinations of distance, rain, and
-wind, but poor agreement for all other combinations of metrics. However, these
-two sets of tropical cyclones represented the minority of all tropical cyclones
-considered; for most tropical cyclones, there was low overlap in the counties
-determined to be exposed to the tropical cyclones between most pairings of
-metrics.
-
-\subsection*{Case study}
-
-Finally, we conducted a case study to investigate how these differences in
-exposure assessment across the exposure metrics could affect estimates of
-expected physical exposure among a vulnerable subpopulation,
-electricity-dependent Medicare beneficiaries.
-Figure~\ref{fig:topelecdependexposure} shows the study counties with the
-highest expected physical exposures in this subpopulation based on each
-exposure metric. A few counties (Miami-Dade County, FL; Harris County, TX) were
-ranked in the top three counties for expected physical exposure in this
-subpopulation for almost all exposure metrics
-(Figure~\ref{fig:topelecdependexposure}). However, there were key differences
-across exposure metrics. For example, most of the top counties based on wind-
-and tornado-based metrics were in Florida, while over half of the counties with
-highest expected physical exposures for the flood-based metric were in
-Mid-Atlantic states (PA, NY).

### 2019-10-09 writing/ehp_abstract.tex
Work on paper, use 34 knot threshold

@@
 \noindent \textbf{Background:} Tropical cyclones bring severe impacts to
-\ac{US} communities, including substantial human health risks, from hazards
-that include wind, rain, flooding, and tornadoes. However, studies quantifying
-tropical cyclone risks and impacts vary widely in how they classify exposure to
-tropical cyclones, using various hazard-based metrics and, in some cases, using
-distance from the storm track as a surrogate. \\ 
-\textbf{Objectives:} (1)~Investigate how well county-level assessments of
-tropical cyclone exposure agree when based on distance versus hazard-based
-measurements, as well as among assessments based on different hazard
-measurements, and (2)~provide software to improve tropical cyclone exposure
-assessment for health-related research. \\ 
-\textbf{Methods:} We measured county-level exposure to tropical cyclones in the
-\ac{US} based on distance from the storm track, maximum sustained wind,
-rainfall, flooding, and tornadoes for all land-falling or near-land Atlantic
-basin storms, for~1996\,--\,2011 for all metrics and up to~1988\,--\,2015 for
-specific metrics. We quantified how well county-level exposure assessment
-agreed for all pairwise combinations of metrics.\\ 
-\textbf{Results:} In most of the storms investigated, agreement in county-level
-exposure assessment was at best moderate, and often poor, when using distance
-compared with four hazard-based measurements. There was also substantial
-disagreement in exposure assessment based on all pairwise combinations of the
-four hazard-based measurements. There were geographic patterns in exposures
-based on these metrics, with most wind-based exposures near the coast, flood-
-and rain-based exposures extending to inland and northern counties, and
-tornado-based exposures in southern counties. \\ 
-\textbf{Discussion:} Our results suggest that when impact studies use distance
-as a surrogate for tropical cyclone exposure, or use one hazard-based metric
-when the impact is partly or fully caused by a different hazard, the study will
-be prone to exposure misclassification.  To facilitate future research, we make
-this multi-hazard storm exposure data available through open-source software. 
+	\ac{US} communities, including substantial human health risks, from
+	hazards that include wind, rain, flooding, and tornadoes. However, 
+	it is difficult to conduct multi-year, multi-storm epidemologic 
+	studies without storm hazard datasets that can easily be linked to 
+	human health data. Further, while some storm exposure metrics are
+	straightforward to measure (e.g., distance from a community to a 
+	storm's track), they may poorly capture exposure to storm hazards. \\ 
+\textbf{Objectives:} (1)~Provide an open-source dataset to improve tropical
+	cyclone exposure assessment for health-related research and
+	(2)~investigate agreement between county-level assessments of tropical cyclone
+	exposure when based on different storm hazards, to clarify 
+	risk of exposure misclassification in epidemiologic studies. \\ 
+\textbf{Methods:} We created an open-source dataset with county-level data on
+	exposure to tropical cyclones in the \ac{US}. The dataset includes
+	measures of distance from the storm track, maximum sustained wind,
+	rainfall, flooding, and tornadoes for all land-falling or near-land
+	Atlantic basin storms over multiple years (1996\,--\,2011 for all
+	metrics, up to~1988\,--\,2015 for specific metrics). [Validation] We measured
+	agreement in county-level exposure assessment for all pairwise
+	combinations of these metrics. \\ 
+\textbf{Results:} We created an open-source dataset ... 
+	[Most data agreed well. Flooding is hard to measure.]
+	In most of the storms investigated, agreement in county-level
+	exposure assessment was at best moderate, and often poor, when using
+	distance compared with four hazard-based measurements. There was also
+	substantial disagreement in exposure assessment based on all pairwise
+	combinations of the four hazard-based measurements. There were
+	geographic patterns in exposures based on these metrics, with most
+	wind-based exposures near the coast, flood- and rain-based exposures
+	extending to inland and northern counties, and tornado-based exposures
+	in southern counties. \\ 
+\textbf{Discussion:} Our results provide a multi-hazard dataset that can be
+	leveraged for epidemiologic research on tropical cyclones.  Further,
+	our results suggest that when epidemiologic studies use distance as a
+	surrogate for tropical cyclone exposure, or use one hazard-based metric
+	when the impact is partly or fully caused by a different hazard, the
+	study will be prone to exposure misclassification.  
 \end{abstract}

### 2019-10-09 writing/ehp_discussion.tex
Work on paper, use 34 knot threshold

--- 2019-07-28 writing/ehp_discussion.tex
@@
 
-Tropical cyclones impact public health in many ways, and epidemiological
-studies of the health risk and impacts associated with tropical cyclone
-exposures could help improve preparedness for and response to future tropical
-cyclones.  However, tropical cyclones are multi-hazard events, making it
-complicated to measure exposure. Here, we provide open-source county-level data
-on multiple storm exposure metrics, and we find that county-level tropical
-cyclone exposure assessments vary substantially when using a distance-based
-metric versus four hazard-based metrics, as well as among different
-hazard-based metrics. Our results can inform exposure assessment for future
-county-level studies of the health risks and impacts associated with tropical
-cyclones exposures, and the open-source data and software we have created
-provides multi-hazard, county-level exposure data for such studies.
+Epidemiological studies can help characterize which health risks are elevated
+during disasters, to what degree, and to whom [Ibrahim editorial; Noji paper].
+As a result, these studies help improve disaster preparedness and response
+[Noji paper].  However, tropical cyclones are multi-hazard events, making it
+complicated to measure exposure.  Here, we provide open-source county-level
+data on multiple tropical cyclone exposure metrics and explore limitations in
+that data.
+
+Further, we find that county-level tropical cyclone exposure assessments vary
+substantially when using different metrics, based either on storm hazards or on
+how close the storm's central track came to the community. Our results can
+inform exposure assessment for future county-level studies of the health risks
+and impacts associated with tropical cyclones exposures, and the open-source
+data and software we have created provides multi-hazard, county-level exposure
+data for such studies.
 
@@
 
-In creating this hurricane exposure dataset, we aimed for data that are
-available for all eastern \ac{US} counties. In addition to providing tropical
-cyclone exposure metrics for individual hazards, this data and its associated
-software allow users not only to access measurements for single hazards, but
-also to create tropical cyclone exposure profiles based on multiple hazards or
-craft exposure indices that combine hazard
-metrics~\parencite{chakraborty2005population, peduzzi2009assessing}. This
-ability can be critical, as different hazards of tropical cyclones often act
-synergistically in causing impacts~\parencite{smith2009}.
-
-\begin{comment}
-To assist with future tropical cyclone studies, we created and published open
-source software with this dataset and accompanying
-tools~\parencite{hurricaneexposuredata, hurricaneexposure}.  Many previous
-studies have used geographical information system software (e.g., ArcGIS) to
-assess exposure to tropical cyclones in the \ac{US}~\parencite{grabich2016,
-zandbergen2009, czajkowski2011, kruk2010}.  Here, we offer methods to map and
-output historic exposure to tropical cyclones that does not require the use of
-proprietary software but instead uses a package written in the R statistical
-programming language~\parencite{R}, which is free and open-source.  Further, by
-including measurements of different hazard exposures in each county for each
-tropical cyclone, this software allows for the development of more complex
-exposure indices or models (e.g., random forests, multivariable generalized
-linear models) that incorporate multiple tropical cyclone hazard measurements.  
-\end{comment}
-
-We investigated how well these data correspond with data from other available
-sources (Figures~\ref{fig:windcomparison}\,--\,\ref{fig:floodcomparison}
-and~S6\,--\,7). While generally in agreement with data from other sources,
-there are a few caveats that should be considered when applying these data in
-future studies. The rainfall data are from re-analysis data are generally
-well-correlated with observed ground-based station data, but may sometimes
+We investigated how well the open-source data correspond with
+data from other available sources
+(Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison} and~S2). While
+generally in agreement, there are a few caveats that should be considered when
+using these data. The rainfall data are generally
+well-correlated with ground-based observations, but may sometimes
 oversmooth very high rainfall values compared to ground-based observations
@@
 as both data sources agree in identifying these as storms with high rainfall,
-but it may be important to consider for cases that include rainfall as a
+but would be important to consider for cases that include rainfall as a
 continuous measurement.
@@
 from \ac{HURDAT2}~\parencite{landsea2013} (Figure~\ref{fig:windcomparison}),
-there were a few storms where the two datasets did not agree as well ($<$90\%
-of counties classified into the same wind categories by the two sources of wind
-estimates). These storms were exceptionally large systems for which high winds
-persisted well inland from landfall, like Hurricanes Sandy and Ike
-(Figures~\ref{fig:windcomparison} and\~S6). Since the wind radii are based on
-\ldots, they have [advantages]; their main disadvantages is that they provide
-categorical, rather than continuous, estimates of wind speed, and in many cases
-researchers might prefer more precise estimates of storm wind speeds. However,
-since the wind radii data may be useful in some cases as either a primary
-metric or for sensitivity analysis, we have include county-level wind estimates
-for each storm based on these wind radii, in addition to the continuous modeled
-windspeeds, in the open-source data \cite{hurricaneexposuredata}.
+there were a few storms with some discrepancies. These storms---for example, Hurricane Sandy
+in 2012 and Hurricane Ike in 2008---were
+unusually large systems for which high winds persisted well inland from
+landfall (Figures~\ref{fig:windcomparison}
+and~S2).
 
-For the flooding events data, we found it often corresponded with measurements
-from \ac{US} Geographical Survey streamgages, with flood event listings more
-likely during storms which elevated stream flows. However, there are
-differences between the two flooding datasets, and these highlight both the
+For the flooding events data, we found that flood event status determined based
+on the NOAA Storm Events listings often agreed with measurements from \ac{USGS}
+streamgages, with a flood event more likely to be listed if a storm elevated
+stream flows in the county (Figure~\ref{fig:floodcomparison}).  However, there
+are differences between the two flooding datasets, and these highlight both the
 difficulty of measuring flood exposure at the county level and inherent
-challenges in using the storm event database for epidemiologic exposure
-assessment.  The flood and tornado data came from the \ac{NOAA} Storm Events
-database, which, while a widely-used database of events maintained by
-\ac{NOAA}, is based on reports, and so may be prone to
-underreporting~\parencite{Ashley2008flood, Curran2000}, especially in less
-populated areas~\parencite{Witt1998, Ashley2007}, as well as to other reporting
-errors.
+challenges in using data from a storm event database for epidemiologic exposure
+assessment.
 
-\begin{comment}
-[For discussion?] Based on these results, the two data sources were generally
-in agreement, but the analysis highlights the difficulty of assessing
-storm-related flooding at a multi-county, multi-year scale, as flooding can be
-very localized within a county or can \ldots .   
-\end{comment}
+For example, there was one storm in Fulton County, GA, for which there was high
+streamflow but not an associated flood event listing (Hurricane Ida, 2009).
+This storm occurred in November 2009, following a month with historic rainfall
+and flooding in Georgia \parencite{shepherd2011overview}.  Therefore, in this
+case the flooding associated with Ida was incorporated into an ongoing flood
+event listing, with a start date well before the five-day window used to
+temporally match storm event listings with tropical cyclone tracks used here.
+
+This disagreement highlights the difficulty of large-scale pairing of storm
+tracks with storm event listings---without criteria for temporally matching
+event dates to storm tracks, it is likely that many false positives would be
+captured, for which the occurrance of a storm in the midst of an ongoing storm
+event listing would be improperly attributed to the storm.  However, distance
+and time restrictions, like those we used in matching NOAA Storm Event listings
+with tropical storm locations and dates, do create the opportunity for
+occasional false negatives, as for Hurricane Ida in Fulton County, GA, where a
+storm contributes meaningfully to an ongoing event, but the event is not
+captured for the storm.
+
+There are further limitations for the flood and tornado data---these data came
+from the \ac{NOAA} Storm Events database, which, while a widely-used database
+of events maintained by \ac{NOAA}, is based on reports, and so may be prone to
+underreporting~\parencite{Ashley2008flood, Curran2000}, especially in
+sparsely-populated areas~\parencite{Witt1998, Ashley2007} [Simmons and Sutter
+ref], as well as to other reporting errors.
 
 While these are important caveats for the data, we selected these data sources
-as among the best currently available for measuring each of these hazards at a
-multi-county, multi-year scale.
+as among the best currently available for measuring each of these hazards
+consistently and comprehensively at a multi-county, multi-year scale. In
+addition to providing tropical cyclone exposure metrics for individual hazards,
+this dataset and its associated software allow users not only to access
+measurements for single hazards, but also to create tropical cyclone exposure
+profiles based on multiple hazards or craft exposure indices that combine
+hazard metrics~\parencite{chakraborty2005population, peduzzi2009assessing}.
+This ability can be critical, as different hazards of tropical cyclones often
+act synergistically in causing impacts~\parencite{smith2009}.
 
@@
 
-We found geographic patterns in tropical cyclones based on the different
-exposure metrics that. These patterns were not unexpected based on what is
-known about tropical cyclone hazards, but still highlight variations across
-metrics that are critical to consider in designing studies for tropical cyclone
-epidemiology, and they demonstrate the need for multi-hazard exposure datasets
-for tropical cyclone epidemiology. Wind-based exposures had a strong coastal
-pattern, which is consistent with the dramatic decrease in wind intensity that
-typically characterizes the landfall of tropical cyclones. Exposures by rain-
-and distance-based metrics often extended further inland compared to the
-wind-based metric, up to the Appalachian mountains. This agrees with previous
-research indicating that the Appalachian mountains' topography both enhances
-precipitation during tropical cyclones and provides hydrological conditions for
-severe flooding~\parencite{rees2001}.  Almost all tornado-based exposures were
-in southern coastal states, echoing previous findings that most noteworthy
+We found differences in geographic patterns in average exposures to different
+tropical cyclone hazards (Figure~\ref{fig:averageexposure}). These patterns
+were not unexpected based on what is known about tropical cyclone hazards, but
+still highlight variations that are critical to consider in designing studies
+and statistical analysis for tropical cyclone epidemiology. Further, they
+demonstrate the need for multi-hazard exposure datasets for tropical cyclone
+epidemiology.
+
+Tropical cyclone wind exposures had a strong coastal pattern, which is
+consistent with the dramatic decrease in wind intensity that typically
+characterizes the landfall of tropical cyclones.  Tropical cyclone rain
+exposures tended to extended further inland compared to wind exposures, up to
+the Appalachian mountains. This agrees with previous research indicating that
+the Appalachian mountains' topography both enhances precipitation during
+tropical cyclones and provides hydrological conditions for severe
+flooding~\parencite{rees2001}.  Almost all tropical cyclone tornado exposures
+were in southern coastal states, consistent with previous evidence that most noteworthy
 tropical cyclone-related tornadoes occur on the right-side of the tropical
-cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.  These
-exposure averages are somewhat limited as estimates of long-term frequencies,
-as tropical cyclones follow decadal patterns~\parencite{kossin2007more} likely
-not adequately captured in the available data. However, exposure assessments
+cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.  The
+exposure averages we calculated are somewhat limited as estimates of long-term
+frequencies, as tropical cyclones follow decadal
+patterns~\parencite{kossin2007more} likely not adequately captured in the
+available data, which covered less than 20 years. However, exposure assessments
 spanning over a century have found similar patterns for distance- and
@@
 
-\begin{comment}
-However, these frequency maps, together with evidence from
-specific tropical cyclones (Figures~\ref{fig:ivanexposure}
-and~\ref{fig:jaccard}),  do illustrate the potential for strong differences in
-spatial patterns in tropical cyclone exposures, depending on which tropical
-cyclone hazards are considered.  A few previous studies have sought to
-determine county-level exposure to tropical cyclones over multi-year periods,
-including~\textcite{zandbergen2009}, which estimated exposure in \ac{US}
-counties to all \ac{US} landfalling Atlantic-basin tropical cyclones
-between~1851 and~2003, using both a distance-based metric and a metric that
-combined distance and windspeed, and~\textcite{kruk2010}, which explored
-exposure to hurricane-related winds in the \ac{US}, including inland areas,
-for~1900\,--\,2008.  Our results suggest that such exposure assessments may
-perform well in capturing some tropical cyclone hazards (e.g., wind), but
-likely miss other potentially dangerous tropical cyclone exposures, especially
-for hazards that repeatedly threaten northern or inland counties (e.g., rain,
-flooding).
-\end{comment}
-
 \subsection*{Agreement between exposure metrics}
 
-We found that there was typically only moderate or low agreement between
-county-level exposure classifications for a tropical cyclone when comparing
-different exposure metrics. This suggests that the use of one of these metrics
-as a proxy for all storm-based hazards in epidemiological research would often
-be problematic, introducing exposure misclassification.
-
-\begin{comment}
-For example, distance from a tropical cyclone's track is relatively easy to
-measure and has been used as an operational metric of exposure to tropical
-cyclones in previous large-scale studies (examples include ...). Since distance
-itself does not constitute a hazard, distance is meant in these cases as a
-surrogate to capture exposure to hazards from the tropical cyclone. However,
-here we found that in assessing \ac{US} county-level exposure to tropical
-cyclones, distance is, at best, a moderate, and often a very poor, surrogate
-for exposure to the specific tropical cyclone hazards of high wind, extreme
-rainfall, flooding, and tornadoes (Figure~\ref{fig:jaccard}). Therefore, use of
-distance to assess tropical cyclone exposure for impact studies could result in
-problematic exposure misclassification, which could mask true associations,
-even strong associations, between tropical cyclone exposure and outcomes of
-interest in impact studies~\parencite{savitz2016interpreting,
-armstrong1998effect}.  
-\end{comment}
+We found that agreement was typically low across hazard-specific tropical
+exposure classifications, and agreement was also typically low between
+distance-based tropical cyclone exposure assessment and each of the single
+hazard-specific exposures (Figures~\ref{fig:ivanexposure}\,--\,
+\ref{fig:jaccard} and S4\,--\,S5). This suggests that the use of a
+distance-based metric to assess exposure to any of these hazards, or the use of
+measurements from one hazard as a proxy for exposure to any of the other
+hazards considered, would often be problematic, introducing exposure
+misclassification.
 
@@
 fields on the characteristics of tropical cyclones. While tropical cyclone
-rainfall and windspeed are often well-correlated when the tropical cyclone is
-over water~\parencite{cerveny2000}, this relationship often does not remain as
-strong once the hurricane has made landfall~\parencite{jiang2008}.  Fast-moving
-tropical cyclones bring higher risks of dangerous winds
-inland~\parencite{kruk2010}, while slow-moving tropical cyclone are likely to
-bring more rain~\parencite{rappaport2000} and cause more damage because of
-sustained hazardous conditions~\parencite{rezapour2014}. Further, while the
-likelihood and extent of flooding during a tropical cyclone is related to the
-tropical cyclone's rainfall, it is also driven by factors like top soil
-saturation and the structure of the water basin's drainage
-network~\parencite{chen2015, rees2001}.
+rainfall and windspeed can be well-correlated when the tropical cyclone is
+over water~\parencite{cerveny2000}, this relationship often weakens once the
+hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
+cyclones bring higher risks of dangerous winds inland~\parencite{kruk2010},
+while slow-moving tropical cyclones are likely to bring more
+rain~\parencite{rappaport2000} and cause more damage because of sustained
+hazardous conditions~\parencite{rezapour2014}. Further, while the likelihood
+and extent of flooding during a tropical cyclone is related to the tropical
+cyclone's rainfall, it is also driven by factors like top soil saturation and
+the structure of the water basin's drainage network~\parencite{chen2015,
+rees2001}.
 
 However, we did find a small set of tropical cyclones for which for which
-agreement between metrics were unusually high (e.g., Floyd in~1999, Irene in
-2011, Hannah in~2008, Bertha in~1996).  Hurricanes Floyd in~1999 and Irene
+agreement was high across single-hazard exposure assessments (e.g., Floyd
+in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996;
+Figure~\ref{fig:jaccard}).  Hurricanes Floyd in~1999 (Figure~S4) and Irene
 in~2011 both made their first \ac{US} landfall in North Carolina at minor
 hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
-eastern coast of the \ac{US} north through New England, bringing large rainfall
-to much of the eastern coast from North Carolina north and causing extensive
-inland flooding in North Carolina (Floyd) and New England
+eastern coast of the \ac{US} north through New England, bringing substantial
+rainfall to much of the eastern coast from North Carolina north and causing
+extensive inland flooding in North Carolina (Floyd) and New England
 (Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
@@
 tropical cyclones' persistent proximity to water may have helped maintain wind
-speeds in similar patterns to rain and distance exposures, resulting in this
-moderate to good agreement among exposure assessments based on different
-metrics.
+speeds in similar patterns to rain and distance-based exposure assessments,
+resulting in more similarities across exposure assessments compared to the
+other tropical cyclones we considered.
 
-In some cases, exposure misclassification from using one of these exposure
-metrics to capture one or more of the other storm hazards may be differential
-(i.e., associated with the outcome of interest or with factors associated with
-risk of the outcome of interest). For example, tropical cyclone wind exposures
-tend to be concentrated in counties near the coast, since most tropical
-cyclones rapidly decrease in their sustained windspeed following landfall.
-However, tropical cyclone exposures based on distance can extend well inland,
-following the tropical cyclone's tracks, but may not adequately capture all
-wind-exposed counties near the coast. In this case, if the
-etiologically-relevant exposure is high wind but exposure is classified based
-on distance, the probability of being misclassified as unexposed would be
-higher in coastal counties, while the probability of being misclassified as
-exposed would be higher in inland counties. If coastal counties differ from
+Since exposure assessments often differ across storm hazards, as well as
+between each storm hazard and a distance-based proxy measurement, exposure
+misclassification is a potential risk. For some studies, such exposure
+misclassification might plausibly be differential (i.e., associated with the
+outcome of interest or with factors associated with risk of the outcome of
+interest).  For example, tropical cyclone wind exposures tend to be
+concentrated in counties near the coast, while tropical cyclone wind exposures
+sometimes extend well inland.  If the etiologically-relevant exposure for a
+health outcome is extreme rainfall but exposure is classified based on
+measurements of wind, the probability of being misclassified as unexposed would
+be higher in inland counties, while the probability of being misclassified as
+exposed would be higher in coastal counties. If coastal counties differ from
 inland counties in either the outcome of interest or in factors associated with
-risk of that outcome, differential exposure misclassification would
-exist~\parencite{savitz2016interpreting}. Such differential exposure
-misclassification can bias estimates of tropical cyclone effects either towards
-the null (estimating a lower or null association compared to the true
+risk of that outcome, exposure misclassification would be
+differential~\parencite{savitz2016interpreting}.  Such differential exposure
+misclassification could bias estimates of tropical cyclone effects either
+towards the null (estimating a lower or null association compared to the true
 association that exists) or away from the null (estimating a larger association
 than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.
-
-\begin{comment}
-The use of a single hazard-based metric (e.g., wind) could cause similar
-problems if the impact is driven, at least in part, by a different hazard or by
-multiple hazards of the tropical cyclone.
-\end{comment}
+[Variation across studies in how exposure is assessed?]
 
@@
 
-The conducted case study demonstrates how differences in exposure assessment
-across storm exposure metrics can result in differing estimates of expected
-physical exposure to tropical cyclones among a population
-(Figure~\ref{fig:topelecdependexposure}).  Previous assessments of physical
-exposure to tropical cyclones have often focused on solely storm winds and
-storm proximity. For example, a global study of exposure and vulnerability to
-natural hazards used a combination of windspeed above a certain threshold and
-distance from the storm's track to assess
-exposure~\parencite{peduzzi2009assessing}. While many storm impacts might be
-most strongly linked to wind hazards, there is a growing recognition of the
-potential risks of adverse health outcomes and property damage from rain- and
-flood-related tropical cyclone hazards in the \ac{US}~\parencite{smith2009}.
-Such assessments may underemphasize potentially dangerous exposures to certain
-regions of the \ac{US}, particularly inland locations. In fact, based on our
-assessment of physical exposures to tropical cyclones among
-electricity-dependent Medicare beneficiaries, the highest expected physical
-exposure under any metric was measured for Philadelphia County, PA, under the
-flood metric (Figure~\ref{fig:topelecdependexposure}).  While this county has a
-similar-sized population of electricity-dependent Medicare beneficiaries as
-Miami-Dade County, FL (Figure~\ref{fig:topelecdependexposure}), the frequency
-of storm-related flood events in this inland county results in a very high
-expected physical exposure among this subpopulation, while this county is not
-included in the top ten counties for physical exposure based on most of the
-other exposure metrics.
+The case study illustrates how differences in exposure assessment
+across storm exposure metrics can result in differing conclusions. In this 
+case, we found differences in estimates of expected
+physical exposure to tropical cyclones among a population when exposure
+assessment was based on different storm exposure metrics
+(Figure~\ref{fig:topelecdependexposure}).
+
+These results are informative because previous assessments of physical exposure
+to tropical cyclones have often concentrated on storm winds and storm
+proximity. For example, a global study of exposure and vulnerability to natural
+hazards used a combination of windspeed above a certain threshold and distance
+from the storm's track to assess exposure~\parencite{peduzzi2009assessing}.
+While many storm impacts might be most strongly linked to wind hazards, there
+is a growing recognition of the potential risks of adverse health outcomes and
+property damage from rain- and flood-related tropical cyclone hazards in the
+\ac{US}~\parencite{smith2009}.  Assessments based solely on wind and distance
+from the storm track therefore may underemphasize potentially dangerous
+exposures to certain regions of the \ac{US}, particularly inland locations.
+
+In fact, based on our assessment of physical exposures to tropical cyclones
+among electricity-dependent Medicare beneficiaries, the highest expected
+physical exposure under any metric was measured for Philadelphia County, PA,
+under the flood metric (Figure~\ref{fig:topelecdependexposure}).  While this
+county has a similar-sized population of electricity-dependent Medicare
+beneficiaries as Miami-Dade County, FL
+(Figure~\ref{fig:topelecdependexposure}), the inland county of Philadelphia has
+frequent storm-related flood events, which results in a very high expected
+physical exposure among electricity-dependent Medicare beneficiaries in the
+county. This exposure is poorly captured by most of the other exposure metrics,
+under which Philadelphia County is not included in the top ten counties for
+physical exposure~(Figure~\ref{fig:topelecdependexposure}).
 
@@
 can have in \ac{US} communities. However, studies of tropical cyclone exposures
-and their impacts have varied widely in how they assess exposure. Here we found
-large differences in which counties are exposed to different hazards of
-tropical cyclones and that distance is, at best, a moderate, and often a very
-poor, surrogate for exposure to the specific tropical cyclone hazards of high
-wind, extreme rainfall, flooding, and tornadoes. Use of distance as a surrogate
-for any of these hazards could lead to exposure misclassification and, in the
-case of tropical cyclone risk and impact studies, including epidemiological
-studies, result in biased estimates.  Similarly, if such studies disagree on
-their findings, it could result from the often poor agreement between exposure
-classifications based on these different metrics, making studies that use
-different metrics hard to compare or combine in meta-analyses. Our findings
+and their impacts have varied widely in how they assess exposure, and there was
+not previously a widely available public dataset that captured county-level
+exposure to multiple storm hazards. Here we provide open-source tropical
+cyclone hazard-related measurements for a large collection of historical
+Atlantic-basin cyclones, aggregated at the county level to align with the
+spatial scale at which much health-related data is collected.
+
+Further, we investigated agreement in exposure assessment across these metrics
+when applied individually. We found large differences in which counties are
+exposed to different hazards of tropical cyclones and that distance is, at
+best, a moderate, and often a very poor, surrogate for exposure to the specific
+tropical cyclone hazards of high wind, extreme rainfall, flooding, and
+tornadoes. Use of distance as a surrogate for any of these hazards, or exposure
+assessment based on one hazard when the pathway for health impacts is in part
+or full through another storm hazard, could lead to exposure misclassification.
+In the case of tropical cyclone risk and impact studies, including
+epidemiological studies, this would result in biased estimates.  Our findings
 highlight the importance of clarifying the potential pathway from tropical
@@
 epidemiological studies, and then basing exposure assessment on these hazards.
-To help, we provide open-source tropical cyclone hazard-related measurements
-for a large collection of historical Atlantic-basin cyclones, aggregated at the
-county level to align with the spatial scale at which much health-related data
-is collected.

### 2019-10-09 writing/ehp_figures.tex
Work on paper, use 34 knot threshold

@@
 \caption{States and storms considered in this study. All counties in the states
-        shown in this map were investigated. The lines show the paths
-	of the study storms, which included all tracked storms in 1988\,--\,2015
-	\ac{HURDAT2}~\parencite{landsea2013} that came within 250~\si{\kilo\metre} 
-	of at least one \ac{US} county. Thicker lines show the tracks 
-	of storms whose names have been retired, indicating that the storm was particularly 
-	severe or had notable impacts~\parencite{retirednames}.
-	}
+	shown in this map were investigated. The lines show the paths of the
+	study storms, which included all tracked storms in 1988\,--\,2015 that
+	are recorded in \ac{HURDAT2}~\parencite{landsea2013} and that came
+	within 250~\si{\kilo\metre} of at least one \ac{US} county. Thicker
+	lines show the tracks of storms whose names have been retired,
+	indicating that the storm was particularly severe or had notable
+	impacts~\parencite{retirednames}.  }
 \label{fig:hurrtracks}
@@
 	and each point shows one tropical storm. The number of storms 
-	within each county and the average number of stations reporting rainfall during 
+	within each county and the number of stations reporting rainfall during 
 	the county's storms are given above each plot. Horizontal and vertical lines 
@@
 	as ``exposed'' in further analysis (Table 1). Note that 
-	x- and y-axis ranges differ by county.
+	ranges of the x and y axes differ across counties.
 	} 
@@
 \includegraphics[width=0.6\linewidth]{windcomparison}
-\caption{Comparison of modeled maximum sustained surface wind speeds 
-	(included as the primary wind exposure metric in the open-source
-	data and used in further analysis) with estimates from the \ac{HURDAT2}
-	wind speed radii. Each point represents a storm, with the x-axis
-	giving the percent of counties classified in the same category
-	of maximum sustained surface wind speeds ($<$34~\si{\knot}; 
-	34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; 
-	$\ge$64~\si{\knot}) by both sets of wind data for that storm. The 
-	color of each point indicates the extent of the storm, 
-	giving the number of study counties that were exposed to 
-	maximum sustained surface wind speeds of at least 34~\si{\knot}
-	(based on modeled wind speeds). Estimates are shown for [all?]
-	study storms since 2004, the earliest year for which 
-	wind speed radii estimates are available in \ac{HURDAT2} [?].
+\caption{Comparison of modeled maximum sustained surface wind speeds (included
+	as the primary wind exposure metric in the open-source data and used in
+	further analysis) with estimates from \ac{HURDAT2}'s wind speed radii.
+	Each point represents a storm, with the x axis giving the percent of
+	counties classified in the same category of maximum sustained surface
+	wind speeds ($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot};
+	50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) by both sources of wind
+	data for that storm. The color of each point indicates the extent of
+	the storm, giving the number of study counties that were exposed to
+	maximum sustained surface wind speeds of at least 34~\si{\knot} (based
+	on modeled wind speeds). Estimates are shown for study storms since
+	2004, the earliest year for which post-storm reanalysis wind speed
+	radii estimates are routinely available in \ac{HURDAT2}, for which at
+	least one study county had a sustained wind speed of $\ge$34~\si{\knot}
+	based on the \ac{HURDAT2} wind radii.
 	}
@@
 \includegraphics[width=0.8\linewidth]{floodcomparison}
-\caption{Comparison for a sample of study counties of NOAA Storm Events 
-	listings versus five-day total streamflows at county streamflow 
-	gages during tropical cyclones. Each small plot shows results for 
-	one of the sample counties. Each point represents a single tropical 
-	storm, and the point's position along the x-axis shows the highest 
-	daily total streamflow (cubic feet per second), summed across all 
-	identified streamgages in the county, for a five-day window centered 
-	on the day of the storm's closest approach to the county. The y-axis 
-	separates storms for which a flood event was reported in NOAA's Storm Events 
-	database for the county with a start date within the five-day window 
-	of the storm's closest approach to the county. The color of each point 
-	gives the percent of streamflow gages in the county with a daily streamflow 
-	that exceeded a threshold of flooding on any day during the five-day window.
-	Comparison for a sample of study counties of NOAA Storm Events 
-	listings versus five-day total streamflows at county streamflow gages during 
-	tropical storms. The number of streamflow gages used in analysis are given in 
-	parentheses beside the county's name in the panel title. Each point represents 
-	a single tropical storm. Note that the x-axis scales differ by county, 
-	depending on the number of streamflow gages and typical flow rates for each 
-	gage, and are on a log-10 scale.}
+\caption{Comparison for a sample of study counties of flood status based on
+	matching NOAA Storm Events listings to \ac{HURDAT2} storm track data
+	versus five-day total streamflows at county streamflow gages during
+	tropical cyclones. Each small plot shows results for one of the sample
+	counties. Each point represents a single tropical storm, and the
+	point's position along the x-axis shows the highest daily total
+	streamflow (cubic feet per second), summed across all identified
+	streamgages in the county, for a five-day window centered on the day of
+	the storm's closest approach to the county. The y-axis separates storms
+	for which a flood event was reported in NOAA's Storm Events database
+	for the county with a start date within the five-day window of the
+	storm's closest approach to the county. The color of each point gives
+	the percent of streamflow gages in the county with a daily streamflow
+	that exceeded a threshold of flooding (the streamgage's median value
+	for annual peak flow) on any day during the five-day window. The number
+	of streamflow gages used in analysis are given in parentheses beside
+	the county's name in the panel title.  Note that the x-axis scales
+	differ by county, depending on the number of streamflow gages and
+	typical flow rates for each gage, and are on a log-10 scale.
+	}
 \label{fig:floodcomparison}
@@
 \centering
-\includegraphics[width=15cm]{averageexposureonly.pdf} 
-\caption{Average number of storm exposures per decade in U.S. counties for each 
-	exposure metric. The criteria behind each of the five metrics is given in 
-	Table \ref{tab:exposuremetrics}. The years used to estimate these averages 
-	are based on years of available exposure data 
-	(distance and wind:~1988\,--\,2015; rain: 1988\,--\,2011; 
-	flood and tornado events:~1996\,--\,2015). } 
+\includegraphics[width=7cm]{averageexposureonly.pdf} 
+\caption{Average number of county-level storm exposures per decade for each
+	single-hazard exposure metric. The criteria behind each of the five
+	metrics is given in Table \ref{tab:exposuremetrics}. The years used to
+	estimate these averages are based on years of available exposure data
+	(rain: 1988\,--\,2011; wind:~1988\,--\,2015; flood and
+	tornado events:~1996\,--\,2015). } 
 \label{fig:averageexposure} 
@@
 \caption{Counties classified as exposed to Hurricane Ivan in 2004 under each
-exposure metric considered (Table~\ref{tab:exposuremetrics}). The red line
-shows the track of Hurricane Ivan based on the revised Atlantic hurricane
-database \ac{HURDAT2}~\parencite{landsea2013}.  Similar maps for other
-large-extent storms are given in Figure~S3.}
+	exposure metric considered (Table~\ref{tab:exposuremetrics}). The red
+	line shows the track of Hurricane Ivan based on
+	\ac{HURDAT2}~\parencite{landsea2013}.  Similar maps for other
+	large-extent storms are given in Figure~S4.
+	}
 \label{fig:ivanexposure} 
@@
 	 and Hurricanes Floyd in~1999, Lee in~2011, Cindy in~2005, and Katrina 
-	 in~2005 (Figure~S3).
+	 in~2005 (Figure~S4).
 } 

### 2019-10-09 writing/ehp_introduction.tex
Work on paper, use 34 knot threshold

--- 2019-01-28 writing/ehp_introduction.tex
@@
 
-Tropical cyclones, including hurricanes, can have disastrous impacts in the
-\acf{US}. A tropical cyclone's high winds can bring health risks
-and property damage through structural damage of houses and other buildings,
-falling trees, and wind-borne debris~\parencite{rappaport2000} and cause power
-outages~\parencite{liu2005, han2009}, which introduce a number of threats to human
-and ecological health, including water quality risks if the outage affects
-wastewater treatment plants~\parencite{mallin2006}.  Other risks can exist without
-severe wind; for example, one study found that most of the direct
-hurricane-related deaths in the \ac{US} between~1970 and~1999 occurred in cases
-when wind was below hurricane strength, including for Tropical Storms Charley
-in~1998 and Alberto in~1994~\parencite{rappaport2000}.  Tropical cyclones can
-produce excessive rain, especially in certain topographies (e.g., near
-mountains), so counties well inland sometimes experience more extreme rain than
-coastal counties. Flood risks from tropical cyclones can result from this rain,
-although the two risks are not perfectly correlated~\parencite{chen2015}. In the
-\ac{US}, over half of hurricane-related direct deaths from~1970 to~1999 from
-Atlantic basin storms were a result of freshwater flooding~\parencite{rappaport2000}. 
-Flooding can also degrade water quality~\parencite{mallin2006}, which can threaten both 
-human and ecological health.
+\acresetall
 
-To measure tropical cyclone exposure for either investigations of exposure
-patterns or for estimation of tropical cyclone risks and impacts, including
-those related to human health, studies have varied widely in how they determine
-tropical cyclone exposure (some examples shown in Figure~S1\todo{Check
-fig.ref}). While some studies have measured exposure based on specific hazards
-of the tropical cyclone (e.g., wind, rain), many have used distance from the
-tropical cyclone's central track as a surrogate in identifying exposure to the
-hazards of a tropical cyclone (e.g.,~\textcite{czajkowski2011, tansel2010,
-kinney2008, caillouet2008increase}).
+Recently, several major tropical cyclones have hit \ac{US} cities, including
+Hurricanes Harvey, Irma, and Maria in 2017 \parencite{blake20182017} and
+Hurricanes Florence and Michael in 2018 \parencite{avila20192018}. Researchers
+are exploring how these storms impacted human health (e.g.,
+\parencite{santos2018use, rivera2018estimating, santos2018differential,
+grineski2019impact, issa2018deaths, tanz2019notes, paul2019brief}), including
+through projects funded through \ac{NIH} Rapid Response grants
+\parencite{nihreporter}. These projects will add to evidence from previous
+hurricanes, including Sandy in 2012 (e.g., \parencite{swerdel2014}) and Katrina
+in 2005 (e.g., \parencite{burton2009health}), characterizing the health impacts
+of hurricanes.
 
-Distance from the tropical cyclone's central track can be easily measured using
-widely-available hurricane tracking data. There are, however, a number of
-limitations to using distance to assess exposure to tropical cyclones, and
-there are reasons to suspect that exposure classifications based on other
-tropical cyclone hazards may disagree with distance-based classifications.
-Tropical cyclones vary dramatically in size: \ac{US} tropical cyclones have
-been observed with radii to maximum winds as small as~20~\si{\kilo\metre} and as
-large as~200~\si{\kilo\metre}~\parencite{mallin2006, quiring2011variations}. While a
+A recent review highlights the need to supplement these single-storm studies
+with multi-year, multi-storm studies, to identify patterns that are common
+across multiple storms [ref]. While multi-year studies of hurriance mortality
+impacts have been conducted using disaster surveillance data, studies based on
+administrative data could identify risks for non-accidental mortality and
+morbidity that those studies likely undercount [ref]. Multi-storm studies with
+administrative data require exposure assessment that is consistent across time
+and space, allowing the administrative data to be linked to disaster exposure
+[ref]. Key to expanding tropical cyclone epidemiology are therefore exposure
+assessment methods that appropriately capture hazards of the storm that are
+relevant to human health, methods that can be applied consistently across
+multiple storms, locations, and years.
+
+The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
+standard for Atlantic-basin tropical cyclone tracks.  It records the central
+position of a tropical cyclone every six hours, as well as the storm's
+\underline{minimum pressure} and \underline{central maximum sustained winds}. This data is openly
+available through the \ac{HURDAT2}, a post-storm assessment conducted
+by the \ac{US} \ac{NHC} that incorporates data from a variety of sources,
+including satellite data and, when available, aircraft reconnaissance
+data~\parencite{landsea2013, jarvinen1988}. With this data, it is
+straightforward to measure whether a community was in the direct path, or
+within a certain distance of the path, of a storm. Some studies have indeed
+used this data to assess tropical cyclone exposure based on how near the
+storm's central track came to the county, either to assess average exposure
+patterns to tropical storms [refs] or to characterize exposure for health
+studies [refs].
+
+When epidemiological studies assess exposure to a tropical cyclone based on 
+how close the storm came to a community, however, they may misclassify exposure. 
+Tropical cyclones vary dramatically in size: \ac{US} tropical
+cyclones have been observed with radii to maximum winds as small
+as~20~\si{\kilo\metre} and as large
+as~200~\si{\kilo\metre}~\parencite{mallin2006, quiring2011variations}.  While a
 number of tropical cyclone hazards are strongly associated with distance from
 the tropical cyclone's center (e.g., wind and, at the coast, storm surge and
-waves~\parencite{rappaport2000, kruk2010}), other hazards like dangerous rain,
+waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
 floods, and tornadoes can occur well away from the tropical cyclone's central
@@
 
-Further, distance-based exposure metrics that use
-buffers tend to use an equal buffer distance on each side of the tropical
-cyclone track (e.g.,~\cite{czajkowski2011, grabich2015, grabich2016,
-zandbergen2009, tansel2010}), but the forces of a tropical cyclone tend to be
-distributed around the center in a non-symmetrical way. For example, extreme
-winds are more common to the right of the track, where counter-clockwise
-cyclonic winds move in concert with the tropical cyclone's forward 
-motion~\parencite{halverson2015}, and the fatal tornadoes associated with \ac{US} tropical
-cyclones between~1995 and~2009 occurred almost exclusively to the right of the
-tropical cyclone's track, mostly in the right front quadrant of the tropical
-cyclone~\parencite{moore2012}. Rain, conversely, is often heaviest to the left of
-the tropical cyclone's track, especially when the tropical cyclone interacts
-with other weather systems~\parencite{atallah2003, atallah2007, zhu2013variations}
-or undergoes an extratropical transition~\parencite{elsberry2002}.
+Further, when studies use distance from the storm's track to assess exposure,
+they often use an equal buffer distance on each side of the tropical cyclone
+track (e.g.,~\cite{czajkowski2011, grabich2015, grabich2016, zandbergen2009,
+tansel2010}). However, the forces of a tropical cyclone tend to be distributed
+around the center in a non-symmetrical way. Extreme winds are more common to
+the right of the track, where counter-clockwise cyclonic winds move in concert
+with the tropical cyclone's forward motion~\parencite{halverson2015}, and the
+fatal tornadoes associated with \ac{US} tropical cyclones between~1995 and~2009
+occurred almost exclusively to the right of the tropical cyclone's track,
+mostly in the right front quadrant of the tropical
+cyclone~\parencite{moore2012}. Rain, conversely, is often heaviest to the left
+of the tropical cyclone's track, especially when the tropical cyclone interacts
+with other weather systems~\parencite{atallah2003, atallah2007,
+zhu2013variations} or undergoes an extratropical
+transition~\parencite{elsberry2002}.
 
-\todo[inline]{Something about difference in other metrics}
+If a study misclassifies exposure to the hazard or hazards that cause the
+health risk being studied, the study will generate biased estimates of tropical
+cyclone risks and impacts. Further, when different studies use different
+methods to assess exposure to tropical cyclones, their results are difficult to
+meaningfully compare and aggregate. If different methods identify
+similar sets of communities as ``exposed'',  these concerns are less serious.
+However, if different methods differ substantially in which communities they
+identify as ``exposed'', it makes it very important that epidemiological
+studies are thoughtful in how they assess exposure.
 
-The use of different exposure metrics across studies hampers the development of
-a scientific consensus on tropical cyclone-related risks, as differences
-observed between studies could result from differences in exposure assessment.
-Further, if a study misclassifies exposure to the hazard or hazards that cause
-an impact, estimates of tropical cyclone risks or impacts will be biased,
-making it hard to identify true associations. These concerns are particularly
-important if there are large differences in the locations that are determined
-to be exposed based on different metrics.
-
-To investigate this, we measured county-level exposure to tropical cyclones in
-all eastern \ac{US} counties using five metrics
+To provide open-source data on tropical cyclone exposures to other researchers,
+we developed a dataset of county-level exposure to tropical cyclones in all
+eastern \ac{US} counties for five different metrics
 (Table~\ref{tab:exposuremetrics}): (1)~distance to tropical cyclone track;
 (2)~maximum sustained wind speed; (3)~cumulative rainfall; (4)~flood events;
-and (5)~tornado events.  We assessed exposure at the county level since data on
-many potential impacts are available at county-level aggregations (e.g., direct
-hurricane-related deaths~\parencite{czajkowski2011}; birth
-outcomes~\parencite{grabich2015, grabich2016}; autism
-prevalence~\parencite{kinney2008}) and since decisions and policies to prepare for,
-and respond to, tropical cyclones are often undertaken at the county
-level~\parencite{zandbergen2009, rappaport2000}.  We explored patterns in these
-exposure assessments and measured how well exposure classification agreed
-across these five metrics in terms of classifying specific counties as exposed
-to a tropical cyclone.  Finally, to make this hazard-specific tropical cyclone
-exposure data available to other researchers, we published open source software
-for county-based hurricane exposure assessment in \ac{US}
-counties~\parencite{hurricaneexposure}.
+and (5)~tornado events~\parencite{hurricaneexposure}. This data is provided at
+the county level since data on many potential impacts are available at
+county-level aggregations (e.g., direct hurricane-related
+deaths~\parencite{czajkowski2011}; birth outcomes~\parencite{grabich2015,
+grabich2016}; autism prevalence~\parencite{kinney2008}) and since decisions and
+policies to prepare for, and respond to, tropical cyclones are often undertaken
+at the county level~\parencite{zandbergen2009, rappaport2000}. Further, we
+explored patterns in these exposure assessments across storm hazards and
+measured how well exposure classification agreed across these five metrics in
+terms of classifying specific counties as exposed to a tropical cyclone.

### 2019-10-09 writing/ehp_methods.tex
Work on paper, use 34 knot threshold

--- 2019-07-29 writing/ehp_methods.tex
@@
 This study covered all counties in the eastern half of the \ac{US}
-(Figure~\ref{fig:hurrtracks}), as of the~2010 \ac{US} Decennial Census. The
+(Figure~\ref{fig:hurrtracks}) as of the~2010 \ac{US} Decennial Census. The
 study covered all tracked land-falling or near-land Atlantic-basin tropical
-cyclones between~1988 and~2015, considering all 136 storms in \ac{HURDAT2}
+cyclones between~1988 and~2015 by considering all 136 storms in \ac{HURDAT2}
 \parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
@@
 
-\begin{comment}
-These data typically give measurements of
-the tropical cyclone center's location [and central pressure / maximum
-windspeed?] at~6-\si{\hour} intervals at synoptic times (i.e., 6:00~am,
-12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}); some landfalling tropical cyclones
-have an additional observation at the time of landfall~\parencite{landsea2013}.
-\end{comment}
-
 \subsection*{Distance-based exposure metric}
 
-We first measured how close each study storm came to each study county. We
-started with tracking data from \ac{HURDAT2}, which records the storm's
-position every~6 \si{\hour}, and interpolated to~15-\si{\minute} intervals
-using natural cubic splines~\parencite{hurricaneexposure}. We then measured the
-distance between the storm's position at each~15-\si{\minute} interval and each
-county's population mean center, as assessed by the 2010 Decennial
-Census~\parencite{countycenters}, using the Great Circle (WGS84 ellipsoid)
-method~\parencite{bivand2013applied}. These measurements captured the distance
-between the storm's central track and each study county throughout the period
-when the storm was tracked. The closest distance between the storm and each
-county was recorded as the distance of the storm's closest approach to the
-county. We also recorded the time storm came closest to each county, for use in
-assessing rain-, flood-, and tornado-based metrics where observed data on the
-hazards needed to be matched by time to storm events.  To allow matching with
-data based on local time, times were converted from \ac{UTC} to local
+We first measured how close each storm came to each county. We used tracking
+data from \ac{HURDAT2}, which records the storm center's position every~6
+\si{\hour} (at the synoptic times of 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am
+\ac{UTC}), and interpolated this position to~15-\si{\minute} intervals using
+natural cubic splines~\parencite{hurricaneexposure}. At each~15-\si{\minute}
+interval, we measured the distance between the storm's center and each county's
+population mean center~\parencite{bivand2013applied, countycenters}. From these
+measurements, we determined how close each storm came to each county.
+
+We also recorded the time the storm came closest to each county. We later used
+this time to link observed data on precipitation, flooding, and tornadoes to
+the storm.  To allow matching with data recorded in local time (e.g., health
+data), we converted these times from \ac{UTC} to local
 time~\parencite{countytimezones}.
@@
 
-To estimate storm-associated rainfall, we used precipitation data from a
-re-analysis dataset, as this data was available for every county and every day
-through a continuous period of the study
-(1988\,--\,2011)~\parencite{alhamdan2014environmental, cdcwonder}.  By
-comparison, observations from ground-based monitoring networks had missing
-values spatially (i.e., for some study counties), temporally (for some days),
-or both.
+We based storm rainfall measurements on precipitation data files from the
+\ac{NLDAS-2} re-analysis dataset, which is available for the continental
+\ac{US}~\parencite{rui2013nldas}. We used data that were previously aggregated
+from this re-analysis dataset to the county level for the Centers for Disease
+Control and Prevention's \ac{WONDER} database \parencite{cdcwonder} by two of
+the coauthors (Al-Hamdan and Crosson) \parencite{alhamdan2014environmental,
+cdcwonder}. To create this aggregated data originally, these coauthors took
+hourly precipitation measurements in the \ac{NLDAS-1} precipitation files,
+which are given at~1/8\si{\degree} grid points across the continential
+\ac{US}, and summed them by day for each grid point.  This created a daily
+precipitation total for each grid point; these were then averaged for all grid
+points within a county's~1990 \ac{US} Census
+boundaries~\parencite{alhamdan2014environmental, cdcwonder}. This process
+generated daily county-level precipitation estimates for each continental
+\ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.  We matched these
+county-level daily measurements by date with storm tracks, using for each
+county and storm the date when the storm was closest to the county. Given the
+location of storm-affected counties and the typical timing of tropical
+cyclones, these precipitation measures primarily represent rainfall, although
+occasionally they may represent snowfall or other types of precipitation.
 
-We estimated daily rainfall for each study county by aggregating
-hourly~1/8\si{\degree} gridded reanalysis data from the \ac{NLDAS-2}
-precipitation data files~\parencite{rui2013nldas}. These data integrate
-satellite-based and land-based monitoring, applying a land-surface model to
-create a reanalysis dataset that is spatially and temporally complete across
-the continental \ac{US}~\parencite{rui2013nldas, alhamdan2014environmental}. To
-aggregate to county-level, the hourly data at each grid point were summed to
-create a daily rainfall total, and these grid point rainfall totals were then
-averaged across all grid points within a county's~1990 \ac{US} Census
-boundaries~\parencite{alhamdan2014environmental, cdcwonder}. These daily
-county-level estimates were then matched temporally with storm tracks, using
-the date when each storm was closest to each county. The cumulative
-storm-associated rainfall was then calculated as the sum of rainfall from two
-days before the storm's closest approach to a county to one day after, as 
-storm-related rains often precede the passing of the storm's center.
+In the open-source data, we provide daily precipitation values for each county
+for the period from five days before to three days after each storm's closest
+approach to the county.  In the software that we published in association with
+this data, we provide the functionality to aggregate these daily values to
+create cumulative precipitation estimates for custom windows as needed. For
+example, a user could determine precipitation only on the day the storm was
+closest to each county, but could also determine the cumulative rainfall for a
+more extended period, like from three days before to three days after the
+storm's closest approach. For later analysis in this study, we calculated
+storm-associated rainfall as the sum of precipitation from two days before the
+storm's closest approach to the county to one day after.
 
-To validate these rainfall estimates, we investigated a subset of counties that
-also had available ground-based precipitation observations. We selected nine
-sample counties geographically spread across storm-prone regions of the eastern
-\ac{US} and for which precipitation data were available from multiple stations
-throughout the study period: Miami-Dade, FL; Harris County, TX; Mobile County,
-AL; Orleans Parish, LA; Fulton County, GA; Charleston County, SC; Wake County,
-NC; Baltimore County, MD; and Philadelphia County, PA. We collected
-ground-based observations of precipitation data from all stations in the county
-with available daily data in the Global Historical Climatology Network
-throughout 1988\,--\,2011~\parencite{menne2012overview, rnoaa, countyweather}.
-We summed daily station-specific measurments for two days before to one day
-after each storm's closest approach and then averaged these cumulative
-station-based precipitation totals for each county as a county-wide estimate of
-cumulative storm-related precipitation. We measured the rank correlation
-(Spearman's~$\rho$~\parencite{spearman1904proof}) between storm-specific
-cumulative precipitation estimates for the two data sources within each sample
-county.
+To validate these precipitation estimates, we compared them with ground-based
+observations in a subset of study counties. We selected nine sample counties
+geographically spread across storm-prone regions of the eastern \ac{US} and for
+which precipitation data were available from multiple ground-based stations
+within the Global Historical Climatology Network throughout the study
+period~\parencite{menne2012overview, rnoaa, countyweather}. These sample
+counties were: Miami-Dade, FL; Harris County, TX; Mobile County, AL; Orleans
+Parish, LA; Fulton County, GA; Charleston County, SC; Wake County, NC;
+Baltimore County, MD; and Philadelphia County, PA. We summed daily
+station-specific measurements from two days before to one day after each
+storm's closest approach and then averaged these cumulative station-based
+precipitation totals for each county to create a county-wide estimate of
+cumulative storm-related precipitation based on ground-based monitoring. We
+measured the rank correlation (Spearman's~$\rho$~\parencite{spearman1904proof})
+between storm-specific cumulative precipitation estimates for the two data
+sources (\ac{NLDAS} re-analysis data versus ground-based monitoring) within
+each sample county.
 
@@
 
-Ground-based observations of wind speed are problematic as a metric of exposure
-to tropical cyclones, as instruments often fail at high wind speed, while
-reanalysis data, often available at hourly or higher resolution, can lack a
-fine enough temporal resolution to capture wind extremes associated with a
-tropical cyclone. Therefore, to create a dataset of county-level sustained
-winds during historical tropical cyclones, we modeled maximum sustained wind
-speeds at each county's population mean center~\parencite{countycenters}, using
-a wind speed model based on results from
-Willoughby~\parencite{willoughby2006parametric}. For each storm, we used the
-interpolated storm tracks generated for the distance metric and, for
-each~15-\si{\minute} increment, we modeled maximum ground-level sustained wind
-speed at each county center using a double exponential wind speed model with
+To create a dataset of county-level sustained winds during historical tropical
+cyclones, we modeled maximum sustained wind speeds at each county's population
+mean center~\parencite{countycenters}, using a wind speed model based on
+results from Willoughby~\parencite{willoughby2006parametric}. For each storm,
+we modeled the maximum ground-level sustained wind speed that the storm
+generated at each county center for each~15-\si{\minute} increment throughout
+the storm tracking period, using a double exponential wind speed model with
 inputs for the storm's forward speed, direction, and central wind
 speed~\parencite{willoughby2006parametric, stormwindmodel}. This model
-incorporated asymmetry in wind speeds around the tropical cyclone center that
-results from the storm's forward movement~\parencite{phadke2003modeling,
-stormwindmodel}. We then determined the maximum value of this estimate over the
-storm tracking period for each study county to determine the maximum sustained
-surface wind speed in each county during each storm.
+incorporates the asymmetry in wind speeds around the tropical cyclone center
+that results from the storm's forward movement~\parencite{phadke2003modeling,
+stormwindmodel}. We then identified the maximum sustained surface wind speed in
+each county during each storm.
 
 To validate these modeled county-level wind speed estimates, we compared them
-to county-level maximum sustained surface wind speeds based on the wind radii
-estimates in \ac{HURDAT2}, which are based on \ldots. The wind radii values
-have been included in \ac{HURDAT2} since 2004 [?] and estimate the distance
-from the storm's center to the furthest point with winds at or above winds of a
-certain speed in each four quadrants of the storm. They give estimates of these
-maximum radii for three thresholds of maximum surface wind speeds: 64,~50,
-and~34 \si{\knot}.  They therefore allow for the classification of counties
-into four categories of maximum sustained wind speeds: $<$34~\si{\knot};
-34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}. We
-interpolated this data to~15-\si{\minute} increments and classified a county as
-exposed to winds in a given wind speed category if its population mean center
-was within~85\% of the maximum radius for that wind speed in its quadrant of
-the storm [clarify why 85\%]. We compared these wind radii-based estimates with
-the modeled wind speed estimates for the study storms since 2004 for which at
-least one study county had a sustained wind of~$\ge$34 \si{\knot} (based on the
-\ac{HURDAT2} wind radii). For each of these storms, we calculated the percent
-of study counties that were classified in the same wind speed category.
+to the  post-season wind radii values, which have been routinely published
+since 2004 \parencite{knaff2016using}. These wind radii characterize storm
+size, estimating the maximum distance from the storm's center that winds of a
+certain intensity extend. They give separate estimates for four quadrants of
+the storm to capture asymmetry in the storm's wind field and are 
+based on a post-season
+re-analysis that incorporates available data (e.g., satellite data, aircraft
+reconnaissance and ground-based data if available)
+\parencite{knaff2016using}. Wind radii are estimated
+for three thresholds of maximum surface wind speeds: 64,~50, and~34 \si{\knot}.
+They therefore allow for the classification of counties into four categories of
+maximum sustained wind speeds: $<$34~\si{\knot}; 34\,--\,49.9~\si{\knot};
+50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}.
+
+We interpolated the wind radii data to~15-\si{\minute} increments and
+classified a county as exposed to winds in a given wind speed category if its
+population mean center was within~85\% of the maximum radius for that wind
+speed in that quadrant of the storm. The~85\% adjustment is based on previous
+research that found that this ratio helps capture average wind extents within a
+quadrant from these maximum wind radii \parencite{knaff2016using}.  We compared
+these wind radii-based estimates of county-level maximum sustained surface
+winds during a storm with the modeled wind speed estimates, comparing all study
+storms since 2004 for which at least one study county experienced sustained
+winds of~$\ge$34 \si{\knot} based on the \ac{HURDAT2} wind radii. For each of
+these storms, we calculated the percent of study counties that were classified
+in the same wind speed category based on both data sources.
 
@@
 counties, we used event listings from the National Oceanic and Atmospheric
-Administration (NOAA)'s Storm Event Database~\parencite{stormevents}. For each
-tropical cyclone, we identified all events with event types related to flooding
-(``Flood", ``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and
-that occurred in a county within~500 \si{\kilo\metre} of the tropical cyclone's
-track and with a start date within a five-day window centered on the date of
-the tropical cyclone's closest approach to the
-county~\parencite{hurricaneexposuredata}. ``Flood", ``Flash Flood" and
-``Tornado" events in this database were reported by county \ac{FIPS} code and
-so could be directly linked to counties.  ``Coastal Flood" events were reported
-by forecast zone; for these, the event was matched to the appropriate county if
-possible using regular expression matching of listed county
-names~\parencite{noaastormevents}. While this database has recorded storm data
-since~1950, its coverage changed substantially in~1996 to cover a wider variety
-of storm events~\parencite{stormevents}. We therefore only considered these
-metrics of hurricane exposure for tropical cyclones in~1996 and later.
+Administration (NOAA)'s Storm Event Database~\parencite{stormevents}. While
+this database has recorded storm data since~1950, its coverage changed
+substantially in~1996 to cover a wider variety of storm
+events~\parencite{stormevents}. We therefore only considered these metrics of
+hurricane exposure for tropical cyclones in~1996 and later.
 
-The tornado observations from this dataset form the traditional tornado event
-database for the \ac{US}, and so we did not conduct analyses to validate the
-tornado event data. It is difficult to characterize flooding at the county
+For each tropical cyclone, we identified all events with event types related to
+flooding (``Flood,'' ``Flash Flood,'' ``Coastal Flood'') and tornadoes
+(``Tornado'') with a start date within a five-day window centered on the date
+of the tropical cyclone's closest approach to the
+county~\parencite{hurricaneexposuredata}. To exclude events that started near
+in time to the storm but far from the storm's track, and so were likely
+unrelated to the storm, we exculed any events that occurred outside~500
+\si{\kilo\metre} of the tropical cyclone's track. ``Flood,'' ``Flash Flood,''
+and ``Tornado'' events in this database were reported by county \ac{FIPS} code
+and so could be directly linked to counties.  ``Coastal Flood'' events were
+reported by forecast zone; for these, the event was matched to the appropriate
+county if possible using regular expression matching of listed county
+names~\parencite{noaastormevents}.
+
+The tornado observations from this dataset form a traditional tornado event
+database for the \ac{US}, and so we did not further validate
+the tornado event data. It is difficult to characterize flooding at the county
 level because flooding can be very localized and can be triggered by a variety
-of causes. To investigate the extent to which the NOAA flood event data
-captures extremes that might be identified with other flooding data sources, we
+of causes. To investigate the extent to which the NOAA flood event listings
+capture extremes that might be identified with other flooding data sources, we
 investigated a sample of study counties, comparing the flood event data during
@@
 through storm-prone areas of the eastern \ac{US} and with multiple streamflow
-gages reporting data during events (Baltimore County, MD; Bergen County, NJ;
-Escambia County, FL; Fairfield County, CT; Fulton County, GA; Harris County,
-TX; Mobile County, AL; Montgomery County, PL; and Wake County, NC). For each
-county, we first identified all streamflow gages in the county that had
-complete data for Jan.~1~1996\,--\,Dec.~31,~2015. If a storm did not come
-within~500~\si{\kilo\metre} of a county, it was excluded from this analysis,
-but all other study storms were considered. For each storm and county, we
-summed daily total streamflow across all streamgages for each day in the
-five-day window around the storm's closest approach. We took the maximum daily
-streamflow total during those five days as a measure of the county's maximum
-streamflow during that storm. We also calculated the percent of streamflow
-gages in the county with a daily streamflow that exceeded a threshold of
-flooding (the streamgage's median value for annual peak
-flow~\parencite{countyfloods}) on any day during the five-day window. We
-investigate how these measurements varied between storms with associated flood
-events in the NOAA Storm Events data versus storms without an event listing, to
-explore if storms with flood event listings tended to be  associated with
-higher streamflows at gages within the county.
+gages reporting data during events. The sample counties were: Baltimore County,
+MD; Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton
+County, GA; Harris County, TX; Mobile County, AL; Montgomery County, PL; and
+Wake County, NC. For each county, we first identified all streamflow gages in
+the county with complete data for Jan.~1~1996\,--\,Dec.~31,~2015. If a storm
+did not come within~500~\si{\kilo\metre} of a county, it was excluded from this
+analysis, but all other study storms were considered.
 
-\subsection*{Binary storm exposure classifications based on different metrics}
+For each storm and county, we summed streamflow measurements across all county
+gages to generate daily totals for the five-day window around the storm's
+closest approach. We took the maximum of these daily streamflow totals as a
+measure of the county's maximum daily streamflow during that storm. We also
+calculated the percent of streamflow gages in the county with a daily
+streamflow that exceeded a threshold of flooding (the streamgage's median value
+for annual peak flow~\parencite{countyfloods}) on any day during the five-day
+window. We investigated how these measurements varied between storms with
+associated flood events in the NOAA Storm Events data versus storms without an
+event listing for the county, to explore if storms with flood event listings
+tended to be associated with higher streamflows at gages within the county.
 
-In our open-source dataset, we provide the closest distance of each storm to
-each county, as well as county-level storm-related winds and rainfall, as
-continuous metrics. However, epidemiologic studies of tropical cyclones often
-compare ``exposed'' versus ``unexposed'' communities to assess storm-related
-health risks [refs], and so we also used these continuous metrics to classify
-counties as ``exposed'' or ``unexposed'' and explored patterns in storm
-exposure based on these binary clasifications.
+\subsection*{Binary storm exposure classifications}
 
-Two of the exposure metrics (flood- and tornado-based) were inherently binary,
-since these metrics were based on whether an event was listed in the NOAA Storm
-Events database.  For the other exposure metrics, each county was classified as
-exposed to a tropical cyclone based on whether the exposure metric exceeded a
-certain threshold (Table~\ref{tab:exposuremetrics}). For the rainfall metric, a
-distance constraint was also necessary, to ensure that rains unrelated and far
-from the storm track were not misattributed to a storm. Through exploratory
-analysis, we set this distance metric at~500 \si{\kilo\metre} (i.e., for a
-county to be classified as exposed based on rainfall, the cumulative rainfall
-had to be over 75 \si{\milli\metre} and the storm must have passed within~500
-\si{\kilo\metre} of the county; Table~\ref{tab:exposuremetrics}). We set this
-distance constraint at a value that was typically large enough to capture
-storm-related rain. However, data users should note that in rare examples of
-exceptionally large storms (e.g., Hurricane Ike in 2008) or storms for which
-storm tracking was stopped at extratropical transition (e.g., Tropical Storm
-Lee in 2011), some storm-related rains may be missed because of this distance
-constraint (Figure~S1).
+In our open-source data, we provide continuous measurements of three exposure
+metrics: closest distance of each storm to each county, modeled storm winds,
+and storm-related rainfall. However, epidemiologic studies of tropical cyclones
+often use a binary exposure classification (``exposed'' versus ``unexposed'')
+to assess storm-related health risks [refs], and so we also explored patterns
+in storm exposure based on binary clasifications of these exposure metrics.
 
-We assessed patterns in county-level tropical cyclone exposure in the eastern
-\ac{US} based on each exposure metric. Depending on available exposure data,
-this assessment included some or all of the period from~1988 to~2015 (Table 1).
-For each binary metric of tropical cyclone exposure (Table 1), we first summed
-the total number of county-level exposures over available years for each
-exposure metric and mapped patterns in these exposures. We next investigated
-agreement between storm exposure classifications based on different exposure
-metrics. We calculated the within-storm Jaccard
+Two of the exposure metrics (flood- and tornado-based) were inherently binary
+in our data, since these metrics were based on whether an event was listed in
+the NOAA Storm Events database.  For the other exposure metrics, each county
+was classified as exposed to a tropical cyclone based on whether the exposure
+metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
+picked reasonable thresholds, but others could be used with the open-source
+data and its associated software.
+
+For the rainfall metric, a distance constraint was also necessary, to ensure
+that rains unrelated and far from the storm track were not misattributed to a
+storm. Through exploratory analysis, we set this distance metric at~500
+\si{\kilo\metre} (i.e., for a county to be classified as exposed based on
+rainfall, the cumulative rainfall in the county had to be over the
+75~\si{\milli\metre} threshold and the storm must have passed within~500
+\si{\kilo\metre} of the county; Table~\ref{tab:exposuremetrics}). This distance
+constraint was typically large enough to capture storm-related rain.  However,
+data users should note that in rare cases---for example, exceptionally large
+storms (e.g., Hurricane Ike in 2008) or storms for which storm tracking was
+stopped at extratropical transition (e.g., Tropical Storm Lee in 2011)---some
+storm-related rains may be missed because of this distance constraint
+(Figure~S1). This distance constraint can be customized using the software
+published in association with the open-source
+data~\parencite{hurricaneexposure}.
+
+We characterized patterns in county-level exposure in the eastern \ac{US} for
+each of the measured tropical cyclone hazards (extreme precipitation, maximum
+sustained winds, flooding, and tornadoes).  Depending on available exposure
+data, this assessment included some or all of the period from~1988 to~2015
+(Table 1). For each binary metric of exposure to a tropical cyclone hazard, we
+first summed the total number of county-level exposures over available years
+for each exposure metric and mapped patterns in these exposures. We next
+investigated agreement between storm exposure classifications based on these
+different single-hazard exposures, as well as between each of these and a
+distance-based proxy assessment of tropical cyclone exposure
+(Table~\ref{tab:exposuremetrics}). We calculated the within-storm Jaccard
 index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between each
@@
 tropical cyclone. We measured these values for all study storms that affected a
-large number of eastern \ac{US} counties (250 or more of the study counties
-classified as exposed by at least one of the exposure metrics) during the years
-when all exposure data were available (1996\,--\,2011).
+large number of eastern \ac{US} counties (those storms with 250 or more of the
+study counties classified as exposed by at least one of the exposure metrics)
+during the years when all exposure data were available (1996\,--\,2011).
 
-\subsection*{Case study: Physical exposure of electricity-dependent Medicare
-beneficiaries to tropical cyclones}
+\subsection*{Case study}
 
-Disasters' societal impacts depend both on the geophysical forces of the
-disaster and on the vulnerability of those living in the affected geographical
-areas~\parencite{chakraborty2005population, anderson2003community,
-cutter1996vulnerability}. As a case study, we explored how differences in
-tropical cyclone exposure assessments across different exposure metrics might
-influence estimates of physical exposures of a susceptible population to
-tropical cyclones. We calculated expected average physical exposures among
-electricity-dependent Medicare beneficiaries in eastern \ac{US} counties to
-tropical cyclones based on each exposure metric. This subpopulation was
-selected since it may be particularly susceptible to health impacts from
-tropical cyclone exposure, especially through the pathway of storm-associated
-power outages and evacuations.
+The societal impact of a disaster depends both on its geophysical forces and on
+the vulnerability of those living in the geographical areas it
+affects~\parencite{chakraborty2005population, anderson2003community,
+cutter1996vulnerability}. An estimate of ``physical exposure''---which
+...---can help to jointly capture these facets of disaster impacts [ref, maybe
+the peduzzi one given below?]. As a case study, we measured physical exposure
+based on each metric of exposure and investigated whether conclusions from
+these assessments varied substantially.
 
@@
 in each study county from the \ac{US} Department of Health \& Human Service's
-emPower Map~2.0~\parencite{empower}. We then calculated the physical exposure
-of the electricity-dependent Medicare population in each county, based on
-tropical cyclone assessments using each hazard metric,
-following~\parencite{peduzzi2009assessing}:
+emPower Map~2.0~\parencite{empower}. This subpopulation was selected since it
+may be particularly susceptible to health impacts from tropical cyclone
+exposure, especially through the pathway of storm-associated power outages and
+evacuations. We calculated the physical exposure of the electricity-dependent
+Medicare population in each county, based on tropical cyclone assessments using
+each exposure metric (Table~\ref{tab:exposuremetrics}), following
+\citeauthor*{peduzzi2009assessing}~(\citeyear{peduzzi2009assessing}):
 
@@
 county~$c$, as of July~2017.
-
-\begin{comment}
-When combined with estimates of vulnerability of a
-population to a natural hazard, such measurements of physical exposure can be
-used to calculate risk of human losses from the
-hazard~\parencite{peduzzi2009assessing}.
-\end{comment}

### 2019-10-09 writing/ehp_results.tex
Work on paper, use 34 knot threshold

--- 2019-07-29 writing/ehp_results.tex
@@
 
-\paragraph{Open-source dataset}
-
-We created and published open source data with multi-year, county-level
-tropical cyclone exposure data for eastern \ac{US}
-counties~\parencite{hurricaneexposuredata}. These data include continuous
+We created and published this tropical cyclone exposure data as an open-source
+dataset, so they can be used by epidemiological
+researchers~\parencite{hurricaneexposuredata}. These data include continuous
 county-level measurements for the closest distance of each storm, cumulative
@@
 include binary data on flood and tornado events associated with each storm in
-each county. With these data, we published software tools to explore and map
-the data and to integrate it with human health
+each county. To accompany these data, we published software tools to explore
+and map the data and to integrate it with human health
 datasets~\parencite{hurricaneexposure}.
@@
 We explored potential limitations in these data by comparing them with data
-from other available sources.
-
-\paragraph{Rainfall data validation}
-
-For estimates of storm-associated rainfall, we compared estimates from NLDAS-2
-re-analysis precipitation data (included as the rainfall estimates in the
-open-source data) with ground-based observations in nine sample counties
-(Figure~\ref{fig:raincomparison}). Within these counties, storm-related
-rainfall measurements were well-correlated between the two data sources, with
-rank correlations (bottom right of each graph in
+from other available sources.  For estimates of storm-associated rainfall, we
+compared estimates from NLDAS-2 re-analysis precipitation data (used as the
+rainfall estimates in the open-source data) with ground-based observations in
+nine sample counties (Figure~\ref{fig:raincomparison}). Within these counties,
+storm-related rainfall measurements were well-correlated between the two data
+sources, with rank correlations (bottom right of each graph in
 Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some evidence
 that the reanalysis data may tend to underestimate rainfall totals in storms
-with extremely high rainfall, based a few heavy-rainfall storms in Harris
+with extremely high rainfall, based on a few heavy-rainfall storms in Harris
 County, TX, Mobile County, AL, Charleston County, SC, and Wake County, NC
 (Figure~\ref{fig:raincomparison}). However, even in these cases, it was rare
-for a storm to be classified differently (based on the precipitation threshold
-of 75~\si{\milli\metre} we used for exposure classification for later analysis)
-based on the source of precipitation data---horizontal and vertical lines in
-each small plot in Figure~\ref{fig:raincomparison} show the threshold of
-75~\si{\milli\metre}, so storms in the lower left and upper right quadrants
+for a storm to be classified differently (under the precipitation threshold
+of 75~\si{\milli\metre} we used for binary exposure classification for later
+analysis) based on the source of precipitation data---horizontal and vertical
+lines in each small plot in Figure~\ref{fig:raincomparison} show the threshold
+of 75~\si{\milli\metre}, so storms in the lower left and upper right quadrants
 would be classified the same (``exposed'' or ``unexposed'') regardless of the
 precipitation data source, while storms in the upper left and lower right
-quadrants would be classified differently.
+quadrants would be classified differently. Such cases are rare.
 
-\paragraph{Wind speed data validation}
-
-For maximum sustained surface wind speed estimates, we found that the modeled
+For maximum sustained surface wind estimates, we found that the modeled
 estimates, which we included as the primary wind exposure metric in the open
@@
 storm's most severe winds at landfall but then underestimated the extent of
-34\,--\,49.9 \si{\knot} storm winds further inland (Figure~S2). For researchers
-who would like to conduct sensitivity analysis between these two sets of wind
-data, we have included estimates from the \ac{HURDAT2} wind radii as a
-secondary measure of wind speeds in the open-source dataset
-\parencite{hurricaneexposuredata}.
-
-\paragraph{Flood data analysis}
+34\,--\,49.9 \si{\knot} storm winds further inland (Figure~S2). For
+epidemiology researchers who would like to conduct sensitivity analysis
+between these two sets of wind data, we have included estimates from the
+\ac{HURDAT2} wind radii as a secondary measure of wind speeds in the
+open-source dataset \parencite{hurricaneexposuredata}.
 
 For flood data, we compared flood status, based on event listings in the NOAA
-Storm Events database, to streamflow measurements at \ac{US} Geological Survey
-gages within nine sample counties (Figure~\ref{fig:floodcomparison}).
-
-For flood data, we compared flood status, based on event listings in the NOAA
-Storm Events database, to streamflow measurements at \ac{US} Geological Survey
-gages within nine sample counties (Figure~\ref{fig:floodcomparison}).  Across
-the sampled counties, streamgage data generally indicated higher discharge in
-on dates identified as flood events, particularly in counties with more
-reporting stream gages (e.g., Harris County, TX, with~10 reporting stream
+Storm Events database and included in the open-source data, to streamflow
+measurements at \ac{US} Geological Survey gages within nine sample counties
+(Figure~\ref{fig:floodcomparison}). Across the sampled counties, streamgage
+data generally indicated higher discharge during periods identified as flood
+events based on the NOAA Storm Events database, particularly in counties with
+more reporting streamgages (e.g., Harris County, TX, with~10 reporting stream
 gages; Figure~\ref{fig:floodcomparison}). There were some cases, however, where
@@
 were one or two tropical cyclones in several of the counties (Mobile County,
-AL, Escambia County, FL, and Montgomery County, PA) which were classified as
-flood events based on NOAA storm event listings but for which the total
-discharge across county streamflow gages was below the 25\textsuperscript{th}
-percentile of measurements taken during storms that did not have a flood event
-listing. For tropical cyclones that did not have an associated flood listing in
-the NOAA Storm Events database, most had lower total streamflow discharge
-across county gages compared to storms with flood event listings, and in all
-cases but one all gages were below the flooding threshold.  The exception was
-one storm in Fulton County, GA (Ida in 2009). [More on which storm and why---I
-think there was already a flood event that had started earlier, so it missed
-the cutoff for the startdate being within the range.]
+AL, Escambia County, FL, and Montgomery County, PA) with associated flood event
+listings but for which the total discharge across county streamflow gages was
+below the 25\textsuperscript{th} percentile of measurements during storms
+without a flood event listing in that county. For storms without a flood
+listing for the county, in most cases the total streamflow discharge in
+the county was lower than during storms for which there were flood
+event listings. Further, in all but one of these cases, all streamgage flows
+were below the flooding threshold. The exception was one storm in Fulton
+County, GA (Hurricane Ida in 2009).
 
-\subsection*{Patterns in tropical cyclone exposures in eastern \ac{US} counties}
+\subsection*{Patterns in tropical cyclone exposures}
 
-We used the storm exposure metrics to classify counties as exposed or unexposed
-to specific tropical cyclones, and we explored patterns in these exposure
-classifications over years with available data
-(Table~\ref{tab:exposuresummaries}). Across the five exposure metrics, there
-was wide variation in the average number of county exposures per year. For the
-tornado metric, there were approximately~50 county exposures per
-year, on average, within our study.  For flood and wind metrics, there were
-substantially more county exposures ($>$200 / yr on average), and
-even more when assessing exposure using the distance and rain metrics ($>$400 /
-yr on average).
+We used this storm exposure data to classify counties as exposed or unexposed
+to four different hazards for each tropical cyclone and then explored patterns
+over years with available data. Across the four storm hazards considered,
+there was wide variation in the average number of county exposures per year
+(Table~\ref{tab:exposuresummaries}). For the tropical cyclone tornado exposure,
+there were on average about 50 county exposures per year within our study. For
+tropical cyclone flood and wind exposures, county exposures were much more
+frequent ($>$200 / yr on average), and even more when assessing tropical
+cyclone rain exposure ($>$400 / yr on average).
 
-We also measured the average number of county exposures per storm under each
-exposure metric (limiting this analysis to the tropical cyclones that exposed
-at least one county to that hazard; Table~\ref{tab:exposuresummaries}).
-Tropical cyclones typically exposed only a few counties, if any, to tornadoes
-(median counties exposed per storm:~7), more counties under the flood and wind
-metrics (median:~26 and~34, respectively), and even more counties for the
-metrics of rain (median:~68) and distance (median:~84).  For every metric
-except the tornado-based metric, we identified at least one tropical cyclone
-with over~300 counties exposed.  However, the largest-extent tropical cyclone
-varied by metric (e.g., Beryl in~1994 exposed the most counties based on
-distance, Frances in~2004 based on rain, and Ike in~2008 based on wind).
+We also measured the average number of counties exposed per storm to each
+hazard (limiting this analysis to the tropical cyclones that exposed at least
+one county to that hazard; Table~\ref{tab:exposuresummaries}).  Tropical
+cyclones typically exposed only a few counties, if any, to tornadoes (median
+counties exposed per storm:~7), more counties to flooding and high winds
+(median:~26 and~34, respectively), and even more counties to extreme rain
+(median:~68).  For every hazard except tropical cyclone tornadoes, we
+identified at least one tropical cyclone that exposed over~300 counties.
+However, the largest-extent tropical cyclone varied across hazards: Frances
+in~2004 exposed the most counties based on rain, Ike in~2008 based on wind, and
+Ivan in 2004 based on flooding and tornadoes
+(Table~\ref{tab:exposuresummaries}).
 
 When we calculated and mapped the average number of exposures per decade in
-each county for each exposure metric (Figure~\ref{fig:averageexposure}), strong
-geographical patterns were clear. Wind-based exposure had a strong coastal
-pattern, with almost all exposures in counties within about~200
-\si{\kilo\metre} (124~mi) of the coastline. While exposures by rain- and
-distance-based metrics were also more common in coastal areas compared to
-inland areas, they also created inland exposures that were not identified by
-the wind-based metric. Rain and, to some extent, flood exposures were
-characterized by a pattern defined by the Appalachian Mountains, with much
-lower exposures west of the mountain range than to the east. Almost all
-tornado-based exposures were in coastal states, with most in Florida, Alabama,
-South Carolina, and North Carolina and almost none north of Maryland.
+each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
+strong geographical patterns were clear. Tropical cyclone wind exposure had a
+strong coastal pattern, with almost all exposures in counties within about~200
+\si{\kilo\metre} (124~mi) of the coastline. While tropical cyclone rain
+exposure was also more frequent in coastal areas compared to inland areas,
+there were also inland exposures in counties not often  that were not
+identified as exposed by the wind-based metric. Rain and, to some extent, flood
+exposures were characterized by a pattern defined by the Appalachian Mountains,
+with many fewer exposures west of the mountain range than to the east. Almost
+all tornado-based exposures were in coastal states, with most in Florida,
+Alabama, South Carolina, and North Carolina and almost none north of Maryland.
 Flood-based exposures were highest in North Carolina and along the South
@@
 distance (Figure~\ref{fig:averageexposure}).  Patterns were similar when
-analysis was restricted to years with all exposure data available
-(1996\,--\,2011; Figure~S3).
+analysis was restricted to years with exposure data for all four hazards
+available (1996\,--\,2011; Figure~S3).
 
-\subsection*{Agreement in county-specific exposure classification between exposure metrics}
+\subsection*{Agreement across exposure metrics}
 
@@
 of metrics. We found that these exposure classifications typically did not
-strongly agree between pairs of metrics, indicating that the set of counties
-identified as exposed based on one metric can have limited overlap with the
-set of counties identified as exposed to the storm by another metric.
+strongly agree between pairs of metrics. The set of counties
+identified as ``exposed'' based on one metric often overlap little with
+the set identified as ``exposed'' by another metric.
 
-Figure~\ref{fig:ivanexposure} shows Hurricane Ivan in 2004 as an example. For
-the distance-based metric, the counties assessed as exposed followed the
+Figure~\ref{fig:ivanexposure} shows shows an example: Hurricane Ivan in 2004.
+For the distance-based metric, the counties assessed as exposed followed the
 tropical cyclone's track, including counties well inland. For the wind metric,
@@
 We drew similar conclusions when we investigated all~46 tropical cyclones
-between~1996 and~2011 (the period for which all five metrics were available)
-for which~250 or more counties were exposed based on at least one metric
-(Figure~\ref{fig:jaccard}). In this figure, each row provides results for a
-specific tropical cyclone, and each box in that row shows the measured Jaccard
-coefficient between two of the exposure metrics for that storm.  For most
-tropical cyclones, distance-based county-level exposure assessment was at best
-in moderate agreement with exposure assessments based on the four hazard-based
-metrics. For almost all tropical cyclones, Jaccard index values between
-distance and any of the four hazard-based metrics were~$\le0.6$ (i.e., out of
-the counties assessed as exposed to the tropical cyclone by at least one of the
-two metrics considered,~60\si{\percent} or fewer of the counties had the same
-exposure assessment under the two metrics). For many tropical cyclones, the
-Jaccard index values were~$\le0.2$, suggesting very poor agreement in exposure
-assessment between the metrics.  Among the four hazard-based metrics, the
-tornado-based metric showed universally poor agreement with other metrics in
-county-level classification across all tropical cyclones considered.  For other
-pairs of metrics, there were also generally large differences in which counties
-were determined to be exposed, with the Jaccard index values~$<0.4$ for most
-metric pairs in most tropical cyclones.
+between~1996 and~2011 (the period for which data for all five metrics were
+available) for which~250 or more counties were exposed based on at least one
+metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
+for one tropical cyclone, and each box in that row shows the measured Jaccard
+coefficient a pair of exposure metrics. For most tropical cyclones,
+distance-based county-level exposure assessment was, at best, in moderate
+agreement with exposure assessments based on the four hazard-based metrics. For
+almost all tropical cyclones, Jaccard index values between distance and any of
+the four hazard-based metrics were~$\le0.6$ (i.e., out of the counties assessed
+as exposed by at least one of the two metrics in the pair,~60\si{\percent} or
+fewer were assessed the same under both metrics in the pair). For many tropical
+cyclones, the Jaccard index values were~$\le0.2$, indicating poor agreement.
+Among the four hazard-based metrics, the tornado-based metric had
+universally poor agreement with other metrics in county-level classification
+across the tropical cyclones considered.  For other pairs of metrics, there was
+also generally disagreement in which counties were determined to be exposed,
+with Jaccard index values~$<0.4$ for most metric pairs in most tropical
+cyclones.
 
-\begin{comment}
-Storms are displayed within
-clusters that have similar patterns in county-level exposure agreement for
-metric pairs, based on hierarchical clustering using the complete link
-method~\parencite{murtagh2012algorithms} (i.e., storms in the same cluster tend
-to have similar patterns for the pairwise strength of agreement among metrics);
-columns are also ordered based on hierarchical clustering. 
-Maps are
-available showing the counties identified as exposed under each of five metrics
-for the widest-extent storm in each cluster: Hurricane Ivan in~2004
-(Figure~\ref{fig:ivanexposure}) and Hurricanes Floyd in~1999, Lee in~2011, Cindy
-in~2005, and Katrina in~2005 (Figure~S4).
-\end{comment}
-
-There were, however, a few exceptions\,---\,tropical cyclones in which similar
-counties were determined to be exposed to the tropical cyclone for two or more
-of the metrics considered.  For Floyd in~1999 (Figure~S4) and Irene in~2011,
-for example, county-level classification agreed moderately to well (Jaccard
-index of~$\sim 0.5\mbox{\,--\,}0.8$) for all pairs of exposure metrics except
-those including the tornado-based metric. For another set of tropical cyclones
-(e.g., Lee in~2011 [Figure~S3], Ernesto in~2006, and Bertha in~1996), there was
-moderate to good agreement for all pairwise combinations of distance, rain, and
-wind, but poor agreement for all other combinations of metrics. However, these
-two sets of tropical cyclones represented the minority of all tropical cyclones
-considered; for most tropical cyclones, there was low overlap in the counties
-determined to be exposed to the tropical cyclones between most pairings of
-metrics.
+There were, however, exceptions\,---\,tropical cyclones in which exposure
+assessment agreed well across several of the metrics considered.  For Floyd
+in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
+agreed moderately to well (Jaccard index of~$\sim 0.5\mbox{\,--\,}0.8$) for all
+pairs of exposure metrics except those including the tornado-based metric. For
+another set of tropical cyclones (e.g., Lee in~2011 [Figure~S3], Ernesto
+in~2006, and Bertha in~1996), there was moderate to good agreement for all
+pairwise combinations of distance, rain, and wind, but poor agreement for all
+other combinations of metrics.  However, these two sets of tropical cyclones
+represented the minority of tropical cyclones considered; for most tropical
+cyclones, there was low overlap in the counties determined to be exposed to the
+tropical cyclones between most pairings of metrics.
 
@@
 
-Finally, we conducted a case study to investigate how these differences in
-exposure assessment across the exposure metrics could affect estimates of
-expected physical exposure among a vulnerable subpopulation,
-electricity-dependent Medicare beneficiaries.
+We conducted a case study comparing estimates of physical exposure to
+tropical cyclones that were measured using each exposure metric.
 Figure~\ref{fig:topelecdependexposure} shows the study counties with the
-highest expected physical exposures in this subpopulation based on each
-exposure metric. A few counties (Miami-Dade County, FL; Harris County, TX) were
-ranked in the top three counties for expected physical exposure in this
-subpopulation for almost all exposure metrics
+highest expected physical exposures to tropical cyclones among
+electricity-dependent Medicare beneficiaries based on each metric. A few
+counties (Miami-Dade County, FL; Harris County, TX) were ranked in the top
+three under almost all exposure metrics
 (Figure~\ref{fig:topelecdependexposure}). However, there were key differences
 across exposure metrics. For example, most of the top counties based on wind-
-and tornado-based metrics were in Florida, while over half of the counties with
-highest expected physical exposures for the flood-based metric were in
-Mid-Atlantic states (PA, NY).
+and tornado-based metrics were in Florida, while over half of the top counties
+based on the flood-based metric were in Mid-Atlantic states (PA, NY).

### 2019-10-09 writing/ehp_tables.tex
Work on paper, use 34 knot threshold

--- 2019-01-27 writing/ehp_tables.tex
@@
 Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 \si{\kilo\metre} of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
-Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
-Tornado & 1996--2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track\\
+Distance & 1988\,--\,2015 & Storm track came within 100~\si{\kilo\metre} of county population mean center\\ 
+Rain & 1988\,--\,2011 & County cumulative rainfall of $\ge$75~\si{\milli\metre} over the period from two days before to 
+one day after the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\ 
+Wind & 1988\,--\,2015 & Modeled storm-associated maximum sustained winds of $\ge$34~kts 
+at the county's population mean center\\ 
+Flood & 1996\,--\,2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center \\
+Tornado & 1996\,--\,2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\
 \bottomrule 

### 2019-10-09 R/average_elec_depend_exposure_figure.R

@@ -9,7 +9,7 @@ library(maptools)
 library(RColorBrewer)
 library(choroplethr)
 data(closest_dist)
-load("clean_empower.RData") # Loads dataframe named `medi`
+load("data/clean_empower.RData") # Loads dataframe named `medi`
 
 my_fips <- unique(closest_dist$fips)
 dist <- county_distance(counties = my_fips, start_year = 1988,
@@ -26,7 +26,7 @@ rain <- county_rain(counties = my_fips, start_year = 1988,
         mutate(metric = "rain")
 
 wind <- county_wind(counties = my_fips, start_year = 1988,
-                             end_year = 2015, wind_limit = 15) %>%
+                             end_year = 2015, wind_limit = 17.5) %>%
         group_by(fips) %>%
         dplyr::summarize(n_exposures = n() / 28) %>%
         mutate(metric = "wind")

### 2019-10-09 R/average_exposure_figure.R

@@ -25,7 +25,7 @@ rain <- county_rain(counties = my_fips, start_year = 1988,
         mutate(metric = "rain")
 
 wind <- county_wind(counties = my_fips, start_year = 1988,
-                             end_year = 2015, wind_limit = 15) %>%
+                             end_year = 2015, wind_limit = 17.4) %>%
         group_by(fips) %>%
         dplyr::summarize(n_exposures = n() / 28) %>%
         mutate(metric = "wind")
@@ -78,6 +78,9 @@ out_data$value <- cut(out_data$n_exposures, c(0,
                                        "6-7", "7-8", "8+"),
                             include.lowest = TRUE, right = FALSE)
 
+out_data <- out_data %>% 
+        dplyr::filter(metric != "Distance-based metric")
+
 
 out <- ggplot2::ggplot() +
         ggplot2::geom_polygon(data = out_data,
@@ -101,14 +104,14 @@ out <- ggplot2::ggplot() +
                                     discrete = TRUE, direction = -1, option = "A") +
         viridis::scale_color_viridis(name = "Hits per\ncounty per\ndecade",
                                      discrete = TRUE, direction = -1, option = "A") +
-        facet_wrap(~ metric) +
-        theme(legend.position = c(.85, .3),
+        facet_wrap(~ metric, ncol = 1) +
+        theme(legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size = unit(0.8, "lines"),
               legend.title = element_text(size = 8),
               legend.text = element_text(size = 7),
-              strip.background = element_rect(colour = "black", fill = "white")) # +
-        # ggtitle("Average county storm exposures per decade by metric")
+              strip.background = element_rect(colour = "black", fill = "white"))  +
+         ggtitle("")
 
 storm <- "Ivan-2004"
 
@@ -131,7 +134,7 @@ rain_data <- filter_storm_data(storm = storm,
 wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
                               output_vars = c("fips", "vmax_sust")) %>%
         `colnames<-`(c("fips", "wind_value")) %>%
-        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 17.4) %>%
         dplyr::select(fips, wind) %>%
         dplyr::tbl_df()
 
[…]

### 2019-10-09 R/flood_comparison_figure.R

@@ -103,7 +103,7 @@ out <- county_gage_data %>%
                county_cd = factor(county_cd)) %>%
         ggplot(aes(x = flood, y = max_total)) +
         geom_boxplot(color = "darkgray", fill = NA, outlier.alpha = 0, width = 0.3) +
-        geom_point(aes(color = perc_flooded), size = 1.7,
+        geom_point(aes(fill = perc_flooded), color = "black", shape = 21, size = 1.7,
                    position = position_jitter(width = 0.1)) +
         facet_wrap(~ county_cd, ncol = 3, scales = "free_x") +
         theme_classic() +
@@ -111,7 +111,7 @@ out <- county_gage_data %>%
         coord_flip() +
         scale_y_log10() +
         theme(legend.position = "bottom") +
-        scale_color_viridis(name = "% streamflow gages over threshold for flooding",
+        scale_fill_viridis(name = "% streamflow gages over threshold for flooding",
                             breaks = c(0, .5, 1), labels = c("0%", "50%", "100%")) +
         theme(plot.margin = unit(c(5.5, 11, 5.5, 11), "points"))
 

### 2019-10-09 R/jaccard_heatmap.R

@@ -19,9 +19,9 @@ calc_jaccard <- function(x){
         jaccard <- proxy::simil(t(x_mat), method = "jaccard")
         out <- combn(colnames(x_mat), 2) %>%
                 t() %>%
-                tbl_df() %>%
-                dplyr::rename(metric_1 = V1,
-                              metric_2 = V2) %>%
+                as_tibble(.name_repair = "unique") %>%
+                dplyr::rename(metric_1 = `...1`,
+                              metric_2 = `...2`) %>%
                 dplyr::mutate(jaccard = as.vector(jaccard))
         return(out)
 }
@@ -44,9 +44,9 @@ calc_metric_1_not_2 <- function(x){
         out_metric <- outer(1:5, 1:5, FUN = Vectorize(met_1_not_2))
         out <- combn(colnames(x_mat), 2) %>%
                 t() %>%
-                tbl_df() %>%
-                dplyr::rename(metric_1 = V1,
-                              metric_2 = V2) %>%
+                as_tibble(.name_repair = "unique") %>%
+                dplyr::rename(metric_1 = `...1`,
+                              metric_2 = `...2`) %>%
                 dplyr::mutate(met_1_not_2 = as.vector(t(out_metric))[t(upper.tri(out_metric))])
         return(out)
 }
@@ -70,9 +70,9 @@ calc_metric_2_not_1 <- function(x){
         out_metric <- outer(1:5, 1:5, FUN = Vectorize(met_2_not_1))
         out <- combn(colnames(x_mat), 2) %>%
                 t() %>%
-                tbl_df() %>%
-                dplyr::rename(metric_1 = V1,
-                              metric_2 = V2) %>%
+                as_tibble(.name_repair = "unique") %>%
+                dplyr::rename(metric_1 = `...1`,
+                              metric_2 = `...2`) %>%
                 dplyr::mutate(met_2_not_1 = as.vector(t(out_metric))[t(upper.tri(out_metric))])
         return(out)
 }
@@ -90,7 +90,7 @@ rain_data <- county_rain(counties = my_fips, start_year = 1996, end_year = 2011,
         dplyr::mutate(rain = TRUE)
 
 wind_data <- county_wind(counties = my_fips, start_year = 1996, end_year = 2011,
-                         wind_limit = 15) %>%
+                         wind_limit = 17.4) %>%
         select(storm_id, fips) %>%
         dplyr::mutate(wind = TRUE)
 
[…]

### 2019-10-09 R/other_top_storm_exposures.R

@@ -26,7 +26,7 @@ rain_data <- filter_storm_data(storm = storm,
 wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
                              output_vars = c("fips", "vmax_sust")) %>%
         `colnames<-`(c("fips", "wind_value")) %>%
-        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 17.5) %>%
         dplyr::select(fips, wind) %>%
         dplyr::tbl_df()
 
@@ -74,6 +74,9 @@ storm_data <- distance_data %>%
 
 out_data <- hurricaneexposure:::get_eastern_map() %>%
         dplyr::left_join(storm_data, by = "fips")
+out_data <- out_data %>% 
+        dplyr::filter(metric != "Distance-based metric")
+
 out <- ggplot2::ggplot() +
         ggplot2::geom_polygon(data = out_data,
                               ggplot2::aes_(x = ~ long, y = ~ lat, group = ~ group,
@@ -101,8 +104,8 @@ out <- ggplot2::ggplot() +
         # ggplot2::scale_fill_manual(name = "",
         #                            values = c("white", "dodgerblue"),
         #                            labels = c("Unexposed", "Exposed")) +
-        facet_wrap(~ metric, ncol = 3) +
-        theme(legend.position=c(.8,.15))
+        facet_wrap(~ metric, ncol = 4) +
+        theme(legend.position="none")
 
 floyd <- map_tracks(storm, plot_object = out) +
         ggtitle("Floyd (1999)")
@@ -176,6 +179,10 @@ storm_data <- distance_data %>%
 
 out_data <- hurricaneexposure:::get_eastern_map() %>%
         dplyr::left_join(storm_data, by = "fips")
+
+out_data <- out_data %>% 
+        dplyr::filter(metric != "Distance-based metric")
+
 out <- ggplot2::ggplot() +
         ggplot2::geom_polygon(data = out_data,
                               ggplot2::aes_(x = ~ long, y = ~ lat, group = ~ group,
@@ -203,8 +210,8 @@ out <- ggplot2::ggplot() +
         # ggplot2::scale_fill_manual(name = "",
         #                            values = c("white", "dodgerblue"),
         #                            labels = c("Unexposed", "Exposed")) +
-        facet_wrap(~ metric, ncol = 3) +
-        theme(legend.position=c(.8,.15))
[…]

### 2019-10-09 R/storm_severity_table.R

@@ -44,7 +44,7 @@ rain_sum <- rain_exposure %>%
             which_max = storm_id[which.max(n)])
 
 wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
-                             end_year = 2015, wind_limit = 15) %>%
+                             end_year = 2015, wind_limit = 17.4) %>%
         select(storm_id, fips) %>%
         mutate(exposed = TRUE,
                storm_id = as.character(storm_id))
@@ -161,6 +161,7 @@ total_sum <- bind_rows(bind_cols(dist_sum, dist_sum2),
   select(metric, yearly_sum, sum, max_storm)
 
 total_sum %>% 
+  filter(metric != "Distance") %>% 
   rename(Metric = metric,
          `Mean (interquartile range) of county exposures per year` = yearly_sum,
          `Median (interquartile range) of county exposures per tropical cyclone` = sum,

### 2019-10-11 R/average_elec_depend_exposure_figure.R
Update to use 34 knot threshold for wind (over 17.4 m/s)

@@ -26,7 +26,7 @@ rain <- county_rain(counties = my_fips, start_year = 1988,
         mutate(metric = "rain")
 
 wind <- county_wind(counties = my_fips, start_year = 1988,
-                             end_year = 2015, wind_limit = 17.5) %>%
+                             end_year = 2015, wind_limit = 17.4) %>%
         group_by(fips) %>%
         dplyr::summarize(n_exposures = n() / 28) %>%
         mutate(metric = "wind")

### 2019-10-11 R/jaccard_heatmap.R

@@ -260,12 +260,12 @@ ha_row <- HeatmapAnnotation(counties_exposed = counties_exposed$counties_exposed
                                                                                    title_gp = gpar(fontsize = 9,
                                                                                                    lineheight = 0.7,
                                                                                                    font = 2))),
-                            # col = list(counties_exposed = colorRamp2(c(100, 275, 450, 625),
-                            #                                          viridis::viridis(4, option = "D", direction = -1))),
+                            col = list(counties_exposed = colorRamp2(c(100, 275, 450, 625),
+                                                                     viridis::viridis(4, option = "D", direction = -1))),
                             which = "row", width = unit(0.25, "cm"))
 
 heat_map <- Heatmap(for_heatmap, col = viridis(256, option = "A", direction = -1),
-        cluster_rows = row_dend, split = 5,
+        cluster_rows = row_dend, split = 5, show_row_dend = FALSE,
         row_names_side = "left", row_names_gp = gpar(fontsize = 8),
         show_column_names = FALSE, show_column_dend = FALSE, #column_dend_side = "bottom",
         column_dend_height = unit(6, "mm"),
@@ -274,7 +274,8 @@ heat_map <- Heatmap(for_heatmap, col = viridis(256, option = "A", direction = -1
         # bottom_annotation = ha_col,
         top_annotation = ha_rot_cn,
         rect_gp = gpar(col = "white", lwd = 1),
-        row_dend_reorder = TRUE)
+        row_dend_reorder = TRUE, 
+        row_title = NULL)
 
 pdf(file = "figures/jaccard_heatmap.pdf", height = 8.4, width = 5)
 draw(heat_map + ha_row, padding = unit(c(2, 2, 18, 2), "mm"), annotation_legend_side = "bottom")

### 2019-10-11 R/other_top_storm_exposures.R

@@ -26,7 +26,7 @@ rain_data <- filter_storm_data(storm = storm,
 wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
                              output_vars = c("fips", "vmax_sust")) %>%
         `colnames<-`(c("fips", "wind_value")) %>%
-        dplyr::mutate_(wind = ~ wind_value >= 17.5) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 17.4) %>%
         dplyr::select(fips, wind) %>%
         dplyr::tbl_df()
 
@@ -110,7 +110,7 @@ out <- ggplot2::ggplot() +
 floyd <- map_tracks(storm, plot_object = out) +
         ggtitle("Floyd (1999)")
 
-storm <- "Lee-2011"
+storm <- "Ike-2008"
 
 distance_data <- filter_storm_data(storm = storm,
                                    output_vars = c("fips", "storm_dist")) %>%
@@ -131,7 +131,7 @@ rain_data <- filter_storm_data(storm = storm,
 wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
                               output_vars = c("fips", "vmax_sust")) %>%
         `colnames<-`(c("fips", "wind_value")) %>%
-        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 17.4) %>%
         dplyr::select(fips, wind) %>%
         dplyr::tbl_df()
 
@@ -213,114 +213,10 @@ out <- ggplot2::ggplot() +
         facet_wrap(~ metric, ncol = 4) +
         theme(legend.position="none")
 
-lee <- map_tracks(storm, plot_object = out) +
-        ggtitle("Lee (2011)")
-
-storm <- "Cindy-2005"
-
-distance_data <- filter_storm_data(storm = storm,
-                                   output_vars = c("fips", "storm_dist")) %>%
-        dplyr::mutate_(distance = ~ storm_dist <= 100) %>%
-        dplyr::select(fips, distance) %>%
-        dplyr::tbl_df()
-
-rain_data <- filter_storm_data(storm = storm,
-                               days_included = c(-2, -1, 0, 1),
-                               include_rain = TRUE,
-                               output_vars = c("fips", "tot_precip",
-                                               "storm_dist")) %>%
-        dplyr::mutate_(rain = ~ tot_precip >= 75 &
-                               storm_dist <= 500) %>%
[…]

### 2019-10-13 tables/metric_definitions.tex
Work on writing

--- 2019-01-26 tables/metric_definitions.tex
@@
 \begin{tabular}{p{0.9cm}p{2.5cm}p{9cm}} 
-Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Tornado & 1996--2015 & Tornado event listed in the NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
+Metric & Available & Criteria for exposure for binary analysis \\ \midrule 
+Distance & 1988\,--\,2015 & County population mean center within 100 kilometers of storm track \\ 
+Rain & 1988\,--\,2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
+Wind & 1988\,--\,2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
+Flood & 1996\,--\,2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
+Tornado & 1996\,--\,2015 & Tornado event listed in the NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
 \bottomrule 

### 2019-10-13 writing/darlings_intro.tex
Work on writing

@@
 non-communicable diseases [refs].
+
+, including Sandy in 2012 (e.g., \parencite{swerdel2014}) and Katrina
+in 2005 (e.g., \parencite{burton2009health}),
+
+A recent review highlights the need to supplement these single-storm studies
+with multi-year, multi-storm studies, to identify patterns that are common
+across multiple storms [ref]. While multi-year studies of hurriance mortality
+impacts have been conducted using disaster surveillance data, studies based on
+administrative data could identify risks for non-accidental mortality and
+morbidity that those studies likely undercount [ref]. Multi-storm studies with
+administrative data require exposure assessment that is consistent across time
+and space, allowing the administrative data to be linked to disaster exposure
+[ref]. Key to expanding tropical cyclone epidemiology are therefore exposure
+assessment methods that appropriately capture hazards of the storm that are
+relevant to human health, methods that can be applied consistently across
+multiple storms, locations, and years.
+
+If a study misclassifies exposure to the hazard or hazards that cause the
+health risk being studied, the study will generate biased estimates of tropical
+cyclone risks and impacts. Further, when different studies use different
+methods to assess exposure to tropical cyclones, their results are difficult to
+meaningfully compare and aggregate. If different methods identify
+similar sets of communities as ``exposed'',  these concerns are less serious.
+However, if different methods differ substantially in which communities they
+identify as ``exposed'', it makes it very important that epidemiological
+studies are thoughtful in how they assess exposure.
+
+ and
+measured how well exposure classification agreed across these five metrics in
+terms of classifying specific counties as exposed to a tropical cyclone.

### 2019-10-13 writing/ehp_discussion.tex
Work on writing

@@
 during disasters, to what degree, and to whom [Ibrahim editorial; Noji paper].
-As a result, these studies help improve disaster preparedness and response
-[Noji paper].  However, tropical cyclones are multi-hazard events, making it
-complicated to measure exposure.  Here, we provide open-source county-level
-data on multiple tropical cyclone exposure metrics and explore limitations in
-that data.
+\parencite{ibrahim2005unfortunate, noji2005disasters}.  As a result, these studies help improve
+disaster preparedness and response \parencite{noji2005disasters}.  However,
+tropical cyclones are multi-hazard events, making it complicated to measure
+exposure.  Here, we provide open-source county-level data on multiple tropical
+cyclone exposure metrics and explore limitations in that data.
 
@@
 
-For the flooding events data, we found that flood event status determined based
+For the flooding data, we found that flood event status determined based
 on the NOAA Storm Events listings often agreed with measurements from \ac{USGS}
 streamgages, with a flood event more likely to be listed if a storm elevated
-stream flows in the county (Figure~\ref{fig:floodcomparison}).  However, there
+streamflow across the county (Figure~\ref{fig:floodcomparison}).  However, there
 are differences between the two flooding datasets, and these highlight both the
@@
 case the flooding associated with Ida was incorporated into an ongoing flood
-event listing, with a start date well before the five-day window used to
-temporally match storm event listings with tropical cyclone tracks used here.
+event listing, with a start date well before the five-day window we used to
+temporally match storm event listings with tropical cyclone tracks.
 
@@
 tracks with storm event listings---without criteria for temporally matching
-event dates to storm tracks, it is likely that many false positives would be
+event start dates to storm dates, many false positives would be
 captured, for which the occurrance of a storm in the midst of an ongoing storm
@@
 storm contributes meaningfully to an ongoing event, but the event is not
-captured for the storm.
+captured for the storm in the exposure data.
 
@@
 
-We found that agreement was typically low across hazard-specific tropical
-exposure classifications, and agreement was also typically low between
+We found that tropical cyclones tended to bring different hazards to 
+different counties and also that agreement was typically low between
 distance-based tropical cyclone exposure assessment and each of the single
@@
 
-However, we did find a small set of tropical cyclones for which for which
-agreement was high across single-hazard exposure assessments (e.g., Floyd
-in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996;
-Figure~\ref{fig:jaccard}).  Hurricanes Floyd in~1999 (Figure~S4) and Irene
-in~2011 both made their first \ac{US} landfall in North Carolina at minor
-hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
-eastern coast of the \ac{US} north through New England, bringing substantial
-rainfall to much of the eastern coast from North Carolina north and causing
-extensive inland flooding in North Carolina (Floyd) and New England
-(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
-in~2008 and Bertha in~1996 followed the eastern coastline north of North
-Carolina in a pattern similar to Floyd and Irene. For this set of storms, the
-tropical cyclones' persistent proximity to water may have helped maintain wind
-speeds in similar patterns to rain and distance-based exposure assessments,
-resulting in more similarities across exposure assessments compared to the
-other tropical cyclones we considered.
-
-Since exposure assessments often differ across storm hazards, as well as
-between each storm hazard and a distance-based proxy measurement, exposure
-misclassification is a potential risk. For some studies, such exposure
-misclassification might plausibly be differential (i.e., associated with the
-outcome of interest or with factors associated with risk of the outcome of
-interest).  For example, tropical cyclone wind exposures tend to be
-concentrated in counties near the coast, while tropical cyclone wind exposures
-sometimes extend well inland.  If the etiologically-relevant exposure for a
-health outcome is extreme rainfall but exposure is classified based on
+Therefore for most tropical cyclones, since exposure assessments often differ
+across storm hazards, as well as between each storm hazard and a distance-based
+proxy measurement, exposure misclassification is a potential risk. For some
+studies, such exposure misclassification might plausibly be differential (i.e.,
+associated with the outcome of interest or with factors associated with risk of
+the outcome of interest).  For example, tropical cyclone wind exposures tend to
+be concentrated in counties near the coast, while tropical cyclone wind
+exposures sometimes extend well inland.  If the etiologically-relevant exposure
+for a health outcome is extreme rainfall but exposure is classified based on
 measurements of wind, the probability of being misclassified as unexposed would
@@
 
+However, we did find a small set of tropical cyclones for which for which
+agreement was high across several single-hazard exposure assessments (e.g.,
+Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto in~2006
+Figure~\ref{fig:jaccard}).  Hurricanes Floyd in~1999 (Figure~S4) and Irene
+in~2011 both made their first \ac{US} landfall in North Carolina at minor
+hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
+eastern coast of the \ac{US} north through New England, bringing substantial
+rainfall to much of the eastern coast from North Carolina north and causing
+extensive inland flooding in North Carolina (Floyd) and New England
+(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
+in~2008, Bertha in~1996, and Ernesto in~2006 also followed the eastern
+coastline. For this set of storms, the tropical cyclones' persistent proximity
+to water may have helped maintain wind speeds in similar patterns to rain and
+distance-based exposure assessments, resulting in more similarities across
+exposure assessments compared to the other tropical cyclones we considered.
+For these storms, it may be possible to assess exposure to multiple hazards of
+the storm using a single metric, in some cases even a proxy like the distance
+between a county and the storm's track.  However, for these storms it may be
+difficult to try to untangle the contribution of each hazard to the overall
+effect of the storm, given that several hazards have similar geographical
+patterns.
+
 \subsection*{Case study}
 
-The case study illustrates how differences in exposure assessment
-across storm exposure metrics can result in differing conclusions. In this 
-case, we found differences in estimates of expected
-physical exposure to tropical cyclones among a population when exposure
-assessment was based on different storm exposure metrics
-(Figure~\ref{fig:topelecdependexposure}).
-
-These results are informative because previous assessments of physical exposure
-to tropical cyclones have often concentrated on storm winds and storm
-proximity. For example, a global study of exposure and vulnerability to natural
-hazards used a combination of windspeed above a certain threshold and distance
-from the storm's track to assess exposure~\parencite{peduzzi2009assessing}.
-While many storm impacts might be most strongly linked to wind hazards, there
-is a growing recognition of the potential risks of adverse health outcomes and
-property damage from rain- and flood-related tropical cyclone hazards in the
+The case study illustrates how differences in exposure assessment across storm
+exposure metrics can result in differing conclusions. In this case, we found
+differences in estimates of expected rates of physical exposure to tropical
+cyclones among a population when exposure assessment was based on different
+storm exposure metrics (Figure~\ref{fig:topelecdependexposure}).  These results
+are informative because previous assessments of physical exposure to tropical
+cyclones have often concentrated on storm winds and storm proximity. For
+example, a global study of exposure and vulnerability to natural hazards used a
+combination of windspeed above a certain threshold and distance from the
+storm's track to assess exposure~\parencite{peduzzi2009assessing}.  While many
+storm impacts might be most strongly linked to wind hazards, there is a growing
+recognition of the potential risks of adverse health outcomes and property
+damage from rain- and flood-related tropical cyclone hazards in the
 \ac{US}~\parencite{smith2009}.  Assessments based solely on wind and distance
 from the storm track therefore may underemphasize potentially dangerous
-exposures to certain regions of the \ac{US}, particularly inland locations.
-
-In fact, based on our assessment of physical exposures to tropical cyclones
-among electricity-dependent Medicare beneficiaries, the highest expected
-physical exposure under any metric was measured for Philadelphia County, PA,
-under the flood metric (Figure~\ref{fig:topelecdependexposure}).  While this
-county has a similar-sized population of electricity-dependent Medicare
-beneficiaries as Miami-Dade County, FL
-(Figure~\ref{fig:topelecdependexposure}), the inland county of Philadelphia has
-frequent storm-related flood events, which results in a very high expected
-physical exposure among electricity-dependent Medicare beneficiaries in the
-county. This exposure is poorly captured by most of the other exposure metrics,
-under which Philadelphia County is not included in the top ten counties for
-physical exposure~(Figure~\ref{fig:topelecdependexposure}).
+exposures to certain regions of the \ac{US}, particularly inland locations.  In
+fact, based on our assessment of physical exposures to tropical cyclones among
+electricity-dependent Medicare beneficiaries, the highest expected physical
+exposure under any metric was measured for Philadelphia County, PA, under the
+flood metric (Figure~\ref{fig:topelecdependexposure}).  While this county has a
+similar-sized population of electricity-dependent Medicare beneficiaries as
+Miami-Dade County, FL (Figure~\ref{fig:topelecdependexposure}), the inland
+county of Philadelphia has frequent storm-related flood events, which results
+in a very high expected physical exposure among electricity-dependent Medicare
+beneficiaries in the county. This exposure is poorly captured by most of the
+other exposure metrics, under which Philadelphia County is not included in the
+top ten counties for expected rates of physical
+exposure~(Figure~\ref{fig:topelecdependexposure}).
 

### 2019-10-13 writing/ehp_figures.tex
Work on writing

@@
          exposure metrics for all storms between~1996 and~2011 for which 
-	 at least~250 counties were exposed based on at least one metric.
+	 at least~100 counties were exposed based on at least one metric.
 	 Each row shows one storm, and the color of each cell shows the 
@@
 \includegraphics[width=15.5cm]{topelecdependexposure}
-\caption{Study counties with the highest expected physical exposure per year among
-	 electricity-dependent Medicare beneficiaries for each exposure metric. 
+\caption{Study counties with the highest expected rate of physical exposure per year among
+	 electricity-dependent Medicare beneficiaries based on each exposure metric. 
 	 The color of each bar indicates the number of Medicare beneficiaries in the 
 	 county reliant on electricity-dependent medical and assistive equipment as 
-	 of July~2017. The length of each bar shows the average expected physical exposure
+	 of July~2017. The length of each bar shows the average expected rate of physical exposure
 	 to tropical cyclones among this population based on a given exposure metric, i.e., 
 	 the expected number of these electricity-dependent Medicare beneficiaries exposed 
-	 to tropical storms per year based on that metric.}
+	 to tropical storms per year based on that metric (Table 1).}
 \label{fig:topelecdependexposure}

### 2019-10-13 writing/ehp_introduction.tex
Work on writing

@@
 grineski2019impact, issa2018deaths, tanz2019notes, paul2019brief}), including
-through projects funded through \ac{NIH} Rapid Response grants
-\parencite{nihreporter}. These projects will add to evidence from previous
-hurricanes, including Sandy in 2012 (e.g., \parencite{swerdel2014}) and Katrina
-in 2005 (e.g., \parencite{burton2009health}), characterizing the health impacts
-of hurricanes.
+through projects funded by \ac{NIH} Rapid Response grants
+\parencite{nihreporter}. These projects will add to evidence characterizing how
+hurricanes impact human health, including evidence from multi-year studies
+based on fatality data from post-disaster surveillance
+\parencite{rappaport2000, rappaport2014fatalities, rappaport2016fatalities,
+czajkowski2010fatal, czajkowski2011, moore2012}.
 
-A recent review highlights the need to supplement these single-storm studies
-with multi-year, multi-storm studies, to identify patterns that are common
-across multiple storms [ref]. While multi-year studies of hurriance mortality
-impacts have been conducted using disaster surveillance data, studies based on
-administrative data could identify risks for non-accidental mortality and
-morbidity that those studies likely undercount [ref]. Multi-storm studies with
-administrative data require exposure assessment that is consistent across time
-and space, allowing the administrative data to be linked to disaster exposure
-[ref]. Key to expanding tropical cyclone epidemiology are therefore exposure
-assessment methods that appropriately capture hazards of the storm that are
-relevant to human health, methods that can be applied consistently across
-multiple storms, locations, and years.
+For several types of climate-related disasters---including heat waves, floods,
+and wildfires---research based on surveillance data has been richly
+supplemented by studies analyzing trends in community-wide administrative or
+vital data, including time series or case-crossover studies (e.g.,
+\cite{anderson2010heat, son2012impact, haikerwal2015impact, liu2017wildfire,
+milojevic2017mental}).  These studies typically compare the daily rate of a
+health outcome in a community during disasters to rates during comparable
+non-disaster periods.  While these studies cannot attribute specific cases
+(e.g., a specific person's death) to a disaster, they can help characterize the
+community-wide change in health risk during disasters, and can help capture
+impacts that might be missed or underestimated in post-disaster surveillance.
+In many cases, these studies analyze multi-year, multi-community data, allowing
+them to estimate average associations over many disasters and to explore how a
+disaster's characteristics, or the characteristics of the community it hits,
+modify associated health risk (e.g., \cite{anderson2010heat, son2012impact,
+liu2017wildfire}).  Some studies have begun to use this approach to study the
+health impacts of tropical cyclones, including several studies of Hurricane
+Maria (e.g., \cite{santos2018use, santos2018differential}) and a study of the
+2004 hurricane season in Florida \parencite{mckinney2011}.  However, to expand
+this approach to study longer time periods or a larger set of communities, it
+is critical that researchers be able to assess exposure to tropical cyclones in
+a way that is consistent and comparable across storms, years, and communities.
 
@@
 standard for Atlantic-basin tropical cyclone tracks.  It records the central
-position of a tropical cyclone every six hours, as well as the storm's
-\underline{minimum pressure} and \underline{central maximum sustained winds}. This data is openly
-available through the \ac{HURDAT2}, a post-storm assessment conducted
-by the \ac{US} \ac{NHC} that incorporates data from a variety of sources,
-including satellite data and, when available, aircraft reconnaissance
-data~\parencite{landsea2013, jarvinen1988}. With this data, it is
-straightforward to measure whether a community was in the direct path, or
-within a certain distance of the path, of a storm. Some studies have indeed
-used this data to assess tropical cyclone exposure based on how near the
-storm's central track came to the county, either to assess average exposure
-patterns to tropical storms [refs] or to characterize exposure for health
-studies [refs].
+position of a tropical cyclone every six hours, as well as the storm's minimum
+central pressure and maximum wind speed.  This data is openly available through
+the \ac{HURDAT2}, a post-storm assessment conducted by the \ac{US} \ac{NHC}
+that incorporates data from a variety of sources, including satellite data and,
+when available, aircraft reconnaissance data~\parencite{landsea2013,
+jarvinen1988}. With this data, it is straightforward to measure whether a
+community was in the direct path, or within a certain distance of the path, of
+a storm. Some studies have indeed used this data to assess tropical cyclone
+exposure based on how near the storm's central track came to the county, either
+to assess average exposure patterns for tropical cyclones [refs] or to
+characterize exposure for health studies (e.g., \cite{currie2013}).
 
@@
 Further, when studies use distance from the storm's track to assess exposure,
-they often use an equal buffer distance on each side of the tropical cyclone
-track (e.g.,~\cite{czajkowski2011, grabich2015, grabich2016, zandbergen2009,
-tansel2010}). However, the forces of a tropical cyclone tend to be distributed
-around the center in a non-symmetrical way. Extreme winds are more common to
-the right of the track, where counter-clockwise cyclonic winds move in concert
+they often use the same buffer constraints on each side of the tropical cyclone
+track (e.g.,~\cite{czajkowski2011, zandbergen2009, grabich2015measuring,
+tansel2010, currie2013}). However, the forces of a tropical cyclone tend to be
+distributed around its center non-symmetrically. Extreme winds are more common
+to the track's right, where counter-clockwise cyclonic winds move in concert
 with the tropical cyclone's forward motion~\parencite{halverson2015}, and the
@@
 occurred almost exclusively to the right of the tropical cyclone's track,
-mostly in the right front quadrant of the tropical
-cyclone~\parencite{moore2012}. Rain, conversely, is often heaviest to the left
-of the tropical cyclone's track, especially when the tropical cyclone interacts
-with other weather systems~\parencite{atallah2003, atallah2007,
-zhu2013variations} or undergoes an extratropical
+mostly in the storm's right front quadrant~\parencite{moore2012}. Rain,
+conversely, is often heaviest to the left of the track, especially when the
+tropical cyclone interacts with other weather systems~\parencite{atallah2003,
+atallah2007, zhu2013variations} or undergoes an extratropical
 transition~\parencite{elsberry2002}.
 
-If a study misclassifies exposure to the hazard or hazards that cause the
-health risk being studied, the study will generate biased estimates of tropical
-cyclone risks and impacts. Further, when different studies use different
-methods to assess exposure to tropical cyclones, their results are difficult to
-meaningfully compare and aggregate. If different methods identify
-similar sets of communities as ``exposed'',  these concerns are less serious.
-However, if different methods differ substantially in which communities they
-identify as ``exposed'', it makes it very important that epidemiological
-studies are thoughtful in how they assess exposure.
+The multi-hazard nature of tropical cyclones therefore makes it hard to assess
+exposure based only on a proxy measurement, like how close the storm's central
+track came to the community. While other approaches have been developed to
+incorporate wind hazards into exposure assessment (e.g.,
+\cite{grabich2015measuring, zandbergen2009}), but these do not incorporate other
+storm hazards. To help epidemiological researchers with exposure assessment for
+tropical cyclones, we here develop a dataset of county-level exposure to
+tropical cyclones in all eastern \ac{US} counties, which we have packaged as an
+open-source dataset~\parencite{hurricaneexposure}, as well as an associated
+package of open-source functions that can be used to explore and map this data
+and to link it with county-level health data~\parencite{hurricaneexposuredata}.
+These data cover five metrics characterizing exposure to tropical cyclone
+hazards (Table~\ref{tab:exposuremetrics}): (1)~closest distance the storm's
+central track came to the county's center (a proxy measurement of storm
+exposure used in some previous studies); (2)~maximum sustained wind speed at
+the county's center over the course of the storm; (3)~cumulative rainfall in
+the county over the course of the storm; (4)~flood events in the county
+associated with the storm; and (5)~tornado events in the county associated with
+the storm.  We aggregated these data at the county level since data on many
+potential impacts are available at county-level aggregations (e.g., direct
+hurricane-related deaths~\parencite{czajkowski2011}; birth
+outcomes~\parencite{grabich2015, grabich2016}; autism
+prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
+for, and respond to, tropical cyclones are often undertaken at the county
+level~\parencite{zandbergen2009, rappaport2000}.
 
-To provide open-source data on tropical cyclone exposures to other researchers,
-we developed a dataset of county-level exposure to tropical cyclones in all
-eastern \ac{US} counties for five different metrics
-(Table~\ref{tab:exposuremetrics}): (1)~distance to tropical cyclone track;
-(2)~maximum sustained wind speed; (3)~cumulative rainfall; (4)~flood events;
-and (5)~tornado events~\parencite{hurricaneexposure}. This data is provided at
-the county level since data on many potential impacts are available at
-county-level aggregations (e.g., direct hurricane-related
-deaths~\parencite{czajkowski2011}; birth outcomes~\parencite{grabich2015,
-grabich2016}; autism prevalence~\parencite{kinney2008}) and since decisions and
-policies to prepare for, and respond to, tropical cyclones are often undertaken
-at the county level~\parencite{zandbergen2009, rappaport2000}. Further, we
-explored patterns in these exposure assessments across storm hazards and
-measured how well exposure classification agreed across these five metrics in
-terms of classifying specific counties as exposed to a tropical cyclone.
+In this paper, we describe how we developed this dataset and explore how the
+measurements in it compare to other data that could be used to characterize
+tropical cyclone hazards in the context of epidemiological studies.  Further,
+we explore and present patterns in binary exposure assessments based on these
+measurements, both for each individual metric and in terms of agreement between
+metrics. These analyses can help epidemiologists as they develop study designs
+and plan statistical analysis for multi-year, multi-community studies that
+investigate how community-wide rates of health outcomes change during exposure
+to the hazards brought by tropical cyclones. These analyses can also help
+researchers compare and interpret results from previous epidemiological studies
+that have assessed tropical cyclone exposure in different ways, helping to
+[synthesize] previous epidemiologic results to form a better picture of how
+these storms can affect human health.

### 2019-10-13 writing/ehp_methods.tex
Work on writing

@@
 
-This study covered all counties in the eastern half of the \ac{US}
-(Figure~\ref{fig:hurrtracks}) as of the~2010 \ac{US} Decennial Census. The
-study covered all tracked land-falling or near-land Atlantic-basin tropical
-cyclones between~1988 and~2015 by considering all 136 storms in \ac{HURDAT2}
-\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
-eastern \ac{US} county in this period (Figure~\ref{fig:hurrtracks}).
+We aggregated data for all counties in the eastern half of the \ac{US}
+(Figure~\ref{fig:hurrtracks}) as of the~2010 \ac{US} Decennial Census and for
+all tracked land-falling or near-land Atlantic-basin tropical cyclones
+between~1988 and~2015---all 136 storms in \ac{HURDAT2} \parencite{landsea2013}
+that came within~250 \si{\kilo\metre} of at least one eastern \ac{US} county in
+this period (Figure~\ref{fig:hurrtracks}).
 
@@
 interval, we measured the distance between the storm's center and each county's
-population mean center~\parencite{bivand2013applied, countycenters}. From these
-measurements, we determined how close each storm came to each county.
-
-We also recorded the time the storm came closest to each county. We later used
-this time to link observed data on precipitation, flooding, and tornadoes to
-the storm.  To allow matching with data recorded in local time (e.g., health
+population mean center~\parencite{bivand2013applied, countycenters}. We took
+the minimum for each county as a measure of how close the storm came to the
+county.  We also recorded the time the storm came closest to each county, which
+we later used to link observed data on precipitation, flooding, and tornadoes
+to the storm.  To allow matching with data recorded in local time (e.g., health
 data), we converted these times from \ac{UTC} to local
@@
 \ac{US}~\parencite{rui2013nldas}. We used data that were previously aggregated
-from this re-analysis dataset to the county level for the Centers for Disease
-Control and Prevention's \ac{WONDER} database \parencite{cdcwonder} by two of
-the coauthors (Al-Hamdan and Crosson) \parencite{alhamdan2014environmental,
-cdcwonder}. To create this aggregated data originally, these coauthors took
-hourly precipitation measurements in the \ac{NLDAS-1} precipitation files,
-which are given at~1/8\si{\degree} grid points across the continential
-\ac{US}, and summed them by day for each grid point.  This created a daily
-precipitation total for each grid point; these were then averaged for all grid
-points within a county's~1990 \ac{US} Census
-boundaries~\parencite{alhamdan2014environmental, cdcwonder}. This process
-generated daily county-level precipitation estimates for each continental
-\ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.  We matched these
-county-level daily measurements by date with storm tracks, using for each
-county and storm the date when the storm was closest to the county. Given the
-location of storm-affected counties and the typical timing of tropical
-cyclones, these precipitation measures primarily represent rainfall, although
-occasionally they may represent snowfall or other types of precipitation.
+from this re-analysis dataset to the county level by two of the coauthors
+(Al-Hamdan and Crosson) for the Centers for Disease Control and Prevention's
+\ac{WONDER} database \parencite{cdcwonder, alhamdan2014environmental}. To
+create this aggregated data originally, these coauthors took hourly
+precipitation measurements in the \ac{NLDAS-2} precipitation files, which are
+given at~1/8\si{\degree} grid points across the continential \ac{US}, and
+summed them by day for each grid point.  This created a daily precipitation
+total for each grid point; these were then averaged for all grid points within
+a county's~1990 \ac{US} Census boundaries~\parencite{alhamdan2014environmental,
+cdcwonder}. This process generated daily county-level precipitation estimates
+for each continental \ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.
+We matched these county-level daily measurements by date with storm tracks,
+using for each county and storm the date when the storm was closest to the
+county. Given the location of storm-affected counties and the typical timing of
+tropical cyclones, these precipitation measures primarily represent rainfall,
+although occasionally they may represent snowfall or other types of
+precipitation.
 
-In the open-source data, we provide daily precipitation values for each county
-for the period from five days before to three days after each storm's closest
-approach to the county.  In the software that we published in association with
-this data, we provide the functionality to aggregate these daily values to
-create cumulative precipitation estimates for custom windows as needed. For
-example, a user could determine precipitation only on the day the storm was
-closest to each county, but could also determine the cumulative rainfall for a
-more extended period, like from three days before to three days after the
-storm's closest approach. For later analysis in this study, we calculated
+In the open-source data~\parencite{hurricaneexposuredata}, we provide daily
+precipitation values for each county for the period from five days before to
+three days after each storm's closest approach to the county.  In the software
+that we published in association with this data~\parencite{hurricaneexposure},
+we provide functionality to aggregate these daily values to create cumulative
+precipitation estimates for custom windows around the date of the storm's
+closest approach to the county. For example, a user could determine
+precipitation only on the day the storm was closest to each county, but could
+also determine the cumulative rainfall for a more extended period, like from
+three days before to three days after the storm's closest approach. For the
+analysis of binary exposure classifications in this study, we calculated
 storm-associated rainfall as the sum of precipitation from two days before the
@@
 between storm-specific cumulative precipitation estimates for the two data
-sources (\ac{NLDAS} re-analysis data versus ground-based monitoring) within
+sources (\ac{NLDAS-2} re-analysis data versus ground-based monitoring) within
 each sample county.
@@
 
-To create a dataset of county-level sustained winds during historical tropical
-cyclones, we modeled maximum sustained wind speeds at each county's population
-mean center~\parencite{countycenters}, using a wind speed model based on
-results from Willoughby~\parencite{willoughby2006parametric}. For each storm,
-we modeled the maximum ground-level sustained wind speed that the storm
+To create a dataset of county-level maximum sustained winds during historical
+tropical cyclones, we modeled maximum sustained wind speeds at each county's
+population mean center~\parencite{countycenters} using a wind speed model based
+on results from Willoughby~\parencite{willoughby2006parametric}. For each
+storm, we modeled the maximum ground-level sustained wind speed that the storm
 generated at each county center for each~15-\si{\minute} increment throughout
 the storm tracking period, using a double exponential wind speed model with
-inputs for the storm's forward speed, direction, and central wind
+inputs for the storm's forward speed, direction, and maximum wind
 speed~\parencite{willoughby2006parametric, stormwindmodel}. This model
@@
 this database has recorded storm data since~1950, its coverage changed
-substantially in~1996 to cover a wider variety of storm
+substantially in~1996 to cover a more types of storm
 events~\parencite{stormevents}. We therefore only considered these metrics of
@@
 window. We investigated how these measurements varied between storms with
-associated flood events in the NOAA Storm Events data versus storms without an
+associated flood events listed in the NOAA Storm Events data versus storms without an
 event listing for the county, to explore if storms with flood event listings
-tended to be associated with higher streamflows at gages within the county.
+tended to be associated with higher streamflow at gages within the county.
 
@@
 In our open-source data, we provide continuous measurements of three exposure
-metrics: closest distance of each storm to each county, modeled storm winds,
-and storm-related rainfall. However, epidemiologic studies of tropical cyclones
-often use a binary exposure classification (``exposed'' versus ``unexposed'')
-to assess storm-related health risks [refs], and so we also explored patterns
-in storm exposure based on binary clasifications of these exposure metrics.
+metrics: closest distance of each storm to each county, maximum sustained winds
+at the county center, and cumulative rainfall. However, epidemiologic studies
+of tropical cyclones often use a binary exposure classification (``exposed''
+versus ``unexposed'') to assess storm-related health risks, and so we
+explored patterns in storm exposure based on binary clasifications of
+these exposure metrics.
 
@@
 in our data, since these metrics were based on whether an event was listed in
-the NOAA Storm Events database.  For the other exposure metrics, each county
+the NOAA Storm Events database.  For other exposure metrics, each county
 was classified as exposed to a tropical cyclone based on whether the exposure
@@
 For the rainfall metric, a distance constraint was also necessary, to ensure
-that rains unrelated and far from the storm track were not misattributed to a
+that rainfall unrelated and far from the storm track were not misattributed to a
 storm. Through exploratory analysis, we set this distance metric at~500
@@
 stopped at extratropical transition (e.g., Tropical Storm Lee in 2011)---some
-storm-related rains may be missed because of this distance constraint
+storm-related rain exposures may be missed because of this distance constraint
 (Figure~S1). This distance constraint can be customized using the software
@@
 We characterized patterns in county-level exposure in the eastern \ac{US} for
-each of the measured tropical cyclone hazards (extreme precipitation, maximum
-sustained winds, flooding, and tornadoes).  Depending on available exposure
-data, this assessment included some or all of the period from~1988 to~2015
-(Table 1). For each binary metric of exposure to a tropical cyclone hazard, we
-first summed the total number of county-level exposures over available years
-for each exposure metric and mapped patterns in these exposures. We next
-investigated agreement between storm exposure classifications based on these
-different single-hazard exposures, as well as between each of these and a
-distance-based proxy assessment of tropical cyclone exposure
-(Table~\ref{tab:exposuremetrics}). We calculated the within-storm Jaccard
-index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between each
-pair of exposure metrics. The Jaccard index~($J_s$) measures similarity between
-two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical cyclone~$s$ as the
+exposure assessment based on each of the measured tropical cyclone hazards
+(cumulative precipitation, maximum sustained winds, flooding, and tornadoes).
+Depending on available exposure data, this assessment included some or all of
+the period from~1988 to~2015 (Table 1). For each binary metric of exposure to a
+tropical cyclone hazard, we first summed the total number of county-level
+exposures over available years and mapped patterns in
+these exposures. We next investigated agreement between storm exposure
+classifications based on these different single-hazard exposures, as well as
+between each of these and a distance-based proxy assessment of tropical cyclone
+exposure (Table~\ref{tab:exposuremetrics}). We calculated the within-storm
+Jaccard index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between
+each pair of exposure metrics. The Jaccard index~($J_s$) measures similarity
+between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical cyclone~$s$ as the
 proportion of counties in which both of the metrics classify the county as
@@
 tropical cyclone. We measured these values for all study storms that affected a
-large number of eastern \ac{US} counties (those storms with 250 or more of the
-study counties classified as exposed by at least one of the exposure metrics)
+large number of eastern \ac{US} counties (those storms with 100 or more
+counties were classified as exposed by at least one of the exposure metrics)
 during the years when all exposure data were available (1996\,--\,2011).
@@
 in each study county from the \ac{US} Department of Health \& Human Service's
-emPower Map~2.0~\parencite{empower}. This subpopulation was selected since it
+emPower Map~2.0~\parencite{empower}. This subpopulation was selected because it
 may be particularly susceptible to health impacts from tropical cyclone

### 2019-10-13 writing/ehp_results.tex
Work on writing

@@
 
-We created and published this tropical cyclone exposure data as an open-source
-dataset, so they can be used by epidemiological
+We created and published this tropical cyclone exposure data as open-source
+data, so they can be used by epidemiological
 researchers~\parencite{hurricaneexposuredata}. These data include continuous
@@
 listings but for which the total discharge across county streamflow gages was
-below the 25\textsuperscript{th} percentile of measurements during storms
-without a flood event listing in that county. For storms without a flood
-listing for the county, in most cases the total streamflow discharge in
-the county was lower than during storms for which there were flood
-event listings. Further, in all but one of these cases, all streamgage flows
-were below the flooding threshold. The exception was one storm in Fulton
-County, GA (Hurricane Ida in 2009).
+relatively low. For storms without a flood listing for the county, in most
+cases the total streamflow discharge in the county was relatively low, and, in
+all but one of these cases all streamgage flows were below the flooding
+threshold. The exception was for one storm in Fulton County, GA (Hurricane Ida
+in 2009).
 
@@
 to four different hazards for each tropical cyclone and then explored patterns
-over years with available data. Across the four storm hazards considered,
-there was wide variation in the average number of county exposures per year
+over years with available data. Across the four storm hazards considered, there
+was wide variation in the average number of county exposures per year
 (Table~\ref{tab:exposuresummaries}). For the tropical cyclone tornado exposure,
-there were on average about 50 county exposures per year within our study. For
-tropical cyclone flood and wind exposures, county exposures were much more
-frequent ($>$200 / yr on average), and even more when assessing tropical
-cyclone rain exposure ($>$400 / yr on average).
+there were on average about 50 county exposures per year within our study.
+County exposures were more frequent for tropical cyclone wind exposures ($>$140
+/ yr on average), even more frequent for tropical cyclone flood exposures
+($>$200 / yr on average), and most frequent for tropical cyclone rain exposure
+($>$400 / yr on average).
 
 We also measured the average number of counties exposed per storm to each
-hazard (limiting this analysis to the tropical cyclones that exposed at least
-one county to that hazard; Table~\ref{tab:exposuresummaries}).  Tropical
-cyclones typically exposed only a few counties, if any, to tornadoes (median
-counties exposed per storm:~7), more counties to flooding and high winds
-(median:~26 and~34, respectively), and even more counties to extreme rain
-(median:~68).  For every hazard except tropical cyclone tornadoes, we
-identified at least one tropical cyclone that exposed over~300 counties.
-However, the largest-extent tropical cyclone varied across hazards: Frances
-in~2004 exposed the most counties based on rain, Ike in~2008 based on wind, and
-Ivan in 2004 based on flooding and tornadoes
-(Table~\ref{tab:exposuresummaries}).
+hazard (limiting this analysis to tropical cyclones that exposed at least one
+county to that hazard; Table~\ref{tab:exposuresummaries}).  If a tropical
+cyclone exposed any counties to tornadoes, it was typically only a few (median
+counties exposed per storm:~7). When a storm exposed any counties to other
+tropical cyclone exposures, the swath of exposed counties tended to be much
+larger: a median of~25 counties exposed for high winds per storm,~26 for
+flooding, and~68 for extreme rainfall.  For every hazard except tropical
+cyclone tornadoes, we identified at least one tropical cyclone that exposed
+over~250 counties (Table~\ref{tab:exposuresummaries}).  However, the
+largest-extent tropical cyclone varied across hazards: Frances in~2004 exposed
+the most counties based on rain, Ike in~2008 based on wind, and Ivan in 2004
+based on flooding and tornadoes (Table~\ref{tab:exposuresummaries}).
 
@@
 exposure was also more frequent in coastal areas compared to inland areas,
-there were also inland exposures in counties not often  that were not
-identified as exposed by the wind-based metric. Rain and, to some extent, flood
-exposures were characterized by a pattern defined by the Appalachian Mountains,
-with many fewer exposures west of the mountain range than to the east. Almost
-all tornado-based exposures were in coastal states, with most in Florida,
-Alabama, South Carolina, and North Carolina and almost none north of Maryland.
-Flood-based exposures were highest in North Carolina and along the South
-Carolina coast, as well as further north in New Jersey, southeastern
-Pennsylvania, and southeastern New York. Notably, some areas of the \ac{US}
-that experienced regular exposures to tropical cyclone-related excessive
-rainfall and floods were not identified as areas of regular risk based on the
-more commonly used tropical cyclone exposure assessment metrics of wind or
-distance (Figure~\ref{fig:averageexposure}).  Patterns were similar when
-analysis was restricted to years with exposure data for all four hazards
+there were also inland exposures in counties that were never to extreme storm
+winds. Rain and, to some extent, flood exposures were characterized by a
+pattern defined by the Appalachian Mountains, with fewer exposures west of the
+mountain range than to the east. Almost all tornado-based exposures were in
+coastal states, with many in Florida, and almost none north of Maryland.
+Flood-based exposures were frequently in the Mid-Atlantic region, with a
+pattern that skewed north compared with other exposures.  Patterns were similar
+when analysis was restricted to years with exposure data for all four hazards
 available (1996\,--\,2011; Figure~S3).
@@
 Next, we measured the agreement between exposure classifications for each pair
-of metrics. We found that these exposure classifications typically did not
-strongly agree between pairs of metrics. The set of counties
-identified as ``exposed'' based on one metric often overlap little with
-the set identified as ``exposed'' by another metric.
+of hazards, as well as between each hazard and exposure assessment based on
+distance from the storm's track as a proxy for exposure to storm hazards. We
+found that these exposure classifications typically did not strongly agree
+between pairs of metrics. The set of counties identified as ``exposed'' based
+on one metric often overlap little with the set identified as ``exposed'' by
+another metric, as storms frequently brought different hazards to different 
+locations.
 
 Figure~\ref{fig:ivanexposure} shows shows an example: Hurricane Ivan in 2004.
-For the distance-based metric, the counties assessed as exposed followed the
-tropical cyclone's track, including counties well inland. For the wind metric,
-only counties near two of the tropical cyclone's landfalls were assessed as
+For the distance-based metric, the counties assessed as exposed follow the
+tropical cyclone's track. For the wind-based metric,
+only counties near the tropical cyclone's first landfall were assessed as
 exposed. For rain- and flood-based metrics, however, exposure extended to the
@@
 cyclones (selected because they exposed many \ac{US}  counties based on at
-least one metric), and Figure~S5 provides four examples of tropical cyclones in
-which the disagreement in exposure assessment between rain- and wind-based
-metrics was particularly notable.
+least one metric).
 
@@
 between~1996 and~2011 (the period for which data for all five metrics were
-available) for which~250 or more counties were exposed based on at least one
+available) for which~100 or more counties were exposed based on at least one
 metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
 for one tropical cyclone, and each box in that row shows the measured Jaccard
-coefficient a pair of exposure metrics. For most tropical cyclones,
+coefficient for a pair of metrics. For most tropical cyclones,
 distance-based county-level exposure assessment was, at best, in moderate
@@
 fewer were assessed the same under both metrics in the pair). For many tropical
-cyclones, the Jaccard index values were~$\le0.2$, indicating poor agreement.
+cyclones, the Jaccard index values were~$\le0.4$, indicating poor agreement.
 Among the four hazard-based metrics, the tornado-based metric had
 universally poor agreement with other metrics in county-level classification
-across the tropical cyclones considered.  For other pairs of metrics, there was
-also generally disagreement in which counties were determined to be exposed,
-with Jaccard index values~$<0.4$ for most metric pairs in most tropical
-cyclones.
+across the tropical cyclones considered.
 
@@
 in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
-agreed moderately to well (Jaccard index of~$\sim 0.5\mbox{\,--\,}0.8$) for all
-pairs of exposure metrics except those including the tornado-based metric. For
-another set of tropical cyclones (e.g., Lee in~2011 [Figure~S3], Ernesto
-in~2006, and Bertha in~1996), there was moderate to good agreement for all
-pairwise combinations of distance, rain, and wind, but poor agreement for all
-other combinations of metrics.  However, these two sets of tropical cyclones
-represented the minority of tropical cyclones considered; for most tropical
-cyclones, there was low overlap in the counties determined to be exposed to the
-tropical cyclones between most pairings of metrics.
+agreed moderately to well for all pairs of exposure metrics except those
+including the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
+of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
+in~2003), there was moderate to good agreement for all pairwise combinations of
+distance, rain, and wind, but poor agreement for all other combinations of
+metrics, while for another set of storms (e.g., Matthew in~2004 and Katrina
+in~2005), there was moderate to good agreement between distance and rain.
+However, these sets of tropical cyclones represented the minority of tropical
+cyclones considered; for most tropical cyclones, there was low overlap in the
+counties determined to be exposed to the tropical cyclones between most
+pairings of metrics.
 
@@
 Figure~\ref{fig:topelecdependexposure} shows the study counties with the
-highest expected physical exposures to tropical cyclones among
+highest expected rates of physical exposure to tropical cyclones among
 electricity-dependent Medicare beneficiaries based on each metric. A few

### 2019-10-13 R/average_elec_depend_exposure_figure.R

@@ -133,7 +133,7 @@ out2 <- top_exposures %>%
         geom_col() +
         coord_flip() +
         labs(x = "", y = "Average # of electricity-dependent Medicare beneficiaries exposed per year",
-             fill = "# of electricity-dependent\nMedicare beneficiaries\nin the county") +
+             fill = "# of electricity-dependent\nMedicare beneficiaries\nin the county (2017)") +
         scale_fill_viridis(breaks = c(6000, 9000, 12000),
                            labels = c("6,000", "9,000", "12,000")) +
         facet_wrap(~ metric, scales = "free_y", drop = TRUE) +

### 2019-10-13 R/other_top_storm_exposures.R

@@ -74,8 +74,8 @@ storm_data <- distance_data %>%
 
 out_data <- hurricaneexposure:::get_eastern_map() %>%
         dplyr::left_join(storm_data, by = "fips")
-out_data <- out_data %>% 
-        dplyr::filter(metric != "Distance-based metric")
+# out_data <- out_data %>% 
+#         dplyr::filter(metric != "Distance-based metric")
 
 out <- ggplot2::ggplot() +
         ggplot2::geom_polygon(data = out_data,
@@ -104,7 +104,7 @@ out <- ggplot2::ggplot() +
         # ggplot2::scale_fill_manual(name = "",
         #                            values = c("white", "dodgerblue"),
         #                            labels = c("Unexposed", "Exposed")) +
-        facet_wrap(~ metric, ncol = 4) +
+        facet_wrap(~ metric, ncol = 5) +
         theme(legend.position="none")
 
 floyd <- map_tracks(storm, plot_object = out) +
@@ -180,8 +180,8 @@ storm_data <- distance_data %>%
 out_data <- hurricaneexposure:::get_eastern_map() %>%
         dplyr::left_join(storm_data, by = "fips")
 
-out_data <- out_data %>% 
-        dplyr::filter(metric != "Distance-based metric")
+# out_data <- out_data %>% 
+#         dplyr::filter(metric != "Distance-based metric")
 
 out <- ggplot2::ggplot() +
         ggplot2::geom_polygon(data = out_data,
@@ -210,7 +210,7 @@ out <- ggplot2::ggplot() +
         # ggplot2::scale_fill_manual(name = "",
         #                            values = c("white", "dodgerblue"),
         #                            labels = c("Unexposed", "Exposed")) +
-        facet_wrap(~ metric, ncol = 4) +
+        facet_wrap(~ metric, ncol = 5) +
         theme(legend.position="none")
 
 ike <- map_tracks(storm, plot_object = out) +
@@ -288,8 +288,8 @@ storm_data <- distance_data %>%
 out_data <- hurricaneexposure:::get_eastern_map() %>%
         dplyr::left_join(storm_data, by = "fips")
 
-out_data <- out_data %>% 
-        dplyr::filter(metric != "Distance-based metric")
+# out_data <- out_data %>% 
+#         dplyr::filter(metric != "Distance-based metric")
 
 out <- ggplot2::ggplot() +
         ggplot2::geom_polygon(data = out_data,
@@ -318,7 +318,7 @@ out <- ggplot2::ggplot() +
[…]

### 2019-11-07 writing/darlings_intro.tex
Make more edits

@@
 terms of classifying specific counties as exposed to a tropical cyclone.
+
+, either to assess average
+exposure patterns for tropical cyclones (e.g., \cite{zandbergen2009}) or
+
+These analyses can also help
+researchers compare and interpret results from previous epidemiological studies
+that have assessed tropical cyclone exposure in different ways, helping to
+[synthesize] previous epidemiologic results to form a better picture of how
+these storms can affect human health.
+
+Tropical cyclones vary
+dramatically in size: \ac{US} tropical cyclones have been observed with radii
+to maximum winds as small as~20~\si{\kilo\metre} and as large
+as~200~\si{\kilo\metre}~\parencite{mallin2006, quiring2011variations}.

### 2019-11-07 writing/darlings_results.tex
Make more edits

@@
 in~2005, and Katrina in~2005 (Figure~S4).
+
+For most tropical cyclones,
+distance-based county-level exposure assessment was, at best, in moderate
+agreement with exposure assessments based on the four hazard-based metrics.
+
+\subsection*{Case study}
+
+We conducted a case study comparing estimates of physical exposure to tropical
+cyclones based on each exposure metric.  Figure~\ref{fig:topelecdependexposure}
+shows the study counties with the highest expected rates of physical exposure
+to tropical cyclones among electricity-dependent Medicare beneficiaries based
+on each metric. A few counties (Miami-Dade County, FL; Harris County, TX) were
+ranked in the top three in expected yearly rate of physical exposure under
+almost all exposure metrics (Figure~\ref{fig:topelecdependexposure}). However,
+there were key differences across exposure metrics. For example, highly-ranked
+based on wind- and tornado-based metrics were in Florida, while over half of
+the highly-ranked counties based on the flood-based metric were in Mid-Atlantic
+states (PA, NY).

### 2019-11-07 writing/ehp_discussion.tex
Make more edits

@@
 Epidemiological studies can help characterize which health risks are elevated
-during disasters, to what degree, and to whom [Ibrahim editorial; Noji paper].
-\parencite{ibrahim2005unfortunate, noji2005disasters}.  As a result, these studies help improve
-disaster preparedness and response \parencite{noji2005disasters}.  However,
-tropical cyclones are multi-hazard events, making it complicated to measure
-exposure.  Here, we provide open-source county-level data on multiple tropical
+during disasters, to what degree, and to
+whom~\parencite{ibrahim2005unfortunate, noji2005disasters}.  As a result, these
+studies help improve disaster preparedness and
+response~\parencite{noji2005disasters}.  However, tropical cyclones are
+multi-hazard events, making it complicated to measure exposure and so to
+conduct multi-community, multi-year studies leveraging large administrative
+datasets.  Here, we provide open-source county-level data on multiple tropical
 cyclone exposure metrics and explore limitations in that data.
 
-Further, we find that county-level tropical cyclone exposure assessments vary
-substantially when using different metrics, based either on storm hazards or on
-how close the storm's central track came to the community. Our results can
-inform exposure assessment for future county-level studies of the health risks
-and impacts associated with tropical cyclones exposures, and the open-source
-data and software we have created provides multi-hazard, county-level exposure
-data for such studies.
+Further, we explore patterns in storm exposure classifications based on
+different metrics, and we find that county-level tropical cyclone exposure
+assessments vary substantially when using different metrics, based either on
+storm hazards or on how close the storm's central track came to the community.
+Our results can inform exposure assessment for future county-level studies of
+the health risks and impacts associated with tropical cyclones exposures as
+well as insights to help in designing studies.  The open-source data and
+software we have created provides multi-hazard, county-level exposure data for
+such studies.
 
@@
 
-We investigated how well the open-source data correspond with
-data from other available sources
+We investigated how well the open-source data correspond with data from other
+potential sources
 (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison} and~S2). While
-generally in agreement, there are a few caveats that should be considered when
-using these data. The rainfall data are generally
-well-correlated with ground-based observations, but may sometimes
-oversmooth very high rainfall values compared to ground-based observations
+generally in agreement, there are some caveats. The rainfall data are generally
+well-correlated with ground-based observations, but may sometimes oversmooth
+very high rainfall values compared to ground-based observations
 (Figure~\ref{fig:raincomparison}). When rainfall data is used to create binary
@@
 
-For the flooding data, we found that flood event status determined based
-on the NOAA Storm Events listings often agreed with measurements from \ac{USGS}
+For the flooding data, we found that flood event status determined based on the
+NOAA Storm Events listings typically agreed with measurements from \ac{USGS}
 streamgages, with a flood event more likely to be listed if a storm elevated
-streamflow across the county (Figure~\ref{fig:floodcomparison}).  However, there
-are differences between the two flooding datasets, and these highlight both the
-difficulty of measuring flood exposure at the county level and inherent
-challenges in using data from a storm event database for epidemiologic exposure
-assessment.
-
-For example, there was one storm in Fulton County, GA, for which there was high
-streamflow but not an associated flood event listing (Hurricane Ida, 2009).
-This storm occurred in November 2009, following a month with historic rainfall
-and flooding in Georgia \parencite{shepherd2011overview}.  Therefore, in this
-case the flooding associated with Ida was incorporated into an ongoing flood
-event listing, with a start date well before the five-day window we used to
-temporally match storm event listings with tropical cyclone tracks.
+streamflow across the county (Figure~\ref{fig:floodcomparison}).  However,
+there are differences between the two flooding datasets, and these highlight
+both the difficulty of measuring flood exposure at the county level and
+inherent challenges in using data from a storm event database for epidemiologic
+exposure assessment.  For example, there was one storm in Fulton County, GA,
+for which there was high streamflow but not an associated flood event listing
+(Hurricane Ida, 2009).  This storm occurred in November 2009, following a month
+with historic rainfall and flooding in
+Georgia~\parencite{shepherd2011overview}.  In this case the flooding
+associated with Ida was incorporated into an ongoing flood event listing, with
+a start date well before the five-day window we used to temporally match storm
+event listings with tropical cyclone tracks.
 
@@
 underreporting~\parencite{Ashley2008flood, Curran2000}, especially in
-sparsely-populated areas~\parencite{Witt1998, Ashley2007} [Simmons and Sutter
-ref], as well as to other reporting errors.
+sparsely-populated areas~\parencite{Witt1998, Ashley2007}, as well as to other
+reporting errors.
 
@@
 measurements for single hazards, but also to create tropical cyclone exposure
-profiles based on multiple hazards or craft exposure indices that combine
+profiles based on multiple hazards or to craft exposure indices that combine
 hazard metrics~\parencite{chakraborty2005population, peduzzi2009assessing}.
-This ability can be critical, as different hazards of tropical cyclones often
+This ability can be critical, as different hazards of tropical cyclones can
 act synergistically in causing impacts~\parencite{smith2009}.
@@
 
-We found differences in geographic patterns in average exposures to different
-tropical cyclone hazards (Figure~\ref{fig:averageexposure}). These patterns
-were not unexpected based on what is known about tropical cyclone hazards, but
-still highlight variations that are critical to consider in designing studies
-and statistical analysis for tropical cyclone epidemiology. Further, they
+We found average exposures to different tropical cyclone hazards differed
+geographically (Figure~\ref{fig:averageexposure}). These patterns were not
+unexpected based on what is known about tropical cyclone hazards, but still
+highlight variations that are critical to consider in designing studies and
+statistical analysis for tropical cyclone epidemiology. Further, they
 demonstrate the need for multi-hazard exposure datasets for tropical cyclone
-epidemiology.
+epidemiology, especially in capturing inland risks.
 
@@
 flooding~\parencite{rees2001}.  Almost all tropical cyclone tornado exposures
-were in southern coastal states, consistent with previous evidence that most noteworthy
-tropical cyclone-related tornadoes occur on the right-side of the tropical
-cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.  The
-exposure averages we calculated are somewhat limited as estimates of long-term
-frequencies, as tropical cyclones follow decadal
+were in southern coastal states, consistent with previous evidence that most
+noteworthy tropical cyclone-related tornadoes occur on the right-side of the
+tropical cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.
+The exposure averages we calculated are somewhat limited as estimates of
+long-term frequencies, as tropical cyclones follow decadal
 patterns~\parencite{kossin2007more} likely not adequately captured in the
@@
 spanning over a century have found similar patterns for distance- and
-wind-based metrics~\textcite{zandbergen2009, kruk2010} [double-check that refs
-found similar patterns].
+wind-based metrics~\textcite{zandbergen2009, kruk2010} \textbf{[double-check
+that refs found similar patterns]}.
 
@@
 
-We found that tropical cyclones tended to bring different hazards to 
-different counties and also that agreement was typically low between
-distance-based tropical cyclone exposure assessment and each of the single
-hazard-specific exposures (Figures~\ref{fig:ivanexposure}\,--\,
-\ref{fig:jaccard} and S4\,--\,S5). This suggests that the use of a
-distance-based metric to assess exposure to any of these hazards, or the use of
-measurements from one hazard as a proxy for exposure to any of the other
-hazards considered, would often be problematic, introducing exposure
-misclassification.
-
-These findings align with previous results from atmospheric science and related
+We found that tropical cyclones tended to bring different hazards to different
+counties and also that agreement was typically low between distance-based
+tropical cyclone exposure assessment and each of the single hazard-specific
+exposures, as well as between pairs of hazard-specific metrics
+(Figures~\ref{fig:ivanexposure}\,--\, \ref{fig:jaccard} and S4\,--\,S5).  These
+findings align with previous results from atmospheric science and related
 fields on the characteristics of tropical cyclones. While tropical cyclone
-rainfall and windspeed can be well-correlated when the tropical cyclone is
-over water~\parencite{cerveny2000}, this relationship often weakens once the
+rainfall and windspeed can be well-correlated when the tropical cyclone is over
+water~\parencite{cerveny2000}, this relationship often weakens once the
 hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
@@
 
-Therefore for most tropical cyclones, since exposure assessments often differ
-across storm hazards, as well as between each storm hazard and a distance-based
-proxy measurement, exposure misclassification is a potential risk. For some
-studies, such exposure misclassification might plausibly be differential (i.e.,
-associated with the outcome of interest or with factors associated with risk of
-the outcome of interest).  For example, tropical cyclone wind exposures tend to
-be concentrated in counties near the coast, while tropical cyclone wind
-exposures sometimes extend well inland.  If the etiologically-relevant exposure
-for a health outcome is extreme rainfall but exposure is classified based on
-measurements of wind, the probability of being misclassified as unexposed would
-be higher in inland counties, while the probability of being misclassified as
-exposed would be higher in coastal counties. If coastal counties differ from
-inland counties in either the outcome of interest or in factors associated with
-risk of that outcome, exposure misclassification would be
-differential~\parencite{savitz2016interpreting}.  Such differential exposure
-misclassification could bias estimates of tropical cyclone effects either
-towards the null (estimating a lower or null association compared to the true
-association that exists) or away from the null (estimating a larger association
-than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.
-[Variation across studies in how exposure is assessed?]
+This suggests that the use of a distance-based metric to assess exposure to any
+of these hazards, or the use of measurements from one hazard as a proxy for
+exposure to any of the other hazards considered, would often be problematic,
+introducing exposure misclassification. \textbf{[Compare with Grabich results?]}
+For some studies, such exposure misclassification might plausibly be
+differential (i.e., associated with the outcome of interest or with factors
+associated with risk of the outcome of interest).  For example, tropical
+cyclone wind exposures tend to be concentrated in counties near the coast,
+while tropical cyclone wind exposures sometimes extend well inland.  If the
+etiologically-relevant exposure for a health outcome is extreme rainfall but
+exposure is classified based on measurements of wind, the probability of being
+misclassified as unexposed would be higher in inland counties, while the
+probability of being misclassified as exposed would be higher in coastal
+counties. If coastal counties differ from inland counties in either the outcome
+of interest or in factors associated with risk of that outcome, exposure
+misclassification would be differential~\parencite{savitz2016interpreting}.
+Such differential exposure misclassification could bias estimates of tropical
+cyclone effects either towards the null (estimating a lower or null association
+compared to the true association that exists) or away from the null (estimating
+a larger association than actually exists)~\parencite{savitz2016interpreting,
+armstrong1998effect}.
 
-However, we did find a small set of tropical cyclones for which for which
-agreement was high across several single-hazard exposure assessments (e.g.,
-Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto in~2006
-Figure~\ref{fig:jaccard}).  Hurricanes Floyd in~1999 (Figure~S4) and Irene
-in~2011 both made their first \ac{US} landfall in North Carolina at minor
+We did, however, find a small set of tropical cyclones for which for
+which agreement was high across several single-hazard exposure assessments
+(e.g., Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto
+in~2006 Figure~\ref{fig:jaccard}).  Hurricanes Floyd in~1999 (Figure~S4) and
+Irene in~2011 both made their first \ac{US} landfall in North Carolina at minor
 hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
@@
 in~2008, Bertha in~1996, and Ernesto in~2006 also followed the eastern
-coastline. For this set of storms, the tropical cyclones' persistent proximity
+coastline. For these storms, the tropical cyclones' persistent proximity
 to water may have helped maintain wind speeds in similar patterns to rain and
@@
 
-\subsection*{Case study}
+\subsection*{Limitations}
 
-The case study illustrates how differences in exposure assessment across storm
-exposure metrics can result in differing conclusions. In this case, we found
-differences in estimates of expected rates of physical exposure to tropical
-cyclones among a population when exposure assessment was based on different
-storm exposure metrics (Figure~\ref{fig:topelecdependexposure}).  These results
-are informative because previous assessments of physical exposure to tropical
-cyclones have often concentrated on storm winds and storm proximity. For
-example, a global study of exposure and vulnerability to natural hazards used a
-combination of windspeed above a certain threshold and distance from the
-storm's track to assess exposure~\parencite{peduzzi2009assessing}.  While many
-storm impacts might be most strongly linked to wind hazards, there is a growing
-recognition of the potential risks of adverse health outcomes and property
-damage from rain- and flood-related tropical cyclone hazards in the
-\ac{US}~\parencite{smith2009}.  Assessments based solely on wind and distance
-from the storm track therefore may underemphasize potentially dangerous
-exposures to certain regions of the \ac{US}, particularly inland locations.  In
-fact, based on our assessment of physical exposures to tropical cyclones among
-electricity-dependent Medicare beneficiaries, the highest expected physical
-exposure under any metric was measured for Philadelphia County, PA, under the
-flood metric (Figure~\ref{fig:topelecdependexposure}).  While this county has a
-similar-sized population of electricity-dependent Medicare beneficiaries as
-Miami-Dade County, FL (Figure~\ref{fig:topelecdependexposure}), the inland
-county of Philadelphia has frequent storm-related flood events, which results
-in a very high expected physical exposure among electricity-dependent Medicare
-beneficiaries in the county. This exposure is poorly captured by most of the
-other exposure metrics, under which Philadelphia County is not included in the
-top ten counties for expected rates of physical
-exposure~(Figure~\ref{fig:topelecdependexposure}).
+The dataset presented here does have several limitations, in addition to the
+caveats previously discussed. First, the dataset is not comprehensive of all
+possible tropical cyclone hazards. For example, coastal counties can experience
+dangerous storm surge, which is not specifically covered in this dataset
+(although some resulting flooding would be captured). We are exploring ways to
+include this in future versions of the dataset, although to date we have
+focused on exposures that could affect any county, whether inland or coastal.
+Second, this data is aggregated to the county level. Many health outcome
+datasets are aggregated at this level, but some may be aggregated at a finer
+spatial resolution (e.g., Census tract or ZIP code). We have published the wind
+model we used to create this dataset as its own open-source R package
+\parencite{stormwindmodel}, and this model can be used to model
+storm-associated winds at a finer spatial resolution; however, measurements of
+other hazards cannot similarly be re-scaled through tools we provide. Finally,
+we provide these data and associated software tools through R packages, and so
+some experience in the R programming language is required to make full use of
+them. However, R is currently a popular programming language for environmental
+epidemiology, allowing the data to reach a large audience, and we are exploring
+options to create a web application using the Shiny platform to allow more
+user-friendly web-based access of the data \parencite{shiny2019}.
 
@@
 
-Previous research has highlighted the range of impacts that tropical cyclones
-can have in \ac{US} communities. However, studies of tropical cyclone exposures
-and their impacts have varied widely in how they assess exposure, and there was
-not previously a widely available public dataset that captured county-level
-exposure to multiple storm hazards. Here we provide open-source tropical
-cyclone hazard-related measurements for a large collection of historical
-Atlantic-basin cyclones, aggregated at the county level to align with the
-spatial scale at which much health-related data is collected.
+To conduct epidemiological studies that span multiple communities and tropical 
+cyclones, it is critical to have consistent and comprehensive measurements of 
+community-level exposure to hazards of the storm. Here we have created and 
+shared a dataset that provides these data for counties in the United States
+over multiple years. Despite some limitations in this dataset, it provides 
+a powerful tool for expanding tropical cyclone epidemiology studies to more
+extensively leverage existing administrative and vital data, allowing researchers
+to improve our understanding of how these storms affect community-wide risk of 
+numerous health outcomes, including outcomes that are also common outside of storm 
+periods and so may be hard to attribute on a case-by-case basis to a storm. 
+Further, this dataset provides hazard measurements that are comparable across
+communities and storms, which allows epidemiological researchers to design studies
+to explore how health risks are modified by characteristics of both the storms and
+the communities that are hit. The data are given in an open-source format, along 
+with associated software tools, which allows them to be freely used and for others
+to explore all associated code and to contribute additions through platforms like 
+GitHub.
 
-Further, we investigated agreement in exposure assessment across these metrics
-when applied individually. We found large differences in which counties are
-exposed to different hazards of tropical cyclones and that distance is, at
-best, a moderate, and often a very poor, surrogate for exposure to the specific
-tropical cyclone hazards of high wind, extreme rainfall, flooding, and
-tornadoes. Use of distance as a surrogate for any of these hazards, or exposure
-assessment based on one hazard when the pathway for health impacts is in part
-or full through another storm hazard, could lead to exposure misclassification.
-In the case of tropical cyclone risk and impact studies, including
-epidemiological studies, this would result in biased estimates.  Our findings
-highlight the importance of clarifying the potential pathway from tropical
-cyclone hazards to health impacts when conducting tropical cyclone
-epidemiological studies, and then basing exposure assessment on these hazards.
+Based on our analysis in this paper, these data sources are typically in agreement
+with other sources of data available to characterize storm-associated hazards
+(e.g., ground-based monitors, streamgages, post-storm wind radii estimates). 
+However, researchers who are planning to use the data should explore the analyses
+presented in this paper to understand the strengths and weaknesses of the data. 
+Our analyses also indicate that a storm often brings different combinations of 
+hazards to different counties along or near its path. Our results in comparing 
+storm hazards suggests that county-level storm exposure is not well-characterized 
+by the closest distance that a storm's central track came to a county, and that
+exposure to one storm hazard within a county (e.g., severe winds) does not imply 
+exposure to other hazards (e.g., excessive rainfall, flooding). As a result, it is 
+critical that researchers consider which storm hazards are likely on the causal 
+pathway for the outcomes they are studying, and to characterize storm exposure
+in a way that captures those specific hazards.

### 2019-11-07 writing/ehp_introduction.tex
Make more edits

--- 2019-10-09 writing/ehp_introduction.tex
@@
 Recently, several major tropical cyclones have hit \ac{US} cities, including
-Hurricanes Harvey, Irma, and Maria in 2017 \parencite{blake20182017} and
+Hurricanes Harvey, Irma, and Maria in 2017~\parencite{blake20182017} and
 Hurricanes Florence and Michael in 2018 \parencite{avila20192018}. Researchers
-are exploring how these storms impacted human health (e.g.,
-\parencite{santos2018use, rivera2018estimating, santos2018differential,
+are exploring how these storms impacted human health
+(e.g.,~\cite{santos2018use, rivera2018estimating, santos2018differential,
 grineski2019impact, issa2018deaths, tanz2019notes, paul2019brief}), including
-through projects funded through \ac{NIH} Rapid Response grants
-\parencite{nihreporter}. These projects will add to evidence from previous
-hurricanes, including Sandy in 2012 (e.g., \parencite{swerdel2014}) and Katrina
-in 2005 (e.g., \parencite{burton2009health}), characterizing the health impacts
-of hurricanes.
+through projects funded by \ac{NIH} Rapid Response
+grants~\parencite{nihreporter}. These projects will add to evidence
+characterizing how hurricanes impact human health, including evidence from
+multi-year studies based on fatality data from post-disaster
+surveillance~\parencite{rappaport2000, rappaport2014fatalities,
+rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.
 
-A recent review highlights the need to supplement these single-storm studies
-with multi-year, multi-storm studies, to identify patterns that are common
-across multiple storms [ref]. While multi-year studies of hurriance mortality
-impacts have been conducted using disaster surveillance data, studies based on
-administrative data could identify risks for non-accidental mortality and
-morbidity that those studies likely undercount [ref]. Multi-storm studies with
-administrative data require exposure assessment that is consistent across time
-and space, allowing the administrative data to be linked to disaster exposure
-[ref]. Key to expanding tropical cyclone epidemiology are therefore exposure
-assessment methods that appropriately capture hazards of the storm that are
-relevant to human health, methods that can be applied consistently across
-multiple storms, locations, and years.
+For several types of climate-related disasters---including heat waves, floods,
+and wildfires---research based on post-disaster surveillance data has been
+richly supplemented by studies analyzing trends in community-wide
+administrative or vital data, including time series or case-crossover studies
+(e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
+liu2017wildfire, milojevic2017mental}).  These studies typically compare the
+rate of a health outcome in a community during a disaster or disasters to rates
+during comparable non-disaster periods.  While these studies cannot attribute
+specific cases (e.g., a specific person's death) to a disaster, they can help
+characterize the community-wide change in health risk during disasters, and can
+help capture impacts that might be missed or underestimated in post-disaster
+surveillance.  In many cases, these studies analyze multi-year, multi-community
+data, allowing them to estimate average associations over many disasters and to
+explore how a disaster's characteristics, or the characteristics of the
+community it hits, modify associated health risk (e.g.,~\cite{anderson2010heat,
+son2012impact, liu2017wildfire}).  Some studies have begun to use this approach
+to study the health impacts of tropical cyclones, including several studies of
+Hurricane Maria (e.g.,~\cite{santos2018use, santos2018differential}), Hurricane
+Sandy (e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane
+season in Florida~\parencite{mckinney2011}.  However, to expand this approach
+to study longer time periods for large sets of communities, researchers must be
+able to assess exposure to tropical cyclones consistently and comparably across
+storms, years, and communities.
 
@@
 standard for Atlantic-basin tropical cyclone tracks.  It records the central
-position of a tropical cyclone every six hours, as well as the storm's
-\underline{minimum pressure} and \underline{central maximum sustained winds}. This data is openly
-available through the \ac{HURDAT2}, a post-storm assessment conducted
-by the \ac{US} \ac{NHC} that incorporates data from a variety of sources,
-including satellite data and, when available, aircraft reconnaissance
+position of a tropical cyclone every six hours, as well as the storm's minimum
+central pressure and maximum wind speed.  This data is openly available through
+the \ac{HURDAT2}, a post-storm assessment that incorporates data from several
+sources, including satellite data and, when available, aircraft reconnaissance
 data~\parencite{landsea2013, jarvinen1988}.  With this data, it is
@@
 within a certain distance of the path, of a storm. Some studies have indeed
-used this data to assess tropical cyclone exposure based on how near the
-storm's central track came to the county, either to assess average exposure
-patterns to tropical storms [refs] or to characterize exposure for health
-studies [refs].
+used this data within epidemiological studies to assess tropical cyclone
+exposure based on whether a storm's track passed through a community or how
+near the storm's central track came to the community (e.g.,~\cite{currie2013,
+kinney2008, caillouet2008increase}).
 
-When epidemiological studies assess exposure to a tropical cyclone based on 
-how close the storm came to a community, however, they may misclassify exposure. 
-Tropical cyclones vary dramatically in size: \ac{US} tropical
-cyclones have been observed with radii to maximum winds as small
-as~20~\si{\kilo\metre} and as large
-as~200~\si{\kilo\metre}~\parencite{mallin2006, quiring2011variations}.  While a
-number of tropical cyclone hazards are strongly associated with distance from
-the tropical cyclone's center (e.g., wind and, at the coast, storm surge and
+When epidemiological studies assess exposure to a tropical cyclone using, as a
+proxy measurement, how close the storm came to a community, they may
+misclassify exposure~\parencite{grabich2015measuring}. While a number of
+tropical cyclone hazards are strongly associated with distance from the
+tropical cyclone's center (e.g., wind and, at the coast, storm surge and
 waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
@@
 between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
-tropical cyclone's center~\parencite{moore2012}.
+tropical cyclone's center~\parencite{moore2012}.  Further, when studies use
+distance from the storm's track to assess exposure, they often use the same
+buffer constraints on each side of the tropical cyclone track
+(e.g.,~\cite{czajkowski2011, kinney2008, currie2013}).  However, the forces of
+a tropical cyclone tend to be distributed around its center non-symmetrically.
+Extreme winds are more common to the track's right, where counter-clockwise
+cyclonic winds move in concert with the tropical cyclone's forward
+motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
+associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
+right of the tropical cyclone's track, mostly in the storm's right front
+quadrant~\parencite{moore2012}. Rain, conversely, is often heaviest to the left
+of the track, especially when the tropical cyclone interacts with other weather
+systems~\parencite{atallah2003, atallah2007, zhu2013variations} or undergoes an
+extratropical transition~\parencite{elsberry2002}.
 
-Further, when studies use distance from the storm's track to assess exposure,
-they often use an equal buffer distance on each side of the tropical cyclone
-track (e.g.,~\cite{czajkowski2011, grabich2015, grabich2016, zandbergen2009,
-tansel2010}). However, the forces of a tropical cyclone tend to be distributed
-around the center in a non-symmetrical way. Extreme winds are more common to
-the right of the track, where counter-clockwise cyclonic winds move in concert
-with the tropical cyclone's forward motion~\parencite{halverson2015}, and the
-fatal tornadoes associated with \ac{US} tropical cyclones between~1995 and~2009
-occurred almost exclusively to the right of the tropical cyclone's track,
-mostly in the right front quadrant of the tropical
-cyclone~\parencite{moore2012}. Rain, conversely, is often heaviest to the left
-of the tropical cyclone's track, especially when the tropical cyclone interacts
-with other weather systems~\parencite{atallah2003, atallah2007,
-zhu2013variations} or undergoes an extratropical
-transition~\parencite{elsberry2002}.
+The multi-hazard nature of tropical cyclones therefore makes it hard to assess
+exposure based on how close the storm's central
+track came to the community.  While other approaches have been developed to
+incorporate storm hazards, particulary wind, into exposure assessment (e.g.,
+\cite{grabich2015measuring, zandbergen2009, czajkowski2011}), there is not yet
+a standard approach, and when different studies use different datasets or storm
+hazards when assessing storm exposure, it becomes difficult to compare and
+aggregate findings from different studies.
 
-If a study misclassifies exposure to the hazard or hazards that cause the
-health risk being studied, the study will generate biased estimates of tropical
-cyclone risks and impacts. Further, when different studies use different
-methods to assess exposure to tropical cyclones, their results are difficult to
-meaningfully compare and aggregate. If different methods identify
-similar sets of communities as ``exposed'',  these concerns are less serious.
-However, if different methods differ substantially in which communities they
-identify as ``exposed'', it makes it very important that epidemiological
-studies are thoughtful in how they assess exposure.
+To help epidemiological researchers assess exposure to tropical cyclones, we
+developed an open-source dataset~\parencite{hurricaneexposure}.  These data
+cover all counties in states in the eastern half of the \ac{US} over multiple
+years and include five metrics characterizing exposure to tropical cyclones
+(Table~\ref{tab:exposuremetrics}): (1)~closest distance the storm's central
+track came to the county's center (a proxy measurement of storm exposure used
+in some previous studies); (2)~maximum sustained wind speed at the county's
+center over the course of the storm; (3)~cumulative rainfall in the county over
+the course of the storm; (4)~flood events in the county associated with the
+storm; and (5)~tornado events in the county associated with the storm.  We
+aggregated these data at the county level since data on many potential impacts
+are available at county-level aggregations (e.g., direct hurricane-related
+deaths~\parencite{czajkowski2011}; birth outcomes~\parencite{grabich2015,
+grabich2015measuring}; autism prevalence~\parencite{kinney2008}) and since
+decisions and policies to prepare for, and respond to, tropical cyclones are
+often undertaken at the county level~\parencite{zandbergen2009, rappaport2000}.
+We also wrote functions to explore and map this data and to link it
+with county-level health data~\parencite{hurricaneexposuredata}
 
-To provide open-source data on tropical cyclone exposures to other researchers,
-we developed a dataset of county-level exposure to tropical cyclones in all
-eastern \ac{US} counties for five different metrics
-(Table~\ref{tab:exposuremetrics}): (1)~distance to tropical cyclone track;
-(2)~maximum sustained wind speed; (3)~cumulative rainfall; (4)~flood events;
-and (5)~tornado events~\parencite{hurricaneexposure}. This data is provided at
-the county level since data on many potential impacts are available at
-county-level aggregations (e.g., direct hurricane-related
-deaths~\parencite{czajkowski2011}; birth outcomes~\parencite{grabich2015,
-grabich2016}; autism prevalence~\parencite{kinney2008}) and since decisions and
-policies to prepare for, and respond to, tropical cyclones are often undertaken
-at the county level~\parencite{zandbergen2009, rappaport2000}. Further, we
-explored patterns in these exposure assessments across storm hazards and
-measured how well exposure classification agreed across these five metrics in
-terms of classifying specific counties as exposed to a tropical cyclone.
+Here, we describe how we developed this dataset and explore how its
+measurements compare to other data that could be used to characterize tropical
+cyclone hazards in the context of epidemiological research. \textit{Further, we expand
+on previous research that found large differences in which communities are
+assessed as exposed to a tropical cyclone depending on whether exposure
+assessment is based on distance to the storm's track or if it uses a metric
+incorporating storm-related winds within the county during the storm,
+concluding that a study may be prone to bias from exposure misclassification if
+distance to the storm track is used as a proxy measurement for exposure
+assessment~\parencite{grabich2015measuring}.}  The previous study investigated
+this question for one hurricane season in Florida. Here we expand to
+investigate this question across all counties in states in the eastern half of
+the \ac{US} over multiple storms and years.  Further, we investigate patterns
+in exposure classification based on other metrics of storm-related
+hazards---rainfall, flooding, and tornadoes---that are important to characterize to
+improve understanding of the inland health impacts of tropical
+cyclones~\parencite{czajkowski2011, moore2012}. These results can help
+epidemiologists design studies and plan statistical analysis for
+multi-year, multi-community studies that investigate changes in community-wide
+rates of health outcomes during exposure to the hazards brought by tropical
+cyclones, including counties inland from a storm's landfall.

### 2019-11-07 writing/ehp_methods.tex
Make more edits

@@
 We aggregated data for all counties in the eastern half of the \ac{US}
-(Figure~\ref{fig:hurrtracks}) as of the~2010 \ac{US} Decennial Census and for
-all tracked land-falling or near-land Atlantic-basin tropical cyclones
-between~1988 and~2015---all 136 storms in \ac{HURDAT2} \parencite{landsea2013}
-that came within~250 \si{\kilo\metre} of at least one eastern \ac{US} county in
-this period (Figure~\ref{fig:hurrtracks}).
+(Figure~\ref{fig:hurrtracks}) as of the~2010 \ac{US} Decennial Census. We
+included all tracked land-falling or near-land Atlantic-basin tropical cyclones
+between~1988 and~2015 by covering all 136 storms in \ac{HURDAT2}
+\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
+eastern \ac{US} county in this period (Figure~\ref{fig:hurrtracks}).
 
@@
 
-In the open-source data~\parencite{hurricaneexposuredata}, we provide daily
-precipitation values for each county for the period from five days before to
-three days after each storm's closest approach to the county.  In the software
-that we published in association with this data~\parencite{hurricaneexposure},
-we provide functionality to aggregate these daily values to create cumulative
-precipitation estimates for custom windows around the date of the storm's
-closest approach to the county. For example, a user could determine
-precipitation only on the day the storm was closest to each county, but could
-also determine the cumulative rainfall for a more extended period, like from
-three days before to three days after the storm's closest approach. For the
-analysis of binary exposure classifications in this study, we calculated
-storm-associated rainfall as the sum of precipitation from two days before the
-storm's closest approach to the county to one day after.
+In the open-source data~\parencite{hurricaneexposuredata}, we provide
+precipitation values at a daily resolution for each county for the period from
+five days before to three days after each storm's closest approach to the
+county.  In the software that we published in association with this
+data~\parencite{hurricaneexposure}, we provide functionality to aggregate these
+daily values to create cumulative precipitation estimates for custom windows
+around the date of the storm's closest approach to the county. For example, a
+user could determine precipitation only on the day the storm was closest to
+each county, but could also determine the cumulative rainfall for a more
+extended period, like from three days before to three days after the storm's
+closest approach. For the analysis of binary exposure classifications in this
+study, we calculated storm-associated rainfall as the sum of precipitation from
+two days before the storm's closest approach to the county to one day after.
 
@@
 generated at each county center for each~15-\si{\minute} increment throughout
-the storm tracking period, using a double exponential wind speed model with
+the storm tracking period. We used a double exponential wind speed model with
 inputs for the storm's forward speed, direction, and maximum wind
-speed~\parencite{willoughby2006parametric, stormwindmodel}. This model
-incorporates the asymmetry in wind speeds around the tropical cyclone center
-that results from the storm's forward movement~\parencite{phadke2003modeling,
-stormwindmodel}. We then identified the maximum sustained surface wind speed in
-each county during each storm.
+speed~\parencite{willoughby2006parametric, stormwindmodel}, incorporating the
+asymmetry in wind speeds around the tropical cyclone center that results from
+the storm's forward movement~\parencite{phadke2003modeling}. From these
+15-\si{\minute} interval estimates, we identified the maximum sustained surface
+wind speed in each county during each storm.
 
@@
 For each tropical cyclone, we identified all events with event types related to
-flooding (``Flood,'' ``Flash Flood,'' ``Coastal Flood'') and tornadoes
+flooding (``Flood,'' ``Flash Flood,'' ``Coastal Flood'') or tornadoes
 (``Tornado'') with a start date within a five-day window centered on the date
@@
 in time to the storm but far from the storm's track, and so were likely
-unrelated to the storm, we exculed any events that occurred outside~500
-\si{\kilo\metre} of the tropical cyclone's track. ``Flood,'' ``Flash Flood,''
+unrelated to the storm, we excluded any events that occurred~$\ge$500
+\si{\kilo\metre} from the tropical cyclone's track. ``Flood,'' ``Flash Flood,''
 and ``Tornado'' events in this database were reported by county \ac{FIPS} code
@@
 For the rainfall metric, a distance constraint was also necessary, to ensure
-that rainfall unrelated and far from the storm track were not misattributed to a
+that rainfall unrelated and far from the storm track was not misattributed to a
 storm. Through exploratory analysis, we set this distance metric at~500
@@
 exposures over available years and mapped patterns in
-these exposures. We next investigated agreement between storm exposure
+these exposures.
+
+\textit{We next investigated agreement between storm exposure
 classifications based on these different single-hazard exposures, as well as
 between each of these and a distance-based proxy assessment of tropical cyclone
-exposure (Table~\ref{tab:exposuremetrics}). We calculated the within-storm
+exposure.} We calculated the within-storm
 Jaccard index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between
@@
 large number of eastern \ac{US} counties (those storms with 100 or more
-counties were classified as exposed by at least one of the exposure metrics)
+counties classified as exposed by at least one of the exposure metrics)
 during the years when all exposure data were available (1996\,--\,2011).
-
-\subsection*{Case study}
-
-The societal impact of a disaster depends both on its geophysical forces and on
-the vulnerability of those living in the geographical areas it
-affects~\parencite{chakraborty2005population, anderson2003community,
-cutter1996vulnerability}. An estimate of ``physical exposure''---which
-...---can help to jointly capture these facets of disaster impacts [ref, maybe
-the peduzzi one given below?]. As a case study, we measured physical exposure
-based on each metric of exposure and investigated whether conclusions from
-these assessments varied substantially.
-
-We collected data on the number of electricity-dependent Medicare beneficiaries
-in each study county from the \ac{US} Department of Health \& Human Service's
-emPower Map~2.0~\parencite{empower}. This subpopulation was selected because it
-may be particularly susceptible to health impacts from tropical cyclone
-exposure, especially through the pathway of storm-associated power outages and
-evacuations. We calculated the physical exposure of the electricity-dependent
-Medicare population in each county, based on tropical cyclone assessments using
-each exposure metric (Table~\ref{tab:exposuremetrics}), following
-\citeauthor*{peduzzi2009assessing}~(\citeyear{peduzzi2009assessing}):
-
-\begin{equation}
-E_c = F_c * P_c
-\end{equation}
-
-\noindent where~$E_c$ is the average yearly physical exposure among
-electricity-dependent Medicare beneficiaries in county~$c$ to tropical cyclone
-exposures based on a given metric,~$F_c$ is the estimated yearly expected
-frequency of tropical cyclone exposures in county~$c$ based on that metric,
-and~$P_c$ is the electricity-dependent Medicare population in
-county~$c$, as of July~2017.

### 2019-11-07 writing/ehp_results.tex
Make more edits

@@
 
-We created and published this tropical cyclone exposure data as open-source
+\textit{We created and published this tropical cyclone exposure data as open-source
 data, so they can be used by epidemiological
@@
 and map the data and to integrate it with human health
-datasets~\parencite{hurricaneexposure}.
+datasets~\parencite{hurricaneexposure}.}
 
@@
 from other available sources.  For estimates of storm-associated rainfall, we
-compared estimates from NLDAS-2 re-analysis precipitation data (used as the
-rainfall estimates in the open-source data) with ground-based observations in
-nine sample counties (Figure~\ref{fig:raincomparison}). Within these counties,
-storm-related rainfall measurements were well-correlated between the two data
-sources, with rank correlations (bottom right of each graph in
+compared estimates used in the open-source data with ground-based observations
+in nine sample counties (Figure~\ref{fig:raincomparison}). Within these
+counties, storm-related rainfall measurements were well-correlated between the
+two data sources, with rank correlations (bottom right of each graph in
 Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some evidence
-that the reanalysis data may tend to underestimate rainfall totals in storms
-with extremely high rainfall, based on a few heavy-rainfall storms in Harris
-County, TX, Mobile County, AL, Charleston County, SC, and Wake County, NC
-(Figure~\ref{fig:raincomparison}). However, even in these cases, it was rare
-for a storm to be classified differently (under the precipitation threshold
-of 75~\si{\milli\metre} we used for binary exposure classification for later
-analysis) based on the source of precipitation data---horizontal and vertical
-lines in each small plot in Figure~\ref{fig:raincomparison} show the threshold
-of 75~\si{\milli\metre}, so storms in the lower left and upper right quadrants
-would be classified the same (``exposed'' or ``unexposed'') regardless of the
-precipitation data source, while storms in the upper left and lower right
-quadrants would be classified differently. Such cases are rare.
+that the reanalysis data included in the open-source dataset may tend to
+underestimate rainfall totals in storms with extremely high rainfall, based on
+a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
+County, SC, and Wake County, NC (Figure~\ref{fig:raincomparison}). However,
+even in these cases, it was rare for a storm to be classified differently
+(under the precipitation threshold of 75~\si{\milli\metre} we used for binary
+exposure classification for later analysis) based on the source of
+precipitation data---horizontal and vertical lines in each small plot in
+Figure~\ref{fig:raincomparison} show the threshold of 75~\si{\milli\metre}, so
+storms in the lower left and upper right quadrants would be classified the same
+(``exposed'' or ``unexposed'') regardless of the precipitation data source,
+while storms in the upper left and lower right quadrants would be classified
+differently. Such cases are rare.
 
-For maximum sustained surface wind estimates, we found that the modeled
-estimates, which we included as the primary wind exposure metric in the open
-source data and used for further analysis in this study, generally agreed well
-with estimates based on the wind radii reported in \ac{HURDAT2}. For most
-storms, $\ge$95\% of counties were assigned the same category of wind speed
-($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot};
-$\ge$64~\si{\knot}) by both data sources (Figure~\ref{fig:windcomparison}).
-Disagreement was limited to large storms that were characterized by high winds
-well inland, like Hurricanes Sandy in 2012 and Ike in 2008. For these two
-storms, the modeled wind speed values somewhat overestimated the extent of the
-storm's most severe winds at landfall but then underestimated the extent of
-34\,--\,49.9 \si{\knot} storm winds further inland (Figure~S2). For
-epidemiology researchers who would like to conduct sensitivity analysis
-between these two sets of wind data, we have included estimates from the
-\ac{HURDAT2} wind radii as a secondary measure of wind speeds in the
+For maximum sustained surface wind estimates, we found that the primary wind
+exposure metric in the open source data generally agreed well with the wind
+radii reported in \ac{HURDAT2}. For most storms, $\ge$95\% of counties were
+assigned the same category of wind speed ($<$34~\si{\knot};
+34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) by both
+data sources (Figure~\ref{fig:windcomparison}).  Disagreement was limited to a
+few storms (e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two
+storms, the modeled wind speed values used in the open-source data somewhat
+overestimated the extent of the storm's most severe winds at landfall but then
+underestimated the extent of 34\,--\,49.9 \si{\knot} storm winds further inland
+(Figure~S2). For epidemiology researchers who would like to conduct sensitivity
+analysis between these two sets of wind data, we have included estimates from
+the \ac{HURDAT2} wind radii as a secondary measure of wind speeds in the
 open-source dataset \parencite{hurricaneexposuredata}.
 
-For flood data, we compared flood status, based on event listings in the NOAA
-Storm Events database and included in the open-source data, to streamflow
-measurements at \ac{US} Geological Survey gages within nine sample counties
-(Figure~\ref{fig:floodcomparison}). Across the sampled counties, streamgage
-data generally indicated higher discharge during periods identified as flood
-events based on the NOAA Storm Events database, particularly in counties with
-more reporting streamgages (e.g., Harris County, TX, with~10 reporting stream
-gages; Figure~\ref{fig:floodcomparison}). There were some cases, however, where
-the two flooding data sources were somewhat inconsistent.  For example, there
-were one or two tropical cyclones in several of the counties (Mobile County,
-AL, Escambia County, FL, and Montgomery County, PA) with associated flood event
-listings but for which the total discharge across county streamflow gages was
-relatively low. For storms without a flood listing for the county, in most
-cases the total streamflow discharge in the county was relatively low, and, in
-all but one of these cases all streamgage flows were below the flooding
-threshold. The exception was for one storm in Fulton County, GA (Hurricane Ida
-in 2009).
+For flood data, we compared the flood status values included in the open-source
+data to streamflow measurements at \ac{US} Geological Survey gages within nine
+sample counties (Figure~\ref{fig:floodcomparison}). Across the sampled
+counties, streamgage data generally indicated higher discharge during periods
+identified as flood events based on the NOAA Storm Events database,
+particularly in counties with more reporting streamgages (e.g., Harris County,
+TX, with~10 reporting stream gages; Figure~\ref{fig:floodcomparison}). There
+were some cases, however, where the two flooding data sources were somewhat
+inconsistent.  For example, there were one or two tropical cyclones in several
+of the counties (Mobile County, AL, Escambia County, FL, and Montgomery County,
+PA) with associated flood event listings but for which the total discharge
+across county streamflow gages was relatively low. For storms without a flood
+listing for the county, in most cases the total streamflow discharge in the
+county was relatively low, and, in all but one of these cases all streamgage
+flows were below the flooding threshold. The exception was for one storm in
+Fulton County, GA (Hurricane Ida in 2009).
 
@@
 
-We also measured the average number of counties exposed per storm to each
+\textit{We also measured the average number of counties exposed per storm to each
 hazard (limiting this analysis to tropical cyclones that exposed at least one
@@
 the most counties based on rain, Ike in~2008 based on wind, and Ivan in 2004
-based on flooding and tornadoes (Table~\ref{tab:exposuresummaries}).
+based on flooding and tornadoes (Table~\ref{tab:exposuresummaries}).}
 
@@
 exposure was also more frequent in coastal areas compared to inland areas,
-there were also inland exposures in counties that were never to extreme storm
-winds. Rain and, to some extent, flood exposures were characterized by a
+there were also inland exposures in counties that were never classified as
+exposed based on winds. Flood-based exposures were frequently in the
+Mid-Atlantic region, with a pattern that skewed north compared with other
+exposures.Rain and, to some extent, flood exposures were characterized by a
 pattern defined by the Appalachian Mountains, with fewer exposures west of the
@@
 coastal states, with many in Florida, and almost none north of Maryland.
-Flood-based exposures were frequently in the Mid-Atlantic region, with a
-pattern that skewed north compared with other exposures.  Patterns were similar
-when analysis was restricted to years with exposure data for all four hazards
-available (1996\,--\,2011; Figure~S3).
+Patterns were similar when analysis was restricted to years with exposure data
+for all four hazards available (1996\,--\,2011; Figure~S3).
 
@@
 of hazards, as well as between each hazard and exposure assessment based on
-distance from the storm's track as a proxy for exposure to storm hazards. We
-found that these exposure classifications typically did not strongly agree
-between pairs of metrics. The set of counties identified as ``exposed'' based
-on one metric often overlap little with the set identified as ``exposed'' by
-another metric, as storms frequently brought different hazards to different 
-locations.
+distance from the storm's track. We found that these exposure classifications
+typically did not strongly agree between pairs of metrics---the set of counties
+identified as ``exposed'' based on one metric often overlapped little with the set
+identified as ``exposed'' by another metric, as storms frequently brought
+different hazards to different locations.
 
-Figure~\ref{fig:ivanexposure} shows shows an example: Hurricane Ivan in 2004.
+Figure~\ref{fig:ivanexposure} shows as an example Hurricane Ivan in 2004.
 For the distance-based metric, the counties assessed as exposed follow the
@@
 for one tropical cyclone, and each box in that row shows the measured Jaccard
-coefficient for a pair of metrics. For most tropical cyclones,
-distance-based county-level exposure assessment was, at best, in moderate
-agreement with exposure assessments based on the four hazard-based metrics. For
-almost all tropical cyclones, Jaccard index values between distance and any of
-the four hazard-based metrics were~$\le0.6$ (i.e., out of the counties assessed
-as exposed by at least one of the two metrics in the pair,~60\si{\percent} or
-fewer were assessed the same under both metrics in the pair). For many tropical
-cyclones, the Jaccard index values were~$\le0.4$, indicating poor agreement.
-Among the four hazard-based metrics, the tornado-based metric had
-universally poor agreement with other metrics in county-level classification
-across the tropical cyclones considered.
+coefficient for a pair of metrics.  For almost all tropical cyclones, Jaccard
+index values between distance and any of the four hazard-based metrics
+were~$\le0.6$ (i.e., out of the counties assessed as exposed by at least one of
+the two metrics in the pair,~60\si{\percent} or fewer were assessed the same
+under both metrics in the pair). For many tropical cyclones, the Jaccard index
+values were~$\le0.5$, indicating at best moderate agreement.  Among the four
+hazard-based metrics, the tornado-based metric had universally poor agreement
+with other metrics in county-level classification across the tropical cyclones
+considered.  Agreement between other pairs of metrics were better, but still
+typically at best moderate (Jaccard index values usually $\le0.5$).
 
@@
 agreed moderately to well for all pairs of exposure metrics except those
-including the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
+involving the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
 of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
 in~2003), there was moderate to good agreement for all pairwise combinations of
-distance, rain, and wind, but poor agreement for all other combinations of
+distance, rain, and wind, but poor agreement for other combinations of
 metrics, while for another set of storms (e.g., Matthew in~2004 and Katrina
@@
 pairings of metrics.
-
-\subsection*{Case study}
-
-We conducted a case study comparing estimates of physical exposure to
-tropical cyclones that were measured using each exposure metric.
-Figure~\ref{fig:topelecdependexposure} shows the study counties with the
-highest expected rates of physical exposure to tropical cyclones among
-electricity-dependent Medicare beneficiaries based on each metric. A few
-counties (Miami-Dade County, FL; Harris County, TX) were ranked in the top
-three under almost all exposure metrics
-(Figure~\ref{fig:topelecdependexposure}). However, there were key differences
-across exposure metrics. For example, most of the top counties based on wind-
-and tornado-based metrics were in Florida, while over half of the top counties
-based on the flood-based metric were in Mid-Atlantic states (PA, NY).

### 2019-11-07 R/average_exposure_figure.R

@@ -222,7 +222,7 @@ pdf("figures/averageexposureonly.pdf", width = 3, height = 9)
 print(out)
 dev.off()
 
-pdf("figures/ivanonly.pdf", width = 9, height = 5)
+pdf("figures/ivanonly.pdf", width = 6, height = 3.3)
 print(out2)
 dev.off()
 

### 2019-11-07 R/flood_comparison_figure.R

@@ -107,15 +107,16 @@ out <- county_gage_data %>%
                    position = position_jitter(width = 0.1)) +
         facet_wrap(~ county_cd, ncol = 3, scales = "free_x") +
         theme_classic() +
-        labs(x = "", y = "Total discharge across county streamflow gages") +
+        labs(x = "", 
+             y = bquote('Total discharge across county streamflow gages '~(ft^3/s))) +
         coord_flip() +
-        scale_y_log10() +
+        scale_y_log10(label = scales::comma) +
         theme(legend.position = "bottom") +
         scale_fill_viridis(name = "% streamflow gages over threshold for flooding",
                             breaks = c(0, .5, 1), labels = c("0%", "50%", "100%")) +
         theme(plot.margin = unit(c(5.5, 11, 5.5, 11), "points"))
 
 
-pdf(file = "figures/floodcomparison.pdf", height = 6, width = 7.5)
+pdf(file = "figures/floodcomparison.pdf", height = 6, width = 9)
 print(out)
 dev.off()

### 2019-11-08 writing/ehp_figures.tex
Make more edits
Work some on figure legends

@@
 \begin{figure}[tbhp!] \centering
-\includegraphics[width=0.7\linewidth]{hurrtracks}
-\caption{States and storms considered in this study. All counties in the states
+\includegraphics[width=\linewidth]{hurrtracks}
+\caption{Study area and storms considered in this study. All counties in the states
 	shown in this map were investigated. The lines show the paths of the
 	study storms, which included all tracked storms in 1988\,--\,2015 that
-	are recorded in \ac{HURDAT2}~\parencite{landsea2013} and that came
+	are recorded in \ac{HURDAT2} and that came
 	within 250~\si{\kilo\metre} of at least one \ac{US} county. Thicker
@@
 	indicating that the storm was particularly severe or had notable
-	impacts~\parencite{retirednames}.  }
+	impacts.  }
 \label{fig:hurrtracks}
@@
 \begin{figure}[tbhp!] \centering
-\includegraphics[width=0.9\linewidth]{raincomparison} 
-	\caption{Comparison of storm-associated rainfall estimates between 
-	a re-analysis dataset (included as the rainfall exposure metric
-	in the open-source data and included in further analysis) and ground-based
-	observations in nine sample counties. The rainfall estimates  
+\includegraphics[width=\linewidth]{raincomparison} 
+	\caption{Comparison in nine sample counties of two sources of storm-associated 
+	rainfall estimates: (1) county-level estimates derived from 
+	a re-analysis dataset and (2) county-level estimates based on ground-based
+	observations in nine sample counties. For both, estimates  
 	include rainfall from two days before to one day after the storm's closest
@@
 	and each point shows one tropical storm. The number of storms 
-	within each county and the number of stations reporting rainfall during 
+	within each county and the number of ground-based monitors reporting precipitation during 
 	the county's storms are given above each plot. Horizontal and vertical lines 
 	in each plot show the threshold of 75~\si{\milli\metre} used to classify a storm 
-	as ``exposed'' in further analysis (Table 1). Note that 
-	ranges of the x and y axes differ across counties.
+	as ``exposed'' for the binary classification considered for this metric 
+	in further analysis (Table 1). Note that 
+	ranges of the x- and y-axes differ across counties.
 	} 
@@
 \centering
-\includegraphics[width=0.6\linewidth]{windcomparison}
-\caption{Comparison of modeled maximum sustained surface wind speeds (included
-	as the primary wind exposure metric in the open-source data and used in
-	further analysis) with estimates from \ac{HURDAT2}'s wind speed radii.
+\includegraphics[width=0.8\linewidth]{windcomparison}
+	\caption{Comparison of two sources of wind exposure estimates: 
+	(1) modeled maximum sustained surface wind speeds 
+	and (2) estimates from \ac{HURDAT2}'s wind speed radii.
 	Each point represents a storm, with the x axis giving the percent of
@@
 	wind speeds ($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot};
-	50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) by both sources of wind
-	data for that storm. The color of each point indicates the extent of
-	the storm, giving the number of study counties that were exposed to
+	50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) by both sources of 
+	data. The color of each point gives the number of study counties that were exposed to
 	maximum sustained surface wind speeds of at least 34~\si{\knot} (based
@@
 	2004, the earliest year for which post-storm reanalysis wind speed
-	radii estimates are routinely available in \ac{HURDAT2}, for which at
+	radii estimates are routinely available in \ac{HURDAT2}, and for which at
 	least one study county had a sustained wind speed of $\ge$34~\si{\knot}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{floodcomparison}
-\caption{Comparison for a sample of study counties of flood status based on
-	matching NOAA Storm Events listings to \ac{HURDAT2} storm track data
-	versus five-day total streamflows at county streamflow gages during
-	tropical cyclones. Each small plot shows results for one of the sample
+\includegraphics[width=\linewidth]{floodcomparison}
+\caption{Comparison in nine sample study counties of flood status based on
+	two data sources: (1) NOAA Storm Events listings matched to \ac{HURDAT2} storm track data
+	and (2) county streamflow gages. Each small plot shows results for one of the sample
 	counties. Each point represents a single tropical storm, and the
 	point's position along the x-axis shows the highest daily total
-	streamflow (cubic feet per second), summed across all identified
-	streamgages in the county, for a five-day window centered on the day of
-	the storm's closest approach to the county. The y-axis separates storms
+	streamflow (cubic feet per second) during the five-day window surrounding the storm, 
+	summed across all identified
+	streamgages in the county. The y-axis separates storms
 	for which a flood event was reported in NOAA's Storm Events database
-	for the county with a start date within the five-day window of the
-	storm's closest approach to the county. The color of each point gives
+	in the county with a start date within the five-day window of the
+	storm's closest approach. The color of each point gives
 	the percent of streamflow gages in the county with a daily streamflow
@@
 \includegraphics[width=7cm]{averageexposureonly.pdf} 
-\caption{Average number of county-level storm exposures per decade for each
-	single-hazard exposure metric. The criteria behind each of the five
-	metrics is given in Table \ref{tab:exposuremetrics}. The years used to
+\caption{Average number of county-level storm exposures per decade for binary 
+	classifications based on each
+	single-hazard exposure metric (Table \ref{tab:exposuremetrics}). The years used to
 	estimate these averages are based on years of available exposure data
@@
 \centering
-\includegraphics[width=16cm]{ivanonly.pdf}
+\includegraphics[width=\textwidth]{ivanonly.pdf}
 \caption{Counties classified as exposed to Hurricane Ivan in 2004 under each
@@
 	line shows the track of Hurricane Ivan based on
-	\ac{HURDAT2}~\parencite{landsea2013}.  Similar maps for other
+	\ac{HURDAT2}.  Similar maps for other
 	large-extent storms are given in Figure~S4.
@@
 	\centering 
-\includegraphics[width = 0.6\linewidth]{jaccard_heatmap.pdf} 
-\caption{Agreement between exposure classifications based on different
-         exposure metrics for all storms between~1996 and~2011 for which 
-	 at least~100 counties were exposed based on at least one metric.
-	 Each row shows one storm, and the color of each cell shows the 
-	 measured Jaccard index for each pair of exposure metrics 
-	 (proportion of counties classified as exposed by both metrics out 
-	 of storms classified as exposed by either metric). The colors to the 
-	 right of the main heatmap for each storm indicate the total number of 
-	 counties classified as exposed to the storm by any of the five metrics, 
-	 providing an indication of each storm's extent. Storms are displayed 
-	 within clusters that have similar patterns in county-level exposure 
-	 agreement for metric pairs, based on hierarchical clustering using the 
-	 complete link method~\parencite{murtagh2012algorithms}; columns are also 
-	 ordered based on hierarchical clustering. Maps are available showing the 
-	 counties identified as exposed under each of five metrics for the widest-extent 
-	 storm in each cluster: Hurricane Ivan in~2004 (Figure~\ref{fig:ivanexposure}) 
-	 and Hurricanes Floyd in~1999, Lee in~2011, Cindy in~2005, and Katrina 
-	 in~2005 (Figure~S4).
-} 
+	\includegraphics[width = 0.8\linewidth]{jaccard_heatmap.pdf}
+	\caption{Agreement between exposure
+	classifications based on different exposure metrics for all storms
+	between~1996 and~2011 for which at least~100 counties were exposed
+	based on at least one metric.  Each row shows one storm, and the color
+	of each cell shows the measured Jaccard index for each pair of exposure
+	metrics (proportion of counties classified as exposed by both metrics
+	out of counties classified as exposed by either metric). For 
+	Grace in 2003 and Ida in 2009, there were no county exposures for
+	either the tornado-based metric or the wind-based metric (indicated 
+	by gray squares). Colors to the right of the
+	heatmap show the number of exposed counties
+	based on any of the metrics. Storms are displayed
+	within clusters that have similar patterns based on hierarchical 
+	clustering (complete link method).}
 \label{fig:jaccard} 
@@
 
+\begin{comment}
 \clearpage
@@
 \end{figure*}
+
+\end{comment}

### 2019-11-08 tables/stormseverity.tex
Work some more on draft

@@
 % latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Tue Oct  8 14:03:24 2019
+% Fri Nov  8 11:15:53 2019
 \begin{table}[ht]
 \centering
-\caption{Summary statistics for the number of county tropical cyclone exposures under each metric.     
-	 The median and interquartile range of number of county exposures per tropical cyclone are based on
-         the tropical cyclones for which at least one US county was exposed. The years for which
-         data are available for each metric are given in Table 1.} 
+\caption{Summary statistics for the number of county tropical cyclone exposures under each metric.} 
 \label{tab:exposuresummaries}
-\begin{tabular}{p{1.75cm}p{3cm}p{4cm}p{4cm}}
+\begin{tabular}{p{4.5cm}p{4.5cm}p{4.5cm}}
   \toprule
-Metric & Mean (interquartile range) of county exposures per year & Median (interquartile range) of county exposures per tropical cyclone & Tropical cyclone with most counties exposures (\# exposed counties) \\
+Metric (years available) & Mean (interquartile range) of county exposures per year & Tropical cyclone with most counties exposures (\# exposed counties) \\ 
   \midrule
-Rain & 401 (152, 553) & 68 (14, 146) & Frances, 2004 (464) \\ 
-  Wind & 143 (51, 232) & 25 (14, 58) & Ike, 2008 (251) \\ 
-  Flood & 213 (84, 248) & 26 (6, 58) & Ivan, 2004 (318) \\ 
-  Tornado & 53 (16, 43) & 7 (2, 16) & Ivan, 2004 (91) \\ 
+Distance (1988--2015) & 433 (224, 503) & Beryl, 1994 (330) \\ 
+  Rain (1988--2015) & 401 (152, 553) & Frances, 2004 (464) \\ 
+  Wind (1988--2011) & 143 (51, 232) & Ike, 2008 (251) \\ 
+  Flood (1996--2015) & 213 (84, 248) & Ivan, 2004 (318) \\ 
+  Tornado (1996--2015) & 53 (16, 43) & Ivan, 2004 (91) \\ 
    \bottomrule

### 2019-11-08 writing/ehp_abstract.tex
Work some more on draft

@@
 	cyclone exposure assessment for health-related research and
-	(2)~investigate agreement between county-level assessments of tropical cyclone
-	exposure when based on different storm hazards, to clarify 
-	risk of exposure misclassification in epidemiologic studies. \\ 
+	(2)~investigate patterns and agreement between county-level assessments of tropical cyclone
+	exposure when based on different storm hazards, to inform the design of tropical cyclone epidemiological studies. \\ 
 \textbf{Methods:} We created an open-source dataset with county-level data on
@@
 	Atlantic basin storms over multiple years (1996\,--\,2011 for all
-	metrics, up to~1988\,--\,2015 for specific metrics). [Validation] We measured
-	agreement in county-level exposure assessment for all pairwise
-	combinations of these metrics. \\ 
-\textbf{Results:} We created an open-source dataset ... 
-	[Most data agreed well. Flooding is hard to measure.]
-	In most of the storms investigated, agreement in county-level
-	exposure assessment was at best moderate, and often poor, when using
-	distance compared with four hazard-based measurements. There was also
-	substantial disagreement in exposure assessment based on all pairwise
-	combinations of the four hazard-based measurements. There were
-	geographic patterns in exposures based on these metrics, with most
-	wind-based exposures near the coast, flood- and rain-based exposures
-	extending to inland and northern counties, and tornado-based exposures
-	in southern counties. \\ 
+	metrics, up to~1988\,--\,2015 for specific metrics). We compared these
+	measurements to other potential data sources and investigated patterns
+	and agreement among binary exposure classifications based on these
+	metrics across multiple years and counties. \\ 
+\textbf{Results:} Here, we created and shared a large-scale dataset with 
+	consistent and comprehensive measurements of county-level exposure
+	to storm hazards. These data typically 
+	agreed with data from other sources, although there are some caveats
+	that should be considered by potential users. Long-term geographic patterns in binary 
+	classifications based on these data agree with previous results from
+	atmospheric science. County-level exposure classification for a storm differs substantially 
+	depending on the hazard metric used. \\ 
 \textbf{Discussion:} Our results provide a multi-hazard dataset that can be
 	leveraged for epidemiologic research on tropical cyclones.  Further,	
-	our results suggest that when epidemiologic studies use distance as a
+	our results suggest that epidemiologic studies that use distance as a
 	surrogate for tropical cyclone exposure, or use one hazard-based metric
-	when the impact is partly or fully caused by a different hazard, the
-	study will be prone to exposure misclassification.  
+	when the impact is partly or fully caused by a different hazard, will 
+	likely be prone to exposure misclassification.  
 \end{abstract}

### 2019-11-08 writing/ehp_discussion.tex
Work some more on draft

@@
 spanning over a century have found similar patterns for distance- and
-wind-based metrics~\textcite{zandbergen2009, kruk2010} \textbf{[double-check
-that refs found similar patterns]}.
+wind-based metrics~\parencite{zandbergen2009, kruk2010}
 
@@
 exposure to any of the other hazards considered, would often be problematic,
-introducing exposure misclassification. \textbf{[Compare with Grabich results?]}
-For some studies, such exposure misclassification might plausibly be
-differential (i.e., associated with the outcome of interest or with factors
-associated with risk of the outcome of interest).  For example, tropical
-cyclone wind exposures tend to be concentrated in counties near the coast,
-while tropical cyclone wind exposures sometimes extend well inland.  If the
-etiologically-relevant exposure for a health outcome is extreme rainfall but
-exposure is classified based on measurements of wind, the probability of being
-misclassified as unexposed would be higher in inland counties, while the
-probability of being misclassified as exposed would be higher in coastal
-counties. If coastal counties differ from inland counties in either the outcome
-of interest or in factors associated with risk of that outcome, exposure
-misclassification would be differential~\parencite{savitz2016interpreting}.
-Such differential exposure misclassification could bias estimates of tropical
-cyclone effects either towards the null (estimating a lower or null association
-compared to the true association that exists) or away from the null (estimating
-a larger association than actually exists)~\parencite{savitz2016interpreting,
-armstrong1998effect}.
+introducing exposure misclassification. This result reinforces similar findings
+for a study of Florida's 2004 storm season on methods of measuring exposure to
+tropical cyclones for epidemiological
+research~\parencite{grabich2015measuring}.  For some studies, such exposure
+misclassification might plausibly be differential (i.e., associated with the
+outcome of interest or with factors associated with risk of the outcome of
+interest).  For example, tropical cyclone wind exposures tend to be
+concentrated in counties near the coast, while tropical cyclone wind exposures
+sometimes extend well inland.  If the etiologically-relevant exposure for a
+health outcome is extreme rainfall but exposure is classified based on
+measurements of wind, the probability of being misclassified as unexposed would
+be higher in inland counties, while the probability of being misclassified as
+exposed would be higher in coastal counties. If coastal counties differ from
+inland counties in either the outcome of interest or in factors associated with
+risk of that outcome, exposure misclassification would be
+differential~\parencite{savitz2016interpreting}.  Such differential exposure
+misclassification could bias estimates of tropical cyclone effects either
+towards the null (estimating a lower or null association compared to the true
+association that exists) or away from the null (estimating a larger association
+than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.
 
@@
 storm-associated winds at a finer spatial resolution; however, measurements of
-other hazards cannot similarly be re-scaled through tools we provide. Finally,
+other hazards cannot similarly be re-scaled through tools we provide. Next,
 we provide these data and associated software tools through R packages, and so
@@
 user-friendly web-based access of the data \parencite{shiny2019}. 
+Finally, to assess patterns and agreement for binary exposure classifications, 
+we have chosen one set of sensible thresholds for binary classifications based
+on continuous metrics (rainfall, maximum sustained winds, and distance from 
+the storm's track). Results and conclusions would differ somewhat with other
+threshold choices. We have published all our code for this analysis online
+[web address], in case other researchers would like to explore other 
+threshold choices for these analyses.
 
@@
 
-To conduct epidemiological studies that span multiple communities and tropical 
-cyclones, it is critical to have consistent and comprehensive measurements of 
-community-level exposure to hazards of the storm. Here we have created and 
-shared a dataset that provides these data for counties in the United States
-over multiple years. Despite some limitations in this dataset, it provides 
-a powerful tool for expanding tropical cyclone epidemiology studies to more
-extensively leverage existing administrative and vital data, allowing researchers
-to improve our understanding of how these storms affect community-wide risk of 
-numerous health outcomes, including outcomes that are also common outside of storm 
-periods and so may be hard to attribute on a case-by-case basis to a storm. 
-Further, this dataset provides hazard measurements that are comparable across
-communities and storms, which allows epidemiological researchers to design studies
-to explore how health risks are modified by characteristics of both the storms and
-the communities that are hit. The data are given in an open-source format, along 
-with associated software tools, which allows them to be freely used and for others
-to explore all associated code and to contribute additions through platforms like 
-GitHub.
+To conduct tropical cyclone epidemiological studies that span multiple
+communities and storms, it is critical to have consistent and comprehensive
+measurements of county-level exposure to storm hazards. Here we have created
+and shared a dataset that provides these data for counties in the United States
+over multiple years. Despite some limitations in these data, they provide a
+powerful tool for expanding tropical cyclone epidemiology studies to more
+extensively leverage existing administrative and vital data, allowing
+researchers to investigate how these storms affect community-wide risk of
+numerous health outcomes, including outcomes that are also common outside of
+storm periods and so may be hard to identify through traditional disaster
+surveillance.  Further, this dataset provides hazard measurements that are
+comparable across communities and storms, which allows epidemiological
+researchers to design studies to explore how health risks are modified by
+characteristics of both the storms and the communities that are hit. The data
+are given in an open-source format, along with associated software tools, which
+allows them to be freely used and for others to explore all associated code and
+to contribute additions through platforms like GitHub.
 

### 2019-11-08 writing/ehp_introduction.tex
Work some more on draft

@@
 measurements compare to other data that could be used to characterize tropical
-cyclone hazards in the context of epidemiological research. \textit{Further, we expand
-on previous research that found large differences in which communities are
-assessed as exposed to a tropical cyclone depending on whether exposure
-assessment is based on distance to the storm's track or if it uses a metric
-incorporating storm-related winds within the county during the storm,
-concluding that a study may be prone to bias from exposure misclassification if
-distance to the storm track is used as a proxy measurement for exposure
-assessment~\parencite{grabich2015measuring}.}  The previous study investigated
-this question for one hurricane season in Florida. Here we expand to
+cyclone hazards in the context of epidemiological research. Further, we expand
+on previous research that investigated methods to measure exposure to tropical
+cyclones for epidemiological research~\parencite{grabich2015measuring}.  The
+previous study investigated differences in which communities were assessed as
+exposed to tropical cyclones during the 2004 hurricane season in Florida,
+depending on whether exposure assessment was based on distance to the storm's
+track or used a metric incorporating storm-related winds within the county
+during the storm.  They found important differences across methods of assessing
+storm exposure, concluding that a study may be prone to bias from exposure
+misclassification if distance to the storm track is used as a proxy measurement
+for exposure assessment~\parencite{grabich2015measuring}.  Here we expand to
 investigate this question across all counties in states in the eastern half of
@@
 in exposure classification based on other metrics of storm-related
-hazards---rainfall, flooding, and tornadoes---that are important to characterize to
-improve understanding of the inland health impacts of tropical
+hazards---rainfall, flooding, and tornadoes---that are important to
+characterize to improve understanding of the inland health impacts of tropical
 cyclones~\parencite{czajkowski2011, moore2012}. These results can help
-epidemiologists design studies and plan statistical analysis for
-multi-year, multi-community studies that investigate changes in community-wide
-rates of health outcomes during exposure to the hazards brought by tropical
-cyclones, including counties inland from a storm's landfall.
+epidemiologists design studies and plan statistical analysis for multi-year,
+multi-community studies that investigate changes in community-wide rates of
+health outcomes during exposure to the hazards brought by tropical cyclones,
+including counties inland from a storm's landfall.

### 2019-11-08 writing/ehp_methods.tex
Work some more on draft

@@
 
-\textit{We next investigated agreement between storm exposure
-classifications based on these different single-hazard exposures, as well as
-between each of these and a distance-based proxy assessment of tropical cyclone
-exposure.} We calculated the within-storm
-Jaccard index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between
-each pair of exposure metrics. The Jaccard index~($J_s$) measures similarity
+Next, we investigated how well exposure assessment agreed across 
+these metrics. For each storm we investigated the degree to which the 
+set of counties assessed as exposed based on one metric overlapped with the set
+assessed as exposed based on each other metric, including a metric that used
+distance of a county from the storm track as a proxy for exposure to storm hazards. 
+We calculated the within-storm
+Jaccard index ($J_s$)~\parencite{jaccard1901distribution, jaccard1908nouvelles} between
+each pair of exposure metrics, which measures similarity
 between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical cyclone~$s$ as the

### 2019-11-08 writing/ehp_results.tex
Work some more on draft

@@
 
-\textit{We created and published this tropical cyclone exposure data as open-source
-data, so they can be used by epidemiological
-researchers~\parencite{hurricaneexposuredata}. These data include continuous
-county-level measurements for the closest distance of each storm, cumulative
-rainfall, and maximum sustained surface wind speed. These continuous metrics
-can be used for classifying counties as exposed or unexposed, using thresholds
-selected by the user, or can be used as continuous metrics. The dataset also
-include binary data on flood and tornado events associated with each storm in
-each county. To accompany these data, we published software tools to explore
-and map the data and to integrate it with human health
-datasets~\parencite{hurricaneexposure}.}
+We created and published this tropical cyclone exposure data as an R
+package~\parencite{hurricaneexposuredata}.  The package size exceeded the
+recommended maximum size for the standard repository for publishing R packages
+(\ac{CRAN}). Therefore, we used the \textit{drat} framework to set up our own
+repository to host the package~\parencite{anderson2017hosting}. These data
+include continuous county-level measurements for the closest distance of each
+storm, cumulative rainfall, and maximum sustained surface wind speed. These
+continuous metrics can be used for classifying counties as exposed or
+unexposed---using thresholds selected by the user---or can be extracted
+directly as continuous metrics. The dataset also include binary data on flood
+and tornado events associated with each storm in each county. To accompany
+these data, we published an associated R package on \ac{CRAN} with software
+tools to explore and map the data and to integrate it with human health
+datasets~\parencite{hurricaneexposure}.
 
@@
 county was relatively low, and, in all but one of these cases all streamgage
-flows were below the flooding threshold. The exception was for one storm in
-Fulton County, GA (Hurricane Ida in 2009).
+flows were below the flooding threshold. The exception was for Hurricane Ida in
+2009 in Fulton County, GA.
 
@@
 ($>$200 / yr on average), and most frequent for tropical cyclone rain exposure
-($>$400 / yr on average).
-
-\textit{We also measured the average number of counties exposed per storm to each
-hazard (limiting this analysis to tropical cyclones that exposed at least one
-county to that hazard; Table~\ref{tab:exposuresummaries}).  If a tropical
-cyclone exposed any counties to tornadoes, it was typically only a few (median
-counties exposed per storm:~7). When a storm exposed any counties to other
-tropical cyclone exposures, the swath of exposed counties tended to be much
-larger: a median of~25 counties exposed for high winds per storm,~26 for
-flooding, and~68 for extreme rainfall.  For every hazard except tropical
-cyclone tornadoes, we identified at least one tropical cyclone that exposed
-over~250 counties (Table~\ref{tab:exposuresummaries}).  However, the
-largest-extent tropical cyclone varied across hazards: Frances in~2004 exposed
-the most counties based on rain, Ike in~2008 based on wind, and Ivan in 2004
-based on flooding and tornadoes (Table~\ref{tab:exposuresummaries}).}
+($>$400 / yr on average). For every hazard except tornadoes, we identified at
+least one tropical cyclone that exposed over~250 counties
+(Table~\ref{tab:exposuresummaries}).  However, the largest-extent tropical
+cyclone varied across hazards: Frances in~2004 exposed the most counties based
+on rain, Ike in~2008 based on wind, and Ivan in 2004 based on flooding and
+tornadoes (Table~\ref{tab:exposuresummaries}).
 

### 2019-11-08 R/storm_severity_table.R

@@ -157,20 +157,17 @@ total_sum <- bind_rows(bind_cols(dist_sum, dist_sum2),
          years = c("1988--2015", "1988--2015", "1988--2011",
                    "1996--2015", "1996--2015"),
          yearly_sum = paste(round(mean_exp), " (", round(p25_exp), 
-                            ", ", round(p75_exp), ")", sep = "")) %>% 
-  select(metric, yearly_sum, sum, max_storm)
+                            ", ", round(p75_exp), ")", sep = ""),
+         metric = paste0(metric, " (", years, ")")) %>% 
+  select(metric, yearly_sum, max_storm)
 
 total_sum %>% 
   filter(metric != "Distance") %>% 
-  rename(Metric = metric,
+  rename(`Metric (years available)` = metric,
          `Mean (interquartile range) of county exposures per year` = yearly_sum,
-         `Median (interquartile range) of county exposures per tropical cyclone` = sum,
          `Tropical cyclone with most counties exposures (# exposed counties)` = max_storm) %>%
-  xtable(align = "cp{1.75cm}p{3cm}p{4cm}p{4cm}",
-         caption = "Summary statistics for the number of county tropical cyclone exposures under each metric.     
-	 The median and interquartile range of number of county exposures per tropical cyclone are based on
-         the tropical cyclones for which at least one US county was exposed. The years for which
-         data are available for each metric are given in Table 1.",
+  xtable(align = "cp{4.5cm}p{4.5cm}p{4.5cm}",
+         caption = "Summary statistics for the number of county tropical cyclone exposures under each metric.",
          label = "tab:exposuresummaries",
          digits = 0) %>%
   print(include.rownames = FALSE, booktabs = TRUE, 

### 2019-11-12 writing/ehp_abstract.tex
Work some more on draft

@@
 
-\noindent \textbf{Background:} Tropical cyclones bring severe impacts to
-	\ac{US} communities, including substantial human health risks, from
-	hazards that include wind, rain, flooding, and tornadoes. However, 
-	it is difficult to conduct multi-year, multi-storm epidemologic 
-	studies without storm hazard datasets that can easily be linked to 
-	human health data. Further, while some storm exposure metrics are
-	straightforward to measure (e.g., distance from a community to a 
-	storm's track), they may poorly capture exposure to storm hazards. \\ 
+\noindent \textbf{Background:} 
+        Tropical cyclone epidemiology can be 
+	advanced through methods to assess 
+	county-level exposure to storm hazards, particularly 
+	methods that are comprehensive and 
+	consistent across space and time, allowing multi-year, multi-storm studies. 
+	Further, an understanding of 
+	patterns in and between hazard-specific exposure metrics can help in 
+	designing tropical cyclone epidemiological studies.\\ 
 \textbf{Objectives:} (1)~Provide an open-source dataset to improve tropical
-	cyclone exposure assessment for health-related research and
+	cyclone exposure assessment for epidemiological research and
 	(2)~investigate patterns and agreement between county-level assessments of tropical cyclone
-	exposure when based on different storm hazards, to inform the design of tropical cyclone epidemiological studies. \\ 
+	exposure based on different storm hazards. \\ 
 \textbf{Methods:} We created an open-source dataset with county-level data on
-	exposure to tropical cyclones in the \ac{US}. The dataset includes
-	measures of distance from the storm track, maximum sustained wind,
-	rainfall, flooding, and tornadoes for all land-falling or near-land
-	Atlantic basin storms over multiple years (1996\,--\,2011 for all
-	metrics, up to~1988\,--\,2015 for specific metrics). We compared these
-	measurements to other potential data sources and investigated patterns
+	exposure to four tropical cyclone hazards: maximum sustained winds,
+	rainfall, flooding, and tornadoes. The data cover all eastern \ac{US} counties 
+	for all land-falling or near-land
+	Atlantic basin storms for~1996\,--\,2011 for all
+	metrics (up to~1988\,--\,2015 for specific metrics). 	
+	We validated 
+	measurements against other data sources and investigated patterns
 	and agreement among binary exposure classifications based on these
-	metrics across multiple years and counties. \\ 
-\textbf{Results:} Here, we created and shared a large-scale dataset with 
-	consistent and comprehensive measurements of county-level exposure
-	to storm hazards. These data typically 
-	agreed with data from other sources, although there are some caveats
-	that should be considered by potential users. Long-term geographic patterns in binary 
+	metrics. We also compared exposure assessment
+	based on using distance between a county and the storm's track as a proxy for
+	storm exposure.\\ 
+\textbf{Results:} Measurements in the open-source dataset created here typically
+        agreed with data from other sources, and we present and discuss areas of
+	disagreement and other caveats for using the data.
+        Long-term geographic patterns in binary 
 	classifications based on these data agree with previous results from
@@
 	leveraged for epidemiologic research on tropical cyclones.  Further,	
-	our results suggest that epidemiologic studies that use distance as a
+	our results suggest that the use of distance as a
 	surrogate for tropical cyclone exposure, or use one hazard-based metric
-	when the impact is partly or fully caused by a different hazard, will 
-	likely be prone to exposure misclassification.  
+	when the impact is partly or fully caused by a different hazard, would often 
+	result in exposure misclassification in epidemiological studies.  
 \end{abstract}

### 2019-11-12 writing/ehp_discussion.tex
Work some more on draft

@@
 cyclone exposure metrics and explore limitations in that data.  
-
 Further, we explore patterns in storm exposure classifications based on
@@
 the health risks and impacts associated with tropical cyclones exposures as
-well as insights to help in designing studies.  The open-source data and
-software we have created provides multi-hazard, county-level exposure data for
-such studies.
+well as insights to help in designing studies.
 
@@
 with historic rainfall and flooding in
-Georgia~\parencite{shepherd2011overview}.  In this case the flooding
+Georgia~\parencite{shepherd2011overview}.  In this case, the flooding
 associated with Ida was incorporated into an ongoing flood event listing, with
@@
 captured, for which the occurrance of a storm in the midst of an ongoing storm
-event listing would be improperly attributed to the storm.  However, distance
+event might be improperly attributed to the storm.  However, distance
 and time restrictions, like those we used in matching NOAA Storm Event listings
@@
 tropical cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.
-The exposure averages we calculated are somewhat limited as estimates of
+The exposure averages we calculated may be limited as estimates of
 long-term frequencies, as tropical cyclones follow decadal
 patterns~\parencite{kossin2007more} likely not adequately captured in the
-available data, which covered less than 20 years. However, exposure assessments
-spanning over a century have found similar patterns for distance- and
-wind-based metrics~\parencite{zandbergen2009, kruk2010}
+available data, which covered less than 20 years.
 
@@
 exposures, as well as between pairs of hazard-specific metrics
-(Figures~\ref{fig:ivanexposure}\,--\, \ref{fig:jaccard} and S4\,--\,S5).  These
+(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard} and S4\,--\,S5).  These
 findings align with previous results from atmospheric science and related
@@
 
-This suggests that the use of a distance-based metric to assess exposure to any
-of these hazards, or the use of measurements from one hazard as a proxy for
-exposure to any of the other hazards considered, would often be problematic,
-introducing exposure misclassification. This result reinforces similar findings
-for a study of Florida's 2004 storm season on methods of measuring exposure to
-tropical cyclones for epidemiological
-research~\parencite{grabich2015measuring}.  For some studies, such exposure
-misclassification might plausibly be differential (i.e., associated with the
-outcome of interest or with factors associated with risk of the outcome of
-interest).  For example, tropical cyclone wind exposures tend to be
-concentrated in counties near the coast, while tropical cyclone wind exposures
-sometimes extend well inland.  If the etiologically-relevant exposure for a
-health outcome is extreme rainfall but exposure is classified based on
-measurements of wind, the probability of being misclassified as unexposed would
-be higher in inland counties, while the probability of being misclassified as
-exposed would be higher in coastal counties. If coastal counties differ from
-inland counties in either the outcome of interest or in factors associated with
-risk of that outcome, exposure misclassification would be
+Therefore, the use of a distance-based metric to assess exposure to any of
+these hazards, or the use of measurements from one hazard as a proxy for
+exposure to any of the other hazards considered, would often introduce exposure
+misclassification. This conclusion reinforces similar findings from a study of
+Florida's 2004 storm season~\parencite{grabich2015measuring}.  For some
+studies, such exposure misclassification might plausibly be differential.  For
+example, tropical cyclone wind exposures tend to be concentrated in counties
+near the coast, while tropical cyclone wind exposures sometimes extend well
+inland.  If the etiologically-relevant exposure for a health outcome is extreme
+rainfall but exposure is classified based on measurements of wind, the
+probability of being misclassified as unexposed would be higher in inland
+counties, while the probability of being misclassified as exposed would be
+higher in coastal counties. If coastal counties differ from inland counties in
+either the outcome of interest or in factors associated with risk of that
+outcome, exposure misclassification would be
 differential~\parencite{savitz2016interpreting}.  Such differential exposure
@@
 (e.g., Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto
-in~2006 Figure~\ref{fig:jaccard}).  Hurricanes Floyd in~1999 (Figure~S4) and
+in~2006 (Figure~\ref{fig:jaccard})).  Hurricanes Floyd in~1999 and
 Irene in~2011 both made their first \ac{US} landfall in North Carolina at minor
@@
 between a county and the storm's track.  However, for these storms it may be
-difficult to try to untangle the contribution of each hazard to the overall
+difficult to untangle the contribution of each hazard to the overall
 effect of the storm, given that several hazards have similar geographical
@@
 dangerous storm surge, which is not specifically covered in this dataset
-(although some resulting flooding would be captured). We are exploring ways to
+(although some resulting coastal flooding is captured). We are exploring ways to
 include this in future versions of the dataset, although to date we have
@@
 datasets are aggregated at this level, but some may be aggregated at a finer
-spatial resolution (e.g., Census tract or ZIP code). We have published the wind
+spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e., point locations
+for each outcome). We have published the wind
 model we used to create this dataset as its own open-source R package
-\parencite{stormwindmodel}, and this model can be used to model
+\parencite{stormwindmodel}, and it can be used to model
 storm-associated winds at a finer spatial resolution; however, measurements of
@@
 communities and storms, it is critical to have consistent and comprehensive
-measurements of county-level exposure to storm hazards. Here we have created
-and shared a dataset that provides these data for counties in the United States
-over multiple years. Despite some limitations in these data, they provide a
-powerful tool for expanding tropical cyclone epidemiology studies to more
-extensively leverage existing administrative and vital data, allowing
-researchers to investigate how these storms affect community-wide risk of
-numerous health outcomes, including outcomes that are also common outside of
-storm periods and so may be hard to identify through traditional disaster
-surveillance.  Further, this dataset provides hazard measurements that are
-comparable across communities and storms, which allows epidemiological
-researchers to design studies to explore how health risks are modified by
-characteristics of both the storms and the communities that are hit. The data
-are given in an open-source format, along with associated software tools, which
-allows them to be freely used and for others to explore all associated code and
-to contribute additions through platforms like GitHub.
+measurements of exposure to storm hazards. Here we have created and shared a
+dataset that provides these data for counties in the United States over
+multiple years. Despite some limitations in these data, they provide a powerful
+tool for expanding tropical cyclone epidemiology studies to more extensively
+leverage existing administrative health data, allowing researchers to
+investigate how these storms affect county-wide health risk, including for
+outcomes also common outside of storm periods.  Further, this dataset provides
+hazard measurements that are comparable across communities and storms, which
+allows epidemiological researchers to design studies to explore how health
+risks are modified by characteristics of both the storms and the communities
+that are hit. The data are given in an open-source format, along with
+associated software tools, which allows them to be freely used and for others
+to explore all associated code and to contribute additions through platforms
+like GitHub.
 
-Based on our analysis in this paper, these data sources are typically in agreement
-with other sources of data available to characterize storm-associated hazards
-(e.g., ground-based monitors, streamgages, post-storm wind radii estimates). 
-However, researchers who are planning to use the data should explore the analyses
-presented in this paper to understand the strengths and weaknesses of the data. 
-Our analyses also indicate that a storm often brings different combinations of 
-hazards to different counties along or near its path. Our results in comparing 
-storm hazards suggests that county-level storm exposure is not well-characterized 
-by the closest distance that a storm's central track came to a county, and that
-exposure to one storm hazard within a county (e.g., severe winds) does not imply 
-exposure to other hazards (e.g., excessive rainfall, flooding). As a result, it is 
-critical that researchers consider which storm hazards are likely on the causal 
-pathway for the outcomes they are studying, and to characterize storm exposure
-in a way that captures those specific hazards.
+Based on our analysis in this paper, these data are typically in agreement with
+measurements from other sources of data available to characterize
+storm-associated hazards (e.g., ground-based monitors, streamgages, post-storm
+wind radii estimates).  However, researchers who are planning to use the data
+should explore the analyses presented in this paper to understand the strengths
+and weaknesses of the data.  Our results suggest that county-level storm
+exposure is not well-characterized by the closest distance that a storm's
+central track came to a county, and that exposure to one storm hazard within a
+county (e.g., severe winds) does not imply exposure to other hazards (e.g.,
+excessive rainfall, flooding). As a result, it is critical that researchers
+consider which storm hazards are likely on the causal pathway for the outcomes
+they are studying, and to characterize storm exposure in a way that captures
+those specific hazards.

### 2019-11-12 writing/ehp_introduction.tex
Work some more on draft

@@
 
-Recently, several major tropical cyclones have hit \ac{US} cities, including
-Hurricanes Harvey, Irma, and Maria in 2017~\parencite{blake20182017} and
-Hurricanes Florence and Michael in 2018 \parencite{avila20192018}. Researchers
-are exploring how these storms impacted human health
-(e.g.,~\cite{santos2018use, rivera2018estimating, santos2018differential,
-grineski2019impact, issa2018deaths, tanz2019notes, paul2019brief}), including
-through projects funded by \ac{NIH} Rapid Response
-grants~\parencite{nihreporter}. These projects will add to evidence
-characterizing how hurricanes impact human health, including evidence from
+Hurricanes and other tropical cyclones can severely impact human health in
+\ac{US} communities, and some of these impacts have been characterized through
 multi-year studies based on fatality data from post-disaster
 surveillance~\parencite{rappaport2000, rappaport2014fatalities,
-rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.
-
-For several types of climate-related disasters---including heat waves, floods,
-and wildfires---research based on post-disaster surveillance data has been
-richly supplemented by studies analyzing trends in community-wide
-administrative or vital data, including time series or case-crossover studies
+rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.  For
+several types of climate-related disasters---including heat waves, floods, and
+wildfires---research based on post-disaster surveillance data has been richly
+supplemented by studies analyzing trends in community-wide administrative or
+vital data, including time series or case-crossover studies
 (e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
 liu2017wildfire, milojevic2017mental}).  These studies typically compare the
-rate of a health outcome in a community during a disaster or disasters to rates
-during comparable non-disaster periods.  While these studies cannot attribute
-specific cases (e.g., a specific person's death) to a disaster, they can help
-characterize the community-wide change in health risk during disasters, and can
-help capture impacts that might be missed or underestimated in post-disaster
+rate of a health outcome in a community during a disaster to rates during
+comparable non-disaster periods.  While these studies cannot attribute specific
+cases (e.g., a specific person's death) to a disaster, they can help
+characterize the community-wide change in health risk during disasters and
+capture impacts that might be missed or underestimated in post-disaster
 surveillance.  In many cases, these studies analyze multi-year, multi-community
@@
 explore how a disaster's characteristics, or the characteristics of the
-community it hits, modify associated health risk (e.g.,~\cite{anderson2010heat,
+community it hits, modify associated health risks (e.g.,~\cite{anderson2010heat,
 son2012impact, liu2017wildfire}).  Some studies have begun to use this approach
@@
 position of a tropical cyclone every six hours, as well as the storm's minimum
-central pressure and maximum wind speed.  This data is openly available through
-the \ac{HURDAT2}, a post-storm assessment that incorporates data from several
-sources, including satellite data and, when available, aircraft reconnaissance
-data~\parencite{landsea2013, jarvinen1988}.  With this data, it is
-straightforward to measure whether a community was in the direct path, or
-within a certain distance of the path, of a storm. Some studies have indeed
-used this data within epidemiological studies to assess tropical cyclone
-exposure based on whether a storm's track passed through a community or how
-near the storm's central track came to the community (e.g.,~\cite{currie2013,
-kinney2008, caillouet2008increase}).
+central pressure and maximum wind speed.  These data are openly available
+through the \ac{HURDAT2}, a post-storm assessment that incorporates data from
+several sources, including satellite data and, when available, aircraft
+reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these data, it
+is straightforward to measure whether a community was in the direct path, or
+within a certain distance of the central track, of a storm, as some previous
+epidemiological studies have done to assess exposure to a tropical storm
+(e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}).
 
-When epidemiological studies assess exposure to a tropical cyclone using, as a
-proxy measurement, how close the storm came to a community, they may
-misclassify exposure~\parencite{grabich2015measuring}. While a number of
+However, when epidemiological studies assess exposure to a tropical cyclone
+using, as a proxy measurement, how close the storm came to a community, they
+may misclassify exposure~\parencite{grabich2015measuring}. While a number of
 tropical cyclone hazards are strongly associated with distance from the
@@
 distance from the storm's track to assess exposure, they often use the same
-buffer constraints on each side of the tropical cyclone track
-(e.g.,~\cite{czajkowski2011, kinney2008, currie2013}).  However, the forces of
-a tropical cyclone tend to be distributed around its center non-symmetrically.
-Extreme winds are more common to the track's right, where counter-clockwise
-cyclonic winds move in concert with the tropical cyclone's forward
+buffer constraints on each side of the storm track
+(e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
+cyclone tend to be distributed around its center non-symmetrically.  Extreme
+winds are more common to the track's right, where counter-clockwise cyclonic
+winds move in concert with the tropical cyclone's forward
 motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
 associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
-right of the tropical cyclone's track, mostly in the storm's right front
-quadrant~\parencite{moore2012}. Rain, conversely, is often heaviest to the left
-of the track, especially when the tropical cyclone interacts with other weather
-systems~\parencite{atallah2003, atallah2007, zhu2013variations} or undergoes an
-extratropical transition~\parencite{elsberry2002}.
-
-The multi-hazard nature of tropical cyclones therefore makes it hard to assess
-exposure based on how close the storm's central
-track came to the community.  While other approaches have been developed to
-incorporate storm hazards, particulary wind, into exposure assessment (e.g.,
-\cite{grabich2015measuring, zandbergen2009, czajkowski2011}), there is not yet
-a standard approach, and when different studies use different datasets or storm
-hazards when assessing storm exposure, it becomes difficult to compare and
-aggregate findings from different studies.
+right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
+is often heaviest to the left of the track, especially when the tropical
+cyclone interacts with other weather systems~\parencite{atallah2003,
+atallah2007, zhu2013variations} or undergoes an extratropical
+transition~\parencite{elsberry2002}.  The multi-hazard nature of tropical
+cyclones therefore makes it hard to assess exposure based on how close the
+storm's central track came to the community.  While other approaches have been
+developed to incorporate storm hazards, particulary wind, into exposure
+assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
+there is not yet a standard approach, and when different studies use different
+datasets or storm hazards when assessing storm exposure, it becomes difficult
+to compare and aggregate findings from different studies.
 
@@
 center over the course of the storm; (3)~cumulative rainfall in the county over
-the course of the storm; (4)~flood events in the county associated with the
-storm; and (5)~tornado events in the county associated with the storm.  We
-aggregated these data at the county level since data on many potential impacts
-are available at county-level aggregations (e.g., direct hurricane-related
-deaths~\parencite{czajkowski2011}; birth outcomes~\parencite{grabich2015,
-grabich2015measuring}; autism prevalence~\parencite{kinney2008}) and since
-decisions and policies to prepare for, and respond to, tropical cyclones are
-often undertaken at the county level~\parencite{zandbergen2009, rappaport2000}.
-We also wrote functions to explore and map this data and to link it
-with county-level health data~\parencite{hurricaneexposuredata}
+the course of the storm; (4)~flooding in the county
+concurrent with the storm; and (5)~tornadoes in the county
+concurrent with the storm.  We aggregated these data at the county level since
+data for epidemiological studies are often available at county-level
+aggregations (e.g., direct hurricane-related deaths~\parencite{czajkowski2011};
+birth outcomes~\parencite{grabich2015, grabich2015measuring}; autism
+prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
+for, and respond to, tropical cyclones are often undertaken at the county
+level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
+explore and map these data and to link it with county-level health
+data~\parencite{hurricaneexposuredata}
 
@@
 measurements compare to other data that could be used to characterize tropical
-cyclone hazards in the context of epidemiological research. Further, we expand
+cyclone hazards for epidemiological research. Further, we expand
 on previous research that investigated methods to measure exposure to tropical
@@
 for exposure assessment~\parencite{grabich2015measuring}.  Here we expand to
-investigate this question across all counties in states in the eastern half of
-the \ac{US} over multiple storms and years.  Further, we investigate patterns
-in exposure classification based on other metrics of storm-related
-hazards---rainfall, flooding, and tornadoes---that are important to
-characterize to improve understanding of the inland health impacts of tropical
-cyclones~\parencite{czajkowski2011, moore2012}. These results can help
-epidemiologists design studies and plan statistical analysis for multi-year,
-multi-community studies that investigate changes in community-wide rates of
-health outcomes during exposure to the hazards brought by tropical cyclones,
-including counties inland from a storm's landfall.
+investigate this question across a large set of counties and storm seasons.
+Further, we investigate patterns in exposure classification based on other
+metrics of storm-related hazards---rainfall, flooding, and tornadoes---that are
+important to characterize to improve understanding of the inland health impacts
+of tropical cyclones~\parencite{czajkowski2011, moore2012}. These results can
+help epidemiologists design studies and plan statistical analysis for
+multi-year, multi-community studies that investigate changes in community-wide
+rates of health outcomes during exposure to the hazards brought by tropical
+cyclones, including counties inland from a storm's landfall.

### 2019-11-12 writing/ehp_methods.tex
Work some more on draft

@@
 We aggregated data for all counties in the eastern half of the \ac{US}
-(Figure~\ref{fig:hurrtracks}) as of the~2010 \ac{US} Decennial Census. We
+(Figure~\ref{fig:hurrtracks}). We
 included all tracked land-falling or near-land Atlantic-basin tropical cyclones
@@
 \parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
-eastern \ac{US} county in this period (Figure~\ref{fig:hurrtracks}).
+eastern \ac{US} county in this period (the tracks of the study storms are shown in 
+Figure~\ref{fig:hurrtracks}).
 
@@
 
-We first measured how close each storm came to each county. We used tracking
-data from \ac{HURDAT2}, which records the storm center's position every~6
-\si{\hour} (at the synoptic times of 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am
-\ac{UTC}), and interpolated this position to~15-\si{\minute} intervals using
-natural cubic splines~\parencite{hurricaneexposure}. At each~15-\si{\minute}
-interval, we measured the distance between the storm's center and each county's
-population mean center~\parencite{bivand2013applied, countycenters}. We took
-the minimum for each county as a measure of how close the storm came to the
+We first measured how close each storm's central track came to each county. We
+used tracking data from \ac{HURDAT2}, which records the storm center's position
+at the synoptic times of 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}, and
+interpolated this position to~15-\si{\minute} intervals using natural cubic
+splines~\parencite{hurricaneexposure}. At each~15-\si{\minute} interval, we
+measured the distance between the storm's center and each county's population
+mean center~\parencite{bivand2013applied, countycenters}. We took the minimum
+distance for each county as a measure of how close the storm came to the
 county.  We also recorded the time the storm came closest to each county, which
@@
 \ac{WONDER} database \parencite{cdcwonder, alhamdan2014environmental}. To
-create this aggregated data originally, these coauthors took hourly
-precipitation measurements in the \ac{NLDAS-2} precipitation files, which are
-given at~1/8\si{\degree} grid points across the continential \ac{US}, and
-summed them by day for each grid point.  This created a daily precipitation
-total for each grid point; these were then averaged for all grid points within
-a county's~1990 \ac{US} Census boundaries~\parencite{alhamdan2014environmental,
-cdcwonder}. This process generated daily county-level precipitation estimates
-for each continental \ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.
-We matched these county-level daily measurements by date with storm tracks,
-using for each county and storm the date when the storm was closest to the
-county. Given the location of storm-affected counties and the typical timing of
-tropical cyclones, these precipitation measures primarily represent rainfall,
-although occasionally they may represent snowfall or other types of
-precipitation.
+create this aggregated data, these coauthors took hourly precipitation
+measurements in the \ac{NLDAS-2} precipitation files, which are given
+at~1/8\si{\degree} grid points across the continential \ac{US}, and summed them
+by day for each grid point.  This created a daily precipitation total for each
+grid point; these were then averaged for all grid points within a county's~1990
+\ac{US} Census boundaries~\parencite{alhamdan2014environmental, cdcwonder}.
+This process generated daily county-level precipitation estimates for each
+continental \ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.  We
+matched these county-level daily measurements by date with storm tracks, using
+the date when the storm was closest to each county. Given the location of
+storm-affected counties and the typical timing of tropical cyclones, these
+precipitation measures primarily represent rainfall, although occasionally they
+may represent snowfall or other types of precipitation.
 
@@
 each county, but could also determine the cumulative rainfall for a more
-extended period, like from three days before to three days after the storm's
+extended period, like from two days before to two days after the storm's
 closest approach. For the analysis of binary exposure classifications in this
@@
 15-\si{\minute} interval estimates, we identified the maximum sustained surface
-wind speed in each county during each storm.
+wind speed in each county over the course of each storm.
 
@@
 quadrant from these maximum wind radii \parencite{knaff2016using}.  We compared
-these wind radii-based estimates of county-level maximum sustained surface
+these wind radii--based estimates of county-level maximum sustained surface
 winds during a storm with the modeled wind speed estimates, comparing all study
 storms since 2004 for which at least one study county experienced sustained
-winds of~$\ge$34 \si{\knot} based on the \ac{HURDAT2} wind radii. For each of
+winds of~$\ge$34 \si{\knot} (based on the \ac{HURDAT2} wind radii). For each of
 these storms, we calculated the percent of study counties that were classified
@@
 To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
-counties, we used event listings from the National Oceanic and Atmospheric
-Administration (NOAA)'s Storm Event Database~\parencite{stormevents}. While
-this database has recorded storm data since~1950, its coverage changed
-substantially in~1996 to cover a more types of storm
-events~\parencite{stormevents}. We therefore only considered these metrics of
-hurricane exposure for tropical cyclones in~1996 and later.
+counties, we matched storm tracks with event listings from the National Oceanic
+and Atmospheric Administration (NOAA)'s Storm Event
+Database~\parencite{stormevents}. While this database has recorded storm data
+since~1950, its coverage changed substantially in~1996 to cover a more types of
+storm events~\parencite{stormevents}. We therefore only considered these
+metrics of hurricane exposure for tropical cyclones in~1996 and later.
 
@@
 county streamflow gages \parencite{usgsgages, countyfloods, dataRetrieval}.  
-
 We considered nine study counties, selecting counties geographically spread
@@
 analysis, but all other study storms were considered. 
-
 For each storm and county, we summed streamflow measurements across all county
@@
 streamflow that exceeded a threshold of flooding (the streamgage's median value
-for annual peak flow~\parencite{countyfloods}) on any day during the five-day
+for annual peak flow) on any day during the five-day
 window. We investigated how these measurements varied between storms with
-associated flood events listed in the NOAA Storm Events data versus storms without an
+associated flood events listings versus storms without an
 event listing for the county, to explore if storms with flood event listings
@@
 
-In our open-source data, we provide continuous measurements of three exposure
-metrics: closest distance of each storm to each county, maximum sustained winds
-at the county center, and cumulative rainfall. However, epidemiologic studies
-of tropical cyclones often use a binary exposure classification (``exposed''
-versus ``unexposed'') to assess storm-related health risks, and so we
-explored patterns in storm exposure based on binary clasifications of
-these exposure metrics.
+In our open-source data, we provide continuous measurements of some of the
+exposure metrics: closest distance of each storm to each county, maximum
+sustained winds at the county center, and cumulative rainfall. However,
+epidemiologic studies of tropical cyclones often use a binary exposure
+classification (``exposed'' versus ``unexposed'') to assess storm-related
+health risks, and so we explored patterns in storm exposure based on binary
+clasifications of these exposure metrics.
 
@@
 rainfall, the cumulative rainfall in the county had to be over the
-75~\si{\milli\metre} threshold and the storm must have passed within~500
-\si{\kilo\metre} of the county; Table~\ref{tab:exposuremetrics}). This distance
-constraint was typically large enough to capture storm-related rain.  However,
-data users should note that in rare cases---for example, exceptionally large
-storms (e.g., Hurricane Ike in 2008) or storms for which storm tracking was
-stopped at extratropical transition (e.g., Tropical Storm Lee in 2011)---some
-storm-related rain exposures may be missed because of this distance constraint
-(Figure~S1). This distance constraint can be customized using the software
-published in association with the open-source
+75~\si{\milli\metre} threshold and the storm must have passed
+within~500 \si{\kilo\metre} of the county; Table~\ref{tab:exposuremetrics}).
+This distance constraint was typically large enough to capture storm-related
+rain.  However, data users should note that in rare cases---for example,
+exceptionally large storms (e.g., Hurricane Ike in 2008) or storms for which
+storm tracking was stopped at extratropical transition (e.g., Tropical Storm
+Lee in 2011)---some storm-related rain exposures may be missed because of this
+distance constraint (Figure~S1). This distance constraint can be customized
+using the software published in association with the open-source
 data~\parencite{hurricaneexposure}.
@@
 Jaccard index ($J_s$)~\parencite{jaccard1901distribution, jaccard1908nouvelles} between
-each pair of exposure metrics, which measures similarity
+each pair of exposure metrics. This measures similarity
 between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical cyclone~$s$ as the
@@
 the two metrics classify exactly the same set of counties as exposed to the
-tropical cyclone. We measured these values for all study storms that affected a
-large number of eastern \ac{US} counties (those storms with 100 or more
-counties classified as exposed by at least one of the exposure metrics)
-during the years when all exposure data were available (1996\,--\,2011).
+tropical cyclone. We calculated these values for all study storms that affected
+$\ge$100 study counties (based on at least one exposure metric) during the
+years when all exposure data were available (1996\,--\,2011).

### 2019-11-12 writing/ehp_results.tex
Work some more on draft

@@
 from other available sources.  For estimates of storm-associated rainfall, we
-compared estimates used in the open-source data with ground-based observations
-in nine sample counties (Figure~\ref{fig:raincomparison}). Within these
-counties, storm-related rainfall measurements were well-correlated between the
-two data sources, with rank correlations (bottom right of each graph in
-Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some evidence
-that the reanalysis data included in the open-source dataset may tend to
-underestimate rainfall totals in storms with extremely high rainfall, based on
-a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
-County, SC, and Wake County, NC (Figure~\ref{fig:raincomparison}). However,
-even in these cases, it was rare for a storm to be classified differently
-(under the precipitation threshold of 75~\si{\milli\metre} we used for binary
-exposure classification for later analysis) based on the source of
-precipitation data---horizontal and vertical lines in each small plot in
-Figure~\ref{fig:raincomparison} show the threshold of 75~\si{\milli\metre}, so
-storms in the lower left and upper right quadrants would be classified the same
-(``exposed'' or ``unexposed'') regardless of the precipitation data source,
-while storms in the upper left and lower right quadrants would be classified
-differently. Such cases are rare.
+compared estimates included in the open-source data with ground-based
+observations in nine sample counties (Figure~\ref{fig:raincomparison}). Within
+these counties, storm-related rainfall measurements were well-correlated
+between the two data sources, with rank correlations (bottom right of each
+graph in Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some
+evidence that the primary rainfall metric may tend to underestimate rainfall
+totals in storms with extremely high rainfall, based on a few heavy-rainfall
+storms in Harris County, TX, Mobile County, AL, Charleston County, SC, and Wake
+County, NC (Figure~\ref{fig:raincomparison}). However, even in these cases, it
+was rare for a storm to be classified differently (under the precipitation
+threshold of 75~\si{\milli\metre} we used for binary exposure classification
+for later analysis) based on the source of precipitation data---horizontal and
+vertical lines in each small plot in Figure~\ref{fig:raincomparison} show the
+threshold of 75~\si{\milli\metre}, so storms in the lower left and upper right
+quadrants would be classified the same (``exposed'' or ``unexposed'')
+regardless of the precipitation data source, while storms in the upper left and
+lower right quadrants would be classified differently. Such cases were rare.
 
@@
 was wide variation in the average number of county exposures per year
-(Table~\ref{tab:exposuresummaries}). For the tropical cyclone tornado exposure,
+(Table~\ref{tab:exposuresummaries}). For tropical cyclone tornadoes,
 there were on average about 50 county exposures per year within our study.
@@
 exposure was also more frequent in coastal areas compared to inland areas,
-there were also inland exposures in counties that were never classified as
+there were also inland rain exposures in counties that were never classified as
 exposed based on winds. Flood-based exposures were frequently in the
 Mid-Atlantic region, with a pattern that skewed north compared with other
-exposures.Rain and, to some extent, flood exposures were characterized by a
+exposures. Rain and, to some extent, flood exposures were characterized by a
 pattern defined by the Appalachian Mountains, with fewer exposures west of the
@@
 Next, we measured the agreement between exposure classifications for each pair
-of hazards, as well as between each hazard and exposure assessment based on
+of hazards, as well as between with a proxy measurement based on
 distance from the storm's track. We found that these exposure classifications
-typically did not strongly agree between pairs of metrics---the set of counties
+typically did not agree strongly between pairs of metrics---the set of counties
 identified as ``exposed'' based on one metric often overlapped little with the set
@@
 metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
-for one tropical cyclone, and each box in that row shows the measured Jaccard
+for one tropical cyclone, and each box in that row shows the Jaccard
 coefficient for a pair of metrics.  For almost all tropical cyclones, Jaccard
-index values between distance and any of the four hazard-based metrics
-were~$\le0.6$ (i.e., out of the counties assessed as exposed by at least one of
-the two metrics in the pair,~60\si{\percent} or fewer were assessed the same
-under both metrics in the pair). For many tropical cyclones, the Jaccard index
-values were~$\le0.5$, indicating at best moderate agreement.  Among the four
-hazard-based metrics, the tornado-based metric had universally poor agreement
-with other metrics in county-level classification across the tropical cyclones
-considered.  Agreement between other pairs of metrics were better, but still
-typically at best moderate (Jaccard index values usually $\le0.5$).
+coefficients were~$\le0.6$ between the distance-based metric and any of the
+four hazard-based metrics (i.e., out of the counties assessed as exposed by at
+least one of the two metrics in the pair,~60\si{\percent} or fewer were
+assessed the same under both metrics). For many tropical cyclones, Jaccard
+coefficients were~$\le0.5$, indicating at best moderate agreement.  Among the
+four hazard-based metrics, the tornado-based metric had universally poor
+agreement with other metrics in county-level classification across the tropical
+cyclones considered.  Agreement between other pairs of metrics were better, but
+still typically at best moderate (Jaccard coefficients usually $\le0.5$).
 
@@
 in~2003), there was moderate to good agreement for all pairwise combinations of
-distance, rain, and wind, but poor agreement for other combinations of
-metrics, while for another set of storms (e.g., Matthew in~2004 and Katrina
-in~2005), there was moderate to good agreement between distance and rain.
-However, these sets of tropical cyclones represented the minority of tropical
-cyclones considered; for most tropical cyclones, there was low overlap in the
-counties determined to be exposed to the tropical cyclones between most
+distance, rain, and wind, but poor agreement for other combinations of metrics,
+while for another set of storms (e.g., Matthew in~2004 and Katrina in~2005),
+there was moderate to good agreement between distance and rain.  However, these
+sets of tropical cyclones represented the minority of tropical cyclones
+considered; for most tropical cyclones, there was low to moderate overlap in
+the counties determined to be exposed to the tropical cyclones between most
 pairings of metrics.

### 2019-11-12 writing/ehp_abstract.tex
Work some more on abstract and small wording

@@
         Tropical cyclone epidemiology can be 
-	advanced through methods to assess 
-	county-level exposure to storm hazards, particularly 
-	methods that are comprehensive and 
+	advanced through county-level exposure assessment methods that are comprehensive and 
 	consistent across space and time, allowing multi-year, multi-storm studies. 
@@
 	designing tropical cyclone epidemiological studies.\\ 
-\textbf{Objectives:} (1)~Provide an open-source dataset to improve tropical
+\textbf{Objectives:} (1)~Provide an open-source dataset for tropical
 	cyclone exposure assessment for epidemiological research and
@@
 	Atlantic basin storms for~1996\,--\,2011 for all
-	metrics (up to~1988\,--\,2015 for specific metrics). 	
+	metrics and up to~1988\,--\,2015 for specific metrics. 	
 	We validated 
@@
 	and agreement among binary exposure classifications based on these
-	metrics. We also compared exposure assessment
-	based on using distance between a county and the storm's track as a proxy for
-	storm exposure.\\ 
-\textbf{Results:} Measurements in the open-source dataset created here typically
+	metrics, as well as between distance from the storm's track as a proxy for
+	exposure.\\ 
+\textbf{Results:} Measurements in the dataset created here typically
         agreed with data from other sources, and we present and discuss areas of
-	disagreement and other caveats for using the data.
-        Long-term geographic patterns in binary 
-	classifications based on these data agree with previous results from
+	disagreement and other caveats for the data.
+        Long-term geographic patterns in exposure classification agree with previous results from
 	atmospheric science. County-level exposure classification for a storm differs substantially 
@@
 \textbf{Discussion:} Our results provide a multi-hazard dataset that can be
-	leveraged for epidemiologic research on tropical cyclones.  Further,	
-	our results suggest that the use of distance as a
-	surrogate for tropical cyclone exposure, or use one hazard-based metric
-	when the impact is partly or fully caused by a different hazard, would often 
-	result in exposure misclassification in epidemiological studies.  
+	leveraged for epidemiologic research on tropical cyclones, as well as 
+	insights that can inform the design of tropical cyclone epidemiological 
+	studies.
 \end{abstract}

### 2019-11-12 writing/ehp_discussion.tex
Work some more on abstract and small wording

@@
 Our results can inform exposure assessment for future county-level studies of
-the health risks and impacts associated with tropical cyclones exposures as
-well as insights to help in designing studies.
+the health risks and impacts associated with tropical cyclones exposure as
+well as insights to inform epidemiological study design.
 
@@
 and time restrictions, like those we used in matching NOAA Storm Event listings
-with tropical storm locations and dates, do create the opportunity for
+with tropical cyclone locations and dates, do create the opportunity for
 occasional false negatives, as for Hurricane Ida in Fulton County, GA, where a
@@
 We found that tropical cyclones tended to bring different hazards to different
-counties and also that agreement was typically low between distance-based
+counties, and so agreement was typically low between distance-based
 tropical cyclone exposure assessment and each of the single hazard-specific
@@
 dangerous storm surge, which is not specifically covered in this dataset
-(although some resulting coastal flooding is captured). We are exploring ways to
-include this in future versions of the dataset, although to date we have
+(although some resulting coastal flooding is captured). We are exploring ways
+to include this in future versions of the dataset, although to date we have
 focused on exposures that could affect any county, whether inland or coastal.
@@
 datasets are aggregated at this level, but some may be aggregated at a finer
-spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e., point locations
-for each outcome). We have published the wind
-model we used to create this dataset as its own open-source R package
-\parencite{stormwindmodel}, and it can be used to model
-storm-associated winds at a finer spatial resolution; however, measurements of
-other hazards cannot similarly be re-scaled through tools we provide. Next,
-we provide these data and associated software tools through R packages, and so
-some experience in the R programming language is required to make full use of
-them. However, R is currently a popular programming language for environmental
-epidemiology, allowing the data to reach a large audience, and we are exploring
-options to create a web application using the Shiny platform to allow more
-user-friendly web-based access of the data \parencite{shiny2019}. 
-Finally, to assess patterns and agreement for binary exposure classifications, 
-we have chosen one set of sensible thresholds for binary classifications based
-on continuous metrics (rainfall, maximum sustained winds, and distance from 
-the storm's track). Results and conclusions would differ somewhat with other
-threshold choices. We have published all our code for this analysis online
-[web address], in case other researchers would like to explore other 
-threshold choices for these analyses.
+spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e.,
+point locations for each outcome). We have published the wind model used to
+create this dataset as its own open-source R package
+\parencite{stormwindmodel}, and it can be used to model storm-associated winds
+at a finer spatial resolution; however, measurements of other hazards cannot
+similarly be re-scaled through tools we provide. Next, we provide these data
+and associated software tools through R packages, and so some experience in the
+R programming language is required to make full use of them. However, R is
+currently a popular programming language for environmental epidemiology,
+allowing the data to reach a large audience, and we are exploring options to
+create a web application using the Shiny platform to allow broader web-based
+access of the data \parencite{shiny2019}.  Finally, to assess patterns and
+agreement for binary exposure classifications, we have chosen one set of
+sensible thresholds for binary classifications based on continuous metrics
+(rainfall, maximum sustained winds, and distance from the storm's track).
+Results and conclusions would differ somewhat with other threshold choices. We
+have published code for this analysis online, allowing
+other researchers can explore other threshold choices for these
+analyses.
 
@@
 leverage existing administrative health data, allowing researchers to
-investigate how these storms affect county-wide health risk, including for
-outcomes also common outside of storm periods.  Further, this dataset provides
-hazard measurements that are comparable across communities and storms, which
-allows epidemiological researchers to design studies to explore how health
-risks are modified by characteristics of both the storms and the communities
-that are hit. The data are given in an open-source format, along with
-associated software tools, which allows them to be freely used and for others
-to explore all associated code and to contribute additions through platforms
-like GitHub.
+investigate how these storms affect county-wide health risk.  Further, this
+dataset provides hazard measurements that are comparable across communities and
+storms, which allows epidemiological researchers to design studies to explore
+how health risks are modified by characteristics of both the storms and the
+communities that are hit. The data are given in an open-source format, along
+with associated software tools, which allows them to be freely used and for
+others to explore all associated code and to contribute additions through
+platforms like GitHub.
 
@@
 they are studying, and to characterize storm exposure in a way that captures
-those specific hazards.
+those specific hazards, to avoid exposure misclassification.

### 2019-11-12 writing/ehp_introduction.tex
Work some more on abstract and small wording

@@
 Hurricanes and other tropical cyclones can severely impact human health in
-\ac{US} communities, and some of these impacts have been characterized through
-multi-year studies based on fatality data from post-disaster
-surveillance~\parencite{rappaport2000, rappaport2014fatalities,
-rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.  For
-several types of climate-related disasters---including heat waves, floods, and
-wildfires---research based on post-disaster surveillance data has been richly
-supplemented by studies analyzing trends in community-wide administrative or
-vital data, including time series or case-crossover studies
-(e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
-liu2017wildfire, milojevic2017mental}).  These studies typically compare the
-rate of a health outcome in a community during a disaster to rates during
-comparable non-disaster periods.  While these studies cannot attribute specific
-cases (e.g., a specific person's death) to a disaster, they can help
-characterize the community-wide change in health risk during disasters and
-capture impacts that might be missed or underestimated in post-disaster
+\ac{US} communities, impacts that have been characterized based on fatality
+data from post-disaster surveillance~\parencite{rappaport2000,
+rappaport2014fatalities, rappaport2016fatalities, czajkowski2010fatal,
+czajkowski2011, moore2012}.  For several types of climate-related
+disasters---including heat waves, floods, and wildfires---research based on
+post-disaster surveillance data has been richly supplemented by studies
+analyzing trends in community-wide administrative or vital data, including time
+series or case-crossover studies (e.g.,~\cite{anderson2010heat, son2012impact,
+haikerwal2015impact, liu2017wildfire, milojevic2017mental}).  These studies
+typically compare the rate of a health outcome in a community during a disaster
+to rates during comparable non-disaster periods.  While such studies cannot
+attribute specific cases (e.g., a specific death) to a disaster, they
+can help characterize the community-wide change in health risk during disasters
+and capture impacts that might be missed or underestimated in post-disaster
 surveillance.  In many cases, these studies analyze multi-year, multi-community
@@
 explore how a disaster's characteristics, or the characteristics of the
-community it hits, modify associated health risks (e.g.,~\cite{anderson2010heat,
-son2012impact, liu2017wildfire}).  Some studies have begun to use this approach
-to study the health impacts of tropical cyclones, including several studies of
-Hurricane Maria (e.g.,~\cite{santos2018use, santos2018differential}), Hurricane
-Sandy (e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane
-season in Florida~\parencite{mckinney2011}.  However, to expand this approach
-to study longer time periods for large sets of communities, researchers must be
-able to assess exposure to tropical cyclones consistently and comparably across
-storms, years, and communities.
+community it hits, modify associated health risks
+(e.g.,~\cite{anderson2010heat, son2012impact, liu2017wildfire}).  Some studies
+have begun to use this approach to study the health impacts of tropical
+cyclones, including several studies of Hurricane Maria
+(e.g.,~\cite{santos2018use, santos2018differential}), Hurricane Sandy
+(e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane season
+in Florida~\parencite{mckinney2011}.  However, to expand this approach to study
+longer time periods for large sets of communities, researchers must be able to
+assess exposure to tropical cyclones consistently and comparably across storms,
+years, and communities.
 
@@
 However, when epidemiological studies assess exposure to a tropical cyclone
-using, as a proxy measurement, how close the storm came to a community, they
+using, as a proxy, how close the storm came to a community, they
 may misclassify exposure~\parencite{grabich2015measuring}. While a number of

### 2019-11-12 writing/ehp_results.tex
Work some more on abstract and small wording

@@
 For maximum sustained surface wind estimates, we found that the primary wind
-exposure metric in the open source data generally agreed well with the wind
+exposure metric in the open-source data generally agreed well with the wind
 radii reported in \ac{HURDAT2}. For most storms, $\ge$95\% of counties were
@@
 Next, we measured the agreement between exposure classifications for each pair
-of hazards, as well as between with a proxy measurement based on
+of hazards, as well as with a proxy measurement based on
 distance from the storm's track. We found that these exposure classifications

### 2019-11-28 writing/anderson_ehp_supp.tex
Update figure references
Work on paper, use 34 knot threshold
Continue work on text

@@
   State University, Fort Collins, CO, 80523} 
-\affil[2]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
+\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, 80045} 
 \affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
@@
 \affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
-\affil[7]{Debian and R Projects; Department of Statistics, University of
+\affil[7]{Department of Statistics, University of
   Illinois at Urbana-Champaign, Champaign, IL, 61820} 
@@
 
+\clearpage
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=\linewidth]{windexamples}
+\caption{Comparison of county-level estimates of peak sustained surface wind 
+	for Hurricane Ike in 2008 (top) and Hurriance Sandy in 2012 (bottom). 
+	Each map shows the estimated peak sustained surface wind 
+	classification ($<$34~kt; 34\,--\,49.9~kt; 
+	50\,--\,63.9~kt; $\ge$64~kt) for each study county. 
+	The maps labelled ``Modeled'' (left)
+	shows the classifications based on modeled peak sustained surface wind,
+	which were included in the open-source data as the main 
+	wind metric and used in further analysis in this research. The maps
+	labelled ``Extended Best Tracks'' (right) show classifications based on the 
+	wind radii given in \ac{HURDAT2} (included as a secondary wind 
+	metric in the open-source data). The red lines show the storms' 
+	tracks.
+	}
+\label{fig:windexamples}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=7cm]{averageexposureonlysupp}
+\caption{Average number of storm exposures per decade in U.S. counties for each 
+	single-hazard exposure metric, limited analysis to years for which data on all five 
+	exposures were available (1996\,--\,2011). The criteria behind each of 
+	the five metrics is given in Table 1 of the main text. 	
+	} 
+\label{fig:averageexposuresupp}
+\end{figure}
+
+\clearpage
+
 \begin{figure}[tbhp!]
@@
 \clearpage
+
+\begin{comment}
 
@@
 
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{averageexposureonlysupp}
-\caption{Average number of storm exposures per decade in U.S. counties for each 
-	exposure metric, limited analysis to years for which data on all five 
-	exposures were available (1996\,--\,2011). The criteria behind each of 
-	the five metrics is given in Table 1 of the main text. 	
-	} 
-\label{fig:averageexposuresupp}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[tbhp!]
-\centering
-\includegraphics[width=\linewidth]{windexamples}
-\caption{Comparison of county-level estimates of maximum sustained surface wind speeds
-	for Hurricane Ike in 2008 (top) and Hurriance Sandy in 2012 (bottom). 
-	Each map shows the estimated maximum sustained surface wind speed 
-	classification ($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot}; 
-	50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) for each study county. 
-	The maps labelled ``Modeled'' (left)
-	shows the classifications based on modeled sustained surface wind
-	speeds, which were included in the open-source data as the main 
-	wind metric and used in further analysis in this research. The maps
-	labelled ``Extended Best Tracks'' (right) show classifications based on the 
-	wind radii given in \ac{HURDAT2} (included as a secondary wind 
-	metric in the open-source data). The red lines show the storms' 
-	tracks.
-	}
-\label{fig:windexamples}
-\end{figure}
-
-\clearpage
+\end{comment}
 

### 2019-11-28 writing/ehp_abstract.tex
Continue work on text

@@
 	advanced through county-level exposure assessment methods that are comprehensive and 
-	consistent across space and time, allowing multi-year, multi-storm studies. 
+	consistent across space and time, facilitating multi-year, multi-storm studies. 
 	Further, an understanding of 
-	patterns in and between hazard-specific exposure metrics can help in 
+	patterns in and between exposure metrics that are based on specific hazards of the storm can help in 
 	designing tropical cyclone epidemiological studies.\\ 
@@
 \textbf{Methods:} We created an open-source dataset with county-level data on
-	exposure to four tropical cyclone hazards: maximum sustained winds,
+	exposure to four tropical cyclone hazards: peak sustained wind,
 	rainfall, flooding, and tornadoes. The data cover all eastern \ac{US} counties 
 	for all land-falling or near-land
-	Atlantic basin storms for~1996\,--\,2011 for all
+	Atlantic basin storms, covering~1996\,--\,2011 for all
 	metrics and up to~1988\,--\,2015 for specific metrics. 	
@@
 	and agreement among binary exposure classifications based on these
-	metrics, as well as between distance from the storm's track as a proxy for
+	metrics, as well as compared them to use of distance from the storm's track as a proxy for
 	exposure.\\ 
-\textbf{Results:} Measurements in the dataset created here typically
-        agreed with data from other sources, and we present and discuss areas of
-	disagreement and other caveats for the data.
-        Long-term geographic patterns in exposure classification agree with previous results from
-	atmospheric science. County-level exposure classification for a storm differs substantially 
-	depending on the hazard metric used. \\ 
+\textbf{Results:} We describe how we developed this dataset and explore how its
+	measurements compare to other data that could be used to characterize
+	tropical cyclone hazards for epidemiological research. Measurements in
+	the dataset created here typically agreed with data from other sources,
+	and we present and discuss areas of disagreement and other caveats for
+	the data. Tropical cyclones typically brought different hazards to different
+	counties, and so agreement was usually low between distance-based tropical
+	cyclone exposure assessment and each hazard-specific metric, as well as 
+	between pairs of hazard-specific metrics. \\
 \textbf{Discussion:} Our results provide a multi-hazard dataset that can be

### 2019-11-28 writing/ehp_cover.tex
Rearrange content
Continue work on text

@@
 
-\author[1,*]{G. Brooke Anderson} 
+\author[1,*]{G.~Brooke Anderson} 
 \author[1,2]{Joshua Ferreri}
@@
 \author[1]{Meilin Yan} 
-\author[8]{Roger D. Peng}
+\author[8]{Roger D.~Peng}
 
@@
   Health, Baltimore, MD, 21205}
-\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu,
+\affil[*]{Corresponding author: G.~Brooke Anderson, brooke.anderson@colostate.edu,
   1681 Campus Delivery, Fort Collins, CO, 80523}

### 2019-11-28 writing/ehp_discussion.tex
Continue work on text

@@
 conduct multi-community, multi-year studies leveraging large administrative
-datasets.  Here, we provide open-source county-level data on multiple tropical
-cyclone exposure metrics and explore limitations in that data.  
+datasets.  Here, we provide open-source county-level data for several tropical
+cyclone exposures and explore limitations in that data.  
 Further, we explore patterns in storm exposure classifications based on
@@
 
-We investigated how well the open-source data correspond with data from other
-potential sources
-(Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison} and~S2). While
-generally in agreement, there are some caveats. The rainfall data are generally
+The open-source data generally correspond well with data from other potential
+sources (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison}
+and~S2), but there are some caveats. The rainfall data are generally
 well-correlated with ground-based observations, but may sometimes oversmooth
-very high rainfall values compared to ground-based observations
-(Figure~\ref{fig:raincomparison}). When rainfall data is used to create binary
-exposure classifications, this disagreement is unlikely to influence results,
-as both data sources agree in identifying these as storms with high rainfall,
-but would be important to consider for cases that include rainfall as a
-continuous measurement.
+very high rainfall values (Figure~\ref{fig:raincomparison}). When rainfall data
+is used to create binary exposure classifications, this disagreement is
+unlikely to influence results, as both data sources agree in identifying these
+as storms with high rainfall, but would be important to consider for cases that
+include rainfall as a continuous measurement.
 
-The wind data are based on modeled, rather than observed, values, and while the
-modeled wind data generally agree well with post-analysis maximum wind radii
-from \ac{HURDAT2}~\parencite{landsea2013} (Figure~\ref{fig:windcomparison}),
-there were a few storms with some discrepancies. These storms---for example, Hurricane Sandy
-in 2012 and Hurricane Ike in 2008---were
-unusually large systems for which high winds persisted well inland from
-landfall (Figures~\ref{fig:windcomparison}
-and~S2).
+The peak sustained wind estimates are based on modeled, rather than observed,
+values, and while the modeled wind data generally agree well with post-analysis
+maximum wind radii from \ac{HURDAT2}~\parencite{landsea2013}
+(Figure~\ref{fig:windcomparison}), there were a few storms with some
+discrepancies. These storms---for example, Hurricane Sandy in 2012 and
+Hurricane Ike in 2008---were unusually large systems for which high winds
+persisted well inland from landfall (Figures~\ref{fig:windcomparison} and~S2).
 
-For the flooding data, we found that flood event status determined based on the
-NOAA Storm Events listings typically agreed with measurements from \ac{USGS}
-streamgages, with a flood event more likely to be listed if a storm elevated
-streamflow across the county (Figure~\ref{fig:floodcomparison}).  However,
-there are differences between the two flooding datasets, and these highlight
-both the difficulty of measuring flood exposure at the county level and
-inherent challenges in using data from a storm event database for epidemiologic
-exposure assessment.  For example, there was one storm in Fulton County, GA,
-for which there was high streamflow but not an associated flood event listing
-(Hurricane Ida, 2009).  This storm occurred in November 2009, following a month
-with historic rainfall and flooding in
+For the flooding data, we found that flood event status, as determined based on
+the NOAA Storm Events listings, typically agreed with measurements from
+\ac{USGS} streamgages, with a flood event more likely to be listed if a storm
+elevated streamflow at streamgages across the county
+(Figure~\ref{fig:floodcomparison}).  However, there are differences between the
+two flooding datasets, and these highlight both the difficulty of measuring
+flood exposure at the county level and inherent challenges in using data from a
+storm event database for epidemiologic exposure assessment.  For example, there
+was one storm in Fulton County, GA, for which there was high streamflow but not
+an associated flood event listing (Hurricane Ida, 2009).  This storm occurred
+in November 2009, following a month with historic rainfall and flooding in
 Georgia~\parencite{shepherd2011overview}.  In this case, the flooding
@@
 hazard metrics~\parencite{chakraborty2005population, peduzzi2009assessing}.
-This ability can be critical, as different hazards of tropical cyclones can
+This functionality can be critical, as different hazards of tropical cyclones can
 act synergistically in causing impacts~\parencite{smith2009}.
@@
 exposures, as well as between pairs of hazard-specific metrics
-(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard} and S4\,--\,S5).  These
+(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard} and S4).  These
 findings align with previous results from atmospheric science and related
@@
 
-Therefore, the use of a distance-based metric to assess exposure to any of
+Based on our results, the use of a distance-based metric to assess exposure to any of
 these hazards, or the use of measurements from one hazard as a proxy for
@@
 
-We did, however, find a small set of tropical cyclones for which for
-which agreement was high across several single-hazard exposure assessments
-(e.g., Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto
-in~2006 (Figure~\ref{fig:jaccard})).  Hurricanes Floyd in~1999 and
-Irene in~2011 both made their first \ac{US} landfall in North Carolina at minor
-hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
-eastern coast of the \ac{US} north through New England, bringing substantial
-rainfall to much of the eastern coast from North Carolina north and causing
-extensive inland flooding in North Carolina (Floyd) and New England
+We did, however, find a small set of tropical cyclones for which for which
+agreement was high across several single-hazard exposure assessments (e.g.,
+Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto in~2006
+(Figure~\ref{fig:jaccard})).  Hurricanes Floyd in~1999 and Irene in~2011 both
+made their first \ac{US} landfall in North Carolina at minor hurricane
+windspeeds (Category~2 and~1, respectively) and then skimmed the eastern coast
+of the \ac{US} north through New England, bringing substantial rainfall to much
+of the eastern coast from North Carolina north and causing extensive inland
+flooding in North Carolina (Floyd) and New England
 (Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
 in~2008, Bertha in~1996, and Ernesto in~2006 also followed the eastern
-coastline. For these storms, the tropical cyclones' persistent proximity
-to water may have helped maintain wind speeds in similar patterns to rain and
+coastline. For these storms, the tropical cyclones' persistent proximity to
+water may have helped maintain wind speeds in similar patterns to rain and
 distance-based exposure assessments, resulting in more similarities across
-exposure assessments compared to the other tropical cyclones we considered.
-For these storms, it may be possible to assess exposure to multiple hazards of
-the storm using a single metric, in some cases even a proxy like the distance
-between a county and the storm's track.  However, for these storms it may be
-difficult to untangle the contribution of each hazard to the overall
-effect of the storm, given that several hazards have similar geographical
-patterns.
+exposure assessments compared to other tropical cyclones.  For these storms, it
+may be possible to assess exposure to multiple hazards of the storm using a
+single metric, in some cases even a proxy like the distance between the county
+and the storm's track.  However, for these storms it may be difficult to
+untangle the contribution of each hazard to the overall effect of the storm,
+given that several hazards have similar geographical patterns.
 
@@
 caveats previously discussed. First, the dataset is not comprehensive of all
-possible tropical cyclone hazards. For example, coastal counties can experience
+tropical cyclone hazards. For example, coastal counties can experience
 dangerous storm surge, which is not specifically covered in this dataset
 (although some resulting coastal flooding is captured). We are exploring ways
-to include this in future versions of the dataset, although to date we have
+to include this in future versions of the dataset, but to date we have
 focused on exposures that could affect any county, whether inland or coastal.
@@
 sensible thresholds for binary classifications based on continuous metrics
-(rainfall, maximum sustained winds, and distance from the storm's track).
+(rainfall, maximum sustained surface wind, and distance from the storm's track).
 Results and conclusions would differ somewhat with other threshold choices. We
 have published code for this analysis online, allowing
-other researchers can explore other threshold choices for these
+other researchers to explore other threshold choices for these
 analyses.
@@
 dataset provides hazard measurements that are comparable across communities and
-storms, which allows epidemiological researchers to design studies to explore
+storms, allowing epidemiological researchers to design studies to explore
 how health risks are modified by characteristics of both the storms and the
@@
 should explore the analyses presented in this paper to understand the strengths
-and weaknesses of the data.  Our results suggest that county-level storm
+and weaknesses of the data.  Further, our results indicate that county-level storm
 exposure is not well-characterized by the closest distance that a storm's

### 2019-11-28 writing/ehp_figures.tex
Work some more on abstract and small wording
Continue work on text

@@
 	\caption{Comparison of two sources of wind exposure estimates: 
-	(1) modeled maximum sustained surface wind speeds 
-	and (2) estimates from \ac{HURDAT2}'s wind speed radii.
+	(1) modeled peak sustained surface wind 
+	and (2) estimates based on \ac{HURDAT2}'s wind speed radii.
 	Each point represents a storm, with the x axis giving the percent of
-	counties classified in the same category of maximum sustained surface
-	wind speeds ($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot};
-	50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) by both sources of 
+	counties classified in the same category of peak sustained surface
+	wind ($<$34~kt; 34\,--\,49.9~kt;
+	50\,--\,63.9~kt; $\ge$64~kt) by both sources of 
 	data. The color of each point gives the number of study counties that were exposed to
-	maximum sustained surface wind speeds of at least 34~\si{\knot} (based
-	on modeled wind speeds). Estimates are shown for study storms since
+	peak sustained surface wind of at least 34~kt (based
+	on modeled wind). Estimates are shown for study storms since
 	2004, the earliest year for which post-storm reanalysis wind speed
 	radii estimates are routinely available in \ac{HURDAT2}, and for which at
-	least one study county had a sustained wind speed of $\ge$34~\si{\knot}
+	least one study county had a peak sustained wind of $\ge$34~\si{\knot}
 	based on the \ac{HURDAT2} wind radii.
@@
 	within clusters that have similar patterns based on hierarchical 
-	clustering (complete link method).}
+	clustering.}
 \label{fig:jaccard} 

### 2019-11-28 writing/ehp_introduction.tex
Continue work on text

@@
 Hurricanes and other tropical cyclones can severely impact human health in
-\ac{US} communities, impacts that have been characterized based on fatality
-data from post-disaster surveillance~\parencite{rappaport2000,
-rappaport2014fatalities, rappaport2016fatalities, czajkowski2010fatal,
-czajkowski2011, moore2012}.  For several types of climate-related
-disasters---including heat waves, floods, and wildfires---research based on
-post-disaster surveillance data has been richly supplemented by studies
-analyzing trends in community-wide administrative or vital data, including time
-series or case-crossover studies (e.g.,~\cite{anderson2010heat, son2012impact,
-haikerwal2015impact, liu2017wildfire, milojevic2017mental}).  These studies
-typically compare the rate of a health outcome in a community during a disaster
-to rates during comparable non-disaster periods.  While such studies cannot
-attribute specific cases (e.g., a specific death) to a disaster, they
-can help characterize the community-wide change in health risk during disasters
-and capture impacts that might be missed or underestimated in post-disaster
+\ac{US} communities, as has been characterized in several long-term national
+studies~\parencite{rappaport2000, rappaport2014fatalities,
+rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.
+These studies characterize tropical cyclone health impacts using fatality data
+aggregated in large part from disaster-related mortality ascertainment and
+surveillance conducted by agencies like the \ac{US} Centers for Disease Control
+and Prevention, the \ac{US} National Weather Service, and local vital
+statistics departments. These fatality data are based on case-by-case
+ascertainment: identifying and characterizing deaths for which there is a clear
+link with a disaster, through death certificate coding or other indicators
+\parencite{rocha2017medicolegal}.
+
+For several types of climate-related disasters---including heat waves, floods,
+and wildfires---such research has been richly supplemented by studies that
+estimate community-wide excess mortality and morbidity associated with the
+disaster (e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
+liu2017wildfire, milojevic2017mental}).  These studies, which often use time
+series or case-crossover study designs, typically compare the rate of a health
+outcome in a community during a disaster to rates during comparable
+non-disaster periods.  While such studies cannot attribute specific cases
+(e.g., a specific death) to a disaster, they can quantify the community-wide
+change in health risk during disasters and capture impacts that might be missed
+or underestimated with traditional disaster-related mortality ascertainment and
 surveillance.  In many cases, these studies analyze multi-year, multi-community
@@
 (e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane season
-in Florida~\parencite{mckinney2011}.  However, to expand this approach to study
-longer time periods for large sets of communities, researchers must be able to
+in Florida~\parencite{mckinney2011}.  However, to expand this approach to
+longer time periods and large sets of communities, researchers must be able to
 assess exposure to tropical cyclones consistently and comparably across storms,
@@
 position of a tropical cyclone every six hours, as well as the storm's minimum
-central pressure and maximum wind speed.  These data are openly available
-through the \ac{HURDAT2}, a post-storm assessment that incorporates data from
-several sources, including satellite data and, when available, aircraft
-reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these data, it
-is straightforward to measure whether a community was in the direct path, or
-within a certain distance of the central track, of a storm, as some previous
-epidemiological studies have done to assess exposure to a tropical storm
+central pressure and maximum sustained surface wind.  These data are openly
+available through the \ac{HURDAT2}, a post-storm assessment that incorporates
+data from several sources, including satellite data and, when available,
+aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these
+data, it is straightforward to measure whether a community was in a storm's
+direct path or within a certain distance of its central track; some previous
+epidemiological studies have done this to assess exposure to a tropical storm
 (e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}).
@@
 datasets or storm hazards when assessing storm exposure, it becomes difficult
-to compare and aggregate findings from different studies.
+to compare and aggregate findings.
 
 To help epidemiological researchers assess exposure to tropical cyclones, we
-developed an open-source dataset~\parencite{hurricaneexposure}.  These data
-cover all counties in states in the eastern half of the \ac{US} over multiple
-years and include five metrics characterizing exposure to tropical cyclones
-(Table~\ref{tab:exposuremetrics}): (1)~closest distance the storm's central
-track came to the county's center (a proxy measurement of storm exposure used
-in some previous studies); (2)~maximum sustained wind speed at the county's
-center over the course of the storm; (3)~cumulative rainfall in the county over
-the course of the storm; (4)~flooding in the county
-concurrent with the storm; and (5)~tornadoes in the county
-concurrent with the storm.  We aggregated these data at the county level since
-data for epidemiological studies are often available at county-level
+developed an open-source dataset~\parencite{hurricaneexposure}, which we
+present here.  These data cover all counties in states in the eastern half of
+the \ac{US} over multiple years and include five metrics characterizing
+exposure to tropical cyclones (Table~\ref{tab:exposuremetrics}): (1)~closest
+distance the storm's central track came to the county's center (a proxy
+measurement of storm exposure used in some previous studies); (2)~peak
+sustained surface wind at the county's center over the course of the storm;
+(3)~cumulative rainfall in the county over the course of the storm;
+(4)~flooding in the county concurrent with the storm; and (5)~tornadoes in the
+county concurrent with the storm.  We aggregated these data at the county level
+since data for epidemiological studies are often available at county-level
 aggregations (e.g., direct hurricane-related deaths~\parencite{czajkowski2011};
@@
 explore and map these data and to link it with county-level health
-data~\parencite{hurricaneexposuredata}
+data~\parencite{hurricaneexposuredata}.
 
@@
 measurements compare to other data that could be used to characterize tropical
-cyclone hazards for epidemiological research. Further, we expand
-on previous research that investigated methods to measure exposure to tropical
-cyclones for epidemiological research~\parencite{grabich2015measuring}.  The
-previous study investigated differences in which communities were assessed as
-exposed to tropical cyclones during the 2004 hurricane season in Florida,
-depending on whether exposure assessment was based on distance to the storm's
-track or used a metric incorporating storm-related winds within the county
-during the storm.  They found important differences across methods of assessing
-storm exposure, concluding that a study may be prone to bias from exposure
-misclassification if distance to the storm track is used as a proxy measurement
-for exposure assessment~\parencite{grabich2015measuring}.  Here we expand to
-investigate this question across a large set of counties and storm seasons.
-Further, we investigate patterns in exposure classification based on other
-metrics of storm-related hazards---rainfall, flooding, and tornadoes---that are
-important to characterize to improve understanding of the inland health impacts
-of tropical cyclones~\parencite{czajkowski2011, moore2012}. These results can
-help epidemiologists design studies and plan statistical analysis for
-multi-year, multi-community studies that investigate changes in community-wide
-rates of health outcomes during exposure to the hazards brought by tropical
-cyclones, including counties inland from a storm's landfall.
+cyclone hazards for epidemiological research. Further, we expand on previous
+research on methods to measure exposure to tropical cyclones for
+epidemiological research~\parencite{grabich2015measuring}.  The previous study
+investigated differences in which communities were assessed as exposed to
+tropical cyclones during the 2004 hurricane season in Florida, depending on
+whether exposure assessment was based on distance to the storm's track or a
+metric incorporating storm-related winds within the county during the storm.
+They found important differences across methods of assessing storm exposure,
+concluding that a study may be prone to bias from exposure misclassification if
+distance to the storm track is used as a proxy measurement for exposure
+assessment~\parencite{grabich2015measuring}.  Here we expand to investigate
+this question across a large set of counties and storm seasons.  Further, we
+investigate patterns in exposure classification based on other metrics of
+storm-related hazards---rainfall, flooding, and tornadoes---which are important
+for inland health impacts of tropical cyclones~\parencite{czajkowski2011,
+moore2012}. These results can help epidemiologists design studies and plan
+statistical analysis for multi-year, multi-community studies that estimate
+excess mortality and morbidity exposure to the hazards brought by tropical
+cyclones.

### 2019-11-28 writing/ehp_methods.tex
Continue work on text

@@
 
-We aggregated data for all counties in the eastern half of the \ac{US}
-(Figure~\ref{fig:hurrtracks}). We
-included all tracked land-falling or near-land Atlantic-basin tropical cyclones
-between~1988 and~2015 by covering all 136 storms in \ac{HURDAT2}
-\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
-eastern \ac{US} county in this period (the tracks of the study storms are shown in 
+We aggregated these data for all counties in the eastern half of the \ac{US}
+(Figure~\ref{fig:hurrtracks}).  We included all tropical cyclones in
+\ac{HURDAT2} \parencite{landsea2013} that came within~250 \si{\kilo\metre} of
+at least one eastern \ac{US} county between 1988 and 2015.  This ensured we
+were including all land-falling or near-land Atlantic basin tropical cyclones
+in this period (the tracks of the study storms are shown in
 Figure~\ref{fig:hurrtracks}).
@@
 cumulative storm-related precipitation based on ground-based monitoring. We
-measured the rank correlation (Spearman's~$\rho$~\parencite{spearman1904proof})
+measured the rank correlation (Spearman's~$\rho$;~\textcite{spearman1904proof})
 between storm-specific cumulative precipitation estimates for the two data
@@
 
-To create a dataset of county-level maximum sustained winds during historical
-tropical cyclones, we modeled maximum sustained wind speeds at each county's
-population mean center~\parencite{countycenters} using a wind speed model based
-on results from Willoughby~\parencite{willoughby2006parametric}. For each
-storm, we modeled the maximum ground-level sustained wind speed that the storm
-generated at each county center for each~15-\si{\minute} increment throughout
-the storm tracking period. We used a double exponential wind speed model with
-inputs for the storm's forward speed, direction, and maximum wind
-speed~\parencite{willoughby2006parametric, stormwindmodel}, incorporating the
-asymmetry in wind speeds around the tropical cyclone center that results from
-the storm's forward movement~\parencite{phadke2003modeling}. From these
-15-\si{\minute} interval estimates, we identified the maximum sustained surface
-wind speed in each county over the course of each storm.
+We created a dataset of county-level peak sustained surface wind during each
+storm. To do this, we first modeled each storm's wind field to each county's
+population mean center~\parencite{countycenters} at 15-minute intervals
+throughout the course of the storm. We used a double exponential wind model
+based on results from Willoughby~\parencite{willoughby2006parametric} to model
+1-minute surface wind (10 meters above ground) at each county center, based on
+inputs of the storm's forward speed, direction, and maximum wind
+speed~\parencite{stormwindmodel}. Our model incorporated the asymmetry in wind
+speeds around the tropical cyclone center that results from the storm's forward
+movement~\parencite{phadke2003modeling}.  From these 15-\si{\minute} interval
+estimates, we identified the peak sustained surface wind in each county over
+the course of each storm.
 
-To validate these modeled county-level wind speed estimates, we compared them
+To validate these modeled county-level peak wind estimates, we compared them
 to the  post-season wind radii values, which have been routinely published
-since 2004 \parencite{knaff2016using}. These wind radii characterize storm
-size, estimating the maximum distance from the storm's center that winds of a
-certain intensity extend. They give separate estimates for four quadrants of
-the storm to capture asymmetry in the storm's wind field and are 
-based on a post-season
-re-analysis that incorporates available data (e.g., satellite data, aircraft
-reconnaissance and ground-based data if available)
-\parencite{knaff2016using}. Wind radii are estimated
-for three thresholds of maximum surface wind speeds: 64,~50, and~34 \si{\knot}.
-They therefore allow for the classification of counties into four categories of
-maximum sustained wind speeds: $<$34~\si{\knot}; 34\,--\,49.9~\si{\knot};
-50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}.
+since 2004 \parencite{knaff2016using}. These wind radii estimate the maximum
+distance from the storm's center that winds of a certain intensity extend. They
+give separate estimates for four quadrants of the storm to capture asymmetry in
+the storm's wind field and are based on a post-season re-analysis that
+incorporates available data (e.g., satellite data, aircraft reconnaissance and
+ground-based data if available) \parencite{knaff2016using}. Wind radii are
+estimated for three thresholds of peak sustained surface wind: 64,~50, and~34
+kt.  They therefore allow for the classification of counties into four
+categories of peak sustained surface wind: $<$34~kt; 34\,--\,49.9~kt;
+50\,--\,63.9~kt; $\ge$64~kt.
 
-We interpolated the wind radii data to~15-\si{\minute} increments and
+We interpolated the wind radii data to~15-minute increments and
 classified a county as exposed to winds in a given wind speed category if its
@@
 quadrant from these maximum wind radii \parencite{knaff2016using}.  We compared
-these wind radii--based estimates of county-level maximum sustained surface
-winds during a storm with the modeled wind speed estimates, comparing all study
-storms since 2004 for which at least one study county experienced sustained
-winds of~$\ge$34 \si{\knot} (based on the \ac{HURDAT2} wind radii). For each of
-these storms, we calculated the percent of study counties that were classified
-in the same wind speed category based on both data sources.
+these wind radii--based estimates of county-level peak sustained surface wind
+during a storm with the modeled wind estimates, comparing all study storms
+since 2004 for which at least one study county experienced peak sustained winds
+of~$\ge$34 kt (based on the \ac{HURDAT2} wind radii). For each of these storms,
+we calculated the percent of study counties that were classified in the same
+wind category based on both data sources.
 
@@
 Database~\parencite{stormevents}. While this database has recorded storm data
-since~1950, its coverage changed substantially in~1996 to cover a more types of
+since~1950, its coverage changed substantially in~1996 to cover more types of
 storm events~\parencite{stormevents}. We therefore only considered these
@@
 In our open-source data, we provide continuous measurements of some of the
-exposure metrics: closest distance of each storm to each county, maximum
-sustained winds at the county center, and cumulative rainfall. However,
+exposure metrics: closest distance of each storm to each county, peak sustained
+surface wind at the county center, and cumulative rainfall. However,
 epidemiologic studies of tropical cyclones often use a binary exposure
 classification (``exposed'' versus ``unexposed'') to assess storm-related
-health risks, and so we explored patterns in storm exposure based on binary
-clasifications of these exposure metrics.
+health risks (e.g., \textcite{grabich2015, mckinney2011,
+caillouet2008increase}), and so we explored patterns in storm exposure based on
+binary clasifications of these exposure metrics.
 
@@
 exposure assessment based on each of the measured tropical cyclone hazards
-(cumulative precipitation, maximum sustained winds, flooding, and tornadoes).
+(cumulative precipitation, peak sustained winds, flooding, and tornadoes).
 Depending on available exposure data, this assessment included some or all of
@@
 
-Next, we investigated how well exposure assessment agreed across 
+Finally, we investigated how well exposure assessment agreed across 
 these metrics. For each storm we investigated the degree to which the 

### 2019-11-28 writing/ehp_results.tex
Continue work on text

@@
 package~\parencite{hurricaneexposuredata}.  The package size exceeded the
-recommended maximum size for the standard repository for publishing R packages
-(\ac{CRAN}). Therefore, we used the \textit{drat} framework to set up our own
+recommended maximum size for \ac{CRAN}, the standard repository for publishing
+R packages. Therefore, we used the \textit{drat} framework to set up our own
 repository to host the package~\parencite{anderson2017hosting}. These data
 include continuous county-level measurements for the closest distance of each
-storm, cumulative rainfall, and maximum sustained surface wind speed. These
-continuous metrics can be used for classifying counties as exposed or
-unexposed---using thresholds selected by the user---or can be extracted
-directly as continuous metrics. The dataset also include binary data on flood
-and tornado events associated with each storm in each county. To accompany
-these data, we published an associated R package on \ac{CRAN} with software
-tools to explore and map the data and to integrate it with human health
+storm, cumulative rainfall, and peak sustained surface wind. These continuous
+metrics can be used for classifying counties as exposed or unexposed---using
+thresholds selected by the user---or can be extracted directly as continuous
+metrics. The dataset also include binary data on flood and tornado events
+associated with each storm in each county. To accompany these data, we
+published an additional R package with software tools to explore and map the
+data and to integrate it with human health
 datasets~\parencite{hurricaneexposure}.
@@
 
-For maximum sustained surface wind estimates, we found that the primary wind
+For peak sustained surface wind estimates, we found that the primary wind
 exposure metric in the open-source data generally agreed well with the wind
 radii reported in \ac{HURDAT2}. For most storms, $\ge$95\% of counties were
-assigned the same category of wind speed ($<$34~\si{\knot};
-34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) by both
-data sources (Figure~\ref{fig:windcomparison}).  Disagreement was limited to a
-few storms (e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two
-storms, the modeled wind speed values used in the open-source data somewhat
-overestimated the extent of the storm's most severe winds at landfall but then
-underestimated the extent of 34\,--\,49.9 \si{\knot} storm winds further inland
-(Figure~S2). For epidemiology researchers who would like to conduct sensitivity
-analysis between these two sets of wind data, we have included estimates from
-the \ac{HURDAT2} wind radii as a secondary measure of wind speeds in the
+assigned the same category of wind speed ($<$34~kt; 34\,--\,49.9~kt;
+50\,--\,63.9~kt; $\ge$64~kt) by both data sources
+(Figure~\ref{fig:windcomparison}).  Disagreement was limited to a few storms
+(e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two storms, the
+modeled wind speed values used in the open-source data somewhat overestimated
+the extent of the storm's most severe winds at landfall but then underestimated
+the extent of 34\,--\,49.9 kt storm winds further inland (Figure~S2). For
+epidemiology researchers who would like to conduct sensitivity analysis between
+these two sets of wind data, we have included estimates from the \ac{HURDAT2}
+wind radii as a secondary measure of county-level peak sustained wind in the
 open-source dataset \parencite{hurricaneexposuredata}.
@@
 each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
-strong geographical patterns were clear. Tropical cyclone wind exposure had a
+strong geographical patterns were clear. Peak sustained wind exposure had a
 strong coastal pattern, with almost all exposures in counties within about~200
@@
 there were also inland rain exposures in counties that were never classified as
-exposed based on winds. Flood-based exposures were frequently in the
+exposed based on wind. Flood-based exposures were frequently in the
 Mid-Atlantic region, with a pattern that skewed north compared with other
@@
 
-Next, we measured the agreement between exposure classifications for each pair
+Finally, we measured the agreement between exposure classifications for each pair
 of hazards, as well as with a proxy measurement based on
@@
 We drew similar conclusions when we investigated all~46 tropical cyclones
-between~1996 and~2011 (the period for which data for all five metrics were
+between~1996 and~2011 (when data for all five metrics were
 available) for which~100 or more counties were exposed based on at least one
@@
 while for another set of storms (e.g., Matthew in~2004 and Katrina in~2005),
-there was moderate to good agreement between distance and rain.  However, these
-sets of tropical cyclones represented the minority of tropical cyclones
-considered; for most tropical cyclones, there was low to moderate overlap in
-the counties determined to be exposed to the tropical cyclones between most
-pairings of metrics.
+there was moderate to good agreement between distance and rain.

### 2019-11-28 writing/ehp_tables.tex
Continue work on text

--- 2019-01-28 writing/ehp_tables.tex
@@
 Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988\,--\,2015 & County population mean center within 100~\si{\kilo\metre} of storm track \\ 
+Distance & 1988\,--\,2015 & Storm track came within 100~\si{\kilo\metre} of county population mean center\\ 
 Rain & 1988\,--\,2011 & County cumulative rainfall of $\ge$75~\si{\milli\metre} over the period from two days before to 
-one day after the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the 
-storm track \\ 
-Wind & 1988\,--\,2015 & Modeled storm-associated maximum sustained winds of $\ge$15~\si{\metre\per\second}
+one day after the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\ 
+Wind & 1988\,--\,2015 & Modeled storm-associated peak sustained surface wind of $\ge$34~kts 
 at the county's population mean center\\ 
-Flood & 1996\,--\,2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the storm track \\
-Tornado & 1996\,--\,2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the storm track\\
+Flood & 1996\,--\,2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center \\
+Tornado & 1996\,--\,2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\
 \bottomrule 

### 2020-02-06 tables/stormseverity.tex
Update based on latest hurricaneexposuredata version

@@
-% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Fri Nov  8 11:15:53 2019
+% latex table generated in R 3.6.2 by xtable 1.8-4 package
+% Thu Feb  6 11:53:51 2020
 \begin{table}[ht]
@@
   \midrule
-Distance (1988--2015) & 433 (224, 503) & Beryl, 1994 (330) \\ 
-  Rain (1988--2015) & 401 (152, 553) & Frances, 2004 (464) \\ 
-  Wind (1988--2011) & 143 (51, 232) & Ike, 2008 (251) \\ 
-  Flood (1996--2015) & 213 (84, 248) & Ivan, 2004 (318) \\ 
-  Tornado (1996--2015) & 53 (16, 43) & Ivan, 2004 (91) \\ 
+Distance (1988--2015) & 369 (72, 466) & Beryl, 1994 (344) \\ 
+  Rain (1988--2015) & 298 (48, 441) & Frances, 2004 (464) \\ 
+  Wind (1988--2011) & 151 (55, 246) & Ike, 2008 (248) \\ 
+  Flood (1996--2015) & 189 (67, 228) & Ivan, 2004 (317) \\ 
+  Tornado (1996--2015) & 48 (15, 43) & Ivan, 2004 (91) \\ 
    \bottomrule

### 2020-02-06 R/extent_disagreement_figure.R

@@ -264,3 +264,4 @@ out_d <- map_tracks(storm, plot_object = out) +
 pdf("figures/extentdisagreement.pdf", width = 9, height = 5)
 grid.arrange(out_a, out_d, out_b, out_c, ncol = 2)
 dev.off()
+

### 2020-02-06 R/jaccard_heatmap.R

@@ -194,7 +194,7 @@ for_heatmap <- as.matrix(a[ , 3:12])
 rownames(for_heatmap) <- a$storm_id
 
 row_dend <- hclust(stats::dist(for_heatmap), method = "complete")
-row_dend <- color_branches(row_dend, k = 5)
+row_dend <- color_branches(row_dend, k = 4)
 
 ha_rot_cn <- HeatmapAnnotation(text = anno_text(colnames(for_heatmap), rot = 45, just = "left",
                                                 gp = gpar(fontsize = 10),
@@ -265,7 +265,7 @@ ha_row <- HeatmapAnnotation(counties_exposed = counties_exposed$counties_exposed
                             which = "row", width = unit(0.25, "cm"))
 
 heat_map <- Heatmap(for_heatmap, col = viridis(256, option = "A", direction = -1),
-        cluster_rows = row_dend, split = 5, show_row_dend = FALSE,
+        cluster_rows = row_dend, split = 4, show_row_dend = FALSE,
         row_names_side = "left", row_names_gp = gpar(fontsize = 8),
         show_column_names = FALSE, show_column_dend = FALSE, #column_dend_side = "bottom",
         column_dend_height = unit(6, "mm"),

### 2020-02-06 R/wind_comparison_figure.R

@@ -1,11 +1,14 @@
 library(hurricaneexposuredata)
 library(dplyr)
+library(stringr)
 library(ggplot2)
 library(polycor)
+library(latex2exp)
 data(storm_winds)
 data(ext_tracks_wind)
 
 wind_comp <- storm_winds %>%
+        filter(str_remove(storm_id, ".+\\-") %>%  as.numeric() <= 2015) %>% 
         select(fips, vmax_sust, storm_id) %>%
         rename(vmax_sust_model = vmax_sust) %>%
         inner_join(select(ext_tracks_wind, fips, storm_id, vmax_sust),
@@ -26,23 +29,16 @@ fig <- wind_comp %>%
         filter(non_zero > 0) %>%
         ggplot(aes(x = perc_same, y = reorder(storm_id, perc_same), fill = non_zero)) +
         geom_point(size = 2, shape = 21, color = "black") +
-        scale_x_continuous(labels = scales::percent) +
+        scale_x_continuous(labels = scales::percent, lim = c(0, 1)) +
         theme_classic() +
         labs(x = "% of counties classified in same wind category\nby modeled storm wind and Extended Best Tracks",
              y = "",
-             fill = "Number of counties with modeled\nsustained winds of \u2265 34 knots",
-             parse = TRUE
+             fill = "Number of counties with modeled\n sustained winds of <= 34 knots"
              ) +
         viridis::scale_fill_viridis(direction = -1, option = "B") +
         theme(legend.position = "bottom")
 
-quartz.save(file = "figures/windcomparison.pdf", type = "pdf", width = 5.25, height = 8)
-print(fig)
-dev.off()
-
-tiff(filename = "figures/windcomparison.tiff", width = 525, height = 700)
-print(fig)
-dev.off()
+ggsave("figures/windcomparison.pdf", fig, width = 5.25, height = 8, units = "in")
 
 wind_comp_ordinal <- wind_comp %>%
         mutate(vmax_sust_ext = factor(vmax_sust_ext, ordered = TRUE,

### 2020-02-19 tables/metric_definitions.tex
Update years in definition table

--- 2019-01-26 tables/metric_definitions.tex
@@
 \begin{tabular}{p{0.9cm}p{2.5cm}p{9cm}} 
-Metric & Available & Criteria for exposure \\ \midrule 
-Distance & 1988--2015 & County population mean center within 100 kilometers of storm track \\ 
-Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
-Flood & 1996--2015 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
-Tornado & 1996--2015 & Tornado event listed in the NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
+Metric & Available & Criteria for exposure for binary analysis \\ \midrule 
+Distance & 1988\,--\,2018 & County population mean center within 100 kilometers of storm track \\ 
+Rain & 1988\,--\,2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
+Wind & 1988\,--\,2018 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
+Flood & 1996\,--\,2018 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
+Tornado & 1988\,--\,2018 & Tornado event listed in the NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
 \bottomrule 

### 2020-02-19 R/flood_comparison_figure.R
Update years for flood comparison figure

@@ -25,10 +25,10 @@ library(scales)
 
 county_fips <- c("37183", "42091", "01097", "13121", "12033", "34003", "48201",
                  "24005", "09001")
-county_floods <- county_events(county_fips, start_year = 1996, end_year = 2015, event_type = "flood") %>%
+county_floods <- county_events(county_fips, start_year = 1996, end_year = 2018, event_type = "flood") %>%
         select(storm_id, fips) %>%
         mutate(flood = TRUE)
-county_storms <- county_distance(county_fips, start_year = 1996, end_year = 2015, dist_limit = 500) %>%
+county_storms <- county_distance(county_fips, start_year = 1996, end_year = 2018, dist_limit = 500) %>%
         left_join(county_floods, by = c("fips", "storm_id")) %>%
         mutate(flood = ifelse(is.na(flood), FALSE, flood),
                closest_date = ymd(closest_date)) %>%
@@ -47,10 +47,10 @@ county_storms <- county_distance(county_fips, start_year = 1996, end_year = 2015
                lag = str_replace(lag, "m", "-"),
                lag = as.numeric(lag))
 
-# county_gages <- get_gages(county_fips, start_date = "1996-01-01", end_date = "2015-12-31")
+# county_gages <- get_gages(county_fips, start_date = "1996-01-01", end_date = "2018-12-31")
 # gage_q2s <- find_q2(county_gages$site_no)
 # county_gages <- full_join(county_gages, gage_q2s, by = "site_no")
-# county_flow_data <- get_flow_data(county_gages, start_date = "1996-01-01", end_date = "2015-12-31")
+# county_flow_data <- get_flow_data(county_gages, start_date = "1996-01-01", end_date = "2018-12-31")
 # save(county_gages, file = "data/county_gages.Rdata")
 # save(county_flow_data, file = "data/county_flow_data.Rdata")
 
@@ -69,7 +69,7 @@ out <- county_gage_data %>%
         dplyr::mutate(min_date = min(date),
                       max_date = max(date),
                       n = n()) %>%
-        filter(min_date == ymd("1996-01-01") & max_date == ymd("2015-12-31") & n == 7305) %>%
+        filter(min_date == ymd("1996-01-01") & max_date == ymd("2018-12-31") & n == 8401) %>%
         ungroup() %>%
         right_join(county_storms, by = c("date" = "date", "county_cd" = "fips")) %>%
[…]

### 2020-02-19 R/hurr_tracks_figure.R
Update years for hurricane tracks figure

@@ -5,7 +5,7 @@ library(hurricaneexposure)
 
 data(closest_dist)
 storms <- unique(closest_dist$storm_id)
-storms <- storms[gsub("*.+-", "", storms) <= 2015 &
+storms <- storms[gsub("*.+-", "", storms) <= 2018 &
                          gsub("*.+-", "", storms) >= 1988]
 a <- map_tracks(storms, plot_points = FALSE, alpha = 0.3, color = "darkcyan")
 # Add storms whose names were retired
@@ -19,7 +19,10 @@ fig <- map_tracks(c("Hugo-1989", "Bob-1991", "Andrew-1992",
              "Dennis-2005", "Katrina-2005", "Rita-2005",
              "Wilma-2005", "Noel-2007",
              "Gustav-2008", "Ike-2008", "Paloma-2008",
-             "Irene-2011", "Sandy-2012"),
+             "Irene-2011", "Sandy-2012", "Matthew-2016",  
+             "Harvey-2017", "Irma-2017", "Nate-2017", 
+             "Florence-2018", "Michael-2018"
+             ),
            plot_object = a, plot_points = FALSE, color = "darkcyan",
            padding = 0) +
         theme_void() +
@@ -30,6 +33,3 @@ pdf(file = "figures/hurrtracks.pdf", width = 4.8, height = 4.0)
 print(fig)
 dev.off()
 
-tiff(filename = "figures/hurrtracks.tiff", width = 480, height = 400)
-print(fig)
-dev.off()

### 2020-02-19 tables/stormseverity.tex
Update year for storm severity table

@@
 % latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Feb  6 11:53:51 2020
+% Wed Feb 19 10:41:15 2020
 \begin{table}[ht]
@@
   \midrule
-Distance (1988--2015) & 369 (72, 466) & Beryl, 1994 (344) \\ 
-  Rain (1988--2015) & 298 (48, 441) & Frances, 2004 (464) \\ 
-  Wind (1988--2011) & 151 (55, 246) & Ike, 2008 (248) \\ 
-  Flood (1996--2015) & 189 (67, 228) & Ivan, 2004 (317) \\ 
-  Tornado (1996--2015) & 48 (15, 43) & Ivan, 2004 (91) \\ 
+Distance (1988--2018) & 399 (78, 503) & Beryl, 1994 (344) \\ 
+  Rain (1988--2018) & 298 (48, 441) & Frances, 2004 (464) \\ 
+  Wind (1988--2011) & 162 (55, 258) & Michael, 2018 (260) \\ 
+  Flood (1996--2018) & 197 (76, 241) & Ivan, 2004 (317) \\ 
+  Tornado (1988--2018) & 38 (8, 42) & Ivan, 2004 (91) \\ 
    \bottomrule

### 2020-02-19 R/storm_severity_table.R

@@ -8,7 +8,7 @@ library(stringr)
 my_fips <- unique(closest_dist$fips)
 
 distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
-                                     end_year = 2015, dist_limit = 100) %>%
+                                     end_year = 2018, dist_limit = 100) %>%
         select(storm_id, fips) %>%
         mutate(exposed = TRUE,
                storm_id = as.character(storm_id))
@@ -17,8 +17,8 @@ dist_sum <- distance_exposure %>%
   group_by(storm_id) %>% 
   count() %>% 
   ungroup() %>% 
-  summarize(n_storms = n() / (2015 - 1988 + 1),
-            n_exposures = sum(n) / (2015 - 1988 + 1),
+  summarize(n_storms = n() / (2018 - 1988 + 1),
+            n_exposures = sum(n) / (2018 - 1988 + 1),
             med = median(n),
             p25 = quantile(n, 0.25),
             p75 = quantile(n, 0.75),
@@ -44,7 +44,7 @@ rain_sum <- rain_exposure %>%
             which_max = storm_id[which.max(n)])
 
 wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
-                             end_year = 2015, wind_limit = 17.4) %>%
+                             end_year = 2018, wind_limit = 17.4) %>%
         select(storm_id, fips) %>%
         mutate(exposed = TRUE,
                storm_id = as.character(storm_id))
@@ -52,8 +52,8 @@ wind_sum <- wind_exposure %>%
   group_by(storm_id) %>% 
   count() %>% 
   ungroup() %>% 
-  summarize(n_storms = n() / (2015 - 1988 + 1),
-            n_exposures = sum(n) / (2015 - 1988 + 1),
+  summarize(n_storms = n() / (2018 - 1988 + 1),
+            n_exposures = sum(n) / (2018 - 1988 + 1),
             med = median(n),
             p25 = quantile(n, 0.25),
             p75 = quantile(n, 0.75),
@@ -61,7 +61,7 @@ wind_sum <- wind_exposure %>%
             which_max = storm_id[which.max(n)])
 
 flood_exposure <- county_events(counties = my_fips,
-                                start_year = 1996, end_year = 2015,
+                                start_year = 1996, end_year = 2018,
                                 event_type = "flood") %>%
         select(storm_id, fips) %>%
         mutate(exposed = TRUE,
[…]

### 2020-02-19 R/wind_comparison_figure.R
Update years in wind comparison figure

@@ -8,7 +8,7 @@ data(storm_winds)
 data(ext_tracks_wind)
 
 wind_comp <- storm_winds %>%
-        filter(str_remove(storm_id, ".+\\-") %>%  as.numeric() <= 2015) %>% 
+        filter(str_remove(storm_id, ".+\\-") %>%  as.numeric() <= 2018) %>% 
         select(fips, vmax_sust, storm_id) %>%
         rename(vmax_sust_model = vmax_sust) %>%
         inner_join(select(ext_tracks_wind, fips, storm_id, vmax_sust),
@@ -36,9 +36,10 @@ fig <- wind_comp %>%
              fill = "Number of counties with modeled\n sustained winds of <= 34 knots"
              ) +
         viridis::scale_fill_viridis(direction = -1, option = "B") +
-        theme(legend.position = "bottom")
+        theme(legend.position = "bottom") + 
+        theme(panel.grid.major.y = element_line(colour = "lightgray"))
 
-ggsave("figures/windcomparison.pdf", fig, width = 5.25, height = 8, units = "in")
+ggsave("figures/windcomparison.pdf", fig, width = 5.25, height = 8.5, units = "in")
 
 wind_comp_ordinal <- wind_comp %>%
         mutate(vmax_sust_ext = factor(vmax_sust_ext, ordered = TRUE,

### 2020-02-19 R/average_exposure_figure.R
Update years in average exposure figures

@@ -12,36 +12,36 @@ data(closest_dist)
 
 my_fips <- unique(closest_dist$fips)
 dist <- county_distance(counties = my_fips, start_year = 1988,
-                                     end_year = 2015, dist_limit = 100) %>%
+                                     end_year = 2018, dist_limit = 100) %>%
         group_by(fips) %>%
-        dplyr::summarize(n_exposures = n() / 28) %>%
+        dplyr::summarize(n_exposures = n() / (2018 - 1988 + 1)) %>%
         mutate(metric = "distance")
 
 rain <- county_rain(counties = my_fips, start_year = 1988,
                              end_year = 2011, dist_limit = 500,
                              rain_limit = 75) %>%
         group_by(fips) %>%
-        dplyr::summarize(n_exposures = n() / 24) %>%
+        dplyr::summarize(n_exposures = n() / (2011 - 1988 + 1)) %>%
         mutate(metric = "rain")
 
 wind <- county_wind(counties = my_fips, start_year = 1988,
-                             end_year = 2015, wind_limit = 17.4) %>%
+                             end_year = 2018, wind_limit = 17.4) %>%
         group_by(fips) %>%
-        dplyr::summarize(n_exposures = n() / 28) %>%
+        dplyr::summarize(n_exposures = n() / (2018 - 1988 + 1)) %>%
         mutate(metric = "wind")
 
 flood <- county_events(counties = my_fips,
-                                start_year = 1996, end_year = 2015,
+                                start_year = 1996, end_year = 2018,
                                 event_type = "flood") %>%
         group_by(fips) %>%
-        dplyr::summarize(n_exposures = n() / 20) %>%
+        dplyr::summarize(n_exposures = n() / (2018 - 1996 + 1)) %>%
         mutate(metric = "flood")
 
 tornado <- county_events(counties = my_fips,
-                                start_year = 1996, end_year = 2015,
+                                start_year = 1988, end_year = 2018,
                                 event_type = "tornado") %>%
         group_by(fips) %>%
-        dplyr::summarize(n_exposures = n() / 20) %>%
+        dplyr::summarize(n_exposures = n() / (2018 - 1988 + 1)) %>%
         mutate(metric = "tornado")
 
 our_fips <- hurricaneexposure:::get_eastern_map() %>% dplyr::select(fips) %>% dplyr::distinct() %>%
[…]

### 2020-02-19 R/ivan_figure.R
Remove tiff code for Ivan figure

@@ -110,7 +110,4 @@ pdf("figures/ivanexposure.pdf", width = 9, height = 6)
 print(out)
 dev.off()
 
-tiff("figures/ivanexposure.tiff", width = 800, height = 530)
-print(out)
-dev.off()
 

### 2020-02-19 R/other_top_storm_exposures.R
Update other top exposure figure with new data

@@ -431,6 +431,6 @@ dennis <- map_tracks(storm, plot_object = out) +
         ggtitle("Dennis (2005)")
 
 pdf("figures/othertopstorms.pdf", width = 10, height = 7)
-grid.arrange(floyd, ike, katrina, dennis, ncol = 1)
+grid.arrange(floyd, ike, katrina, ncol = 1)
 dev.off()
 

### 2020-02-19 writing/anderson_ehp_supp.tex
Update years and other details in text

@@
 \author[7]{Dirk Eddelbuettel} 
-\author[1]{Meilin Yan} 
-\author[8]{Roger D. Peng}
+\author[1,8]{Meilin Yan} 
+\author[9]{Roger D. Peng}
 
@@
   Illinois at Urbana-Champaign, Champaign, IL, 61820} 
-\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
+\affil[8]{Beijing Innovation Center for Engineering Science and Advanced Technology 
+  and State Key Joint Laboratory of Environmental Simulation and Pollution Control, 
+  Peking University, Beijing, China, 100871}
+\affil[9]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
   Health, Baltimore, MD, 21205}

### 2020-02-19 writing/ehp_abstract.tex
Update years and other details in text

@@
         Tropical cyclone epidemiology can be 
-	advanced through county-level exposure assessment methods that are comprehensive and 
-	consistent across space and time, facilitating multi-year, multi-storm studies. 
+	advanced through exposure assessment methods that are comprehensive and 
+	consistent across space and time, as these facilite multi-year, multi-storm studies. 
 	Further, an understanding of 
 	patterns in and between exposure metrics that are based on specific hazards of the storm can help in 
-	designing tropical cyclone epidemiological studies.\\ 
+	designing tropical cyclone epidemiologic research.\\ 
 \textbf{Objectives:} (1)~Provide an open-source dataset for tropical
@@
 	Atlantic basin storms, covering~1996\,--\,2011 for all
-	metrics and up to~1988\,--\,2015 for specific metrics. 	
+	metrics and up to~1988\,--\,2018 for specific metrics. 	
 	We validated 
@@
 	exposure.\\ 
-\textbf{Results:} We describe how we developed this dataset and explore how its
-	measurements compare to other data that could be used to characterize
-	tropical cyclone hazards for epidemiological research. Measurements in
-	the dataset created here typically agreed with data from other sources,
-	and we present and discuss areas of disagreement and other caveats for
-	the data. Tropical cyclones typically brought different hazards to different
-	counties, and so agreement was usually low between distance-based tropical
-	cyclone exposure assessment and each hazard-specific metric, as well as 
-	between pairs of hazard-specific metrics. \\
+\textbf{Results:} Our open-source dataset was typically consistent with data from other sources,
+	and we present and discuss areas of disagreement and other caveats. Over the 
+	study period and area, tropical cyclones typically brought different hazards to different
+	counties. Therefore, when comparing exposure assessment between
+	different hazard-specific metrics, agreement was usually low, as it also was when 
+	comparing exposure assessment based on a distance-based proxy measurement and 
+	any of the hazard-specific metrics.\\
 \textbf{Discussion:} Our results provide a multi-hazard dataset that can be
 	leveraged for epidemiologic research on tropical cyclones, as well as 
-	insights that can inform the design of tropical cyclone epidemiological 
-	studies.
+	insights that can inform the design and analysis for tropical cyclone epidemiologic 
+	research.
 \end{abstract}

### 2020-02-19 writing/ehp_cover.tex
Update years and other details in text

@@
 \author[7]{Dirk Eddelbuettel} 
-\author[1]{Meilin Yan} 
-\author[8]{Roger D.~Peng}
+\author[1,8]{Meilin Yan} 
+\author[9]{Roger D.~Peng}
 

### 2020-02-19 writing/ehp_figures.tex
Update years and other details in text

@@
 	shown in this map were investigated. The lines show the paths of the
-	study storms, which included all tracked storms in 1988\,--\,2015 that
+	study storms, which included all tracked storms in 1988\,--\,2018 that
 	are recorded in \ac{HURDAT2} and that came
@@
 	estimate these averages are based on years of available exposure data
-	(rain: 1988\,--\,2011; wind:~1988\,--\,2015; flood and
-	tornado events:~1996\,--\,2015). } 
+	(rain: 1988\,--\,2011; wind:~1988\,--\,2018; flood events:~1996\,--\,2018, and
+	tornado events:~1988\,--\,2018). } 
 \label{fig:averageexposure} 

### 2020-02-19 writing/ehp_methods.tex
Update years and other details in text

@@
 \ac{HURDAT2} \parencite{landsea2013} that came within~250 \si{\kilo\metre} of
-at least one eastern \ac{US} county between 1988 and 2015.  This ensured we
+at least one eastern \ac{US} county between 1988 and 2018.  This ensured we
 were including all land-falling or near-land Atlantic basin tropical cyclones
@@
 and Atmospheric Administration (NOAA)'s Storm Event
-Database~\parencite{stormevents}. While this database has recorded storm data
-since~1950, its coverage changed substantially in~1996 to cover more types of
-storm events~\parencite{stormevents}. We therefore only considered these
-metrics of hurricane exposure for tropical cyclones in~1996 and later.
+Database~\parencite{stormevents}. While this database has recorded storm data,
+particularly tornadoes, since~1950, its coverage changed substantially in~1996
+to cover more types of storm events, including flood
+events~\parencite{stormevents}. We therefore only considered flood metrics of
+hurricane exposure for tropical cyclones in~1996 and later.
 
@@
 Wake County, NC. For each county, we first identified all streamflow gages in
-the county with complete data for Jan.~1~1996\,--\,Dec.~31,~2015. If a storm
+the county with complete data for Jan.~1~1996\,--\,Dec.~31,~2018. If a storm
 did not come within~500~\si{\kilo\metre} of a county, it was excluded from this
@@
 Depending on available exposure data, this assessment included some or all of
-the period from~1988 to~2015 (Table 1). For each binary metric of exposure to a
+the period from~1988 to~2018 (Table 1). For each binary metric of exposure to a
 tropical cyclone hazard, we first summed the total number of county-level

### 2020-02-19 writing/ehp_supp.tex
Update years and other details in text

@@
 The study included all tracked storms in the revised Atlantic hurricane database (HURDAT2 \cite{landsea2013}) that came within 250 kilometers of at least on U.S.
-county, 1988--2015.
+county, 1988--2018.
 Thicker lines show the tracks of storms whose names were retired, indicating that the storm was particularly severe or had notable impacts \cite{retirednames}.}
@@
 \caption{Comparison of estimated wind categories from modeled sustained wind speed (as used for the wind exposure metric in the main text; \cite{stormwindmodel}) versus sustained wind speed determined from the revised Atlantic hurricane database (HURDAT2 \cite{landsea2013}).
-In HURDAT2, wind radii were available for study storms from 2004--2015 and gives estimates of the maximum radii to which storm winds of 34 knots, 50 knots, and 64 knots extended in each four quadrant of the storm at each of the six-hour storm track observations.
+In HURDAT2, wind radii were available for study storms from 2004--2018 and gives estimates of the maximum radii to which storm winds of 34 knots, 50 knots, and 64 knots extended in each four quadrant of the storm at each of the six-hour storm track observations.
 We interpolated this data to 15-minute increments and assessed a county as exposed to winds at or above one of these cutoffs if its population mean center was within 85\% of the maximum radius for that wind speed in its quadrant of the storm.
@@
 1 1996--Dec.
-31, 2015.
+31, 2018.
 The number of streamflow gages meeting this criterion are given in parentheses beside the county's name in the panel title.

### 2020-02-20 tables/stormseverity.tex
Remove "distance" row from severity table

--- 2020-02-06 tables/stormseverity.tex
@@
 % latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Feb  6 11:53:51 2020
+% Thu Feb 20 12:50:06 2020
 \begin{table}[ht]
@@
   \midrule
-Distance (1988--2015) & 369 (72, 466) & Beryl, 1994 (344) \\ 
-  Rain (1988--2015) & 298 (48, 441) & Frances, 2004 (464) \\ 
-  Wind (1988--2011) & 151 (55, 246) & Ike, 2008 (248) \\ 
-  Flood (1996--2015) & 189 (67, 228) & Ivan, 2004 (317) \\ 
-  Tornado (1996--2015) & 48 (15, 43) & Ivan, 2004 (91) \\ 
+Rain (1988--2018) & 298 (48, 441) & Frances, 2004 (464) \\ 
+  Wind (1988--2011) & 162 (55, 258) & Michael, 2018 (260) \\ 
+  Flood (1996--2018) & 197 (76, 241) & Ivan, 2004 (317) \\ 
+  Tornado (1988--2018) & 38 (8, 42) & Ivan, 2004 (91) \\ 
    \bottomrule

### 2020-02-20 R/storm_severity_table.R

@@ -162,7 +162,7 @@ total_sum <- bind_rows(bind_cols(dist_sum, dist_sum2),
   select(metric, yearly_sum, max_storm)
 
 total_sum %>% 
-  filter(metric != "Distance") %>% 
+  slice(2:5) %>% 
   rename(`Metric (years available)` = metric,
          `Mean (interquartile range) of county exposures per year` = yearly_sum,
          `Tropical cyclone with most counties exposures (# exposed counties)` = max_storm) %>%

### 2020-02-20 writing/ehp_discussion.tex
Edit the text

@@
 
-Epidemiological studies can help characterize which health risks are elevated
-during disasters, to what degree, and to
+Epidemiologic studies can help characterize which health risks are elevated
+during disasters, to what degree, and for
 whom~\parencite{ibrahim2005unfortunate, noji2005disasters}.  As a result, these
@@
 different metrics, and we find that county-level tropical cyclone exposure
-assessments vary substantially when using different metrics, based either on
-storm hazards or on how close the storm's central track came to the community.
+assessments vary substantially when using different metrics.
 Our results can inform exposure assessment for future county-level studies of
-the health risks and impacts associated with tropical cyclones exposure as
-well as insights to inform epidemiological study design.
+the health risk and impacts associated with tropical cyclones exposure as
+well as provide insights to inform epidemiologic study design.
 
@@
 and~S2), but there are some caveats. The rainfall data are generally
-well-correlated with ground-based observations, but may sometimes oversmooth
+well-correlated with ground-based observations, but may sometimes underestimate
 very high rainfall values (Figure~\ref{fig:raincomparison}). When rainfall data
@@
 values, and while the modeled wind data generally agree well with post-analysis
-maximum wind radii from \ac{HURDAT2}~\parencite{landsea2013}
+maximum wind radii \parencite{landsea2013}
 (Figure~\ref{fig:windcomparison}), there were a few storms with some
@@
 a start date well before the five-day window we used to temporally match storm
-event listings with tropical cyclone tracks.
+event listings with tropical cyclone tracks for our data.
 
@@
 and time restrictions, like those we used in matching NOAA Storm Event listings
-with tropical cyclone locations and dates, do create the opportunity for
+with tropical cyclone locations and dates, do cause 
 occasional false negatives, as for Hurricane Ida in Fulton County, GA, where a
@@
 flooding~\parencite{rees2001}.  Almost all tropical cyclone tornado exposures
-were in southern coastal states, consistent with previous evidence that most
-noteworthy tropical cyclone-related tornadoes occur on the right-side of the
-tropical cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.
-The exposure averages we calculated may be limited as estimates of
-long-term frequencies, as tropical cyclones follow decadal
-patterns~\parencite{kossin2007more} likely not adequately captured in the
-available data, which covered less than 20 years.
+were in southern coastal states, consistent with previous evidence that
+tropical cyclone-related tornadoes typically occur to the right of tropical
+cyclone tracks in Atlantic-basin \ac{US} storms~\parencite{moore2012}.  It is
+important to note, however, that the exposure averages we calculated may be
+limited as estimates of long-term frequencies, as tropical cyclones follow
+decadal patterns~\parencite{kossin2007more} that may not adequately captured in
+the available data.
 
@@
 while slow-moving tropical cyclones are likely to bring more
-rain~\parencite{rappaport2000} and cause more damage because of sustained
+rain~\parencite{rappaport2000} and may cause more damage because of sustained
 hazardous conditions~\parencite{rezapour2014}. Further, while the likelihood
@@
 example, tropical cyclone wind exposures tend to be concentrated in counties
-near the coast, while tropical cyclone wind exposures sometimes extend well
+near the coast, while tropical cyclone rain exposures sometimes extend well
 inland.  If the etiologically-relevant exposure for a health outcome is extreme
@@
 
-We did, however, find a small set of tropical cyclones for which for which
+We did find a small set of tropical cyclones for which for which
 agreement was high across several single-hazard exposure assessments (e.g.,
@@
 water may have helped maintain wind speeds in similar patterns to rain and
-distance-based exposure assessments, resulting in more similarities across
+distance-based exposures, resulting in more similarities across
 exposure assessments compared to other tropical cyclones.  For these storms, it
 may be possible to assess exposure to multiple hazards of the storm using a
-single metric, in some cases even a proxy like the distance between the county
+single metric, perhaps even a proxy like the distance between the county
 and the storm's track.  However, for these storms it may be difficult to
@@
 focused on exposures that could affect any county, whether inland or coastal.
-Second, this data is aggregated to the county level. Many health outcome
+Second, these data are aggregated to the county level. Many health outcome
 datasets are aggregated at this level, but some may be aggregated at a finer

### 2020-02-20 writing/ehp_introduction.tex
Edit the text

@@
 Hurricanes and other tropical cyclones can severely impact human health in
-\ac{US} communities, as has been characterized in several long-term national
+\ac{US} communities, as shown in several long-term national
 studies~\parencite{rappaport2000, rappaport2014fatalities,
@@
 disaster (e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
-liu2017wildfire, milojevic2017mental}).  These studies, which often use time
-series or case-crossover study designs, typically compare the rate of a health
-outcome in a community during a disaster to rates during comparable
+liu2017wildfire, milojevic2017mental}).  These studies, which typically use time
+series or case-crossover designs, compare the community-wide rate of a health
+outcome during a disaster to rates during comparable
 non-disaster periods.  While such studies cannot attribute specific cases
@@
 in Florida~\parencite{mckinney2011}.  However, to expand this approach to
-longer time periods and large sets of communities, researchers must be able to
+longer time periods and larger sets of communities, researchers must be able to
 assess exposure to tropical cyclones consistently and comparably across storms,
@@
 aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these
-data, it is straightforward to measure whether a community was in a storm's
-direct path or within a certain distance of its central track; some previous
-epidemiological studies have done this to assess exposure to a tropical storm
-(e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}).
+data, it is straightforward to measure a storm's direct path, and so to measure
+whether a community was on or within a certain distance of that central track.
+Some previous epidemiological studies have done this to assess exposure to a
+tropical storm (e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}), 
+and this method can be applied consistently across years and communities 
+for large-scale studies.
 
 However, when epidemiological studies assess exposure to a tropical cyclone
-using, as a proxy, how close the storm came to a community, they
-may misclassify exposure~\parencite{grabich2015measuring}. While a number of
+using, as a proxy, how close the storm came to a community, they may
+misclassify exposure~\parencite{grabich2015measuring}. While a number of
 tropical cyclone hazards are strongly associated with distance from the
@@
 floods, and tornadoes can occur well away from the tropical cyclone's central
-track~\parencite{rappaport2000, atallah2007, moore2012}.  For example, fatal
-tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
-between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
-tropical cyclone's center~\parencite{moore2012}.  Further, when studies use
-distance from the storm's track to assess exposure, they often use the same
-buffer constraints on each side of the storm track
-(e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
-cyclone tend to be distributed around its center non-symmetrically.  Extreme
-winds are more common to the track's right, where counter-clockwise cyclonic
-winds move in concert with the tropical cyclone's forward
-motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
+track~\parencite{rappaport2000, atallah2007, moore2012}.  For example,
+tornadoes generated by tropical cyclones, which were linked to over~300 deaths
+in the \ac{US} between~1995 and~2009, most often
+occur~200\,--\,500~\si{\kilo\metre} from the tropical cyclone's
+center~\parencite{moore2012}.  Further, when studies use distance from the
+storm's track to assess exposure, they often use the same buffer constraints on
+each side of the storm track (e.g.,~\cite{kinney2008, currie2013}).  However,
+the forces of a tropical cyclone tend to be distributed around its center
+non-symmetrically.  Extreme winds are more common to the track's right, where
+counter-clockwise cyclonic winds move in concert with the tropical cyclone's
+forward motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
 associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
@@
 assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
-there is not yet a standard approach, and when different studies use different
-datasets or storm hazards when assessing storm exposure, it becomes difficult
-to compare and aggregate findings.
+there is not yet a standard approach, and when different studies use
+different datasets or storm hazards when assessing storm exposure, it becomes
+difficult to compare and aggregate findings.
 
@@
 county concurrent with the storm.  We aggregated these data at the county level
-since data for epidemiological studies are often available at county-level
-aggregations (e.g., direct hurricane-related deaths~\parencite{czajkowski2011};
-birth outcomes~\parencite{grabich2015, grabich2015measuring}; autism
+since data for epidemiological studies are often available at this level (e.g.,
+direct hurricane-related deaths~\parencite{czajkowski2011}; birth
+outcomes~\parencite{grabich2015, grabich2015measuring}; autism
 prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
@@
 level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
-explore and map these data and to link it with county-level health
+explore and map these data and to link them with county-level health
 data~\parencite{hurricaneexposuredata}.
@@
 measurements compare to other data that could be used to characterize tropical
-cyclone hazards for epidemiological research. Further, we expand on previous
-research on methods to measure exposure to tropical cyclones for
-epidemiological research~\parencite{grabich2015measuring}.  The previous study
-investigated differences in which communities were assessed as exposed to
-tropical cyclones during the 2004 hurricane season in Florida, depending on
-whether exposure assessment was based on distance to the storm's track or a
-metric incorporating storm-related winds within the county during the storm.
-They found important differences across methods of assessing storm exposure,
-concluding that a study may be prone to bias from exposure misclassification if
-distance to the storm track is used as a proxy measurement for exposure
+cyclone hazards for epidemiologic research. Further, we expand on previous
+research on methods to measure exposure to tropical cyclones for epidemiologic
+research~\parencite{grabich2015measuring}. This previous study investigated
+differences in which communities were assessed as exposed to tropical cyclones
+during the 2004 hurricane season in Florida, comparing exposure assessment
+based on distance to the storm's track or versus assessment based on a metric
+that incorporated storm-generated winds within the county.  They found
+important differences across methods of assessing storm exposure, concluding
+that a study may be prone to bias from exposure misclassification if distance
+to the storm track is used as a proxy measurement for exposure
 assessment~\parencite{grabich2015measuring}.  Here we expand to investigate
-this question across a large set of counties and storm seasons.  Further, we
+this question across a larger set of counties and storm seasons.  Further, we
 investigate patterns in exposure classification based on other metrics of
-storm-related hazards---rainfall, flooding, and tornadoes---which are important
-for inland health impacts of tropical cyclones~\parencite{czajkowski2011,
-moore2012}. These results can help epidemiologists design studies and plan
-statistical analysis for multi-year, multi-community studies that estimate
-excess mortality and morbidity exposure to the hazards brought by tropical
-cyclones.
+storm-related hazards---rainfall, flooding, and tornadoes---all of which are
+important for inland health impacts of tropical
+cyclones~\parencite{czajkowski2011, moore2012}. These results can help
+epidemiologists design studies and plan statistical analysis for multi-year,
+multi-community studies that estimate excess mortality and morbidity associated
+with exposure to the hazards brought by tropical cyclones.

### 2020-02-20 writing/ehp_methods.tex
Edit the text

@@
 
-We aggregated these data for all counties in the eastern half of the \ac{US}
-(Figure~\ref{fig:hurrtracks}).  We included all tropical cyclones in
-\ac{HURDAT2} \parencite{landsea2013} that came within~250 \si{\kilo\metre} of
-at least one eastern \ac{US} county between 1988 and 2018.  This ensured we
-were including all land-falling or near-land Atlantic basin tropical cyclones
-in this period (the tracks of the study storms are shown in
+For our study domain, we aggregated data for all counties in the eastern half
+of the \ac{US} (Figure~\ref{fig:hurrtracks}).  In terms of storm coverage, we
+included all tropical cyclones between 1988 and 2018 in \ac{HURDAT2}
+\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
+eastern \ac{US} county, therefore including all land-falling or near-land
+Atlantic basin tropical cyclones while excluding storms that never neared the
+\ac{US}. The tracks of the study storms are shown in
 Figure~\ref{fig:hurrtracks}).
@@
 used tracking data from \ac{HURDAT2}, which records the storm center's position
-at the synoptic times of 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}, and
+at the synoptic times of 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We
 interpolated this position to~15-\si{\minute} intervals using natural cubic
-splines~\parencite{hurricaneexposure}. At each~15-\si{\minute} interval, we
-measured the distance between the storm's center and each county's population
-mean center~\parencite{bivand2013applied, countycenters}. We took the minimum
-distance for each county as a measure of how close the storm came to the
-county.  We also recorded the time the storm came closest to each county, which
-we later used to link observed data on precipitation, flooding, and tornadoes
-to the storm.  To allow matching with data recorded in local time (e.g., health
-data), we converted these times from \ac{UTC} to local
+spline interpolation~\parencite{hurricaneexposure}. At each~15-\si{\minute}
+interval, we measured the distance between the storm's center and each county's
+population mean center~\parencite{bivand2013applied, countycenters}. We took
+the minimum distance for each county as a measure of how close the storm came
+to the county over its lifetime.  We also recorded the time when this occurred
+for each county, which we later used to link observed data on precipitation,
+flooding, and tornadoes.  To allow matching with data recorded in
+local time (e.g., health data), we converted these times from \ac{UTC} to local
 time~\parencite{countytimezones}.
@@
 \ac{WONDER} database \parencite{cdcwonder, alhamdan2014environmental}. To
-create this aggregated data, these coauthors took hourly precipitation
+create these aggregated data, these coauthors took hourly precipitation
 measurements in the \ac{NLDAS-2} precipitation files, which are given
@@
 each county, but could also determine the cumulative rainfall for a more
-extended period, like from two days before to two days after the storm's
-closest approach. For the analysis of binary exposure classifications in this
+extended period (e.g., two days before to two days after the storm's
+closest approach). For the analysis of binary exposure classifications in this
 study, we calculated storm-associated rainfall as the sum of precipitation from
@@
 which precipitation data were available from multiple ground-based stations
-within the Global Historical Climatology Network throughout the study
+in the Global Historical Climatology Network throughout the study
 period~\parencite{menne2012overview, rnoaa, countyweather}. These sample
@@
 the storm's wind field and are based on a post-season re-analysis that
-incorporates available data (e.g., satellite data, aircraft reconnaissance and
+incorporates all available data (e.g., satellite data, aircraft reconnaissance and
 ground-based data if available) \parencite{knaff2016using}. Wind radii are
@@
 categories of peak sustained surface wind: $<$34~kt; 34\,--\,49.9~kt;
-50\,--\,63.9~kt; $\ge$64~kt.
+50\,--\,63.9~kt; and $\ge$64~kt.
 
@@
 MD; Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton
-County, GA; Harris County, TX; Mobile County, AL; Montgomery County, PL; and
+County, GA; Harris County, TX; Mobile County, AL; Montgomery County, PA; and
 Wake County, NC. For each county, we first identified all streamflow gages in
@@
 \noindent This metric can range from~0, in the case of no overlap between the
-counties classified as exposed based on the two metrics, to~1, in the case that
-the two metrics classify exactly the same set of counties as exposed to the
+counties classified as exposed based on the two metrics, to~1, if
+the two metrics classify an identical set of counties as exposed to the
 tropical cyclone. We calculated these values for all study storms that affected

### 2020-02-20 writing/ehp_results.tex
Edit the text

@@
 thresholds selected by the user---or can be extracted directly as continuous
-metrics. The dataset also include binary data on flood and tornado events
+metrics. The dataset also includes binary data on flood and tornado events
 associated with each storm in each county. To accompany these data, we
@@
 between the two data sources, with rank correlations (bottom right of each
-graph in Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some
-evidence that the primary rainfall metric may tend to underestimate rainfall
-totals in storms with extremely high rainfall, based on a few heavy-rainfall
-storms in Harris County, TX, Mobile County, AL, Charleston County, SC, and Wake
-County, NC (Figure~\ref{fig:raincomparison}). However, even in these cases, it
-was rare for a storm to be classified differently (under the precipitation
-threshold of 75~\si{\milli\metre} we used for binary exposure classification
-for later analysis) based on the source of precipitation data---horizontal and
-vertical lines in each small plot in Figure~\ref{fig:raincomparison} show the
-threshold of 75~\si{\milli\metre}, so storms in the lower left and upper right
-quadrants would be classified the same (``exposed'' or ``unexposed'')
-regardless of the precipitation data source, while storms in the upper left and
-lower right quadrants would be classified differently. Such cases were rare.
+graph in Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. 
+There was some evidence that our primary rainfall metric may tend to
+underestimate rainfall totals in storms with extremely high rainfall, based on
+a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
+County, SC, and Wake County, NC (Figure~\ref{fig:raincomparison}). However,
+even in these cases, it was rare for a storm to be classified differently
+(under the precipitation threshold of 75~\si{\milli\metre} we used for binary
+exposure classification for later analysis) based on the source of
+precipitation data. Horizontal and vertical lines in each small plot in
+Figure~\ref{fig:raincomparison} show the threshold of 75~\si{\milli\metre}, so
+storms in the lower left and upper right quadrants would be classified the same
+(``exposed'' or ``unexposed'') regardless of the precipitation data source,
+while storms in the upper left and lower right quadrants would be classified
+differently. Such cases were rare.
 
@@
 exposure metric in the open-source data generally agreed well with the wind
-radii reported in \ac{HURDAT2}. For most storms, $\ge$95\% of counties were
+radii reported in \ac{HURDAT2}. For most storms, $\ge$90\% of counties were
 assigned the same category of wind speed ($<$34~kt; 34\,--\,49.9~kt;
@@
 modeled wind speed values used in the open-source data somewhat overestimated
-the extent of the storm's most severe winds at landfall but then underestimated
-the extent of 34\,--\,49.9 kt storm winds further inland (Figure~S2). For
-epidemiology researchers who would like to conduct sensitivity analysis between
-these two sets of wind data, we have included estimates from the \ac{HURDAT2}
-wind radii as a secondary measure of county-level peak sustained wind in the
-open-source dataset \parencite{hurricaneexposuredata}.
+the severity of the storm's winds at landfall but then underestimated the
+extent of 34\,--\,49.9 kt winds from the storm's central track further inland
+(Figure~S2). For epidemiology researchers who would like to conduct sensitivity
+analysis between these two sets of wind data, we have included estimates from
+the \ac{HURDAT2} wind radii as a secondary measure of county-level peak
+sustained wind in the open-source dataset \parencite{hurricaneexposuredata}.
 
@@
 counties, streamgage data generally indicated higher discharge during periods
-identified as flood events based on the NOAA Storm Events database,
-particularly in counties with more reporting streamgages (e.g., Harris County,
-TX, with~10 reporting stream gages; Figure~\ref{fig:floodcomparison}). There
-were some cases, however, where the two flooding data sources were somewhat
-inconsistent.  For example, there were one or two tropical cyclones in several
-of the counties (Mobile County, AL, Escambia County, FL, and Montgomery County,
-PA) with associated flood event listings but for which the total discharge
-across county streamflow gages was relatively low. For storms without a flood
-listing for the county, in most cases the total streamflow discharge in the
-county was relatively low, and, in all but one of these cases all streamgage
-flows were below the flooding threshold. The exception was for Hurricane Ida in
-2009 in Fulton County, GA.
+identified as flood events based on the NOAA Storm Events database
+(Figure~\ref{fig:floodcomparison}). There were some cases, however, where the
+two flooding data sources were somewhat inconsistent.  For example, there were
+one or two tropical cyclones in several of the counties (Mobile County, AL,
+Escambia County, FL, Fairfield County, CT, and Fulton County, GA) with
+associated flood event listings but for which the total discharge across county
+streamflow gages was relatively low. For storms without a flood listing for the
+county, in most cases the total streamflow discharge in the county was
+relatively low, and, in all but two cases all streamgage flows were
+below the flooding threshold. The exceptions were for Hurricane Ida in 2009 in
+Fulton County, GA, and Hurricane Isaac in 2012 in Mobile County, AL.
 
@@
 (Table~\ref{tab:exposuresummaries}). For tropical cyclone tornadoes,
-there were on average about 50 county exposures per year within our study.
-County exposures were more frequent for tropical cyclone wind exposures ($>$140
+there were on average about 40 county exposures per year within our study.
+County exposures were more frequent for tropical cyclone wind exposures ($>$160
 / yr on average), even more frequent for tropical cyclone flood exposures
-($>$200 / yr on average), and most frequent for tropical cyclone rain exposure
-($>$400 / yr on average). For every hazard except tornadoes, we identified at
+($>$190 / yr on average), and most frequent for tropical cyclone rain exposure
+($>$290 / yr on average). For every hazard except tornadoes, we identified at
 least one tropical cyclone that exposed over~250 counties
@@
 cyclone varied across hazards: Frances in~2004 exposed the most counties based
-on rain, Ike in~2008 based on wind, and Ivan in 2004 based on flooding and
+on rain, Michael in~2018 based on wind, and Ivan in 2004 based on flooding and
 tornadoes (Table~\ref{tab:exposuresummaries}).
@@
 
-Finally, we measured the agreement between exposure classifications for each pair
+Finally, we assessed agreement between exposure classifications for each pair
 of hazards, as well as with a proxy measurement based on
@@
 and Georgia that were not identified as exposed to Ivan based on any other
-metric. Figure~S4 provides similar maps for four other example tropical
+metric. Figure~S4 provides similar maps for three other example tropical
 cyclones (selected because they exposed many \ac{US}  counties based on at
@@
 We drew similar conclusions when we investigated all~46 tropical cyclones
-between~1996 and~2011 (when data for all five metrics were
-available) for which~100 or more counties were exposed based on at least one
-metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
-for one tropical cyclone, and each box in that row shows the Jaccard
-coefficient for a pair of metrics.  For almost all tropical cyclones, Jaccard
-coefficients were~$\le0.6$ between the distance-based metric and any of the
-four hazard-based metrics (i.e., out of the counties assessed as exposed by at
-least one of the two metrics in the pair,~60\si{\percent} or fewer were
-assessed the same under both metrics). For many tropical cyclones, Jaccard
-coefficients were~$\le0.5$, indicating at best moderate agreement.  Among the
-four hazard-based metrics, the tornado-based metric had universally poor
-agreement with other metrics in county-level classification across the tropical
-cyclones considered.  Agreement between other pairs of metrics were better, but
-still typically at best moderate (Jaccard coefficients usually $\le0.5$).
+between~1996 and~2011 (when data for all five metrics were available) for
+which~100 or more counties were exposed based on at least one metric
+(Figure~\ref{fig:jaccard}). In this figure, each row provides results for one
+tropical cyclone, and each box in that row shows the Jaccard coefficient for a
+pair of metrics.  For almost all tropical cyclones, Jaccard coefficients
+were~$\le0.6$ between the distance-based metric and any of the four
+hazard-based metrics (i.e., out of the counties assessed as exposed by at least
+one of the two metrics in the pair,~60\si{\percent} or fewer were assessed the
+same under both metrics). Among the four hazard-based metrics, the
+tornado-based metric had universally poor agreement with other metrics in
+county-level classification across the tropical cyclones considered.  Agreement
+between other pairs of hazard-based metrics were better, but still typically at
+best moderate (Jaccard coefficients usually $\le0.5$).
 
@@
 of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
-in~2003), there was moderate to good agreement for all pairwise combinations of
+in~2003), there was moderate to good agreement for pairwise combinations of
 distance, rain, and wind, but poor agreement for other combinations of metrics,

### 2020-02-21 writing/anderson_ehp_supp.tex
Make a few edits to main text

--- 2019-11-28 writing/anderson_ehp_supp.tex
@@
 \author[7]{Dirk Eddelbuettel} 
-\author[1]{Meilin Yan} 
-\author[8]{Roger D. Peng}
+\author[1,8]{Meilin Yan} 
+\author[9]{Roger D. Peng}
 
 \affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
-  State University, Fort Collins, CO, 80523} 
-\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, 80045} 
+  State University, Fort Collins, CO, USA} 
+\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, USA} 
 \affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
-  Center, Huntsville, AL, 35805}
+  Center, Huntsville, AL, USA}
 \affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
-  University, Fort Collins, CO, 80523} 
+  University, Fort Collins, CO, USA} 
 \affil[5]{Department of Industrial and Operations Engineering, University of 
-  Michigan, Ann Arbor, MI, 48109}
-\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
+  Michigan, Ann Arbor, MI, USA}
+\affil[6]{Department of Geography, Ohio State University, Columbus, OH, USA}
 \affil[7]{Department of Statistics, University of
-  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
-\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
-  Health, Baltimore, MD, 21205}
+  Illinois at Urbana-Champaign, Champaign, IL, USA} 
+\affil[8]{Beijing Innovation Center for Engineering Science and Advanced Technology 
+  and State Key Joint Laboratory of Environmental Simulation and Pollution Control, 
+  Peking University, Beijing, China}
+\affil[9]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
+  Health, Baltimore, MD, USA}
 \affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}

### 2020-02-21 writing/ehp_cover.tex
Make a few edits to main text

--- 2019-11-28 writing/ehp_cover.tex
@@
 \author[7]{Dirk Eddelbuettel} 
-\author[1]{Meilin Yan} 
-\author[8]{Roger D.~Peng}
+\author[1,8]{Meilin Yan} 
+\author[9]{Roger D.~Peng}
 
 \affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
-  State University, Fort Collins, CO, 80523} 
-\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, 80045} 
+  State University, Fort Collins, CO, USA} 
+\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, USA} 
 \affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
-  Center, Huntsville, AL, 35805}
+  Center, Huntsville, AL, USA}
 \affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
-  University, Fort Collins, CO, 80523} 
+  University, Fort Collins, CO, USA} 
 \affil[5]{Department of Industrial and Operations Engineering, University of 
-  Michigan, Ann Arbor, MI, 48109}
-\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
+  Michigan, Ann Arbor, MI, USA}
+\affil[6]{Department of Geography, Ohio State University, Columbus, OH, USA}
 \affil[7]{Department of Statistics, University of
-  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
-\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
-  Health, Baltimore, MD, 21205}
+  Illinois at Urbana-Champaign, Champaign, IL, USA} 
+\affil[8]{Beijing Innovation Center for Engineering Science and Advanced Technology 
+  and State Key Joint Laboratory of Environmental Simulation and Pollution Control, 
+  Peking University, Beijing, China}
+\affil[9]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
+  Health, Baltimore, MD, USA}
 \affil[*]{Corresponding author: G.~Brooke Anderson, brooke.anderson@colostate.edu,
-  1681 Campus Delivery, Fort Collins, CO, 80523}
+  1681 Campus Delivery, Fort Collins, CO, USA}
 

### 2020-02-21 writing/ehp_methods.tex
Make a few edits to main text

@@
 in our data, since these metrics were based on whether an event was listed in
-the NOAA Storm Events database.  For other exposure metrics, each county
-was classified as exposed to a tropical cyclone based on whether the exposure
+the NOAA Storm Events database.  For other exposure metrics, each county was
+classified as exposed to a tropical cyclone based on whether the exposure
 metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
-picked reasonable thresholds, but others could be used with the open-source
-data and its associated software.
+picked reasonable thresholds (e.g., the threshold for gale-force wind for the
+wind exposure), but others could be used with the open-source data and its
+associated software.
 
@@
 years when all exposure data were available (1996\,--\,2011).
+
+\subsection*{Data and code access}
+
+We have posted study data as an open-source R package
+\cite{hurricaneexposuredata}. We have posted additional data and code at
+\url{https://github.com/geanders/county_hurricane_exposure}.

### 2020-02-21 writing/ehp_results.tex
Make a few edits to main text

@@
 tropical cyclone, and each box in that row shows the Jaccard coefficient for a
-pair of metrics.  For almost all tropical cyclones, Jaccard coefficients
-were~$\le0.6$ between the distance-based metric and any of the four
-hazard-based metrics (i.e., out of the counties assessed as exposed by at least
-one of the two metrics in the pair,~60\si{\percent} or fewer were assessed the
-same under both metrics). Among the four hazard-based metrics, the
-tornado-based metric had universally poor agreement with other metrics in
-county-level classification across the tropical cyclones considered.  Agreement
-between other pairs of hazard-based metrics were better, but still typically at
-best moderate (Jaccard coefficients usually $\le0.5$).
+pair of metrics. For all pairs of metrics, agreement in exposure assessment
+was, at best, moderate for all but a few storms. When comparing distance- and
+wind-based exposure assessment, only about 10\% of storms had Jaccard indices
+higher than 0.6 (i.e., out of the counties assessed as exposed by at least one
+of the two metrics in the pair,~60\si{\percent} or more were assessed the same
+under both metrics). For comparisons of assessments based on other combinations
+of distance-, wind-, rain-, and flood-based metrics, fewer than 5\% of storms
+had Jaccard indices above 0.6.  The tornado-based metric had universally poor
+agreement with other metrics in county-level classification across the tropical
+cyclones considered.
 
-There were, however, exceptions\,---\,tropical cyclones in which exposure
-assessment agreed well across several of the metrics considered.  For Floyd
-in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
-agreed moderately to well for all pairs of exposure metrics except those
-involving the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
-of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
-in~2003), there was moderate to good agreement for pairwise combinations of
-distance, rain, and wind, but poor agreement for other combinations of metrics,
-while for another set of storms (e.g., Matthew in~2004 and Katrina in~2005),
-there was moderate to good agreement between distance and rain.
+There were a few tropical cyclones in which exposure assessment agreed well
+across several of the metrics considered.  For Floyd in~1999 (Figure~S4) and
+Irene in~2011, for example, county-level classification agreed moderately to
+well for all pairs of exposure metrics except those involving the tornado-based
+metric (Figure~\ref{fig:jaccard}). For another set of tropical cyclones (e.g.,
+Ernesto in~2006, and Bertha in~1996, Isabel in~2003), there was moderate to
+good agreement for pairwise combinations of distance, rain, and wind, but poor
+agreement for other combinations of metrics, while for another set of storms
+(e.g., Matthew in~2004 and Katrina in~2005), there was moderate to good
+agreement between distance and rain.

### 2020-02-21 writing/ehp_abstract.tex
Fix some typos in abstract

@@
 	advanced through exposure assessment methods that are comprehensive and 
-	consistent across space and time, as these facilite multi-year, multi-storm studies. 
+	consistent across space and time, as these facilitate multi-year, multi-storm studies. 
 	Further, an understanding of 
@@
 \textbf{Objectives:} (1)~Provide an open-source dataset for tropical
-	cyclone exposure assessment for epidemiological research and
+	cyclone exposure assessment for epidemiologic research and
 	(2)~investigate patterns and agreement between county-level assessments of tropical cyclone

### 2020-02-21 writing/ehp_discussion.tex
Make some small edits

@@
 event start dates to storm dates, many false positives would be
-captured, for which the occurrance of a storm in the midst of an ongoing storm
+captured, for which the occurrence of a storm in the midst of an ongoing storm
 event might be improperly attributed to the storm.  However, distance
@@
 characterizes the landfall of tropical cyclones.  Tropical cyclone rain
-exposures tended to extended further inland compared to wind exposures, up to
+exposures tended to extend further inland compared to wind exposures, up to
 the Appalachian mountains. This agrees with previous research indicating that
@@
 (although some resulting coastal flooding is captured). We are exploring ways
-to include this in future versions of the dataset, but to date we have
-focused on exposures that could affect any county, whether inland or coastal.
-Second, these data are aggregated to the county level. Many health outcome
-datasets are aggregated at this level, but some may be aggregated at a finer
-spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e.,
-point locations for each outcome). We have published the wind model used to
-create this dataset as its own open-source R package
-\parencite{stormwindmodel}, and it can be used to model storm-associated winds
-at a finer spatial resolution; however, measurements of other hazards cannot
-similarly be re-scaled through tools we provide. Next, we provide these data
-and associated software tools through R packages, and so some experience in the
-R programming language is required to make full use of them. However, R is
-currently a popular programming language for environmental epidemiology,
-allowing the data to reach a large audience, and we are exploring options to
-create a web application using the Shiny platform to allow broader web-based
-access of the data \parencite{shiny2019}.  Finally, to assess patterns and
-agreement for binary exposure classifications, we have chosen one set of
-sensible thresholds for binary classifications based on continuous metrics
-(rainfall, maximum sustained surface wind, and distance from the storm's track).
-Results and conclusions would differ somewhat with other threshold choices. We
-have published code for this analysis online, allowing
-other researchers to explore other threshold choices for these
-analyses.
+to include this in future versions of the dataset, but to date we have focused
+on exposures that could affect any county, whether inland or coastal.  Second,
+these data are aggregated to the county level. Many health outcome datasets are
+aggregated at this level, but some may be aggregated at a finer spatial
+resolution (e.g., census tract or ZIP code) or unaggregated (i.e., point
+locations for each outcome). We have published the wind model used to create
+this dataset as its own open-source R package \parencite{stormwindmodel}, and
+it can be used to model storm-associated winds at a finer spatial resolution;
+however, measurements of other hazards cannot similarly be re-scaled through
+tools we provide. Next, we provide these data and associated software tools
+through R packages, and so some experience in the R programming language is
+required to make full use of them. However, R is currently a popular
+programming language for environmental epidemiology, allowing the data to reach
+a large audience, and we are exploring options to create a web application
+using the Shiny platform to allow broader web-based access of the data
+\parencite{shiny2019}.  Finally, to assess patterns and agreement for binary
+exposure classifications, we have chosen one set of sensible thresholds for
+binary classifications based on continuous metrics (rainfall, maximum sustained
+surface wind, and distance from the storm's track).  Results and conclusions
+would differ somewhat with other threshold choices. We have published code for
+this analysis online
+(\url{https://github.com/geanders/county_hurricane_exposure}), allowing other
+researchers to explore other threshold choices for these analyses.
 

### 2020-02-21 writing/ehp_introduction.tex
Make some small edits

@@
 storm's central track came to the community.  While other approaches have been
-developed to incorporate storm hazards, particulary wind, into exposure
+developed to incorporate storm hazards, particularly wind, into exposure
 assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),

### 2020-02-21 writing/ehp_methods.tex
Make some small edits

@@
 measurements in the \ac{NLDAS-2} precipitation files, which are given
-at~1/8\si{\degree} grid points across the continential \ac{US}, and summed them
+at~1/8\si{\degree} grid points across the continental \ac{US}, and summed them
 by day for each grid point.  This created a daily precipitation total for each
@@
 caillouet2008increase}), and so we explored patterns in storm exposure based on
-binary clasifications of these exposure metrics.
+binary classifications of these exposure metrics.
 
@@
 We have posted study data as an open-source R package
-\cite{hurricaneexposuredata}. We have posted additional data and code at
+\parencite{hurricaneexposuredata}. We have posted additional data and code at
 \url{https://github.com/geanders/county_hurricane_exposure}.

### 2020-02-24 writing/anderson_ehp_supp.tex
Work on some final edits throughout

@@
 	number of counties exposed based on any metric) from each of the 
-	clusters shown in the Jaccard heatmap in the main text (Figure 7).}
+	clusters shown in the Jaccard heatmap in the main text (Figure 7; a similar
+	map for Hurricane Ivan in 2004 is shown in Figure 6 of the main text).}
 \label{fig:othertopstorms}

### 2020-02-24 writing/ehp_figures.tex
Work on some final edits throughout

@@
 	and (2) estimates based on \ac{HURDAT2}'s wind speed radii.
-	Each point represents a storm, with the x axis giving the percent of
+	Each point represents a storm, with the x-axis giving the percent of
 	counties classified in the same category of peak sustained surface
@@
 	2004, the earliest year for which post-storm reanalysis wind speed
-	radii estimates are routinely available in \ac{HURDAT2}, and for which at
+	radii are routinely available in \ac{HURDAT2}, and for which at
 	least one study county had a peak sustained wind of $\ge$34~\si{\knot}
-	based on the \ac{HURDAT2} wind radii.
+	based on the post-storm wind radii.
 	}
@@
 	estimate these averages are based on years of available exposure data
-	(rain: 1988\,--\,2011; wind:~1988\,--\,2018; flood events:~1996\,--\,2018, and
+	(rain: 1988\,--\,2011; wind:~1988\,--\,2018; flood events:~1996\,--\,2018; and
 	tornado events:~1988\,--\,2018). } 
@@
 	\caption{Agreement between exposure
-	classifications based on different exposure metrics for all storms
+	classifications based on different single-hazard exposure metrics for all storms
 	between~1996 and~2011 for which at least~100 counties were exposed

### 2020-02-24 writing/ehp_introduction.tex
Work on some final edits throughout

@@
 ascertainment: identifying and characterizing deaths for which there is a clear
-link with a disaster, through death certificate coding or other indicators
+link with a disaster through death certificate coding or other indicators
 \parencite{rocha2017medicolegal}.
@@
 whether a community was on or within a certain distance of that central track.
-Some previous epidemiological studies have done this to assess exposure to a
+Some previous epidemiologic studies have done this to assess exposure to a
 tropical storm (e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}), 
@@
 
-However, when epidemiological studies assess exposure to a tropical cyclone
+However, when epidemiologic studies assess exposure to a tropical cyclone
 using, as a proxy, how close the storm came to a community, they may
@@
 occur~200\,--\,500~\si{\kilo\metre} from the tropical cyclone's
-center~\parencite{moore2012}.  Further, when studies use distance from the
-storm's track to assess exposure, they often use the same buffer constraints on
-each side of the storm track (e.g.,~\cite{kinney2008, currie2013}).  However,
-the forces of a tropical cyclone tend to be distributed around its center
-non-symmetrically.  Extreme winds are more common to the track's right, where
-counter-clockwise cyclonic winds move in concert with the tropical cyclone's
-forward motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
+center~\parencite{moore2012}.
+
+Further, when studies use distance from the storm's track to assess exposure,
+they often use the same buffer constraints on each side of the storm track
+(e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
+cyclone tend to be distributed around its center non-symmetrically.  Extreme
+winds are more common to the track's right, where counterclockwise cyclonic
+winds move in concert with the tropical cyclone's forward
+motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
 associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
@@
 atallah2007, zhu2013variations} or undergoes an extratropical
-transition~\parencite{elsberry2002}.  The multi-hazard nature of tropical
-cyclones therefore makes it hard to assess exposure based on how close the
-storm's central track came to the community.  While other approaches have been
-developed to incorporate storm hazards, particularly wind, into exposure
-assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
-there is not yet a standard approach, and when different studies use
-different datasets or storm hazards when assessing storm exposure, it becomes
-difficult to compare and aggregate findings.
+transition~\parencite{elsberry2002}.
+
+The multi-hazard nature of tropical cyclones therefore makes it hard to assess
+exposure based on how close the storm's central track came to the community.
+While other approaches have been developed to incorporate storm hazards,
+particularly wind, into exposure assessment (e.g., \cite{grabich2015measuring,
+zandbergen2009, czajkowski2011}), there is not yet a standard approach, and
+when different studies use different datasets or storm hazards when assessing
+storm exposure, it becomes difficult to compare and aggregate findings.
 
@@
 county concurrent with the storm.  We aggregated these data at the county level
-since data for epidemiological studies are often available at this level (e.g.,
+since data for epidemiologic studies are often available at this level (e.g.,
 direct hurricane-related deaths~\parencite{czajkowski2011}; birth
@@
 cyclone hazards for epidemiologic research. Further, we expand on previous
-research on methods to measure exposure to tropical cyclones for epidemiologic
+research on measuring exposure to tropical cyclones for epidemiologic
 research~\parencite{grabich2015measuring}. This previous study investigated
@@
 during the 2004 hurricane season in Florida, comparing exposure assessment
-based on distance to the storm's track or versus assessment based on a metric
-that incorporated storm-generated winds within the county.  They found
-important differences across methods of assessing storm exposure, concluding
-that a study may be prone to bias from exposure misclassification if distance
-to the storm track is used as a proxy measurement for exposure
-assessment~\parencite{grabich2015measuring}.  Here we expand to investigate
-this question across a larger set of counties and storm seasons.  Further, we
+based on distance to the storm's track or versus a metric that incorporated
+storm-generated winds within the county.  They found important differences,
+concluding that a study may be prone to bias from exposure misclassification if
+distance to the storm track is used as a proxy for
+exposure~\parencite{grabich2015measuring}.  Here we expand to investigate this
+question across a larger set of counties and storm seasons.  Further, we
 investigate patterns in exposure classification based on other metrics of
-storm-related hazards---rainfall, flooding, and tornadoes---all of which are
-important for inland health impacts of tropical
-cyclones~\parencite{czajkowski2011, moore2012}. These results can help
-epidemiologists design studies and plan statistical analysis for multi-year,
-multi-community studies that estimate excess mortality and morbidity associated
-with exposure to the hazards brought by tropical cyclones.
+storm-related hazards---rainfall, flooding, and tornadoes---important for
+inland health impacts of tropical cyclones~\parencite{czajkowski2011,
+moore2012}. These results can help epidemiologists design studies and plan
+statistical analysis for multi-year, multi-community studies that estimate
+excess mortality and morbidity associated with exposure to the hazards brought
+by tropical cyclones.

### 2020-02-24 writing/ehp_methods.tex
Work on some final edits throughout

@@
 
-For our study domain, we aggregated data for all counties in the eastern half
-of the \ac{US} (Figure~\ref{fig:hurrtracks}).  In terms of storm coverage, we
-included all tropical cyclones between 1988 and 2018 in \ac{HURDAT2}
-\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
-eastern \ac{US} county, therefore including all land-falling or near-land
-Atlantic basin tropical cyclones while excluding storms that never neared the
-\ac{US}. The tracks of the study storms are shown in
-Figure~\ref{fig:hurrtracks}).
-
-\subsection*{Distance-based exposure metric}
+For our study domain, we used counties in the eastern half of the \ac{US}
+(Figure~\ref{fig:hurrtracks}).  We included all tropical cyclones between 1988
+and 2018 that were tracked in \ac{HURDAT2} \parencite{landsea2013} and came
+within~250 \si{\kilo\metre} of at least one eastern \ac{US} county
+(Figure~\ref{fig:hurrtracks}). We therefore including all land-falling or
+near-land Atlantic basin tropical cyclones while excluding storms that never
+neared the \ac{US}.  \subsection*{Distance-based exposure metric}
 
@@
 interval, we measured the distance between the storm's center and each county's
-population mean center~\parencite{bivand2013applied, countycenters}. We took
+population mean center~\parencite{countycenters}. We took
 the minimum distance for each county as a measure of how close the storm came
 to the county over its lifetime.  We also recorded the time when this occurred
-for each county, which we later used to link observed data on precipitation,
+for each county so we could later link observed data on precipitation,
 flooding, and tornadoes.  To allow matching with data recorded in
@@
 This process generated daily county-level precipitation estimates for each
-continental \ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.  We
-matched these county-level daily measurements by date with storm tracks, using
-the date when the storm was closest to each county. Given the location of
-storm-affected counties and the typical timing of tropical cyclones, these
-precipitation measures primarily represent rainfall, although occasionally they
-may represent snowfall or other types of precipitation.
+continental \ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.  In this
+study, we matched these county-level daily measurements by date with storm
+tracks, using the date when the storm was closest to each county. Given the
+location of storm-affected counties and the typical timing of tropical
+cyclones, these precipitation measures primarily represent rainfall, although
+occasionally they may represent snowfall or other types of precipitation.
 
@@
 To validate these modeled county-level peak wind estimates, we compared them
-to the  post-season wind radii values, which have been routinely published
-since 2004 \parencite{knaff2016using}. These wind radii estimate the maximum
+to the  post-season wind radii, which have been routinely published
+since 2004 \parencite{knaff2016using}. These estimate the maximum
 distance from the storm's center that winds of a certain intensity extend. They
@@
 the storm's wind field and are based on a post-season re-analysis that
-incorporates all available data (e.g., satellite data, aircraft reconnaissance and
+incorporates all available data (e.g., satellite data, aircraft reconnaissance, and
 ground-based data if available) \parencite{knaff2016using}. Wind radii are
@@
 
-We interpolated the wind radii data to~15-minute increments and
-classified a county as exposed to winds in a given wind speed category if its
-population mean center was within~85\% of the maximum radius for that wind
-speed in that quadrant of the storm. The~85\% adjustment is based on previous
-research that found that this ratio helps capture average wind extents within a
-quadrant from these maximum wind radii \parencite{knaff2016using}.  We compared
-these wind radii--based estimates of county-level peak sustained surface wind
-during a storm with the modeled wind estimates, comparing all study storms
-since 2004 for which at least one study county experienced peak sustained winds
-of~$\ge$34 kt (based on the \ac{HURDAT2} wind radii). For each of these storms,
-we calculated the percent of study counties that were classified in the same
-wind category based on both data sources.
+We interpolated the wind radii data to~15-minute increments using linear
+interpolation \parencite{hurricaneexposuredata} and classified a county as
+exposed to winds in a given wind speed category if its population mean center
+was within~85\% of the maximum radius for that wind speed in that quadrant of
+the storm. The~85\% adjustment is based on previous research that found that
+this ratio helps capture average wind extents within a quadrant based on these
+maximum wind radii \parencite{knaff2016using}.  We compared these wind
+radii--based estimates of county-level peak sustained surface wind during a
+storm with the modeled wind estimates, comparing all study storms since 2004
+for which at least one study county experienced peak sustained winds of~$\ge$34
+kt (based on the post-season wind radii). For each of these storms, we
+calculated the percent of study counties that were classified in the same wind
+category based on both data sources.
 
@@
 events~\parencite{stormevents}. We therefore only considered flood metrics of
-hurricane exposure for tropical cyclones in~1996 and later.
+tropical cyclone exposure for storms in~1996 and later.
 
@@
 We considered nine study counties, selecting counties geographically spread
-through storm-prone areas of the eastern \ac{US} and with multiple streamflow
-gages reporting data during events. The sample counties were: Baltimore County,
+through storm-prone areas of the eastern \ac{US} and with at least one streamflow
+gage reporting data during all events. The sample counties were: Baltimore County,
 MD; Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton
 County, GA; Harris County, TX; Mobile County, AL; Montgomery County, PA; and
-Wake County, NC. For each county, we first identified all streamflow gages in
-the county with complete data for Jan.~1~1996\,--\,Dec.~31,~2018. If a storm
-did not come within~500~\si{\kilo\metre} of a county, it was excluded from this
-analysis, but all other study storms were considered. 
-For each storm and county, we summed streamflow measurements across all county
-gages to generate daily totals for the five-day window around the storm's
-closest approach. We took the maximum of these daily streamflow totals as a
-measure of the county's maximum daily streamflow during that storm. We also
-calculated the percent of streamflow gages in the county with a daily
-streamflow that exceeded a threshold of flooding (the streamgage's median value
-for annual peak flow) on any day during the five-day
-window. We investigated how these measurements varied between storms with
-associated flood events listings versus storms without an
-event listing for the county, to explore if storms with flood event listings
-tended to be associated with higher streamflow at gages within the county.
+Wake County, NC.
+
+For each county, we first identified all streamflow gages in the county with
+complete data for Jan.~1~1996\,--\,Dec.~31,~2018. If a storm did not come
+within~500~\si{\kilo\metre} of a county, it was excluded from this analysis,
+but all other study storms were considered.  For each storm and county, we
+summed streamflow measurements across all county gages to generate daily totals
+for the five-day window around the storm's closest approach. We took the
+maximum of these daily streamflow totals as a measure of the county's maximum
+daily streamflow during that storm. We also calculated the percent of
+streamflow gages in the county with a daily streamflow that exceeded a
+threshold of flooding (the streamgage's median value for annual peak flow) on
+any day during the five-day window. We investigated how these measurements
+varied between storms with associated flood events listings versus storms
+without an event listing for the county, to explore if storms with flood event
+listings tended to be associated with higher streamflow at gages within the
+county.
 
@@
 metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
-picked reasonable thresholds (e.g., the threshold for gale-force wind for the
+picked reasonable thresholds (e.g., the threshold for gale-force wind for 
 wind exposure), but others could be used with the open-source data and its
@@
 exposure assessment based on each of the measured tropical cyclone hazards
-(cumulative precipitation, peak sustained winds, flooding, and tornadoes).
+(cumulative precipitation, peak sustained wind, flooding, and tornadoes).
 Depending on available exposure data, this assessment included some or all of
@@
 
-We have posted study data as an open-source R package
-\parencite{hurricaneexposuredata}. We have posted additional data and code at
+We have posted most study data as an open-source R package
+\parencite{hurricaneexposuredata}. We have posted remaining data and code at
 \url{https://github.com/geanders/county_hurricane_exposure}.

### 2020-02-24 writing/ehp_results.tex
Work on some final edits throughout

@@
 from other available sources.  For estimates of storm-associated rainfall, we
-compared estimates included in the open-source data with ground-based
+compared data in the open-source package with ground-based
 observations in nine sample counties (Figure~\ref{fig:raincomparison}). Within
@@
 exposure metric in the open-source data generally agreed well with the wind
-radii reported in \ac{HURDAT2}. For most storms, $\ge$90\% of counties were
+radii reported in \ac{HURDAT2}. For most storms, $\>$90\% of counties were
 assigned the same category of wind speed ($<$34~kt; 34\,--\,49.9~kt;
@@
 (e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two storms, the
-modeled wind speed values used in the open-source data somewhat overestimated
-the severity of the storm's winds at landfall but then underestimated the
-extent of 34\,--\,49.9 kt winds from the storm's central track further inland
-(Figure~S2). For epidemiology researchers who would like to conduct sensitivity
-analysis between these two sets of wind data, we have included estimates from
-the \ac{HURDAT2} wind radii as a secondary measure of county-level peak
-sustained wind in the open-source dataset \parencite{hurricaneexposuredata}.
+modeled wind speed in the open-source data somewhat overestimated the severity
+of the storm's winds at landfall but then underestimated, particularly for
+Hurricane Sandy, how far 34\,--\,49.9 kt winds extended from the storm's
+central track further inland (Figure~S2). For epidemiology researchers who
+would like to conduct sensitivity analysis using both sets of wind data,
+we have included estimates from the \ac{HURDAT2} wind radii as a secondary
+measure of county-level peak sustained wind in the open-source dataset
+\parencite{hurricaneexposuredate}.
 
@@
 exposure was also more frequent in coastal areas compared to inland areas,
-there were also inland rain exposures in counties that were never classified as
-exposed based on wind. Flood-based exposures were frequently in the
-Mid-Atlantic region, with a pattern that skewed north compared with other
+there were also inland rain exposures in counties that were rarely or never
+classified as exposed based on wind. Flood-based exposures were frequently in
+the Mid-Atlantic region, with a pattern that skewed north compared with other
 exposures. Rain and, to some extent, flood exposures were characterized by a
@@
 
-Finally, we assessed agreement between exposure classifications for each pair
-of hazards, as well as with a proxy measurement based on
-distance from the storm's track. We found that these exposure classifications
-typically did not agree strongly between pairs of metrics---the set of counties
-identified as ``exposed'' based on one metric often overlapped little with the set
-identified as ``exposed'' by another metric, as storms frequently brought
+Finally, we assessed within-storm agreement between exposure classifications
+for each pair of hazards, as well as with a proxy measurement based on distance
+from the storm's track. We found that these exposure classifications typically
+did not agree strongly between pairs of metrics---the set of counties
+identified as ``exposed'' based on one metric often overlapped little with the
+set identified as ``exposed'' by another metric, as storms frequently brought
 different hazards to different locations.
@@
 
-There were a few tropical cyclones in which exposure assessment agreed well
-across several of the metrics considered.  For Floyd in~1999 (Figure~S4) and
-Irene in~2011, for example, county-level classification agreed moderately to
-well for all pairs of exposure metrics except those involving the tornado-based
-metric (Figure~\ref{fig:jaccard}). For another set of tropical cyclones (e.g.,
-Ernesto in~2006, and Bertha in~1996, Isabel in~2003), there was moderate to
-good agreement for pairwise combinations of distance, rain, and wind, but poor
-agreement for other combinations of metrics, while for another set of storms
-(e.g., Matthew in~2004 and Katrina in~2005), there was moderate to good
-agreement between distance and rain.
+There were a few exceptions---tropical cyclones in which exposure assessment
+agreed well across several of the metrics considered.  For Floyd in~1999
+(Figure~S4) and Irene in~2011, for example, county-level classification agreed
+moderately to well for all pairs of exposure metrics except those involving the
+tornado-based metric (Figure~\ref{fig:jaccard}). For another set of tropical
+cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel in~2003), there was
+moderate to good agreement for pairwise combinations of distance, rain, and
+wind, but poor agreement for other combinations of metrics, while for another
+set of storms (e.g., Matthew in~2004 and Katrina in~2005), there was moderate
+to good agreement between distance and rain.

### 2020-02-24 R/storm_severity_table.R

@@ -154,7 +154,7 @@ total_sum <- bind_rows(bind_cols(dist_sum, dist_sum2),
          sum = paste(round(med), " (", round(p25), ", ", round(p75), ")", sep = ""),
          max_storm = paste(str_replace(which_max, "-", ", "),
                            " (", max, ")", sep = ""),
-         years = c("1988--2018", "1988--2018", "1988--2011",
+         years = c("1988--2018", "1988--2011", "1988--2018",
                    "1996--2018", "1988--2018"),
          yearly_sum = paste(round(mean_exp), " (", round(p25_exp), 
                             ", ", round(p75_exp), ")", sep = ""),
@@ -164,8 +164,8 @@ total_sum <- bind_rows(bind_cols(dist_sum, dist_sum2),
 total_sum %>% 
   slice(2:5) %>% 
   rename(`Metric (years available)` = metric,
-         `Mean (interquartile range) of county exposures per year` = yearly_sum,
-         `Tropical cyclone with most counties exposures (# exposed counties)` = max_storm) %>%
+         `Mean of county exposures per year (interquartile range)` = yearly_sum,
+         `Tropical cyclone with most counties exposed (# exposed counties)` = max_storm) %>%
   xtable(align = "cp{4.5cm}p{4.5cm}p{4.5cm}",
          caption = "Summary statistics for the number of county tropical cyclone exposures under each metric.",
          label = "tab:exposuresummaries",

### 2020-07-07 ehp_revision/response_to_reviewers.Rmd
Start response

---
title: "Response to reviewers' comments for *\"Assessing United States county-level exposure for research on tropical cyclones and human health\"*"
output: bookdown::pdf_document2
toc: false
bibliography: hurr_exposure.bib
header-includes:
- \usepackage{xcolor}
- \usepackage{framed}
- \usepackage{bm}
- \usepackage{subcaption}
- \usepackage{gensymb}
- \renewcommand\familydefault{\sfdefault}
- \usepackage{sansmath}
- \usepackage{bm}
- \usepackage{soul}
- \sansmath
- \usepackage{booktabs}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{pdflscape}
- \usepackage{float}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{fullpage}
- \usepackage{pdflscape}
- \usepackage{tcolorbox}
- \tcbuselibrary{skins,breakable}
- \usepackage{amsmath}
- \usepackage{xtab}
- \usepackage{rotating}
- \allowdisplaybreaks
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overall comments {-#overall-comments}

\colorlet{shadecolor}{blue!8}
\begin{shaded}
\textbf{Associate Editor's Overview:}
Developing a consistent and comparable metric for exposure to tropical cyclones
would be a valuable addition to the literature. All reviewers had questions
about the exposure metric, including the challenges of using county-level data
where not everyone in a county would have been exposed and the difficulties of
going from country-level to individual -level risks. Further the comparison of
different storm hazards could be clarified. The reviewers identified several
issues where further explanation would increase the accessibility of the
manuscript.
\end{shaded}

# Reviewer 1: {-#reviewer-1:}

## Overall comments {-#overall-comments}
\colorlet{shadecolor}{blue!8}
\begin{shaded}
\textbf{Reviewer 1 Overview:}
I applaud the authors for this work, and their efforts to develop standardized
ways to measure exposure to tropical cyclones. As a disaster scientist, I very
much appreciate efforts that promote our collective ability to learn from and
across disasters to build cumulative science. 
\end{shaded}

**Response:**

\begin{shaded}
\textbf{Reviewer 1 (R1) Comment 1:}
I very much appreciate the rationale for authors' decision to focus at the
county level, especially given the available health-related data. But, since
tropical cyclones don't necessarily have consistent impacts across an entire
county, I think the limitations of this decision are worthy of further
discussion in the discussion section.
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R1 Comment 2:}
Are there opportunities to expand this approach internationally? Some decisions
(e.g., county-level) data may restrict cross-country comparisons. Can the
authors discuss this? 
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R1 Comment 3:}
In non-health fields, a county's inclusion in a FEMA major disaster declaration
or eligibility for public/individual assistance (all of which are somewhat
reflective of level of damage) is sometimes used as an "exposure" proxy. Have
you considered if/how your exposure assessments relate to inclusion in a FEMA
disaster declaration? In other words, does exposure to any of the hazards align
with the level of damage required for a county to be considered a "disaster
area" and/or worthy of federal assistance by current federal policy? Does this
matter?
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R1 Comment 4:}
Line 95---wondering if there is a better word for "hits" here. Maybe "exposed?"

\textit{Lines 92--95 in the original manuscript:}

\begin{quotation}
\noindent
\textit{"In many cases, these studies analyze multi-year, multi-community data,
allowing them to estimate average associations over many disasters and to
explore how a disaster’s characteristics, or the characteristics of the
community it hits, modify associated health risks (e.g., Anderson and Bell 2010;
Son et al. 2012; Liu et al. 2017)."}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R1 Comment 5:}
Lines 137--139, suggest breaking into two sentences

\textit{Lines 137--141 from original manuscript:}

\begin{quotation}
\noindent
\textit{"While other approaches have been developed to incorporate storm
hazards, particularly wind, into exposure assessment (e.g., Grabich et al.
2015a; Zandbergen 2009; Czajkowski et al. 2011), there is not yet a standard
approach, and when different studies use different datasets or storm hazards
when assessing storm exposure, it becomes difficult to compare and aggregate
findings."}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R1 Comment 6:}
This is not a comment on the paper, but the authors may also want to consider
publishing their work in DesignSafe, which is "the web-based research platform
of the NHERI  [NSF-Funded Natural Hazards Engineering Research Infrastructure]
Network that provides the computational tools needed to manage, analyze, and
understand critical data for natural hazards research." It is the data
repository widely used by interdisciplinary hazards and disaster researchers.
\end{shaded}

**Response:**

# Reviewer 2: {-#reviewer-2}

\begin{shaded}
\textbf{Reviewer 2 overview:}
Exposure assessment is essential to study health impacts from disasters, but it
is also a great source of biases due to a lack of a good and fine measure of
exposures during the chaos. Often, epidemiologists use existing data at large
population level such as county level weather data which provide no individual
level exposures but give average exposure at the population level. So it is very
challenging to draw a good causal inference to health outcomes that measured at
the individual level. This manuscript is a well-written report with generously
shared data and programs on US tropical cyclones. This will contribute
significantly to disaster epidemiology which is still a new area and suffering
from good exposure ascertainment.

While I enjoyed reading the manuscript very much, I have one question I could
not figure out clearly.

One objective is to investigate patterns and agreement between different
exposure measures. I do not clearly understand the value of studying agreement
between exposures, while each exposure should be measured separately, so both
individual and overall effects can be assessed.  I do understand using "distance
from the storm" as exposure is only proxy measure and highly sensitive to
misclassification, but because direct storm hazards such as rainfall, flood,
storm surge, and tornado are publicly accessible as authors indicated, authors
could just simply guide readers to use the best exposure data rather than
conducting a quite intense analysis. This may be due to my limited knowledge, so
it would be appreciated if the authors provide explicit explanations on the
purpose and benefit of the analysis.

\end{shaded}

**Response:**

\begin{shaded}
\textbf{Reviewer 2 (R2) Comment 1:}
Line 164: Isn’t it obvious to use better exposure measures than proxies? Wind,
rainfall, tornado, and flood are direct products of storms while distance is
only proxy of those exposures. Rather than calculating agreement among the
exposures, it would be important to note that these are independent hazards that
can come together or come with different intensity and at different time
windows.

\textit{Lines 164--166 from original manuscript:}

\begin{quotation}
\noindent
\textit{"They found important differences, concluding that a study may be prone
to bias from exposure misclassification if distance to the storm track is used
as a proxy for exposure (Grabich et al. 2015a)."}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 2:}
Line 354: I am not sure why different exposures should be compared and measured
the agreement. I would just treat each exposure as single independent exposure
and give a summary score by summing all the exposure with some weighting for
multiple exposures.

I may not understand the importance of this approach. If so, adding the
explanation would help readers not to questioning.

\textit{Line 354 from original manuscript:}

\begin{quotation}
\noindent
\textit{"Finally, we investigated how well exposure assessment agreed across
these metrics."}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 3:}
Line 394: It looks like the correlation coefficients can be very low if measured
from 75mm or somewhere around.  This seems to be a significant issue for
analyzing storms with heavier rainfall using NLDAS-2 data.

\textit{Line 392--394 from original manuscript:}

\begin{quotation}
\noindent
\textit{"Within these counties, storm-related rainfall measurements were
well-correlated between the two data sources, with rank correlations (bottom
right of each graph in Figure 2) between 0.87 and 0.98. "}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 4:}
Figure 4: I would add Y-axis label “NOAA classification."
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 5:}
Table 2: This is another confusing table. Not sure what compared here. The
second column is a mean \# of exposed counties over years and the third column
seems like showing single storm with the max \# of exposed counties. Not sure
what info I should digest from this table.
\end{shaded}

> **Text from Results of original manuscript:**
"Across the four storm hazards considered, there was wide variation in the
average number of county exposures per year (Table 2). For tropical cyclone
tornadoes, there were on average about 40 county exposures per year within our
study. County exposures were more frequent for tropical cyclone wind exposures
(>160/year on average), even more frequent for tropical cyclone flood exposures
(>190/year on average), and most frequent for tropical cyclone rain exposure
(>290/year on average). For every hazard except tornadoes, we identified at
least one tropical cyclone that exposed over 250 counties (Table 2). However,
the largest-extent tropical cyclone varied across hazards: Frances in 2004
exposed the most counties based on rain, Michael in 2018 based on wind, and Ivan
in 2004 based on flooding and tornadoes (Table 2)."

**Response:**

\begin{shaded}
\textbf{R2 Comment 6:}
Subsection of Results on \textit{Agreement across exposure metrics}: 
As stated below, storms bring different hazards to different locations with
different time. What would be new learning by comparing these different pairs of
exposure metrics. Somewhat apple and orange comparison. Maybe providing more
explanation can help what the important learning from this is. I may not catch
it. Similar comment was added earlier.
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 7:}
Lines 494--495: Why this agreement is important? Same questions added above
already, but still wonder what I missed. It seems like the authors try to
develop an exposure matrix that is common in occupational epidemiology, but it
is only valid method when the exposures are similar and lead to the same health
outcomes.  Rain, wind, flood, and tornado should be treated separately, maybe
except rain and flood. It is possible to be an extreme storm with only severe
wind without major impacts from other exposures. Tornado is a good example.

\textit{Line 491--495 from original manuscript:}

\begin{quotation}
\noindent
\textit{"For another set of tropical cyclones (e.g., Ernesto in 2006, and Bertha
in 1996, Isabel in 2003), there was moderate to good agreement for pairwise
combinations of distance, rain, and wind, but poor agreement for other
combinations of metrics, while for another set of storms (e.g., Matthew in 2004
and Katrina in 2005), there was moderate to good agreement between distance and
rain."}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 8:}
Line 515: It would be informative to provide correlation coefficient for only
above 75 mm measures.

\textit{Line 514--515 from original manuscript:}

\begin{quotation}
\noindent
\textit{"The rainfall data are generally well-correlated with ground-based
observations, but may sometimes underestimate very high rainfall values (Figure
2)."}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 9:}
Line 602: This is not a new finding, but how much misclassified would be a good
question to answer.

\textit{Line 600--602 from original manuscript:}

\begin{quotation}
\noindent
\textit{"Based on our results, the use of a distance-based metric to assess
exposure to any of these hazards, or the use of measurements from one hazard as
a proxy for exposure to any of the other hazards considered, would often
introduce exposure misclassification."}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 10:}
Lines 626--627: It would still be beneficial to treat each exposure as a single
variable first and then consider overall impact. Only if each exposure data are
available.

\textit{Line 626--628 from original manuscript:}

\begin{quotation}
\noindent
\textit{"For these storms, it may be possible to assess exposure to multiple
hazards of the storm using a single metric, perhaps even a proxy like the
distance between the county and the storm’s track."}
\end{quotation}
\end{shaded}

**Response:**

# Reviewer 3: {-#reviewer-3}

\begin{shaded}
\textbf{Reviewer 3 overview:}
This paper describes the development of a dataset and R package that can be used
to assess exposure to tropical cyclone-related hazards (e.g., wind,
precipitation, flooding) at the county level in the eastern United States, as
well as a descriptive analysis of county-level exposure to these hazards. This R
package is novel and I believe of high interest to researchers interested in the
health effects of tropical cyclones.  I have a few minor suggestions to further
improve the clarity of the manuscript.
\end{shaded}

**Response:**

\begin{shaded}
\textbf{Reviewer 3 (R3) Comment 1:}
Starting on line 128, the authors write "Extreme winds are more common to the
track's right, where counterclockwise cyclonic winds move in concert with the
tropical cyclone's forward motion". I think this statement may be specific to
tropical cyclones in the Northern hemisphere (i.e., those relevant to exposure
in the US). Would be good to clarify this.
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R3 Comment 2:}
Line 186: What is meant by the phrase "synoptic times"?

\textit{Lines 137--139 from original manuscript:}

\begin{quotation}
\noindent
\textit{"\ldots"}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 3:}
Several of the exposures are calculated for the population mean center of each
county. From the reference list, it looks like the authors consistently used
population centers from 2010 throughout the years included in the dataset/R
package. If so, I suggest clarifying this in the text. 
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 4:}
In the methods section starting on line 276, the authors describe the flooding
and tornado data from the NOAA storm events database. I'd suggest providing some
information about where the events in this database come from (e.g., that they
are reported events) here in addition to covering it in the discussion.

\begin{quotation}
\noindent
\textit{"\ldots"}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 5:}
Line 294: What is meant by the phrase "traditional tornado event database"?

\begin{quotation}
\noindent
\textit{"\ldots"}
\end{quotation}
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 6:}
Table 1: I suggest providing the rationale for picking the thresholds for
classifying continuous exposures into binary ones within the table. 
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 7:}
Figure 3: The x-axis refers to one of the two wind data sources as "Extended
Best Tracks." Is this the same thing as the wind radii dataset described in the
text? Is so, I suggest clarifying this in the figure label.
\end{shaded}

**Response:**

# Requests from EHP: {-#requests-from-ehp}

\begin{shaded}
\textbf{EHP Request 1:}
Figure 3: Please truncate x-axis scale as you do for Figure S7. Since smallest
value is $\sim75\%$, this would make the data in the relevant range much easier to
appreciate.
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 2:}
Main text figures (general)
To ensure that figures are accessible to readers with impaired color vision, do not use color as the sole means of conveying information in a figure unless absolutely necessary. Use different symbols, shading/textures, and line patterns instead of (or in addition to) color to distinguish among different data points. Use contrasting colors that can be easily distinguished from each other when the figure is printed in black and white. For details, see: \url{https://ehp.niehs.nih.gov/authors/figures}.
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 3:}
Figure 3: Please check the contrast between the colors used to indicate numbers
of counties. If you can see the gradient when printed in black and white it
should be ok, but some of the differences are very subtle. Truncating the x-axis
might help with this (if possible.)
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 4:}
Figure 4: Even with nominally normal color vision and a large high-resolution
monitor, it is difficult for me to make out the color differences in the
symbols-- almost all appear black or very dark blue, a dozen or so look yellow,
and a handful look dark grey. Is there any way to increase the contrast in these
colors? Would you consider binning into categories instead of using a
continuous color gradient?
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 5:}
Figures 5--6: Please check to see if the colors used can be distinguished if
printed in black and white, and if not, use a “color-blind friendly” palette if
possible. If this is not feasible, we can request and exemption from
accessibility requirements, but we try to do this only when alternatives  would
degrade the information content of  the figure.
\end{shaded}

**Response:**

\begin{shaded}
\textbf{R2 Comment 6:}
Supplemental Material: Please provide your supplemental material file in Word
(.docx) format. For additional information, see:
\url{https://ehp.niehs.nih.gov/authors/supplemental-material}. \end{shaded}

**Response:**

# References for the response {-#references-for-the-response}

### 2020-07-10 ehp_revision/response_to_reviewers.Rmd
Start some on response

@@
 toc: false
-bibliography: hurr_exposure.bib
+bibliography: writing/hurr_exposure.bib
 header-includes:
@@
 
+Thank you for the opportunity to revise and resubmit this manuscript. We
+appreciate the helpful suggestions from the reviewers for clarifying and
+deepening the discussion in the paper.
+
+Broadly, we have:
+
+1. Added discussion on the question of county-level versus individual-level
+data. 
+2. Added explanations for how our comparisons among different storm hazards
+can help epidemiologists in designing studies and statistical analysis.
+3. Added a small analysis of the correlation in rainfall metrics specifically
+in cases of high rainfall ($\ge75$ mm of cumulative rainfall in the county
+associated with the storm), to deepen the discussion of potential disagreement
+in continuous measures of rainfall during extreme conditions. 
+4. Changed labeling, axis range, and other presentation details for some
+figures, as well as confirmed that the color scales will be accessible.
+5. Corrected typos and improved wording throughout based on suggestions of the
+reviewers.
+
 # Reviewer 1: {-#reviewer-1:}
@@
 
+- Ecological fallacy vs individual fallacy, especially in the disaster context
+- Spatial misalignment / ecological effects of aggregating data at the county
+(or other geopolitical) level. 
+- Could be particularly problematic in some high-risk areas, because of the 
+"long and narrow" characteristics of island counties. Dare County, Key West.
+
+We think this is a very interesting question for disaster epidemiology, as
+while there remains potential for ecological bias, the potential pathways for
+health effects may also pass through damage in the individual's community beyond
+their residence. For example, damage to major roads in the community might make
+it harder for many residents to reach medical care in the aftermath of the
+storm, even if their home did not experience extensive destruction. Similarly,
+since power is provided over a network, conditions in an individual's community
+can be important in determining risk of long-term outages (does Seth have
+something on this?). As another example, if hospitals in the community are 
+over capacity or have to evacuate, this could increase health risk for people in 
+a fairly large "catchment" area for that hospital.
+
+Therefore, while we agree that individual-level data would in many cases be 
+helpful for tropical cyclone epidemiology, the ideal might be to design studies 
+that integrate both community-level and individual-level metrics of exposure. 
+This could help in determining if there are risks introduced to individuals based
+on storm-related hazards in their broader community, after controlling for the 
+hazards experienced directly at their homes.
+
 \begin{shaded}
@@
 **Response:**
+
+- Differs by exposure metric. Easier for some (wind, rain) than for others (flood, 
+tornado)
+- For rain, we use a re-analysis product only available for the continental US (right?).
+However, there are similar gridded re-analysis products available for other regions 
+worldwide that could be used in a similar way.
+- For wind, the wind model could be generalizable with some tweaks. These include: 
+  + More landmasks
+  + Direction of storm movement for Northern vs Southern Hemispheres
+  + Different wind averaging periods for the recorded best tracks data in different 
+  storm basins.
 
@@
 **Response:**
+
+This is a great question, and we are aware of a number of studies on the
+societal impacts of tropical cyclones that have used FEMA disaster declarations
+to assess county-level exposure to tropical cyclones [refs].
+
+However, evidence from prior tropical cyclone epidemiology that doing so is
+problematic and should be avoided [@grabich2015measuring]. FEMA's disaster
+declarations are issued to provide assistance, based on damage assessment, to
+individuals or the entire public in certain locations where catastrophes
+overwhelm the state or local government [@mccarthy2014fema], and any use of them
+for exposure assessment is secondary. Due to the political nature of disaster
+declaration process, these declarations are often subject to many political and
+economic factors [@mccarthy2014fema; @logue1981research]. One epidemiological
+study specifically focused on exposure assessment for tropical cyclone
+epidemiology and compared use of FEMA declarations with other methods of
+exposure assessment, including one based directly on the storm hazard of wind
+[@grabich2015measuring]. They found that exposure assessment based on FEMA
+declarations tended to overassign counties to be "exposed," resulting in false
+positives [@grabich2015measuring]. A further concern is that, given the
+political and economic nature of these declarations, there is likely variation
+across time and geography in the likelihood of a given exposure resulting in a
+disaster declaration. This is particularly worrisome for a study that seeks to
+explore risk across multiple years and affected communities.
+
+We have added some discussion of this point...
+
+- Future research could expand the work by [@grabich2015measuring] to determine
+the extent of these issues with use of FEMA disaster declarations for
+epidemiological exposure assessment. This would need to go beyond the comparison
+with exposure assessment based on storm hazards, because sometimes the
+characteristics of the affected community modifies the amount of disaster that
+results from hazards of a certain severity. For example, a community that it
+exposed often might be more "hardened" against damage from the storm (any work
+from Seth on this) and so might experience less infrastructure damage from a
+tropical cyclone with certain characteristics compared to a community that is
+rarely exposed to tropical cyclones. One interesting direction would be to
+investigate if there is evidence of differences in probablity of disaster
+declarations at geopolitical boundaries, like when comparing counties on either
+side of a state border.
 
@@
 **Response:**
+
+Thank you for this suggestion. We have changed the wording in this sentence:
+
+> **Text in revised manuscript (relevant change in bold):**
+"In many cases, these studies analyze multi-year, multi-community data, allowing
+them to estimate average associations over many disasters and to explore how a
+disaster’s characteristics, or the characteristics ofthe **affected
+communities**, modify associated health risks (e.g., Anderson and Bell 2010; Son
+et al. 2012; Liu et al. 2017)."
 
@@
 
+This is a great suggestion. We have split the cited sentence into two sentences:
+
+> **Text in revised manuscript:**
+"While other approaches have been developed to incorporate storm
+hazards, particularly wind, into exposure assessment (e.g., Grabich et al.
+2015a; Zandbergen 2009; Czajkowski et al. 2011), there is not yet a standard
+approach. When different studies use different datasets or storm hazards
+when assessing storm exposure, it becomes difficult to compare and aggregate
+findings."
+
 \begin{shaded}
@@
 **Response:**
+
+Great idea. We are investigating sharing this work in this repository.
 
@@
 \end{shaded}
+
+Thank you for this suggestion, and we can see that we could have more clearly 
+articulated in the original manuscript how our assessment of patterns in and 
+among exposures to specific hazards can help epidemiologists to design their
+studies and the statistical methods to analyze the resulting data.
+
+1. If exposure to the different hazards strongly agrees (i.e., when a county is 
+exposed to one hazard, they are typically also exposed to a second), this would
+have several implications for epidemiological research, especially for multi-year
+studies: 
+  + From a statistical modeling point of view, if you tried to include each 
+  hazard as a separate independent variable, you would run into problems with 
+  collinearity if fitting the study data with a generalized linear model or other
+  regression model. One implication of this is that the confidence intervals on 
+  effect estimates for the two variables would be very inflated, because the model
+  would struggle to figure out which of the two variables the weight should be placed
+  on, and so this uncertainty would be reflected in the size of the associated 
+  confidence intervals.
+  + If you see an effect, it would be very hard to try to determine if the causal 
+  pathway goes through the first hazard or the second. To mitigate effects, it is 
+  important to understand the pathway through which they happen, and it might be easier
+  to get clues towards pathways if the hazards are better separated. (Analogy: if you
+  were able to conduct a controlled experiment to look at the effects of each hazard, 
+  jointly and in combination, you would design separation in the assignment of these
+  treatments, so you could distinguish. We can't do that because we are limited to 
+  an observational analysis, but the area of causal inference is doing a lot to 
+  explore how we can design observational studies and analyze data to come closer to 
+  what we might be able to see with a randomized control study, and which of those methods
+  we might be able to use for tropical cyclone epidemiology will depend on the interplay of
+  these patterns of exposures.)
+2. These results will not be surprising to an atmospheric scientist / disaster researcher,
+but there are environmental epidemiologists who have done research on other exposures and
+are now starting to study tropical cyclones (session at ISEE, for example). It is useful 
+for epidemiologists entering into tropical cyclone research to understand these facets of
+tropical cyclone exposure, including the distinction between the very coastal wind exposures
+and other hazards that tend to reach further inland. 
+3. There are epidemiological studies that have used only one hazard for exposure assessment.
+This has most often been the case for wind. If this exposure assessment is meant to 
+capture "tropical cyclone exposure" in general, with all the hazards that are related to 
+tropical cyclones, then there is the chance for exposure misclassification if this single
+hazard is used. 
+4. In the past, it was a very complex issue to try to pull together data on 
+all these hazards. We agree that now that we have provided a dataset with all of them, 
+epidemiologists should be thinking about multi-hazard analyses. However, there still 
+will be some barriers to that. [Recent work on multi-pollutant exposures and the complexities
+associated with that.] If one of the hazards dominates in explaining associated health 
+risk, then in some cases it might be reasonable to focus on that hazard, but in that 
+case it is critical, based on our results, to remember that the exposure assessment is
+not capturing all hazards of the storms.
+
+We have moved other suggestions that you sent in "Track Changes" of the Word 
+document for the original manuscript and address each in detail below.
 
@@
 
+As we note in our general response to your comments, we agree in general with your
+instinct here. However, there are some subtle points to note.
+
+First, we understand the potential appeal of creating a single exposure 
+summary score based on a summation of exposures. This simplifies some of the 
+statistical modeling, for example, and also provides a single exposure estimate to 
+provide as the key study conclusion. However, in many areas of tropical cyclone epidemiology, 
+it is unlikely that exposure to each type of hazard should be given equal weight. For 
+example, an analysis of respiratory health in the months after the storm might hypothesize
+a strong connection with flooding, and smaller connections with rainfall and winds (through
+the pathway of property damage). Use of a pre-defined summation of exposure to each hazard
+would mask some of the potential association.
+
+While a summary statistic could be defined that adds a weight to each exposure, it would
+often be unclear *a priori* what weight each hazard exposure should be given. A better 
+method might be to use a regression framework, where each hazard exposure is included as
+a separate explanatory variable, and the model uses the observed data to determine the 
+appropriate weight for each hazard in terms of modifying risk to the outcome of interest. 
+To explore the influence of multiple concurrent hazards, interaction terms could be added.
+
+However, if our analysis had found that hazard exposures follow very similar patterns, in 
+terms of the counties exposed during a storm, then this strong agreement would result in 
+a number of problems in fitting and interpreting these types of models. First, strong
+agreement in two or more of the explanatory variables in a regression model leads to 
+collinearity, which among other implications often leads to a large inflation of confidence
+intervals for the effect estimates of the correlated explanatory variables. Further, if 
+the hazard exposures were in strong agreement, then it would be difficult or impossible to
+disentagle the effects of the separate hazards, as a strong effect for one would be 
+impossible to interpret with confidence, since the strong agreement in the data (and resulting
+low co-variability in the variables in the data) could cause the model to mistakenly 
+attribute the effect of one hazard to the other. Further, it may become impossible in this
+case to estimate synergistic effects between the two, as an indicator of an interaction
+between the two exposures would be almost collinear with the two main effects.
+
+In our analysis, we find that this is usually not the case for the four tropical cyclone 
+hazards that we consider. Instead, we find that for most storms, the hazards are fairly 
+well-separated in terms of the counties they effect. There is a particularly strong pattern
+where inland counties can be exposed to extreme rain and flooding, but not wind. This 
+separation opens the door for some interesting study designs and statistical modeling. 
+For example, these patterns suggest that a study could investigate whether storm-associated
+rains alone create a health risk, as there are many exposures with only rain in inland 
+counties, and so health associations could be explored here and contrasted with coastal 
+areas that experienced both wind and extreme rain.
+
 \begin{shaded}
@@
 **Response:**
+
+This is a great idea. In the original manuscript, we describe this lower correlation
+at higher precipitation values through the use of examples from specific storms, and
+discuss the potential reasons for it, but measuring the correlation specifically within
+these higher-precipitation events would add more quantitative evidence to this discussion.
+
+Here is the text where we discuss this point in the original manuscript:
+
+> **Text in original manuscript:**
+"There was some evidence that our primary rainfall metric may tend to
+underestimate rainfall totals in storms with extremely high rainfall, based on
+a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
+County, SC, and Wake County, NC (Figure 2)."
+
+> **Text in original manuscript:**
+"The rainfall data are generally well-correlated with ground-based observations,
+but may sometimes underestimate very high rainfall values (Figure 2). When
+rainfall data is used to create binary exposure classifications, this
+disagreement is unlikely to influence results, as both data sources agree in
+identifying these as storms with high rainfall, but would be important to
+consider for cases that include rainfall as a continuous measurement."
+
+[Why we think this is happening... Re-analysis product uses a model to integrate
+data from multiple sources and so might be likely to pull values back toward 
+the "typical", and so oversmooth very extreme values.]
+
+Here is the text and analysis we have added on this point in the revised manuscript:
 
@@
 **Response:**
+
+[Check: *exposure matrix* in occupational epidemiology...]
+
+[New advances in multi-pollutant analysis.]
+
+[Some of the hazards could create similar risks, if they converge on a common
+pathway of risk. For example, if pathway is through stress from property damage,
+individually and in the county, then several hazards could indeed lead to the
+same health outcomes (e.g., PTSD, other psychological markers, risk of acute
+cardiovascular outcomes, etc.). One example is extreme level of destruction both
+from wind (Hurricane Andrew, Hurricane Hugo), from river flooding associated
+with among of precipitation (Hurricane Floyd, Hurricane Florence), and from
+flash flooding associated with the rate of precipitation (Virginia storms).]
+
+[A better future direction might be through regression models with multiple
+explanatory pathways. This allows the model to fit different weights to each. Other
+models can also be explored, like regression trees, which might have an easier time
+capturing non-linearity (if exposure is included as a continuous value, as with 
+wind and rain) and interactions when there are co-exposures to two or more hazards.
+Both of these are able (and will) either exclude or give a very low weight to an 
+explanatory variable that does not help much in explaining variation in the health 
+outcome.]
 
@@
 
+You are right, the opposite would be try in the Southern hemisphere. This
+difference is driven by the Coriolis effect on cyclonic rotation, in which the
+rotation of the earth diverts winds as they move toward the low pressure center
+of the cyclone. We have edited the text to specify that this statement is
+specific to the Northern Hemisphere:
+
+> **Text in revised manuscript (relevant addition in bold):**
+"**In the Northern Hemisphere,** extreme winds are more common to the track's
+right, where counterclockwise cyclonic winds move in concert with the tropical
+cyclone's forward motion."
+
 \begin{shaded}
@@
 
-\textit{Lines 137--139 from original manuscript:}
+\textit{Lines 185--187 from original manuscript:}
 
@@
 \noindent
-\textit{"\ldots"}
+\textit{"We used tracking data from HURDAT2, which records the storm center’s
+position at the synoptic times of 6:00 am, 12:00 pm, 6:00 pm, and 12:00 am
+Coordinated Universal Time (UTC)."}
 \end{quotation}
@@
 
+\textit{Lines 276--281 from original manuscript:}
+
 \begin{quotation}
 \noindent
-\textit{"\ldots"}
+\textit{"To identify flood- and tornado-based tropical cyclone exposures in US
+counties, we matched storm tracks with event listings from the National Oceanic
+and Atmospheric Administration (NOAA)’s Storm Event Database (NOAA NCEI 2020).
+While this database has recorded storm data, particularly tornadoes, since 1950,
+its coverage changed substantially in 1996 to cover more types of storm events,
+including flood events (NOAA NCEI 2020). We therefore only considered flood
+metrics of tropical cyclone exposure for storms in 1996 and later."}
 \end{quotation}
@@
 
+\textit{Lines 294--295 from original manuscript:}
+
 \begin{quotation}
 \noindent
-\textit{"\ldots"}
+\textit{"The tornado observations from this dataset form a traditional tornado
+event database for the US, and so we did not further validate the tornado event
+data."}
 \end{quotation}

### 2020-07-10 ehp_revision/tables/metric_definitions.tex
Start some on response

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{0.9cm}p{2.5cm}p{9cm}} 
Metric & Available & Criteria for exposure for binary analysis \\ \midrule 
Distance & 1988\,--\,2018 & County population mean center within 100 kilometers of storm track \\ 
Rain & 1988\,--\,2011 & County cumulative rainfall of 75 millimeters or more from two days before to one day after the storm's closest approach and county population mean center within 500 kilometers of the storm track \\ 
Wind & 1988\,--\,2018 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
Flood & 1996\,--\,2018 & Flood event listed in the National Oceanic and Atmospheric Administration (NOAA) Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track \\
Tornado & 1988\,--\,2018 & Tornado event listed in the NOAA Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 kilometers of the storm track\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}

### 2020-07-10 ehp_revision/tables/stormseverity.tex
Start some on response

--- 2020-02-20 tables/stormseverity.tex
@@
 % latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Feb 20 12:50:06 2020
+% Fri Feb 21 12:25:11 2020
 \begin{table}[ht]
@@
   \toprule
-Metric (years available) & Mean (interquartile range) of county exposures per year & Tropical cyclone with most counties exposures (\# exposed counties) \\ 
+Metric (years available) & Mean of county exposures per year (interquartile range) & Tropical cyclone with most counties exposed (\# exposed counties) \\ 
   \midrule
-Rain (1988--2018) & 298 (48, 441) & Frances, 2004 (464) \\ 
-  Wind (1988--2011) & 162 (55, 258) & Michael, 2018 (260) \\ 
+Rain (1988--2011) & 298 (48, 441) & Frances, 2004 (464) \\ 
+  Wind (1988--2018) & 162 (55, 258) & Michael, 2018 (260) \\ 
   Flood (1996--2018) & 197 (76, 241) & Ivan, 2004 (317) \\ 

### 2020-07-10 ehp_revision/writing/darlings_discussion.tex
Start some on response

--- 2019-10-09 writing/darlings_discussion.tex
@@
 interest or with factors associated with risk of the outcome of interest).
+
+Therefore for most tropical cyclones, since exposure assessments often differ
+across storm hazards, as well as between each storm hazard and a distance-based
+proxy measurement, exposure misclassification is a potential risk.
+
+\subsection*{Case study}
+
+In the case study, we found differences in estimates of expected rates of
+physical exposure to tropical cyclones among a population when exposure
+assessment was based on different storm exposure metrics
+(Figure~\ref{fig:topelecdependexposure}).  These results are informative
+because previous assessments of physical exposure to tropical cyclones have
+often concentrated on storm winds and storm proximity. For example, a global
+study of exposure and vulnerability to natural hazards used a combination of
+windspeed above a certain threshold and distance from the storm's track to
+assess exposure~\parencite{peduzzi2009assessing}. \textbf{[Also check other
+Peduzzi study]}  While many storm impacts might be most strongly linked to wind
+hazards, there is a growing recognition of the potential risks of adverse
+health outcomes and property damage from rain- and flood-related tropical
+cyclone hazards in the \ac{US}~\parencite{smith2009}.  Assessments based solely
+on wind and distance from the storm track therefore may underemphasize
+potentially dangerous exposures to certain regions of the \ac{US}, particularly
+inland locations.  In fact, based on our assessment of physical exposures to
+tropical cyclones among electricity-dependent Medicare beneficiaries, the
+highest expected physical exposure under any metric was measured for
+Philadelphia County, PA, under the flood metric
+(Figure~\ref{fig:topelecdependexposure}).  While this county has a
+similar-sized population of electricity-dependent Medicare beneficiaries as
+Miami-Dade County, FL (Figure~\ref{fig:topelecdependexposure}), the inland
+county of Philadelphia has frequent storm-related flood events, which results
+in a very high expected physical exposure among electricity-dependent Medicare
+beneficiaries in the county. This exposure is poorly captured by most of the
+other exposure metrics, under which Philadelphia County is not included in the
+top ten counties for expected rates of physical
+exposure~(Figure~\ref{fig:topelecdependexposure}).
+
+\textit{Previous research has highlighted the range of impacts that tropical cyclones
+can have in \ac{US} communities. However, studies of tropical cyclone exposures
+and their impacts have varied widely in how they assess exposure, and there was
+not previously a widely available public dataset that captured county-level
+exposure to multiple storm hazards. Here we provide open-source tropical
+cyclone hazard-related measurements for a large collection of historical
+Atlantic-basin cyclones, aggregated at the county level to align with the
+spatial scale at which much health-related data is collected.}
+
+\textit{Further, we investigated agreement in exposure assessment across these metrics
+when applied individually. We found large differences in which counties are
+exposed to different hazards of tropical cyclones and that distance is, at
+best, a moderate, and often a very poor, surrogate for exposure to the specific
+tropical cyclone hazards of high wind, extreme rainfall, flooding, and
+tornadoes. Use of distance as a surrogate for any of these hazards, or exposure
+assessment based on one hazard when the pathway for health impacts is in part
+or full through another storm hazard, could lead to exposure misclassification.
+In the case of tropical cyclone risk and impact studies, including
+epidemiological studies, this would result in biased estimates.  Our findings
+highlight the importance of clarifying the potential pathway from tropical
+cyclone hazards to health impacts when conducting tropical cyclone
+epidemiological studies, and then basing exposure assessment on these hazards.}

### 2020-07-10 ehp_revision/writing/darlings_intro.tex
Start some on response

--- 2019-11-07 writing/darlings_intro.tex
@@
 as~200~\si{\kilo\metre}~\parencite{mallin2006, quiring2011variations}.
+
+Recently, several major tropical cyclones have hit \ac{US} cities, including
+Hurricanes Harvey, Irma, and Maria in 2017~\parencite{blake20182017} and
+Hurricanes Florence and Michael in 2018 \parencite{avila20192018}. Researchers
+are exploring how these storms impacted human health
+(e.g.,~\cite{santos2018use, rivera2018estimating, santos2018differential,
+grineski2019impact, issa2018deaths, tanz2019notes, paul2019brief}), including
+through projects funded by \ac{NIH} Rapid Response
+grants~\parencite{nihreporter}. These projects will add to evidence
+characterizing how hurricanes impact human health, including evidence from
+
+. Some studies have
+indeed used these data within epidemiological studies to assess tropical cyclone
+exposure based on how near the storm's central track came to the community
+(e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}).

### 2020-07-10 ehp_revision/writing/darlings_methods.tex
Start some on response

--- 2019-10-09 writing/darlings_methods.tex
@@
 storm's approach.
+
+\subsection*{Case study}
+
+\textit{The societal impact of a disaster depends both on its geophysical
+forcesi, on the population exposed to the hazard, and on the vulnerability of
+those living in the geographical areas it
+affects~\parencite{chakraborty2005population, anderson2003community,
+cutter1996vulnerability}. An estimate of ``physical exposure''---which
+...---can help to jointly capture these facets of disaster impacts [ref, maybe
+the peduzzi one given below?]. As a case study, we measured physical exposure
+based on each metric of exposure and investigated whether conclusions from
+these assessments varied substantially.}
+
+We collected data on the number of electricity-dependent Medicare beneficiaries
+in each study county from the \ac{US} Department of Health \& Human Service's
+emPower Map~2.0~\parencite{empower}. This subpopulation was selected because it
+may be particularly susceptible to health impacts from tropical cyclone
+exposure, especially through the pathway of storm-associated power outages and
+evacuations. We calculated the physical exposure of the electricity-dependent
+Medicare population in each county, based on tropical cyclone assessments using
+each exposure metric (Table~\ref{tab:exposuremetrics}), following
+\citeauthor*{peduzzi2009assessing}~(\citeyear{peduzzi2009assessing}):
+
+\begin{equation}
+E_c = F_c * P_c
+\end{equation}
+
+\textit{\noindent where~$E_c$ is the average yearly physical exposure among
+electricity-dependent Medicare beneficiaries in county~$c$ to tropical cyclone
+exposures based on a given metric,~$F_c$ is the estimated yearly expected
+frequency of tropical cyclone exposures in county~$c$ based on that metric,
+and~$P_c$ is the electricity-dependent Medicare population in
+county~$c$, as of July~2017.}

### 2020-07-10 ehp_revision/writing/ehp_abstract.tex
Start some on response

--- 2020-02-21 writing/ehp_abstract.tex
@@
 	exposure based on different storm hazards. \\ 
-\textbf{Methods:} We created an open-source dataset with county-level data on
+\textbf{Methods:} We created an open-source dataset with data at the county level on
 	exposure to four tropical cyclone hazards: peak sustained wind,
@@
 	and agreement among binary exposure classifications based on these
-	metrics, as well as compared them to use of distance from the storm's track as a proxy for
-	exposure.\\ 
+	metrics, as well as compared them to use of distance from the storm's track, which has 
+	been used as a proxy for exposure in some epidemiologic studies.\\ 
 \textbf{Results:} Our open-source dataset was typically consistent with data from other sources,

### 2020-07-10 ehp_revision/writing/ehp_discussion.tex
Start some on response

--- 2020-02-20 writing/ehp_discussion.tex
@@
 The open-source data generally correspond well with data from other potential
-sources (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison}
-and~S2), but there are some caveats. The rainfall data are generally
-well-correlated with ground-based observations, but may sometimes underestimate
-very high rainfall values (Figure~\ref{fig:raincomparison}). When rainfall data
-is used to create binary exposure classifications, this disagreement is
-unlikely to influence results, as both data sources agree in identifying these
-as storms with high rainfall, but would be important to consider for cases that
-include rainfall as a continuous measurement.
+sources (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison}), but
+there are caveats. The rainfall data are generally well-correlated with
+ground-based observations, but may sometimes underestimate very high rainfall
+values (Figure~\ref{fig:raincomparison}). When rainfall data is used to create
+binary exposure classifications, this disagreement is unlikely to influence
+results, as both data sources agree in identifying these as storms with high
+rainfall, but would be important to consider for cases that include rainfall as
+a continuous measurement.
 
 The peak sustained wind estimates are based on modeled, rather than observed,
-values, and while the modeled wind data generally agree well with post-analysis
-maximum wind radii \parencite{landsea2013}
-(Figure~\ref{fig:windcomparison}), there were a few storms with some
-discrepancies. These storms---for example, Hurricane Sandy in 2012 and
-Hurricane Ike in 2008---were unusually large systems for which high winds
-persisted well inland from landfall (Figures~\ref{fig:windcomparison} and~S2).
+values, and while the modeled wind data generally agree well
+(Figure~\ref{fig:windcomparison}) with post-analysis maximum wind radii
+\parencite{landsea2013}, there were a few storms with some discrepancies. These
+storms---for example, Hurricane Sandy in 2012 and Hurricane Ike in 2008---were
+unusually large systems for which high winds persisted well inland from
+landfall (Figures~\ref{fig:windcomparison} and~S2).
 
@@
 event start dates to storm dates, many false positives would be
-captured, for which the occurrance of a storm in the midst of an ongoing storm
+captured, for which the occurrence of a storm in the midst of an ongoing 
 event might be improperly attributed to the storm.  However, distance
@@
 storm contributes meaningfully to an ongoing event, but the event is not
-captured for the storm in the exposure data.
+captured for the storm in the exposure data because its start date is not
+close in time to the storm's date.
 
@@
 hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
-cyclones bring higher risks of dangerous winds inland~\parencite{kruk2010},
+cyclones heighten risk of dangerous winds inland~\parencite{kruk2010},
 while slow-moving tropical cyclones are likely to bring more
@@
 made their first \ac{US} landfall in North Carolina at minor hurricane
-windspeeds (Category~2 and~1, respectively) and then skimmed the eastern coast
+intensity (Category~2 and~1, respectively) and then skimmed the eastern coast
 of the \ac{US} north through New England, bringing substantial rainfall to much
@@
 (although some resulting coastal flooding is captured). We are exploring ways
-to include this in future versions of the dataset, but to date we have
-focused on exposures that could affect any county, whether inland or coastal.
+to include this in future versions of the dataset, but to date we have focused
+on exposures that could affect any county, whether inland or coastal.
+
 Second, these data are aggregated to the county level. Many health outcome
 datasets are aggregated at this level, but some may be aggregated at a finer
-spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e.,
+spatial resolution (e.g., census tract or ZIP code) or unaggregated (i.e.,
 point locations for each outcome). We have published the wind model used to
@@
 at a finer spatial resolution; however, measurements of other hazards cannot
-similarly be re-scaled through tools we provide. Next, we provide these data
-and associated software tools through R packages, and so some experience in the
-R programming language is required to make full use of them. However, R is
-currently a popular programming language for environmental epidemiology,
-allowing the data to reach a large audience, and we are exploring options to
-create a web application using the Shiny platform to allow broader web-based
-access of the data \parencite{shiny2019}.  Finally, to assess patterns and
-agreement for binary exposure classifications, we have chosen one set of
-sensible thresholds for binary classifications based on continuous metrics
-(rainfall, maximum sustained surface wind, and distance from the storm's track).
-Results and conclusions would differ somewhat with other threshold choices. We
-have published code for this analysis online, allowing
-other researchers to explore other threshold choices for these
-analyses.
+similarly be re-scaled through tools we provide.
+
+Next, we provide these data and associated software tools through R packages,
+and so some experience in the R programming language is required to make full
+use of them. However, R is currently a popular programming language for
+environmental epidemiology, allowing the data to reach a large audience, and we
+are exploring options to create a web application using the Shiny platform to
+allow broader web-based access of the data \parencite{shiny2019}.
+
+Finally, to assess patterns and agreement for binary exposure classifications,
+we have chosen one set of sensible thresholds for binary classifications based
+on continuous metrics (rainfall, maximum sustained surface wind, and distance
+from the storm's track).  Results and conclusions would differ somewhat with
+other threshold choices. We have published code for this analysis online
+(\url{https://github.com/geanders/county_hurricane_exposure}), allowing other
+researchers to explore other threshold choices for these analyses.
 
@@
 storm-associated hazards (e.g., ground-based monitors, streamgages, post-storm
-wind radii estimates).  However, researchers who are planning to use the data
+wind radii).  However, researchers who are planning to use the data
 should explore the analyses presented in this paper to understand the strengths

### 2020-07-10 ehp_revision/writing/ehp_introduction.tex
Start some on response

--- 2020-02-24 writing/ehp_introduction.tex
@@
 data, allowing them to estimate average associations over many disasters and to
-explore how a disaster's characteristics, or the characteristics of the
-community it hits, modify associated health risks
+explore how a disaster's characteristics, or the characteristics of 
+\textbf{affected communities}, modify associated health risks
 (e.g.,~\cite{anderson2010heat, son2012impact, liu2017wildfire}).  Some studies
@@
 (e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
-cyclone tend to be distributed around its center non-symmetrically.  Extreme
-winds are more common to the track's right, where counterclockwise cyclonic
-winds move in concert with the tropical cyclone's forward
-motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
-associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
-right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
-is often heaviest to the left of the track, especially when the tropical
-cyclone interacts with other weather systems~\parencite{atallah2003,
-atallah2007, zhu2013variations} or undergoes an extratropical
-transition~\parencite{elsberry2002}.
+cyclone tend to be distributed around its center non-symmetrically.  \textbf{In
+the Northern Hemisphere,} extreme winds are more common to the track's right,
+where counterclockwise cyclonic winds move in concert with the tropical
+cyclone's forward motion~\parencite{halverson2015}, and almost all of the fatal
+tornadoes associated with \ac{US} tropical cyclones between~1995 and~2009
+occurred to the right of the tropical cyclone's track~\parencite{moore2012}.
+Rain, conversely, is often heaviest to the left of the track, especially when
+the tropical cyclone interacts with other weather
+systems~\parencite{atallah2003, atallah2007, zhu2013variations} or undergoes an
+extratropical transition~\parencite{elsberry2002}.
 
@@
 particularly wind, into exposure assessment (e.g., \cite{grabich2015measuring,
-zandbergen2009, czajkowski2011}), there is not yet a standard approach, and
-when different studies use different datasets or storm hazards when assessing
+zandbergen2009, czajkowski2011}), there is not yet a standard approach. 
+When different studies use different datasets or storm hazards when assessing
 storm exposure, it becomes difficult to compare and aggregate findings.

### 2020-07-10 ehp_revision/writing/ehp_tables.tex
Start some on response

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{1.5cm}p{2.5cm}p{10cm}} 
\toprule
Metric & Available & Criteria for exposure \\ \midrule 
Distance & 1988\,--\,2018 & Storm track came within 100~\si{\kilo\metre} of county population mean center\\ 
Rain & 1988\,--\,2011 & County cumulative rainfall of $\ge$75~\si{\milli\metre} over the period from two days before to 
one day after the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\ 
Wind & 1988\,--\,2018 & Modeled storm-associated peak sustained surface wind of $\ge$34~kts 
at the county's population mean center\\ 
Flood & 1996\,--\,2018 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center \\
Tornado & 1988\,--\,2018 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}

\clearpage

\input{../tables/stormseverity.tex}

### 2020-07-10 ehp_revision/R/average_exposure_figure.R

@@ -1,0 +1,336 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+library(rgdal)
+library(maptools)
+library(RColorBrewer)
+library(choroplethr)
+data(closest_dist)
+
+my_fips <- unique(closest_dist$fips)
+dist <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2018, dist_limit = 100) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / (2018 - 1988 + 1)) %>%
+        mutate(metric = "distance")
+
+rain <- county_rain(counties = my_fips, start_year = 1988,
+                             end_year = 2011, dist_limit = 500,
+                             rain_limit = 75) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / (2011 - 1988 + 1)) %>%
+        mutate(metric = "rain")
+
+wind <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2018, wind_limit = 17.4) %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / (2018 - 1988 + 1)) %>%
+        mutate(metric = "wind")
+
+flood <- county_events(counties = my_fips,
+                                start_year = 1996, end_year = 2018,
+                                event_type = "flood") %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / (2018 - 1996 + 1)) %>%
+        mutate(metric = "flood")
+
+tornado <- county_events(counties = my_fips,
+                                start_year = 1988, end_year = 2018,
+                                event_type = "tornado") %>%
+        group_by(fips) %>%
+        dplyr::summarize(n_exposures = n() / (2018 - 1988 + 1)) %>%
+        mutate(metric = "tornado")
+
+our_fips <- hurricaneexposure:::get_eastern_map() %>% dplyr::select(fips) %>% dplyr::distinct() %>%
+        dplyr::mutate(fips = as.numeric(fips))
+all_exposures <- bind_rows(dist, rain, wind, flood, tornado) %>%
+        mutate(n_exposures = n_exposures * 10,
+               fips = as.numeric(fips)) %>%
+        right_join(our_fips, by = "fips") %>%
[…]

### 2020-07-10 ehp_revision/R/extent_disagreement_figure.R

@@ -1,0 +1,267 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+
+storm <- "Frances-2004"
+
+rain_data <- filter_storm_data(storm = storm,
+                               days_included = c(-2, -1, 0, 1),
+                               include_rain = TRUE,
+                               output_vars = c("fips", "tot_precip",
+                                               "storm_dist")) %>%
+        dplyr::mutate_(rain = ~ tot_precip >= 75 &
+                               storm_dist <= 500) %>%
+        dplyr::select(fips, rain) %>%
+        dplyr::tbl_df()
+
+wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
+                              output_vars = c("fips", "vmax_sust")) %>%
+        `colnames<-`(c("fips", "wind_value")) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::select(fips, wind) %>%
+        dplyr::tbl_df()
+
+
+storm_data <- rain_data %>%
+        full_join(wind_data, by = "fips") %>%
+        gather(key = metric, value = Exposed, -fips) %>%
+        mutate(metric = factor(metric,
+                               levels = c("rain", "wind"),
+                               labels = c("Rain-based metric", "Wind-based metric")))
+
+out_data <- hurricaneexposure:::get_eastern_map() %>%
+        dplyr::left_join(storm_data, by = "fips")
+out <- ggplot2::ggplot() +
+        ggplot2::geom_polygon(data = out_data,
+                              ggplot2::aes_(x = ~ long, y = ~ lat, group = ~ group,
+                                            fill = ~ Exposed, color = ~ Exposed),
+                              size = 0.2) +
+        ggplot2::borders("state", regions = c("virginia", "north carolina", "south carolina",
+                                              "georgia", "florida", "alabama", "kentucky",
+                                              "tennessee", "maryland", "west virginia",
+                                              "district of columbia", "pennsylvania",
+                                              "new jersey", "delaware", "mississippi",
+                                              "louisiana", "texas", "oklahoma", "arkansas",
+                                              "new york", "connecticut", "rhode island",
+                                              "massachusetts", "new hampshire", "vermont",
+                                              "maine", "kansas", "missouri", "iowa", "michigan",
[…]

### 2020-07-10 ehp_revision/R/flood_comparison_figure.R

@@ -1,0 +1,122 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(countyfloods)
+library(lubridate)
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(ggplot2)
+library(viridis)
+library(scales)
+
+## Fairfield County, Connecticut 09001
+# Baltimore County, MD 24005
+## Bergen County, New Jersey 34003
+## Montgomery County, Pennsylvania 42091
+## Mobile County, Alabama 01097
+## DC 11001
+## Wake County, North Carolina 37183
+## Harris County, TX, 48201
+## Fulton County, GA 13121
+## Escambia County, FL 12033
+# Spartanburg, SC 45083
+# Prince William County, VA 51153
+# Forrest County, MS 28035
+
+county_fips <- c("37183", "42091", "01097", "13121", "12033", "34003", "48201",
+                 "24005", "09001")
+county_floods <- county_events(county_fips, start_year = 1996, end_year = 2018, event_type = "flood") %>%
+        select(storm_id, fips) %>%
+        mutate(flood = TRUE)
+county_storms <- county_distance(county_fips, start_year = 1996, end_year = 2018, dist_limit = 500) %>%
+        left_join(county_floods, by = c("fips", "storm_id")) %>%
+        mutate(flood = ifelse(is.na(flood), FALSE, flood),
+               closest_date = ymd(closest_date)) %>%
+        dplyr::select(storm_id, fips, closest_date, flood) %>%
+        rename(date_0 = closest_date) %>%
+        mutate(date_m2 = date_0 - 2,
+               date_m1 = date_0 - 1,
+               date_1 = date_0 + 1,
+               date_2 = date_0 + 2,
+               date_3 = date_0 + 3,
+               date_4 = date_0 + 4,
+               date_5 = date_0 + 5,
+               date_6 = date_0 + 6) %>%
+        gather(key = lag, value = date, -storm_id, -fips, -flood) %>%
+        mutate(lag = str_replace(lag, "date_", ""),
+               lag = str_replace(lag, "m", "-"),
+               lag = as.numeric(lag))
+
+# county_gages <- get_gages(county_fips, start_date = "1996-01-01", end_date = "2018-12-31")
+# gage_q2s <- find_q2(county_gages$site_no)
+# county_gages <- full_join(county_gages, gage_q2s, by = "site_no")
[…]

### 2020-07-10 ehp_revision/R/hurr_tracks_figure.R

@@ -1,0 +1,35 @@
+## Map of hurricane tracks
+library(ggplot2)
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+
+data(closest_dist)
+storms <- unique(closest_dist$storm_id)
+storms <- storms[gsub("*.+-", "", storms) <= 2018 &
+                         gsub("*.+-", "", storms) >= 1988]
+a <- map_tracks(storms, plot_points = FALSE, alpha = 0.3, color = "darkcyan")
+# Add storms whose names were retired
+# Source: http://www.nhc.noaa.gov/aboutnames_history.shtml
+fig <- map_tracks(c("Hugo-1989", "Bob-1991", "Andrew-1992",
+             "Opal-1995", "Fran-1996", "Georges-1998", "Mitch-1998",
+             "Floyd-1999", "Allison-2001", "Michelle-2001",
+             "Isidore-2002", "Lili-2002", "Isabel-2003",
+             "Charley-2004", "Frances-2004",
+             "Ivan-2004", "Jeanne-2004",
+             "Dennis-2005", "Katrina-2005", "Rita-2005",
+             "Wilma-2005", "Noel-2007",
+             "Gustav-2008", "Ike-2008", "Paloma-2008",
+             "Irene-2011", "Sandy-2012", "Matthew-2016",  
+             "Harvey-2017", "Irma-2017", "Nate-2017", 
+             "Florence-2018", "Michael-2018"
+             ),
+           plot_object = a, plot_points = FALSE, color = "darkcyan",
+           padding = 0) +
+        theme_void() +
+        theme(panel.background = element_rect(fill = "white")) +
+        coord_map()
+
+pdf(file = "figures/hurrtracks.pdf", width = 4.8, height = 4.0)
+print(fig)
+dev.off()
+

### 2020-07-10 ehp_revision/R/ivan_figure.R

@@ -1,0 +1,113 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+
+storm <- "Ivan-2004"
+
+distance_data <- filter_storm_data(storm = storm,
+                              output_vars = c("fips", "storm_dist")) %>%
+        dplyr::mutate_(distance = ~ storm_dist <= 100) %>%
+        dplyr::select(fips, distance) %>%
+        dplyr::tbl_df()
+
+rain_data <- filter_storm_data(storm = storm,
+                              days_included = c(-2, -1, 0, 1),
+                              include_rain = TRUE,
+                              output_vars = c("fips", "tot_precip",
+                                              "storm_dist")) %>%
+        dplyr::mutate_(rain = ~ tot_precip >= 75 &
+                               storm_dist <= 500) %>%
+        dplyr::select(fips, rain) %>%
+        dplyr::tbl_df()
+
+wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
+                             output_vars = c("fips", "vmax_sust")) %>%
+        `colnames<-`(c("fips", "wind_value")) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 15) %>%
+        dplyr::select(fips, wind) %>%
+        dplyr::tbl_df()
+
+storm_year <- gsub("*.+-", "", storm)
+counties <- hurricaneexposuredata::closest_dist %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips), quote(storm_dist))
+
+flood_data <- county_events(counties = counties$fips,
+                          start_year = storm_year,
+                          end_year = storm_year,
+                          event_type = "flood") %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips)) %>%
+        dplyr::mutate_(event = ~ 1) %>%
+        dplyr::right_join(counties, by = "fips") %>%
+        dplyr::mutate_(event = ~ !is.na(event)) %>%
+        dplyr::rename_(flood = ~ event) %>%
+        dplyr::select_(quote(-storm_dist))
+
+tornado_data <- county_events(counties = counties$fips,
+                            start_year = storm_year,
+                            end_year = storm_year,
+                            event_type = "tornado") %>%
[…]

### 2020-07-10 ehp_revision/R/jaccard_heatmap.R

@@ -1,0 +1,284 @@
+library(hurricaneexposure)
+library(hurricaneexposuredata)
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(purrr)
+library(viridis)
+library(dendextend)
+library(ComplexHeatmap)
+library(circlize)
+
+data(closest_dist)
+
+calc_jaccard <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        jaccard <- proxy::simil(t(x_mat), method = "jaccard")
+        out <- combn(colnames(x_mat), 2) %>%
+                t() %>%
+                as_tibble(.name_repair = "unique") %>%
+                dplyr::rename(metric_1 = `...1`,
+                              metric_2 = `...2`) %>%
+                dplyr::mutate(jaccard = as.vector(jaccard))
+        return(out)
+}
+
+calc_metric_1_not_2 <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        met_1_not_2 <- function(i, j){
+                x <- x_mat[ , i]
+                y <- x_mat[ , j]
+                if(sum(x | y) == 0) return(NA)
+                a <- sum(x & y)
+                b <- sum(x & !y)
+                c <- sum(!x & y)
+                metric <- b / (a + b + c)
+                return(metric)
+        }
+        out_metric <- outer(1:5, 1:5, FUN = Vectorize(met_1_not_2))
+        out <- combn(colnames(x_mat), 2) %>%
+                t() %>%
+                as_tibble(.name_repair = "unique") %>%
+                dplyr::rename(metric_1 = `...1`,
+                              metric_2 = `...2`) %>%
+                dplyr::mutate(met_1_not_2 = as.vector(t(out_metric))[t(upper.tri(out_metric))])
+        return(out)
+}
+
+
+calc_metric_2_not_1 <- function(x){
+        x_mat <- x %>%
+                dplyr::select(-fips) %>%
+                dplyr::mutate_all(funs(as.numeric(.))) %>%
+                as.matrix()
+        met_2_not_1 <- function(i, j){
+                x <- x_mat[ , i]
+                y <- x_mat[ , j]
+                if(sum(x | y) == 0) return(NA)
+                a <- sum(x & y)
[…]

### 2020-07-10 ehp_revision/R/other_top_storm_exposures.R

@@ -1,0 +1,436 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gridExtra)
+
+storm <- "Floyd-1999"
+
+distance_data <- filter_storm_data(storm = storm,
+                              output_vars = c("fips", "storm_dist")) %>%
+        dplyr::mutate_(distance = ~ storm_dist <= 100) %>%
+        dplyr::select(fips, distance) %>%
+        dplyr::tbl_df()
+
+rain_data <- filter_storm_data(storm = storm,
+                              days_included = c(-2, -1, 0, 1),
+                              include_rain = TRUE,
+                              output_vars = c("fips", "tot_precip",
+                                              "storm_dist")) %>%
+        dplyr::mutate_(rain = ~ tot_precip >= 75 &
+                               storm_dist <= 500) %>%
+        dplyr::select(fips, rain) %>%
+        dplyr::tbl_df()
+
+wind_data <- filter_wind_data(storm = storm, wind_source = "modeled",
+                             output_vars = c("fips", "vmax_sust")) %>%
+        `colnames<-`(c("fips", "wind_value")) %>%
+        dplyr::mutate_(wind = ~ wind_value >= 17.4) %>%
+        dplyr::select(fips, wind) %>%
+        dplyr::tbl_df()
+
+storm_year <- gsub("*.+-", "", storm)
+counties <- hurricaneexposuredata::closest_dist %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips), quote(storm_dist))
+
+flood_data <- county_events(counties = counties$fips,
+                          start_year = storm_year,
+                          end_year = storm_year,
+                          event_type = "flood") %>%
+        dplyr::filter_(~ storm_id == storm) %>%
+        dplyr::select_(quote(fips)) %>%
+        dplyr::mutate_(event = ~ 1) %>%
+        dplyr::right_join(counties, by = "fips") %>%
+        dplyr::mutate_(event = ~ !is.na(event)) %>%
+        dplyr::rename_(flood = ~ event) %>%
+        dplyr::select_(quote(-storm_dist))
+
+tornado_data <- county_events(counties = counties$fips,
+                            start_year = storm_year,
+                            end_year = storm_year,
+                            event_type = "tornado") %>%
[…]

### 2020-07-10 ehp_revision/R/rain_examples_figure.R

@@ -1,0 +1,108 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(ggplot2)
+library(sp)
+library(maptools)
+library(rgeos)
+
+ike_data <- filter_storm_data(storm = c("Ike-2008"), include_rain = TRUE,
+                              days_included = c(-2, -1, 0, 1),
+                              output_vars = c("storm_id", "fips", "tot_precip", "storm_dist"))
+lee_data <- filter_storm_data(storm = c("Lee-2011"), include_rain = TRUE,
+                              days_included = c(-2, -1, 0, 1),
+                              output_vars = c("storm_id", "fips", "tot_precip", "storm_dist"))
+map_data <- bind_rows(ike_data, lee_data) %>%
+        mutate(storm_id = factor(storm_id, levels = c("Ike-2008", "Lee-2011"),
+                                 labels = c("Hurricane Ike (2008)", "Tropical Storm Lee (2011)")),
+               exposed = tot_precip >= 75 & storm_dist <= 500)
+
+map_dim <- apply(matrix(c(-106.65037, 25.12993, -67.00742, 47.48101),
+                        byrow = TRUE, ncol = 2),
+                 MARGIN = 2,
+                 function(x) range(x) + c(-1, 1) * 2)
+tracks <- hurricaneexposuredata::hurr_tracks %>%
+        dplyr::select_(~ latitude, ~ longitude, ~ storm_id,
+                       ~ date) %>%
+        dplyr::filter_(~ as.character(storm_id) %in% c("Ike-2008", "Lee-2011") &
+                               longitude > map_dim[1, 1] &
+                               longitude < map_dim[2, 1] &
+                               latitude > map_dim[1, 2] &
+                               latitude < map_dim[2, 2]) %>%
+        dplyr::mutate_(date = ~ lubridate::ymd_hm(date)) %>%
+        mutate(storm_id = factor(storm_id, levels = c("Ike-2008", "Lee-2011"),
+                                 labels = c("Hurricane Ike (2008)", "Tropical Storm Lee (2011)")))
+splt_tracks <- split(tracks, tracks$storm_id)
+full_tracks <- lapply(splt_tracks, hurricaneexposure:::interp_track)
+full_tracks <- do.call("rbind", full_tracks)
+
+out_data <- hurricaneexposure:::get_eastern_map() %>%
+        dplyr::full_join(map_data, by = "fips")
+
+exposed_ike <- out_data %>%
[…]

### 2020-07-10 ehp_revision/R/severity_agreement_figure.R

@@ -1,0 +1,386 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(xtable)
+library(stringr)
+library(psych)
+
+my_fips <- unique(closest_dist$fips)
+distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2015, dist_limit = 100) %>%
+        select(storm_id, fips) %>%
+        mutate(distance_exposed = TRUE,
+               storm_id = as.character(storm_id))
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500) %>%
+        select(storm_id, fips) %>%
+        mutate(rain_exposed = TRUE,
+               storm_id = as.character(storm_id))
+wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2015, wind_limit = 15) %>%
+        select(storm_id, fips) %>%
+        mutate(wind_exposed = TRUE,
+               storm_id = as.character(storm_id))
+flood_exposure <- county_events(counties = my_fips,
+                                start_year = 1996, end_year = 2015,
+                                event_type = "flood") %>%
+        select(storm_id, fips) %>%
+        mutate(flood_exposed = TRUE,
+               storm_id = as.character(storm_id))
+tornado_exposure <- county_events(counties = my_fips,
+                                  start_year = 1996, end_year = 2015,
+                                  event_type = "tornado") %>%
+        select(storm_id, fips) %>%
+        mutate(tornado_exposed = TRUE,
+               storm_id = as.character(storm_id))
+total_exposure <- distance_exposure %>%
+        full_join(rain_exposure, by = c("storm_id", "fips")) %>%
+        full_join(wind_exposure, by = c("storm_id", "fips")) %>%
+        full_join(flood_exposure, by = c("storm_id", "fips")) %>%
+        full_join(tornado_exposure, by = c("storm_id", "fips")) %>%
+        gather(metric, exposed,
+               -storm_id, -fips, na.rm = FALSE) %>%
+        mutate(exposed = ifelse(is.na(exposed), FALSE, exposed))
+
+storm_severity <- total_exposure %>%
+        dplyr::mutate(metric = str_replace(metric, "_exposed", ""),
+               metric = str_to_title(metric)) %>%
+        dplyr::group_by(storm_id, metric) %>%
[…]

### 2020-07-10 ehp_revision/R/severity_agreement_table.R

@@ -1,0 +1,70 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(xtable)
+library(stringr)
+library(psych)
+
+my_fips <- unique(closest_dist$fips)
+distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2015, dist_limit = 100) %>%
+        select(storm_id, fips) %>%
+        mutate(distance_exposed = TRUE,
+               storm_id = as.character(storm_id))
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500) %>%
+        select(storm_id, fips) %>%
+        mutate(rain_exposed = TRUE,
+               storm_id = as.character(storm_id))
+wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2015, wind_limit = 15) %>%
+        select(storm_id, fips) %>%
+        mutate(wind_exposed = TRUE,
+               storm_id = as.character(storm_id))
+flood_exposure <- county_events(counties = my_fips,
+                                start_year = 1996, end_year = 2015,
+                                event_type = "flood") %>%
+        select(storm_id, fips) %>%
+        mutate(flood_exposed = TRUE,
+               storm_id = as.character(storm_id))
+tornado_exposure <- county_events(counties = my_fips,
+                                  start_year = 1996, end_year = 2015,
+                                  event_type = "tornado") %>%
+        select(storm_id, fips) %>%
+        mutate(tornado_exposed = TRUE,
+               storm_id = as.character(storm_id))
+total_exposure <- distance_exposure %>%
+        full_join(rain_exposure, by = c("storm_id", "fips")) %>%
+        full_join(wind_exposure, by = c("storm_id", "fips")) %>%
+        full_join(flood_exposure, by = c("storm_id", "fips")) %>%
+        full_join(tornado_exposure, by = c("storm_id", "fips")) %>%
+        gather(metric, exposed,
+               -storm_id, -fips, na.rm = FALSE) %>%
+        mutate(exposed = ifelse(is.na(exposed), FALSE, exposed))
+
+storm_severity <- total_exposure %>%
+        mutate(metric = str_replace(metric, "_exposed", ""),
[…]

### 2020-07-10 ehp_revision/R/storm_severity_table.R

@@ -1,0 +1,177 @@
+library(hurricaneexposuredata)
+library(hurricaneexposure)
+library(dplyr)
+library(tidyr)
+library(xtable)
+library(stringr)
+
+my_fips <- unique(closest_dist$fips)
+
+distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2018, dist_limit = 100) %>%
+        select(storm_id, fips) %>%
+        mutate(exposed = TRUE,
+               storm_id = as.character(storm_id))
+
+dist_sum <- distance_exposure %>%  
+  group_by(storm_id) %>% 
+  count() %>% 
+  ungroup() %>% 
+  summarize(n_storms = n() / (2018 - 1988 + 1),
+            n_exposures = sum(n) / (2018 - 1988 + 1),
+            med = median(n),
+            p25 = quantile(n, 0.25),
+            p75 = quantile(n, 0.75),
+            max = max(n),
+            which_max = storm_id[which.max(n)])
+
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500) %>%
+        select(storm_id, fips) %>%
+        mutate(exposed = TRUE,
+               storm_id = as.character(storm_id))
+rain_sum <- rain_exposure %>%  
+  group_by(storm_id) %>% 
+  count() %>% 
+  ungroup() %>% 
+  summarize(n_storms = n() / (2011 - 1988 + 1),
+            n_exposures = sum(n) / (2011 - 1988 + 1),
+            med = median(n),
+            p25 = quantile(n, 0.25),
+            p75 = quantile(n, 0.75),
+            max = max(n),
+            which_max = storm_id[which.max(n)])
+
+wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2018, wind_limit = 17.4) %>%
+        select(storm_id, fips) %>%
+        mutate(exposed = TRUE,
+               storm_id = as.character(storm_id))
+wind_sum <- wind_exposure %>%  
+  group_by(storm_id) %>% 
+  count() %>% 
+  ungroup() %>% 
+  summarize(n_storms = n() / (2018 - 1988 + 1),
+            n_exposures = sum(n) / (2018 - 1988 + 1),
+            med = median(n),
+            p25 = quantile(n, 0.25),
[…]

### 2020-07-10 ehp_revision/R/wind_comparison_figure.R

@@ -1,0 +1,62 @@
+library(hurricaneexposuredata)
+library(dplyr)
+library(stringr)
+library(ggplot2)
+library(polycor)
+library(latex2exp)
+data(storm_winds)
+data(ext_tracks_wind)
+
+wind_comp <- storm_winds %>%
+        filter(str_remove(storm_id, ".+\\-") %>%  as.numeric() <= 2018) %>% 
+        select(fips, vmax_sust, storm_id) %>%
+        rename(vmax_sust_model = vmax_sust) %>%
+        inner_join(select(ext_tracks_wind, fips, storm_id, vmax_sust),
+                  by = c("fips", "storm_id")) %>%
+        rename(vmax_sust_ext = vmax_sust) %>%
+        mutate(vmax_sust_model_cat = ifelse(vmax_sust_model < 17.4896, 0, NA),
+               vmax_sust_model_cat = ifelse(vmax_sust_model >= 17.4896 & vmax_sust_model < 25.7200,
+                                            17.4896, vmax_sust_model_cat),
+               vmax_sust_model_cat = ifelse(vmax_sust_model >= 25.7200 & vmax_sust_model < 32.9216,
+                                            25.7200, vmax_sust_model_cat),
+               vmax_sust_model_cat = ifelse(vmax_sust_model >= 32.9216, 32.9216, vmax_sust_model_cat),
+               cats_equal = vmax_sust_model_cat == vmax_sust_ext)
+
+fig <- wind_comp %>%
+        group_by(storm_id) %>%
+        summarize(perc_same = sum(cats_equal) / n(),
+                  non_zero = sum(vmax_sust_ext != 0)) %>%
+        filter(non_zero > 0) %>%
+        ggplot(aes(x = perc_same, y = reorder(storm_id, perc_same), fill = non_zero)) +
+        geom_point(size = 2, shape = 21, color = "black") +
+        scale_x_continuous(labels = scales::percent, lim = c(0, 1)) +
+        theme_classic() +
+        labs(x = "% of counties classified in same wind category\nby modeled storm wind and Extended Best Tracks",
+             y = "",
+             fill = "Number of counties with modeled\n sustained winds of <= 34 knots"
+             ) +
+        viridis::scale_fill_viridis(direction = -1, option = "B") +
+        theme(legend.position = "bottom") + 
+        theme(panel.grid.major.y = element_line(colour = "lightgray"))
+
[…]

### 2020-07-12 ehp_revision/response_to_reviewers.Rmd
Add in some on synoptic times

@@
 
+"Synoptic meteorology" refers to ...
+A key idea is that observations from a widespread geographical area can be
+integrate to provide a picture of large-scale weather [features?], including
+large areas of high or low pressure [?].
+
+> "In meteorology, the term synoptic denotes simultaneous observations taken
+globally, or at least over a large area." [@willoughby2007hurricane]
+
+> "Nearly 20 years ago [from 1937] [Bjerknes] declared that numerous and 
+important problems might be solved with the aid of observations already 
+assembled in the archives of different bureaus, but that the difficulties
+in making these observations serve a special purpose were unsurmountable
+for the individual scholar. He added that, in particular, the synoptic 
+charts representing the momentary states of the atmosphere over large
+areas of the earth, in all the necessary details and exactitude, were 
+excessively difficult to assemble. His proposal was that there be entered
+at the central bureau of each country, on the appropriate maps, all the available
+observations taken in that country and that there be employed, by the different
+bureaus, maps that may be superimposed or juxtaposed." [@gregg1937international]
+
+However, to integrate data from many sources to provide inputs for broadscale,
+synoptic analysis, it is helpful to use some standards in collecting these data
+at different locations. To facilitate numerical weather forecasting and other
+synoptic meteorology, weather data are regularly collected at standardized
+times, so that the data can be more easily integrated as an input to models.
+This practice extends back over 100 years, when data were collected from many
+[offices?] in the US by telegram [?] at regular times during the day and used to
+create synoptic weather maps that captured major weather systems in the country.
+
+> "In order to allow comparison of surface conditions at the same time throughout 
+the world, the groups regulating the collection of hydrometeorological data have
+established a standard time for taking observations. These observations will have
+times observed within the 10 minutes preceding the UTC standard time. Hence 
+surface observations for meteorological use have timestamps reflecting 0000, 
+0300, 0600, 0900, 1200, 1500, 1800, 2100 UTC." [@usnwstime]
+
+> "Since the end of World War II, observers have launched rawinsonde observation
+worldwide at 00 and 12 UTC (coordinated universal time). High-threat situations, 
+such as impending TC landfall, may dictate observations at 6-h or even 3-h intervals. 
+Interpolating rawinsonde observations to the model's mesh and starting the calculation
+at a synoptic time is the simplest way to prepare a model initial condition."
+[@willoughby2007hurricane]
+
+> "Because simple interpolation in space and time is a less-than-optimum way to 
+use observations, nonsynoptic data require elaborate four-dimensional variational
+data assimilation schemes." [@willoughby2007hurricane]
+
+> "By 1941, charts for the 750, 500, and 250 mbar surfaces were being prepared daily
+for 0400, 1200, and 2000 GMT." [@grahame2000development]
+
+> "During the period considered here, NHC issued advisories 3 h after the standard
+synoptic time (0300, 0900, 1500, and 2100 UTC)." [@torn2012uncertainty]
+
+> "The operation data assimilation and forecast cycle interval is 6 h" [@aberson201010]
+
+> "Dan Petersen asked whether a data deficiency really exists with all the 
+observational sources in the Atlantic region, or are the data not being
+properly used in the numerical models? ... Steve Lord emphasized the difficulty of the 
+task because many of these observations are not obtained at the synoptic times..."
+[@elsberry1992there]
+
+> "In the modern era, forecasters rely on purely dynamical models that integrate the 
+Navier-Stokes equations for atmospheric motions to represent both the storm itself and
+its surroundings. Numerical models are structured on a computational grid with 
+temperature, moisture, and wind tabulated in more-or-less rectangular cells." 
+[@willoughby2007hurricane]
+
+> "Predicting the weather required a widespread but highly coordinated team of 
+observers and forecasters. Organization was key... They transmitted data from the 
+field using forms and protocols adapted from those that Abbe had developed while 
+forecasting the weather from Cincinnati, Ohio. The information arrived by telegraph
+coded to maximize accuracy and minimize the word count and changes. The reports piled
+up in a rush at the appointed hours." [@willis2006cleveland]
+
+> "Three times daily the duty forecaster broadcast a synopsis of prevailing weather
+conditions nationwide..." [@willis2006cleveland]
+
+> "The requirement of synoptic data collection is easily met by the
+'conventional' observing system now serving as the main source of meteorological 
+data throughout the world, since it is based on human observations which could 
+be scheduled at convenient synoptic times." [@morel1971initialization]
+
+Since large-scale weather features are very larger, relevant data for a synoptic
+map often comes from locations with different local times. The need for
+synchronized weather data collection therefore actually played an important role
+in establishing standardized time zones in the United States in [year]. Synoptic
+weather times are typically set based on the Universal Time Constant, with times 
+recorded in Coordinated Universal Time (i.e.,
+Greenwich ...), and time stamps on weather maps and data based on this time zone
+often is given a "Z" (for "Zulu", an earlier name for this timezone) after the
+time.
+
+> "Accuracy in forecasting depended on the accurate timing of observations. At
+first weather observers used local time because no system of standard time
+existed, either in the United States or abroad. Local time often meant railroad
+time, and that could vary as much as 20 minutes from the time proper to the
+local meridian. Abbe called for change, emphasizing the need for simultaneous
+readings by Signal Service observers in order to construct accurate charts and
+weather maps. In 1875, he recommended to the president of the American
+Meteorological Society the establishment of a committee on standard time. ...
+Abbe's report in 1879 recommended the system of meridians that is now in general
+use ... a system of time meridians one hour apart across the country, anchored
+on the Greenwich prime meridian." [@willis2006cleveland]
+
+> "From Abbe's perspective, however, the full potential for dynamic weather
+prediction would not be achieved without standard time worldwide."
+[@willis2006cleveland]
+
+> "The time in all hydrometeorological products is expressed according to 
+a single standard, which is the Universal Coordinated Time (UTC) 
+(formerly known as Greenwich Mean Time [GMT] or Zulu (zero) Time [Z])."
+[@usnwstime]
+
+> "The U.S. National Weather Service base all data and product times on the 
+UTC standard. ... A common practice is to include a time of the observation 
+or event denoting the time with a following capital letter 'Z' to indicate the 
+time as UTC." [@usnwstime]
+
+We have added text to clarify the meaning of "synoptic times" in the text:
+
+> **Text in revised manuscript (relevant change in bold):**
+"We used tracking data from HURDAT2, which records the storm center’s position
+at **four standardized times for weather data collection (synoptic times),**
+6:00 am, 12:00 pm, 6:00 pm, and 12:00 am Coordinated Universal Time (UTC)."
+
 \begin{shaded}
@@
 **Response:**
+
+The NOAA Storm Events database originated as a database for recording tornadoes
+[? anything else] in the US in 195[x], and was originally called ... . The
+database has gone through a number of expansions to cover more events, and it
+now includes reports of other types of [hydrometeorological?] disaster events,
+including floods, extreme heat and cold, wildfires, frost, hail, ... . However,
+it has maintained its status as the database of record [?] for tornadoes in the
+United States.
+
+Examples of research papers that leverage this database as a record of tornado
+events, or the version of it shared by the [Severe Storm ...] through their
+website [ref], include ... .
+
+[Is this database used to validate tornado forecasts / warnings?]
 

### 2020-07-12 ehp_revision/writing/ehp_methods.tex
Add in some on synoptic times

--- 2020-02-24 writing/ehp_methods.tex
@@
 used tracking data from \ac{HURDAT2}, which records the storm center's position
-at the synoptic times of 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We
-interpolated this position to~15-\si{\minute} intervals using natural cubic
-spline interpolation~\parencite{hurricaneexposure}. At each~15-\si{\minute}
-interval, we measured the distance between the storm's center and each county's
-population mean center~\parencite{countycenters}. We took
-the minimum distance for each county as a measure of how close the storm came
-to the county over its lifetime.  We also recorded the time when this occurred
-for each county so we could later link observed data on precipitation,
-flooding, and tornadoes.  To allow matching with data recorded in
-local time (e.g., health data), we converted these times from \ac{UTC} to local
-time~\parencite{countytimezones}.
+at four standardized times for weather data collection (synoptic times),
+6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We interpolated this
+position to~15-\si{\minute} intervals using natural cubic spline
+interpolation~\parencite{hurricaneexposure}. At each~15-\si{\minute} interval,
+we measured the distance between the storm's center and each county's
+population mean center~\parencite{countycenters}. We took the minimum distance
+for each county as a measure of how close the storm came to the county over its
+lifetime.  We also recorded the time when this occurred for each county so we
+could later link observed data on precipitation, flooding, and tornadoes.  To
+allow matching with data recorded in local time (e.g., health data), we
+converted these times from \ac{UTC} to local time~\parencite{countytimezones}.
 

### 2020-07-18 ehp_revision/response_to_reviewers.Rmd
Get started on response

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+```
+
+```{r}
+library(knitr)
 ```
@@
 hazards experienced directly at their homes.
+
+There is already some discussion on this topic in the Discussion [exposure 
+measurement error / misclassification].
+
+There are two interrelated mechanisms at play regarding this question. The
+first is the potential for ecological bias. This bias results from inferring
+individual-level associations from data that is available at an aggregated level. 
+In the case of epidemiological research, these data include health outcomes, 
+exposure, and covariates if the model is adjusting for confounding from them.
+For example, in tropical cyclone epidemiology, an ecological study might 
+compare county-wide rates of COPD hospitalizations to county-wide measures
+of tropical cyclone exposure while controlling for county-level smoking 
+rates and age distribution. The model in this case is fit at the county level, 
+and so the relevant inference is at that level. If this county-level estimate
+is inferred to represent the individual-level association between tropical 
+cyclone exposure and risk of COPD hospitalization, controlled for smoking and
+age, then the possibility of ecological bias arises. Given that this bias
+results from inferring from a model with data at one level of aggregation
+to another, this bias is also called cross-level bias.
+
+There are some subtleties that arise regarding this question for disaster
+epidemiology. First, it has been argued that, for some ambient exposures
+like weather and air pollution, contextual effects (that is, at the community
+level) are helpful to estimate in their own right, not just as a potential 
+surrogate for individual-level associations. This is because regulations
+(in the case of pollutants), interventions, and planning often happen at
+this level.
+
+Second, for a disaster, the relevant exposure might be not just at the
+individual level (e.g., winds or flooding at the individual's residence), but
+also throughout a broader area surrounding the individual. If storm damage
+closes roads in a county, for example, or damages the power grid, then these
+could create pathways for health risks for individuals throughout the county,
+whether damage was high at their residence or not. The causal pathways for
+tropical cyclones to affect human health differ from those for a dangerous
+substance, like air pollutants, in which the substance itself must enter the
+body to cause harm. While some health risk comes directly from the storm (e.g.,
+deaths and injuries from trees falling on homes or drowning from flooding),
+there are many more pathways that are indirect. These include pathways that go
+through the way that the storm's damage affects community infrastructure and
+access to medical care. Several papers have discussed the potential for 
+"individualistic fallacy", as a counterpoint to "ecological fallacy", when 
+such contextual-level exposure pathways are ignored.
+
+There are study designs that can be used to leverage ecological data while 
+minimizing risk from ecological bias. For example, potential confounders
+like age distribution and smoking rates vary much less within a county over
+time than comparing between counties. The mechanism for ecological bias 
+depends on the joint distribution between individual exposure, outcome, and 
+covariates, if any are included in the model. Time series-style study designs, 
+which compare a county to itself over time, allow for very similar covariate
+distributions between exposure and non-exposure. Other studies add to this
+design by stabilizing for temporal confounding through the addition of 
+counties that were never exposed, allowing for a differences-in-differences
+style approach to calibrate for seasonal or longer-term trends that might
+otherwise create confounding. For example, many health outcomes have a strong
+seasonal trend, with peak rates in the winter and lows in the summer. Since
+the hurricane season stretches from summer into fall, a study design that compares
+the rate of a health outcome in an exposed county to the rate two weeks before
+the exposure might be biased away from the null, since baseline rates of the
+health outcome will typically be moving up over most of the hurricane season.
+
+Also, there may be confounders that are relevant at the contextual, rather than
+individual level, as well as modifiers. Example: whether the county is coastal
+could be a contextual-level confounder and effect modifier. This will influence
+whether the county is exposed to that storm or not, since storms usually weaken
+rapidly when the center is over land. In terms of counfounding pathways, coastal
+communities might tend to have lower levels of air pollution, because sea
+breezes clear the pollution regularly. They might also be wealthier on average,
+since property on or near the beach is desirable. Finally, they might be better
+prepared for or more hardened against tropical cyclones at the community-wide
+level (e.g., through hardier power infrastructure, more rigorous building codes,
+higher likelihood of evacuating in advance of a threatening storm) compared to
+nearby inland counties.
+
+Other approaches have also been suggested to reduce ecological bias while 
+leveraging ecological data, which might be available at a much larger scale---and
+so provide more power---than is practical in collecting individual data. These
+include multi-level approaches in which individual-level data from a sample
+of individuals in each community is incorporated.
+
+The second issue is exposure measurement error/misclassification. When the 
+exposure (e.g., binary exposure or continuous measure of intensity of exposure)
+is only measured at the county level, all the individuals in that county are 
+assigned that same exposure. However, the hazards included in this study will 
+vary across each county. For some hazards, this variation will be more stark. 
+For example, tornadoes can be very local, causing extreme damage to a row of 
+homes and none to the homes across the street. Some types of floods can also 
+be very localized, rather than creating widespread flooding throughout 
+the county. In both cases, when a county-level exposure assessment is made
+based on an event in the county in association with the storm, there may be
+many individuals in the county who did not experience that hazard at their 
+residence. The other hazards we included---storm-associated winds and rainfall---will
+also have some variation across counties, but this will often not be as stark. 
+The rainfields for tropical cyclones are very large, and while there are rainbands
+within the storm that might have particularly high rates of precipitation, these
+progess over the course of the storm, and it is unlikely that a county will have
+one area that experienced very extreme precipitation while another experienced
+very little. Winds might have more variation, especially in coastal counties and 
+counties near the storm's central track, since winds decay quickly as you move
+from the storm's center, and storms typically weaken rapidly once they make 
+landfall and can no longer draw on warm ocean waters for energy.
 
@@
 et al. 2012; Liu et al. 2017)."
+
+We also used the word "Hits per county per decade" in the legend for Figure 5 in
+the main text and Figure S3 of the Supplemental Material. We have revised the
+wording in these figures to "Exposures per county per decade".
+
+```{r fig5, fig.cap="Reproduction of Figure 5 from the original manuscript (left) and revised version in the resubmitted manuscript (right).
+In the revised version, the figure legend was changed from '\\textbf{Hits} per county per decade' to '\\textbf{Exposures} per county per decade' (bold used to highlight change).
+Figure caption in the revised manuscript: \\textit{Figure caption.}", fig.show = "hold", out.width="40%", fig.align = "center"}
+include_graphics("../figures/averageexposureonly.pdf")
+include_graphics("figures/averageexposureonly.pdf")
+```
 
@@
 
+> "The extent and the hour of the observations changed over time. Originally the
+observers took the readings at 0735, 1635, and 2335, Washington time, but the
+last reading was soon changed to 2300 so that the information could be included
+in the morning newspapers. Before the introduction of standard time and time
+zones within the United States, a confusing multiplicity of local times existed.
+In 1870 there were over 100 such regional times. For the purposes of the
+meteorological observations, observers used Washington time until 1885.
+Thereafter the observers took their readings according to eastern time, or that
+at the 75th meridian west of Greenwich, England. Observers also made local-time
+readings from 1876 to 1881.[30]" [@raines1996getting]
+
+> "Once the observers had gathered the weather data, the means of reporting and
+disseminating it became most important. Like the Smithsonian, the chief signal
+officer made arrangements with the leading commercial telegraph companies to
+carry the tri-daily reports. Civilian telegraph experts established special
+circuits routed to the Signal Office. The initial arrangement with Western Union
+regarding transmission was only temporary, and at the end of the trial period
+the company refused to continue service. The House Appropriations Committee held
+hearings over the dispute and ruled that the company had a mandate to transmit
+the weather information as government business. The Signal Corps compensated the
+company, however, at rates determined by the postmaster general.[34]" [@raines1996getting]
+
+> "When making the daily telegraphic reports, the weather observers used special
+codes to reduce their length to twenty words in the morning and ten in each of
+the other two reports, thereby saving the government both time and money.[35]
+Regular transmission of the reports began at 0735 on 1 November 1870 from
+twenty-four stations stretching from Boston, Massachusetts, south to Key West,
+Florida, and west to Cheyenne in the Wyoming Territory. In addition to the
+station atop Mount Washington (opened in December 1870), the Signal Corps soon
+reached new heights in weather reporting with the station on Pikes Peak that
+began reporting in November 1873.[36]" [@raines1996getting]
+
+> "To provide a picture of weather conditions across the country, the observers
+made their reports as nearly simultaneous as possible. The weather service did
+not initially make forecasts, and the enabling legislation did not specifically
+call for it to do so. Eventually general forecasts, referred to as
+probabilities, emanated from the Signal Office in Washington. Locally, the
+observers posted bulletins and maps in the offices of boards of trade and
+chambers of commerce to provide weather information to the public. Post offices
+also displayed daily bulletins, and observers supplied local newspapers with
+data. Some communities appointed meteorological committees to confer with the
+chief signal officer and to serve as a check upon the operations of the local
+weather station. On the national level, the Signal Office in Washington issued
+daily weather maps compiled from the reports received from all the stations. It
+also published the Daily Weather Bulletin, Weekly Weather Chronicle, and the
+Monthly Weather Review. All were available for sale to the public. Myer
+estimated that through these various means at least one third of American
+households received the Signal Corps' weather information in some form. A
+railway bulletin service, initiated in 1879, enabled stations along many major
+railroads to display weather information.[37]" [@raines1996getting]
+
+> "The Signal Corps' telegraph network soon expanded beyond the eastern
+seaboard. In 1874 Congress enacted legislation directing the War Department to
+build lines to connect military posts and protect frontier settlements in Texas
+against Indians and Mexicans. Other acts authorized lines in Arizona and New
+Mexico and, somewhat later, the Northwest. These lines were intended to serve
+sparsely settled areas where commercial lines were not yet available. For the
+most part, soldiers maintained and operated the lines, but the Corps employed
+some civilians. As the telegraph extended its reach, the weather system also
+grew, because the operators doubled as weather observers." [@raines1996getting]
+
+> "Until 1934 the Weather Bureau offices operated 12--15 hours a day with two
+basic observations taken at 8 a.m. and 8 p.m. The observations were transmitted
+via telegraph. There were no satellite images and few upper air observations."
+\url{https://www.weather.gov/dvn/armistice_day_blizzard}
+
 We have added text to clarify the meaning of "synoptic times" in the text:
@@
 \begin{shaded}
-\textbf{R2 Comment 3:}
+\textbf{R3 Comment 3:}
 Several of the exposures are calculated for the population mean center of each
@@
 \begin{shaded}
-\textbf{R2 Comment 4:}
+\textbf{R3 Comment 4:}
 In the methods section starting on line 276, the authors describe the flooding
@@
 \begin{shaded}
-\textbf{R2 Comment 5:}
+\textbf{R3 Comment 5:}
 Line 294: What is meant by the phrase "traditional tornado event database"?
@@
 
-\begin{shaded}
-\textbf{R2 Comment 6:}
+This database has evolved through several stages but has, since its earliest
+iteration in the 1950s, served as the National Weather Service's [?] data of
+record on tornado events. While there is also a database maintained by the 
+[Severe Storm ...], the two databases have different names but almost 
+identical information once data are aggregated to a daily, county-level 
+measure of whether there was a tornado recorded in that county on that date. 
+[Difference between the two and why there are two.]
+
+There is no other tornado base at the same temporal and spatial scale that
+we could use for validation.
+
+This database serves as the basis for much of the large-scale (e.g., assessing
+patterns over the US over long time periods) tornado-based research for the US. 
+It does have some well-documented limitations. One is a notable increase over 
+time in the number of tornadoes reported, particularly for the weakest, F0 [?]
+tornadoes. [This stabilizes some in the age of radar?]
+
+\begin{shaded}
+\textbf{R3 Comment 6:}
 Table 1: I suggest providing the rationale for picking the thresholds for
@@
 
-\begin{shaded}
-\textbf{R2 Comment 7:}
+Wind: gale force?
+
+Precipitation:
+
+Distance:
+
+\begin{shaded}
+\textbf{R3 Comment 7:}
 Figure 3: The x-axis refers to one of the two wind data sources as "Extended
@@
 **Response:**
+
+We have revised the x-axis on the figure to clarify that the comparison is based
+on the wind radii dataset described in the text, as requested. The original and
+revised versions of this paper are reproduced in this response document as
+Figure \@ref(fig:fig3).
+
+[Also revise figure legend?]
+
+```{r fig3, fig.cap="Reproduction of Figure 3 from the original manuscript (left) and revised version in the resubmitted manuscript (right).
+In the revised version, the x-axis has been labelled to clarify that this axis is showing results based on the wind radii dataset described in the text, in response to Reviewer 3's Comment 7.
+The revised version also uses a more focused range for the x axis, as requested in EHP's Request 1.
+Figure caption in the revised manuscript: \\textit{Figure caption.}", fig.show = "hold", out.width="50%"}
+include_graphics("../figures/windcomparison.pdf")
+include_graphics("figures/windcomparison.pdf")
+```
 
@@
 **Response:**
+
+[Typo here? I don't think there's a Figure S7.]
+
+We have revised this figure to show an x-axis truncated at the minimum observed 
+value ($\sim75\%$), as requested. The original and revised versions of this paper 
+are reproduced in this response document as Figure \@ref(fig:fig3).
 
@@
 
+In all these figures, we have used a color scale that is both accessible to readers
+with the most common forms of impaired color vision and can be easily distinguished 
+when the figure is printed in black and white.
+
+For these figures, we selected a color palette that was created to meet both these
+criteria. [More on viridis]
+
+In Figure [x] of this response document, we have reproduced the current version of 
+each figure as it would appear to someone with [vision impairment].
+
+In Figure [x] of this response document, we have reproduced the current version of 
+each figure as it will appear when printed using grayscale (i.e., with most black and 
+white printers).
+
+```{r fig5check1, fig.cap="Reproduction of revised version of Figure 5 to check accessibility to readers with two types of colorblindness, deuteranopia (left) and protanopia (right).", out.width="90%", fig.align = "center"}
+include_graphics("figures/averageexposureonly_check1.pdf")
+```
+
+```{r fig5check2, fig.cap="Reproduction of revised version of Figure 5 to check accessibility to readers with a third type of colorblindness, tritanopia (left) and for printing the figure in grayscale (right).", out.width="90%", fig.align = "center"}
+include_graphics("figures/averageexposureonly_check2.pdf")
+```
+
 \begin{shaded}
@@
 **Response:**
+
+As we note in our response to the previous comment, we used a color palette
+designed to be perceptible both for those with common types of color blindness
+and when printed in grayscale. We have increased the size of the circles in
+Figure 3 of the revised manuscript (reproduced in Figure \@ref(fig:fig3) of this
+response) so that the colors are clearer, as well as truncated the x-axis as
+requested in EHP Request 1. Also, we show in Figures \@ref(fig:fig3check1) and
+\@ref(fig:fig3check2) of this response how the revised Figure 3 would look with
+three common types of color blindness as well as when printed in black and
+white. The gradient is clear when this figure is printed in black and white.
+
+```{r fig3check1, fig.cap="Reproduction of revised version of Figure 3 to check accessibility to readers with two types of colorblindness, deuteranopia (left) and protanopia (right).", out.width="100%"}
+include_graphics("figures/windcomparison_check1.pdf")
+```
+
+```{r fig3check2, fig.cap="Reproduction of revised version of Figure 3 to check accessibility to readers with a third type of colorblindness, tritanopia (left) and for printing the figure in grayscale (right).", out.width="100%"}
+include_graphics("figures/windcomparison_check2.pdf")
+```
 

### 2020-07-18 ehp_revision/R/average_exposure_figure.R

@@ -100,9 +100,9 @@ out <- ggplot2::ggplot() +
                          colour = "black", fill = NA, size = 0.2, alpha = 0.5) +
         ggplot2::theme_void() +
         ggplot2::coord_map() +
-        viridis::scale_fill_viridis(name = "Hits per\ncounty per\ndecade",
+        viridis::scale_fill_viridis(name = "Exposures\nper county\nper decade",
                                     discrete = TRUE, direction = -1, option = "A") +
-        viridis::scale_color_viridis(name = "Hits per\ncounty per\ndecade",
+        viridis::scale_color_viridis(name = "Exposures\nper county\nper decade",
                                      discrete = TRUE, direction = -1, option = "A") +
         facet_wrap(~ metric, ncol = 1) +
         theme(legend.position = "bottom",
@@ -218,13 +218,92 @@ pdf("figures/averageexposure.pdf", width = 9, height = 11)
 grid.arrange(out, out2, ncol = 1)
 dev.off()
 
-pdf("figures/averageexposureonly.pdf", width = 3, height = 9)
+pdf("ehp_revision/figures/averageexposureonly.pdf", width = 3, height = 9)
 print(out)
 dev.off()
 
 pdf("figures/ivanonly.pdf", width = 6, height = 3.3)
 print(out2)
 dev.off()
+
+
+#### Check for accessibility
+
+library(dichromat)
+library(colorspace)
+library(gridExtra)
+
+out <- ggplot2::ggplot() +
+        ggplot2::geom_polygon(data = out_data,
+                              ggplot2::aes_(x = ~ long, y = ~ lat, group = ~ group,
+                                            fill = ~ value, color = ~ value),
+                              size = 0.2) +
+        ggplot2::borders("state", regions = c("virginia", "north carolina", "south carolina",
+                                              "georgia", "florida", "alabama", "kentucky",
+                                              "tennessee", "maryland", "west virginia",
+                                              "district of columbia", "pennsylvania",
+                                              "new jersey", "delaware", "mississippi",
+                                              "louisiana", "texas", "oklahoma", "arkansas",
+                                              "new york", "connecticut", "rhode island",
+                                              "massachusetts", "new hampshire", "vermont",
+                                              "maine", "kansas", "missouri", "iowa", "michigan",
+                                              "illinois", "ohio", "wisconsin", "indiana"),
[…]

### 2020-07-18 ehp_revision/R/wind_comparison_figure.R

@@ -28,10 +28,10 @@ fig <- wind_comp %>%
                   non_zero = sum(vmax_sust_ext != 0)) %>%
         filter(non_zero > 0) %>%
         ggplot(aes(x = perc_same, y = reorder(storm_id, perc_same), fill = non_zero)) +
-        geom_point(size = 2, shape = 21, color = "black") +
-        scale_x_continuous(labels = scales::percent, lim = c(0, 1)) +
+        geom_point(size = 3, shape = 21, color = "black") +
+        scale_x_continuous(labels = scales::percent) +
         theme_classic() +
-        labs(x = "% of counties classified in same wind category\nby modeled storm wind and Extended Best Tracks",
+        labs(x = "% of counties classified in same wind category\nby modeled storm wind and based on HURDAT2 wind radii",
              y = "",
              fill = "Number of counties with modeled\n sustained winds of <= 34 knots"
              ) +
@@ -39,7 +39,7 @@ fig <- wind_comp %>%
         theme(legend.position = "bottom") + 
         theme(panel.grid.major.y = element_line(colour = "lightgray"))
 
-ggsave("figures/windcomparison.pdf", fig, width = 5.25, height = 8.5, units = "in")
+ggsave("ehp_revision/figures/windcomparison.pdf", fig, width = 5.25, height = 8.5, units = "in")
 
 wind_comp_ordinal <- wind_comp %>%
         mutate(vmax_sust_ext = factor(vmax_sust_ext, ordered = TRUE,
@@ -60,3 +60,52 @@ wind_comp_ordinal %>%
         theme_classic() +
         labs(x ="Polychoric correlation between wind\ncategorization within counties",
              y = "")
+
+# Check figure for accessibility
+
+library(dichromat)
+library(colorspace)
+library(gridExtra)
+
+fig_check <- wind_comp %>%
+        group_by(storm_id) %>%
+        summarize(perc_same = sum(cats_equal) / n(),
+                  non_zero = sum(vmax_sust_ext != 0)) %>%
+        filter(non_zero > 0) %>%
+        ggplot(aes(x = perc_same, y = reorder(storm_id, perc_same), fill = non_zero)) +
+        geom_point(size = 3, shape = 21, color = "black") +
+        scale_x_continuous(labels = scales::percent) +
+        theme_classic() +
+        labs(x = "% of counties classified in same wind category\nby modeled storm wind and based on HURDAT2 wind radii",
+             y = "",
[…]

### 2020-07-19 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 \begin{shaded}
-\textbf{R2 Comment 2:}
+\textbf{EHP Request 2:}
 Main text figures (general)
@@
 
-In Figure [x] of this response document, we have reproduced the current version of 
-each figure as it would appear to someone with [vision impairment].
-
-In Figure [x] of this response document, we have reproduced the current version of 
-each figure as it will appear when printed using grayscale (i.e., with most black and 
-white printers).
+We have shown an example with one of the figures in the manuscript that uses
+this scale. In Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2) of this
+response document, we have reproduced the current version of Figure 5 as it
+would appear to someone with three common types of color blindness as well as
+when printed using grayscale (i.e., with most black and white printers).
 
@@
 \begin{shaded}
-\textbf{R2 Comment 3:}
+\textbf{EHP Request 3:}
 Figure 3: Please check the contrast between the colors used to indicate numbers
@@
 \begin{shaded}
-\textbf{R2 Comment 4:}
+\textbf{EHP Request 4:}
 Figure 4: Even with nominally normal color vision and a large high-resolution
@@
 \begin{shaded}
-\textbf{R2 Comment 5:}
+\textbf{EHP Request 5:}
 Figures 5--6: Please check to see if the colors used can be distinguished if
@@
 
+As we explain in our response to EHP Request 2, these figures' color palette is
+color-blind friendly and also remains perceptible when printed in grayscale.
+Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check1) show what Figure 5 would
+look like under three common types of color blindness and when printed in black
+and white. Figures \@ref(fig:fig6check1) and \@ref(fig:fig6check2) show the same
+thing for Figure 6. Therefore, these figures should not require any exemption
+from accessibility requirements, as they should satisfy these requirements.
+
+```{r fig6check1, fig.cap="Reproduction of revised version of Figure 6 to check accessibility to readers with two types of colorblindness, deuteranopia (left) and protanopia (right).", fig.show = "hold", out.width="50%"}
+include_graphics("figures/jaccard_heatmap_check1.pdf")
+
+include_graphics("figures/jaccard_heatmap_check2.pdf")
+```
+
+```{r fig6check2, fig.cap="Reproduction of revised version of Figure 6 to check accessibility to readers with a third type of colorblindness, tritanopia (left) and for printing the figure in grayscale (right).", fig.show = "hold", out.width="50%"}
+include_graphics("figures/jaccard_heatmap_check3.pdf")
+include_graphics("figures/jaccard_heatmap_check4.pdf")
+```
+
 \begin{shaded}
@@
 
+We have provided the Supplemental Material in the Word file format in our
+revised submission.
+
 # References for the response {-#references-for-the-response}

### 2020-07-19 ehp_revision/R/jaccard_heatmap.R

@@ -282,3 +282,138 @@ draw(heat_map + ha_row, padding = unit(c(2, 2, 18, 2), "mm"), annotation_legend_
 # decorate_annotation("first_metric", {grid.text("Exposure metric #1", unit(-1, "mm"), just = "right", gp = gpar(fontsize = 8, font = 2))})
 # decorate_annotation("second_metric", {grid.text("Exposure metric #2", unit(-1, "mm"), just = "right", gp = gpar(fontsize = 8, font = 2))})
 dev.off()
+
+
+### Check for accessibility
+
+library(dichromat)
+library(colorspace)
+
+ha_row <- HeatmapAnnotation(counties_exposed = counties_exposed$counties_exposed,
+                            show_annotation_name = FALSE,
+                            annotation_legend_param = list(counties_exposed = list(title = "\n# of\ncounties\nexposed\nby any\nmetric",
+                                                                                   title_gp = gpar(fontsize = 9,
+                                                                                                   lineheight = 0.7,
+                                                                                                   font = 2))),
+                            col = list(counties_exposed = colorRamp2(c(100, 275, 450, 625),
+                                                                     viridis::viridis(4, option = "D", direction = -1) %>% 
+                                                                             dichromat(type = "deutan"))),
+                            which = "row", width = unit(0.25, "cm"))
+heat_map <- Heatmap(for_heatmap, col = viridis(256, option = "A", direction = -1) %>% 
+                            dichromat(type = "deutan"),
+                    cluster_rows = row_dend, split = 4, show_row_dend = FALSE,
+                    row_names_side = "left", row_names_gp = gpar(fontsize = 8),
+                    show_column_names = FALSE, show_column_dend = FALSE, #column_dend_side = "bottom",
+                    column_dend_height = unit(6, "mm"),
+                    name = "Jaccard\nindex\nbetween\nmetrics",
+                    heatmap_legend_param = list(title_gp = gpar(fontsize = 9, lineheight = 0.7, font = 2)),
+                    # bottom_annotation = ha_col,
+                    top_annotation = ha_rot_cn,
+                    rect_gp = gpar(col = "white", lwd = 1),
+                    row_dend_reorder = TRUE, 
+                    row_title = NULL, 
+                    column_title = "Green-Blind (Deuteranopia)")
+
+pdf(file = "ehp_revision/figures/jaccard_heatmap_check1.pdf", height = 8.4, width = 5)
[…]

### 2020-07-21 ehp_revision/response_to_reviewers.Rmd
Work some more on revisions

@@
 worldwide that could be used in a similar way.
+- For tornadoes, there may be national databases in other spots similar to the tornado
+database that we use, but as far as we're aware, no international one. 
+- Same for floods. 
 - For wind, the wind model could be generalizable with some tweaks. These include: 
@@
   storm basins.
+  + Our wind model was derived with US data and so may need bias-correction or a shift
+  to a West Pacific--based model
+
+> "Tornadoes long have been recognized as a global phenomenon (e.g., Wegener 1917; 
+Feuerstein et al. 2005), having been recorded in every continent except Antarctica. 
+As such, numerous nationas share an interest in improving assessment of their 
+damage. International research and involvement in tornado survey work and damage
+and intensity scales is well underway." [@edwards2013tornado]
 
@@
 side of a state border.
+
+There are some cases where damage reports are directly used to infer the intensity 
+of the weather event. For example, damage surveying is used following a tornado 
+to assign the tornado's class on the [enhanced] Fujita scale. However, this is not the
+intent of FEMA damage reports, which instead are used to determine post-disaster
+aid [?] and are influenced by a number of political factors.
+
+> "Because of the historic lack of direct measurements and remotely sensed tornado 
+wind speeds at or near group level, damage surveying has remained the most common 
+form for indicating tornado strength. ... The occurrence of a direct tornado strict 
+upon a fixed, sufficiently sturdy, and well-calibrated wind measuring station is quite
+rare. Only 31 direct in situ tornado observations are evident between 1894 and 2011."
+[@edwards2013tornado]
 
@@
 
-You are right, the opposite would be try in the Southern hemisphere. This
+You are right, the opposite would be true in the Southern hemisphere. This
 difference is driven by the Coriolis effect on cyclonic rotation, in which the
@@
 > **Text in revised manuscript (relevant addition in bold):**
-"**In the Northern Hemisphere,** extreme winds are more common to the track's
-right, where counterclockwise cyclonic winds move in concert with the tropical
-cyclone's forward motion."
+"**In the Northern Hemisphere, cyclonic winds are counterclockwise,** so
+extreme winds are more common to the track's right where cyclonic winds move in
+concert with the tropical cyclone's forward motion."
 
@@
 
+This is correct---the population centers from the US 2010 Decennial Census were 
+used throughout. We have added text to the manuscript to clarify this:
+
+> **Text in revised manuscript (relevant addition in bold):**
+"At each 15- interval, we measured the distance between the storm's
+center and each county's population mean center, **as of the 2010 US Decennial
+Census** (US Census Bureau 2020)."
+
 \begin{shaded}
@@
 
+> "Weather events listed in the publication Storm Data are considered to be
+the official database for tornadoes by the National Weather Service." 
+[written by someone at the NWS Storm Prediction Center] [@mccarthy2003nws]
+
+The US National Weather Service's Storm Prediction Center has a tornado database
+called the "National Tornado Database" [@center2020storm]. This uses the NOAA
+Storm Events database as its primary source [@center2020storm]. They note on their
+website:
+
+> "The tables below provide the links to comma separated values (.csv) files for
+tornadoes, hail, and damaging winds, as compiled from *NWS Storm Data*. Tornado 
+reports exist back to 1950 while hail and damaging wind events date from 1955.
+
+The NOAA Storm Events database, or the Storm Prediction Center's version of the
+tornado database have been used in a number of studies, including in a number of
+studies of tornado climatology [@tippett2016more; @brooks2003climatological;
+@strader2015climatology].
+
+> "The dataset we will use is the so-called smooth log of severe-weather reports
+collected by the SPC and archived in the National Oceanic and Atmospheric 
+Administration publication *Storm Data*." [@brooks2003climatological]
+
+> "Data such as tornado magnitude, path length, path width, and so on are recorded
+by the National Climatic Data Center (NCDC) as a public service and for use
+in meteorological, climatological, and engineering studies (Edwards et al. 2013)."
+[@strader2015climatology]
+
+Sometimes used in conjunction with other NWS data products:
+
+> "Initially, tornado cases available in the NWS's Damage Assessment Toolkit
+(NOAA, 2014) were employed, which is a GIS-based framework for collecting, storing, 
+and retrieving damage survey data (Camp et al., 2014). This toolkit provides a 
+variety of tornado event filtering (tornado survey point, track, footprint, and/or
+swath) and download options [key mark-up language (.kml) or shapefile (.shp)], supplying
+an initial sample of georeferenced damage assessments. In this study, a tornado 
+footprint is defined as the maximum areal extent of tornado intensity inferred by 
+the damage, wind speed measured directly by mobile Doppler radar, or assessed
+theoretically as the recorded length multiplied with the maximum width as reported in 
+Storm Data." [@strader2015climatology]
+
+> "To serve the public interest and the National Climatic Data Center (NCDC) storm 
+data record, the National Weather Service (NWS) documents the path length, width, 
+and maximum damage rating for every tornado county segment (NOAA 2007).[Footnote
+in manuscript: County segments of tornado paths are then combined at the NWS Storm
+Prediction Center to yield a unified one-tornado (ONETOR) dataset of whole-tornado 
+records (Schaefer and Edwards 1999).] Such data are used in meteorological and 
+climatological research, as well as in determining construction standard for 
+critical infrastructure such as high-tension electric lines and nuclear power plants
+(e.g., Ramsdell et al. 2007)." [@edwards2013tornado]
+
+Sometimes used in conjunction with other datasets:
+
+> "The contiguous U.S. tornado fatality dataset utilized in this study was 
+transcribed and compiled from two primary sources: (1) a long-term study of 
+U.S. tornados by Grazulis (1993, 1997, hereafter Grazulis dataset) and 
+(2) the National Climatic Data Center's Storm Data (NCDC 1959--2005) and 
+'Storm Events' datavase. The Grazulis dataset included 'significant' tornado
+events, including all events that are known to have produced a fatality, 
+from 1680 to 1995. Storm Data reports from 1959 to 2005 were utilized to 
+supplement the existing record of killer tornado events documented by 
+Frazulis." [in a study of tornado fatalities] [@Ashley2007]
+
+History:
+
+> "During the early 1950s, the US Weather Bureau began a concerted effort to 
+count all US tornadoes. In 1950, a list of tornadoes was provided in the 
+*Climatological Data National Summary*, but it was not until 1953, the first 
+full year of the issuance of Weather Bureau tornado watches, that the agency 
+began to formally count all tornadoes. The *Climaticalogical Data National Summary*
+document evolved into *Storm Data* by 1959, which, in addition to tornado 
+data, included information on other storm perils, such as severe hail, high 
+winds, and floods (Grazulis 1993). Over the years, the tornado dataset 
+has been formatted and adjusted by the National Severe Storms Forecast 
+Center and Storm Prediction Center (SPC), and currently the National Tornado 
+Database is administered by the NWS Headquarters, SPC, and NCDC 
+(McCarthy 2003)." [@Ashley2007]
+
 \begin{shaded}
@@
 
-In all these figures, we have used a color scale that is both accessible to readers
-with the most common forms of impaired color vision and can be easily distinguished 
-when the figure is printed in black and white.
-
-For these figures, we selected a color palette that was created to meet both these
-criteria. [More on viridis]
-
-We have shown an example with one of the figures in the manuscript that uses
-this scale. In Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2) of this
+In all these figures, we have used colormaps that are both accessible to readers
+with several forms of color vision deficiency, including the most common type
+(deuteranomaly, a type of red-green color vision deficiency), and can be easily
+distinguished when the figure is printed in black and white [@viridis;
+@van2015mpl].
+
+For these figures, we selected colormaps that were created to meet both these
+criteria. These colormaps are all multi-hue sequential maps that avoid red-green
+contrasts, while maximizing dynamic range in comparison to something
+single-hued, like grayscale [@nunez2018optimizing; @viridis]. They are all
+perceptually uniform across the scale for those with normal color vision
+[@liu2018somewhere] and are close to perceptually uniform for those with common
+forms of color deficiency [@nunez2018optimizing]. They reproduce their original
+sequential gradient when printed in grayscale and so are perceptable when
+printed on a black-and-white printer [@van2015mpl; @nunez2018optimizing].
+Previous research has found that colormaps in this family allow viewers to more
+quickly and accurately judge relative distances in scientific figures
+[@liu2018somewhere]. The family of colormaps we used in the figures is currently
+considered the gold standard for scientific figures [@nunez2018optimizing].
+
+For example, in Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2) (left) of this
 response document, we have reproduced the current version of Figure 5 as it
-would appear to someone with three common types of color blindness as well as
-when printed using grayscale (i.e., with most black and white printers).
-
-```{r fig5check1, fig.cap="Reproduction of revised version of Figure 5 to check accessibility to readers with two types of colorblindness, deuteranopia (left) and protanopia (right).", out.width="90%", fig.align = "center"}
+would appear to someone with three common types of color vision deficiencies:
+deuteranomaly (defective green cone cells, the most common type), as well as
+protanopia (defective red cone cells) and tritanopia (defective blue cone
+cells). We have also shown how the Figure would appear in grayscale in
+\@ref(fig:fig5check2) (right). We have included similar versions of other 
+figures from the manuscript in response to later requests from EHP in this response.
+
+```{r fig5check1, fig.cap="Reproduction of revised version of Figure 5 to check accessibility to readers with two types of color vision deficiency, deuteranopia (left) and protanopia (right).", out.width="90%", fig.align = "center"}
 include_graphics("figures/averageexposureonly_check1.pdf")
@@
 
-```{r fig5check2, fig.cap="Reproduction of revised version of Figure 5 to check accessibility to readers with a third type of colorblindness, tritanopia (left) and for printing the figure in grayscale (right).", out.width="90%", fig.align = "center"}
+```{r fig5check2, fig.cap="Reproduction of revised version of Figure 5 to check accessibility to readers with a third type of color vision deficiency, tritanopia (left) and for printing the figure in grayscale (right).", out.width="90%", fig.align = "center"}
 include_graphics("figures/averageexposureonly_check2.pdf")
@@
 
-As we note in our response to the previous comment, we used a color palette
-designed to be perceptible both for those with common types of color blindness
-and when printed in grayscale. We have increased the size of the circles in
-Figure 3 of the revised manuscript (reproduced in Figure \@ref(fig:fig3) of this
-response) so that the colors are clearer, as well as truncated the x-axis as
-requested in EHP Request 1. Also, we show in Figures \@ref(fig:fig3check1) and
-\@ref(fig:fig3check2) of this response how the revised Figure 3 would look with
-three common types of color blindness as well as when printed in black and
-white. The gradient is clear when this figure is printed in black and white.
-
-```{r fig3check1, fig.cap="Reproduction of revised version of Figure 3 to check accessibility to readers with two types of colorblindness, deuteranopia (left) and protanopia (right).", out.width="100%"}
+As we note in our response to EHP Request 2, we used colormaps for all
+figures that are designed to be perceptible both for those with common types of
+color vision deficiency and when printed in grayscale. We have increased the
+size of the circles in Figure 3 of the revised manuscript (reproduced in Figure
+\@ref(fig:fig3) of this response) so that the colors are clearer, as well as
+truncated the x-axis as requested in EHP Request 1. Also, we show in Figures
+\@ref(fig:fig3check1) and \@ref(fig:fig3check2) of this response how the revised
+Figure 3 would look with three common types of color vision deficiency as well
+as when printed in black and white. The gradient is clear when this figure is
+printed in black and white.
+
+```{r fig3check1, fig.cap="Reproduction of revised version of Figure 3 to check accessibility to readers with two types of color vision deficiency, deuteranopia (left) and protanopia (right).", out.width="100%"}
 include_graphics("figures/windcomparison_check1.pdf")
@@
 
-```{r fig3check2, fig.cap="Reproduction of revised version of Figure 3 to check accessibility to readers with a third type of colorblindness, tritanopia (left) and for printing the figure in grayscale (right).", out.width="100%"}
+```{r fig3check2, fig.cap="Reproduction of revised version of Figure 3 to check accessibility to readers with a third type of color vision deficiency, tritanopia (left) and for printing the figure in grayscale (right).", out.width="100%"}
 include_graphics("figures/windcomparison_check2.pdf")
@@
 
-As we explain in our response to EHP Request 2, these figures' color palette is
-color-blind friendly and also remains perceptible when printed in grayscale.
-Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check1) show what Figure 5 would
-look like under three common types of color blindness and when printed in black
-and white. Figures \@ref(fig:fig6check1) and \@ref(fig:fig6check2) show the same
-thing for Figure 6. Therefore, these figures should not require any exemption
-from accessibility requirements, as they should satisfy these requirements.
-
-```{r fig6check1, fig.cap="Reproduction of revised version of Figure 6 to check accessibility to readers with two types of colorblindness, deuteranopia (left) and protanopia (right).", fig.show = "hold", out.width="50%"}
+As we note in our response to EHP Request 2, we used colormaps for all figures
+that are designed to be perceptible both for those with common types of color
+vision deficiency and when printed in grayscale Figures \@ref(fig:fig5check1)
+and \@ref(fig:fig5check1) show what Figure 5 would look like under three common
+types of color vision deficiency and when printed in black and white. Figures
+\@ref(fig:fig6check1) and \@ref(fig:fig6check2) show the same thing for Figure
+6. Therefore, these figures should not require any exemption from accessibility
+requirements, as they should satisfy these requirements.
+
+```{r fig6check1, fig.cap="Reproduction of revised version of Figure 6 to check accessibility to readers with two types of color vision deficiency, deuteranopia (left) and protanopia (right).", fig.show = "hold", out.width="50%"}
 include_graphics("figures/jaccard_heatmap_check1.pdf")
@@
 
-```{r fig6check2, fig.cap="Reproduction of revised version of Figure 6 to check accessibility to readers with a third type of colorblindness, tritanopia (left) and for printing the figure in grayscale (right).", fig.show = "hold", out.width="50%"}
+```{r fig6check2, fig.cap="Reproduction of revised version of Figure 6 to check accessibility to readers with a third type of color vision deficiency, tritanopia (left) and for printing the figure in grayscale (right).", fig.show = "hold", out.width="50%"}
 include_graphics("figures/jaccard_heatmap_check3.pdf")

### 2020-07-21 ehp_revision/writing/ehp_introduction.tex
Work some more on revisions

@@
 cyclone tend to be distributed around its center non-symmetrically.  \textbf{In
-the Northern Hemisphere,} extreme winds are more common to the track's right,
-where counterclockwise cyclonic winds move in concert with the tropical
-cyclone's forward motion~\parencite{halverson2015}, and almost all of the fatal
-tornadoes associated with \ac{US} tropical cyclones between~1995 and~2009
-occurred to the right of the tropical cyclone's track~\parencite{moore2012}.
-Rain, conversely, is often heaviest to the left of the track, especially when
-the tropical cyclone interacts with other weather
+the Northern Hemisphere, cyclonic winds are counterclockwise,} so extreme
+winds are more common to the track's right where cyclonic winds move in
+concert with the tropical cyclone's forward motion. extreme winds are more
+common to the track's right, where counterclockwise cyclonic winds move in
+concert with the tropical cyclone's forward motion~\parencite{halverson2015},
+and almost all of the fatal tornadoes associated with \ac{US} tropical cyclones
+between~1995 and~2009 occurred to the right of the tropical cyclone's
+track~\parencite{moore2012}.  Rain, conversely, is often heaviest to the left
+of the track, especially when the tropical cyclone interacts with other weather
 systems~\parencite{atallah2003, atallah2007, zhu2013variations} or undergoes an

### 2020-07-21 ehp_revision/writing/ehp_methods.tex
Work some more on revisions

@@
 we measured the distance between the storm's center and each county's
-population mean center~\parencite{countycenters}. We took the minimum distance
+population mean center, **as of the 2010 US Decennial Census**~\parencite{countycenters}. We took the minimum distance
 for each county as a measure of how close the storm came to the county over its

### 2020-07-21 DraftExposurePaper/dade_storm_plot.R

@@ -1,0 +1,168 @@
+options("noaakey" = Sys.getenv("noaakey"))
+library(countyweather)
+library(tidyr)
+library(dplyr)
+library(hurricaneexposure)
+library(hurricaneexposuredata)
+library(lubridate)
+library(data.table)
+library(ggplot2)
+library(ggthemes)
+data("closest_dist")
+
+
+# Create a new dataset for precip with lag -5 to 3 for fips 12086 only
+
+check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
+        dplyr::mutate(closest_date = ymd(closest_date)) %>%
+        dplyr::rename(day_0 = closest_date) %>%
+        dplyr::mutate(fips = as.integer(fips),
+                      day_0 = day_0 + days(0),
+                      day_b1 = day_0 - days(1),
+                      day_b2 = day_0 - days(2),
+                      day_a1 = day_0 + days(1),
+                      day_a2 = day_0 + days(2),
+                      day_a3 = day_0 + days(3)) %>%
+        dplyr::select(storm_id, fips, day_b2, day_b1,
+                      day_0, day_a1, day_a2, day_a3) %>%
+        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
+        dplyr::rename(date = day) %>%
+        dplyr::filter(fips == "12086") %>%
+        dplyr::mutate(fips = as.character(fips))
+        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
+
+all_dates <- unique(check_dates$date)
+all_fips <- unique(check_dates$fips) # has Miami as "12086", and is still in check_dates here
+
+#create timeseries for Dade precipitation data from the `countyweather` package
+
+fips <- "12086"
+# I've commented this out because we're saving the data, so we only need
+# to run it once.
+# county_timeseries(fips, percent_coverage = 0,
+#                   date_min = "1988-01-01", date_max = "2011-12-31",
+#                   var = "PRCP",
+#                   out_directory = "dade_data/")
+#                   #out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")
+
+# miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
+dade_dir <- "dade_data/"
+dade_weather <- readRDS(paste0(dade_dir, fips, ".rds"))
+
[…]

### 2020-07-21 ehp_revision/R/rain_comparison_figure.R

@@ -283,3 +283,63 @@ county_rain_compare_data(ex_fips = "37183", ex_dir = "../DraftExposurePaper/wake
         arrange(desc(prcp)) %>%
         select(storm_id, tot_precip, prcp) %>%
         slice(1:3)
+
+# Correlations at high rain values
+
+
+check_high_corr <- function(ex_fips, ex_dir,
+                                     get_data = FALSE){
+        if(get_data){
+                write_daily_timeseries(ex_fips, coverage = 0,
+                                       date_min = "1988-01-01",
+                                       date_max = "2011-12-31",
+                                       var = "PRCP",
+                                       out_directory = ex_dir,
+                                       keep_map = FALSE)
+        }
+
+        check_dates <- closest_dist %>%
+                dplyr::filter(fips == ex_fips) %>%
+                dplyr::select(-storm_dist) %>%
+                dplyr::mutate(closest_date = ymd(closest_date)) %>%
+                dplyr::rename(day_0 = closest_date) %>%
+                dplyr::mutate(fips = as.integer(fips),
+                              day_0 = day_0 + days(0),
+                              day_b1 = day_0 - days(1),
+                              day_b2 = day_0 - days(2),
+                              day_a1 = day_0 + days(1)) %>%
+                dplyr::select(storm_id, day_b2, day_b1, day_0, day_a1) %>%
+                tidyr::gather(key = lag, value = day, -storm_id) %>%
+                dplyr::rename(date = day)
+        all_dates <- unique(check_dates$date)
+
+        ex_weather <- readRDS(paste0(ex_dir, ex_fips, ".rds"))
+
+        ex_weather <- ex_weather %>%
+                dplyr::filter(date %in% all_dates) %>%
+                dplyr::right_join(check_dates, by = "date") %>%
+                dplyr::group_by(storm_id) %>%
+                dplyr::summarize(prcp = sum(prcp),
+                                 ave_n = mean(prcp_reporting))
+
+        ex_rain <- county_rain(counties = ex_fips,
+                               start_year = 1988, end_year = 2011,
+                               rain_limit = 0, dist_limit = 500) %>%
+                full_join(ex_weather, by = "storm_id") %>%
+                filter(!is.na(prcp) & !is.na(tot_precip)) %>%
+                mutate(prcp = prcp / 10) ## Units for countyweather are now 10ths
[…]

### 2020-07-21 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 
-[Why we think this is happening... Re-analysis product uses a model to integrate
-data from multiple sources and so might be likely to pull values back toward 
-the "typical", and so oversmooth very extreme values.]
-
-Here is the text and analysis we have added on this point in the revised manuscript:
+We have added estimates of these correlations specifically for high-precipitation
+events in a new table in the Supplemental Material (Table S1, reproduced 
+as Table \@ref(tab:highprecipcorr) of this response document).
+
+\input{tables/precip_high_corr.tex}
 

### 2020-07-21 ehp_revision/tables/precip_high_corr.tex
Work some more on response

\begin{table}

\caption{\label{tab:highprecipcorr}Precipitation correlation during all versus high-precipitation events.
                     The same sample of counties is shown as in Figure ... of the main text.
                     Events are cases where a tropical cyclone came within ... km of each of the 
                     listed counties. The number of total events gives the sum of all points
                     shown on the main plot for the county in Figure ... of the main text. 
                     The Spearman correlation for all events is the same as that shown in Figure ...
                     of the main text. High-precipitation events are those for which storm-associated
                     precipitation was 75 mm or higher based on at least one of the two measures
                     considered in this comparison (NLDAS and monitors). The Spearman correlation
                     for these high-precipitation events is given in the last column of the 
                     table.}
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{All events} & \multicolumn{2}{c}{High-precipitation events} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
County & \makecell[c]{Number\\of events} & \makecell[c]{Spearman\\correlation} & \makecell[c]{Number\\of events} & \makecell[c]{Spearman\\correlation}\\
\midrule
Miami-Dade, FL & 65 & 0.94 & 18 & 0.49\\
Harris, FL & 38 & 0.93 & 10 & 0.84\\
Mobile, AL & 50 & 0.95 & 20 & 0.57\\
Orleans, LA & 55 & 0.89 & 13 & 0.95\\
Fulton, GA & 48 & 0.95 & 12 & 0.69\\
\addlinespace
Charleston, SC & 73 & 0.94 & 17 & 0.65\\
Wake, NC & 61 & 0.98 & 12 & 0.84\\
Baltimore, MD & 33 & 0.92 & 5 & 0.70\\
Philadelphia, PA & 52 & 0.96 & 6 & 0.77\\
\bottomrule
\end{tabular}
\end{table}

### 2020-07-21 ehp_revision/R/rain_comparison_figure.R

@@ -338,8 +338,67 @@ check_high_corr <- function(ex_fips, ex_dir,
                          nldas = ex_rain$prcp) %>% 
                 as_tibble() %>% 
                 dplyr::filter(monitor >= 75 | nldas >= 75)
-        cor(rain_df$monitor, rain_df$nldas, method = "spearman")
+        print(paste(ex_fips, ex_dir))
+        print(paste("Spearman cor, high events:", 
+                    round(cor(rain_df$monitor, rain_df$nldas, method = "spearman"), 2)))
+        print(paste("N events:", nrow(rain_df)))
+        print(paste("Spearman cor, all events:", round(rain_rho, 2)))
+        print(paste("N all events:", nrow(ex_rain)))
 }
 
 
+
+check_high_corr(ex_fips = "12086", ex_dir = "../DraftExposurePaper/dade_data/data/") 
+check_high_corr(ex_fips = "48201", ex_dir = "../DraftExposurePaper/harris_data/data/") 
+check_high_corr(ex_fips = "01097", ex_dir = "../DraftExposurePaper/mobile_data/data/")
 check_high_corr(ex_fips = "22071", ex_dir = "../DraftExposurePaper/orleans_data/data/") 
+check_high_corr(ex_fips = "13121", ex_dir = "../DraftExposurePaper/fulton_data/data/") 
+check_high_corr(ex_fips = "45019", ex_dir = "../DraftExposurePaper/charleston_data/data/")
+check_high_corr(ex_fips = "37183", ex_dir = "../DraftExposurePaper/wake_data/data/") 
+check_high_corr(ex_fips = "24005", ex_dir = "../DraftExposurePaper/baltimore_data/data/") 
+check_high_corr(ex_fips = "42101", ex_dir = "../DraftExposurePaper/philadelphia_data/data/") 
+
+# Based on running previous functions
+high_corr_table <- tribble(
+        ~ county, ~ n_events, ~ corr_events, ~ n_high_events, ~ cor_high_events,
+        "Miami-Dade, FL", "65", "0.94", "18", "0.49", 
+        "Harris, FL", "38", "0.93", "10", "0.84", 
+        "Mobile, AL", "50", "0.95", "20", "0.57", 
+        "Orleans, LA", "55", "0.89", "13", "0.95", 
+        "Fulton, GA", "48", "0.95", "12", "0.69", 
+        "Charleston, SC", "73", "0.94", "17", "0.65", 
[…]

### 2020-07-22 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 
+This is a great suggestion. We have made this change, as well as several changes
+in response to EHP's Request 4 later in this response.
+
+The original and revised versions of Figure 4 of the paper are reproduced in
+Figure \@ref(fig:fig4) of this response.
+
+```{r fig4, fig.cap="Reproduction of Figure 4 from the original manuscript (top) and revised version in the resubmitted manuscript (bottom).
+In the revised version, we have added a title to the y-axis ('Classification based on NOAA Storm Events Data'), binned the values shown by each point's fill, so these colors are easier to distinguish, increased the size of the points, and changed the algorithm for jittering, so none of the points overlap.", fig.show = "hold", fig.align = "center", out.width="100%"}
+include_graphics("../figures/floodcomparison.pdf")
+include_graphics("figures/floodcomparison.pdf")
+```
+
 \begin{shaded}
@@
 monitor, it is difficult for me to make out the color differences in the
-symbols-- almost all appear black or very dark blue, a dozen or so look yellow,
+symbols---almost all appear black or very dark blue, a dozen or so look yellow,
 and a handful look dark grey. Is there any way to increase the contrast in these
@@
 **Response:**
+
+Thank you for these suggestions.
+
+For the original figure, most of the events resulted in 0% of flood gages
+exceeding the flood threshold, especially for events that were not recorded as
+floods based on NOAA Storm Events. Therefore, most of the points in the original
+were indeed dark blue, representing 0%, but we can see that readers might have
+questioned if this was the case or if the scale made it harder to see subtler
+differences.
+
+We have reproduced both the original and revised versions of Figure 4 as Figure
+\@ref(fig:fig4) of this response. We have made several changes to make this
+figure easier for readers to see and interpret. These changes include:
+
+- Binning the values shown by the points fill, so that readers only need to 
+be able to distinguish among four colors, rather than across a larger scale.
+- Increasing the size of the points, so the fill in each point is easier to 
+see. 
+- Using a different algorithm to "jitter" the points, so there are no points that
+overlap each other.
 

### 2020-07-22 ehp_revision/R/flood_comparison_figure.R

@@ -63,6 +63,7 @@ county_gage_data <- county_gages %>%
         inner_join(county_flow_data, by = "site_no") %>%
         mutate(flood_usgs = discharge > flood_val)
 
+library(ggbeeswarm)
 out <- county_gage_data %>%
         mutate(discharge = ifelse(discharge == -999999, NA, discharge)) %>%
         group_by(site_no) %>%
@@ -101,22 +102,35 @@ out <- county_gage_data %>%
                               labels = c("No flood event", "Flood event")),
                county_cd = paste0(as.character(county_cd), " (", n_gages, ")"),
                county_cd = factor(county_cd)) %>%
+        mutate(perc_flood_cat = case_when(
+                perc_flooded == 0 ~ "0%",
+                perc_flooded < 0.5 ~ "1 to 49%",
+                perc_flooded < 1 ~ "50 to 99%", 
+                perc_flooded == 1 ~ "100%", 
+                TRUE ~ NA_character_
+        )) %>% 
+        mutate(perc_flood_cat = fct_relevel(perc_flood_cat, 
+                                            "0%", "1 to 49%", "50 to 99%", "100%")) %>% 
         ggplot(aes(x = flood, y = max_total)) +
-        geom_boxplot(color = "darkgray", fill = NA, outlier.alpha = 0, width = 0.3) +
-        geom_point(aes(fill = perc_flooded), color = "black", shape = 21, size = 1.7,
-                   position = position_jitter(width = 0.1)) +
+        #geom_point(aes(fill = perc_flood_cat), color = "black", shape = 21, size = 1.9,
+        #           position = position_jitter(width = 0.15)) +
+        geom_boxplot(color = "black", fill = NA, outlier.alpha = 0, width = 0.4) +
+        geom_quasirandom(aes(fill = perc_flood_cat), 
+                         color = "black", shape = 21, size = 1.9, alpha = 0.7) +
         facet_wrap(~ county_cd, ncol = 3, scales = "free_x") +
         theme_classic() +
-        labs(x = "", 
+        labs(x = "Classification based on NOAA Storm Events Data", 
              y = bquote('Total discharge across county streamflow gages '~(ft^3/s))) +
         coord_flip() +
         scale_y_log10(label = scales::comma) +
         theme(legend.position = "bottom") +
-        scale_fill_viridis(name = "% streamflow gages over threshold for flooding",
[…]

### 2020-07-23 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 and intensity scales is well underway." [@edwards2013tornado]
+
+For flood and tornado events, we used a database of severe weather events that
+is specific to the United States (the US NOAA's Storm Events database). To 
+expand our exposure data for these hazards internationally, we would need to find
+similar databases for other affected countries. 
+[international flood and tornado databases? EM-DAT? Others?]
+
+The wind exposure data is based on modeling wind fields based on records of
+the storm's track and central intensity throughout the tracking period. These
+values are tracked for tropical cyclones around the world, and so it would be
+possible to expand the dataset to encompass other countries for wind exposure
+measurements. However, there are some practical issues that would need to 
+be addressed to do so with the approach we've taken.
+
+The software we are using to model wind [@stormwindmodel] has several features that
+make it specific to Atlantic basin storms. First, some steps in the model depend
+on whether the storm's center or the location being modeled is over land or water. 
+These steps use a landmask---a grid with a binary indicator of land or sea by 
+latitude and longitude. Currently, the software only provides a landmask that 
+covers the United States, the Gulf of Mexico, and areas of the Atlantic Ocean 
+near the US. Second, the model assumes that cyclonic winds are counterclockwise, 
+and so is valid only for the Northern Hemisphere. Third, the model assumes that
+the central intensity of the storm is based on [wind averaging period], which is 
+the standard used in the tracking data for Atlantic-basin storms, but is not
+a common standard around the world. Instead, some basins record central intensity
+measurements using averaging periods of ....
+
+Therefore, it would be relatively straightforward to expand the wind exposure data
+to cover other countries affected by Atlantic-basin storms, although this would 
+require some expansion of the area covered by our current landmask. More modifications
+would need to be made to the wind model software before it could be extended
+to storms in other basins, particularly basins that use different wind averaging 
+periods for their main tracking data and basins in the Southern Hemisphere. 
+Alternatively, other wind modeling software could be used for this exposure 
+assessment, as other groups have developed software that is more extensible to 
+multiple storm basins [Australian group's software].
+
+The rain exposure data would also be fairly easy to extend to cover other countries, 
+although again there would be a few practical limitations to extend the approach
+we've taken in developing the dataset covered in this manuscript. For this exposure, 
+we've used a re-analysis dataset. We've taken some steps to translate this gridded data
+to the county scale, so it can be integrated with the scale for many human impacts
+measurements. However, these steps have been well documents [ref] and would be easy to
+reproduce. The re-analysis dataset that we use is specific to the contiguous US [ref
+on NLDAS], but other similar re-analysis datasets exist for other locations [example
+of one with global coverage?].
+
+For all of these, we have provided exposure assessments at the county scale. Any of the 
+steps we have taken could be done for other countries at a similar geopolitical scale, 
+although the name of that areal unit will vary by country (e.g., municipalities in Mexico, 
+districts in India). For wind and rain exposures, the approach we take here could be
+used with any areal unit, as long as there are geographical shape files available for
+the area boundaries (although the modeling for wind may be to the area's geographic
+mean center rather than population mean center).
+
+> "We initially developed software to model exclusively Atlantic-basin storms
+[@tormwindmodel]. There are two characteristics of Northwestern Pacific Ocean
+basin storm data that prevent direct use of the current version of the software.
+First, the software currently requires Western Hemisphere longitude inputs.
+Second, the ``Best Tracks'' data from the China Meteorological Administration
+include central tropical cyclone intensity using a 2-minute averaging period,
+while the software currently assumes that central intensities are given for a
+1-minute averaging period, as is the case for storm track data from the US
+National Hurricane Center. With the assistance of Ms. Schumacher, we will update
+our tropical cyclone wind model software to appropriately handle 'Best Tracks' data from
+the China Meteorological Administration, incorporating conversion factors
+[@harper2010guidelines] to convert central storm maximum sustained wind
+measurements to a common averaging period prior to applying further steps of the
+wind model."
 
@@
 
+It is obvious that it is typically better, but in cases where the proxy is much more
+practical to measure, it may need to be more than just a bit better, but 
+instead substantially better. Also, there could be some cases where the proxy 
+might actually be better---for example, if the proxy can be measured with 
+very low level, while the direct exposure measure might only be measurable 
+with high error. In that case, the proxy could be the better metric and might
+provide a more precise estimate of the association with health risk.
+
+[Env. Epi. book---advantages and disadvantages of different types of measures]
+
+We agree that it is helpful to clarify for readers that these independent hazards
+can come together in a storm event at different intensities and at different time
+windows. We have added some text on this:
+
+[Added text]
+
+However, we disagree that this should be down as an alternative to calculating
+agreement. Epidemiologists need to understand how these exposure measurements
+agree to help in planning studies and statistical approaches to interpreting the
+resulting data. In this sense, "agreement" is providing a measure similar to
+correlation for continuous variables; our choice of agreement with the binary
+exposure classifications reflects the more common use of binary classifications
+for most tropical cyclone epidemiology to date.
+
+Epidemiological studies might focus on tropical cyclone exposures at two levels.
+First, they might wish to focus on a specific hazard. For example, a study might
+hypothesize that mold growth following flooding from a tropical storm might
+increase risk of respiratory health outcomes in the months after a storm. In
+this case, the clear causal pathway recommends a focus on one of the storm
+hazards, rather than all storm hazards. However, other storm hazards could
+confound the association between flood exposure and respiratory health risk if
+the other hazards are ignored *and* if they are associated with both the
+exposure of interest (storm-associated flooding) and the health outcome. For
+example, if wind exposure assessments are in high agreement with flood exposure
+assessments for storms, and if wind affects risk of respiratory outcomes by
+causing extensive property damage and creating extreme psychological stress,
+which can have repercussions on respiratory health, then the wind hazard could
+confound the estimated association between storm-related flood exposure and the
+health outcome.
+
+This bias would be higher when the confounder and exposure of interest are more
+strongly correlated (for continuous variables) or in higher agreement (for
+categorical variables). Conversely, if the exposure and the potential confounder
+are, in practice, only weakly associated, then the potential for confounding is
+low. The Jaccard index is one measure of similarities between categorical
+(including binary) coefficients.
+
+Another consideration is collinearity. If one hazard is the exposure of interest
+and another hazard is a potential confounder, then the model should adjust the
+confounder or it will be misspecified and be prone to bias. However, if the 
+exposure of interest and the potential confounder are very strongly correlated/in
+agreement, then including both as independent variables in a regression model
+often causes numerical problems in fitting the regression model, including 
+instability in the estimated coefficients and inflated variance. In very extreme
+cases, the statistical software may not be able to invert the matrix [? only for
+linear regression?].
+
+Therefore, it is helpful for epidemiologists to understand if these independent
+storm-associated hazards tend to be strongly associated (in high agreement), 
+both within specific storms (for single-storm studies) and on average across
+most large storms (for multi-storm studies). This knowledge will help them in 
+identifying potential confounders if they are interested in studying the 
+effects of a single hazard of the storm and also in determining if control for
+these potential confounders among other storm hazards might cause numerical 
+problems or other collinearity-associated problems when fitting a regression 
+model.
+
+If a researcher wishes to study the storm as a whole, then the researcher 
+could either seek a single measure to capture exposure to the storm as a whole
+or investigate more complex models that model the influence of the storm through
+pathways of several different hazards.
+
+If a researcher wants a single measure for the storm as a whole, then it is helpful
+to know how well exposure assessments based on the hazard components of the storm
+agree with each other. If there is strong agreement, then any of those hazards could
+be used as a primary marker of exposure assessment. This is conceptually similar to 
+checking for observer / rater agreement.
+
+> "When the row and column variable srepresent different observers' rating the same
+subjects or objects, interest is focused on *observer agreement* rather than mere 
+association. In this case, measures and tests of agreement provide a method of
+assessing the reliability of a subjective classification or assessment procedure."
+[@friendly2015discrete]
+
+> "In assessing the strength of *agreement* we usually have a more stringent 
+criterion than in measuring the strength of *association*, because observers 
+ratings can be strongly associated without strong agreement. For example, one
+rater could use a more stringent criterion and thus consistently rate subjects
+one category lower (on an ordinal scale) than another rater. More generally, 
+measures of agreement must take account of the marginal frequencies with which two 
+raters use the categories. If observers tend to use the categories with different
+frequency, this will affect measures of agreement." [@friendly2015discrete]
+
+When you have an array of two or more exposures that are strongly correlated, 
+including them together in a regression model can cause problems from collinearity
+[@schisterman2017collinearity].
+
+> "Environmental exposures ... often consist of an array of related individual 
+factors that may be highly correlated with each other, causing concern about the
+impact of collinearity when attempting to identify individual effects." [@schisterman2017collinearity]
+
+**Commonly have multiple factors of interest, correlated**
+
+In some areas of epidemiology, researchers have moved toward the idea of a 
+"web of causality":
+
+> "'Multiple causation' is the canon of contemporary epidemiology, and its
+metaphor and model is the 'web of causation'. Expressed through the notion
+of 'multifactorial etiology' and embedded in the statistical techniques
+of 'multivariate analysis', the belief that population patterns of health 
+and disease can be explained by a complex web of numerous interconnected 
+risk and protective factors has become one of this discipline's central 
+concepts. Equally entrenched is the corrolary that epidemiology's power
+to improve the public's health rests upon its ability to identify---and predict
+the results of breaking---selected strands of this causal web." [@krieger1994epidemiology]
+
+This is in contrast to focusing on single factors:
+
+> "Expressly challenging the still-pervasive tendency of epidemiologists to think
+in terms of single 'agents' causing discrete diseases, the provocative metaphor
+and model of the 'web' invited epidemiologists to embrace a more sophisticated
+view of causality." [@krieger1994epidemiology]
+
+In this framework, epidemiology helps by identifying where it would help to
+break strands of the web:
+
+> "Using this model, MacMahon et al. drew several important inferences about
+prevention and research that remain part of epidemiologic thinking to this 
+day. Arguing that 'to effect preventative measures, it is not necessary to 
+understand causal mechanisms in their entirity', they stated that 'even knowledge
+of one small component may allow some degree of prevention' since 'wherever the
+chain is broken the disease will be prevented'." [@krieger1994epidemiology]
+
+In other words, if you can identify *necessary* causes, you can craft 
+*practical* interventions.
+
+In some fields of epidemiology (e.g., nutritional epidemiology), it is common 
+to study cases where there are multiple potential risk factors, and these 
+factors are often strongly correlated [@schisterman2017collinearity].
+
+> "The dynamics of many important ecological and evolutionary processes are 
+often influenced by multiple interacting factors, and attempting to understand 
+the dynamics of such systems presents difficult technical and philosophical 
+problems." [@petraitis1996inferring]
+
+> "Often in practice, investigators are faced with decisions of how to handle
+highly correlated variables." [@schisterman2017collinearity]
+
+**Methods to tackle multiple causality**
+
+When the aim is to explore the influences of multiple associated factors, there
+are some different strategies that can be used. These include multiple regression
+and path analysis:
+
+> "Evolutionary ecologists have recently turned to path analysis, multiple regression,
+and related techniques to analyse systems of multiple causality. While these
+techniques are often seen as competitors, they are, in fact, related, and as a 
+result the problems of these techniques are similar." [@petraitis1996inferring]
+
+**Implications of collinearity**
+
+> "Two variables are defined as collinear if one can be expressed as an exact
+or near linear combination of the other. ... In general, collinearity between
+variables is described by the magnitude of their correlation." 
+[@schisterman2017collinearity]
+
+It can complicate inference of model coefficients. This is always a concern for 
+inferential models. For predictive models, it can be a problem if you later use
+the model to predict in a case where the correlation structure is different than
+in the training data.
+
+> "The dependence of regression and path coefficients on the correlation structure
+is often glossed over. Yet this dependence means that inferences about the strengths
+of path coefficients are conditional upon the correlational structure among the 
+predictor variables. If the correlation structure of the sample does not
+match the correlational structure of the population then regression and path 
+coefficients will not estimate relative strengths accurately. This can occur if
+sampling is not random or if analysis is based on data from experiments with 
+factorial designs, which break the correlational structure amnog the predictor
+variables." [@petraitis1996inferring]
+
+These implications can depend on whether the modeling goal is predictive or 
+inferential.
+
+> "When the goal is to choose covariates based on their predictive value, 
+concerns about collinearity stem from the desire to improve parameter estimates'
+precision, which can be inflated when highly correlated covariates are included
+in the model." [@schisterman2017collinearity]
+
+> "In a causal framework, on the other hand, interest is often centered on a
+specific effect from a correctly specified causal model, even in the presence of
+highly correlated data, despite a potential loss of statistical efficiency.
+Under such a viewpoint, bias due to misspecification of the causal associations
+is a more devastating result than reduced precision of parameter estimates
+correctly estimated. Furthermore, a precisely but incorrectly estimated
+parameter may lead to invalid inferences. ... Relying on conventional
+variable-selection algorithms and standard analyses can ignore important
+confounders and even produce extremely imprecise confidence intervals."
+[@schisterman2017collinearity]
+
+The amount of bias in the unadjusted model (in terms of bias from an unadjusted
+confounder) depends on the strength of association between the exposure of
+interest and the potential confounder [@schisterman2017collinearity].
+
+> "This is especially important for the confounding variable scenario, where 
+correctly specifying the model may result in adjusting for a nearly 
+collinear variable." [@schisterman2017collinearity]
+
+The degree of agreement will influence the degree of confounding in an 
+unadjusted model, for either one or two confounders:
+
+> "In this confounding scenario, increasing correlation exacerbates the bias
+due to misspecification, similar to results for a single confounder."
+[@schisterman2017collinearity]
+
+Extreme collinearity can create numerical problems [@schisterman2017collinearity].
+
+> "Under the extremely unlikely scenario of extreme collinearity [(abs. rho of 0.999 
+or higher)], the numerical matrix inversion required for estimation fails, resulting
+in unstable and biased predictors." [@schisterman2017collinearity]
+
+> "Only for extreme values of rho cloes to 1 or -1 is beta-squareed nonidentifiable
+due to numerical instability; otherwise, the correctly specified model produces
+the most desirable (e.g., unbiased) estimates of the total effect of the exposure
+on the outcome." [@schisterman2017collinearity]
+
+Collinearity can cause imprecision in coefficient estimates, even if those
+estimates are unbiased [@schisterman2017collinearity].
+
+**Techniques to address collinearity**
+
+There are study designs that might help to separate different effects. For example, 
+some matching designs might help in creating a subset of data in which the association 
+that exists in the full dataset is broken (e.g., @leal2012multicollinearity).
+
+> "Matching is typically employed to reduce model dependence and estimate 
+associations in a more empirical way than would be necessary without 
+matching. In line with this practice, matching was employed as a diagnostic
+tool to verify whether the effects of environmental variables that are highly 
+correlated with each other could be disentangled." [@leal2012multicollinearity]
+
+You could also explore techniques of dimension reduction, to create a smaller
+subset of independent variables for which the correlation is broken. For example, 
+you could use PCA [@leal2012multicollinearity].
+
+There are other techniques that allow some bias in estimates in an attempt to reduce
+problems from collinearity, like ridge regression [@petraitis1996inferring].
+
+> "Several biased estimation techniquest ahve been proposed as a way to combat
+collinearity (Myers 1990). These are controversial methods. Ridge regression, for
+example, biases the regression coefficients in order to improve the model's 
+ability to predict the criterion variable. The estimates of the individual regression
+coefficients may be very biased even though when taken together they provide a 
+very good prediction of the criterion variable." [@petraitis1996inferring]
+
+> "When correct specification of the models still leads to 
+problems for conventional regression methods, one can apply modern techniques
+such as Bayesian hierarchical models with shrinkage estimators."
+[@schisterman2017collinearity]
+
+**Are effects of different factors 'separable'?**
+
+Epidemiology often aims to inform policy decisions. In this case, it can be important
+to characterize the separate effects of different associated factors that contribute
+to the health risk [@leal2012multicollinearity].
+
+> "To develop efficient public health interventions addressing the obesity 
+epidemic, it is important to identify exactly which aspects of the
+environment influence obesity risk. For example, demonstrating that the density 
+of fast-food restaurants was associated with obesity risk and demnstrating that
+the density of sports facilities was associated with obesity risk would lead to 
+different interventions." [@leal2012multicollinearity]
+
+However, it can be very difficult to separate the effects of different factors
+if those factors are strongly correlated [@leal2012multicollinearity]:
+
+> "However, many neighborhood characteristics, especially those related to the 
+densities of physical features and services, are strongly correlated with 
+each other. Therefore, it is not clear whether it is even possible to disentangle
+the effects of these different environmental dimensions, even if they are hypothesized
+to influence obesity risk through distinct causal pathways." [@leal2012multicollinearity]
+
 \begin{shaded}

### 2020-07-29 ehp_revision/response_to_reviewers.Rmd
Work some more on revisions

@@
 landfall and can no longer draw on warm ocean waters for energy.
+
+Data that is aggregated across a population---for example, the total number of 
+deaths in a geographic area in a certain time period---is known as *ecological
+data*, and studies that base inference on this type of data are known as *ecological
+studies*. Terms include *ecological*, *aggregate*, *contextual level*.
+
+> "In an ecological study the unit of analysis is a group of people rather than 
+the individual." [@sedgwick2014ecological]
+
+> "In his seminal paper, Robinson (1950) distinguished between two types of 
+correlation---ecological and individual. The former is obtained for a group
+of people, while the latter is estimated for indivisible units, such as individuals."
+[@portnov2007ecological]
+
+Here is one challenge when working with aggregated data. If you use aggregated
+data to infer individual-level associations, this can lead to *ecological fallacy*. 
+This results from cross-level inference---the data used to model the association
+is at the contextual level (e.g., county-level), while the inference is for the
+individual association between exposure and outcome.
+
+One definition of the ecological fallacy:
+
+> "This spurious result provides an example of the ecological fallacy, in 
+which conclusions ... based on ecologic data are opposite of those drawn 
+on the basis of individual-level data." [@wakefield2008overcoming]
+
+> "It is a bias produced when analyses realised in an ecological (group level)
+analysis are used to make inferences at the individual level."
+[@delgado2004bias]
+
+> "Results from ecological studies are prone to the ecological fallacy. The 
+ecological fallacy is a term used when collected data are analysed at a 
+group level and the results are assumed to apply to associations at the 
+individual level." [@sedgwick2014ecological]
+
+Importantly, aggregated data can be used both to infer a contextual effect
+(e.g., the association between county-wide rates of a health outcome and 
+exposure) as well as an individual-level effect (e.g., how a person's
+risk of the outcome is associated with that person's exposure to the hazard). 
+The first inference will *not* be prone to ecological bias, as the inference
+is at the same level as the data. The second inference (individual-level) 
+will be susceptible to bias when aggregated data is used to fit the model. 
+In this case, the association estimated based on aggregated data can be 
+different, and even reversed, compared to the true individual-level effect.
+
+> "I propose three criteria for the identification of ecological fallacy; 
+all three of these should be present to confirm its existence: (1) Results
+must be obtrained with ecological (population) data; (2) Data must be inferred
+to individuals. One use of ecological studies is to explore individual-level 
+association when individual data are not available. When the focus of the 
+study was contextual or based on population effects and there is no inference
+to individuals, ecological fallacy is not possible. When only the first two 
+criteria are present---which is insufficient to affirm ecological fallacy---it
+is appropriate to acknowledge that there is a possible relationship and that further 
+study is required; (3) Results obtained with individual data are contradictory."
+[@idrovo2011three]
+
+> "One important breakdown in the analogy between ecologic-bias and individual-level
+confounding occurs because in etiologic research, the target of inference for
+both ecologic and individual-level studies is the same: Both study effects at the 
+individual level. For an individual-level study, these target effects are at the 
+same level as the units of analysis. But, for an ecologic study, these target 
+effects are at a finer level than the units of analysis. As a result, an ecologic 
+study can be unbiased for ecologic effects (in particular, ecologic confounding
+may be absent) and yet still be biased for individual-level effects. Many of the 
+classic social science examples of ecologic bias are of this form." [@greenland1994invited]
+
+> "We refer to an incorrect extrapolation from unbiased estimates of ecologic effects
+to unobserved individual-level effects as 'cross-level bias' (although the latter
+term is sometimes used to refer to any ecologic bias). As the examples suggest, 
+it is possible to detect cross-level bias if one can identify and observe 
+homogeneously exposed regions." [@greenland1994invited]
+
+> "assuming that associations observed at the level of the area hold for the 
+individuals within the areas can lead to the so-called ecological fallacy 
+(Selvin, 1958)." [@wakefield2006health]
+
+> "A broad definition of ecological fallacy is that certain data (e.g., illiteracy 
+rates by foreign-born) are unavailable, leading to the use of proxies and 
+erroneous estimates. However, we advocate a more narrow definition of this 
+phenomenon, according to which 'ecological fallacy' refers solely to differences
+in conclusions which may be drawn from group-level data (e.g., average illiteracy 
+rates by foreign-born), as opposed to data obtained for individuals."
+[@portnov2007ecological]
+
+> "If different conclusions are drawn from the analysis of data upon their aggregations
+into units of different sizes (e.g., from individuals to townships and regions), these
+differences are commonly referred to as 'ecological fallacy'. In epidemiology, these
+differences are also known as ecological or cross-level bias [refs]. The ecological 
+bias is difficult to control and may lead, under certain circumstances, to spurious 
+effects [refs]. If epidemiological study does not consider the possibility of 
+ecological bias, misguided policy decisions may follow [refs]." [@portnov2007ecological]
+
+Ecological bias can *only* happen if there is, indeed, an individual-level effect:
+
+> "If exposure has no effect on any individual, there will be
+no individual-level or ecologic effects, and so cross-level bias cannot occur. 
+Thus, cross-level bias will not affect the validity of an ecologic test of the 
+null hypothesis, although it still must be considered in interpreting a 
+significant effect." [@greenland1994invited]
+
+In some cases, the ecologic-level effect (contextual effect) will be directly 
+of interest, and so we may be interested in inferring at that level (in which 
+case use of ecologic-level data is appropriate and will not be prone to 
+ecological bias):
+
+> "The bias phenomenon in example 7 may be ascribed to the fact that an 
+ecologic (regional) variable (percent Protestant) had effects on individual 
+risk, in addition to effects of the corresponding individual-level variable
+(religion) that the ecological variable summarized. In both social and 
+infectious disease epidemiology, as well as in community intervention studies, 
+ecologic effects may be of direct interest; a classic example is the 
+phenomenon of 'herd immunity', in which the overall prevalence of 
+immunity in a region, as well as individual immune status, determines the 
+risk of individuals within the region (16). For noncontagious diseases, 
+however, ecologic effects may not be of direct interest, and may in fact
+obscure the individual-level effect of interest." [@greenland1994invited]
+
+In some cases, the estimated association based on ecological data will be very 
+different from the association that exists at the individual level:
+
+> "We see that the ecological relative risk associated with low birth 
+weight is completely incomparable with the individual-level coefficient."
+[@wakefield2008overcoming]
+
+Ecological bias can also complicate estimation of effect modification, which 
+otherwise could help in identifying vulnerabilities and susceptibilities among
+certain subpopulations:
+
+> "Furthermore, non-White race now appears protective, rather than detrimental
+as the individual-level analysis suggests." [@wakefield2008overcoming]
+
+Ecologic bias crops up because the aggregated data fails to capture variation
+within each aggregation level (county, for example) for both the exposure 
+and potential confounders. For example, during a tropical cyclone, different
+parts of a county will experience different peak wind speeds. If the county 
+is coastal, the most severe winds will likely be near the coast, as the 
+storm will weaken once it makes landfall and moves inland, as the storm 
+generates its power from the water. When we assign the entire county a
+single estimate of intensity, it smooths over the fact that locations in 
+some parts of the county were subject to higher peak winds while locations
+in other parts of the county had lower winds. At the least, this smoothing
+of variation in exposure levels across the county could lower the power of 
+the study to detect a clear association between exposure and health risk, as
+this smoothing drops information inherent in the variation within the county
+in exposure levels.
+
+This, however, would not result in ecological fallacy unless another factor
+comes into play---variation within each aggregation level in confounders. 
+Just as aggregation smooths over within-area variation in exposure levels, it
+also smooths over within-area variation in levels of potential confounders. 
+Depending on the patterns of this within-area variation, a result could be
+that ecological-level control of the confounders does not, in fact, control
+for their role at the individual level, and so the association inferred at
+the ecologic level continues to be confounded by them when inferred to the
+individual level.
+
+Essentially, this comes down to a question of the joint distribution, at the
+individual level, of exposure, outcome, and covariates [?].
+
+> "As various discussants have pointed out (3--6, 8, 9), ecologic relative-risk
+estimates can be subject to biases not present in estimates from individual-level
+observational studies of the same populations (case-control and cohort studies).
+Unlike an individual-level study, an ecologic study does not link individual
+outcome events to individual exposure or covariate histories, nor does it
+link individual exposure and covariate histories to one another. It is these
+linkage failures that are the source of the special biases of ecologic studies
+(1, 2, 6)." [@greenland1994invited]
+
+> "In the extreme, ecologic bias in comparisons limited to homogeneously exposed
+regions (as in examples 3--6) can be viewed as purely an issue of confounding,
+albeit with special complexities of measurement and control of confounders."
+[@greenland1994invited]
+
+> "More generally, one may show that cross-level bias will not occur if the 
+individual-level effects of all variables (including unmeasured background
+factors) follow a multiple linear-regression model with no regional effects and
+no interactions (17). Nevertheless, given the usual inappropriateness of the 
+multiple-linear model and the absence of homogeneously exposed regions, the 
+possibility of cross-level bias adds a dimension to ecologic bias beyond
+that of simple confounding." [@greenland1994invited]
+
+> "Covariate control in ecologic studies requires attention to details not 
+ordinarily of concern in individual-level studies. When, as is usually the
+case, important nonlinearity or nonadditivity can be expected among 
+exposure and covariate effects, it may be necessary to obtain and 
+control for multiple summaries of joint covariate distributions in order
+to insure that control is adequate." [@greenland1994invited]
+
+> "Ecological fallacy can be produced by within group (individual level)
+biases, such as confounding, selection bias, or misclassification, and
+by confounding by group or effect modification by group." [@delgado2004bias]
+
+> "**Confounding by group**: it is produced in an ecological study, when 
+the exposure prevalence of each community (group) is correlated with 
+the disease risk in non-exposed of the same community. It can be 
+a mechanism for producing ecological fallacy." [@delgado2004bias]
+
+If individual-level inference is the aim, and population-level data is
+available, there are some methods for using it while still aiming to avoid
+ecological bias. Indeed, it can be helpful to use population-level data, as it
+is often available for a large population, improving the power and precision of
+the study. Further, the level of exposure might vary a lot more over the
+population captured with population-level data compared to the variation that
+captured in a smaller sample of individual-level data. This can contribute both 
+to statistical power and improve external validity (as the study data will cover
+more of the range of exposure that might ever be expected). However, if
+individual-level inference is the goal, then there are ways to supplement
+population-level data with samples of individual-level data through two-level, 
+semi-ecologic study designs.
+
+> "It is generally well recognized that in order for ecologic data to provide
+reliable inferences, they need to be supplemented with individual-level data."
+[@wakefield2008overcoming]
+
+> "For this example, we have access to complete individual-level data, permitting
+a 'gold standard' individual-level analysis." [@wakefield2008overcoming]
+
+> "The only reliable way to characterize within-area variation in exposures
+and confounders, and hence control ecologic bias, is to collect and incorporate
+individual-level data. To help epidemiologists achieve this goal, in this paper
+we describe the use of the two-phase design in an ecologic setting."
+[@wakefield2008overcoming]
+
+> "The only reliable way to characterize within-area variation in exposures
+and confounders, and hence control ecologic bias, is to collect and incorporate
+individual-level data. To help epidemiologists achieve this goal, in this 
+paper we describe the use of the two-phase design in an ecologic setting."
+[@wakefield2008overcoming]
+
+> "In a semi-ecologic study, an ecologic exposure is combined with individual-level
+outcomes and confounders. A two-phase approach is particularly useful for
+such a study, with the phase 2 data corresponding to stratified sampling of 
+individual exposures." [@wakefield2008overcoming]
+
+> "To implement a two-phase design, an initial phase 1 cross-classification by 
+the binary disease outcome and stratification variables is required; in phase
+2, samples of individuals are drawn from each of the cross-classification cells, 
+with data on additional variables being drawn from the subsamples of individuals. 
+Intuitively, the stratified sampling is focused on informative cells, and 
+estimation methods use both phases of data for efficiency and to acknowledge
+the outcome-dependent sampling. In the simplest ecologic setting, the cross-classification
+is by outcome and area only, and if area is a surrogate for important risk factors
+this design will be efficient. We are particularly interested in situations where
+an initial classification is available by outcome, area, and confounders such as
+age and gender---this is the case in a semi-ecologic study. Phase 2 may then provide
+detailed exposure information on a subset of the phase 1 individuals." 
+[@wakefield2008overcoming]
+
+> "Two-phase study designs are a generalization of matched case-control designs
+in which, initially, the entire sample population is cross-classified according
+to case/control status and some stratification variable, S. The latter depends
+on covariates observed in all individuals and may include exposures of interest, 
+proxy exposure measures, or potential confounders. In settings like those we
+are considering, such as environmental epidemiology, S may also depend on geographic
+area, which can act as a surrogate for the totality of confounders associated
+with each area, as well as provide a well-defined sampling frame for the controls."
+[@wakefield2008overcoming]
+
+> "The results of our simulation study point to the benefits associated with combining
+the two sources of data, in terms of both bias and efficiency. Rather than supplement
+an ecologic study with individual-level data, it may be of interest to combine 
+existing individual-level data with external group-level data. Strategies that
+combine both types of data have been shown to alleviate participation bias and
+improve efficiency in case-control studies with missing data (31)." 
+[@wakefield2008overcoming]
+
+> "As with other observational studies, ecologic studies can give useful results if 
+biases such as those discussed here can be ruled out or quantified. Nevertheless, 
+bias evaluation can be especially difficult in ecologic studies of geographic
+regions because of the many potentially interacting covariates that may differ
+across regions. When biases cannot be ruled out with available data, further
+exploration will require individual-level studies." [@greenland1994invited]
+
+Why we might sometimes use (and want to use) population-level data:
+
+> "Because of the unavailability of individual-level data, ecological data
+may be resorted to and may come in a variety of forms." [@wakefield2008overcoming]
+
+> "Epidemiologists continue to use ecologic and aggregate data. Despite their 
+known drawbacks, these data, often aggregated across geographic areas, offer the
+advantages of widespread availability and gains in statistical power from large
+populations and increased exposure ranges." [@wakefield2008overcoming]
+
+> "Population and health data are often routinely available in ecological, 
+that is group, form while the exposure data typically consist of a set of
+values recorded at monitor sites, or via one-off sampling." [@wakefield2006health]
+
+It can be important to think about the scale that the process happens at. For
+something very local, you might lose a lot more with aggregated data.
+
+> "Data availability and exposure variability often determine the scale of
+examination and the suitability of a study. Exposures arising from a point or
+line source offer exposure contrasts on small scales, requiring small-area data;
+in constrast, dietary variables show little variation across small scales, and
+consequently international studies are used." [@wakefield2008overcoming]
+
+There can also be an *individualistic fallacy*, if emphasis is limited to the
+association between individual-level factors, without considering the potential
+role of contextual factors in an individual's risk of a health outcome. In other
+words, cross-level bias can go in both directions [@idrovo2011three]. While the
+fallacy of inferring individual-level associations from ecologic-level data is
+called *ecological fallacy*, the fallacy of inferring ecologic-level
+(contextual) effects from individual-level data is known as the *atomistic
+fallacy* [@idrovo2011three]. There are also fallacies that can come in, it
+seems, from inferring at the same level as the data but failing to consider
+influences from factors at the other level. When inferring individual-level
+associations from individual-level data, without considering an additional role
+of ecological-level factors, this is known as the *psychologistic* or
+*individualistic fallacy* [@idrovo2011three]. When inferring contextual effects
+from ecologic data without considering individual-level factors, this is known
+as the *sociologistic fallacy* [@idrovo2011three].
+
+> "Although ecological studies are important to epidemiology (especially in 
+environmental and social epidemiology), public health practitioners seem 
+afraid of ecological studies. It is a common practice to assume the presence of 
+ecological fallacy (Robinson 1950) and low-level validity when analyzing an 
+ecological study. Most epidemiologists prefer an exclusive individualistic
+approach, although the importance of a multilevel causal approach is widely 
+recognized (Diez-Roux 2002). In this sense, some authors suggest that it is
+as important to recognize the presence of ecological fallacy as to 
+recognize psychologistic or individualistic fallacy (Subramanian et al. 2009)."
+[@idrovo2011three]
 

### 2020-07-29 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 
-Great idea. We are investigating sharing this work in this repository.
+Great idea!
+
+We have investigated this. DesignSafe is ...
+
+As part of its platform, DesignSafe allows users to work in a number of 
+programming languages, including R (through Jupyter Notebooks). Therefore, 
+DesignSafe users already have access to any data shared through an R package, 
+including the data we present in this manuscript, without the data needing 
+to be separately loaded and published through DesignSafe's Data Depot.
+
+We have added a note in the manuscript to let readers know that, because the
+data is shared through an R package, and because DesignSafe's [Workspace]
+allows users to work in R, the data can be fully accessed and used for those working
+on the DesignSafe platform by using the same code in a DesignSafe Jupyter notebook
+that would otherwise be used on a local computer.
+
+This is good, because otherwise there would be a good deal of overlap between
+the data on DesignSafe and that in the R package. R packages already created 
+a citable source, comparable to the DOI assigned to data by DesignSafe when
+it is published. Further, R packages include licensing information, as does
+DesignSafe for published data. Finally, if the data were shared through both 
+platforms, it would be critical to make sure that it was clear that the two 
+datasets are analogous and also to keep them up to date with each other, which
+would add an extra layer of work with each update to the R package (we have 
+been updating several hazards as new data becomes available). It is wonderful 
+that DesignSafe's interface with R avoids this problem by allowing its users
+direct access to our package and its data through that path, as it allows for 
+a single, citable, licensed version of the data stored and shared from a single
+location, with no question of syncing separate versions of the data or clarifying
+licensing and citation practices across two storage locations.
+
+*What is DesignSafe:*
+
+> "DesignSafe is the cyberinfrastructure platform that has been developed as
+part of NHERI to support natural hazards engineering research, and
+it succeeds the NEEShub cyberinfrastructure that was developed for the
+earthquake engineering community through the Network for Earthquake
+Engineering Simulation (NEES) program." [@rathje2017designsafe]
+
+> "DesignSafe plays an important role in integrating the various NHERI components
+and the research taking place at NHERI facilities, but also has the broader
+goal of enabling transformative research in natural hazards and engineering
+across the numerous technical disciplines engaged in this field." 
+[@rathje2017designsafe]
+
+> "DesignSafe has been developed as a flexible, extensible, community-driven
+cyberinfrastructure, and it embraces a cloud strategy for the big data generated
+in natural hazards engineering. DesignSafe provides a comprehensive CI that
+supports the full research lifecycle from planning to execution to analysis
+to publication and curation." [@rathje2017designsafe]
+
+> "The future of natural hazards engineering research requires the integration
+of diverse data sets from a variety of sources including experiments, computational
+simulation, and field reconnaissance, as well as a variety of research disciplines
+including earth science, social science, building science, and architecture. 
+The DesignSafe cyberinfrastructure has been designed to provide the functionalities
+that will enable transformative research in natural hazards engineering. By adopting
+a cloud strategy, DesignSafe allows for a fundamental change in the way that
+research is performed. It provides a comprehensive cyberinfrastructure that
+supports research workflows, data analysis, and visualization, as well as the 
+full lifecycle of experimental, field, and computational research required by 
+engineers and scientists to effectively address the threats posed to civil 
+infrastructure by natural hazards. The integration of data and computation 
+in the cloud will enable new research discoveries in natural hazards engineering, 
+which in turn can lead to more hazard-resilient civil infrastructure." [@rathje2017designsafe]
+
+*Access to R:*
+
+> "The real revolution takes place downstream of data generation. Here, data 
+analytics and visualization are performed in the cloud within the *Discovery
+Workspace* while accessing any data within the *Data Depot*. Researchers can 
+invoke common analysis programs, such as MATLAB, as well as other analysis/
+visualization tools, such as Jupyter notebooks. A Jupyter notebook is an 
+electronic notebook that allows users to embed rich text elements as well
+as computer code, graphs, and visualizations within a single notebook that
+can be shared through the web. Over 40 different programming languages are 
+supported in Jupyter, including Python and R, and MATLAB code can be easily 
+converted, making Jupyter a versatile tool for research. Performing analysis in 
+the cloud allows researchers to integrate and explore various data without
+tedious downloads. Additionally, using a seamless cyberinfrastructure to 
+complete all research tasks enables tracking and relating of the processes 
+applied to data." [@rathje2017designsafe]
+
+*Data Depot:*
+
+> "the Data Depot, a flexible data repository with streamlined data management
+tools." [@rathje2017designsafe]
+
+> "As data progresses towards publication the requirements for metadata 
+increase as metadata provides users with search and discovery functions. At
+the end of the research project the user may edit the information for 
+publication and complete the process of assigning digital object identifiers
+(DOIs) and applying the appropriate license. On-demand assistance from a 
+curator is available to provide training and guide users through their data
+curation and publication needs." [@rathje2017designsafe]
 

### 2020-08-09 ehp_revision/response_to_reviewers.Rmd
Fix a typo in a citation

@@
 
+This exposure data is provided at the county level, and this can lead to 
+exposure measurement error / exposure misclassification in studies that 
+infer to the individual level. In this case, there will be several processes
+at play. First, the hazards of the storm will not be uniform across the
+county. The degree of this within-county spatial heterogeneity will usually 
+vary by hazard. The within-county heterogeneity is likely lowest for rain and for
+wind at lower levels. For rain, there can be particularly severe rain bands, 
+but often a storm's rainfield will bring high precipitation to a large 
+area. Wind can have a high heterogeneity close to the storm's center 
+(storm winds decay most quickly with distance near the storm's center)
+and near landfall (winds decay rapidly after landfall). Other hazards can 
+be very local within the county, only affecting a small proportion of the total
+county area. A tornado is a clear example---while it causes severe destruction
+in its track, areas just off the track can be unharmed, creating a dramatic
+spatial gradient in damage within a county. Floods can follow both. Some hurricanes
+have caused widespread flooding (e.g., Floyd, Katrina) while some have caused
+flooding that is more spatially focused (inland, with flashfloods? coast flooding
+just near the coast?). For tidewater/coastal areas, there is often a dense 
+network of waterway. When rains occur further up the watershed, these all
+become flooded.
+
+Second is the question of whether exposure measured at the resident's location
+would be ideal. This is the case typically when studying an exposure that is 
+a chemical substance. In this case, the pathway depends on the substance entering 
+the body. For tropical cyclones, health risk could result from stress, problems
+accessing medical care, or repercussions of power outages (higher exposure to 
+outdoor hazards like heat and air pollution, problems storing food and using 
+some medical equipment). For these, high levels of wind and other hazards in the
+person's community may be as or more important than at their precise location.
+
+Measurement error can be either random or systematic. Systematic error can often 
+be corrected with adjustment if the direction and typical size of the error is 
+understood. Random error cannot in the same way. Either type of measurement error
+can be either differential (associated with the probability of the outcome) or 
+non-differential (independent of the distribution / probability of the outcome).
+In simpler models, this characteristic might help in predicting whether the resulting
+bias is likely toward the null; however, more complex models (e.g., statistical 
+models with adjustment for potential confounders) are trickier to diagnose in terms
+of the likely implications of differential versus non-differential measurement 
+error [?].
+
+Not only can the exposure be measured with error, but so too can potential confounders
+and potential effect modifiers.
+
+When a proxy exposure estimates is used for many people covered by a study, this
+can result in a type of exposure measurement error called *Berkson error*. In this 
+case, the true exposure of each individual is randomly distributed around this 
+proxy/aggregate-level exposure estimate, with a mean exposure level among the 
+subjects that is equal to the proxy/aggregate-level measure. In other words, the 
+group as a whole is assigned a common exposure level, when in fact this is the 
+average exposure level across members of that group, but the individuals' 
+true exposure levels are randomly distributed around this mean group level to 
+which they are all assigned for modeling.
+
+Berkson:
+
+$$
+T = X + E
+$$
+where:
+
+- $T$: true individual exposure
+- $X$: proxy exposure, assigned to everyone in the group
+- $E$: error, with mean of 0, independent of $X$
+
+The dataset provides exposure by group/area (county), not by individual. These
+ecological exposure estimates require us to assume that the exposure level is
+the same for all individuals in that county. It ignores/smooths over within-county
+variability in exposure.
+
+For wind and rain, we're essentially basing exposure assessment on modeling of
+exposure. For rain, this is a based on a re-analysis model that
+incorporates/integrates observed data from the area and time, using a model to
+integrate data from diverse sources. For wind, the exposure estimates are based
+on a model, with a core based on the Willoughby wind model. For flooding and
+tornadoes, the exposure is based on records of events anywhere in the county,
+and in many cases may be based, essentially, on exposure assessment, as in many
+cases someone like someone at a NOAA [?] center have determined whether the
+event should be reported in the database.
+
+Even exposure at each person's residence would not suffice, since people move 
+around during the day (although perhaps less for most people during a severe
+storm).
+
 Data that is aggregated across a population---for example, the total number of 
@@
 
+Confounding in general:
+
+> "A confounder ... is associated with an exposure or risk factor for the outcome
+and with the outcome independent of the exposure, but is not on the causal pathway
+between the exposure and the outcome. Confounding occurs when the relationship
+between an exposure variable and the outcome variable is contaminated, so that the
+measure of association between these two variables is actually also capturing the
+effects of a third variable, the confounding variable. For example, many factors 
+are associated with income (exposure) and health (outcome), such as education, 
+employment or wealth? It may be that these factors explain part of the observed
+assocation." [@gunasekara2008glossary]
+
+Measurement error in general:
+
+> "Measurement error: The different between an observed variable and the variable 
+that belongs in a multiple regression equation. Measurement error is the random 
+or systematic error arising during data collection of variables. The measured
+variable x is measured with error eta, which is the distance from x to the 
+'true' value of x. Measurement error can produce bias such as attenuating the
+estimators of exposure variables, but may also have more complex effects. Longitudinal
+data analyses such as fixed effects models can actually augment the problem of 
+measurement error in mis-measured explanatory variables that change little over
+time." [@gunasekara2008glossary]
+
+Study validity:
+
+> "Study validity refers to the degree to which the inferences drawn from a study are
+warranted when account is taken of the study methods; the representativeness of
+the study sample; and the nature of the population from which it is drawn.[5]
+There are two types of study validity. *Internal validity* is the degree to which the
+results of a study are correct for the sample of people being studied. *External 
+validity* (generalisability) is the degree to which the study results hold true
+for a population beyond the subjects in the study or other settings.[20]"
+
 \begin{shaded}
@@
 > "We initially developed software to model exclusively Atlantic-basin storms
-[@tormwindmodel]. There are two characteristics of Northwestern Pacific Ocean
+[@stormwindmodel]. There are two characteristics of Northwestern Pacific Ocean
 basin storm data that prevent direct use of the current version of the software.
@@
 allowing them to estimate average associations over many disasters and to
-explore how a disaster’s characteristics, or the characteristics of the
+explore how a disaster's characteristics, or the characteristics of the
 community it hits, modify associated health risks (e.g., Anderson and Bell 2010;
@@
 them to estimate average associations over many disasters and to explore how a
-disaster’s characteristics, or the characteristics ofthe **affected
+disaster's characteristics, or the characteristics ofthe **affected
 communities**, modify associated health risks (e.g., Anderson and Bell 2010; Son
@@
 in the cloud will enable new research discoveries in natural hazards engineering, 
-which in turn can lead to more hazard-resilient civil infrastructure." [@rathje2017designsafe]
+which in turn can lead to more hazard-resilient civil infrastructure." 
+[@rathje2017designsafe]
 
@@
 \textbf{Reviewer 2 (R2) Comment 1:}
-Line 164: Isn’t it obvious to use better exposure measures than proxies? Wind,
+Line 164: Isn't it obvious to use better exposure measures than proxies? Wind,
 rainfall, tornado, and flood are direct products of storms while distance is
@@
 
+> "Multicollinearity: A term that refers to correlation among the independent
+variables in a multiple regression model; it is usually invoked when some 
+correlations are 'large', but an actual magnitude test is not well defined.[1]"
+[@gunasekara2008glossary]
+
+> "Collinearity occurs in multiple regression models when two (or more---multicollinearity)
+exposure variables are included in the model but are so similar to each other that
+they are essentially measuring at least part of the same thing. Collinearity 
+creates problems with interpreting the analysis by affecting the standard errors
+of the variables (as it increases variance)[24] and by causing biased estimates for one
+or both of the collinear terms (sometimes dramatically so) [25]. If there is perfect
+collinearity between two variables then one should be dropped from the model. However, 
+dropping exposure variables to reduce multicollinearity may lead to bias in the model
+from dropping useful information." [@gunasekara2008glossary]
+
+> "It might be argued in some cases that it is appropriate to include (collinear) 
+variables in a model even when they are highly correlated, because the theoretical 
+basis for including the variables is strong and the results of the model are consistent
+with expectations [10]. In such cases, models with and without the variables would be
+tested and compared. Such consideraion of alternative models tends to be described
+as an issue of (mis)-specification in econometrics. However, in epidemiology this
+tends to be described in terms of mediating and *intermediary variables*. Thus, an
+epidemiologist might argue for including an intermediate variable in a model (and 
+testing it against a model without the variable) because including that variable might 
+elucidate direct and indirect pathways between the exposure and outcome variables. 
+An econometrician might do the same thing but describe this in terms of finding the
+correct model specification." [@gunasekara2008glossary]
+
 **Techniques to address collinearity**
@@
 \textbf{R2 Comment 4:}
-Figure 4: I would add Y-axis label “NOAA classification."
+Figure 4: I would add Y-axis label "NOAA classification."
 \end{shaded}
@@
 
-[Check: *exposure matrix* in occupational epidemiology...]
+> "A job exposure matrix (JEM) may be defined as a cross classification of jobs
+and occupational exposures. Some matrices are based on tasks, instead of jobs. 
+... Usually there is a group of experts who build the matrix based on their 
+own knowledge and experience on occupational exopsures. The experts should 
+assess exposure to an open or fixed list of occupational exposures (that can 
+include chemical, physical, biological, and/or psychosocial agents) for an
+open or fixed list of jobs (or tasks) usually coded according to some established
+national or international classification system. Some JEMs also include 
+information for different calendar time periods." [@garcia2003glossary]
 

### 2020-08-09 ehp_revision/response_to_reviewers.Rmd
Fix line lengths
Work some more on revision

@@
 that integrate both community-level and individual-level metrics of exposure.
-This could help in determining if there are risks introduced to individuals based
-on storm-related hazards in their broader community, after controlling for the 
-hazards experienced directly at their homes.
+This could help in determining if there are risks introduced to individuals
+based on storm-related hazards in their broader community, after controlling for
+the hazards experienced directly at their homes.
 
@@
 There are study designs that can be used to leverage ecological data while
-minimizing risk from ecological bias. For example, potential confounders
-like age distribution and smoking rates vary much less within a county over
-time than comparing between counties. The mechanism for ecological bias 
-depends on the joint distribution between individual exposure, outcome, and 
-covariates, if any are included in the model. Time series-style study designs, 
-which compare a county to itself over time, allow for very similar covariate
-distributions between exposure and non-exposure. Other studies add to this
-design by stabilizing for temporal confounding through the addition of 
-counties that were never exposed, allowing for a differences-in-differences
-style approach to calibrate for seasonal or longer-term trends that might
-otherwise create confounding. For example, many health outcomes have a strong
-seasonal trend, with peak rates in the winter and lows in the summer. Since
-the hurricane season stretches from summer into fall, a study design that compares
-the rate of a health outcome in an exposed county to the rate two weeks before
-the exposure might be biased away from the null, since baseline rates of the
-health outcome will typically be moving up over most of the hurricane season.
+minimizing risk from ecological bias. For example, potential confounders like
+age distribution and smoking rates vary much less within a county over time than
+comparing between counties. The mechanism for ecological bias depends on the
+joint distribution between individual exposure, outcome, and covariates, if any
+are included in the model. Time series-style study designs, which compare a
+county to itself over time, allow for very similar covariate distributions
+between exposure and non-exposure. Other studies add to this design by
+stabilizing for temporal confounding through the addition of counties that were
+never exposed, allowing for a differences-in-differences style approach to
+calibrate for seasonal or longer-term trends that might otherwise create
+confounding. For example, many health outcomes have a strong seasonal trend,
+with peak rates in the winter and lows in the summer. Since the hurricane season
+stretches from summer into fall, a study design that compares the rate of a
+health outcome in an exposed county to the rate two weeks before the exposure
+might be biased away from the null, since baseline rates of the health outcome
+will typically be moving up over most of the hurricane season.
 
@@
 Other approaches have also been suggested to reduce ecological bias while
-leveraging ecological data, which might be available at a much larger scale---and
-so provide more power---than is practical in collecting individual data. These
-include multi-level approaches in which individual-level data from a sample
-of individuals in each community is incorporated.
+leveraging ecological data, which might be available at a much larger
+scale---and so provide more power---than is practical in collecting individual
+data. These include multi-level approaches in which individual-level data from a
+sample of individuals in each community is incorporated.
 
@@
 For example, tornadoes can be very local, causing extreme damage to a row of
-homes and none to the homes across the street. Some types of floods can also 
-be very localized, rather than creating widespread flooding throughout 
-the county. In both cases, when a county-level exposure assessment is made
-based on an event in the county in association with the storm, there may be
-many individuals in the county who did not experience that hazard at their 
-residence. The other hazards we included---storm-associated winds and rainfall---will
-also have some variation across counties, but this will often not be as stark. 
-The rainfields for tropical cyclones are very large, and while there are rainbands
-within the storm that might have particularly high rates of precipitation, these
-progess over the course of the storm, and it is unlikely that a county will have
-one area that experienced very extreme precipitation while another experienced
-very little. Winds might have more variation, especially in coastal counties and 
+homes and none to the homes across the street. Some types of floods can also be
+very localized, rather than creating widespread flooding throughout the county.
+In both cases, when a county-level exposure assessment is made based on an event
+in the county in association with the storm, there may be many individuals in
+the county who did not experience that hazard at their residence. The other
+hazards we included---storm-associated winds and rainfall---will also have some
+variation across counties, but this will often not be as stark. The rainfields
+for tropical cyclones are very large, and while there are rainbands within the
+storm that might have particularly high rates of precipitation, these progess
+over the course of the storm, and it is unlikely that a county will have one
+area that experienced very extreme precipitation while another experienced very
+little. Winds might have more variation, especially in coastal counties and
 counties near the storm's central track, since winds decay quickly as you move
@@
 This exposure data is provided at the county level, and this can lead to
-exposure measurement error / exposure misclassification in studies that 
-infer to the individual level. In this case, there will be several processes
-at play. First, the hazards of the storm will not be uniform across the
-county. The degree of this within-county spatial heterogeneity will usually 
-vary by hazard. The within-county heterogeneity is likely lowest for rain and for
-wind at lower levels. For rain, there can be particularly severe rain bands, 
-but often a storm's rainfield will bring high precipitation to a large 
-area. Wind can have a high heterogeneity close to the storm's center 
-(storm winds decay most quickly with distance near the storm's center)
-and near landfall (winds decay rapidly after landfall). Other hazards can 
-be very local within the county, only affecting a small proportion of the total
-county area. A tornado is a clear example---while it causes severe destruction
-in its track, areas just off the track can be unharmed, creating a dramatic
-spatial gradient in damage within a county. Floods can follow both. Some hurricanes
-have caused widespread flooding (e.g., Floyd, Katrina) while some have caused
-flooding that is more spatially focused (inland, with flashfloods? coast flooding
-just near the coast?). For tidewater/coastal areas, there is often a dense 
-network of waterway. When rains occur further up the watershed, these all
-become flooded.
+exposure measurement error / exposure misclassification in studies that infer to
+the individual level. In this case, there will be several processes at play.
+First, the hazards of the storm will not be uniform across the county. The
+degree of this within-county spatial heterogeneity will usually vary by hazard.
+The within-county heterogeneity is likely lowest for rain and for wind at lower
+levels. For rain, there can be particularly severe rain bands, but often a
+storm's rainfield will bring high precipitation to a large area. Wind can have a
+high heterogeneity close to the storm's center (storm winds decay most quickly
+with distance near the storm's center) and near landfall (winds decay rapidly
+after landfall). Other hazards can be very local within the county, only
+affecting a small proportion of the total county area. A tornado is a clear
+example---while it causes severe destruction in its track, areas just off the
+track can be unharmed, creating a dramatic spatial gradient in damage within a
+county. Floods can follow both. Some hurricanes have caused widespread flooding
+(e.g., Floyd, Katrina) while some have caused flooding that is more spatially
+focused (inland, with flashfloods? coast flooding just near the coast?). For
+tidewater/coastal areas, there is often a dense network of waterway. When rains
+occur further up the watershed, these all become flooded.
 
 Second is the question of whether exposure measured at the resident's location
-would be ideal. This is the case typically when studying an exposure that is 
-a chemical substance. In this case, the pathway depends on the substance entering 
+would be ideal. This is the case typically when studying an exposure that is a
+chemical substance. In this case, the pathway depends on the substance entering
 the body. For tropical cyclones, health risk could result from stress, problems
@@
 be corrected with adjustment if the direction and typical size of the error is
-understood. Random error cannot in the same way. Either type of measurement error
-can be either differential (associated with the probability of the outcome) or 
-non-differential (independent of the distribution / probability of the outcome).
-In simpler models, this characteristic might help in predicting whether the resulting
-bias is likely toward the null; however, more complex models (e.g., statistical 
-models with adjustment for potential confounders) are trickier to diagnose in terms
-of the likely implications of differential versus non-differential measurement 
-error [?].
-
-Not only can the exposure be measured with error, but so too can potential confounders
-and potential effect modifiers.
+understood. Random error cannot in the same way. Either type of measurement
+error can be either differential (associated with the probability of the
+outcome) or non-differential (independent of the distribution / probability of
+the outcome). In simpler models, this characteristic might help in predicting
+whether the resulting bias is likely toward the null; however, more complex
+models (e.g., statistical models with adjustment for potential confounders) are
+trickier to diagnose in terms of the likely implications of differential versus
+non-differential measurement error [?].
+
+Not only can the exposure be measured with error, but so too can potential
+confounders and potential effect modifiers.
+
+We provide exposure assessments at the county level. This spatial scale allows
+for easy integration with health outcome data aggregated at the county level.
+However, there are also some things to consider when modeling associations
+at an ecological rather than individual level. The considerations are 
+particularly nuanced when investigating the health risks associated with a 
+disaster.
+
+Often in epidemiology, a study aims to estimate how an exposure changes the
+risk of a health outcome for an individual. The data might be collected for
+lots of people, but the key question is this---if a person is exposed, how 
+does his or her risk of the outcome change. If you are exposed to this hazard, 
+how much more likely are you to have the outcome? To get sick, or injured, or
+to die? This inference is at the individual level.
+
+Even when individual-level inference is the goal, sometimes data is only 
+available at an aggregated level. For example, privacy concerns might
+mean that health outcomes are only made available as a county-wide daily 
+count. You can fit a mode with this aggregated data, which will estimate the
+association between exposure and health risk at the ecological level. This
+contextual effect estimate is sometimes used as an estimate of individual-level
+association, but doing so can introduce ecological/multi-level bias. The 
+individual-level effect estimate can be very biased from the true association, 
+even to the point of reversing the effect estimate---estimating a protective
+effect when the true effect is detrimental, for example.
+
+This ecological fallacy plays out in two ways. First, if you use a single 
+exposure measurement for everyone in an area, some people will be misclassified
+(or have exposure measured with error, for a continuous metric) unless the 
+exposure is perfectly homogeneous across the area. This will typically not
+be the case. This exposure misclassification can bias estimates of the 
+association between exposure and outcome in the same way exposure 
+misclassification through any other mechanism would.
+
+Next, ecological bias could result from confounding, even if the confounders
+are controlled in the ecological-level model. When you aggregate data, you 
+lose information about how both the exposure and potential confounders vary
+across individuals within the area of aggregation. As a result, a factor
+could still confound the inference of an individual-level association, even 
+if it is controlled for at a population level in an ecological model. For 
+example, a study of the association between risk of pre-term birth and 
+tropical cyclone exposure could control for county-level smoking when 
+modeling county-level storm exposure and county-level rates of pre-term births.
+Even with this control, an observed association could result from differences
+in individual smoking status, if there is within-county variation in 
+smoking and if this has a different pattern across people in the county than
+variation in exposure from the county-wide exposure estimate.
+
+For disasters, there are added nuances. First, for exposure it may not just 
+matter what the level of exposure is in a person's immediate scope. Disasters
+bring physical hazards that can harm people directly, but also through indirect
+pathways. For example, a tropical cyclone can bring high winds that cause 
+power outages, and as a result those affected could be exposed to more outdoor
+hazards (outdoor air pollution, heat), struggle to safely store perishable
+food and medications, and lose means to power medical equipment. While extreme
+winds at a person's residence would increase their risk of a power outage, 
+outages could also be caused by damage to the grid in another part of the 
+community. In some cases, then, the level of exposure in a person's community
+may be as important in opening a pathway of risk as exposure at the person's
+immediate location.
+
+What's more, if the disaster has a large health impact, the health outcome of
+one person in the community could affect the risk of the outcome (or other
+adverse outcomes) for others. This situation is often only the case for 
+infectious diseases, where one person with the disease can spread it to others.
+However, if the community-wide impact is large enough, it can affect access
+to and effectiveness of medical care for everyone in the community. This
+effect has been seen recently with Covid 19---attempts to "flatten the curve"
+aim to avoid moving into a state where a community's health system becomes
+overwhelmed and can no longer deliver a typical level of care to those in the
+community. This effect could happen with either infectious or non-infectious
+diseases.
 
 When a proxy exposure estimates is used for many people covered by a study, this
-can result in a type of exposure measurement error called *Berkson error*. In this 
-case, the true exposure of each individual is randomly distributed around this 
-proxy/aggregate-level exposure estimate, with a mean exposure level among the 
-subjects that is equal to the proxy/aggregate-level measure. In other words, the 
-group as a whole is assigned a common exposure level, when in fact this is the 
-average exposure level across members of that group, but the individuals' 
+can result in a type of exposure measurement error called *Berkson error*. In
+this case, the true exposure of each individual is randomly distributed around
+this proxy/aggregate-level exposure estimate, with a mean exposure level among
+the subjects that is equal to the proxy/aggregate-level measure. In other words,
+the group as a whole is assigned a common exposure level, when in fact this is
+the average exposure level across members of that group, but the individuals'
 true exposure levels are randomly distributed around this mean group level to
@@
 ecological exposure estimates require us to assume that the exposure level is
-the same for all individuals in that county. It ignores/smooths over within-county
-variability in exposure.
+the same for all individuals in that county. It ignores/smooths over
+within-county variability in exposure.
 
@@
 deaths in a geographic area in a certain time period---is known as *ecological
-data*, and studies that base inference on this type of data are known as *ecological
-studies*. Terms include *ecological*, *aggregate*, *contextual level*.
+data*, and studies that base inference on this type of data are known as
+*ecological studies*. Terms include *ecological*, *aggregate*, *contextual
+level*.
 
@@
 > "In his seminal paper, Robinson (1950) distinguished between two types of
-correlation---ecological and individual. The former is obtained for a group
-of people, while the latter is estimated for indivisible units, such as individuals."
-[@portnov2007ecological]
+correlation---ecological and individual. The former is obtained for a group of
+people, while the latter is estimated for indivisible units, such as
+individuals." [@portnov2007ecological]
 
@@
 
-> "I propose three criteria for the identification of ecological fallacy; 
-all three of these should be present to confirm its existence: (1) Results
-must be obtrained with ecological (population) data; (2) Data must be inferred
-to individuals. One use of ecological studies is to explore individual-level 
-association when individual data are not available. When the focus of the 
-study was contextual or based on population effects and there is no inference
-to individuals, ecological fallacy is not possible. When only the first two 
+> "I propose three criteria for the identification of ecological fallacy; all
+three of these should be present to confirm its existence: (1) Results must be
+obtrained with ecological (population) data; (2) Data must be inferred to
+individuals. One use of ecological studies is to explore individual-level
+association when individual data are not available. When the focus of the study
+was contextual or based on population effects and there is no inference to
+individuals, ecological fallacy is not possible. When only the first two
 criteria are present---which is insufficient to affirm ecological fallacy---it
-is appropriate to acknowledge that there is a possible relationship and that further 
-study is required; (3) Results obtained with individual data are contradictory."
-[@idrovo2011three]
-
-> "One important breakdown in the analogy between ecologic-bias and individual-level
-confounding occurs because in etiologic research, the target of inference for
-both ecologic and individual-level studies is the same: Both study effects at the 
-individual level. For an individual-level study, these target effects are at the 
-same level as the units of analysis. But, for an ecologic study, these target 
-effects are at a finer level than the units of analysis. As a result, an ecologic 
-study can be unbiased for ecologic effects (in particular, ecologic confounding
-may be absent) and yet still be biased for individual-level effects. Many of the 
-classic social science examples of ecologic bias are of this form." [@greenland1994invited]
-
-> "We refer to an incorrect extrapolation from unbiased estimates of ecologic effects
-to unobserved individual-level effects as 'cross-level bias' (although the latter
-term is sometimes used to refer to any ecologic bias). As the examples suggest, 
-it is possible to detect cross-level bias if one can identify and observe 
-homogeneously exposed regions." [@greenland1994invited]
+is appropriate to acknowledge that there is a possible relationship and that
+further study is required; (3) Results obtained with individual data are
+contradictory." [@idrovo2011three]
+
+> "One important breakdown in the analogy between ecologic-bias and
+individual-level confounding occurs because in etiologic research, the target of
+inference for both ecologic and individual-level studies is the same: Both study
+effects at the individual level. For an individual-level study, these target
+effects are at the same level as the units of analysis. But, for an ecologic
+study, these target effects are at a finer level than the units of analysis. As
+a result, an ecologic study can be unbiased for ecologic effects (in particular,
+ecologic confounding may be absent) and yet still be biased for individual-level
+effects. Many of the classic social science examples of ecologic bias are of
+this form." [@greenland1994invited]
+
+> "We refer to an incorrect extrapolation from unbiased estimates of ecologic
+effects to unobserved individual-level effects as 'cross-level bias' (although
+the latter term is sometimes used to refer to any ecologic bias). As the
+examples suggest, it is possible to detect cross-level bias if one can identify
+and observe homogeneously exposed regions." [@greenland1994invited]
 
@@
 
-> "A broad definition of ecological fallacy is that certain data (e.g., illiteracy 
-rates by foreign-born) are unavailable, leading to the use of proxies and 
-erroneous estimates. However, we advocate a more narrow definition of this 
+> "A broad definition of ecological fallacy is that certain data (e.g.,
+illiteracy rates by foreign-born) are unavailable, leading to the use of proxies
+and erroneous estimates. However, we advocate a more narrow definition of this
 phenomenon, according to which 'ecological fallacy' refers solely to differences
-in conclusions which may be drawn from group-level data (e.g., average illiteracy 
-rates by foreign-born), as opposed to data obtained for individuals."
+in conclusions which may be drawn from group-level data (e.g., average
+illiteracy rates by foreign-born), as opposed to data obtained for individuals."
 [@portnov2007ecological]
 
-> "If different conclusions are drawn from the analysis of data upon their aggregations
-into units of different sizes (e.g., from individuals to townships and regions), these
-differences are commonly referred to as 'ecological fallacy'. In epidemiology, these
-differences are also known as ecological or cross-level bias [refs]. The ecological 
-bias is difficult to control and may lead, under certain circumstances, to spurious 
-effects [refs]. If epidemiological study does not consider the possibility of 
-ecological bias, misguided policy decisions may follow [refs]." [@portnov2007ecological]
-
-Ecological bias can *only* happen if there is, indeed, an individual-level effect:
+> "If different conclusions are drawn from the analysis of data upon their
+aggregations into units of different sizes (e.g., from individuals to townships
+and regions), these differences are commonly referred to as 'ecological
+fallacy'. In epidemiology, these differences are also known as ecological or
+cross-level bias [refs]. The ecological bias is difficult to control and may
+lead, under certain circumstances, to spurious effects [refs]. If
+epidemiological study does not consider the possibility of ecological bias,
+misguided policy decisions may follow [refs]." [@portnov2007ecological]
+
+Ecological bias can *only* happen if there is, indeed, an individual-level
+effect:
 
@@
 
-> "For this example, we have access to complete individual-level data, permitting
-a 'gold standard' individual-level analysis." [@wakefield2008overcoming]
+> "For this example, we have access to complete individual-level data,
+permitting a 'gold standard' individual-level analysis."
+[@wakefield2008overcoming]
 
@@
 
-> "In a semi-ecologic study, an ecologic exposure is combined with individual-level
-outcomes and confounders. A two-phase approach is particularly useful for
-such a study, with the phase 2 data corresponding to stratified sampling of 
-individual exposures." [@wakefield2008overcoming]
+> "In a semi-ecologic study, an ecologic exposure is combined with
+individual-level outcomes and confounders. A two-phase approach is particularly
+useful for such a study, with the phase 2 data corresponding to stratified
+sampling of individual exposures." [@wakefield2008overcoming]
 
 > "To implement a two-phase design, an initial phase 1 cross-classification by
-the binary disease outcome and stratification variables is required; in phase
-2, samples of individuals are drawn from each of the cross-classification cells, 
-with data on additional variables being drawn from the subsamples of individuals. 
-Intuitively, the stratified sampling is focused on informative cells, and 
-estimation methods use both phases of data for efficiency and to acknowledge
-the outcome-dependent sampling. In the simplest ecologic setting, the cross-classification
-is by outcome and area only, and if area is a surrogate for important risk factors
-this design will be efficient. We are particularly interested in situations where
-an initial classification is available by outcome, area, and confounders such as
-age and gender---this is the case in a semi-ecologic study. Phase 2 may then provide
-detailed exposure information on a subset of the phase 1 individuals." 
-[@wakefield2008overcoming]
+the binary disease outcome and stratification variables is required; in phase 2,
+samples of individuals are drawn from each of the cross-classification cells,
+with data on additional variables being drawn from the subsamples of
+individuals. Intuitively, the stratified sampling is focused on informative
+cells, and estimation methods use both phases of data for efficiency and to
+acknowledge the outcome-dependent sampling. In the simplest ecologic setting,
+the cross-classification is by outcome and area only, and if area is a surrogate
+for important risk factors this design will be efficient. We are particularly
+interested in situations where an initial classification is available by
+outcome, area, and confounders such as age and gender---this is the case in a
+semi-ecologic study. Phase 2 may then provide detailed exposure information on a
+subset of the phase 1 individuals." [@wakefield2008overcoming]
 
@@
 
-> "The results of our simulation study point to the benefits associated with combining
-the two sources of data, in terms of both bias and efficiency. Rather than supplement
-an ecologic study with individual-level data, it may be of interest to combine 
-existing individual-level data with external group-level data. Strategies that
-combine both types of data have been shown to alleviate participation bias and
-improve efficiency in case-control studies with missing data (31)." 
-[@wakefield2008overcoming]
-
-> "As with other observational studies, ecologic studies can give useful results if 
-biases such as those discussed here can be ruled out or quantified. Nevertheless, 
-bias evaluation can be especially difficult in ecologic studies of geographic
-regions because of the many potentially interacting covariates that may differ
-across regions. When biases cannot be ruled out with available data, further
-exploration will require individual-level studies." [@greenland1994invited]
+> "The results of our simulation study point to the benefits associated with
+combining the two sources of data, in terms of both bias and efficiency. Rather
+than supplement an ecologic study with individual-level data, it may be of
+interest to combine existing individual-level data with external group-level
+data. Strategies that combine both types of data have been shown to alleviate
+participation bias and improve efficiency in case-control studies with missing
+data (31)." [@wakefield2008overcoming]
+
+> "As with other observational studies, ecologic studies can give useful results
+if biases such as those discussed here can be ruled out or quantified.
+Nevertheless, bias evaluation can be especially difficult in ecologic studies of
+geographic regions because of the many potentially interacting covariates that
+may differ across regions. When biases cannot be ruled out with available data,
+further exploration will require individual-level studies."
+[@greenland1994invited]
 
@@
 
-> "Because of the unavailability of individual-level data, ecological data
-may be resorted to and may come in a variety of forms." [@wakefield2008overcoming]
+> "Because of the unavailability of individual-level data, ecological data may
+be resorted to and may come in a variety of forms." [@wakefield2008overcoming]
 
@@
 
-> "Population and health data are often routinely available in ecological, 
-that is group, form while the exposure data typically consist of a set of
-values recorded at monitor sites, or via one-off sampling." [@wakefield2006health]
+> "Population and health data are often routinely available in ecological, that
+is group, form while the exposure data typically consist of a set of values
+recorded at monitor sites, or via one-off sampling." [@wakefield2006health]
 
@@
 > "Although ecological studies are important to epidemiology (especially in
-environmental and social epidemiology), public health practitioners seem 
-afraid of ecological studies. It is a common practice to assume the presence of 
+environmental and social epidemiology), public health practitioners seem afraid
+of ecological studies. It is a common practice to assume the presence of
 ecological fallacy (Robinson 1950) and low-level validity when analyzing an
@@
 approach, although the importance of a multilevel causal approach is widely
-recognized (Diez-Roux 2002). In this sense, some authors suggest that it is
-as important to recognize the presence of ecological fallacy as to 
-recognize psychologistic or individualistic fallacy (Subramanian et al. 2009)."
+recognized (Diez-Roux 2002). In this sense, some authors suggest that it is as
+important to recognize the presence of ecological fallacy as to recognize
+psychologistic or individualistic fallacy (Subramanian et al. 2009)."
 [@idrovo2011three]
@@
 
-> "A confounder ... is associated with an exposure or risk factor for the outcome
-and with the outcome independent of the exposure, but is not on the causal pathway
-between the exposure and the outcome. Confounding occurs when the relationship
-between an exposure variable and the outcome variable is contaminated, so that the
-measure of association between these two variables is actually also capturing the
-effects of a third variable, the confounding variable. For example, many factors 
-are associated with income (exposure) and health (outcome), such as education, 
-employment or wealth? It may be that these factors explain part of the observed
-assocation." [@gunasekara2008glossary]
+> "A confounder ... is associated with an exposure or risk factor for the
+outcome and with the outcome independent of the exposure, but is not on the
+causal pathway between the exposure and the outcome. Confounding occurs when the
+relationship between an exposure variable and the outcome variable is
+contaminated, so that the measure of association between these two variables is
+actually also capturing the effects of a third variable, the confounding
+variable. For example, many factors are associated with income (exposure) and
+health (outcome), such as education, employment or wealth? It may be that these
+factors explain part of the observed assocation." [@gunasekara2008glossary]
 
@@
 
-> "Measurement error: The different between an observed variable and the variable 
-that belongs in a multiple regression equation. Measurement error is the random 
-or systematic error arising during data collection of variables. The measured
-variable x is measured with error eta, which is the distance from x to the 
-'true' value of x. Measurement error can produce bias such as attenuating the
-estimators of exposure variables, but may also have more complex effects. Longitudinal
-data analyses such as fixed effects models can actually augment the problem of 
-measurement error in mis-measured explanatory variables that change little over
-time." [@gunasekara2008glossary]
+> "Measurement error: The different between an observed variable and the
+variable that belongs in a multiple regression equation. Measurement error is
+the random or systematic error arising during data collection of variables. The
+measured variable x is measured with error eta, which is the distance from x to
+the 'true' value of x. Measurement error can produce bias such as attenuating
+the estimators of exposure variables, but may also have more complex effects.
+Longitudinal data analyses such as fixed effects models can actually augment the
+problem of measurement error in mis-measured explanatory variables that change
+little over time." [@gunasekara2008glossary]
 
@@
 
-> "Study validity refers to the degree to which the inferences drawn from a study are
-warranted when account is taken of the study methods; the representativeness of
-the study sample; and the nature of the population from which it is drawn.[5]
-There are two types of study validity. *Internal validity* is the degree to which the
-results of a study are correct for the sample of people being studied. *External 
-validity* (generalisability) is the degree to which the study results hold true
-for a population beyond the subjects in the study or other settings.[20]"
+> "Study validity refers to the degree to which the inferences drawn from a
+study are warranted when account is taken of the study methods; the
+representativeness of the study sample; and the nature of the population from
+which it is drawn.[5] There are two types of study validity. *Internal validity*
+is the degree to which the results of a study are correct for the sample of
+people being studied. *External validity* (generalisability) is the degree to
+which the study results hold true for a population beyond the subjects in the
+study or other settings.[20]"
 
@@
 
-- Differs by exposure metric. Easier for some (wind, rain) than for others (flood, 
-tornado)
-- For rain, we use a re-analysis product only available for the continental US (right?).
-However, there are similar gridded re-analysis products available for other regions 
-worldwide that could be used in a similar way.
-- For tornadoes, there may be national databases in other spots similar to the tornado
-database that we use, but as far as we're aware, no international one. 
+- Differs by exposure metric. Easier for some (wind, rain) than for others
+(flood, tornado)
+- For rain, we use a re-analysis product only available for the continental US
+(right?). However, there are similar gridded re-analysis products available for
+other regions worldwide that could be used in a similar way.
+- For tornadoes, there may be national databases in other spots similar to the
+tornado database that we use, but as far as we're aware, no international one.
 - Same for floods. 
-- For wind, the wind model could be generalizable with some tweaks. These include: 
+- For wind, the wind model could be generalizable with some tweaks. These
+include:
   + More landmasks
   + Direction of storm movement for Northern vs Southern Hemispheres
-  + Different wind averaging periods for the recorded best tracks data in different 
-  storm basins.
-  + Our wind model was derived with US data and so may need bias-correction or a shift
-  to a West Pacific--based model
-
-> "Tornadoes long have been recognized as a global phenomenon (e.g., Wegener 1917; 
-Feuerstein et al. 2005), having been recorded in every continent except Antarctica. 
-As such, numerous nationas share an interest in improving assessment of their 
-damage. International research and involvement in tornado survey work and damage
-and intensity scales is well underway." [@edwards2013tornado]
+  + Different wind averaging periods for the recorded best tracks data in
+  different storm basins.
+  + Our wind model was derived with US data and so may need bias-correction or a
+  shift to a West Pacific--based model
+
+> "Tornadoes long have been recognized as a global phenomenon (e.g., Wegener
+1917; Feuerstein et al. 2005), having been recorded in every continent except
+Antarctica. As such, numerous nationas share an interest in improving assessment
+of their damage. International research and involvement in tornado survey work
+and damage and intensity scales is well underway." [@edwards2013tornado]
 
@@
 is specific to the United States (the US NOAA's Storm Events database). To
-expand our exposure data for these hazards internationally, we would need to find
-similar databases for other affected countries. 
-[international flood and tornado databases? EM-DAT? Others?]
-
-The wind exposure data is based on modeling wind fields based on records of
-the storm's track and central intensity throughout the tracking period. These
-values are tracked for tropical cyclones around the world, and so it would be
-possible to expand the dataset to encompass other countries for wind exposure
-measurements. However, there are some practical issues that would need to 
-be addressed to do so with the approach we've taken.
-
-The software we are using to model wind [@stormwindmodel] has several features that
-make it specific to Atlantic basin storms. First, some steps in the model depend
-on whether the storm's center or the location being modeled is over land or water. 
-These steps use a landmask---a grid with a binary indicator of land or sea by 
-latitude and longitude. Currently, the software only provides a landmask that 
-covers the United States, the Gulf of Mexico, and areas of the Atlantic Ocean 
-near the US. Second, the model assumes that cyclonic winds are counterclockwise, 
-and so is valid only for the Northern Hemisphere. Third, the model assumes that
-the central intensity of the storm is based on [wind averaging period], which is 
-the standard used in the tracking data for Atlantic-basin storms, but is not
-a common standard around the world. Instead, some basins record central intensity
-measurements using averaging periods of ....
-
-Therefore, it would be relatively straightforward to expand the wind exposure data
-to cover other countries affected by Atlantic-basin storms, although this would 
-require some expansion of the area covered by our current landmask. More modifications
-would need to be made to the wind model software before it could be extended
-to storms in other basins, particularly basins that use different wind averaging 
-periods for their main tracking data and basins in the Southern Hemisphere. 
-Alternatively, other wind modeling software could be used for this exposure 
-assessment, as other groups have developed software that is more extensible to 
-multiple storm basins [Australian group's software].
-
-The rain exposure data would also be fairly easy to extend to cover other countries, 
-although again there would be a few practical limitations to extend the approach
-we've taken in developing the dataset covered in this manuscript. For this exposure, 
-we've used a re-analysis dataset. We've taken some steps to translate this gridded data
-to the county scale, so it can be integrated with the scale for many human impacts
-measurements. However, these steps have been well documents [ref] and would be easy to
-reproduce. The re-analysis dataset that we use is specific to the contiguous US [ref
-on NLDAS], but other similar re-analysis datasets exist for other locations [example
-of one with global coverage?].
-
-For all of these, we have provided exposure assessments at the county scale. Any of the 
-steps we have taken could be done for other countries at a similar geopolitical scale, 
-although the name of that areal unit will vary by country (e.g., municipalities in Mexico, 
-districts in India). For wind and rain exposures, the approach we take here could be
-used with any areal unit, as long as there are geographical shape files available for
-the area boundaries (although the modeling for wind may be to the area's geographic
-mean center rather than population mean center).
+expand our exposure data for these hazards internationally, we would need to
+find similar databases for other affected countries. [international flood and
+tornado databases? EM-DAT? Others?]
+
+The wind exposure data is based on modeling wind fields based on records of the
+storm's track and central intensity throughout the tracking period. These values
+are tracked for tropical cyclones around the world, and so it would be possible
+to expand the dataset to encompass other countries for wind exposure
+measurements. However, there are some practical issues that would need to be
+addressed to do so with the approach we've taken.
+
+The software we are using to model wind [@stormwindmodel] has several features
+that make it specific to Atlantic basin storms. First, some steps in the model
+depend on whether the storm's center or the location being modeled is over land
+or water. These steps use a landmask---a grid with a binary indicator of land or
+sea by latitude and longitude. Currently, the software only provides a landmask
+that covers the United States, the Gulf of Mexico, and areas of the Atlantic
+Ocean near the US. Second, the model assumes that cyclonic winds are
+counterclockwise, and so is valid only for the Northern Hemisphere. Third, the
+model assumes that the central intensity of the storm is based on [wind
+averaging period], which is the standard used in the tracking data for
+Atlantic-basin storms, but is not a common standard around the world. Instead,
+some basins record central intensity measurements using averaging periods of
+....
+
+Therefore, it would be relatively straightforward to expand the wind exposure
+data to cover other countries affected by Atlantic-basin storms, although this
+would require some expansion of the area covered by our current landmask. More
+modifications would need to be made to the wind model software before it could
+be extended to storms in other basins, particularly basins that use different
+wind averaging periods for their main tracking data and basins in the Southern
+Hemisphere. Alternatively, other wind modeling software could be used for this
+exposure assessment, as other groups have developed software that is more
+extensible to multiple storm basins [Australian group's software].
+
+The rain exposure data would also be fairly easy to extend to cover other
+countries, although again there would be a few practical limitations to extend
+the approach we've taken in developing the dataset covered in this manuscript.
+For this exposure, we've used a re-analysis dataset. We've taken some steps to
+translate this gridded data to the county scale, so it can be integrated with
+the scale for many human impacts measurements. However, these steps have been
+well documents [ref] and would be easy to reproduce. The re-analysis dataset
+that we use is specific to the contiguous US [ref on NLDAS], but other similar
+re-analysis datasets exist for other locations [example of one with global
+coverage?].
+
+For all of these, we have provided exposure assessments at the county scale. Any
+of the steps we have taken could be done for other countries at a similar
+geopolitical scale, although the name of that areal unit will vary by country
+(e.g., municipalities in Mexico, districts in India). For wind and rain
+exposures, the approach we take here could be used with any areal unit, as long
+as there are geographical shape files available for the area boundaries
+(although the modeling for wind may be to the area's geographic mean center
+rather than population mean center).
 
@@
 National Hurricane Center. With the assistance of Ms. Schumacher, we will update
-our tropical cyclone wind model software to appropriately handle 'Best Tracks' data from
-the China Meteorological Administration, incorporating conversion factors
-[@harper2010guidelines] to convert central storm maximum sustained wind
+our tropical cyclone wind model software to appropriately handle 'Best Tracks'
+data from the China Meteorological Administration, incorporating conversion
+factors [@harper2010guidelines] to convert central storm maximum sustained wind
 measurements to a common averaging period prior to applying further steps of the
@@
 
-There are some cases where damage reports are directly used to infer the intensity 
-of the weather event. For example, damage surveying is used following a tornado 
-to assign the tornado's class on the [enhanced] Fujita scale. However, this is not the
-intent of FEMA damage reports, which instead are used to determine post-disaster
-aid [?] and are influenced by a number of political factors.
-
-> "Because of the historic lack of direct measurements and remotely sensed tornado 
-wind speeds at or near group level, damage surveying has remained the most common 
-form for indicating tornado strength. ... The occurrence of a direct tornado strict 
-upon a fixed, sufficiently sturdy, and well-calibrated wind measuring station is quite
-rare. Only 31 direct in situ tornado observations are evident between 1894 and 2011."
-[@edwards2013tornado]
+There are some cases where damage reports are directly used to infer the
+intensity of the weather event. For example, damage surveying is used following
+a tornado to assign the tornado's class on the [enhanced] Fujita scale. However,
+this is not the intent of FEMA damage reports, which instead are used to
+determine post-disaster aid [?] and are influenced by a number of political
+factors.
+
+> "Because of the historic lack of direct measurements and remotely sensed
+tornado wind speeds at or near group level, damage surveying has remained the
+most common form for indicating tornado strength. ... The occurrence of a direct
+tornado strict upon a fixed, sufficiently sturdy, and well-calibrated wind
+measuring station is quite rare. Only 31 direct in situ tornado observations are
+evident between 1894 and 2011." [@edwards2013tornado]
 
@@
 We have added a note in the manuscript to let readers know that, because the
-data is shared through an R package, and because DesignSafe's [Workspace]
-allows users to work in R, the data can be fully accessed and used for those working
-on the DesignSafe platform by using the same code in a DesignSafe Jupyter notebook
+data is shared through an R package, and because DesignSafe's [Workspace] allows
+users to work in R, the data can be fully accessed and used for those working on
+the DesignSafe platform by using the same code in a DesignSafe Jupyter notebook
 that would otherwise be used on a local computer.
@@
 This is good, because otherwise there would be a good deal of overlap between
-the data on DesignSafe and that in the R package. R packages already created 
-a citable source, comparable to the DOI assigned to data by DesignSafe when
-it is published. Further, R packages include licensing information, as does
-DesignSafe for published data. Finally, if the data were shared through both 
-platforms, it would be critical to make sure that it was clear that the two 
-datasets are analogous and also to keep them up to date with each other, which
-would add an extra layer of work with each update to the R package (we have 
-been updating several hazards as new data becomes available). It is wonderful 
-that DesignSafe's interface with R avoids this problem by allowing its users
-direct access to our package and its data through that path, as it allows for 
-a single, citable, licensed version of the data stored and shared from a single
-location, with no question of syncing separate versions of the data or clarifying
+the data on DesignSafe and that in the R package. R packages already created a
+citable source, comparable to the DOI assigned to data by DesignSafe when it is
+published. Further, R packages include licensing information, as does DesignSafe
+for published data. Finally, if the data were shared through both platforms, it
+would be critical to make sure that it was clear that the two datasets are
+analogous and also to keep them up to date with each other, which would add an
+extra layer of work with each update to the R package (we have been updating
+several hazards as new data becomes available). It is wonderful that
+DesignSafe's interface with R avoids this problem by allowing its users direct
+access to our package and its data through that path, as it allows for a single,
+citable, licensed version of the data stored and shared from a single location,
+with no question of syncing separate versions of the data or clarifying
 licensing and citation practices across two storage locations.
@@
 
-> "DesignSafe plays an important role in integrating the various NHERI components
-and the research taking place at NHERI facilities, but also has the broader
-goal of enabling transformative research in natural hazards and engineering
-across the numerous technical disciplines engaged in this field." 
+> "DesignSafe plays an important role in integrating the various NHERI
+components and the research taking place at NHERI facilities, but also has the
+broader goal of enabling transformative research in natural hazards and
+engineering across the numerous technical disciplines engaged in this field."
 [@rathje2017designsafe]
@@
 in natural hazards engineering. DesignSafe provides a comprehensive CI that
-supports the full research lifecycle from planning to execution to analysis
-to publication and curation." [@rathje2017designsafe]
+supports the full research lifecycle from planning to execution to analysis to
+publication and curation." [@rathje2017designsafe]
 
 > "The future of natural hazards engineering research requires the integration
-of diverse data sets from a variety of sources including experiments, computational
-simulation, and field reconnaissance, as well as a variety of research disciplines
-including earth science, social science, building science, and architecture. 
-The DesignSafe cyberinfrastructure has been designed to provide the functionalities
-that will enable transformative research in natural hazards engineering. By adopting
-a cloud strategy, DesignSafe allows for a fundamental change in the way that
-research is performed. It provides a comprehensive cyberinfrastructure that
-supports research workflows, data analysis, and visualization, as well as the 
-full lifecycle of experimental, field, and computational research required by 
-engineers and scientists to effectively address the threats posed to civil 
-infrastructure by natural hazards. The integration of data and computation 
-in the cloud will enable new research discoveries in natural hazards engineering, 
-which in turn can lead to more hazard-resilient civil infrastructure." 
-[@rathje2017designsafe]
+of diverse data sets from a variety of sources including experiments,
+computational simulation, and field reconnaissance, as well as a variety of
+research disciplines including earth science, social science, building science,
+and architecture. The DesignSafe cyberinfrastructure has been designed to
+provide the functionalities that will enable transformative research in natural
+hazards engineering. By adopting a cloud strategy, DesignSafe allows for a
+fundamental change in the way that research is performed. It provides a
+comprehensive cyberinfrastructure that supports research workflows, data
+analysis, and visualization, as well as the full lifecycle of experimental,
+field, and computational research required by engineers and scientists to
+effectively address the threats posed to civil infrastructure by natural
+hazards. The integration of data and computation in the cloud will enable new
+research discoveries in natural hazards engineering, which in turn can lead to
+more hazard-resilient civil infrastructure." [@rathje2017designsafe]
 
@@
 exposed to one hazard, they are typically also exposed to a second), this would
-have several implications for epidemiological research, especially for multi-year
-studies: 
+have several implications for epidemiological research, especially for
+multi-year studies:
   + From a statistical modeling point of view, if you tried to include each
   hazard as a separate independent variable, you would run into problems with
-  collinearity if fitting the study data with a generalized linear model or other
-  regression model. One implication of this is that the confidence intervals on 
-  effect estimates for the two variables would be very inflated, because the model
-  would struggle to figure out which of the two variables the weight should be placed
-  on, and so this uncertainty would be reflected in the size of the associated 
-  confidence intervals.
-  + If you see an effect, it would be very hard to try to determine if the causal 
-  pathway goes through the first hazard or the second. To mitigate effects, it is 
-  important to understand the pathway through which they happen, and it might be easier
-  to get clues towards pathways if the hazards are better separated. (Analogy: if you
-  were able to conduct a controlled experiment to look at the effects of each hazard, 
-  jointly and in combination, you would design separation in the assignment of these
-  treatments, so you could distinguish. We can't do that because we are limited to 
-  an observational analysis, but the area of causal inference is doing a lot to 
-  explore how we can design observational studies and analyze data to come closer to 
-  what we might be able to see with a randomized control study, and which of those methods
-  we might be able to use for tropical cyclone epidemiology will depend on the interplay of
-  these patterns of exposures.)
-2. These results will not be surprising to an atmospheric scientist / disaster researcher,
-but there are environmental epidemiologists who have done research on other exposures and
-are now starting to study tropical cyclones (session at ISEE, for example). It is useful 
-for epidemiologists entering into tropical cyclone research to understand these facets of
-tropical cyclone exposure, including the distinction between the very coastal wind exposures
-and other hazards that tend to reach further inland. 
-3. There are epidemiological studies that have used only one hazard for exposure assessment.
-This has most often been the case for wind. If this exposure assessment is meant to 
-capture "tropical cyclone exposure" in general, with all the hazards that are related to 
-tropical cyclones, then there is the chance for exposure misclassification if this single
-hazard is used. 
-4. In the past, it was a very complex issue to try to pull together data on 
-all these hazards. We agree that now that we have provided a dataset with all of them, 
-epidemiologists should be thinking about multi-hazard analyses. However, there still 
-will be some barriers to that. [Recent work on multi-pollutant exposures and the complexities
-associated with that.] If one of the hazards dominates in explaining associated health 
-risk, then in some cases it might be reasonable to focus on that hazard, but in that 
-case it is critical, based on our results, to remember that the exposure assessment is
-not capturing all hazards of the storms.
+  collinearity if fitting the study data with a generalized linear model or
+  other regression model. One implication of this is that the confidence
+  intervals on effect estimates for the two variables would be very inflated,
+  because the model would struggle to figure out which of the two variables the
+  weight should be placed on, and so this uncertainty would be reflected in the
+  size of the associated confidence intervals.
+  + If you see an effect, it would be very hard to try to determine if the
+  causal pathway goes through the first hazard or the second. To mitigate
+  effects, it is important to understand the pathway through which they happen,
+  and it might be easier to get clues towards pathways if the hazards are better
+  separated. (Analogy: if you were able to conduct a controlled experiment to
+  look at the effects of each hazard, jointly and in combination, you would
+  design separation in the assignment of these treatments, so you could
+  distinguish. We can't do that because we are limited to an observational
+  analysis, but the area of causal inference is doing a lot to explore how we
+  can design observational studies and analyze data to come closer to what we
+  might be able to see with a randomized control study, and which of those
+  methods we might be able to use for tropical cyclone epidemiology will depend
+  on the interplay of these patterns of exposures.)
+2. These results will not be surprising to an atmospheric scientist / disaster
+researcher, but there are environmental epidemiologists who have done research
+on other exposures and are now starting to study tropical cyclones (session at
+ISEE, for example). It is useful for epidemiologists entering into tropical
+cyclone research to understand these facets of tropical cyclone exposure,
+including the distinction between the very coastal wind exposures and other
+hazards that tend to reach further inland.
+3. There are epidemiological studies that have used only one hazard for exposure
+assessment. This has most often been the case for wind. If this exposure
+assessment is meant to capture "tropical cyclone exposure" in general, with all
+the hazards that are related to tropical cyclones, then there is the chance for
+exposure misclassification if this single hazard is used.
+4. In the past, it was a very complex issue to try to pull together data on all
+these hazards. We agree that now that we have provided a dataset with all of
+them, epidemiologists should be thinking about multi-hazard analyses. However,
+there still will be some barriers to that. [Recent work on multi-pollutant
+exposures and the complexities associated with that.] If one of the hazards
+dominates in explaining associated health risk, then in some cases it might be
+reasonable to focus on that hazard, but in that case it is critical, based on
+our results, to remember that the exposure assessment is not capturing all
+hazards of the storms.
 
@@
 
-It is obvious that it is typically better, but in cases where the proxy is much more
-practical to measure, it may need to be more than just a bit better, but 
+It is obvious that it is typically better, but in cases where the proxy is much
+more practical to measure, it may need to be more than just a bit better, but
 instead substantially better. Also, there could be some cases where the proxy
-might actually be better---for example, if the proxy can be measured with 
-very low level, while the direct exposure measure might only be measurable 
-with high error. In that case, the proxy could be the better metric and might
-provide a more precise estimate of the association with health risk.
+might actually be better---for example, if the proxy can be measured with very
+low level, while the direct exposure measure might only be measurable with high
+error. In that case, the proxy could be the better metric and might provide a
+more precise estimate of the association with health risk.
 
@@
 
-We agree that it is helpful to clarify for readers that these independent hazards
-can come together in a storm event at different intensities and at different time
-windows. We have added some text on this:
+We agree that it is helpful to clarify for readers that these independent
+hazards can come together in a storm event at different intensities and at
+different time windows. We have added some text on this:
 
@@
 confounder or it will be misspecified and be prone to bias. However, if the
-exposure of interest and the potential confounder are very strongly correlated/in
-agreement, then including both as independent variables in a regression model
-often causes numerical problems in fitting the regression model, including 
-instability in the estimated coefficients and inflated variance. In very extreme
-cases, the statistical software may not be able to invert the matrix [? only for
-linear regression?].
+exposure of interest and the potential confounder are very strongly
+correlated/in agreement, then including both as independent variables in a
+regression model often causes numerical problems in fitting the regression
+model, including instability in the estimated coefficients and inflated
+variance. In very extreme cases, the statistical software may not be able to
+invert the matrix [? only for linear regression?].
 
@@
 most large storms (for multi-storm studies). This knowledge will help them in
-identifying potential confounders if they are interested in studying the 
-effects of a single hazard of the storm and also in determining if control for
-these potential confounders among other storm hazards might cause numerical 
-problems or other collinearity-associated problems when fitting a regression 
-model.
+identifying potential confounders if they are interested in studying the effects
+of a single hazard of the storm and also in determining if control for these
+potential confounders among other storm hazards might cause numerical problems
+or other collinearity-associated problems when fitting a regression model.
 
@@
 
-If a researcher wants a single measure for the storm as a whole, then it is helpful
-to know how well exposure assessments based on the hazard components of the storm
-agree with each other. If there is strong agreement, then any of those hazards could
-be used as a primary marker of exposure assessment. This is conceptually similar to 
-checking for observer / rater agreement.
-
-> "When the row and column variable srepresent different observers' rating the same
-subjects or objects, interest is focused on *observer agreement* rather than mere 
-association. In this case, measures and tests of agreement provide a method of
-assessing the reliability of a subjective classification or assessment procedure."
-[@friendly2015discrete]
+If a researcher wants a single measure for the storm as a whole, then it is
+helpful to know how well exposure assessments based on the hazard components of
+the storm agree with each other. If there is strong agreement, then any of those
+hazards could be used as a primary marker of exposure assessment. This is
+conceptually similar to checking for observer / rater agreement.
+
+> "When the row and column variable srepresent different observers' rating the
+same subjects or objects, interest is focused on *observer agreement* rather
+than mere association. In this case, measures and tests of agreement provide a
+method of assessing the reliability of a subjective classification or assessment
+procedure." [@friendly2015discrete]
 
@@
 one category lower (on an ordinal scale) than another rater. More generally,
-measures of agreement must take account of the marginal frequencies with which two 
-raters use the categories. If observers tend to use the categories with different
-frequency, this will affect measures of agreement." [@friendly2015discrete]
+measures of agreement must take account of the marginal frequencies with which
+two raters use the categories. If observers tend to use the categories with
+different frequency, this will affect measures of agreement."
+[@friendly2015discrete]
 
 When you have an array of two or more exposures that are strongly correlated,
-including them together in a regression model can cause problems from collinearity
-[@schisterman2017collinearity].
+including them together in a regression model can cause problems from
+collinearity [@schisterman2017collinearity].
 
@@
 factors that may be highly correlated with each other, causing concern about the
-impact of collinearity when attempting to identify individual effects." [@schisterman2017collinearity]
+impact of collinearity when attempting to identify individual effects."
+[@schisterman2017collinearity]
 
@@
 
-> "Expressly challenging the still-pervasive tendency of epidemiologists to think
-in terms of single 'agents' causing discrete diseases, the provocative metaphor
-and model of the 'web' invited epidemiologists to embrace a more sophisticated
-view of causality." [@krieger1994epidemiology]
+> "Expressly challenging the still-pervasive tendency of epidemiologists to
+think in terms of single 'agents' causing discrete diseases, the provocative
+metaphor and model of the 'web' invited epidemiologists to embrace a more
+sophisticated view of causality." [@krieger1994epidemiology]
 
@@
 > "Using this model, MacMahon et al. drew several important inferences about
-prevention and research that remain part of epidemiologic thinking to this 
-day. Arguing that 'to effect preventative measures, it is not necessary to 
-understand causal mechanisms in their entirity', they stated that 'even knowledge
-of one small component may allow some degree of prevention' since 'wherever the
-chain is broken the disease will be prevented'." [@krieger1994epidemiology]
+prevention and research that remain part of epidemiologic thinking to this day.
+Arguing that 'to effect preventative measures, it is not necessary to understand
+causal mechanisms in their entirity', they stated that 'even knowledge of one
+small component may allow some degree of prevention' since 'wherever the chain
+is broken the disease will be prevented'." [@krieger1994epidemiology]
 
@@
 When the aim is to explore the influences of multiple associated factors, there
-are some different strategies that can be used. These include multiple regression
-and path analysis:
-
-> "Evolutionary ecologists have recently turned to path analysis, multiple regression,
-and related techniques to analyse systems of multiple causality. While these
-techniques are often seen as competitors, they are, in fact, related, and as a 
-result the problems of these techniques are similar." [@petraitis1996inferring]
+are some different strategies that can be used. These include multiple
+regression and path analysis:
+
+> "Evolutionary ecologists have recently turned to path analysis, multiple
+regression, and related techniques to analyse systems of multiple causality.
+While these techniques are often seen as competitors, they are, in fact,
+related, and as a result the problems of these techniques are similar."
+[@petraitis1996inferring]
 
@@
 
-> "The dependence of regression and path coefficients on the correlation structure
-is often glossed over. Yet this dependence means that inferences about the strengths
-of path coefficients are conditional upon the correlational structure among the 
-predictor variables. If the correlation structure of the sample does not
-match the correlational structure of the population then regression and path 
-coefficients will not estimate relative strengths accurately. This can occur if
-sampling is not random or if analysis is based on data from experiments with 
-factorial designs, which break the correlational structure amnog the predictor
-variables." [@petraitis1996inferring]
+> "The dependence of regression and path coefficients on the correlation
+structure is often glossed over. Yet this dependence means that inferences about
+the strengths of path coefficients are conditional upon the correlational
+structure among the predictor variables. If the correlation structure of the
+sample does not match the correlational structure of the population then
+regression and path coefficients will not estimate relative strengths
+accurately. This can occur if sampling is not random or if analysis is based on
+data from experiments with factorial designs, which break the correlational
+structure amnog the predictor variables." [@petraitis1996inferring]
 
@@
 
-Extreme collinearity can create numerical problems [@schisterman2017collinearity].
-
-> "Under the extremely unlikely scenario of extreme collinearity [(abs. rho of 0.999 
-or higher)], the numerical matrix inversion required for estimation fails, resulting
-in unstable and biased predictors." [@schisterman2017collinearity]
-
-> "Only for extreme values of rho cloes to 1 or -1 is beta-squareed nonidentifiable
-due to numerical instability; otherwise, the correctly specified model produces
-the most desirable (e.g., unbiased) estimates of the total effect of the exposure
-on the outcome." [@schisterman2017collinearity]
+Extreme collinearity can create numerical problems
+[@schisterman2017collinearity].
+
+> "Under the extremely unlikely scenario of extreme collinearity [(abs. rho of
+0.999 or higher)], the numerical matrix inversion required for estimation fails,
+resulting in unstable and biased predictors." [@schisterman2017collinearity]
+
+> "Only for extreme values of rho cloes to 1 or -1 is beta-squareed
+nonidentifiable due to numerical instability; otherwise, the correctly specified
+model produces the most desirable (e.g., unbiased) estimates of the total effect
+of the exposure on the outcome." [@schisterman2017collinearity]
 
@@
 
-> "Collinearity occurs in multiple regression models when two (or more---multicollinearity)
-exposure variables are included in the model but are so similar to each other that
-they are essentially measuring at least part of the same thing. Collinearity 
-creates problems with interpreting the analysis by affecting the standard errors
-of the variables (as it increases variance)[24] and by causing biased estimates for one
-or both of the collinear terms (sometimes dramatically so) [25]. If there is perfect
-collinearity between two variables then one should be dropped from the model. However, 
-dropping exposure variables to reduce multicollinearity may lead to bias in the model
-from dropping useful information." [@gunasekara2008glossary]
-
-> "It might be argued in some cases that it is appropriate to include (collinear) 
-variables in a model even when they are highly correlated, because the theoretical 
-basis for including the variables is strong and the results of the model are consistent
-with expectations [10]. In such cases, models with and without the variables would be
-tested and compared. Such consideraion of alternative models tends to be described
-as an issue of (mis)-specification in econometrics. However, in epidemiology this
-tends to be described in terms of mediating and *intermediary variables*. Thus, an
-epidemiologist might argue for including an intermediate variable in a model (and 
-testing it against a model without the variable) because including that variable might 
-elucidate direct and indirect pathways between the exposure and outcome variables. 
-An econometrician might do the same thing but describe this in terms of finding the
-correct model specification." [@gunasekara2008glossary]
+> "Collinearity occurs in multiple regression models when two (or
+more---multicollinearity) exposure variables are included in the model but are
+so similar to each other that they are essentially measuring at least part of
+the same thing. Collinearity creates problems with interpreting the analysis by
+affecting the standard errors of the variables (as it increases variance)[24]
+and by causing biased estimates for one or both of the collinear terms
+(sometimes dramatically so) [25]. If there is perfect collinearity between two
+variables then one should be dropped from the model. However, dropping exposure
+variables to reduce multicollinearity may lead to bias in the model from
+dropping useful information." [@gunasekara2008glossary]
+
+> "It might be argued in some cases that it is appropriate to include
+(collinear) variables in a model even when they are highly correlated, because
+the theoretical basis for including the variables is strong and the results of
+the model are consistent with expectations [10]. In such cases, models with and
+without the variables would be tested and compared. Such consideraion of
+alternative models tends to be described as an issue of (mis)-specification in
+econometrics. However, in epidemiology this tends to be described in terms of
+mediating and *intermediary variables*. Thus, an epidemiologist might argue for
+including an intermediate variable in a model (and testing it against a model
+without the variable) because including that variable might elucidate direct and
+indirect pathways between the exposure and outcome variables. An econometrician
+might do the same thing but describe this in terms of finding the correct model
+specification." [@gunasekara2008glossary]
 
@@
 
-There are study designs that might help to separate different effects. For example, 
-some matching designs might help in creating a subset of data in which the association 
-that exists in the full dataset is broken (e.g., @leal2012multicollinearity).
+There are study designs that might help to separate different effects. For
+example, some matching designs might help in creating a subset of data in which
+the association that exists in the full dataset is broken (e.g.,
+@leal2012multicollinearity).
 
@@
 You could also explore techniques of dimension reduction, to create a smaller
-subset of independent variables for which the correlation is broken. For example, 
-you could use PCA [@leal2012multicollinearity].
-
-There are other techniques that allow some bias in estimates in an attempt to reduce
-problems from collinearity, like ridge regression [@petraitis1996inferring].
+subset of independent variables for which the correlation is broken. For
+example, you could use PCA [@leal2012multicollinearity].
+
+There are other techniques that allow some bias in estimates in an attempt to
+reduce problems from collinearity, like ridge regression
+[@petraitis1996inferring].
 
 > "Several biased estimation techniquest ahve been proposed as a way to combat
-collinearity (Myers 1990). These are controversial methods. Ridge regression, for
-example, biases the regression coefficients in order to improve the model's 
-ability to predict the criterion variable. The estimates of the individual regression
-coefficients may be very biased even though when taken together they provide a 
-very good prediction of the criterion variable." [@petraitis1996inferring]
+collinearity (Myers 1990). These are controversial methods. Ridge regression,
+for example, biases the regression coefficients in order to improve the model's
+ability to predict the criterion variable. The estimates of the individual
+regression coefficients may be very biased even though when taken together they
+provide a very good prediction of the criterion variable."
+[@petraitis1996inferring]
 
@@
 
-Epidemiology often aims to inform policy decisions. In this case, it can be important
-to characterize the separate effects of different associated factors that contribute
-to the health risk [@leal2012multicollinearity].
+Epidemiology often aims to inform policy decisions. In this case, it can be
+important to characterize the separate effects of different associated factors
+that contribute to the health risk [@leal2012multicollinearity].
 
@@
 > "However, many neighborhood characteristics, especially those related to the
-densities of physical features and services, are strongly correlated with 
-each other. Therefore, it is not clear whether it is even possible to disentangle
-the effects of these different environmental dimensions, even if they are hypothesized
-to influence obesity risk through distinct causal pathways." [@leal2012multicollinearity]
+densities of physical features and services, are strongly correlated with each
+other. Therefore, it is not clear whether it is even possible to disentangle the
+effects of these different environmental dimensions, even if they are
+hypothesized to influence obesity risk through distinct causal pathways."
+[@leal2012multicollinearity]
 
@@
 
-As we note in our general response to your comments, we agree in general with your
-instinct here. However, there are some subtle points to note.
-
-First, we understand the potential appeal of creating a single exposure 
-summary score based on a summation of exposures. This simplifies some of the 
-statistical modeling, for example, and also provides a single exposure estimate to 
-provide as the key study conclusion. However, in many areas of tropical cyclone epidemiology, 
-it is unlikely that exposure to each type of hazard should be given equal weight. For 
-example, an analysis of respiratory health in the months after the storm might hypothesize
-a strong connection with flooding, and smaller connections with rainfall and winds (through
-the pathway of property damage). Use of a pre-defined summation of exposure to each hazard
-would mask some of the potential association.
-
-While a summary statistic could be defined that adds a weight to each exposure, it would
-often be unclear *a priori* what weight each hazard exposure should be given. A better 
-method might be to use a regression framework, where each hazard exposure is included as
-a separate explanatory variable, and the model uses the observed data to determine the 
-appropriate weight for each hazard in terms of modifying risk to the outcome of interest. 
-To explore the influence of multiple concurrent hazards, interaction terms could be added.
-
-However, if our analysis had found that hazard exposures follow very similar patterns, in 
-terms of the counties exposed during a storm, then this strong agreement would result in 
-a number of problems in fitting and interpreting these types of models. First, strong
-agreement in two or more of the explanatory variables in a regression model leads to 
-collinearity, which among other implications often leads to a large inflation of confidence
-intervals for the effect estimates of the correlated explanatory variables. Further, if 
-the hazard exposures were in strong agreement, then it would be difficult or impossible to
-disentagle the effects of the separate hazards, as a strong effect for one would be 
-impossible to interpret with confidence, since the strong agreement in the data (and resulting
-low co-variability in the variables in the data) could cause the model to mistakenly 
-attribute the effect of one hazard to the other. Further, it may become impossible in this
-case to estimate synergistic effects between the two, as an indicator of an interaction
-between the two exposures would be almost collinear with the two main effects.
-
-In our analysis, we find that this is usually not the case for the four tropical cyclone 
-hazards that we consider. Instead, we find that for most storms, the hazards are fairly 
-well-separated in terms of the counties they effect. There is a particularly strong pattern
-where inland counties can be exposed to extreme rain and flooding, but not wind. This 
-separation opens the door for some interesting study designs and statistical modeling. 
-For example, these patterns suggest that a study could investigate whether storm-associated
-rains alone create a health risk, as there are many exposures with only rain in inland 
-counties, and so health associations could be explored here and contrasted with coastal 
-areas that experienced both wind and extreme rain.
+As we note in our general response to your comments, we agree in general with
+your instinct here. However, there are some subtle points to note.
+
+First, we understand the potential appeal of creating a single exposure summary
+score based on a summation of exposures. This simplifies some of the statistical
+modeling, for example, and also provides a single exposure estimate to provide
+as the key study conclusion. However, in many areas of tropical cyclone
+epidemiology, it is unlikely that exposure to each type of hazard should be
+given equal weight. For example, an analysis of respiratory health in the months
+after the storm might hypothesize a strong connection with flooding, and smaller
+connections with rainfall and winds (through the pathway of property damage).
+Use of a pre-defined summation of exposure to each hazard would mask some of the
+potential association.
+
+While a summary statistic could be defined that adds a weight to each exposure,
+it would often be unclear *a priori* what weight each hazard exposure should be
+given. A better method might be to use a regression framework, where each hazard
+exposure is included as a separate explanatory variable, and the model uses the
+observed data to determine the appropriate weight for each hazard in terms of
+modifying risk to the outcome of interest. To explore the influence of multiple
+concurrent hazards, interaction terms could be added.
+
+However, if our analysis had found that hazard exposures follow very similar
+patterns, in terms of the counties exposed during a storm, then this strong
+agreement would result in a number of problems in fitting and interpreting these
+types of models. First, strong agreement in two or more of the explanatory
+variables in a regression model leads to collinearity, which among other
+implications often leads to a large inflation of confidence intervals for the
+effect estimates of the correlated explanatory variables. Further, if the hazard
+exposures were in strong agreement, then it would be difficult or impossible to
+disentagle the effects of the separate hazards, as a strong effect for one would
+be impossible to interpret with confidence, since the strong agreement in the
+data (and resulting low co-variability in the variables in the data) could cause
+the model to mistakenly attribute the effect of one hazard to the other.
+Further, it may become impossible in this case to estimate synergistic effects
+between the two, as an indicator of an interaction between the two exposures
+would be almost collinear with the two main effects.
+
+In our analysis, we find that this is usually not the case for the four tropical
+cyclone hazards that we consider. Instead, we find that for most storms, the
+hazards are fairly well-separated in terms of the counties they effect. There is
+a particularly strong pattern where inland counties can be exposed to extreme
+rain and flooding, but not wind. This separation opens the door for some
+interesting study designs and statistical modeling. For example, these patterns
+suggest that a study could investigate whether storm-associated rains alone
+create a health risk, as there are many exposures with only rain in inland
+counties, and so health associations could be explored here and contrasted with
+coastal areas that experienced both wind and extreme rain.
 
@@
 
-This is a great idea. In the original manuscript, we describe this lower correlation
-at higher precipitation values through the use of examples from specific storms, and
-discuss the potential reasons for it, but measuring the correlation specifically within
-these higher-precipitation events would add more quantitative evidence to this discussion.
+This is a great idea. In the original manuscript, we describe this lower
+correlation at higher precipitation values through the use of examples from
+specific storms, and discuss the potential reasons for it, but measuring the
+correlation specifically within these higher-precipitation events would add more
+quantitative evidence to this discussion.
 
@@
 
-We have added estimates of these correlations specifically for high-precipitation
-events in a new table in the Supplemental Material (Table S1, reproduced 
-as Table \@ref(tab:highprecipcorr) of this response document).
+We have added estimates of these correlations specifically for
+high-precipitation events in a new table in the Supplemental Material (Table S1,
+reproduced as Table \@ref(tab:highprecipcorr) of this response document).
 
@@
 [A better future direction might be through regression models with multiple
-explanatory pathways. This allows the model to fit different weights to each. Other
-models can also be explored, like regression trees, which might have an easier time
-capturing non-linearity (if exposure is included as a continuous value, as with 
-wind and rain) and interactions when there are co-exposures to two or more hazards.
-Both of these are able (and will) either exclude or give a very low weight to an 
-explanatory variable that does not help much in explaining variation in the health 
-outcome.]
+explanatory pathways. This allows the model to fit different weights to each.
+Other models can also be explored, like regression trees, which might have an
+easier time capturing non-linearity (if exposure is included as a continuous
+value, as with wind and rain) and interactions when there are co-exposures to
+two or more hazards. Both of these are able (and will) either exclude or give a
+very low weight to an explanatory variable that does not help much in explaining
+variation in the health outcome.]
 
@@
 **Response:** 
+This is a great suggestion and helps provided quantitative support to the text
+in the paper that the reviewer highlighted when making this comment. We have
+added a table in the Supplemental Material that provides the Spearman
+correlation coefficients for both all tropical cyclones to come near the sample
+counties (replicating the analysis shown in Figure 2 of the main text) and
+then when the analysis is limited to only those events with cumulative
+precipitation of 75 mm or higher based on either data source.
+
+In several cases, the correlation was much weaker when limiting analysis to
+these severe rainfall events. For example, in Miami-Dade County, FL, the
+Spearman correlation coefficient was 0.94 for all 64 tropical cyclones that came
+near the county over the analysis period, while it was only 0.49 among the 18
+storms with cumulative precipitation of 75 mm or more. There was a similar
+reduction in Mobile County, AL. However, for some counties, the correlation
+remained high (e.g., Harris County, TX; Orleans Parish, LA), although in all
+cases there are small numbers of extreme precipitation events (5--20 across
+these sample counties), and so there is likely substantial random variation in
+these correlation estimates, since they are based on small numbers of 
+observations.
+
+[Include copy of table here.]
+
+Finally, we have added a bit to the text that the reviewer cited to mention 
+this point and to direct readers to the added supplemental table:
+
+> Text in revised manuscript (relevant addition in bold):
+"The rainfall data are generally well-correlated with ground-based
+observations, but may sometimes underestimate very high rainfall values (Figure
+2), **and in some counties the correlation was substantially lower when 
+considering only tropical cyclones with cumulative local rainfall of 75 mm
+or more (Supplemental Table 1)**."
 
@@
 **Response:** 
+[Add new analysis of what percent of counties misclassified if distance 
+is used as proxy for each hazard?]
 
@@
 **Response:** 
+We agree. This type of approach has not always been easy in the past, but can
+be easily implemented with the dataset we publish here. We have added to the
+text on this point:
+
+> Text in revised manuscript (relevant addition in bold):
+"For these storms, it may be possible to assess exposure to multiple
+hazards of the storm using a single metric, perhaps even a proxy like the
+distance between the county and the storm’s track. **With the dataset we 
+describe in this paper, however, there is little need to limit analysis
+based on exposure to a single hazard or proxy, although multi-hazard studies
+of storms with high agreement among hazard exposures should look out for 
+modeling issues from multicollinearity.**"
 
@@
 
-> "In order to allow comparison of surface conditions at the same time throughout 
-the world, the groups regulating the collection of hydrometeorological data have
-established a standard time for taking observations. These observations will have
-times observed within the 10 minutes preceding the UTC standard time. Hence 
-surface observations for meteorological use have timestamps reflecting 0000, 
-0300, 0600, 0900, 1200, 1500, 1800, 2100 UTC." [@usnwstime]
+> "In order to allow comparison of surface conditions at the same time
+throughout the world, the groups regulating the collection of
+hydrometeorological data have established a standard time for taking
+observations. These observations will have times observed within the 10 minutes
+preceding the UTC standard time. Hence surface observations for meteorological
+use have timestamps reflecting 0000, 0300, 0600, 0900, 1200, 1500, 1800, 2100
+UTC." [@usnwstime]
 
@@
 worldwide at 00 and 12 UTC (coordinated universal time). High-threat situations,
-such as impending TC landfall, may dictate observations at 6-h or even 3-h intervals. 
-Interpolating rawinsonde observations to the model's mesh and starting the calculation
-at a synoptic time is the simplest way to prepare a model initial condition."
-[@willoughby2007hurricane]
+such as impending TC landfall, may dictate observations at 6-h or even 3-h
+intervals. Interpolating rawinsonde observations to the model's mesh and
+starting the calculation at a synoptic time is the simplest way to prepare a
+model initial condition." [@willoughby2007hurricane]
 
 > "Because simple interpolation in space and time is a less-than-optimum way to
-use observations, nonsynoptic data require elaborate four-dimensional variational
-data assimilation schemes." [@willoughby2007hurricane]
-
-> "By 1941, charts for the 750, 500, and 250 mbar surfaces were being prepared daily
-for 0400, 1200, and 2000 GMT." [@grahame2000development]
-
-> "During the period considered here, NHC issued advisories 3 h after the standard
-synoptic time (0300, 0900, 1500, and 2100 UTC)." [@torn2012uncertainty]
-
-> "The operation data assimilation and forecast cycle interval is 6 h" [@aberson201010]
+use observations, nonsynoptic data require elaborate four-dimensional
+variational data assimilation schemes." [@willoughby2007hurricane]
+
+> "By 1941, charts for the 750, 500, and 250 mbar surfaces were being prepared
+daily for 0400, 1200, and 2000 GMT." [@grahame2000development]
+
+> "During the period considered here, NHC issued advisories 3 h after the
+standard synoptic time (0300, 0900, 1500, and 2100 UTC)." [@torn2012uncertainty]
+
+> "The operation data assimilation and forecast cycle interval is 6 h"
+[@aberson201010]
 
 > "Dan Petersen asked whether a data deficiency really exists with all the
-observational sources in the Atlantic region, or are the data not being
-properly used in the numerical models? ... Steve Lord emphasized the difficulty of the 
-task because many of these observations are not obtained at the synoptic times..."
-[@elsberry1992there]
-
-> "In the modern era, forecasters rely on purely dynamical models that integrate the 
-Navier-Stokes equations for atmospheric motions to represent both the storm itself and
-its surroundings. Numerical models are structured on a computational grid with 
-temperature, moisture, and wind tabulated in more-or-less rectangular cells." 
-[@willoughby2007hurricane]
+observational sources in the Atlantic region, or are the data not being properly
+used in the numerical models? ... Steve Lord emphasized the difficulty of the
+task because many of these observations are not obtained at the synoptic
+times..." [@elsberry1992there]
+
+> "In the modern era, forecasters rely on purely dynamical models that integrate
+the Navier-Stokes equations for atmospheric motions to represent both the storm
+itself and its surroundings. Numerical models are structured on a computational
+grid with temperature, moisture, and wind tabulated in more-or-less rectangular
+cells." [@willoughby2007hurricane]
 
 > "Predicting the weather required a widespread but highly coordinated team of
-observers and forecasters. Organization was key... They transmitted data from the 
-field using forms and protocols adapted from those that Abbe had developed while 
-forecasting the weather from Cincinnati, Ohio. The information arrived by telegraph
-coded to maximize accuracy and minimize the word count and changes. The reports piled
-up in a rush at the appointed hours." [@willis2006cleveland]
-
-> "Three times daily the duty forecaster broadcast a synopsis of prevailing weather
-conditions nationwide..." [@willis2006cleveland]
+observers and forecasters. Organization was key... They transmitted data from
+the field using forms and protocols adapted from those that Abbe had developed
+while forecasting the weather from Cincinnati, Ohio. The information arrived by
+telegraph coded to maximize accuracy and minimize the word count and changes.
+The reports piled up in a rush at the appointed hours." [@willis2006cleveland]
+
+> "Three times daily the duty forecaster broadcast a synopsis of prevailing
+weather conditions nationwide..." [@willis2006cleveland]
 
@@
 weather times are typically set based on the Universal Time Constant, with times
-recorded in Coordinated Universal Time (i.e.,
-Greenwich ...), and time stamps on weather maps and data based on this time zone
-often is given a "Z" (for "Zulu", an earlier name for this timezone) after the
-time.
+recorded in Coordinated Universal Time (i.e., Greenwich ...), and time stamps on
+weather maps and data based on this time zone often is given a "Z" (for "Zulu",
+an earlier name for this timezone) after the time.
 
@@
 
-> "Data such as tornado magnitude, path length, path width, and so on are recorded
-by the National Climatic Data Center (NCDC) as a public service and for use
-in meteorological, climatological, and engineering studies (Edwards et al. 2013)."
+> "Data such as tornado magnitude, path length, path width, and so on are
+recorded by the National Climatic Data Center (NCDC) as a public service and for
+use in meteorological, climatological, and engineering studies (Edwards et al.
+2013)." [@strader2015climatology]
+
+Sometimes used in conjunction with other NWS data products:
+
+> "Initially, tornado cases available in the NWS's Damage Assessment Toolkit
+(NOAA, 2014) were employed, which is a GIS-based framework for collecting,
+storing, and retrieving damage survey data (Camp et al., 2014). This toolkit
+provides a variety of tornado event filtering (tornado survey point, track,
+footprint, and/or swath) and download options [key mark-up language (.kml) or
+shapefile (.shp)], supplying an initial sample of georeferenced damage
+assessments. In this study, a tornado footprint is defined as the maximum areal
+extent of tornado intensity inferred by the damage, wind speed measured directly
+by mobile Doppler radar, or assessed theoretically as the recorded length
+multiplied with the maximum width as reported in Storm Data."
 [@strader2015climatology]
 
-Sometimes used in conjunction with other NWS data products:
-
-> "Initially, tornado cases available in the NWS's Damage Assessment Toolkit
-(NOAA, 2014) were employed, which is a GIS-based framework for collecting, storing, 
-and retrieving damage survey data (Camp et al., 2014). This toolkit provides a 
-variety of tornado event filtering (tornado survey point, track, footprint, and/or
-swath) and download options [key mark-up language (.kml) or shapefile (.shp)], supplying
-an initial sample of georeferenced damage assessments. In this study, a tornado 
-footprint is defined as the maximum areal extent of tornado intensity inferred by 
-the damage, wind speed measured directly by mobile Doppler radar, or assessed
-theoretically as the recorded length multiplied with the maximum width as reported in 
-Storm Data." [@strader2015climatology]
-
-> "To serve the public interest and the National Climatic Data Center (NCDC) storm 
-data record, the National Weather Service (NWS) documents the path length, width, 
-and maximum damage rating for every tornado county segment (NOAA 2007).[Footnote
-in manuscript: County segments of tornado paths are then combined at the NWS Storm
-Prediction Center to yield a unified one-tornado (ONETOR) dataset of whole-tornado 
-records (Schaefer and Edwards 1999).] Such data are used in meteorological and 
-climatological research, as well as in determining construction standard for 
-critical infrastructure such as high-tension electric lines and nuclear power plants
-(e.g., Ramsdell et al. 2007)." [@edwards2013tornado]
+> "To serve the public interest and the National Climatic Data Center (NCDC)
+storm data record, the National Weather Service (NWS) documents the path length,
+width, and maximum damage rating for every tornado county segment (NOAA
+2007).[Footnote in manuscript: County segments of tornado paths are then
+combined at the NWS Storm Prediction Center to yield a unified one-tornado
+(ONETOR) dataset of whole-tornado records (Schaefer and Edwards 1999).] Such
+data are used in meteorological and climatological research, as well as in
+determining construction standard for critical infrastructure such as
+high-tension electric lines and nuclear power plants (e.g., Ramsdell et al.
+2007)." [@edwards2013tornado]
 
@@
 full year of the issuance of Weather Bureau tornado watches, that the agency
-began to formally count all tornadoes. The *Climaticalogical Data National Summary*
-document evolved into *Storm Data* by 1959, which, in addition to tornado 
-data, included information on other storm perils, such as severe hail, high 
-winds, and floods (Grazulis 1993). Over the years, the tornado dataset 
-has been formatted and adjusted by the National Severe Storms Forecast 
-Center and Storm Prediction Center (SPC), and currently the National Tornado 
-Database is administered by the NWS Headquarters, SPC, and NCDC 
-(McCarthy 2003)." [@Ashley2007]
+began to formally count all tornadoes. The *Climaticalogical Data National
+Summary* document evolved into *Storm Data* by 1959, which, in addition to
+tornado data, included information on other storm perils, such as severe hail,
+high winds, and floods (Grazulis 1993). Over the years, the tornado dataset has
+been formatted and adjusted by the National Severe Storms Forecast Center and
+Storm Prediction Center (SPC), and currently the National Tornado Database is
+administered by the NWS Headquarters, SPC, and NCDC (McCarthy 2003)."
+[@Ashley2007]
 
@@
 We have revised this figure to show an x-axis truncated at the minimum observed
-value ($\sim75\%$), as requested. The original and revised versions of this paper 
-are reproduced in this response document as Figure \@ref(fig:fig3).
+value ($\sim75\%$), as requested. The original and revised versions of this
+paper are reproduced in this response document as Figure \@ref(fig:fig3).
 
@@
 
-For example, in Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2) (left) of this
-response document, we have reproduced the current version of Figure 5 as it
-would appear to someone with three common types of color vision deficiencies:
+For example, in Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2) (left)
+of this response document, we have reproduced the current version of Figure 5 as
+it would appear to someone with three common types of color vision deficiencies:
 deuteranomaly (defective green cone cells, the most common type), as well as
@@
 \@ref(fig:fig5check2) (right). We have included similar versions of other
-figures from the manuscript in response to later requests from EHP in this response.
+figures from the manuscript in response to later requests from EHP in this
+response.
 

### 2020-08-09 ehp_revision/tables/precip_high_corr.tex
Work some more on revision

@@
 \caption{\label{tab:highprecipcorr}Precipitation correlation during all versus high-precipitation events.
-                     The same sample of counties is shown as in Figure ... of the main text.
+                     The same sample of counties is shown as in Figure 2 of the main text.
                      Events are cases where a tropical cyclone came within ... km of each of the 
                      listed counties. The number of total events gives the sum of all points
-                     shown on the main plot for the county in Figure ... of the main text. 
-                     The Spearman correlation for all events is the same as that shown in Figure ...
+                     shown on the main plot for the county in Figure 2 of the main text. 
+                     The Spearman correlation for all events is the same as that shown in Figure 2
                      of the main text. High-precipitation events are those for which storm-associated
@@
 Miami-Dade, FL & 65 & 0.94 & 18 & 0.49\\
-Harris, FL & 38 & 0.93 & 10 & 0.84\\
+Harris, TX & 38 & 0.93 & 10 & 0.84\\
 Mobile, AL & 50 & 0.95 & 20 & 0.57\\

### 2020-08-09 ehp_revision/writing/ehp_discussion.tex
Work some more on revision

@@
 ground-based observations, but may sometimes underestimate very high rainfall
-values (Figure~\ref{fig:raincomparison}). When rainfall data is used to create
-binary exposure classifications, this disagreement is unlikely to influence
-results, as both data sources agree in identifying these as storms with high
-rainfall, but would be important to consider for cases that include rainfall as
-a continuous measurement.
+values (Figure~\ref{fig:raincomparison}), and in some counties the correlation
+was substantially lower when considering only tropical cyclones with cumulative
+local rainfall of 75 mm or more (Supplemental Table 1). When rainfall data is
+used to create binary exposure classifications, this disagreement is unlikely
+to influence results, as both data sources agree in identifying these as storms
+with high rainfall, but would be important to consider for cases that include
+rainfall as a continuous measurement.
 
@@
 
-We did find a small set of tropical cyclones for which for which
-agreement was high across several single-hazard exposure assessments (e.g.,
-Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto in~2006
+We did find a small set of tropical cyclones for which for which agreement was
+high across several single-hazard exposure assessments (e.g., Floyd in~1999,
+Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto in~2006
 (Figure~\ref{fig:jaccard})).  Hurricanes Floyd in~1999 and Irene in~2011 both
@@
 water may have helped maintain wind speeds in similar patterns to rain and
-distance-based exposures, resulting in more similarities across
-exposure assessments compared to other tropical cyclones.  For these storms, it
-may be possible to assess exposure to multiple hazards of the storm using a
-single metric, perhaps even a proxy like the distance between the county
-and the storm's track.  However, for these storms it may be difficult to
-untangle the contribution of each hazard to the overall effect of the storm,
-given that several hazards have similar geographical patterns.
+distance-based exposures, resulting in more similarities across exposure
+assessments compared to other tropical cyclones.  For these storms, it may be
+possible to assess exposure to multiple hazards of the storm using a single
+metric, perhaps even a proxy like the distance between the county and the
+storm's track. \textbf{With the dataset we describe in this paper, however,
+there is little need to limit analysis based on exposure to a single hazard or
+proxy, although multi-hazard studies of storms with high agreement among hazard
+exposures should look out for modeling issues from multicollinearity.} Further,
+for these storms it may be difficult to untangle the contribution of each
+hazard to the overall effect of the storm, given that several hazards have
+similar geographical patterns.
 

### 2020-08-10 ehp_revision/response_to_reviewers.Rmd
A few more edits

@@
 
-Data that is aggregated across a population---for example, the total number of
+Data that are aggregated across a population---for example, the total number of
 deaths in a geographic area in a certain time period---is known as *ecological

### 2020-08-15 ehp_revision/response_to_reviewers.Rmd
Work some on response

--- 2020-08-09 ehp_revision/response_to_reviewers.Rmd
@@
 **Response:**
+
+This is an excellent point. We have add text to the Discussion:
+
+The dataset we present here provides exposure assessments at the county level.
+This spatial scale allows for easy integration with health outcome data
+aggregated at the county level. Such aggregated or ecological data is often used
+for disaster epidemiology, as aggregated data may be easier to access than
+individual-level data, especially at a scale that covers many locations and a
+long time period and so allows higher statistical power and a broader range of
+exposure levels within the study [@wakefield2008overcoming].
+
+These ecological exposure estimates, however, set a common exposure level
+throughout the county, ignoring within-county variability, even though such
+variability often exists. For some hazards, this within-county variation in
+exposure level could be stark. For example, tornadoes cause very localized
+damage, directly along the tornado's path---a tornado can destroy homes on one
+side of a street while leaving those on the opposite side untouched. Levels of
+other hazards, like storm-associated winds and rain, will also vary within a
+county, but it will be unlikely that a county will have one area that is exposed
+to extremes of these hazards from a tropical cyclone while other parts of the
+county are completely unexposed, as both the wind fields and rain fields of
+tropical cyclones tend to be large in comparison to the size of a county.
+
+Aggregated data can be used to infer contextual level associations---for
+example, the association between county-level exposure to a storm hazard and
+county-wide rates of a health outcome. However, ecological data is also
+sometimes used to infer individual-level associations (e.g., the association
+between personal exposure to a storm hazard and personal risk of  experiencing
+the health outcome). Individual-level inference from ecological/aggregated data
+is prone to ecological bias [].
+
+-----------------------------------------------------------------------------
 

### 2020-08-15 ehp_revision/response_to_reviewers.Rmd
Add some refs
Work some on revisions

--- 2020-08-10 ehp_revision/response_to_reviewers.Rmd
@@
 **Response:**
+
+This is an excellent point. We have add text to the Discussion on the potential
+limitations of using aggregated, county-level exposure data in a research study:
+
+> **Text added to Discussion (added text in bold):**
+**"The dataset we present here provides exposure assessments at the county level.
+This spatial scale allows for easy integration with health outcome data
+aggregated at the county level. Such aggregated or ecological data is often used
+for disaster epidemiology, as aggregated data may be easier to access than
+individual-level data, especially at a scale that covers many locations and a
+long time period and so allows higher statistical power and a broader range of
+exposure levels within the study [@wakefield2008overcoming].**
+
+> "**These ecological exposure estimates, however, set a common exposure level
+throughout the county, ignoring within-county variability, even though such
+variability often exists. For some hazards, this within-county variation in
+exposure level could be stark. For example, tornadoes cause very localized
+damage, directly along the tornado's path---a tornado can destroy homes on one
+side of a street while leaving those on the opposite side untouched. Levels of
+other hazards, like storm-associated winds and rain, will also vary within a
+county, but typically with smoother variation. In particular, it will be
+unlikely that a county will have one area that is exposed to extremes of these
+hazards from a tropical cyclone while other parts of the county are completely
+unexposed, as both the wind fields and rain fields of tropical cyclones tend to
+be large in comparison to the size of a county.**
+
+> "**Aggregated data can be used to infer contextual level associations---for
+example, the association between county-level exposure to a storm hazard and
+county-wide rates of a health outcome. However, ecological data is also
+sometimes used to infer individual-level associations (e.g., the association
+between personal exposure to a storm hazard and personal risk of  experiencing
+the health outcome). Individual-level inference from ecological/aggregated data
+is susceptible to ecological bias [@greenland1994invited; @portnov2007ecological;
+@idrovo2011three]. Researchers who use the data provided here for ecological
+studies---with the aim of making individual-level inferences---should be aware
+of this potential and could explore approaches for minimizing risk of ecological
+bias (e.g., @wakefield2008overcoming).**"
+
+-----------------------------------------------------------------------------
 

### 2020-08-15 ehp_revision/writing/ehp_discussion.tex
Work some on revisions

@@
 
-Second, these data are aggregated to the county level. Many health outcome
-datasets are aggregated at this level, but some may be aggregated at a finer
-spatial resolution (e.g., census tract or ZIP code) or unaggregated (i.e.,
-point locations for each outcome). We have published the wind model used to
-create this dataset as its own open-source R package
-\parencite{stormwindmodel}, and it can be used to model storm-associated winds
-at a finer spatial resolution; however, measurements of other hazards cannot
-similarly be re-scaled through tools we provide.
+Second, these data are aggregated to the county level. \textbf{This spatial
+scale allows for easy integration with health outcome data aggregated at the
+county level. Such aggregated or ecological data is often used for disaster
+epidemiology, as aggregated data may be easier to access than individual-level
+data, especially at a scale that covers many locations and a long time period
+and so allows higher statistical power and a broader range of exposure levels
+within the study \parencite{wakefield2008overcoming}.}
+
+\textbf{These ecological exposure estimates, however, set a common exposure
+level throughout the county, ignoring within-county variability, even though
+such variability often exists. For some hazards, this within-county variation
+in exposure level could be stark. For example, tornadoes cause very localized
+damage, directly along the tornado's path---a tornado can destroy homes on one
+side of a street while leaving those on the opposite side untouched. Levels of
+other hazards, like storm-associated winds and rain, will also vary within a
+county, but typically with smoother variation. In particular, it will be
+unlikely that a county will have one area that is exposed to extremes of these
+hazards from a tropical cyclone while other parts of the county are completely
+unexposed, as both the wind fields and rain fields of tropical cyclones tend to
+be large in comparison to the size of a county.}
+
+\textbf{Aggregated data can be used to infer contextual level
+associations---for example, the association between county-level exposure to a
+storm hazard and county-wide rates of a health outcome. However, ecological
+data is also sometimes used to infer individual-level associations (e.g., the
+association between personal exposure to a storm hazard and personal risk of
+experiencing the health outcome). Individual-level inference from
+ecological/aggregated data is susceptible to ecological bias
+\parencite{greenland1994invited, portnov2007ecological, idrovo2011three}.
+Researchers who use the data provided here for ecological studies---with the
+aim of making individual-level inferences---should be aware of this potential
+and could explore approaches for minimizing risk of ecological bias (e.g.,
+\cite{wakefield2008overcoming}).}
+
+Further, while many health outcome datasets are aggregated by county, 
+some may be aggregated at a finer spatial resolution (e.g., census tract or ZIP
+code) or unaggregated (i.e., point locations for each outcome). We have
+published the wind model used to create this dataset as its own open-source R
+package \parencite{stormwindmodel}, and it can be used to model
+storm-associated winds at a finer spatial resolution; however, measurements of
+other hazards cannot similarly be re-scaled through tools we provide.
 

### 2020-08-16 ehp_revision/response_to_reviewers.Rmd
Work on cleaning up a response

@@
 
-This is an excellent point. We have add text to the Discussion on the potential
-limitations of using aggregated, county-level exposure data in a research study:
-
-> **Text added to Discussion (added text in bold):**
-**"The dataset we present here provides exposure assessments at the county level.
-This spatial scale allows for easy integration with health outcome data
-aggregated at the county level. Such aggregated or ecological data is often used
-for disaster epidemiology, as aggregated data may be easier to access than
-individual-level data, especially at a scale that covers many locations and a
-long time period and so allows higher statistical power and a broader range of
-exposure levels within the study [@wakefield2008overcoming].**
+This is an excellent point. We have added text in the Limitations section of the
+Discussion on the potential limitations of using aggregated, county-level
+exposure data in a research study:
+
+> **Text added to Discussion (added text in bold):** "Second, these data are
+aggregated to the county level. **This spatial scale allows for easy integration
+with health outcome data aggregated at the county level. Such aggregated or
+ecological data is often used for disaster epidemiology, as aggregated data may
+be easier to access than individual-level data, especially at a scale that
+covers many locations and a long time period and so allows higher statistical
+power and a broader range of exposure levels within the study
+[@wakefield2008overcoming].**
 
@@
 
------------------------------------------------------------------------------
-
-- Ecological fallacy vs individual fallacy, especially in the disaster context
-- Spatial misalignment / ecological effects of aggregating data at the county
-(or other geopolitical) level. 
-- Could be particularly problematic in some high-risk areas, because of the 
-"long and narrow" characteristics of island counties. Dare County, Key West.
-
-We think this is a very interesting question for disaster epidemiology, as
-while there remains potential for ecological bias, the potential pathways for
-health effects may also pass through damage in the individual's community beyond
-their residence. For example, damage to major roads in the community might make
-it harder for many residents to reach medical care in the aftermath of the
-storm, even if their home did not experience extensive destruction. Similarly,
-since power is provided over a network, conditions in an individual's community
-can be important in determining risk of long-term outages (does Seth have
-something on this?). As another example, if hospitals in the community are 
-over capacity or have to evacuate, this could increase health risk for people in 
-a fairly large "catchment" area for that hospital.
-
-Therefore, while we agree that individual-level data would in many cases be
-helpful for tropical cyclone epidemiology, the ideal might be to design studies
-that integrate both community-level and individual-level metrics of exposure.
-This could help in determining if there are risks introduced to individuals
-based on storm-related hazards in their broader community, after controlling for
-the hazards experienced directly at their homes.
-
-There is already some discussion on this topic in the Discussion [exposure 
-measurement error / misclassification].
-
-There are two interrelated mechanisms at play regarding this question. The
-first is the potential for ecological bias. This bias results from inferring
-individual-level associations from data that is available at an aggregated level. 
-In the case of epidemiological research, these data include health outcomes, 
-exposure, and covariates if the model is adjusting for confounding from them.
-For example, in tropical cyclone epidemiology, an ecological study might 
-compare county-wide rates of COPD hospitalizations to county-wide measures
-of tropical cyclone exposure while controlling for county-level smoking 
-rates and age distribution. The model in this case is fit at the county level, 
-and so the relevant inference is at that level. If this county-level estimate
-is inferred to represent the individual-level association between tropical 
-cyclone exposure and risk of COPD hospitalization, controlled for smoking and
-age, then the possibility of ecological bias arises. Given that this bias
-results from inferring from a model with data at one level of aggregation
-to another, this bias is also called cross-level bias.
-
-There are some subtleties that arise regarding this question for disaster
-epidemiology. First, it has been argued that, for some ambient exposures
-like weather and air pollution, contextual effects (that is, at the community
-level) are helpful to estimate in their own right, not just as a potential 
-surrogate for individual-level associations. This is because regulations
-(in the case of pollutants), interventions, and planning often happen at
-this level.
-
-Second, for a disaster, the relevant exposure might be not just at the
-individual level (e.g., winds or flooding at the individual's residence), but
-also throughout a broader area surrounding the individual. If storm damage
-closes roads in a county, for example, or damages the power grid, then these
-could create pathways for health risks for individuals throughout the county,
-whether damage was high at their residence or not. The causal pathways for
-tropical cyclones to affect human health differ from those for a dangerous
-substance, like air pollutants, in which the substance itself must enter the
-body to cause harm. While some health risk comes directly from the storm (e.g.,
-deaths and injuries from trees falling on homes or drowning from flooding),
-there are many more pathways that are indirect. These include pathways that go
-through the way that the storm's damage affects community infrastructure and
-access to medical care. Several papers have discussed the potential for 
-"individualistic fallacy", as a counterpoint to "ecological fallacy", when 
-such contextual-level exposure pathways are ignored.
-
-There are study designs that can be used to leverage ecological data while
-minimizing risk from ecological bias. For example, potential confounders like
-age distribution and smoking rates vary much less within a county over time than
-comparing between counties. The mechanism for ecological bias depends on the
-joint distribution between individual exposure, outcome, and covariates, if any
-are included in the model. Time series-style study designs, which compare a
-county to itself over time, allow for very similar covariate distributions
-between exposure and non-exposure. Other studies add to this design by
-stabilizing for temporal confounding through the addition of counties that were
-never exposed, allowing for a differences-in-differences style approach to
-calibrate for seasonal or longer-term trends that might otherwise create
-confounding. For example, many health outcomes have a strong seasonal trend,
-with peak rates in the winter and lows in the summer. Since the hurricane season
-stretches from summer into fall, a study design that compares the rate of a
-health outcome in an exposed county to the rate two weeks before the exposure
-might be biased away from the null, since baseline rates of the health outcome
-will typically be moving up over most of the hurricane season.
-
-Also, there may be confounders that are relevant at the contextual, rather than
-individual level, as well as modifiers. Example: whether the county is coastal
-could be a contextual-level confounder and effect modifier. This will influence
-whether the county is exposed to that storm or not, since storms usually weaken
-rapidly when the center is over land. In terms of counfounding pathways, coastal
-communities might tend to have lower levels of air pollution, because sea
-breezes clear the pollution regularly. They might also be wealthier on average,
-since property on or near the beach is desirable. Finally, they might be better
-prepared for or more hardened against tropical cyclones at the community-wide
-level (e.g., through hardier power infrastructure, more rigorous building codes,
-higher likelihood of evacuating in advance of a threatening storm) compared to
-nearby inland counties.
-
-Other approaches have also been suggested to reduce ecological bias while
-leveraging ecological data, which might be available at a much larger
-scale---and so provide more power---than is practical in collecting individual
-data. These include multi-level approaches in which individual-level data from a
-sample of individuals in each community is incorporated.
-
-The second issue is exposure measurement error/misclassification. When the
-exposure (e.g., binary exposure or continuous measure of intensity of exposure)
-is only measured at the county level, all the individuals in that county are
-assigned that same exposure. However, the hazards included in this study will
-vary across each county. For some hazards, this variation will be more stark.
-For example, tornadoes can be very local, causing extreme damage to a row of
-homes and none to the homes across the street. Some types of floods can also be
-very localized, rather than creating widespread flooding throughout the county.
-In both cases, when a county-level exposure assessment is made based on an event
-in the county in association with the storm, there may be many individuals in
-the county who did not experience that hazard at their residence. The other
-hazards we included---storm-associated winds and rainfall---will also have some
-variation across counties, but this will often not be as stark. The rainfields
-for tropical cyclones are very large, and while there are rainbands within the
-storm that might have particularly high rates of precipitation, these progess
-over the course of the storm, and it is unlikely that a county will have one
-area that experienced very extreme precipitation while another experienced very
-little. Winds might have more variation, especially in coastal counties and
-counties near the storm's central track, since winds decay quickly as you move
-from the storm's center, and storms typically weaken rapidly once they make
-landfall and can no longer draw on warm ocean waters for energy.
-
-This exposure data is provided at the county level, and this can lead to
-exposure measurement error / exposure misclassification in studies that infer to
-the individual level. In this case, there will be several processes at play.
-First, the hazards of the storm will not be uniform across the county. The
-degree of this within-county spatial heterogeneity will usually vary by hazard.
-The within-county heterogeneity is likely lowest for rain and for wind at lower
-levels. For rain, there can be particularly severe rain bands, but often a
-storm's rainfield will bring high precipitation to a large area. Wind can have a
-high heterogeneity close to the storm's center (storm winds decay most quickly
-with distance near the storm's center) and near landfall (winds decay rapidly
-after landfall). Other hazards can be very local within the county, only
-affecting a small proportion of the total county area. A tornado is a clear
-example---while it causes severe destruction in its track, areas just off the
-track can be unharmed, creating a dramatic spatial gradient in damage within a
-county. Floods can follow both. Some hurricanes have caused widespread flooding
-(e.g., Floyd, Katrina) while some have caused flooding that is more spatially
-focused (inland, with flashfloods? coast flooding just near the coast?). For
-tidewater/coastal areas, there is often a dense network of waterway. When rains
-occur further up the watershed, these all become flooded.
-
-Second is the question of whether exposure measured at the resident's location
-would be ideal. This is the case typically when studying an exposure that is a
-chemical substance. In this case, the pathway depends on the substance entering
-the body. For tropical cyclones, health risk could result from stress, problems
-accessing medical care, or repercussions of power outages (higher exposure to
-outdoor hazards like heat and air pollution, problems storing food and using
-some medical equipment). For these, high levels of wind and other hazards in the
-person's community may be as or more important than at their precise location.
-
-Measurement error can be either random or systematic. Systematic error can often
-be corrected with adjustment if the direction and typical size of the error is
-understood. Random error cannot in the same way. Either type of measurement
-error can be either differential (associated with the probability of the
-outcome) or non-differential (independent of the distribution / probability of
-the outcome). In simpler models, this characteristic might help in predicting
-whether the resulting bias is likely toward the null; however, more complex
-models (e.g., statistical models with adjustment for potential confounders) are
-trickier to diagnose in terms of the likely implications of differential versus
-non-differential measurement error [?].
-
-Not only can the exposure be measured with error, but so too can potential
-confounders and potential effect modifiers.
-
-We provide exposure assessments at the county level. This spatial scale allows
-for easy integration with health outcome data aggregated at the county level.
-However, there are also some things to consider when modeling associations
-at an ecological rather than individual level. The considerations are 
-particularly nuanced when investigating the health risks associated with a 
-disaster.
-
-Often in epidemiology, a study aims to estimate how an exposure changes the
-risk of a health outcome for an individual. The data might be collected for
-lots of people, but the key question is this---if a person is exposed, how 
-does his or her risk of the outcome change. If you are exposed to this hazard, 
-how much more likely are you to have the outcome? To get sick, or injured, or
-to die? This inference is at the individual level.
-
-Even when individual-level inference is the goal, sometimes data is only 
-available at an aggregated level. For example, privacy concerns might
-mean that health outcomes are only made available as a county-wide daily 
-count. You can fit a mode with this aggregated data, which will estimate the
-association between exposure and health risk at the ecological level. This
-contextual effect estimate is sometimes used as an estimate of individual-level
-association, but doing so can introduce ecological/multi-level bias. The 
-individual-level effect estimate can be very biased from the true association, 
-even to the point of reversing the effect estimate---estimating a protective
-effect when the true effect is detrimental, for example.
-
-This ecological fallacy plays out in two ways. First, if you use a single 
-exposure measurement for everyone in an area, some people will be misclassified
-(or have exposure measured with error, for a continuous metric) unless the 
-exposure is perfectly homogeneous across the area. This will typically not
-be the case. This exposure misclassification can bias estimates of the 
-association between exposure and outcome in the same way exposure 
-misclassification through any other mechanism would.
-
-Next, ecological bias could result from confounding, even if the confounders
-are controlled in the ecological-level model. When you aggregate data, you 
-lose information about how both the exposure and potential confounders vary
-across individuals within the area of aggregation. As a result, a factor
-could still confound the inference of an individual-level association, even 
-if it is controlled for at a population level in an ecological model. For 
-example, a study of the association between risk of pre-term birth and 
-tropical cyclone exposure could control for county-level smoking when 
-modeling county-level storm exposure and county-level rates of pre-term births.
-Even with this control, an observed association could result from differences
-in individual smoking status, if there is within-county variation in 
-smoking and if this has a different pattern across people in the county than
-variation in exposure from the county-wide exposure estimate.
-
-For disasters, there are added nuances. First, for exposure it may not just 
-matter what the level of exposure is in a person's immediate scope. Disasters
-bring physical hazards that can harm people directly, but also through indirect
-pathways. For example, a tropical cyclone can bring high winds that cause 
-power outages, and as a result those affected could be exposed to more outdoor
-hazards (outdoor air pollution, heat), struggle to safely store perishable
-food and medications, and lose means to power medical equipment. While extreme
-winds at a person's residence would increase their risk of a power outage, 
-outages could also be caused by damage to the grid in another part of the 
-community. In some cases, then, the level of exposure in a person's community
-may be as important in opening a pathway of risk as exposure at the person's
-immediate location.
-
-What's more, if the disaster has a large health impact, the health outcome of
-one person in the community could affect the risk of the outcome (or other
-adverse outcomes) for others. This situation is often only the case for 
-infectious diseases, where one person with the disease can spread it to others.
-However, if the community-wide impact is large enough, it can affect access
-to and effectiveness of medical care for everyone in the community. This
-effect has been seen recently with Covid 19---attempts to "flatten the curve"
-aim to avoid moving into a state where a community's health system becomes
-overwhelmed and can no longer deliver a typical level of care to those in the
-community. This effect could happen with either infectious or non-infectious
-diseases.
-
-When a proxy exposure estimates is used for many people covered by a study, this
-can result in a type of exposure measurement error called *Berkson error*. In
-this case, the true exposure of each individual is randomly distributed around
-this proxy/aggregate-level exposure estimate, with a mean exposure level among
-the subjects that is equal to the proxy/aggregate-level measure. In other words,
-the group as a whole is assigned a common exposure level, when in fact this is
-the average exposure level across members of that group, but the individuals'
-true exposure levels are randomly distributed around this mean group level to
-which they are all assigned for modeling.
-
-Berkson:
-
-$$
-T = X + E
-$$
-where:
-
-- $T$: true individual exposure
-- $X$: proxy exposure, assigned to everyone in the group
-- $E$: error, with mean of 0, independent of $X$
-
-The dataset provides exposure by group/area (county), not by individual. These
-ecological exposure estimates require us to assume that the exposure level is
-the same for all individuals in that county. It ignores/smooths over
-within-county variability in exposure.
-
-For wind and rain, we're essentially basing exposure assessment on modeling of
-exposure. For rain, this is a based on a re-analysis model that
-incorporates/integrates observed data from the area and time, using a model to
-integrate data from diverse sources. For wind, the exposure estimates are based
-on a model, with a core based on the Willoughby wind model. For flooding and
-tornadoes, the exposure is based on records of events anywhere in the county,
-and in many cases may be based, essentially, on exposure assessment, as in many
-cases someone like someone at a NOAA [?] center have determined whether the
-event should be reported in the database.
-
-Even exposure at each person's residence would not suffice, since people move 
-around during the day (although perhaps less for most people during a severe
-storm).
-
-Data that are aggregated across a population---for example, the total number of
-deaths in a geographic area in a certain time period---is known as *ecological
-data*, and studies that base inference on this type of data are known as
-*ecological studies*. Terms include *ecological*, *aggregate*, *contextual
-level*.
-
-> "In an ecological study the unit of analysis is a group of people rather than 
-the individual." [@sedgwick2014ecological]
-
-> "In his seminal paper, Robinson (1950) distinguished between two types of
-correlation---ecological and individual. The former is obtained for a group of
-people, while the latter is estimated for indivisible units, such as
-individuals." [@portnov2007ecological]
-
-Here is one challenge when working with aggregated data. If you use aggregated
-data to infer individual-level associations, this can lead to *ecological fallacy*. 
-This results from cross-level inference---the data used to model the association
-is at the contextual level (e.g., county-level), while the inference is for the
-individual association between exposure and outcome.
-
-One definition of the ecological fallacy:
-
-> "This spurious result provides an example of the ecological fallacy, in 
-which conclusions ... based on ecologic data are opposite of those drawn 
-on the basis of individual-level data." [@wakefield2008overcoming]
-
-> "It is a bias produced when analyses realised in an ecological (group level)
-analysis are used to make inferences at the individual level."
-[@delgado2004bias]
-
-> "Results from ecological studies are prone to the ecological fallacy. The 
-ecological fallacy is a term used when collected data are analysed at a 
-group level and the results are assumed to apply to associations at the 
-individual level." [@sedgwick2014ecological]
-
-Importantly, aggregated data can be used both to infer a contextual effect
-(e.g., the association between county-wide rates of a health outcome and 
-exposure) as well as an individual-level effect (e.g., how a person's
-risk of the outcome is associated with that person's exposure to the hazard). 
-The first inference will *not* be prone to ecological bias, as the inference
-is at the same level as the data. The second inference (individual-level) 
-will be susceptible to bias when aggregated data is used to fit the model. 
-In this case, the association estimated based on aggregated data can be 
-different, and even reversed, compared to the true individual-level effect.
-
-> "I propose three criteria for the identification of ecological fallacy; all
-three of these should be present to confirm its existence: (1) Results must be
-obtrained with ecological (population) data; (2) Data must be inferred to
-individuals. One use of ecological studies is to explore individual-level
-association when individual data are not available. When the focus of the study
-was contextual or based on population effects and there is no inference to
-individuals, ecological fallacy is not possible. When only the first two
-criteria are present---which is insufficient to affirm ecological fallacy---it
-is appropriate to acknowledge that there is a possible relationship and that
-further study is required; (3) Results obtained with individual data are
-contradictory." [@idrovo2011three]
-
-> "One important breakdown in the analogy between ecologic-bias and
-individual-level confounding occurs because in etiologic research, the target of
-inference for both ecologic and individual-level studies is the same: Both study
-effects at the individual level. For an individual-level study, these target
-effects are at the same level as the units of analysis. But, for an ecologic
-study, these target effects are at a finer level than the units of analysis. As
-a result, an ecologic study can be unbiased for ecologic effects (in particular,
-ecologic confounding may be absent) and yet still be biased for individual-level
-effects. Many of the classic social science examples of ecologic bias are of
-this form." [@greenland1994invited]
-
-> "We refer to an incorrect extrapolation from unbiased estimates of ecologic
-effects to unobserved individual-level effects as 'cross-level bias' (although
-the latter term is sometimes used to refer to any ecologic bias). As the
-examples suggest, it is possible to detect cross-level bias if one can identify
-and observe homogeneously exposed regions." [@greenland1994invited]
-
-> "assuming that associations observed at the level of the area hold for the 
-individuals within the areas can lead to the so-called ecological fallacy 
-(Selvin, 1958)." [@wakefield2006health]
-
-> "A broad definition of ecological fallacy is that certain data (e.g.,
-illiteracy rates by foreign-born) are unavailable, leading to the use of proxies
-and erroneous estimates. However, we advocate a more narrow definition of this
-phenomenon, according to which 'ecological fallacy' refers solely to differences
-in conclusions which may be drawn from group-level data (e.g., average
-illiteracy rates by foreign-born), as opposed to data obtained for individuals."
-[@portnov2007ecological]
-
-> "If different conclusions are drawn from the analysis of data upon their
-aggregations into units of different sizes (e.g., from individuals to townships
-and regions), these differences are commonly referred to as 'ecological
-fallacy'. In epidemiology, these differences are also known as ecological or
-cross-level bias [refs]. The ecological bias is difficult to control and may
-lead, under certain circumstances, to spurious effects [refs]. If
-epidemiological study does not consider the possibility of ecological bias,
-misguided policy decisions may follow [refs]." [@portnov2007ecological]
-
-Ecological bias can *only* happen if there is, indeed, an individual-level
-effect:
-
-> "If exposure has no effect on any individual, there will be
-no individual-level or ecologic effects, and so cross-level bias cannot occur. 
-Thus, cross-level bias will not affect the validity of an ecologic test of the 
-null hypothesis, although it still must be considered in interpreting a 
-significant effect." [@greenland1994invited]
-
-In some cases, the ecologic-level effect (contextual effect) will be directly 
-of interest, and so we may be interested in inferring at that level (in which 
-case use of ecologic-level data is appropriate and will not be prone to 
-ecological bias):
-
-> "The bias phenomenon in example 7 may be ascribed to the fact that an 
-ecologic (regional) variable (percent Protestant) had effects on individual 
-risk, in addition to effects of the corresponding individual-level variable
-(religion) that the ecological variable summarized. In both social and 
-infectious disease epidemiology, as well as in community intervention studies, 
-ecologic effects may be of direct interest; a classic example is the 
-phenomenon of 'herd immunity', in which the overall prevalence of 
-immunity in a region, as well as individual immune status, determines the 
-risk of individuals within the region (16). For noncontagious diseases, 
-however, ecologic effects may not be of direct interest, and may in fact
-obscure the individual-level effect of interest." [@greenland1994invited]
-
-In some cases, the estimated association based on ecological data will be very 
-different from the association that exists at the individual level:
-
-> "We see that the ecological relative risk associated with low birth 
-weight is completely incomparable with the individual-level coefficient."
-[@wakefield2008overcoming]
-
-Ecological bias can also complicate estimation of effect modification, which 
-otherwise could help in identifying vulnerabilities and susceptibilities among
-certain subpopulations:
-
-> "Furthermore, non-White race now appears protective, rather than detrimental
-as the individual-level analysis suggests." [@wakefield2008overcoming]
-
-Ecologic bias crops up because the aggregated data fails to capture variation
-within each aggregation level (county, for example) for both the exposure 
-and potential confounders. For example, during a tropical cyclone, different
-parts of a county will experience different peak wind speeds. If the county 
-is coastal, the most severe winds will likely be near the coast, as the 
-storm will weaken once it makes landfall and moves inland, as the storm 
-generates its power from the water. When we assign the entire county a
-single estimate of intensity, it smooths over the fact that locations in 
-some parts of the county were subject to higher peak winds while locations
-in other parts of the county had lower winds. At the least, this smoothing
-of variation in exposure levels across the county could lower the power of 
-the study to detect a clear association between exposure and health risk, as
-this smoothing drops information inherent in the variation within the county
-in exposure levels.
-
-This, however, would not result in ecological fallacy unless another factor
-comes into play---variation within each aggregation level in confounders. 
-Just as aggregation smooths over within-area variation in exposure levels, it
-also smooths over within-area variation in levels of potential confounders. 
-Depending on the patterns of this within-area variation, a result could be
-that ecological-level control of the confounders does not, in fact, control
-for their role at the individual level, and so the association inferred at
-the ecologic level continues to be confounded by them when inferred to the
-individual level.
-
-Essentially, this comes down to a question of the joint distribution, at the
-individual level, of exposure, outcome, and covariates [?].
-
-> "As various discussants have pointed out (3--6, 8, 9), ecologic relative-risk
-estimates can be subject to biases not present in estimates from individual-level
-observational studies of the same populations (case-control and cohort studies).
-Unlike an individual-level study, an ecologic study does not link individual
-outcome events to individual exposure or covariate histories, nor does it
-link individual exposure and covariate histories to one another. It is these
-linkage failures that are the source of the special biases of ecologic studies
-(1, 2, 6)." [@greenland1994invited]
-
-> "In the extreme, ecologic bias in comparisons limited to homogeneously exposed
-regions (as in examples 3--6) can be viewed as purely an issue of confounding,
-albeit with special complexities of measurement and control of confounders."
-[@greenland1994invited]
-
-> "More generally, one may show that cross-level bias will not occur if the 
-individual-level effects of all variables (including unmeasured background
-factors) follow a multiple linear-regression model with no regional effects and
-no interactions (17). Nevertheless, given the usual inappropriateness of the 
-multiple-linear model and the absence of homogeneously exposed regions, the 
-possibility of cross-level bias adds a dimension to ecologic bias beyond
-that of simple confounding." [@greenland1994invited]
-
-> "Covariate control in ecologic studies requires attention to details not 
-ordinarily of concern in individual-level studies. When, as is usually the
-case, important nonlinearity or nonadditivity can be expected among 
-exposure and covariate effects, it may be necessary to obtain and 
-control for multiple summaries of joint covariate distributions in order
-to insure that control is adequate." [@greenland1994invited]
-
-> "Ecological fallacy can be produced by within group (individual level)
-biases, such as confounding, selection bias, or misclassification, and
-by confounding by group or effect modification by group." [@delgado2004bias]
-
-> "**Confounding by group**: it is produced in an ecological study, when 
-the exposure prevalence of each community (group) is correlated with 
-the disease risk in non-exposed of the same community. It can be 
-a mechanism for producing ecological fallacy." [@delgado2004bias]
-
-If individual-level inference is the aim, and population-level data is
-available, there are some methods for using it while still aiming to avoid
-ecological bias. Indeed, it can be helpful to use population-level data, as it
-is often available for a large population, improving the power and precision of
-the study. Further, the level of exposure might vary a lot more over the
-population captured with population-level data compared to the variation that
-captured in a smaller sample of individual-level data. This can contribute both 
-to statistical power and improve external validity (as the study data will cover
-more of the range of exposure that might ever be expected). However, if
-individual-level inference is the goal, then there are ways to supplement
-population-level data with samples of individual-level data through two-level, 
-semi-ecologic study designs.
-
-> "It is generally well recognized that in order for ecologic data to provide
-reliable inferences, they need to be supplemented with individual-level data."
-[@wakefield2008overcoming]
-
-> "For this example, we have access to complete individual-level data,
-permitting a 'gold standard' individual-level analysis."
-[@wakefield2008overcoming]
-
-> "The only reliable way to characterize within-area variation in exposures
-and confounders, and hence control ecologic bias, is to collect and incorporate
-individual-level data. To help epidemiologists achieve this goal, in this paper
-we describe the use of the two-phase design in an ecologic setting."
-[@wakefield2008overcoming]
-
-> "The only reliable way to characterize within-area variation in exposures
-and confounders, and hence control ecologic bias, is to collect and incorporate
-individual-level data. To help epidemiologists achieve this goal, in this 
-paper we describe the use of the two-phase design in an ecologic setting."
-[@wakefield2008overcoming]
-
-> "In a semi-ecologic study, an ecologic exposure is combined with
-individual-level outcomes and confounders. A two-phase approach is particularly
-useful for such a study, with the phase 2 data corresponding to stratified
-sampling of individual exposures." [@wakefield2008overcoming]
-
-> "To implement a two-phase design, an initial phase 1 cross-classification by
-the binary disease outcome and stratification variables is required; in phase 2,
-samples of individuals are drawn from each of the cross-classification cells,
-with data on additional variables being drawn from the subsamples of
-individuals. Intuitively, the stratified sampling is focused on informative
-cells, and estimation methods use both phases of data for efficiency and to
-acknowledge the outcome-dependent sampling. In the simplest ecologic setting,
-the cross-classification is by outcome and area only, and if area is a surrogate
-for important risk factors this design will be efficient. We are particularly
-interested in situations where an initial classification is available by
-outcome, area, and confounders such as age and gender---this is the case in a
-semi-ecologic study. Phase 2 may then provide detailed exposure information on a
-subset of the phase 1 individuals." [@wakefield2008overcoming]
-
-> "Two-phase study designs are a generalization of matched case-control designs
-in which, initially, the entire sample population is cross-classified according
-to case/control status and some stratification variable, S. The latter depends
-on covariates observed in all individuals and may include exposures of interest, 
-proxy exposure measures, or potential confounders. In settings like those we
-are considering, such as environmental epidemiology, S may also depend on geographic
-area, which can act as a surrogate for the totality of confounders associated
-with each area, as well as provide a well-defined sampling frame for the controls."
-[@wakefield2008overcoming]
-
-> "The results of our simulation study point to the benefits associated with
-combining the two sources of data, in terms of both bias and efficiency. Rather
-than supplement an ecologic study with individual-level data, it may be of
-interest to combine existing individual-level data with external group-level
-data. Strategies that combine both types of data have been shown to alleviate
-participation bias and improve efficiency in case-control studies with missing
-data (31)." [@wakefield2008overcoming]
-
-> "As with other observational studies, ecologic studies can give useful results
-if biases such as those discussed here can be ruled out or quantified.
-Nevertheless, bias evaluation can be especially difficult in ecologic studies of
-geographic regions because of the many potentially interacting covariates that
-may differ across regions. When biases cannot be ruled out with available data,
-further exploration will require individual-level studies."
-[@greenland1994invited]
-
-Why we might sometimes use (and want to use) population-level data:
-
-> "Because of the unavailability of individual-level data, ecological data may
-be resorted to and may come in a variety of forms." [@wakefield2008overcoming]
-
-> "Epidemiologists continue to use ecologic and aggregate data. Despite their 
-known drawbacks, these data, often aggregated across geographic areas, offer the
-advantages of widespread availability and gains in statistical power from large
-populations and increased exposure ranges." [@wakefield2008overcoming]
-
-> "Population and health data are often routinely available in ecological, that
-is group, form while the exposure data typically consist of a set of values
-recorded at monitor sites, or via one-off sampling." [@wakefield2006health]
-
-It can be important to think about the scale that the process happens at. For
-something very local, you might lose a lot more with aggregated data.
-
-> "Data availability and exposure variability often determine the scale of
-examination and the suitability of a study. Exposures arising from a point or
-line source offer exposure contrasts on small scales, requiring small-area data;
-in constrast, dietary variables show little variation across small scales, and
-consequently international studies are used." [@wakefield2008overcoming]
-
-There can also be an *individualistic fallacy*, if emphasis is limited to the
-association between individual-level factors, without considering the potential
-role of contextual factors in an individual's risk of a health outcome. In other
-words, cross-level bias can go in both directions [@idrovo2011three]. While the
-fallacy of inferring individual-level associations from ecologic-level data is
-called *ecological fallacy*, the fallacy of inferring ecologic-level
-(contextual) effects from individual-level data is known as the *atomistic
-fallacy* [@idrovo2011three]. There are also fallacies that can come in, it
-seems, from inferring at the same level as the data but failing to consider
-influences from factors at the other level. When inferring individual-level
-associations from individual-level data, without considering an additional role
-of ecological-level factors, this is known as the *psychologistic* or
-*individualistic fallacy* [@idrovo2011three]. When inferring contextual effects
-from ecologic data without considering individual-level factors, this is known
-as the *sociologistic fallacy* [@idrovo2011three].
-
-> "Although ecological studies are important to epidemiology (especially in
-environmental and social epidemiology), public health practitioners seem afraid
-of ecological studies. It is a common practice to assume the presence of
-ecological fallacy (Robinson 1950) and low-level validity when analyzing an
-ecological study. Most epidemiologists prefer an exclusive individualistic
-approach, although the importance of a multilevel causal approach is widely
-recognized (Diez-Roux 2002). In this sense, some authors suggest that it is as
-important to recognize the presence of ecological fallacy as to recognize
-psychologistic or individualistic fallacy (Subramanian et al. 2009)."
-[@idrovo2011three]
-
-Confounding in general:
-
-> "A confounder ... is associated with an exposure or risk factor for the
-outcome and with the outcome independent of the exposure, but is not on the
-causal pathway between the exposure and the outcome. Confounding occurs when the
-relationship between an exposure variable and the outcome variable is
-contaminated, so that the measure of association between these two variables is
-actually also capturing the effects of a third variable, the confounding
-variable. For example, many factors are associated with income (exposure) and
-health (outcome), such as education, employment or wealth? It may be that these
-factors explain part of the observed assocation." [@gunasekara2008glossary]
-
-Measurement error in general:
-
-> "Measurement error: The different between an observed variable and the
-variable that belongs in a multiple regression equation. Measurement error is
-the random or systematic error arising during data collection of variables. The
-measured variable x is measured with error eta, which is the distance from x to
-the 'true' value of x. Measurement error can produce bias such as attenuating
-the estimators of exposure variables, but may also have more complex effects.
-Longitudinal data analyses such as fixed effects models can actually augment the
-problem of measurement error in mis-measured explanatory variables that change
-little over time." [@gunasekara2008glossary]
-
-Study validity:
-
-> "Study validity refers to the degree to which the inferences drawn from a
-study are warranted when account is taken of the study methods; the
-representativeness of the study sample; and the nature of the population from
-which it is drawn.[5] There are two types of study validity. *Internal validity*
-is the degree to which the results of a study are correct for the sample of
-people being studied. *External validity* (generalisability) is the degree to
-which the study results hold true for a population beyond the subjects in the
-study or other settings.[20]"
-
 \begin{shaded}
@@
 
-- Differs by exposure metric. Easier for some (wind, rain) than for others
-(flood, tornado)
-- For rain, we use a re-analysis product only available for the continental US
-(right?). However, there are similar gridded re-analysis products available for
-other regions worldwide that could be used in a similar way.
-- For tornadoes, there may be national databases in other spots similar to the
-tornado database that we use, but as far as we're aware, no international one.
-- Same for floods. 
-- For wind, the wind model could be generalizable with some tweaks. These
-include:
-  + More landmasks
-  + Direction of storm movement for Northern vs Southern Hemispheres
-  + Different wind averaging periods for the recorded best tracks data in
-  different storm basins.
-  + Our wind model was derived with US data and so may need bias-correction or a
-  shift to a West Pacific--based model
-
-> "Tornadoes long have been recognized as a global phenomenon (e.g., Wegener
-1917; Feuerstein et al. 2005), having been recorded in every continent except
-Antarctica. As such, numerous nationas share an interest in improving assessment
-of their damage. International research and involvement in tornado survey work
-and damage and intensity scales is well underway." [@edwards2013tornado]
-
-For flood and tornado events, we used a database of severe weather events that
-is specific to the United States (the US NOAA's Storm Events database). To
-expand our exposure data for these hazards internationally, we would need to
-find similar databases for other affected countries. [international flood and
-tornado databases? EM-DAT? Others?]
-
-The wind exposure data is based on modeling wind fields based on records of the
-storm's track and central intensity throughout the tracking period. These values
-are tracked for tropical cyclones around the world, and so it would be possible
-to expand the dataset to encompass other countries for wind exposure
-measurements. However, there are some practical issues that would need to be
-addressed to do so with the approach we've taken.
-
-The software we are using to model wind [@stormwindmodel] has several features
-that make it specific to Atlantic basin storms. First, some steps in the model
-depend on whether the storm's center or the location being modeled is over land
-or water. These steps use a landmask---a grid with a binary indicator of land or
-sea by latitude and longitude. Currently, the software only provides a landmask
-that covers the United States, the Gulf of Mexico, and areas of the Atlantic
-Ocean near the US. Second, the model assumes that cyclonic winds are
-counterclockwise, and so is valid only for the Northern Hemisphere. Third, the
-model assumes that the central intensity of the storm is based on [wind
-averaging period], which is the standard used in the tracking data for
-Atlantic-basin storms, but is not a common standard around the world. Instead,
-some basins record central intensity measurements using averaging periods of
-....
-
-Therefore, it would be relatively straightforward to expand the wind exposure
-data to cover other countries affected by Atlantic-basin storms, although this
-would require some expansion of the area covered by our current landmask. More
-modifications would need to be made to the wind model software before it could
-be extended to storms in other basins, particularly basins that use different
-wind averaging periods for their main tracking data and basins in the Southern
-Hemisphere. Alternatively, other wind modeling software could be used for this
-exposure assessment, as other groups have developed software that is more
-extensible to multiple storm basins [Australian group's software].
-
-The rain exposure data would also be fairly easy to extend to cover other
-countries, although again there would be a few practical limitations to extend
-the approach we've taken in developing the dataset covered in this manuscript.
-For this exposure, we've used a re-analysis dataset. We've taken some steps to
-translate this gridded data to the county scale, so it can be integrated with
-the scale for many human impacts measurements. However, these steps have been
-well documents [ref] and would be easy to reproduce. The re-analysis dataset
-that we use is specific to the contiguous US [ref on NLDAS], but other similar
-re-analysis datasets exist for other locations [example of one with global
-coverage?].
-
-For all of these, we have provided exposure assessments at the county scale. Any
-of the steps we have taken could be done for other countries at a similar
-geopolitical scale, although the name of that areal unit will vary by country
-(e.g., municipalities in Mexico, districts in India). For wind and rain
-exposures, the approach we take here could be used with any areal unit, as long
-as there are geographical shape files available for the area boundaries
-(although the modeling for wind may be to the area's geographic mean center
-rather than population mean center).
-
-> "We initially developed software to model exclusively Atlantic-basin storms
-[@stormwindmodel]. There are two characteristics of Northwestern Pacific Ocean
-basin storm data that prevent direct use of the current version of the software.
-First, the software currently requires Western Hemisphere longitude inputs.
-Second, the ``Best Tracks'' data from the China Meteorological Administration
-include central tropical cyclone intensity using a 2-minute averaging period,
-while the software currently assumes that central intensities are given for a
-1-minute averaging period, as is the case for storm track data from the US
-National Hurricane Center. With the assistance of Ms. Schumacher, we will update
-our tropical cyclone wind model software to appropriately handle 'Best Tracks'
-data from the China Meteorological Administration, incorporating conversion
-factors [@harper2010guidelines] to convert central storm maximum sustained wind
-measurements to a common averaging period prior to applying further steps of the
-wind model."
+While there would be some limitations in comparisons based on differences in the
+scale of county-equivalents, many countries affected by tropical cyclones do
+have geopolitical areas that are fairly comparable to US counties, like
+municipalities in Mexico, districts in India. There would be a few challenges,
+however, in developing the exposure dataset for the hazards that we cover in the
+US in the data presented in this paper. We have added text to the paper's
+Discussion on opportunities and barriers to expanding this approach
+internationally:
+
+> Text added to the Discussion (added text in bold): 
+"**This dataset is limited to the contiguous US. Some exposure estimates would
+be fairly straightforward to extend internationally. The precipitation data is
+extracted and aggregated to the county level from a re-analysis data product.
+While the re-analysis product used here (NLDAS-2) only covers the contiguous US,
+other similar re-analysis products, as well as other types of precipitation
+datasets, have global coverage [@sun2018review]. The wind data are generated
+based on a model that is currently US-focused [@stormwindmodel], but could be
+extended to other areas, although this would require adding new land/sea masks
+within the associated software, as well as accounting for differences across
+storm basins in wind averaging periods [@harper2010guidelines] and in the
+direction of cyclonic winds in the Northern versus Southern Hemisphere. Further,
+since the core of the wind model was developed based on data from Atlantic-basin
+storms [@willoughby2006parametric], an extension to other areas should include
+separate validation and calibration to ensure it performs appropriately in those
+settings. Other wind field modeling software is also available that already
+provides a global coverage, like Geoscience Australia's Tropical Cyclone Risk
+Model (http://geoscienceaustralia.github.io/tcrm/). For tornado and flood
+events, the data described in this paper drew on a US-focused storm events
+database, and so international extension of data on these hazards would require
+access to similar databases covering other countries. Finally, the relevant
+geopolitical boundaries to use for aggregation would vary by country (e.g.,
+municipalities in Mexico, districts in India).**"
 

### 2020-08-16 ehp_revision/writing/ehp_discussion.tex
Work on cleaning up a response

@@
 act synergistically in causing impacts~\parencite{smith2009}.
+
+\textbf{This dataset is limited to the contiguous US. The precipitation data is
+extracted and aggregated to the county level from a re-analysis data product.
+While the re-analysis product used here (NLDAS-2) only covers the contiguous
+US, other similar re-analysis products, as well as other types of precipitation
+datasets, have global coverage \parencite{sun2018review}. The wind data are
+generated based on a model that is currently US-focused
+\parencite{stormwindmodel}, but could be extended to other areas, although this
+would require adding new land/sea masks within the associated software, as well
+as accounting for differences across storm basins in wind averaging periods
+\parencite{harper2010guidelines} and in the direction of cyclonic winds in the
+Northern versus Southern Hemisphere. Further, since the core of the wind model
+was developed based on data from Atlantic-basin storms
+\parencite{willoughby2006parametric}, an extension to other areas should
+include separate validation and calibration to ensure it performs appropriately
+in those settings. Other wind field modeling software is also available that
+already provides a global coverage, like Geoscience Australia's Tropical
+Cyclone Risk Model (http://geoscienceaustralia.github.io/tcrm/). For tornado
+and flood events, the data described in this paper drew on a US-focused storm
+events database, and so international extension of data on these hazards would
+require access to similar databases covering other countries. Finally, the
+relevant geopolitical boundaries to use for aggregation would vary by country
+(e.g., municipalities in Mexico, districts in India).}
 

### 2020-08-16 ehp_revision/writing/ehp_response_darlings.tex
Work on cleaning up a response

# R1 Comment 1

- Ecological fallacy vs individual fallacy, especially in the disaster context
- Spatial misalignment / ecological effects of aggregating data at the county
(or other geopolitical) level. 
- Could be particularly problematic in some high-risk areas, because of the 
"long and narrow" characteristics of island counties. Dare County, Key West.

We think this is a very interesting question for disaster epidemiology, as
while there remains potential for ecological bias, the potential pathways for
health effects may also pass through damage in the individual's community beyond
their residence. For example, damage to major roads in the community might make
it harder for many residents to reach medical care in the aftermath of the
storm, even if their home did not experience extensive destruction. Similarly,
since power is provided over a network, conditions in an individual's community
can be important in determining risk of long-term outages (does Seth have
something on this?). As another example, if hospitals in the community are 
over capacity or have to evacuate, this could increase health risk for people in 
a fairly large "catchment" area for that hospital.

Therefore, while we agree that individual-level data would in many cases be
helpful for tropical cyclone epidemiology, the ideal might be to design studies
that integrate both community-level and individual-level metrics of exposure.
This could help in determining if there are risks introduced to individuals
based on storm-related hazards in their broader community, after controlling for
the hazards experienced directly at their homes.

There is already some discussion on this topic in the Discussion [exposure 
measurement error / misclassification].

There are two interrelated mechanisms at play regarding this question. The
first is the potential for ecological bias. This bias results from inferring
individual-level associations from data that is available at an aggregated level. 
In the case of epidemiological research, these data include health outcomes, 
exposure, and covariates if the model is adjusting for confounding from them.
For example, in tropical cyclone epidemiology, an ecological study might 
compare county-wide rates of COPD hospitalizations to county-wide measures
of tropical cyclone exposure while controlling for county-level smoking 
rates and age distribution. The model in this case is fit at the county level, 
and so the relevant inference is at that level. If this county-level estimate
is inferred to represent the individual-level association between tropical 
cyclone exposure and risk of COPD hospitalization, controlled for smoking and
age, then the possibility of ecological bias arises. Given that this bias
results from inferring from a model with data at one level of aggregation
to another, this bias is also called cross-level bias.

There are some subtleties that arise regarding this question for disaster
epidemiology. First, it has been argued that, for some ambient exposures
like weather and air pollution, contextual effects (that is, at the community
level) are helpful to estimate in their own right, not just as a potential 
surrogate for individual-level associations. This is because regulations
(in the case of pollutants), interventions, and planning often happen at
this level.

Second, for a disaster, the relevant exposure might be not just at the
individual level (e.g., winds or flooding at the individual's residence), but
also throughout a broader area surrounding the individual. If storm damage
closes roads in a county, for example, or damages the power grid, then these
could create pathways for health risks for individuals throughout the county,
whether damage was high at their residence or not. The causal pathways for
tropical cyclones to affect human health differ from those for a dangerous
substance, like air pollutants, in which the substance itself must enter the
body to cause harm. While some health risk comes directly from the storm (e.g.,
deaths and injuries from trees falling on homes or drowning from flooding),
there are many more pathways that are indirect. These include pathways that go
through the way that the storm's damage affects community infrastructure and
access to medical care. Several papers have discussed the potential for 
"individualistic fallacy", as a counterpoint to "ecological fallacy", when 
such contextual-level exposure pathways are ignored.

There are study designs that can be used to leverage ecological data while
minimizing risk from ecological bias. For example, potential confounders like
age distribution and smoking rates vary much less within a county over time than
comparing between counties. The mechanism for ecological bias depends on the
joint distribution between individual exposure, outcome, and covariates, if any
are included in the model. Time series-style study designs, which compare a
county to itself over time, allow for very similar covariate distributions
between exposure and non-exposure. Other studies add to this design by
stabilizing for temporal confounding through the addition of counties that were
never exposed, allowing for a differences-in-differences style approach to
calibrate for seasonal or longer-term trends that might otherwise create
confounding. For example, many health outcomes have a strong seasonal trend,
with peak rates in the winter and lows in the summer. Since the hurricane season
stretches from summer into fall, a study design that compares the rate of a
health outcome in an exposed county to the rate two weeks before the exposure
might be biased away from the null, since baseline rates of the health outcome
will typically be moving up over most of the hurricane season.

Also, there may be confounders that are relevant at the contextual, rather than
individual level, as well as modifiers. Example: whether the county is coastal
could be a contextual-level confounder and effect modifier. This will influence
whether the county is exposed to that storm or not, since storms usually weaken
rapidly when the center is over land. In terms of counfounding pathways, coastal
communities might tend to have lower levels of air pollution, because sea
breezes clear the pollution regularly. They might also be wealthier on average,
since property on or near the beach is desirable. Finally, they might be better
prepared for or more hardened against tropical cyclones at the community-wide
level (e.g., through hardier power infrastructure, more rigorous building codes,
higher likelihood of evacuating in advance of a threatening storm) compared to
nearby inland counties.

Other approaches have also been suggested to reduce ecological bias while
leveraging ecological data, which might be available at a much larger
scale---and so provide more power---than is practical in collecting individual
data. These include multi-level approaches in which individual-level data from a
sample of individuals in each community is incorporated.

The second issue is exposure measurement error/misclassification. When the
exposure (e.g., binary exposure or continuous measure of intensity of exposure)
is only measured at the county level, all the individuals in that county are
assigned that same exposure. However, the hazards included in this study will
vary across each county. For some hazards, this variation will be more stark.
For example, tornadoes can be very local, causing extreme damage to a row of
homes and none to the homes across the street. Some types of floods can also be
very localized, rather than creating widespread flooding throughout the county.
In both cases, when a county-level exposure assessment is made based on an event
in the county in association with the storm, there may be many individuals in
the county who did not experience that hazard at their residence. The other
hazards we included---storm-associated winds and rainfall---will also have some
variation across counties, but this will often not be as stark. The rainfields
for tropical cyclones are very large, and while there are rainbands within the
storm that might have particularly high rates of precipitation, these progess
over the course of the storm, and it is unlikely that a county will have one
area that experienced very extreme precipitation while another experienced very
little. Winds might have more variation, especially in coastal counties and
counties near the storm's central track, since winds decay quickly as you move
from the storm's center, and storms typically weaken rapidly once they make
landfall and can no longer draw on warm ocean waters for energy.

This exposure data is provided at the county level, and this can lead to
exposure measurement error / exposure misclassification in studies that infer to
the individual level. In this case, there will be several processes at play.
First, the hazards of the storm will not be uniform across the county. The
degree of this within-county spatial heterogeneity will usually vary by hazard.
The within-county heterogeneity is likely lowest for rain and for wind at lower
levels. For rain, there can be particularly severe rain bands, but often a
storm's rainfield will bring high precipitation to a large area. Wind can have a
high heterogeneity close to the storm's center (storm winds decay most quickly
with distance near the storm's center) and near landfall (winds decay rapidly
after landfall). Other hazards can be very local within the county, only
affecting a small proportion of the total county area. A tornado is a clear
example---while it causes severe destruction in its track, areas just off the
track can be unharmed, creating a dramatic spatial gradient in damage within a
county. Floods can follow both. Some hurricanes have caused widespread flooding
(e.g., Floyd, Katrina) while some have caused flooding that is more spatially
focused (inland, with flashfloods? coast flooding just near the coast?). For
tidewater/coastal areas, there is often a dense network of waterway. When rains
occur further up the watershed, these all become flooded.

Second is the question of whether exposure measured at the resident's location
would be ideal. This is the case typically when studying an exposure that is a
chemical substance. In this case, the pathway depends on the substance entering
the body. For tropical cyclones, health risk could result from stress, problems
accessing medical care, or repercussions of power outages (higher exposure to
outdoor hazards like heat and air pollution, problems storing food and using
some medical equipment). For these, high levels of wind and other hazards in the
person's community may be as or more important than at their precise location.

Measurement error can be either random or systematic. Systematic error can often
be corrected with adjustment if the direction and typical size of the error is
understood. Random error cannot in the same way. Either type of measurement
error can be either differential (associated with the probability of the
outcome) or non-differential (independent of the distribution / probability of
the outcome). In simpler models, this characteristic might help in predicting
whether the resulting bias is likely toward the null; however, more complex
models (e.g., statistical models with adjustment for potential confounders) are
trickier to diagnose in terms of the likely implications of differential versus
non-differential measurement error [?].

Not only can the exposure be measured with error, but so too can potential
confounders and potential effect modifiers.

We provide exposure assessments at the county level. This spatial scale allows
for easy integration with health outcome data aggregated at the county level.
However, there are also some things to consider when modeling associations
at an ecological rather than individual level. The considerations are 
particularly nuanced when investigating the health risks associated with a 
disaster.

Often in epidemiology, a study aims to estimate how an exposure changes the
risk of a health outcome for an individual. The data might be collected for
lots of people, but the key question is this---if a person is exposed, how 
does his or her risk of the outcome change. If you are exposed to this hazard, 
how much more likely are you to have the outcome? To get sick, or injured, or
to die? This inference is at the individual level.

Even when individual-level inference is the goal, sometimes data is only 
available at an aggregated level. For example, privacy concerns might
mean that health outcomes are only made available as a county-wide daily 
count. You can fit a mode with this aggregated data, which will estimate the
association between exposure and health risk at the ecological level. This
contextual effect estimate is sometimes used as an estimate of individual-level
association, but doing so can introduce ecological/multi-level bias. The 
individual-level effect estimate can be very biased from the true association, 
even to the point of reversing the effect estimate---estimating a protective
effect when the true effect is detrimental, for example.

This ecological fallacy plays out in two ways. First, if you use a single 
exposure measurement for everyone in an area, some people will be misclassified
(or have exposure measured with error, for a continuous metric) unless the 
exposure is perfectly homogeneous across the area. This will typically not
be the case. This exposure misclassification can bias estimates of the 
association between exposure and outcome in the same way exposure 
misclassification through any other mechanism would.

Next, ecological bias could result from confounding, even if the confounders
are controlled in the ecological-level model. When you aggregate data, you 
lose information about how both the exposure and potential confounders vary
across individuals within the area of aggregation. As a result, a factor
could still confound the inference of an individual-level association, even 
if it is controlled for at a population level in an ecological model. For 
example, a study of the association between risk of pre-term birth and 
tropical cyclone exposure could control for county-level smoking when 
modeling county-level storm exposure and county-level rates of pre-term births.
Even with this control, an observed association could result from differences
in individual smoking status, if there is within-county variation in 
smoking and if this has a different pattern across people in the county than
variation in exposure from the county-wide exposure estimate.

For disasters, there are added nuances. First, for exposure it may not just 
matter what the level of exposure is in a person's immediate scope. Disasters
bring physical hazards that can harm people directly, but also through indirect
pathways. For example, a tropical cyclone can bring high winds that cause 
power outages, and as a result those affected could be exposed to more outdoor
hazards (outdoor air pollution, heat), struggle to safely store perishable
food and medications, and lose means to power medical equipment. While extreme
winds at a person's residence would increase their risk of a power outage, 
outages could also be caused by damage to the grid in another part of the 
community. In some cases, then, the level of exposure in a person's community
may be as important in opening a pathway of risk as exposure at the person's
immediate location.

What's more, if the disaster has a large health impact, the health outcome of
one person in the community could affect the risk of the outcome (or other
adverse outcomes) for others. This situation is often only the case for 
infectious diseases, where one person with the disease can spread it to others.
However, if the community-wide impact is large enough, it can affect access
to and effectiveness of medical care for everyone in the community. This
effect has been seen recently with Covid 19---attempts to "flatten the curve"
aim to avoid moving into a state where a community's health system becomes
overwhelmed and can no longer deliver a typical level of care to those in the
community. This effect could happen with either infectious or non-infectious
diseases.

When a proxy exposure estimates is used for many people covered by a study, this
can result in a type of exposure measurement error called *Berkson error*. In
this case, the true exposure of each individual is randomly distributed around
this proxy/aggregate-level exposure estimate, with a mean exposure level among
the subjects that is equal to the proxy/aggregate-level measure. In other words,
the group as a whole is assigned a common exposure level, when in fact this is
the average exposure level across members of that group, but the individuals'
true exposure levels are randomly distributed around this mean group level to
which they are all assigned for modeling.

Berkson:

$$
T = X + E
$$
where:

- $T$: true individual exposure
- $X$: proxy exposure, assigned to everyone in the group
- $E$: error, with mean of 0, independent of $X$

The dataset provides exposure by group/area (county), not by individual. These
ecological exposure estimates require us to assume that the exposure level is
the same for all individuals in that county. It ignores/smooths over
within-county variability in exposure.

For wind and rain, we're essentially basing exposure assessment on modeling of
exposure. For rain, this is a based on a re-analysis model that
incorporates/integrates observed data from the area and time, using a model to
integrate data from diverse sources. For wind, the exposure estimates are based
on a model, with a core based on the Willoughby wind model. For flooding and
tornadoes, the exposure is based on records of events anywhere in the county,
and in many cases may be based, essentially, on exposure assessment, as in many
cases someone like someone at a NOAA [?] center have determined whether the
event should be reported in the database.

Even exposure at each person's residence would not suffice, since people move 
around during the day (although perhaps less for most people during a severe
storm).

Data that are aggregated across a population---for example, the total number of
deaths in a geographic area in a certain time period---is known as *ecological
data*, and studies that base inference on this type of data are known as
*ecological studies*. Terms include *ecological*, *aggregate*, *contextual
level*.

> "In an ecological study the unit of analysis is a group of people rather than 
the individual." [@sedgwick2014ecological]

> "In his seminal paper, Robinson (1950) distinguished between two types of
correlation---ecological and individual. The former is obtained for a group of
people, while the latter is estimated for indivisible units, such as
individuals." [@portnov2007ecological]

Here is one challenge when working with aggregated data. If you use aggregated
data to infer individual-level associations, this can lead to *ecological fallacy*. 
This results from cross-level inference---the data used to model the association
is at the contextual level (e.g., county-level), while the inference is for the
individual association between exposure and outcome.

One definition of the ecological fallacy:

> "This spurious result provides an example of the ecological fallacy, in 
which conclusions ... based on ecologic data are opposite of those drawn 
on the basis of individual-level data." [@wakefield2008overcoming]

> "It is a bias produced when analyses realised in an ecological (group level)
analysis are used to make inferences at the individual level."
[@delgado2004bias]

> "Results from ecological studies are prone to the ecological fallacy. The 
ecological fallacy is a term used when collected data are analysed at a 
group level and the results are assumed to apply to associations at the 
individual level." [@sedgwick2014ecological]

Importantly, aggregated data can be used both to infer a contextual effect
(e.g., the association between county-wide rates of a health outcome and 
exposure) as well as an individual-level effect (e.g., how a person's
risk of the outcome is associated with that person's exposure to the hazard). 
The first inference will *not* be prone to ecological bias, as the inference
is at the same level as the data. The second inference (individual-level) 
will be susceptible to bias when aggregated data is used to fit the model. 
In this case, the association estimated based on aggregated data can be 
different, and even reversed, compared to the true individual-level effect.

> "I propose three criteria for the identification of ecological fallacy; all
three of these should be present to confirm its existence: (1) Results must be
obtrained with ecological (population) data; (2) Data must be inferred to
individuals. One use of ecological studies is to explore individual-level
association when individual data are not available. When the focus of the study
was contextual or based on population effects and there is no inference to
individuals, ecological fallacy is not possible. When only the first two
criteria are present---which is insufficient to affirm ecological fallacy---it
is appropriate to acknowledge that there is a possible relationship and that
further study is required; (3) Results obtained with individual data are
contradictory." [@idrovo2011three]

> "One important breakdown in the analogy between ecologic-bias and
individual-level confounding occurs because in etiologic research, the target of
inference for both ecologic and individual-level studies is the same: Both study
effects at the individual level. For an individual-level study, these target
effects are at the same level as the units of analysis. But, for an ecologic
study, these target effects are at a finer level than the units of analysis. As
a result, an ecologic study can be unbiased for ecologic effects (in particular,
ecologic confounding may be absent) and yet still be biased for individual-level
effects. Many of the classic social science examples of ecologic bias are of
this form." [@greenland1994invited]

> "We refer to an incorrect extrapolation from unbiased estimates of ecologic
effects to unobserved individual-level effects as 'cross-level bias' (although
the latter term is sometimes used to refer to any ecologic bias). As the
examples suggest, it is possible to detect cross-level bias if one can identify
and observe homogeneously exposed regions." [@greenland1994invited]

> "assuming that associations observed at the level of the area hold for the 
individuals within the areas can lead to the so-called ecological fallacy 
(Selvin, 1958)." [@wakefield2006health]

> "A broad definition of ecological fallacy is that certain data (e.g.,
illiteracy rates by foreign-born) are unavailable, leading to the use of proxies
and erroneous estimates. However, we advocate a more narrow definition of this
phenomenon, according to which 'ecological fallacy' refers solely to differences
in conclusions which may be drawn from group-level data (e.g., average
illiteracy rates by foreign-born), as opposed to data obtained for individuals."
[@portnov2007ecological]

> "If different conclusions are drawn from the analysis of data upon their
aggregations into units of different sizes (e.g., from individuals to townships
and regions), these differences are commonly referred to as 'ecological
fallacy'. In epidemiology, these differences are also known as ecological or
cross-level bias [refs]. The ecological bias is difficult to control and may
lead, under certain circumstances, to spurious effects [refs]. If
epidemiological study does not consider the possibility of ecological bias,
misguided policy decisions may follow [refs]." [@portnov2007ecological]

Ecological bias can *only* happen if there is, indeed, an individual-level
effect:

> "If exposure has no effect on any individual, there will be
no individual-level or ecologic effects, and so cross-level bias cannot occur. 
Thus, cross-level bias will not affect the validity of an ecologic test of the 
null hypothesis, although it still must be considered in interpreting a 
significant effect." [@greenland1994invited]

In some cases, the ecologic-level effect (contextual effect) will be directly 
of interest, and so we may be interested in inferring at that level (in which 
case use of ecologic-level data is appropriate and will not be prone to 
ecological bias):

> "The bias phenomenon in example 7 may be ascribed to the fact that an 
ecologic (regional) variable (percent Protestant) had effects on individual 
risk, in addition to effects of the corresponding individual-level variable
(religion) that the ecological variable summarized. In both social and 
infectious disease epidemiology, as well as in community intervention studies, 
ecologic effects may be of direct interest; a classic example is the 
phenomenon of 'herd immunity', in which the overall prevalence of 
immunity in a region, as well as individual immune status, determines the 
risk of individuals within the region (16). For noncontagious diseases, 
however, ecologic effects may not be of direct interest, and may in fact
obscure the individual-level effect of interest." [@greenland1994invited]

In some cases, the estimated association based on ecological data will be very 
different from the association that exists at the individual level:

> "We see that the ecological relative risk associated with low birth 
weight is completely incomparable with the individual-level coefficient."
[@wakefield2008overcoming]

Ecological bias can also complicate estimation of effect modification, which 
otherwise could help in identifying vulnerabilities and susceptibilities among
certain subpopulations:

> "Furthermore, non-White race now appears protective, rather than detrimental
as the individual-level analysis suggests." [@wakefield2008overcoming]

Ecologic bias crops up because the aggregated data fails to capture variation
within each aggregation level (county, for example) for both the exposure 
and potential confounders. For example, during a tropical cyclone, different
parts of a county will experience different peak wind speeds. If the county 
is coastal, the most severe winds will likely be near the coast, as the 
storm will weaken once it makes landfall and moves inland, as the storm 
generates its power from the water. When we assign the entire county a
single estimate of intensity, it smooths over the fact that locations in 
some parts of the county were subject to higher peak winds while locations
in other parts of the county had lower winds. At the least, this smoothing
of variation in exposure levels across the county could lower the power of 
the study to detect a clear association between exposure and health risk, as
this smoothing drops information inherent in the variation within the county
in exposure levels.

This, however, would not result in ecological fallacy unless another factor
comes into play---variation within each aggregation level in confounders. 
Just as aggregation smooths over within-area variation in exposure levels, it
also smooths over within-area variation in levels of potential confounders. 
Depending on the patterns of this within-area variation, a result could be
that ecological-level control of the confounders does not, in fact, control
for their role at the individual level, and so the association inferred at
the ecologic level continues to be confounded by them when inferred to the
individual level.

Essentially, this comes down to a question of the joint distribution, at the
individual level, of exposure, outcome, and covariates [?].

> "As various discussants have pointed out (3--6, 8, 9), ecologic relative-risk
estimates can be subject to biases not present in estimates from individual-level
observational studies of the same populations (case-control and cohort studies).
Unlike an individual-level study, an ecologic study does not link individual
outcome events to individual exposure or covariate histories, nor does it
link individual exposure and covariate histories to one another. It is these
linkage failures that are the source of the special biases of ecologic studies
(1, 2, 6)." [@greenland1994invited]

> "In the extreme, ecologic bias in comparisons limited to homogeneously exposed
regions (as in examples 3--6) can be viewed as purely an issue of confounding,
albeit with special complexities of measurement and control of confounders."
[@greenland1994invited]

> "More generally, one may show that cross-level bias will not occur if the 
individual-level effects of all variables (including unmeasured background
factors) follow a multiple linear-regression model with no regional effects and
no interactions (17). Nevertheless, given the usual inappropriateness of the 
multiple-linear model and the absence of homogeneously exposed regions, the 
possibility of cross-level bias adds a dimension to ecologic bias beyond
that of simple confounding." [@greenland1994invited]

> "Covariate control in ecologic studies requires attention to details not 
ordinarily of concern in individual-level studies. When, as is usually the
case, important nonlinearity or nonadditivity can be expected among 
exposure and covariate effects, it may be necessary to obtain and 
control for multiple summaries of joint covariate distributions in order
to insure that control is adequate." [@greenland1994invited]

> "Ecological fallacy can be produced by within group (individual level)
biases, such as confounding, selection bias, or misclassification, and
by confounding by group or effect modification by group." [@delgado2004bias]

> "**Confounding by group**: it is produced in an ecological study, when 
the exposure prevalence of each community (group) is correlated with 
the disease risk in non-exposed of the same community. It can be 
a mechanism for producing ecological fallacy." [@delgado2004bias]

If individual-level inference is the aim, and population-level data is
available, there are some methods for using it while still aiming to avoid
ecological bias. Indeed, it can be helpful to use population-level data, as it
is often available for a large population, improving the power and precision of
the study. Further, the level of exposure might vary a lot more over the
population captured with population-level data compared to the variation that
captured in a smaller sample of individual-level data. This can contribute both 
to statistical power and improve external validity (as the study data will cover
more of the range of exposure that might ever be expected). However, if
individual-level inference is the goal, then there are ways to supplement
population-level data with samples of individual-level data through two-level, 
semi-ecologic study designs.

> "It is generally well recognized that in order for ecologic data to provide
reliable inferences, they need to be supplemented with individual-level data."
[@wakefield2008overcoming]

> "For this example, we have access to complete individual-level data,
permitting a 'gold standard' individual-level analysis."
[@wakefield2008overcoming]

> "The only reliable way to characterize within-area variation in exposures
and confounders, and hence control ecologic bias, is to collect and incorporate
individual-level data. To help epidemiologists achieve this goal, in this paper
we describe the use of the two-phase design in an ecologic setting."
[@wakefield2008overcoming]

> "The only reliable way to characterize within-area variation in exposures
and confounders, and hence control ecologic bias, is to collect and incorporate
individual-level data. To help epidemiologists achieve this goal, in this 
paper we describe the use of the two-phase design in an ecologic setting."
[@wakefield2008overcoming]

> "In a semi-ecologic study, an ecologic exposure is combined with
individual-level outcomes and confounders. A two-phase approach is particularly
useful for such a study, with the phase 2 data corresponding to stratified
sampling of individual exposures." [@wakefield2008overcoming]

> "To implement a two-phase design, an initial phase 1 cross-classification by
the binary disease outcome and stratification variables is required; in phase 2,
samples of individuals are drawn from each of the cross-classification cells,
with data on additional variables being drawn from the subsamples of
individuals. Intuitively, the stratified sampling is focused on informative
cells, and estimation methods use both phases of data for efficiency and to
acknowledge the outcome-dependent sampling. In the simplest ecologic setting,
the cross-classification is by outcome and area only, and if area is a surrogate
for important risk factors this design will be efficient. We are particularly
interested in situations where an initial classification is available by
outcome, area, and confounders such as age and gender---this is the case in a
semi-ecologic study. Phase 2 may then provide detailed exposure information on a
subset of the phase 1 individuals." [@wakefield2008overcoming]

> "Two-phase study designs are a generalization of matched case-control designs
in which, initially, the entire sample population is cross-classified according
to case/control status and some stratification variable, S. The latter depends
on covariates observed in all individuals and may include exposures of interest, 
proxy exposure measures, or potential confounders. In settings like those we
are considering, such as environmental epidemiology, S may also depend on geographic
area, which can act as a surrogate for the totality of confounders associated
with each area, as well as provide a well-defined sampling frame for the controls."
[@wakefield2008overcoming]

> "The results of our simulation study point to the benefits associated with
combining the two sources of data, in terms of both bias and efficiency. Rather
than supplement an ecologic study with individual-level data, it may be of
interest to combine existing individual-level data with external group-level
data. Strategies that combine both types of data have been shown to alleviate
participation bias and improve efficiency in case-control studies with missing
data (31)." [@wakefield2008overcoming]

> "As with other observational studies, ecologic studies can give useful results
if biases such as those discussed here can be ruled out or quantified.
Nevertheless, bias evaluation can be especially difficult in ecologic studies of
geographic regions because of the many potentially interacting covariates that
may differ across regions. When biases cannot be ruled out with available data,
further exploration will require individual-level studies."
[@greenland1994invited]

Why we might sometimes use (and want to use) population-level data:

> "Because of the unavailability of individual-level data, ecological data may
be resorted to and may come in a variety of forms." [@wakefield2008overcoming]

> "Epidemiologists continue to use ecologic and aggregate data. Despite their 
known drawbacks, these data, often aggregated across geographic areas, offer the
advantages of widespread availability and gains in statistical power from large
populations and increased exposure ranges." [@wakefield2008overcoming]

> "Population and health data are often routinely available in ecological, that
is group, form while the exposure data typically consist of a set of values
recorded at monitor sites, or via one-off sampling." [@wakefield2006health]

It can be important to think about the scale that the process happens at. For
something very local, you might lose a lot more with aggregated data.

> "Data availability and exposure variability often determine the scale of
examination and the suitability of a study. Exposures arising from a point or
line source offer exposure contrasts on small scales, requiring small-area data;
in constrast, dietary variables show little variation across small scales, and
consequently international studies are used." [@wakefield2008overcoming]

There can also be an *individualistic fallacy*, if emphasis is limited to the
association between individual-level factors, without considering the potential
role of contextual factors in an individual's risk of a health outcome. In other
words, cross-level bias can go in both directions [@idrovo2011three]. While the
fallacy of inferring individual-level associations from ecologic-level data is
called *ecological fallacy*, the fallacy of inferring ecologic-level
(contextual) effects from individual-level data is known as the *atomistic
fallacy* [@idrovo2011three]. There are also fallacies that can come in, it
seems, from inferring at the same level as the data but failing to consider
influences from factors at the other level. When inferring individual-level
associations from individual-level data, without considering an additional role
of ecological-level factors, this is known as the *psychologistic* or
*individualistic fallacy* [@idrovo2011three]. When inferring contextual effects
from ecologic data without considering individual-level factors, this is known
as the *sociologistic fallacy* [@idrovo2011three].

> "Although ecological studies are important to epidemiology (especially in
environmental and social epidemiology), public health practitioners seem afraid
of ecological studies. It is a common practice to assume the presence of
ecological fallacy (Robinson 1950) and low-level validity when analyzing an
ecological study. Most epidemiologists prefer an exclusive individualistic
approach, although the importance of a multilevel causal approach is widely
recognized (Diez-Roux 2002). In this sense, some authors suggest that it is as
important to recognize the presence of ecological fallacy as to recognize
psychologistic or individualistic fallacy (Subramanian et al. 2009)."
[@idrovo2011three]

Confounding in general:

> "A confounder ... is associated with an exposure or risk factor for the
outcome and with the outcome independent of the exposure, but is not on the
causal pathway between the exposure and the outcome. Confounding occurs when the
relationship between an exposure variable and the outcome variable is
contaminated, so that the measure of association between these two variables is
actually also capturing the effects of a third variable, the confounding
variable. For example, many factors are associated with income (exposure) and
health (outcome), such as education, employment or wealth? It may be that these
factors explain part of the observed assocation." [@gunasekara2008glossary]

Measurement error in general:

> "Measurement error: The different between an observed variable and the
variable that belongs in a multiple regression equation. Measurement error is
the random or systematic error arising during data collection of variables. The
measured variable x is measured with error eta, which is the distance from x to
the 'true' value of x. Measurement error can produce bias such as attenuating
the estimators of exposure variables, but may also have more complex effects.
Longitudinal data analyses such as fixed effects models can actually augment the
problem of measurement error in mis-measured explanatory variables that change
little over time." [@gunasekara2008glossary]

Study validity:

> "Study validity refers to the degree to which the inferences drawn from a
study are warranted when account is taken of the study methods; the
representativeness of the study sample; and the nature of the population from
which it is drawn.[5] There are two types of study validity. *Internal validity*
is the degree to which the results of a study are correct for the sample of
people being studied. *External validity* (generalisability) is the degree to
which the study results hold true for a population beyond the subjects in the
study or other settings.[20]"

# R1 Comment 2

- Differs by exposure metric. Easier for some (wind, rain) than for others
(flood, tornado)
- For rain, we use a re-analysis product only available for the continental US
(right?). However, there are similar gridded re-analysis products available for
other regions worldwide that could be used in a similar way.
- For tornadoes, there may be national databases in other spots similar to the
tornado database that we use, but as far as we're aware, no international one.
- Same for floods. 
- For wind, the wind model could be generalizable with some tweaks. These
include:
  + More landmasks
  + Direction of storm movement for Northern vs Southern Hemispheres
  + Different wind averaging periods for the recorded best tracks data in
  different storm basins.
  + Our wind model was derived with US data and so may need bias-correction or a
  shift to a West Pacific--based model

> "Tornadoes long have been recognized as a global phenomenon (e.g., Wegener
1917; Feuerstein et al. 2005), having been recorded in every continent except
Antarctica. As such, numerous nationas share an interest in improving assessment
of their damage. International research and involvement in tornado survey work
and damage and intensity scales is well underway." [@edwards2013tornado]

For flood and tornado events, we used a database of severe weather events that
is specific to the United States (the US NOAA's Storm Events database). To
expand our exposure data for these hazards internationally, we would need to
find similar databases for other affected countries. [international flood and
tornado databases? EM-DAT? Others?]

The wind exposure data is based on modeling wind fields based on records of the
storm's track and central intensity throughout the tracking period. These values
are tracked for tropical cyclones around the world, and so it would be possible
to expand the dataset to encompass other countries for wind exposure
measurements. However, there are some practical issues that would need to be
addressed to do so with the approach we've taken.

The software we are using to model wind [@stormwindmodel] has several features
that make it specific to Atlantic basin storms. First, some steps in the model
depend on whether the storm's center or the location being modeled is over land
or water. These steps use a landmask---a grid with a binary indicator of land or
sea by latitude and longitude. Currently, the software only provides a landmask
that covers the United States, the Gulf of Mexico, and areas of the Atlantic
Ocean near the US. Second, the model assumes that cyclonic winds are
counterclockwise, and so is valid only for the Northern Hemisphere. Third, the
model assumes that the central intensity of the storm is based on [wind
averaging period], which is the standard used in the tracking data for
Atlantic-basin storms, but is not a common standard around the world. Instead,
some basins record central intensity measurements using averaging periods of
....

Therefore, it would be relatively straightforward to expand the wind exposure
data to cover other countries affected by Atlantic-basin storms, although this
would require some expansion of the area covered by our current landmask. More
modifications would need to be made to the wind model software before it could
be extended to storms in other basins, particularly basins that use different
wind averaging periods for their main tracking data and basins in the Southern
Hemisphere. Alternatively, other wind modeling software could be used for this
exposure assessment, as other groups have developed software that is more
extensible to multiple storm basins [Australian group's software].

The rain exposure data would also be fairly easy to extend to cover other
countries, although again there would be a few practical limitations to extend
the approach we've taken in developing the dataset covered in this manuscript.
For this exposure, we've used a re-analysis dataset. We've taken some steps to
translate this gridded data to the county scale, so it can be integrated with
the scale for many human impacts measurements. However, these steps have been
well documents [ref] and would be easy to reproduce. The re-analysis dataset
that we use is specific to the contiguous US [ref on NLDAS], but other similar
re-analysis datasets exist for other locations [example of one with global
coverage?].

For all of these, we have provided exposure assessments at the county scale. Any
of the steps we have taken could be done for other countries at a similar
geopolitical scale, although the name of that areal unit will vary by country
(e.g., municipalities in Mexico, districts in India). For wind and rain
exposures, the approach we take here could be used with any areal unit, as long
as there are geographical shape files available for the area boundaries
(although the modeling for wind may be to the area's geographic mean center
rather than population mean center).

> "We initially developed software to model exclusively Atlantic-basin storms
[@stormwindmodel]. There are two characteristics of Northwestern Pacific Ocean
basin storm data that prevent direct use of the current version of the software.
First, the software currently requires Western Hemisphere longitude inputs.
Second, the ``Best Tracks'' data from the China Meteorological Administration
include central tropical cyclone intensity using a 2-minute averaging period,
while the software currently assumes that central intensities are given for a
1-minute averaging period, as is the case for storm track data from the US
National Hurricane Center. With the assistance of Ms. Schumacher, we will update
our tropical cyclone wind model software to appropriately handle 'Best Tracks'
data from the China Meteorological Administration, incorporating conversion
factors [@harper2010guidelines] to convert central storm maximum sustained wind
measurements to a common averaging period prior to applying further steps of the
wind model."

### 2020-08-16 ehp_revision/response_to_reviewers.Rmd
Add some more ideas for a response

@@
 evident between 1894 and 2011." [@edwards2013tornado]
+
+This question could be explored to some degree through a study that uses the 
+physical hazard data provided in this dataset as the exposure and FEMA disaster
+declarations as the outcome. Such a study could clarify the degree to which 
+each hazard---and its intensity---helps in explaining that probablity that 
+a disaster declaration is issued. It could also clarify the degree to which 
+other factors, beyond the intensity of these hazards, drive the probability of
+a county receiving a disaster declaration. For example, if two counties 
+are exposed to a storm-associated hazard of the same intensity, they may 
+still have different probabilities of associated societal impacts, as a 
+community that is regularly exposed to that hazard may be more "hardened" 
+against it. Other differences in infrastructure and factors associated with 
+vulnerability could similarly introduce variability.
+
+At a more complex level, studies could explore the causal pathways from the 
+initial physical hazard to resulting damage/property loss through to associations
+with health outcomes, to provide a clearer idea of the direct and indirect 
+pathways leading to health impacts from tropical cyclones.
+
+For both of these studies, however, there may be better sources of damage
+estimates to use than FEMA disaster declarations. FEMA disaster declarations 
+can be prone to political and other influences, and since their primary function
+is not to record damage but rather to be a step in getting help to communities, 
+they can be problematic to use as an unbiased estimate of damage levels in 
+epidmiological research. One study, for example, found ... . There are other
+estimates of property damage that might be less problematic, including the 
+estimates recorded in NOAA's Storm Events database and data from private 
+insurers, although the quality of disaster loss data is well known to be
+less than ideal for secondary research [can be prone to bias through a number
+of mechanisms] [ref to Gall paper?].
 

### 2020-08-17 ehp_revision/response_to_reviewers.Rmd
Pull in for response from darlings

@@
 of mechanisms] [ref to Gall paper?].
+
+FEMA declaration is largely subjective, requiring state governors and/or tribal
+chiefs to submit a proposal for government aid on the basis that local resources
+are overwhelmed [@FEMA]. As a result, using FEMA disaster declarations as
+exposure classification to hurricanes restricts exposure to 1) those states
+whose resources are overwhelmed and 2) those states who apply for FEMA aid.
+(FEMA is likely less objective given societal effects (politics) [@Kunkel1999])
+As a result, other severely impacted counties and states that may not fit FEMA
+criteria are excluded and may therefore be incorrectly identified as unexposed,
+preventing researchers from assessing the full impact of a given hurricane
+[JMF-- the Grabich 2016 study actually found that FEMA assigned the highest
+exposure, and may overestimate the health effects of a hurricane; GBA-- let's
+think about why that might have had "false positives". I think that would be an
+interesting point to include here.]. Researchers have found heterogeneity in
+exposure assignments of counties, both within and between storms, when different
+metrics of exposure are used [@Grabich2016; @Grabich2015]. Such
+missclassification can have important consequences in assessments of the public
+health and economic impacts of tropical storms.
+
+One study used FEMA disaster declarations as a binary indicator of tropical
+storm exposure [@Grabich2016]. They found that this metric tended to assess a
+lot more counties as "exposed" to a storm than distance- or wind-based metrics
+[@Grabich2016], although they only compared the two exposure metrics as applied
+to four storms (Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in
+2004).
+
+Data on storm impacts, including property damage or insurance claims data, are
+sometimes available through storm databases or publications (e.g., NOAA's Storm
+Events data, *Monthly Weather Review*). While such data could potentially be
+used to create an exposure metric, tropical storm damage data requires some
+normalization, to incorporate both changes in dollar value over time and also
+changes in development in at-risk areas, to be comparable over extended time
+periods [@Pielke1998].
 
@@
 
+The largest-scale county-level exposure study is likely that of @Zandbergen2009,
+which estimated exposure in US counties to all US landfalling Atlantic basin
+tropical storms between 1851 and 2003, using both distance and an exposure
+metric that incorporated distance and windspeed. They used this exposure
+evaluation to create maps of total exposure to tropical storms within US
+counties, as well as to explore associations between a county's long-term
+exposure to tropical storms and its location, distance from the coast, size, and
+shape [@Zandbergen2009]. @Kruk2010 explored exposure to hurricane-related winds
+in the United States, including inland areas, for 1900-2008.
+
+Geographical patterns can differ for these different
+storm hazards, and so exposure assessment based on one hazard could create
+exposure misclassification if the pathway for health risks is through a
+different storm hazard.
+
+A tropical cyclone's high winds can bring health risks and property damage
+through structural damage of houses and other buildings, falling trees, and
+wind-borne debris [@rappaport2000] and cause power
+outages [@liu2005; han2009], which introduce a number of threats to
+human and ecological health, including water quality risks if the outage affects
+wastewater treatment plants [@mallin2006].  Other risks can exist
+without severe wind; for example, one study found that most of the direct
+hurricane-related deaths in the US between 1970 and 1999 occurred in cases
+when wind was below hurricane strength, including for Tropical Storms Charley
+in 1998 and Alberto in 1994 [@rappaport2000].  Tropical cyclones can
+produce excessive rain, especially in certain topographies (e.g., near
+mountains), so counties well inland sometimes experience more extreme rain than
+coastal counties. Flood risks from tropical cyclones can result from this rain,
+although the two risks are not perfectly correlated [@chen2015]. In the
+US, over half of hurricane-related direct deaths from 1970 to 1999 from
+Atlantic basin storms were a result of freshwater
+flooding [@rappaport2000]. Flooding can also degrade water
+quality [@mallin2006], which can threaten both human and ecological
+health.
+
+If a study misclassifies exposure to the hazard or hazards that cause the
+health risk being studied, the study will generate biased estimates of tropical
+cyclone risks and impacts. Further, when different studies use different
+methods to assess exposure to tropical cyclones, their results are difficult to
+meaningfully compare and aggregate. If different methods identify
+similar sets of communities as ``exposed'',  these concerns are less serious.
+However, if different methods differ substantially in which communities they
+identify as ``exposed'', it makes it very important that epidemiological
+studies are thoughtful in how they assess exposure.
+
+These analyses can also help researchers compare and interpret results from
+previous epidemiological studies that have assessed tropical cyclone exposure in
+different ways, helping to [synthesize] previous epidemiologic results to form a
+better picture of how these storms can affect human health.
+
+However, these frequency maps, together with evidence from
+specific tropical cyclones (Figures \ref{fig:ivanexposure}
+and \ref{fig:jaccard}),  do illustrate the potential for strong differences in
+spatial patterns in tropical cyclone exposures, depending on which tropical
+cyclone hazards are considered.  A few previous studies have sought to
+determine county-level exposure to tropical cyclones over multi-year periods,
+including @zandbergen2009, which estimated exposure in US
+counties to all US landfalling Atlantic-basin tropical cyclones
+between 1851 and 2003, using both a distance-based metric and a metric that
+combined distance and windspeed, and [@kruk2010], which explored
+exposure to hurricane-related winds in the US, including inland areas,
+for 1900--2008.  Our results suggest that such exposure assessments may
+perform well in capturing some tropical cyclone hazards (e.g., wind), but
+likely miss other potentially dangerous tropical cyclone exposures, especially
+for hazards that repeatedly threaten northern or inland counties (e.g., rain,
+flooding).
+
+Therefore, use of distance to assess tropical cyclone exposure for impact
+studies could result in problematic exposure misclassification, which could mask
+true associations, even strong associations, between tropical cyclone exposure
+and outcomes of interest in impact studies [@savitz2016interpreting;
+armstrong1998effect]. The use of a single hazard-based metric (e.g., wind) could
+cause similar problems if the impact is driven, at least in part, by a different
+hazard or by multiple hazards of the tropical cyclone.
+
+If measurements of one storm hazard are used as a proxy to capture exposure to
+a different hazard, or a non-hazard proxy like distance to the storm track is
+used to assess exposure to storm hazards, the resulting exposure
+misclassification could be differential (i.e., associated with the outcome of
+interest or with factors associated with risk of the outcome of interest).
+
+Therefore for most tropical cyclones, since exposure assessments often differ
+across storm hazards, as well as between each storm hazard and a distance-based
+proxy measurement, exposure misclassification is a potential risk.
+
+Further, we investigated agreement in exposure assessment across these metrics
+when applied individually. We found large differences in which counties are
+exposed to different hazards of tropical cyclones and that distance is, at
+best, a moderate, and often a very poor, surrogate for exposure to the specific
+tropical cyclone hazards of high wind, extreme rainfall, flooding, and
+tornadoes. Use of distance as a surrogate for any of these hazards, or exposure
+assessment based on one hazard when the pathway for health impacts is in part
+or full through another storm hazard, could lead to exposure misclassification.
+In the case of tropical cyclone risk and impact studies, including
+epidemiological studies, this would result in biased estimates.  Our findings
+highlight the importance of clarifying the potential pathway from tropical
+cyclone hazards to health impacts when conducting tropical cyclone
+epidemiological studies, and then basing exposure assessment on these hazards.
+
 **Response:**
@@
 
+As suggested, we have added that these events are based on reports in the
+Methods, in addition to our previous discussion of the point in the Discussion:
+
+> **Revised text (relevant addition in bold):**
+To identify flood- and tornado-based tropical cyclone exposures in US counties,
+we matched storm tracks with event listings from the National Oceanic and
+Atmospheric Administration (NOAA)’s Storm Event Database (NOAA NCEI 2020).
+**Events are listed in this database based on reports from ... and other
+sources.** While this database has recorded storm data, particularly tornadoes,
+since 1950, its coverage changed substantially in 1996 to cover more types of
+storm events, including flood events (NOAA NCEI 2020). We therefore only
+considered flood metrics of tropical cyclone exposure for storms in 1996 and
+later."
+
 \begin{shaded}
@@
 **Response:**
+
+This database is NOAA's database of records on tornado events in the United 
+States. It is the original source for the only other national tornado database
+available for the country (the SPC's ...), which provides a version of the 
+same data in which tornado events are joined into a single record in cases where 
+they cross county lines. There is no other comparable source of data on tornado
+occurrences that could be used for a comparison as validation.
 
@@
 
-Wind: gale force?
-
-Precipitation:
-
-Distance:
+For all outcomes, there are various reasonable choices that can be picked for
+the threshold in defining exposure. There is not room to go into these details
+within Table 1, so instead we added a separate table to the Supplemental
+Material that provides both some reasoning behind the thresholds selected in
+this study and also other reasonable thresholds that could be considered.
+Ultimately, this choice will depend on what a researcher considers as plausible
+causal pathways for the outcome he or she is studying. For example, if power
+outages form a plausible step in the pathway, then the threshold for gale-force
+or strong gale-force winds might be reasonable, while if the pathway is more
+likely to be through larger-scale building destruction, a higher threshold would
+be more reasonable. Some studies have done sensitivity analyses to consider how
+results changed using different exposure metrics. For example, @Czajkowski2011
+performed a sensitivity analysis using different assumptions about wind decay
+away from the center of a storm, while @Currie2013, who used a distance-based
+exposure metric, considered different buffer distances as a sensitivity
+analysis.
+
+- **Wind:** This threshold is the threshold for gale-force and higher winds
+based on the Beaufort wind scale. This value is used by the National Hurricane
+Center [?] as the outer limit of winds for its wind radii product in storm 
+tracking data. Other reasonable choices for a threshold of wind exposure 
+could include the two more intense thresholds also used in the NHC [?] wind
+radii estimates of storm extent (50 knots and 64 knots, the later of which 
+is the threshold for hurricane-force winds on the Beaufort Scale) or other
+thresholds on the Beaufort Scale (e.g., "strong gale", which starts at 
+41 knots; "storm", which starts at 48 knots; "violent storm", which starts
+at 56 knots).
+
+- **Rain:** This threshold is ... Other reasonable thresholds might be ... or
+might incorporate different ranges of days around the storm's closest approach
+to a county in calculating the cumulative precipitation. For example, ... For
+the most part, using our window (two days before to one day after a storm's
+closest approach) captured most of the storm-associated precipitation that would
+have been counted using a wider window from five days before to two days after a
+storm's closest approach, even for a large, slow storm. However, for a storm
+that loops and so passes some counties twice, such as Allison (2001), this
+four-day window may not be adequate, and lead to an underestimation of total
+rainfall associated with the storm relative to estimates for faster-moving
+storms. Other storms that looped over or near land and so passed close to US
+counties twice include Allison (1989), Alberto (1994), Gordon (1994), Dennis
+(1999), Edouard (2002), Fay (2002), Alex (2004), Ivan (2004), Dennis (2005), and
+Ophelia (2005). One difficulty in measuring rain exposure consistently across
+many different storms is how many days to include in the rain total. Some
+storms, especially ones associated with dangerous rainfalls, can move very
+slowly, meaning a window of several days needs to be used to fully capture
+storm-related rainfall. For example, one study used a period of over four days
+to provide storm rainfall totals from the slow-moving Hurricane Danny in 1997
+[@Medlin2007]. Single-storm studies often measure rainfall as the total rainfall
+over the full "event" (e.g., Medlin2007); however, the timing of the event is
+determine in a customized way for each storm, making it difficult to extend to
+multi-storm analyses. We included more days prior to the storm because
+storm-related rains often precede the passing of the storm's center [ref].
+
+- **Distance:** This threshold is one that has been commonly used in studies in
+which distance from the storm's track is used as a proxy for tropical cyclone
+exposure (for example, [refs]). Further, [anything in relation to storm size?].
+Other reasonable thresholds for distance include ..., all of which have also
+been used in previous research in assessing tropical cyclone exposure using
+distance from the storm's track as a proxy. Other distance-based exposure
+metrics have defined "exposure" to occur only when a storm's center track passes
+through a county's boundary (e.g., @Zandbergen2009, who describe this metric of
+exposure as a "hit" on the county; @Kinney2008 is another example that
+considered highest exposures for counties directly on the storm's path). This
+way of measuring exposure can exclude nearby counties that suffered extreme
+conditions from the storm but were not directly on the hurricane's track
+[@Kruk2010], especially since hurricanes can have a width ([typical storm width;
+citation])[JMF-- Weatherford and Gray paper referenced in Grabich 2016 may be a
+good place to start. Suggest 30-60km width of the eye of the storm itself] that
+is much larger than a typical county's width and since a hurricane's maximum
+winds can extend well away from the storm's center. Further, exposure assessed
+using this method is associated with the size [@Zandbergen2009; @Kruk2010] and
+shape [@Zandbergen2009] of a county. (While some studies have avoided this
+problem by determining exposure for equal-area grid cells (e.g., [@Kruk2010]),
+in studies where exposure needs to be matched with a county-level outcome, some
+process of aggregation from grid cells to counties would still need to be
+performed.) By contrast, assessing distance-based exposure based on measuring
+distance between the storm track and each county's center would be less
+susceptible to differences in county size and shape and so might be a more
+reliable metric of exposure than determining if a storm track ever crossed a
+county boundary. Further, assessing the minimum distance between county center
+and the storm track, as we do here, allows for a continuous, rather than binary,
+metric of distance-based exposure to a storm. To date, distance parameters
+involved with assessing risk of a particular storm have been rather arbitrary,
+contributing to the necessity in understanding how such parameters influence a
+county's exposure status. In assessing the risk of a given storm based on
+distance, recent studies have defined the distance from the storm track affected
+by a given storm differently. @Czajkowski2011 assessed county-level risk and
+exposure based on a three-tierd definition, with primary counties being those
+closest to the storm track on either side, secondary counties being adjacent to
+primary coutnies, and tertiary counties adjacent to secondary counties. Such a
+definition resulted in an exposure defenition based on an average distance
+radius of 120 km [75 miles] on either side of the storm track [@Czajkowski2011].
+Such a distance is slightly larger than that commonly used by public health
+departments (i.e., 100 km) [@Czajkowski2011]. Other studies used distance
+buffers of 30 kilometers [@Currie2013] or of 30, 60, and 100 kilometers as
+thresholds for identifying distance-based exposure [@Grabich2016] [check the
+Grabich paper-- was this distance to any part of a county, or to the center of
+the county?; JMF-- I think any part of the county]. Another study used a
+distance criterion of 60 kilometers from the storm track, classifying all
+counties for which the 60 km buffer fully or partially passed through the
+county's boundaries as exposed [@Grabich2015] [again, check to see if this was
+to any point in the county or to some center point of the county][JMF-- it is
+not abundantly clear in the Grabich 2016 article; based on their graphs it seems
+like they considered a county exposed if the definition they used crossed any
+county boundary. The Grabich 2015 paper specifies both full and partial counties
+being included]. In a study assessing the association between hurricanes and
+undesireable birth outcomes, researchers found that results were not sensitive
+to the omission of residences 100 km from the storm path, and that results
+varied insignificantly from 30-75 km [@Currie2013]. One study used distance to
+determine exposure to landfalling US hurricanes by assigning counties along the
+storm's tracks as "primary" exposures, counties neighboring primary counties as
+"secondary" counties, and counties neighboring secondary counties as "tertiary"
+counties; this method correspond to a distance buffer of approximately 75 miles
+(120 kilometers) from the storm track for a county to qualify as "exposed" to a
+storm [@Czajkowski2011]. The study found that about 30 direct hurricane
+fatalities over their study period (1970--2007) occurred outside of this
+distance range from the storm center [@Czajkowski2011], suggesting that this
+distance choice can cause some counties with dangerous exposures to a storm to
+be mis-identified as "unexposed". Tropical cyclones vary dramatically in size:
+US tropical cyclones have been observed with radii to maximum winds as small as
+20-km and as large as 200-km [@mallin2006; @quiring2011variations].
 
@@
 
-[Also revise figure legend?]
-
 ```{r fig3, fig.cap="Reproduction of Figure 3 from the original manuscript (left) and revised version in the resubmitted manuscript (right).
@@
 The revised version also uses a more focused range for the x axis, as requested in EHP's Request 1.
-Figure caption in the revised manuscript: \\textit{Figure caption.}", fig.show = "hold", out.width="50%"}
+Figure caption in the revised manuscript: \\textit{Comparison of two sources of wind exposure estimates: (1) modeled peak sustained surface wind and (2) estimates based on HURDAT2's wind speed radii.
+Each point represents a storm, with the x-axis giving the percent of counties classified in the same category of peak sustained surface wind (< 34 kt; 34--49.9 kt; 50--63.9 kt; $\\ge$64 kt) by both sources of data.
+The color of each point gives the number of study counties that were exposed to peak sustained surface wind of at least 34 kt (based on modeled wind).
+Estimates are shown for study storms since 2004, the earliest year for which post-storm reanalysis wind speed radii are routinely available in HURDAT2, and for which at least one study county had a peak sustained wind of $\\ge$34 kt based on the post-storm wind radii.}", fig.show = "hold", out.width="50%"}
 include_graphics("../figures/windcomparison.pdf")

### 2020-08-17 writing/darlings_methods.tex
Pull in for response from darlings

@@
 tropical cyclone.
+
 A tropical cyclone's rainfall can can begin a day or more before a

### 2020-08-17 ehp_revision/response_to_reviewers.Rmd
Add some more to response

@@
 **Response:**
+
+We hope that these results can help epidemiologists who are planning new
+multi-year studies of tropical cyclones and associated health risks. For 
+planning a multi-year study, it is helpful to know how many exposures the 
+study might incorporate, as this can help in giving a general idea of whether
+the study might have adequate statistical power to investigate a certain 
+research question. This table provides some basic summaries of the extent
+of exposure to different tropical cyclone hazards, both on average for the
+storms in the study and also for those storms that were most extensive. 
+Further, as with other results in the study, it provides evidence that 
+patterns in different hazards can vary substantially, both because some 
+hazards are much less likely to expose counties (e.g., torndaoes are a fairly 
+rare exposure, based on this table) and also because there is variation in 
+the most extensive storms across the hazards considered.
+
+We have added text to help the readers interpret this table ...
 

### 2020-08-17 ehp_revision/response_to_reviewers.Rmd
Work more on response

@@
 
-Great idea!
-
-We have investigated this. DesignSafe is ...
+It turns out that, because we have shared the data as an R package, it should
+already be accessible to use on the DesignSafe platform.
+
+DesignSafe is an online platform sponsered by the National Science Foundation as
+part of the Natural Hazards Engineering Research Infrastructure
+[@rathje2017designsafe]. This platform can be used to store, share, and publish
+citable data related to natural hazards through its Data Depot. It goes beyond a
+data repository, however, and also allows researchers to conduct data
+integration and analysis in the cloud through its Discovery Workspace, while
+drawing on data stored in the Data Depot and while using scientific programming
+tools.
 
 As part of its platform, DesignSafe allows users to work in a number of
-programming languages, including R (through Jupyter Notebooks). Therefore, 
-DesignSafe users already have access to any data shared through an R package, 
-including the data we present in this manuscript, without the data needing 
-to be separately loaded and published through DesignSafe's Data Depot.
+programming languages, including R (through Jupyter Notebooks)
+[@rathje2017designsafe]. Therefore, DesignSafe users already have access to any
+data shared through an R package, including the data we present in this
+manuscript, without the data needing to be separately loaded and published
+through DesignSafe's Data Depot.
+
+This is great, because it might otherwise be difficult to try to maintain copies
+of the dataset in different repositories (i.e., one copy as an R package and
+another as a DesignSafe dataset), and there could be confusion about things like
+the equivalence of the two copies. R packages are citable through a standard
+format that gives the R package version, while data publicly shared on
+DesignSafe is assigned a digital object identifiers (DOI)
+[@rathje2017designsafe], and so different citations would be used for the same
+dataset depending on how the user accessed it. If updates of one copy were not
+immediately updated for the other copy, the copies could get out of sync, and
+maintaining two synced copies would add would add an extra layer of work with
+each update to the R package (we have been updating several hazards as new data
+becomes available). DesignSafe's interface with R avoids this problem by
+allowing its users direct access to our package and its data through that path,
+as it allows for a single, citable, licensed version of the data stored and
+shared from a single location, with no question of syncing separate versions of
+the data or clarifying licensing and citation practices across two storage
+locations.
 
 We have added a note in the manuscript to let readers know that, because the
-data is shared through an R package, and because DesignSafe's [Workspace] allows
-users to work in R, the data can be fully accessed and used for those working on
-the DesignSafe platform by using the same code in a DesignSafe Jupyter notebook
-that would otherwise be used on a local computer.
-
-This is good, because otherwise there would be a good deal of overlap between
-the data on DesignSafe and that in the R package. R packages already created a
-citable source, comparable to the DOI assigned to data by DesignSafe when it is
-published. Further, R packages include licensing information, as does DesignSafe
-for published data. Finally, if the data were shared through both platforms, it
-would be critical to make sure that it was clear that the two datasets are
-analogous and also to keep them up to date with each other, which would add an
-extra layer of work with each update to the R package (we have been updating
-several hazards as new data becomes available). It is wonderful that
-DesignSafe's interface with R avoids this problem by allowing its users direct
-access to our package and its data through that path, as it allows for a single,
-citable, licensed version of the data stored and shared from a single location,
-with no question of syncing separate versions of the data or clarifying
-licensing and citation practices across two storage locations.
-
-*What is DesignSafe:*
-
-> "DesignSafe is the cyberinfrastructure platform that has been developed as
-part of NHERI to support natural hazards engineering research, and
-it succeeds the NEEShub cyberinfrastructure that was developed for the
-earthquake engineering community through the Network for Earthquake
-Engineering Simulation (NEES) program." [@rathje2017designsafe]
-
-> "DesignSafe plays an important role in integrating the various NHERI
-components and the research taking place at NHERI facilities, but also has the
-broader goal of enabling transformative research in natural hazards and
-engineering across the numerous technical disciplines engaged in this field."
-[@rathje2017designsafe]
-
-> "DesignSafe has been developed as a flexible, extensible, community-driven
-cyberinfrastructure, and it embraces a cloud strategy for the big data generated
-in natural hazards engineering. DesignSafe provides a comprehensive CI that
-supports the full research lifecycle from planning to execution to analysis to
-publication and curation." [@rathje2017designsafe]
-
-> "The future of natural hazards engineering research requires the integration
-of diverse data sets from a variety of sources including experiments,
-computational simulation, and field reconnaissance, as well as a variety of
-research disciplines including earth science, social science, building science,
-and architecture. The DesignSafe cyberinfrastructure has been designed to
-provide the functionalities that will enable transformative research in natural
-hazards engineering. By adopting a cloud strategy, DesignSafe allows for a
-fundamental change in the way that research is performed. It provides a
-comprehensive cyberinfrastructure that supports research workflows, data
-analysis, and visualization, as well as the full lifecycle of experimental,
-field, and computational research required by engineers and scientists to
-effectively address the threats posed to civil infrastructure by natural
-hazards. The integration of data and computation in the cloud will enable new
-research discoveries in natural hazards engineering, which in turn can lead to
-more hazard-resilient civil infrastructure." [@rathje2017designsafe]
-
-*Access to R:*
-
-> "The real revolution takes place downstream of data generation. Here, data 
-analytics and visualization are performed in the cloud within the *Discovery
-Workspace* while accessing any data within the *Data Depot*. Researchers can 
-invoke common analysis programs, such as MATLAB, as well as other analysis/
-visualization tools, such as Jupyter notebooks. A Jupyter notebook is an 
-electronic notebook that allows users to embed rich text elements as well
-as computer code, graphs, and visualizations within a single notebook that
-can be shared through the web. Over 40 different programming languages are 
-supported in Jupyter, including Python and R, and MATLAB code can be easily 
-converted, making Jupyter a versatile tool for research. Performing analysis in 
-the cloud allows researchers to integrate and explore various data without
-tedious downloads. Additionally, using a seamless cyberinfrastructure to 
-complete all research tasks enables tracking and relating of the processes 
-applied to data." [@rathje2017designsafe]
-
-*Data Depot:*
-
-> "the Data Depot, a flexible data repository with streamlined data management
-tools." [@rathje2017designsafe]
-
-> "As data progresses towards publication the requirements for metadata 
-increase as metadata provides users with search and discovery functions. At
-the end of the research project the user may edit the information for 
-publication and complete the process of assigning digital object identifiers
-(DOIs) and applying the appropriate license. On-demand assistance from a 
-curator is available to provide training and guide users through their data
-curation and publication needs." [@rathje2017designsafe]
+data is shared through an R package, and because DesignSafe's Discovery
+Workspace allows users to work in R, the data can be fully accessed and used for
+those working on the DesignSafe platform by using the same code in a DesignSafe
+Jupyter notebook that would otherwise be used on a local computer.
+
+> **Text added to the Methods of the revised manuscript (added text in bold):**
+"**By sharing this data as an R package, it is also accessible through 
+cloud-based computing platforms that incorporate Jupyter notebooks, including
+the National Science Foundations's DesignSafe platform for natural hazards
+engineering research [@rathje2017designsafe].**"
 
@@
 
-"Synoptic meteorology" refers to ...
-A key idea is that observations from a widespread geographical area can be
-integrate to provide a picture of large-scale weather [features?], including
-large areas of high or low pressure [?].
-
-> "In meteorology, the term synoptic denotes simultaneous observations taken
-globally, or at least over a large area." [@willoughby2007hurricane]
-
-> "Nearly 20 years ago [from 1937] [Bjerknes] declared that numerous and 
-important problems might be solved with the aid of observations already 
-assembled in the archives of different bureaus, but that the difficulties
-in making these observations serve a special purpose were unsurmountable
-for the individual scholar. He added that, in particular, the synoptic 
-charts representing the momentary states of the atmosphere over large
-areas of the earth, in all the necessary details and exactitude, were 
-excessively difficult to assemble. His proposal was that there be entered
-at the central bureau of each country, on the appropriate maps, all the available
-observations taken in that country and that there be employed, by the different
-bureaus, maps that may be superimposed or juxtaposed." [@gregg1937international]
-
-However, to integrate data from many sources to provide inputs for broadscale,
-synoptic analysis, it is helpful to use some standards in collecting these data
-at different locations. To facilitate numerical weather forecasting and other
-synoptic meteorology, weather data are regularly collected at standardized
-times, so that the data can be more easily integrated as an input to models.
-This practice extends back over 100 years, when data were collected from many
-[offices?] in the US by telegram [?] at regular times during the day and used to
-create synoptic weather maps that captured major weather systems in the country.
-
-> "In order to allow comparison of surface conditions at the same time
-throughout the world, the groups regulating the collection of
-hydrometeorological data have established a standard time for taking
-observations. These observations will have times observed within the 10 minutes
-preceding the UTC standard time. Hence surface observations for meteorological
-use have timestamps reflecting 0000, 0300, 0600, 0900, 1200, 1500, 1800, 2100
-UTC." [@usnwstime]
-
-> "Since the end of World War II, observers have launched rawinsonde observation
-worldwide at 00 and 12 UTC (coordinated universal time). High-threat situations,
-such as impending TC landfall, may dictate observations at 6-h or even 3-h
-intervals. Interpolating rawinsonde observations to the model's mesh and
-starting the calculation at a synoptic time is the simplest way to prepare a
-model initial condition." [@willoughby2007hurricane]
-
-> "Because simple interpolation in space and time is a less-than-optimum way to
-use observations, nonsynoptic data require elaborate four-dimensional
-variational data assimilation schemes." [@willoughby2007hurricane]
-
-> "By 1941, charts for the 750, 500, and 250 mbar surfaces were being prepared
-daily for 0400, 1200, and 2000 GMT." [@grahame2000development]
-
-> "During the period considered here, NHC issued advisories 3 h after the
-standard synoptic time (0300, 0900, 1500, and 2100 UTC)." [@torn2012uncertainty]
-
-> "The operation data assimilation and forecast cycle interval is 6 h"
-[@aberson201010]
-
-> "Dan Petersen asked whether a data deficiency really exists with all the
-observational sources in the Atlantic region, or are the data not being properly
-used in the numerical models? ... Steve Lord emphasized the difficulty of the
-task because many of these observations are not obtained at the synoptic
-times..." [@elsberry1992there]
-
-> "In the modern era, forecasters rely on purely dynamical models that integrate
-the Navier-Stokes equations for atmospheric motions to represent both the storm
-itself and its surroundings. Numerical models are structured on a computational
-grid with temperature, moisture, and wind tabulated in more-or-less rectangular
-cells." [@willoughby2007hurricane]
-
-> "Predicting the weather required a widespread but highly coordinated team of
-observers and forecasters. Organization was key... They transmitted data from
-the field using forms and protocols adapted from those that Abbe had developed
-while forecasting the weather from Cincinnati, Ohio. The information arrived by
-telegraph coded to maximize accuracy and minimize the word count and changes.
-The reports piled up in a rush at the appointed hours." [@willis2006cleveland]
-
-> "Three times daily the duty forecaster broadcast a synopsis of prevailing
-weather conditions nationwide..." [@willis2006cleveland]
-
-> "The requirement of synoptic data collection is easily met by the
-'conventional' observing system now serving as the main source of meteorological 
-data throughout the world, since it is based on human observations which could 
-be scheduled at convenient synoptic times." [@morel1971initialization]
-
-Since large-scale weather features are very larger, relevant data for a synoptic
-map often comes from locations with different local times. The need for
-synchronized weather data collection therefore actually played an important role
-in establishing standardized time zones in the United States in [year]. Synoptic
-weather times are typically set based on the Universal Time Constant, with times
-recorded in Coordinated Universal Time (i.e., Greenwich ...), and time stamps on
-weather maps and data based on this time zone often is given a "Z" (for "Zulu",
-an earlier name for this timezone) after the time.
-
-> "Accuracy in forecasting depended on the accurate timing of observations. At
-first weather observers used local time because no system of standard time
-existed, either in the United States or abroad. Local time often meant railroad
-time, and that could vary as much as 20 minutes from the time proper to the
-local meridian. Abbe called for change, emphasizing the need for simultaneous
-readings by Signal Service observers in order to construct accurate charts and
-weather maps. In 1875, he recommended to the president of the American
-Meteorological Society the establishment of a committee on standard time. ...
-Abbe's report in 1879 recommended the system of meridians that is now in general
-use ... a system of time meridians one hour apart across the country, anchored
-on the Greenwich prime meridian." [@willis2006cleveland]
-
-> "From Abbe's perspective, however, the full potential for dynamic weather
-prediction would not be achieved without standard time worldwide."
-[@willis2006cleveland]
-
-> "The time in all hydrometeorological products is expressed according to 
-a single standard, which is the Universal Coordinated Time (UTC) 
-(formerly known as Greenwich Mean Time [GMT] or Zulu (zero) Time [Z])."
-[@usnwstime]
-
-> "The U.S. National Weather Service base all data and product times on the 
-UTC standard. ... A common practice is to include a time of the observation 
-or event denoting the time with a following capital letter 'Z' to indicate the 
-time as UTC." [@usnwstime]
-
-> "The extent and the hour of the observations changed over time. Originally the
-observers took the readings at 0735, 1635, and 2335, Washington time, but the
-last reading was soon changed to 2300 so that the information could be included
-in the morning newspapers. Before the introduction of standard time and time
-zones within the United States, a confusing multiplicity of local times existed.
-In 1870 there were over 100 such regional times. For the purposes of the
-meteorological observations, observers used Washington time until 1885.
-Thereafter the observers took their readings according to eastern time, or that
-at the 75th meridian west of Greenwich, England. Observers also made local-time
-readings from 1876 to 1881.[30]" [@raines1996getting]
-
-> "Once the observers had gathered the weather data, the means of reporting and
-disseminating it became most important. Like the Smithsonian, the chief signal
-officer made arrangements with the leading commercial telegraph companies to
-carry the tri-daily reports. Civilian telegraph experts established special
-circuits routed to the Signal Office. The initial arrangement with Western Union
-regarding transmission was only temporary, and at the end of the trial period
-the company refused to continue service. The House Appropriations Committee held
-hearings over the dispute and ruled that the company had a mandate to transmit
-the weather information as government business. The Signal Corps compensated the
-company, however, at rates determined by the postmaster general.[34]" [@raines1996getting]
-
-> "When making the daily telegraphic reports, the weather observers used special
-codes to reduce their length to twenty words in the morning and ten in each of
-the other two reports, thereby saving the government both time and money.[35]
-Regular transmission of the reports began at 0735 on 1 November 1870 from
-twenty-four stations stretching from Boston, Massachusetts, south to Key West,
-Florida, and west to Cheyenne in the Wyoming Territory. In addition to the
-station atop Mount Washington (opened in December 1870), the Signal Corps soon
-reached new heights in weather reporting with the station on Pikes Peak that
-began reporting in November 1873.[36]" [@raines1996getting]
-
-> "To provide a picture of weather conditions across the country, the observers
-made their reports as nearly simultaneous as possible. The weather service did
-not initially make forecasts, and the enabling legislation did not specifically
-call for it to do so. Eventually general forecasts, referred to as
-probabilities, emanated from the Signal Office in Washington. Locally, the
-observers posted bulletins and maps in the offices of boards of trade and
-chambers of commerce to provide weather information to the public. Post offices
-also displayed daily bulletins, and observers supplied local newspapers with
-data. Some communities appointed meteorological committees to confer with the
-chief signal officer and to serve as a check upon the operations of the local
-weather station. On the national level, the Signal Office in Washington issued
-daily weather maps compiled from the reports received from all the stations. It
-also published the Daily Weather Bulletin, Weekly Weather Chronicle, and the
-Monthly Weather Review. All were available for sale to the public. Myer
-estimated that through these various means at least one third of American
-households received the Signal Corps' weather information in some form. A
-railway bulletin service, initiated in 1879, enabled stations along many major
-railroads to display weather information.[37]" [@raines1996getting]
-
-> "The Signal Corps' telegraph network soon expanded beyond the eastern
-seaboard. In 1874 Congress enacted legislation directing the War Department to
-build lines to connect military posts and protect frontier settlements in Texas
-against Indians and Mexicans. Other acts authorized lines in Arizona and New
-Mexico and, somewhat later, the Northwest. These lines were intended to serve
-sparsely settled areas where commercial lines were not yet available. For the
-most part, soldiers maintained and operated the lines, but the Corps employed
-some civilians. As the telegraph extended its reach, the weather system also
-grew, because the operators doubled as weather observers." [@raines1996getting]
-
-> "Until 1934 the Weather Bureau offices operated 12--15 hours a day with two
-basic observations taken at 8 a.m. and 8 p.m. The observations were transmitted
-via telegraph. There were no satellite images and few upper air observations."
-\url{https://www.weather.gov/dvn/armistice_day_blizzard}
+Some weather data are collected at standardized times, so that observations from
+different locations can be meaningfully integrated to provide a view of
+large-scale (synoptic) weather systems and patterns [@usnwstime;
+@willoughby2007hurricane; @willis2006cleveland; @raines1996getting]. This
+practice extends back over 100 years, when weather data for the US were
+collected by local weather offices and sent by telegram at regular times each
+day into a central office, which used this information to create synoptic
+weather maps that captured major weather systems in the county and from this
+sent back forecasts to local weather offices, also at standardized times
+[@willis2006cleveland; @raines1996getting]. Synoptic weather times are typically
+set based on the Universal Time Constant, with times recorded in Coordinated
+Universal Time (i.e., Greenwich Mean Time), and time stamps on weather maps and
+data based on this time zone often is given a "Z" (for "Zulu", an earlier name
+for this timezone) after the time [@usnwstime;
+@morris2008time].
 
@@
 Atmospheric Administration (NOAA)’s Storm Event Database (NOAA NCEI 2020).
-**Events are listed in this database based on reports from ... and other
-sources.** While this database has recorded storm data, particularly tornadoes,
-since 1950, its coverage changed substantially in 1996 to cover more types of
-storm events, including flood events (NOAA NCEI 2020). We therefore only
-considered flood metrics of tropical cyclone exposure for storms in 1996 and
-later."
+**Events are listed in this database based on reports from National Weather
+Service personnel and other sources.** While this database has recorded storm
+data, particularly tornadoes, since 1950, its coverage changed substantially in
+1996 to cover more types of storm events, including flood events (NOAA NCEI
+2020). We therefore only considered flood metrics of tropical cyclone exposure
+for storms in 1996 and later."
 
@@
 
-We have revised the x-axis on the figure to clarify that the comparison is based
-on the wind radii dataset described in the text, as requested. The original and
-revised versions of this paper are reproduced in this response document as
-Figure \@ref(fig:fig3).
+Yes, it is the same. We have revised the x-axis on the figure to clarify that
+the comparison is based on the wind radii dataset described in the text, as
+requested. The original and revised versions of this paper are reproduced in
+this response document as Figure \@ref(fig:fig3).
 

### 2020-08-17 ehp_revision/writing/ehp_methods.tex
Work more on response

@@
 and Atmospheric Administration (NOAA)'s Storm Event
-Database~\parencite{stormevents}. While this database has recorded storm data,
-particularly tornadoes, since~1950, its coverage changed substantially in~1996
-to cover more types of storm events, including flood
-events~\parencite{stormevents}. We therefore only considered flood metrics of
-tropical cyclone exposure for storms in~1996 and later.
+Database~\parencite{stormevents}. \textbf{Events are listed in this database
+based on reports from National Weather Service personnel and other sources.}
+While this database has recorded storm data, particularly tornadoes,
+since~1950, its coverage changed substantially in~1996 to cover more types of
+storm events, including flood events~\parencite{stormevents}. We therefore only
+considered flood metrics of tropical cyclone exposure for storms in~1996 and
+later.
 
@@
 We have posted most study data as an open-source R package
-\parencite{hurricaneexposuredata}. We have posted remaining data and code at
+\parencite{hurricaneexposuredata}. \textbf{By sharing this data as an R package, it
+is also accessible through cloud-based computing platforms that
+incorporate Jupyter notebooks, including the National Science Foundations's
+DesignSafe platform for natural hazards engineering research
+\parencite{rathje2017designsafe}.} We have posted remaining data and code at
 \url{https://github.com/geanders/county_hurricane_exposure}.

### 2020-08-17 ehp_revision/writing/ehp_response_darlings.tex
Work more on response

@@
 wind model."
+
+# R1 Comment 6
+
+*What is DesignSafe:*
+
+> "DesignSafe is the cyberinfrastructure platform that has been developed as
+part of NHERI to support natural hazards engineering research, and
+it succeeds the NEEShub cyberinfrastructure that was developed for the
+earthquake engineering community through the Network for Earthquake
+Engineering Simulation (NEES) program." [@rathje2017designsafe]
+
+> "DesignSafe plays an important role in integrating the various NHERI
+components and the research taking place at NHERI facilities, but also has the
+broader goal of enabling transformative research in natural hazards and
+engineering across the numerous technical disciplines engaged in this field."
+[@rathje2017designsafe]
+
+> "DesignSafe has been developed as a flexible, extensible, community-driven
+cyberinfrastructure, and it embraces a cloud strategy for the big data generated
+in natural hazards engineering. DesignSafe provides a comprehensive CI that
+supports the full research lifecycle from planning to execution to analysis to
+publication and curation." [@rathje2017designsafe]
+
+> "The future of natural hazards engineering research requires the integration
+of diverse data sets from a variety of sources including experiments,
+computational simulation, and field reconnaissance, as well as a variety of
+research disciplines including earth science, social science, building science,
+and architecture. The DesignSafe cyberinfrastructure has been designed to
+provide the functionalities that will enable transformative research in natural
+hazards engineering. By adopting a cloud strategy, DesignSafe allows for a
+fundamental change in the way that research is performed. It provides a
+comprehensive cyberinfrastructure that supports research workflows, data
+analysis, and visualization, as well as the full lifecycle of experimental,
+field, and computational research required by engineers and scientists to
+effectively address the threats posed to civil infrastructure by natural
+hazards. The integration of data and computation in the cloud will enable new
+research discoveries in natural hazards engineering, which in turn can lead to
+more hazard-resilient civil infrastructure." [@rathje2017designsafe]
+
+*Access to R:*
+
+> "The real revolution takes place downstream of data generation. Here, data 
+analytics and visualization are performed in the cloud within the *Discovery
+Workspace* while accessing any data within the *Data Depot*. Researchers can 
+invoke common analysis programs, such as MATLAB, as well as other analysis/
+visualization tools, such as Jupyter notebooks. A Jupyter notebook is an 
+electronic notebook that allows users to embed rich text elements as well
+as computer code, graphs, and visualizations within a single notebook that
+can be shared through the web. Over 40 different programming languages are 
+supported in Jupyter, including Python and R, and MATLAB code can be easily 
+converted, making Jupyter a versatile tool for research. Performing analysis in 
+the cloud allows researchers to integrate and explore various data without
+tedious downloads. Additionally, using a seamless cyberinfrastructure to 
+complete all research tasks enables tracking and relating of the processes 
+applied to data." [@rathje2017designsafe]
+
+*Data Depot:*
+
+> "the Data Depot, a flexible data repository with streamlined data management
+tools." [@rathje2017designsafe]
+
+> "As data progresses towards publication the requirements for metadata 
+increase as metadata provides users with search and discovery functions. At
+the end of the research project the user may edit the information for 
+publication and complete the process of assigning digital object identifiers
+(DOIs) and applying the appropriate license. On-demand assistance from a 
+curator is available to provide training and guide users through their data
+curation and publication needs." [@rathje2017designsafe]
+
+# R3 Comment 2
+
+"Synoptic meteorology" refers to ...
+A key idea is that observations from a widespread geographical area can be
+integrate to provide a picture of large-scale weather [features?], including
+large areas of high or low pressure [?].
+
+> "In meteorology, the term synoptic denotes simultaneous observations taken
+globally, or at least over a large area." [@willoughby2007hurricane]
+
+> "Nearly 20 years ago [from 1937] [Bjerknes] declared that numerous and 
+important problems might be solved with the aid of observations already 
+assembled in the archives of different bureaus, but that the difficulties
+in making these observations serve a special purpose were unsurmountable
+for the individual scholar. He added that, in particular, the synoptic 
+charts representing the momentary states of the atmosphere over large
+areas of the earth, in all the necessary details and exactitude, were 
+excessively difficult to assemble. His proposal was that there be entered
+at the central bureau of each country, on the appropriate maps, all the available
+observations taken in that country and that there be employed, by the different
+bureaus, maps that may be superimposed or juxtaposed." [@gregg1937international]
+
+However, to integrate data from many sources to provide inputs for broadscale,
+synoptic analysis, it is helpful to use some standards in collecting these data
+at different locations. To facilitate numerical weather forecasting and other
+synoptic meteorology, weather data are regularly collected at standardized
+times, so that the data can be more easily integrated as an input to models.
+This practice extends back over 100 years, when data were collected from many
+[offices?] in the US by telegram [?] at regular times during the day and used to
+create synoptic weather maps that captured major weather systems in the country.
+
+> "In order to allow comparison of surface conditions at the same time
+throughout the world, the groups regulating the collection of
+hydrometeorological data have established a standard time for taking
+observations. These observations will have times observed within the 10 minutes
+preceding the UTC standard time. Hence surface observations for meteorological
+use have timestamps reflecting 0000, 0300, 0600, 0900, 1200, 1500, 1800, 2100
+UTC." [@usnwstime]
+
+> "Since the end of World War II, observers have launched rawinsonde observation
+worldwide at 00 and 12 UTC (coordinated universal time). High-threat situations,
+such as impending TC landfall, may dictate observations at 6-h or even 3-h
+intervals. Interpolating rawinsonde observations to the model's mesh and
+starting the calculation at a synoptic time is the simplest way to prepare a
+model initial condition." [@willoughby2007hurricane]
+
+> "Because simple interpolation in space and time is a less-than-optimum way to
+use observations, nonsynoptic data require elaborate four-dimensional
+variational data assimilation schemes." [@willoughby2007hurricane]
+
+> "By 1941, charts for the 750, 500, and 250 mbar surfaces were being prepared
+daily for 0400, 1200, and 2000 GMT." [@grahame2000development]
+
+> "During the period considered here, NHC issued advisories 3 h after the
+standard synoptic time (0300, 0900, 1500, and 2100 UTC)." [@torn2012uncertainty]
+
+> "The operation data assimilation and forecast cycle interval is 6 h"
+[@aberson201010]
+
+> "Dan Petersen asked whether a data deficiency really exists with all the
+observational sources in the Atlantic region, or are the data not being properly
+used in the numerical models? ... Steve Lord emphasized the difficulty of the
+task because many of these observations are not obtained at the synoptic
+times..." [@elsberry1992there]
+
+> "In the modern era, forecasters rely on purely dynamical models that integrate
+the Navier-Stokes equations for atmospheric motions to represent both the storm
+itself and its surroundings. Numerical models are structured on a computational
+grid with temperature, moisture, and wind tabulated in more-or-less rectangular
+cells." [@willoughby2007hurricane]
+
+> "Predicting the weather required a widespread but highly coordinated team of
+observers and forecasters. Organization was key... They transmitted data from
+the field using forms and protocols adapted from those that Abbe had developed
+while forecasting the weather from Cincinnati, Ohio. The information arrived by
+telegraph coded to maximize accuracy and minimize the word count and changes.
+The reports piled up in a rush at the appointed hours." [@willis2006cleveland]
+
+> "Three times daily the duty forecaster broadcast a synopsis of prevailing
+weather conditions nationwide..." [@willis2006cleveland]
+
+> "The requirement of synoptic data collection is easily met by the
+'conventional' observing system now serving as the main source of meteorological 
+data throughout the world, since it is based on human observations which could 
+be scheduled at convenient synoptic times." [@morel1971initialization]
+
+Since large-scale weather features are very larger, relevant data for a synoptic
+map often comes from locations with different local times. The need for
+synchronized weather data collection therefore actually played an important role
+in establishing standardized time zones in the United States in [year]. Synoptic
+weather times are typically set based on the Universal Time Constant, with times
+recorded in Coordinated Universal Time (i.e., Greenwich ...), and time stamps on
+weather maps and data based on this time zone often is given a "Z" (for "Zulu",
+an earlier name for this timezone) after the time.
+
+> "Accuracy in forecasting depended on the accurate timing of observations. At
+first weather observers used local time because no system of standard time
+existed, either in the United States or abroad. Local time often meant railroad
+time, and that could vary as much as 20 minutes from the time proper to the
+local meridian. Abbe called for change, emphasizing the need for simultaneous
+readings by Signal Service observers in order to construct accurate charts and
+weather maps. In 1875, he recommended to the president of the American
+Meteorological Society the establishment of a committee on standard time. ...
+Abbe's report in 1879 recommended the system of meridians that is now in general
+use ... a system of time meridians one hour apart across the country, anchored
+on the Greenwich prime meridian." [@willis2006cleveland]
+
+> "From Abbe's perspective, however, the full potential for dynamic weather
+prediction would not be achieved without standard time worldwide."
+[@willis2006cleveland]
+
+> "The time in all hydrometeorological products is expressed according to 
+a single standard, which is the Universal Coordinated Time (UTC) 
+(formerly known as Greenwich Mean Time [GMT] or Zulu (zero) Time [Z])."
+[@usnwstime]
+
+> "The U.S. National Weather Service base all data and product times on the 
+UTC standard. ... A common practice is to include a time of the observation 
+or event denoting the time with a following capital letter 'Z' to indicate the 
+time as UTC." [@usnwstime]
+
+> "The extent and the hour of the observations changed over time. Originally the
+observers took the readings at 0735, 1635, and 2335, Washington time, but the
+last reading was soon changed to 2300 so that the information could be included
+in the morning newspapers. Before the introduction of standard time and time
+zones within the United States, a confusing multiplicity of local times existed.
+In 1870 there were over 100 such regional times. For the purposes of the
+meteorological observations, observers used Washington time until 1885.
+Thereafter the observers took their readings according to eastern time, or that
+at the 75th meridian west of Greenwich, England. Observers also made local-time
+readings from 1876 to 1881.[30]" [@raines1996getting]
+
+> "Once the observers had gathered the weather data, the means of reporting and
+disseminating it became most important. Like the Smithsonian, the chief signal
+officer made arrangements with the leading commercial telegraph companies to
+carry the tri-daily reports. Civilian telegraph experts established special
+circuits routed to the Signal Office. The initial arrangement with Western Union
+regarding transmission was only temporary, and at the end of the trial period
+the company refused to continue service. The House Appropriations Committee held
+hearings over the dispute and ruled that the company had a mandate to transmit
+the weather information as government business. The Signal Corps compensated the
+company, however, at rates determined by the postmaster general.[34]" [@raines1996getting]
+
+> "When making the daily telegraphic reports, the weather observers used special
+codes to reduce their length to twenty words in the morning and ten in each of
+the other two reports, thereby saving the government both time and money.[35]
+Regular transmission of the reports began at 0735 on 1 November 1870 from
+twenty-four stations stretching from Boston, Massachusetts, south to Key West,
+Florida, and west to Cheyenne in the Wyoming Territory. In addition to the
+station atop Mount Washington (opened in December 1870), the Signal Corps soon
+reached new heights in weather reporting with the station on Pikes Peak that
+began reporting in November 1873.[36]" [@raines1996getting]
+
+> "To provide a picture of weather conditions across the country, the observers
+made their reports as nearly simultaneous as possible. The weather service did
+not initially make forecasts, and the enabling legislation did not specifically
+call for it to do so. Eventually general forecasts, referred to as
+probabilities, emanated from the Signal Office in Washington. Locally, the
+observers posted bulletins and maps in the offices of boards of trade and
+chambers of commerce to provide weather information to the public. Post offices
+also displayed daily bulletins, and observers supplied local newspapers with
+data. Some communities appointed meteorological committees to confer with the
+chief signal officer and to serve as a check upon the operations of the local
+weather station. On the national level, the Signal Office in Washington issued
+daily weather maps compiled from the reports received from all the stations. It
+also published the Daily Weather Bulletin, Weekly Weather Chronicle, and the
+Monthly Weather Review. All were available for sale to the public. Myer
+estimated that through these various means at least one third of American
+households received the Signal Corps' weather information in some form. A
+railway bulletin service, initiated in 1879, enabled stations along many major
+railroads to display weather information.[37]" [@raines1996getting]
+
+> "The Signal Corps' telegraph network soon expanded beyond the eastern
+seaboard. In 1874 Congress enacted legislation directing the War Department to
+build lines to connect military posts and protect frontier settlements in Texas
+against Indians and Mexicans. Other acts authorized lines in Arizona and New
+Mexico and, somewhat later, the Northwest. These lines were intended to serve
+sparsely settled areas where commercial lines were not yet available. For the
+most part, soldiers maintained and operated the lines, but the Corps employed
+some civilians. As the telegraph extended its reach, the weather system also
+grew, because the operators doubled as weather observers." [@raines1996getting]
+
+> "Until 1934 the Weather Bureau offices operated 12--15 hours a day with two
+basic observations taken at 8 a.m. and 8 p.m. The observations were transmitted
+via telegraph. There were no satellite images and few upper air observations."
+\url{https://www.weather.gov/dvn/armistice_day_blizzard}

### 2020-09-02 ehp_revision/response_to_reviewers.Rmd
Work some more on revision

@@
 States. It is the original source for the only other national tornado database
-available for the country (the SPC's ...), which provides a version of the 
-same data in which tornado events are joined into a single record in cases where 
-they cross county lines. There is no other comparable source of data on tornado
-occurrences that could be used for a comparison as validation.
+available for the country (the Storm Prediction Center [SPC]'s National Tornado
+Database), which provides a version of the same data in which tornado events are
+joined into a single record in cases where they cross county lines. There is no
+other comparable source of data on tornado occurrences that could be used for a
+comparison as validation.
 
 The NOAA Storm Events database originated as a database for recording tornadoes
-[? anything else] in the US in 195[x], and was originally called ... . The
-database has gone through a number of expansions to cover more events, and it
-now includes reports of other types of [hydrometeorological?] disaster events,
-including floods, extreme heat and cold, wildfires, frost, hail, ... . However,
-it has maintained its status as the database of record [?] for tornadoes in the
-United States.
+in the US in 1953, and was originally called the Climatological Data National
+Summary [@Ashley2007]. The database has gone through a number of expansions to
+cover more events, and it now includes reports of other types of
+hydrometeorological disaster events, including floods, extreme heat and cold,
+wildfires, frost, and hail. However, it has maintained its status as the
+database of record for tornadoes in the United States.
+
+We have changed the text in the manuscript to make this clearer:
+
+> "The tornado observations from this dataset---**along with the derived version
+of the data available through the Storm Prediction Center's National Tornado
+Database---are the tornado event data of record** for the US. **There are no other
+collections of tornado data comparable in temporal or geographical scope,** and so
+we did not further validate the tornado event data."
 

### 2020-09-02 ehp_revision/tables/rain_misclass.tex
Work some more on revision

% latex table generated in R 4.0.2 by xtable 1.8-4 package
% Tue Sep  1 17:11:55 2020
\begin{table}[ht]
\centering
\caption{Caption. Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.  Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.).} 
\begin{tabular}{cp{3cm}p{3cm}p{3cm}p{3cm}}
  \toprule
Storm & Classified as exposed based on both distance metric and rain metric & Classified as exposed based on distance metric but unexposed by rain metric & Classified as exposed based on rain metric but unexposed by distance metric & Classified as unexposed based on both distance metric and rain metric \\ 
  \midrule
Frances-2004 & 207 & 114 & 257 & 1818 \\ 
  Ivan-2004 & 124 & 160 & 288 & 1824 \\ 
  Isidore-2002 & 134 & 134 & 186 & 1942 \\ 
  Ike-2008 & 113 & 120 & 213 & 1950 \\ 
  Dennis-2005 &  48 & 244 & 124 & 1980 \\ 
  Fay-2008 &  72 & 180 & 163 & 1981 \\ 
  Jeanne-2004 & 144 & 153 & 113 & 1986 \\ 
  Lee-2011 & 106 &  38 & 230 & 2022 \\ 
  Gustav-2008 & 126 & 108 & 137 & 2025 \\ 
  Allison-2001 & 106 & 181 &  67 & 2042 \\ 
  Bill-2003 & 110 & 120 & 117 & 2049 \\ 
  Cindy-2005 &  78 & 243 &  11 & 2064 \\ 
  Floyd-1999 & 129 &  10 & 170 & 2087 \\ 
  Katrina-2005 & 195 &  46 &  58 & 2097 \\ 
  Danny-1997 &  82 & 147 &  54 & 2113 \\ 
  Rita-2005 &  64 & 101 & 103 & 2128 \\ 
  Gordon-2000 &   5 & 250 &   7 & 2134 \\ 
  Arlene-2005 &  39 & 197 &  23 & 2137 \\ 
  Dennis-1999 &  90 & 110 &  56 & 2140 \\ 
  Ernesto-2006 & 124 &  68 &  63 & 2141 \\ 
  Irene-2011 & 119 &   1 & 106 & 2170 \\ 
  Hanna-2008 & 100 &  92 &  33 & 2171 \\ 
  Hermine-2010 &  44 &  79 & 102 & 2171 \\ 
  Fran-1996 &  89 &  62 &  73 & 2172 \\ 
  Earl-1998 &  72 &  97 &  38 & 2189 \\ 
  Bertha-1996 & 103 &  80 &  20 & 2193 \\ 
  Helene-2000 &  28 & 164 &  11 & 2193 \\ 
  Frances-1998 &  27 &  35 & 139 & 2195 \\ 
  Josephine-1996 &  94 &  29 &  77 & 2196 \\ 
  Bonnie-2004 &  56 &  53 &  56 & 2231 \\ 
  Isabel-2003 &  88 &  59 &  17 & 2232 \\ 
   \bottomrule
\end{tabular}
\end{table}

### 2020-09-02 ehp_revision/tables/wind_misclass.tex
Work some more on revision

% latex table generated in R 4.0.2 by xtable 1.8-4 package
% Tue Sep  1 17:11:57 2020
\begin{table}[ht]
\centering
\caption{Caption. Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns). Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.} 
\begin{tabular}{cp{3cm}p{3cm}p{3cm}p{3cm}}
  \toprule
Storm & Classified as exposed based on both distance metric and wind metric & Classified as exposed based on distance metric but unexposed by wind metric & Classified as exposed based on wind metric but unexposed by distance metric & Classified as unexposed based on both distance metric and wind metric \\ 
  \midrule
Frances-2004 &  44 & 277 &  13 & 2062 \\ 
  Cindy-2005 &  17 & 304 &   4 & 2071 \\ 
  Dennis-2005 &  18 & 274 &  26 & 2078 \\ 
  Ike-2008 & 166 &  67 &  82 & 2081 \\ 
  Ivan-2004 &  29 & 255 &  27 & 2085 \\ 
  Jeanne-2004 &  41 & 256 &   7 & 2092 \\ 
  Allison-2001 &  21 & 266 &   0 & 2109 \\ 
  Isidore-2002 &  29 & 239 &   7 & 2121 \\ 
  Katrina-2005 &  66 & 175 &  32 & 2123 \\ 
  Gordon-2000 &  11 & 244 &   6 & 2135 \\ 
  Fay-2008 &  54 & 198 &   7 & 2137 \\ 
  Gustav-2008 &  36 & 198 &  23 & 2139 \\ 
  Bertha-1996 & 179 &   4 &  62 & 2151 \\ 
  Danny-1997 &  45 & 184 &  12 & 2155 \\ 
  Arlene-2005 &   4 & 232 &   0 & 2160 \\ 
  Bill-2003 &  19 & 211 &   0 & 2166 \\ 
  Dennis-1999 &  17 & 183 &  13 & 2183 \\ 
  Hanna-2008 & 135 &  57 &  12 & 2192 \\ 
  Isabel-2003 & 105 &  42 &  56 & 2193 \\ 
  Ernesto-2006 &  93 &  99 &   4 & 2200 \\ 
  Helene-2000 &  18 & 174 &   0 & 2204 \\ 
  Rita-2005 &  26 & 139 &  20 & 2211 \\ 
  Fran-1996 &  46 & 105 &  30 & 2215 \\ 
  Earl-1998 & 133 &  36 &  11 & 2216 \\ 
  Floyd-1999 & 124 &  15 &  40 & 2217 \\ 
  Irene-2011 & 114 &   6 &  46 & 2230 \\ 
  Lee-2011 &  34 & 110 &   0 & 2252 \\ 
  Josephine-1996 &  62 &  61 &   3 & 2270 \\ 
  Hermine-2010 &  22 & 101 &   1 & 2272 \\ 
  Bonnie-2004 &   0 & 109 &   0 & 2287 \\ 
  Frances-1998 &  15 &  47 &   0 & 2334 \\ 
   \bottomrule
\end{tabular}
\end{table}

### 2020-09-02 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 
-**Response:**
+**Response:** Thank you for your helpful suggestions. We have responded to 
+each in detail below.
 
@@
 
+**Response:**
+
 Thank you for this suggestion, and we can see that we could have more clearly 
@@
 among exposures to specific hazards can help epidemiologists to design their
-studies and the statistical methods to analyze the resulting data.
+studies and the statistical methods to analyze the resulting data. We provide
+a detailed response in our response below to Comment 1 from this reviewer.
+
+We have moved other suggestions that you sent in "Track Changes" of the Word 
+document for the original manuscript and address each in detail below. When 
+there were several comments on a related topic, we have grouped them together
+to provide a more cohesive response to the point.
+
+\begin{shaded}
+\textbf{Reviewer 2 (R2) Comment 1:}
+
+Several of the comments added by the reviewer in the Word document focused on the
+purpose of calculating agreement among hazard-based metrics. These were:
+
+\begin{itemize}
+\item Line 164: Isn't it obvious to use better exposure measures than proxies?
+Wind, rainfall, tornado, and flood are direct products of storms while distance
+is only proxy of those exposures. Rather than calculating agreement among the
+exposures, it would be important to note that these are independent hazards that
+can come together or come with different intensity and at different time
+windows.
+\item Line 354: I am not sure why different exposures should be compared and
+measured the agreement. I would just treat each exposure as single independent
+exposure and give a summary score by summing all the exposure with some
+weighting for multiple exposures. I may not understand the importance of this
+approach. If so, adding the explanation would help readers not to questioning.
+\item Subsection of Results on \textit{Agreement across exposure metrics}: 
+As stated below, storms bring different hazards to different locations with
+different time. What would be new learning by comparing these different pairs of
+exposure metrics. Somewhat apple and orange comparison. Maybe providing more
+explanation can help what the important learning from this is. I may not catch
+it. Similar comment was added earlier.
+\item Lines 494--495: Why this agreement is important? Same questions added above
+already, but still wonder what I missed. It seems like the authors try to
+develop an exposure matrix that is common in occupational epidemiology, but it
+is only valid method when the exposures are similar and lead to the same health
+outcomes.  Rain, wind, flood, and tornado should be treated separately, maybe
+except rain and flood. It is possible to be an extreme storm with only severe
+wind without major impacts from other exposures. Tornado is a good example.
+\end{itemize}
+
+The text in the manuscript for comments with line numbers are reproduced below:
+
+\textit{Lines 164--166 from original manuscript:}
+
+\begin{quotation}
+\noindent
+\textit{"They found important differences, concluding that a study may be prone
+to bias from exposure misclassification if distance to the storm track is used
+as a proxy for exposure (Grabich et al. 2015a)."}
+\end{quotation}
+
+\textit{Line 354 from original manuscript:}
+
+\begin{quotation}
+\noindent
+\textit{"Finally, we investigated how well exposure assessment agreed across
+these metrics."}
+\end{quotation}
+
+\textit{Line 491--495 from original manuscript:}
+
+\begin{quotation}
+\noindent
+\textit{"For another set of tropical cyclones (e.g., Ernesto in 2006, and Bertha
+in 1996, Isabel in 2003), there was moderate to good agreement for pairwise
+combinations of distance, rain, and wind, but poor agreement for other
+combinations of metrics, while for another set of storms (e.g., Matthew in 2004
+and Katrina in 2005), there was moderate to good agreement between distance and
+rain."}
+\end{quotation}
+
+\end{shaded}
+
+**Response:**
 
@@
 
-We have moved other suggestions that you sent in "Track Changes" of the Word 
-document for the original manuscript and address each in detail below.
-
 The largest-scale county-level exposure study is likely that of @Zandbergen2009,
@@
 
-**Response:**
-
-\begin{shaded}
-\textbf{Reviewer 2 (R2) Comment 1:}
-Line 164: Isn't it obvious to use better exposure measures than proxies? Wind,
-rainfall, tornado, and flood are direct products of storms while distance is
-only proxy of those exposures. Rather than calculating agreement among the
-exposures, it would be important to note that these are independent hazards that
-can come together or come with different intensity and at different time
-windows.
-
-\textit{Lines 164--166 from original manuscript:}
-
-\begin{quotation}
-\noindent
-\textit{"They found important differences, concluding that a study may be prone
-to bias from exposure misclassification if distance to the storm track is used
-as a proxy for exposure (Grabich et al. 2015a)."}
-\end{quotation}
-\end{shaded}
-
-**Response:**
+---------------------------------------------------------------------------
 
@@
 
-\begin{shaded}
-\textbf{R2 Comment 2:}
-Line 354: I am not sure why different exposures should be compared and measured
-the agreement. I would just treat each exposure as single independent exposure
-and give a summary score by summing all the exposure with some weighting for
-multiple exposures.
-
-I may not understand the importance of this approach. If so, adding the
-explanation would help readers not to questioning.
-
-\textit{Line 354 from original manuscript:}
-
-\begin{quotation}
-\noindent
-\textit{"Finally, we investigated how well exposure assessment agreed across
-these metrics."}
-\end{quotation}
-\end{shaded}
-
-**Response:**
+-----------------------------------------------------------------------------
 
@@
 
-\begin{shaded}
-\textbf{R2 Comment 3:}
-Line 394: It looks like the correlation coefficients can be very low if measured
+-------------------------------------------------------------------------------
+
+> "A job exposure matrix (JEM) may be defined as a cross classification of jobs
+and occupational exposures. Some matrices are based on tasks, instead of jobs. 
+... Usually there is a group of experts who build the matrix based on their 
+own knowledge and experience on occupational exopsures. The experts should 
+assess exposure to an open or fixed list of occupational exposures (that can 
+include chemical, physical, biological, and/or psychosocial agents) for an
+open or fixed list of jobs (or tasks) usually coded according to some established
+national or international classification system. Some JEMs also include 
+information for different calendar time periods." [@garcia2003glossary]
+
+[New advances in multi-pollutant analysis.]
+
+[Some of the hazards could create similar risks, if they converge on a common
+pathway of risk. For example, if pathway is through stress from property damage,
+individually and in the county, then several hazards could indeed lead to the
+same health outcomes (e.g., PTSD, other psychological markers, risk of acute
+cardiovascular outcomes, etc.). One example is extreme level of destruction both
+from wind (Hurricane Andrew, Hurricane Hugo), from river flooding associated
+with among of precipitation (Hurricane Floyd, Hurricane Florence), and from
+flash flooding associated with the rate of precipitation (Virginia storms).]
+
+[A better future direction might be through regression models with multiple
+explanatory pathways. This allows the model to fit different weights to each.
+Other models can also be explored, like regression trees, which might have an
+easier time capturing non-linearity (if exposure is included as a continuous
+value, as with wind and rain) and interactions when there are co-exposures to
+two or more hazards. Both of these are able (and will) either exclude or give a
+very low weight to an explanatory variable that does not help much in explaining
+variation in the health outcome.]
+
+\begin{shaded}
+\textbf{R2 Comment 2:}
+
+Other comments were related to correlations estimated in the validation 
+analysis for the precipitation data:
+
+\begin{itemize}
+\item Line 394: It looks like the correlation coefficients can be very low if measured
 from 75mm or somewhere around.  This seems to be a significant issue for
 analyzing storms with heavier rainfall using NLDAS-2 data.
+\item Line 515: It would be informative to provide correlation coefficient for only
+above 75 mm measures.
+\end{itemize}
+
+The text from the referenced lines are reproduced below:
 
@@
 right of each graph in Figure 2) between 0.87 and 0.98. "}
+\end{quotation}
+
+\textit{Line 514--515 from original manuscript:}
+
+\begin{quotation}
+\noindent
+\textit{"The rainfall data are generally well-correlated with ground-based
+observations, but may sometimes underestimate very high rainfall values (Figure
+2)."}
 \end{quotation}
@@
 
-\begin{shaded}
-\textbf{R2 Comment 4:}
+---------------------------------------------------------------------------
+
+This is a great suggestion and helps provided quantitative support to the text
+in the paper that the reviewer highlighted when making this comment. We have
+added a table in the Supplemental Material that provides the Spearman
+correlation coefficients for both all tropical cyclones to come near the sample
+counties (replicating the analysis shown in Figure 2 of the main text) and
+then when the analysis is limited to only those events with cumulative
+precipitation of 75 mm or higher based on either data source.
+
+In several cases, the correlation was much weaker when limiting analysis to
+these severe rainfall events. For example, in Miami-Dade County, FL, the
+Spearman correlation coefficient was 0.94 for all 64 tropical cyclones that came
+near the county over the analysis period, while it was only 0.49 among the 18
+storms with cumulative precipitation of 75 mm or more. There was a similar
+reduction in Mobile County, AL. However, for some counties, the correlation
+remained high (e.g., Harris County, TX; Orleans Parish, LA), although in all
+cases there are small numbers of extreme precipitation events (5--20 across
+these sample counties), and so there is likely substantial random variation in
+these correlation estimates, since they are based on small numbers of 
+observations.
+
+[Include copy of table here.]
+
+Finally, we have added a bit to the text that the reviewer cited to mention 
+this point and to direct readers to the added supplemental table:
+
+> Text in revised manuscript (relevant addition in bold):
+"The rainfall data are generally well-correlated with ground-based
+observations, but may sometimes underestimate very high rainfall values (Figure
+2), **and in some counties the correlation was substantially lower when 
+considering only tropical cyclones with cumulative local rainfall of 75 mm
+or more (Supplemental Table 1)**."
+
+\begin{shaded}
+\textbf{R2 Comment 3:}
 Figure 4: I would add Y-axis label "NOAA classification."
@@
 \begin{shaded}
-\textbf{R2 Comment 5:}
+\textbf{R2 Comment 4:}
 Table 2: This is another confusing table. Not sure what compared here. The
@@
 \begin{shaded}
-\textbf{R2 Comment 6:}
-Subsection of Results on \textit{Agreement across exposure metrics}: 
-As stated below, storms bring different hazards to different locations with
-different time. What would be new learning by comparing these different pairs of
-exposure metrics. Somewhat apple and orange comparison. Maybe providing more
-explanation can help what the important learning from this is. I may not catch
-it. Similar comment was added earlier.
-\end{shaded}
-
-**Response:**
-
-\begin{shaded}
-\textbf{R2 Comment 7:}
-Lines 494--495: Why this agreement is important? Same questions added above
-already, but still wonder what I missed. It seems like the authors try to
-develop an exposure matrix that is common in occupational epidemiology, but it
-is only valid method when the exposures are similar and lead to the same health
-outcomes.  Rain, wind, flood, and tornado should be treated separately, maybe
-except rain and flood. It is possible to be an extreme storm with only severe
-wind without major impacts from other exposures. Tornado is a good example.
-
-\textit{Line 491--495 from original manuscript:}
-
-\begin{quotation}
-\noindent
-\textit{"For another set of tropical cyclones (e.g., Ernesto in 2006, and Bertha
-in 1996, Isabel in 2003), there was moderate to good agreement for pairwise
-combinations of distance, rain, and wind, but poor agreement for other
-combinations of metrics, while for another set of storms (e.g., Matthew in 2004
-and Katrina in 2005), there was moderate to good agreement between distance and
-rain."}
-\end{quotation}
-\end{shaded}
-
-**Response:**
-
-> "A job exposure matrix (JEM) may be defined as a cross classification of jobs
-and occupational exposures. Some matrices are based on tasks, instead of jobs. 
-... Usually there is a group of experts who build the matrix based on their 
-own knowledge and experience on occupational exopsures. The experts should 
-assess exposure to an open or fixed list of occupational exposures (that can 
-include chemical, physical, biological, and/or psychosocial agents) for an
-open or fixed list of jobs (or tasks) usually coded according to some established
-national or international classification system. Some JEMs also include 
-information for different calendar time periods." [@garcia2003glossary]
-
-[New advances in multi-pollutant analysis.]
-
-[Some of the hazards could create similar risks, if they converge on a common
-pathway of risk. For example, if pathway is through stress from property damage,
-individually and in the county, then several hazards could indeed lead to the
-same health outcomes (e.g., PTSD, other psychological markers, risk of acute
-cardiovascular outcomes, etc.). One example is extreme level of destruction both
-from wind (Hurricane Andrew, Hurricane Hugo), from river flooding associated
-with among of precipitation (Hurricane Floyd, Hurricane Florence), and from
-flash flooding associated with the rate of precipitation (Virginia storms).]
-
-[A better future direction might be through regression models with multiple
-explanatory pathways. This allows the model to fit different weights to each.
-Other models can also be explored, like regression trees, which might have an
-easier time capturing non-linearity (if exposure is included as a continuous
-value, as with wind and rain) and interactions when there are co-exposures to
-two or more hazards. Both of these are able (and will) either exclude or give a
-very low weight to an explanatory variable that does not help much in explaining
-variation in the health outcome.]
-
-\begin{shaded}
-\textbf{R2 Comment 8:}
-Line 515: It would be informative to provide correlation coefficient for only
-above 75 mm measures.
-
-\textit{Line 514--515 from original manuscript:}
-
-\begin{quotation}
-\noindent
-\textit{"The rainfall data are generally well-correlated with ground-based
-observations, but may sometimes underestimate very high rainfall values (Figure
-2)."}
-\end{quotation}
-\end{shaded}
-
-**Response:** 
-This is a great suggestion and helps provided quantitative support to the text
-in the paper that the reviewer highlighted when making this comment. We have
-added a table in the Supplemental Material that provides the Spearman
-correlation coefficients for both all tropical cyclones to come near the sample
-counties (replicating the analysis shown in Figure 2 of the main text) and
-then when the analysis is limited to only those events with cumulative
-precipitation of 75 mm or higher based on either data source.
-
-In several cases, the correlation was much weaker when limiting analysis to
-these severe rainfall events. For example, in Miami-Dade County, FL, the
-Spearman correlation coefficient was 0.94 for all 64 tropical cyclones that came
-near the county over the analysis period, while it was only 0.49 among the 18
-storms with cumulative precipitation of 75 mm or more. There was a similar
-reduction in Mobile County, AL. However, for some counties, the correlation
-remained high (e.g., Harris County, TX; Orleans Parish, LA), although in all
-cases there are small numbers of extreme precipitation events (5--20 across
-these sample counties), and so there is likely substantial random variation in
-these correlation estimates, since they are based on small numbers of 
-observations.
-
-[Include copy of table here.]
-
-Finally, we have added a bit to the text that the reviewer cited to mention 
-this point and to direct readers to the added supplemental table:
-
-> Text in revised manuscript (relevant addition in bold):
-"The rainfall data are generally well-correlated with ground-based
-observations, but may sometimes underestimate very high rainfall values (Figure
-2), **and in some counties the correlation was substantially lower when 
-considering only tropical cyclones with cumulative local rainfall of 75 mm
-or more (Supplemental Table 1)**."
-
-\begin{shaded}
-\textbf{R2 Comment 9:}
+\textbf{R2 Comment 5:}
 Line 602: This is not a new finding, but how much misclassified would be a good
@@
 \begin{shaded}
-\textbf{R2 Comment 10:}
+\textbf{R2 Comment 6:}
 Lines 626--627: It would still be beneficial to treat each exposure as a single
@@
 **Response:**
+
+Thank you for the helpful suggestions and recommendations. We have responded in 
+detail to each comment or question below.
 
@@
 Atmospheric Administration (NOAA)’s Storm Event Database (NOAA NCEI 2020).
-**Events are listed in this database based on reports from National Weather
+**Events are included in this database based on reports from National Weather
 Service personnel and other sources.** While this database has recorded storm
@@
 
-This database is NOAA's database of records on tornado events in the United
-States. It is the original source for the only other national tornado database
-available for the country (the Storm Prediction Center [SPC]'s National Tornado
-Database), which provides a version of the same data in which tornado events are
-joined into a single record in cases where they cross county lines. There is no
-other comparable source of data on tornado occurrences that could be used for a
-comparison as validation.
+This database is NOAA's official database on tornado events in the United States
+[@mccarthy2003nws]. It is the original source for the only other national
+tornado database available for the country (the Storm Prediction Center [SPC]'s
+National Tornado Database [@center2020storm]), which provides a version of the
+same data in which tornado events are joined into a single record in cases where
+they cross county lines. There is no other comparable source of data on tornado
+occurrences that could be used for a comparison as validation.
 
@@
 
-We have changed the text in the manuscript to make this clearer:
-
-> "The tornado observations from this dataset---**along with the derived version
+We have changed the text in the manuscript to make this clearer, changing the
+wording to clarify that this is the official NOAA data on tornado events in the
+US and that there are no other data sources of comparable scope that could be
+used for validation:
+
+> **Text added to the revised manuscript (relevant additions in bold):**
+"The tornado observations from this dataset---**along with the derived version
 of the data available through the Storm Prediction Center's National Tornado
-Database---are the tornado event data of record** for the US. **There are no other
-collections of tornado data comparable in temporal or geographical scope,** and so
-we did not further validate the tornado event data."
-
-Examples of research papers that leverage this database as a record of tornado
-events, or the version of it shared by the [Severe Storm ...] through their
-website [ref], include ... .
-
-[Is this database used to validate tornado forecasts / warnings?]
-
-This database has evolved through several stages but has, since its earliest
-iteration in the 1950s, served as the National Weather Service's [?] data of
-record on tornado events. While there is also a database maintained by the 
-[Severe Storm ...], the two databases have different names but almost 
-identical information once data are aggregated to a daily, county-level 
-measure of whether there was a tornado recorded in that county on that date. 
-[Difference between the two and why there are two.]
-
-There is no other tornado base at the same temporal and spatial scale that
-we could use for validation.
-
-This database serves as the basis for much of the large-scale (e.g., assessing
-patterns over the US over long time periods) tornado-based research for the US. 
-It does have some well-documented limitations. One is a notable increase over 
-time in the number of tornadoes reported, particularly for the weakest, F0 [?]
-tornadoes. [This stabilizes some in the age of radar?]
-
-> "Weather events listed in the publication Storm Data are considered to be
-the official database for tornadoes by the National Weather Service." 
-[written by someone at the NWS Storm Prediction Center] [@mccarthy2003nws]
-
-The US National Weather Service's Storm Prediction Center has a tornado database
-called the "National Tornado Database" [@center2020storm]. This uses the NOAA
-Storm Events database as its primary source [@center2020storm]. They note on their
-website:
-
-> "The tables below provide the links to comma separated values (.csv) files for
-tornadoes, hail, and damaging winds, as compiled from *NWS Storm Data*. Tornado 
-reports exist back to 1950 while hail and damaging wind events date from 1955.
-
-The NOAA Storm Events database, or the Storm Prediction Center's version of the
-tornado database have been used in a number of studies, including in a number of
-studies of tornado climatology [@tippett2016more; @brooks2003climatological;
-@strader2015climatology].
-
-> "The dataset we will use is the so-called smooth log of severe-weather reports
-collected by the SPC and archived in the National Oceanic and Atmospheric 
-Administration publication *Storm Data*." [@brooks2003climatological]
-
-> "Data such as tornado magnitude, path length, path width, and so on are
-recorded by the National Climatic Data Center (NCDC) as a public service and for
-use in meteorological, climatological, and engineering studies (Edwards et al.
-2013)." [@strader2015climatology]
-
-Sometimes used in conjunction with other NWS data products:
-
-> "Initially, tornado cases available in the NWS's Damage Assessment Toolkit
-(NOAA, 2014) were employed, which is a GIS-based framework for collecting,
-storing, and retrieving damage survey data (Camp et al., 2014). This toolkit
-provides a variety of tornado event filtering (tornado survey point, track,
-footprint, and/or swath) and download options [key mark-up language (.kml) or
-shapefile (.shp)], supplying an initial sample of georeferenced damage
-assessments. In this study, a tornado footprint is defined as the maximum areal
-extent of tornado intensity inferred by the damage, wind speed measured directly
-by mobile Doppler radar, or assessed theoretically as the recorded length
-multiplied with the maximum width as reported in Storm Data."
-[@strader2015climatology]
-
-> "To serve the public interest and the National Climatic Data Center (NCDC)
-storm data record, the National Weather Service (NWS) documents the path length,
-width, and maximum damage rating for every tornado county segment (NOAA
-2007).[Footnote in manuscript: County segments of tornado paths are then
-combined at the NWS Storm Prediction Center to yield a unified one-tornado
-(ONETOR) dataset of whole-tornado records (Schaefer and Edwards 1999).] Such
-data are used in meteorological and climatological research, as well as in
-determining construction standard for critical infrastructure such as
-high-tension electric lines and nuclear power plants (e.g., Ramsdell et al.
-2007)." [@edwards2013tornado]
-
-Sometimes used in conjunction with other datasets:
-
-> "The contiguous U.S. tornado fatality dataset utilized in this study was 
-transcribed and compiled from two primary sources: (1) a long-term study of 
-U.S. tornados by Grazulis (1993, 1997, hereafter Grazulis dataset) and 
-(2) the National Climatic Data Center's Storm Data (NCDC 1959--2005) and 
-'Storm Events' datavase. The Grazulis dataset included 'significant' tornado
-events, including all events that are known to have produced a fatality, 
-from 1680 to 1995. Storm Data reports from 1959 to 2005 were utilized to 
-supplement the existing record of killer tornado events documented by 
-Frazulis." [in a study of tornado fatalities] [@Ashley2007]
-
-History:
-
-> "During the early 1950s, the US Weather Bureau began a concerted effort to
-count all US tornadoes. In 1950, a list of tornadoes was provided in the
-*Climatological Data National Summary*, but it was not until 1953, the first
-full year of the issuance of Weather Bureau tornado watches, that the agency
-began to formally count all tornadoes. The *Climaticalogical Data National
-Summary* document evolved into *Storm Data* by 1959, which, in addition to
-tornado data, included information on other storm perils, such as severe hail,
-high winds, and floods (Grazulis 1993). Over the years, the tornado dataset has
-been formatted and adjusted by the National Severe Storms Forecast Center and
-Storm Prediction Center (SPC), and currently the National Tornado Database is
-administered by the NWS Headquarters, SPC, and NCDC (McCarthy 2003)."
-[@Ashley2007]
+Database [@center2020storm]---are the official tornado event data** for the US.
+**There are no other collections of tornado data comparable in temporal or
+geographical scope,** and so we did not further validate the tornado event
+data."
 
@@
 the comparison is based on the wind radii dataset described in the text, as
-requested. The original and revised versions of this paper are reproduced in
-this response document as Figure \@ref(fig:fig3).
+requested. The original and revised versions of Figure 3 of the manuscript are
+reproduced in this response document as Figure \@ref(fig:fig3).
 
@@
 **Response:**
-
-[Typo here? I don't think there's a Figure S7.]
 

### 2020-09-02 ehp_revision/writing/ehp_methods.tex
Work some more on response

@@
 and Atmospheric Administration (NOAA)'s Storm Event
-Database~\parencite{stormevents}. \textbf{Events are listed in this database
+Database~\parencite{stormevents}. \textbf{Events are included in this database
 based on reports from National Weather Service personnel and other sources.}
@@
 
-The tornado observations from this dataset form a traditional tornado event
-database for the \ac{US}, and so we did not further validate
+The tornado observations from this dataset---\textbf{along with the derived
+version of the data available through the Storm Prediction Center's National
+Tornado Database~\parencite{center2020storm}---are the official tornado event
+data} for the \ac{US}. \textbf{There are no other collections of tornado data
+comparable in emporal or geographic scope,} and so we did not further validate
 the tornado event data. It is difficult to characterize flooding at the county
@@
 tropical storms with streamflow measurements at \ac{US} Geological Survey
-county streamflow gages \parencite{usgsgages, countyfloods, dataRetrieval}.  
-We considered nine study counties, selecting counties geographically spread
-through storm-prone areas of the eastern \ac{US} and with at least one streamflow
-gage reporting data during all events. The sample counties were: Baltimore County,
-MD; Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton
-County, GA; Harris County, TX; Mobile County, AL; Montgomery County, PA; and
-Wake County, NC.
+county streamflow gages \parencite{usgsgages, countyfloods, dataRetrieval}.  We
+considered nine study counties, selecting counties geographically spread
+through storm-prone areas of the eastern \ac{US} and with at least one
+streamflow gage reporting data during all events. The sample counties were:
+Baltimore County, MD; Bergen County, NJ; Escambia County, FL; Fairfield County,
+CT; Fulton County, GA; Harris County, TX; Mobile County, AL; Montgomery County,
+PA; and Wake County, NC.
 

### 2020-09-02 ehp_revision/writing/ehp_response_darlings.tex
Work some more on response

@@
 \url{https://www.weather.gov/dvn/armistice_day_blizzard}
+
+R3 Comment 5
+
+This database serves as the basis for much of the large-scale (e.g., assessing
+patterns over the US over long time periods) tornado-based research for the US. 
+It does have some well-documented limitations. One is a notable increase over 
+time in the number of tornadoes reported, particularly for the weakest, F0 [?]
+tornadoes. [This stabilizes some in the age of radar?]
+
+> "Weather events listed in the publication Storm Data are considered to be
+the official database for tornadoes by the National Weather Service." 
+[written by someone at the NWS Storm Prediction Center] [@mccarthy2003nws]
+
+The US National Weather Service's Storm Prediction Center has a tornado database
+called the "National Tornado Database" [@center2020storm]. This uses the NOAA
+Storm Events database as its primary source [@center2020storm]. They note on their
+website:
+
+> "The tables below provide the links to comma separated values (.csv) files for
+tornadoes, hail, and damaging winds, as compiled from *NWS Storm Data*. Tornado 
+reports exist back to 1950 while hail and damaging wind events date from 1955.
+
+The NOAA Storm Events database, or the Storm Prediction Center's version of the
+tornado database have been used in a number of studies, including in a number of
+studies of tornado climatology [@tippett2016more; @brooks2003climatological;
+@strader2015climatology].
+
+> "The dataset we will use is the so-called smooth log of severe-weather reports
+collected by the SPC and archived in the National Oceanic and Atmospheric 
+Administration publication *Storm Data*." [@brooks2003climatological]
+
+> "Data such as tornado magnitude, path length, path width, and so on are
+recorded by the National Climatic Data Center (NCDC) as a public service and for
+use in meteorological, climatological, and engineering studies (Edwards et al.
+2013)." [@strader2015climatology]
+
+Sometimes used in conjunction with other NWS data products:
+
+> "Initially, tornado cases available in the NWS's Damage Assessment Toolkit
+(NOAA, 2014) were employed, which is a GIS-based framework for collecting,
+storing, and retrieving damage survey data (Camp et al., 2014). This toolkit
+provides a variety of tornado event filtering (tornado survey point, track,
+footprint, and/or swath) and download options [key mark-up language (.kml) or
+shapefile (.shp)], supplying an initial sample of georeferenced damage
+assessments. In this study, a tornado footprint is defined as the maximum areal
+extent of tornado intensity inferred by the damage, wind speed measured directly
+by mobile Doppler radar, or assessed theoretically as the recorded length
+multiplied with the maximum width as reported in Storm Data."
+[@strader2015climatology]
+
+> "To serve the public interest and the National Climatic Data Center (NCDC)
+storm data record, the National Weather Service (NWS) documents the path length,
+width, and maximum damage rating for every tornado county segment (NOAA
+2007).[Footnote in manuscript: County segments of tornado paths are then
+combined at the NWS Storm Prediction Center to yield a unified one-tornado
+(ONETOR) dataset of whole-tornado records (Schaefer and Edwards 1999).] Such
+data are used in meteorological and climatological research, as well as in
+determining construction standard for critical infrastructure such as
+high-tension electric lines and nuclear power plants (e.g., Ramsdell et al.
+2007)." [@edwards2013tornado]
+
+Sometimes used in conjunction with other datasets:
+
+> "The contiguous U.S. tornado fatality dataset utilized in this study was 
+transcribed and compiled from two primary sources: (1) a long-term study of 
+U.S. tornados by Grazulis (1993, 1997, hereafter Grazulis dataset) and 
+(2) the National Climatic Data Center's Storm Data (NCDC 1959--2005) and 
+'Storm Events' datavase. The Grazulis dataset included 'significant' tornado
+events, including all events that are known to have produced a fatality, 
+from 1680 to 1995. Storm Data reports from 1959 to 2005 were utilized to 
+supplement the existing record of killer tornado events documented by 
+Frazulis." [in a study of tornado fatalities] [@Ashley2007]
+
+History:
+
+> "During the early 1950s, the US Weather Bureau began a concerted effort to
+count all US tornadoes. In 1950, a list of tornadoes was provided in the
+*Climatological Data National Summary*, but it was not until 1953, the first
+full year of the issuance of Weather Bureau tornado watches, that the agency
+began to formally count all tornadoes. The *Climaticalogical Data National
+Summary* document evolved into *Storm Data* by 1959, which, in addition to
+tornado data, included information on other storm perils, such as severe hail,
+high winds, and floods (Grazulis 1993). Over the years, the tornado dataset has
+been formatted and adjusted by the National Severe Storms Forecast Center and
+Storm Prediction Center (SPC), and currently the National Tornado Database is
+administered by the NWS Headquarters, SPC, and NCDC (McCarthy 2003)."
+[@Ashley2007]

### 2020-09-03 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 societal impacts of tropical cyclones that have used FEMA disaster declarations
-to assess county-level exposure to tropical cyclones [refs].
-
-However, evidence from prior tropical cyclone epidemiology that doing so is
-problematic and should be avoided [@grabich2015measuring]. FEMA's disaster
-declarations are issued to provide assistance, based on damage assessment, to
-individuals or the entire public in certain locations where catastrophes
-overwhelm the state or local government [@mccarthy2014fema], and any use of them
-for exposure assessment is secondary. Due to the political nature of disaster
-declaration process, these declarations are often subject to many political and
-economic factors [@mccarthy2014fema; @logue1981research]. One epidemiological
-study specifically focused on exposure assessment for tropical cyclone
-epidemiology and compared use of FEMA declarations with other methods of
+to assess county-level exposure to tropical cyclones.
+
+There are two elements that are important related to this question. First, many
+of the health impacts of tropical cyclones will likely come through indirect
+pathways, including through damage to property and infrastructure. Ultimately,
+it would be useful to have large-scale databases of these secondary exposure
+pathways to use for more complex studies of the pathways between the physical
+hazards of these storms and how they affect human health. In this sense,
+something like the FEMA declarations could be considered not as a proxy for
+assessing exposure to physical hazards, but rather as a proxy of damage caused
+by the storm, and so a measurement informative for one of these indirect
+pathways from physical hazards of the storm to health impacts, through something
+like a mediation analysis. We have added text to the manuscript on this point:
+
+...
+
+In terms of using FEMA declarations as a source for this, however, there are 
+some limitations. These declarations are on average tied to the amount of 
+damage and other impacts experienced following a storm, but there are a number
+of other considerations that drive whether a disaster is declared or not. 
+These contributing factors could vary by location (especially from one state
+to another) and also over time, especially over the long period we've included
+for our other data. Given that, these may not be the best data source to use
+to try to capture estimates of the damage or impacts to a community in a multi-year, 
+multi-site dataset like the one we present.
+
+FEMA's disaster declarations are issued to provide assistance, based on damage
+assessment, to individuals or the entire public in certain locations where
+catastrophes overwhelm the state or local government [@mccarthy2014fema], and
+any use of them for exposure assessment is secondary. Due to the political
+nature of disaster declaration process, these declarations are often subject to
+many political and economic factors [@mccarthy2014fema; @logue1981research]. One
+epidemiological study specifically focused on exposure assessment for tropical
+cyclone epidemiology and compared use of FEMA declarations with other methods of
 exposure assessment, including one based directly on the storm hazard of wind
@@
 explore risk across multiple years and affected communities.
+
+Other sources of data may be better suited to use as a marker of damage to
+assess that pathway of indirect health impacts from a tropical cyclone. For
+example, data on storm impacts, including property damage or insurance claims
+data, are sometimes available through storm databases or publications (e.g.,
+NOAA's Storm Events data, *Monthly Weather Review*). While such data could
+potentially be used to create an exposure metric, tropical storm damage data
+requires some normalization, to incorporate both changes in dollar value over
+time and also changes in development in at-risk areas, to be comparable over
+extended time periods [@Pielke1998].
+
+---------------------------------------------------------------------------
+
+However, evidence from prior tropical cyclone epidemiology that doing so is
+problematic and should be avoided [@grabich2015measuring].
 
@@
 
-It turns out that, because we have shared the data as an R package, it should
-already be accessible to use on the DesignSafe platform.
+It turns out that it should already be accessible to use on the DesignSafe
+platform, because we have shared the data as an R package.
 
@@
 
-This is a great idea. In the original manuscript, we describe this lower
+This is a great suggestion. In the original manuscript, we describe this lower
 correlation at higher precipitation values through the use of examples from
@@
 
-Here is the text where we discuss this point in the original manuscript:
-
-> **Text in original manuscript:**
-"There was some evidence that our primary rainfall metric may tend to
-underestimate rainfall totals in storms with extremely high rainfall, based on
-a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
-County, SC, and Wake County, NC (Figure 2)."
-
-> **Text in original manuscript:**
-"The rainfall data are generally well-correlated with ground-based observations,
-but may sometimes underestimate very high rainfall values (Figure 2). When
-rainfall data is used to create binary exposure classifications, this
-disagreement is unlikely to influence results, as both data sources agree in
-identifying these as storms with high rainfall, but would be important to
-consider for cases that include rainfall as a continuous measurement."
-
 We have added estimates of these correlations specifically for
@@
 \input{tables/precip_high_corr.tex}
-
----------------------------------------------------------------------------
-
-This is a great suggestion and helps provided quantitative support to the text
-in the paper that the reviewer highlighted when making this comment. We have
-added a table in the Supplemental Material that provides the Spearman
-correlation coefficients for both all tropical cyclones to come near the sample
-counties (replicating the analysis shown in Figure 2 of the main text) and
-then when the analysis is limited to only those events with cumulative
-precipitation of 75 mm or higher based on either data source.
 
@@
 
-[Include copy of table here.]
-
-Finally, we have added a bit to the text that the reviewer cited to mention 
-this point and to direct readers to the added supplemental table:
-
-> Text in revised manuscript (relevant addition in bold):
+We have added text to mention this point and to direct readers to the added
+supplemental table:
+
+> **Text in revised manuscript (relevant addition in bold):**
 "The rainfall data are generally well-correlated with ground-based
@@
 **Response:**
-[Add new analysis of what percent of counties misclassified if distance 
-is used as proxy for each hazard?]
+
+Figure 7 of the main text is meant to help illustrate the potential for 
+exposure misclassification, including when using a distance-based metric as
+a proxy. We selected to use the Jaccard index because we are using data for 
+counties at a large spatial scale. For any storm, there will be large parts of
+our study area that are clearly unaffected, because they are so far from the 
+storm. If we measured misclassification using a metric that includes these
+counties in the denominator, it would be overly optimistic, and wouldn't help
+to pinpoint the potential for disagreement in areas nearer the storm path.
+
+However, one downside of this metric is that it does not help readers understand
+the direction that disagreement is going. For example, it does not provide 
+information about how many counties were classified as "exposed" based on a 
+distance proxy but "unexposed" by a hazard-based proxy. This could be useful in 
+interpreting our results in Figure 7 of the main table.
+
+We have added four new tables in the Supplemental Material, one for each of the 
+hazard-based metrics we considered. We have reproduced these tables in this response
+document as Tables [x]--[x] Each breaks down agreement and disagreement
+between the distance-based proxy metric and that hazard-based metric within several
+storms in our study period, focusing on storms with at least 250 counties exposed
+by at least one of the metrics considered. All of these storms are included in 
+Figure 7.
+
+\input{tables/wind_misclass.tex}
+
+\input{tables/rain_misclass.tex}
+
+\input{tables/flood_misclass.tex}
+
+\input{tables/tornado_misclass.tex}
+
+These tables allow readers both to investigate general patterns in disagreement
+and exposure misclassification across these storms and also to see cases of unusual 
+storms. For example, for most storms, disagreement between the distance proxy 
+metric and the wind-based metric resulted mainly from mis-classification of 
+counties as exposed based on distance while they did not experience gale-force winds.
+However, there were a few storms, like Hurricane Bertha in 1996, where the 
+inverse was the case, and a distance-based metric would have missed a large number
+of counties that experienced gale-force winds from the storm (misclassified them
+as unexposed), but not misclassified many counties as exposed when they in fact
+did not experience winds at this level.
+
+We have added a reference to these new tables in the main text of the manuscript:
+
+> **Text in the revised manuscript (relevant addition in bold):** 
+"We drew similar conclusions when we investigated all 46 tropical cyclones
+between 1996 and 2011 (when data for all five metrics were available) for which
+100 or more counties were exposed based on at least one metric (Figure 7;
+**Tables S[x]--S[x] provide the underlying numbers for the most extensive of
+these, the storms for which 250 of more counties were exposed based on at least
+one metric)**."
+
+> **Text in the revised manuscript (relevant addition in bold):**
+"Based on our results, the use of a distance-based metric to assess exposure to
+any of these hazards, or the use of measurements from one hazard as a proxy for
+exposure to any of the other hazards considered, would often introduce exposure
+misclassification **(Tables S[x]--S[x])**."
 
@@
 > **Text in revised manuscript (relevant addition in bold):**
-"At each 15- interval, we measured the distance between the storm's
+"At each 15-min interval, we measured the distance between the storm's
 center and each county's population mean center, **as of the 2010 US Decennial

### 2020-09-03 ehp_revision/tables/flood_misclass.tex
Work some more on response

% latex table generated in R 4.0.2 by xtable 1.8-4 package
% Thu Sep  3 12:19:54 2020
\begin{table}[ht]
\centering
\caption{Caption. Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns). Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.} 
\begin{tabular}{cp{3cm}p{3cm}p{3cm}p{3cm}}
  \toprule
Storm & Exposed for both distance metric and flood metric & Exposed for distance metric but unexposed for flood metric & Exposed for flood metric but unexposed for distance metric & Unexposed for both distance metric and flood metric \\ 
  \midrule
Ivan (2004) & 79 & 205 & 238 & 1,874 \\ 
  Frances (2004) & 103 & 218 & 122 & 1,953 \\ 
  Dennis (2005) & 21 & 271 & 87 & 2,017 \\ 
  Jeanne (2004) & 80 & 217 & 77 & 2,022 \\ 
  Allison (2001) & 68 & 219 & 71 & 2,038 \\ 
  Cindy (2005) & 52 & 269 & 29 & 2,046 \\ 
  Ike (2008) & 48 & 185 & 101 & 2,062 \\ 
  Isidore (2002) & 30 & 238 & 64 & 2,064 \\ 
  Fay (2008) & 23 & 229 & 60 & 2,084 \\ 
  Bill (2003) & 39 & 191 & 65 & 2,101 \\ 
  Gustav (2008) & 36 & 198 & 57 & 2,105 \\ 
  Katrina (2005) & 47 & 194 & 47 & 2,108 \\ 
  Gordon (2000) & 7 & 248 & 19 & 2,122 \\ 
  Arlene (2005) & 11 & 225 & 29 & 2,131 \\ 
  Danny (1997) & 39 & 190 & 31 & 2,136 \\ 
  Floyd (1999) & 100 & 39 & 120 & 2,137 \\ 
  Fran (1996) & 46 & 105 & 77 & 2,168 \\ 
  Dennis (1999) & 38 & 162 & 19 & 2,177 \\ 
  Ernesto (2006) & 44 & 148 & 27 & 2,177 \\ 
  Bertha (1996) & 27 & 156 & 17 & 2,196 \\ 
  Rita (2005) & 4 & 161 & 26 & 2,205 \\ 
  Lee (2011) & 18 & 126 & 36 & 2,216 \\ 
   \bottomrule
\end{tabular}
\end{table}

### 2020-09-03 ehp_revision/tables/rain_misclass.tex
Work some more on response

@@
 % latex table generated in R 4.0.2 by xtable 1.8-4 package
-% Tue Sep  1 17:11:55 2020
+% Thu Sep  3 12:19:54 2020
 \begin{table}[ht]
@@
 \caption{Caption.
-Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
+Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study.
 Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
@@
   \toprule
-Storm & Classified as exposed based on both distance metric and rain metric & Classified as exposed based on distance metric but unexposed by rain metric & Classified as exposed based on rain metric but unexposed by distance metric & Classified as unexposed based on both distance metric and rain metric \\
+Storm & Exposed for both distance metric and rain metric & Exposed for distance metric but unexposed for rain metric & Exposed for rain metric but unexposed for distance metric & Unexposed for both distance metric and rain metric \\
   \midrule
-Frances-2004 & 207 & 114 & 257 & 1818 \\ 
-  Ivan-2004 & 124 & 160 & 288 & 1824 \\ 
-  Isidore-2002 & 134 & 134 & 186 & 1942 \\ 
-  Ike-2008 & 113 & 120 & 213 & 1950 \\ 
-  Dennis-2005 &  48 & 244 & 124 & 1980 \\ 
-  Fay-2008 &  72 & 180 & 163 & 1981 \\ 
-  Jeanne-2004 & 144 & 153 & 113 & 1986 \\ 
-  Lee-2011 & 106 &  38 & 230 & 2022 \\ 
-  Gustav-2008 & 126 & 108 & 137 & 2025 \\ 
-  Allison-2001 & 106 & 181 &  67 & 2042 \\ 
-  Bill-2003 & 110 & 120 & 117 & 2049 \\ 
-  Cindy-2005 &  78 & 243 &  11 & 2064 \\ 
-  Floyd-1999 & 129 &  10 & 170 & 2087 \\ 
-  Katrina-2005 & 195 &  46 &  58 & 2097 \\ 
-  Danny-1997 &  82 & 147 &  54 & 2113 \\ 
-  Rita-2005 &  64 & 101 & 103 & 2128 \\ 
-  Gordon-2000 &   5 & 250 &   7 & 2134 \\ 
-  Arlene-2005 &  39 & 197 &  23 & 2137 \\ 
-  Dennis-1999 &  90 & 110 &  56 & 2140 \\ 
-  Ernesto-2006 & 124 &  68 &  63 & 2141 \\ 
-  Irene-2011 & 119 &   1 & 106 & 2170 \\ 
-  Hanna-2008 & 100 &  92 &  33 & 2171 \\ 
-  Hermine-2010 &  44 &  79 & 102 & 2171 \\ 
-  Fran-1996 &  89 &  62 &  73 & 2172 \\ 
-  Earl-1998 &  72 &  97 &  38 & 2189 \\ 
-  Bertha-1996 & 103 &  80 &  20 & 2193 \\ 
-  Helene-2000 &  28 & 164 &  11 & 2193 \\ 
-  Frances-1998 &  27 &  35 & 139 & 2195 \\ 
-  Josephine-1996 &  94 &  29 &  77 & 2196 \\ 
-  Bonnie-2004 &  56 &  53 &  56 & 2231 \\ 
-  Isabel-2003 &  88 &  59 &  17 & 2232 \\ 
+Frances (2004) & 207 & 114 & 257 & 1,818 \\ 
+  Ivan (2004) & 124 & 160 & 288 & 1,824 \\ 
+  Isidore (2002) & 134 & 134 & 186 & 1,942 \\ 
+  Ike (2008) & 113 & 120 & 213 & 1,950 \\ 
+  Dennis (2005) & 48 & 244 & 124 & 1,980 \\ 
+  Fay (2008) & 72 & 180 & 163 & 1,981 \\ 
+  Jeanne (2004) & 144 & 153 & 113 & 1,986 \\ 
+  Lee (2011) & 106 & 38 & 230 & 2,022 \\ 
+  Gustav (2008) & 126 & 108 & 137 & 2,025 \\ 
+  Allison (2001) & 106 & 181 & 67 & 2,042 \\ 
+  Bill (2003) & 110 & 120 & 117 & 2,049 \\ 
+  Cindy (2005) & 78 & 243 & 11 & 2,064 \\ 
+  Floyd (1999) & 129 & 10 & 170 & 2,087 \\ 
+  Katrina (2005) & 195 & 46 & 58 & 2,097 \\ 
+  Danny (1997) & 82 & 147 & 54 & 2,113 \\ 
+  Rita (2005) & 64 & 101 & 103 & 2,128 \\ 
+  Gordon (2000) & 5 & 250 & 7 & 2,134 \\ 
+  Arlene (2005) & 39 & 197 & 23 & 2,137 \\ 
+  Dennis (1999) & 90 & 110 & 56 & 2,140 \\ 
+  Ernesto (2006) & 124 & 68 & 63 & 2,141 \\ 
+  Fran (1996) & 89 & 62 & 73 & 2,172 \\ 
+  Bertha (1996) & 103 & 80 & 20 & 2,193 \\ 
    \bottomrule

### 2020-09-03 ehp_revision/tables/tornado_misclass.tex
Work some more on response

--- 2020-09-03 ehp_revision/tables/flood_misclass.tex
@@
   \toprule
-Storm & Exposed for both distance metric and flood metric & Exposed for distance metric but unexposed for flood metric & Exposed for flood metric but unexposed for distance metric & Unexposed for both distance metric and flood metric \\
+Storm & Exposed for both distance metric and tornado metric & Exposed for distance metric but unexposed for tornado metric & Exposed for tornado metric but unexposed for distance metric & Unexposed for both distance metric and tornado metric \\
   \midrule
-Ivan (2004) & 79 & 205 & 238 & 1,874 \\ 
-  Frances (2004) & 103 & 218 & 122 & 1,953 \\ 
-  Dennis (2005) & 21 & 271 & 87 & 2,017 \\ 
-  Jeanne (2004) & 80 & 217 & 77 & 2,022 \\ 
-  Allison (2001) & 68 & 219 & 71 & 2,038 \\ 
-  Cindy (2005) & 52 & 269 & 29 & 2,046 \\ 
-  Ike (2008) & 48 & 185 & 101 & 2,062 \\ 
-  Isidore (2002) & 30 & 238 & 64 & 2,064 \\ 
-  Fay (2008) & 23 & 229 & 60 & 2,084 \\ 
-  Bill (2003) & 39 & 191 & 65 & 2,101 \\ 
-  Gustav (2008) & 36 & 198 & 57 & 2,105 \\ 
-  Katrina (2005) & 47 & 194 & 47 & 2,108 \\ 
-  Gordon (2000) & 7 & 248 & 19 & 2,122 \\ 
-  Arlene (2005) & 11 & 225 & 29 & 2,131 \\ 
-  Danny (1997) & 39 & 190 & 31 & 2,136 \\ 
-  Floyd (1999) & 100 & 39 & 120 & 2,137 \\ 
-  Fran (1996) & 46 & 105 & 77 & 2,168 \\ 
-  Dennis (1999) & 38 & 162 & 19 & 2,177 \\ 
-  Ernesto (2006) & 44 & 148 & 27 & 2,177 \\ 
-  Bertha (1996) & 27 & 156 & 17 & 2,196 \\ 
-  Rita (2005) & 4 & 161 & 26 & 2,205 \\ 
-  Lee (2011) & 18 & 126 & 36 & 2,216 \\ 
+Frances (2004) & 6 & 315 & 66 & 2,009 \\ 
+  Ivan (2004) & 36 & 248 & 55 & 2,057 \\ 
+  Cindy (2005) & 27 & 294 & 11 & 2,064 \\ 
+  Jeanne (2004) & 13 & 284 & 22 & 2,077 \\ 
+  Dennis (2005) & 0 & 292 & 6 & 2,098 \\ 
+  Allison (2001) & 14 & 273 & 9 & 2,100 \\ 
+  Katrina (2005) & 8 & 233 & 38 & 2,117 \\ 
+  Fay (2008) & 15 & 237 & 26 & 2,118 \\ 
+  Isidore (2002) & 1 & 267 & 6 & 2,122 \\ 
+  Ike (2008) & 4 & 229 & 31 & 2,132 \\ 
+  Gordon (2000) & 1 & 254 & 8 & 2,133 \\ 
+  Gustav (2008) & 9 & 225 & 28 & 2,134 \\ 
+  Bill (2003) & 4 & 226 & 22 & 2,144 \\ 
+  Arlene (2005) & 2 & 234 & 7 & 2,153 \\ 
+  Danny (1997) & 9 & 220 & 4 & 2,163 \\ 
+  Rita (2005) & 8 & 157 & 41 & 2,190 \\ 
+  Dennis (1999) & 0 & 200 & 1 & 2,195 \\ 
+  Ernesto (2006) & 2 & 190 & 1 & 2,203 \\ 
+  Bertha (1996) & 13 & 170 & 4 & 2,209 \\ 
+  Fran (1996) & 1 & 150 & 4 & 2,241 \\ 
+  Lee (2011) & 14 & 130 & 11 & 2,241 \\ 
+  Floyd (1999) & 11 & 128 & 1 & 2,256 \\ 
    \bottomrule

### 2020-09-03 ehp_revision/tables/wind_misclass.tex
Work some more on response

@@
 % latex table generated in R 4.0.2 by xtable 1.8-4 package
-% Tue Sep  1 17:11:57 2020
+% Thu Sep  3 12:19:54 2020
 \begin{table}[ht]
@@
   \toprule
-Storm & Classified as exposed based on both distance metric and wind metric & Classified as exposed based on distance metric but unexposed by wind metric & Classified as exposed based on wind metric but unexposed by distance metric & Classified as unexposed based on both distance metric and wind metric \\
+Storm & Exposed for both distance metric and wind metric & Exposed for distance metric but unexposed for wind metric & Exposed for wind metric but unexposed for distance metric & Unexposed for both distance metric and wind metric \\
   \midrule
-Frances-2004 &  44 & 277 &  13 & 2062 \\ 
-  Cindy-2005 &  17 & 304 &   4 & 2071 \\ 
-  Dennis-2005 &  18 & 274 &  26 & 2078 \\ 
-  Ike-2008 & 166 &  67 &  82 & 2081 \\ 
-  Ivan-2004 &  29 & 255 &  27 & 2085 \\ 
-  Jeanne-2004 &  41 & 256 &   7 & 2092 \\ 
-  Allison-2001 &  21 & 266 &   0 & 2109 \\ 
-  Isidore-2002 &  29 & 239 &   7 & 2121 \\ 
-  Katrina-2005 &  66 & 175 &  32 & 2123 \\ 
-  Gordon-2000 &  11 & 244 &   6 & 2135 \\ 
-  Fay-2008 &  54 & 198 &   7 & 2137 \\ 
-  Gustav-2008 &  36 & 198 &  23 & 2139 \\ 
-  Bertha-1996 & 179 &   4 &  62 & 2151 \\ 
-  Danny-1997 &  45 & 184 &  12 & 2155 \\ 
-  Arlene-2005 &   4 & 232 &   0 & 2160 \\ 
-  Bill-2003 &  19 & 211 &   0 & 2166 \\ 
-  Dennis-1999 &  17 & 183 &  13 & 2183 \\ 
-  Hanna-2008 & 135 &  57 &  12 & 2192 \\ 
-  Isabel-2003 & 105 &  42 &  56 & 2193 \\ 
-  Ernesto-2006 &  93 &  99 &   4 & 2200 \\ 
-  Helene-2000 &  18 & 174 &   0 & 2204 \\ 
-  Rita-2005 &  26 & 139 &  20 & 2211 \\ 
-  Fran-1996 &  46 & 105 &  30 & 2215 \\ 
-  Earl-1998 & 133 &  36 &  11 & 2216 \\ 
-  Floyd-1999 & 124 &  15 &  40 & 2217 \\ 
-  Irene-2011 & 114 &   6 &  46 & 2230 \\ 
-  Lee-2011 &  34 & 110 &   0 & 2252 \\ 
-  Josephine-1996 &  62 &  61 &   3 & 2270 \\ 
-  Hermine-2010 &  22 & 101 &   1 & 2272 \\ 
-  Bonnie-2004 &   0 & 109 &   0 & 2287 \\ 
-  Frances-1998 &  15 &  47 &   0 & 2334 \\ 
+Frances (2004) & 44 & 277 & 13 & 2,062 \\ 
+  Cindy (2005) & 17 & 304 & 4 & 2,071 \\ 
+  Dennis (2005) & 18 & 274 & 26 & 2,078 \\ 
+  Ike (2008) & 166 & 67 & 82 & 2,081 \\ 
+  Ivan (2004) & 29 & 255 & 27 & 2,085 \\ 
+  Jeanne (2004) & 41 & 256 & 7 & 2,092 \\ 
+  Allison (2001) & 21 & 266 & 0 & 2,109 \\ 
+  Isidore (2002) & 29 & 239 & 7 & 2,121 \\ 
+  Katrina (2005) & 66 & 175 & 32 & 2,123 \\ 
+  Gordon (2000) & 11 & 244 & 6 & 2,135 \\ 
+  Fay (2008) & 54 & 198 & 7 & 2,137 \\ 
+  Gustav (2008) & 36 & 198 & 23 & 2,139 \\ 
+  Bertha (1996) & 179 & 4 & 62 & 2,151 \\ 
+  Danny (1997) & 45 & 184 & 12 & 2,155 \\ 
+  Arlene (2005) & 4 & 232 & 0 & 2,160 \\ 
+  Bill (2003) & 19 & 211 & 0 & 2,166 \\ 
+  Dennis (1999) & 17 & 183 & 13 & 2,183 \\ 
+  Ernesto (2006) & 93 & 99 & 4 & 2,200 \\ 
+  Rita (2005) & 26 & 139 & 20 & 2,211 \\ 
+  Fran (1996) & 46 & 105 & 30 & 2,215 \\ 
+  Floyd (1999) & 124 & 15 & 40 & 2,217 \\ 
+  Lee (2011) & 34 & 110 & 0 & 2,252 \\ 
    \bottomrule

### 2020-09-03 ehp_revision/writing/ehp_discussion.tex
Work some more on response

@@
 
-Based on our results, the use of a distance-based metric to assess exposure to any of
-these hazards, or the use of measurements from one hazard as a proxy for
+Based on our results, the use of a distance-based metric to assess exposure to
+any of these hazards, or the use of measurements from one hazard as a proxy for
 exposure to any of the other hazards considered, would often introduce exposure
-misclassification. This conclusion reinforces similar findings from a study of
-Florida's 2004 storm season~\parencite{grabich2015measuring}.  For some
-studies, such exposure misclassification might plausibly be differential.  For
-example, tropical cyclone wind exposures tend to be concentrated in counties
-near the coast, while tropical cyclone rain exposures sometimes extend well
-inland.  If the etiologically-relevant exposure for a health outcome is extreme
-rainfall but exposure is classified based on measurements of wind, the
-probability of being misclassified as unexposed would be higher in inland
-counties, while the probability of being misclassified as exposed would be
-higher in coastal counties. If coastal counties differ from inland counties in
-either the outcome of interest or in factors associated with risk of that
-outcome, exposure misclassification would be
-differential~\parencite{savitz2016interpreting}.  Such differential exposure
-misclassification could bias estimates of tropical cyclone effects either
-towards the null (estimating a lower or null association compared to the true
-association that exists) or away from the null (estimating a larger association
-than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.
+misclassification \textbf{(Tables S[x]--S[x])}. This conclusion reinforces
+similar findings from a study of Florida's 2004 storm
+season~\parencite{grabich2015measuring}.  For some studies, such exposure
+misclassification might plausibly be differential.  For example, tropical
+cyclone wind exposures tend to be concentrated in counties near the coast,
+while tropical cyclone rain exposures sometimes extend well inland.  If the
+etiologically-relevant exposure for a health outcome is extreme rainfall but
+exposure is classified based on measurements of wind, the probability of being
+misclassified as unexposed would be higher in inland counties, while the
+probability of being misclassified as exposed would be higher in coastal
+counties. If coastal counties differ from inland counties in either the outcome
+of interest or in factors associated with risk of that outcome, exposure
+misclassification would be differential~\parencite{savitz2016interpreting}.
+Such differential exposure misclassification could bias estimates of tropical
+cyclone effects either towards the null (estimating a lower or null association
+compared to the true association that exists) or away from the null (estimating
+a larger association than actually exists)~\parencite{savitz2016interpreting,
+armstrong1998effect}.
 

### 2020-09-03 ehp_revision/writing/ehp_results.tex
Start some on response
Work some more on response

--- 2020-02-24 writing/ehp_results.tex
@@
 measure of county-level peak sustained wind in the open-source dataset
-\parencite{hurricaneexposuredate}.
+\parencite{hurricaneexposuredata}.
 
@@
 which~100 or more counties were exposed based on at least one metric
-(Figure~\ref{fig:jaccard}). In this figure, each row provides results for one
-tropical cyclone, and each box in that row shows the Jaccard coefficient for a
-pair of metrics. For all pairs of metrics, agreement in exposure assessment
-was, at best, moderate for all but a few storms. When comparing distance- and
-wind-based exposure assessment, only about 10\% of storms had Jaccard indices
-higher than 0.6 (i.e., out of the counties assessed as exposed by at least one
-of the two metrics in the pair,~60\si{\percent} or more were assessed the same
-under both metrics). For comparisons of assessments based on other combinations
-of distance-, wind-, rain-, and flood-based metrics, fewer than 5\% of storms
-had Jaccard indices above 0.6.  The tornado-based metric had universally poor
-agreement with other metrics in county-level classification across the tropical
-cyclones considered.
+(Figure~\ref{fig:jaccard}; \textbf{Tables S[x]--S[x] provide the underlying
+numbers for the most extensive of these, the storms for which 250 of more
+counties were exposed based on at least one metric}). In this figure, each row
+provides results for one tropical cyclone, and each box in that row shows the
+Jaccard coefficient for a pair of metrics. For all pairs of metrics, agreement
+in exposure assessment was, at best, moderate for all but a few storms. When
+comparing distance- and wind-based exposure assessment, only about 10\% of
+storms had Jaccard indices higher than 0.6 (i.e., out of the counties assessed
+as exposed by at least one of the two metrics in the pair,~60\si{\percent} or
+more were assessed the same under both metrics). For comparisons of assessments
+based on other combinations of distance-, wind-, rain-, and flood-based
+metrics, fewer than 5\% of storms had Jaccard indices above 0.6.  The
+tornado-based metric had universally poor agreement with other metrics in
+county-level classification across the tropical cyclones considered.
 

### 2020-09-03 ehp_revision/R/misclassification_tables.R

@@ -1,0 +1,249 @@
+library(hurricaneexposure)
+library(hurricaneexposuredata)
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(purrr)
+library(viridis)
+library(dendextend)
+library(ComplexHeatmap)
+library(circlize)
+library(stringr)
+library(xtable)
+
+data(closest_dist)
+
+calc_jaccard <- function(x){
+  x_mat <- x %>%
+    dplyr::select(-fips) %>%
+    dplyr::mutate_all(funs(as.numeric(.))) %>%
+    as.matrix()
+  jaccard <- proxy::simil(t(x_mat), method = "jaccard")
+  out <- combn(colnames(x_mat), 2) %>%
+    t() %>%
+    as_tibble(.name_repair = "unique") %>%
+    dplyr::rename(metric_1 = `...1`,
+                  metric_2 = `...2`) %>%
+    dplyr::mutate(jaccard = as.vector(jaccard))
+  return(out)
+}
+
+calc_metric_1_not_2 <- function(x){
+  x_mat <- x %>%
+    dplyr::select(-fips) %>%
+    dplyr::mutate_all(funs(as.numeric(.))) %>%
+    as.matrix()
+  met_1_not_2 <- function(i, j){
+    x <- x_mat[ , i]
+    y <- x_mat[ , j]
+    if(sum(x | y) == 0) return(NA)
+    a <- sum(x & y)
+    b <- sum(x & !y)
+    c <- sum(!x & y)
+    metric <- b / (a + b + c)
+    return(metric)
+  }
+  out_metric <- outer(1:5, 1:5, FUN = Vectorize(met_1_not_2))
+  out <- combn(colnames(x_mat), 2) %>%
+    t() %>%
+    as_tibble(.name_repair = "unique") %>%
+    dplyr::rename(metric_1 = `...1`,
+                  metric_2 = `...2`) %>%
+    dplyr::mutate(met_1_not_2 = as.vector(t(out_metric))[t(upper.tri(out_metric))])
+  return(out)
+}
+
+
+calc_metric_2_not_1 <- function(x){
+  x_mat <- x %>%
+    dplyr::select(-fips) %>%
+    dplyr::mutate_all(funs(as.numeric(.))) %>%
+    as.matrix()
+  met_2_not_1 <- function(i, j){
+    x <- x_mat[ , i]
+    y <- x_mat[ , j]
[…]

### 2020-09-03 ehp_revision/response_to_reviewers.Rmd
Work some more on revision

@@
 - \allowdisplaybreaks
+- \usepackage[markers,nolists]{endfloat}
+- \pagenumbering{gobble}
 ---

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd
Work some more on revision

--- 2020-09-03 ehp_revision/response_to_reviewers.Rmd
@@
 - \allowdisplaybreaks
+- \usepackage[nomarkers]{endfloat}
+- \pagenumbering{gobble}
 ---
@@
 
-# Overall comments {-#overall-comments}
+# Overview {-#overview}
 
@@
 time and also changes in development in at-risk areas, to be comparable over
-extended time periods [@Pielke1998].
+extended time periods [@pielke1998].
 
@@
 whose resources are overwhelmed and 2) those states who apply for FEMA aid.
-(FEMA is likely less objective given societal effects (politics) [@Kunkel1999])
+(FEMA is likely less objective given societal effects (politics) [@kunkel1999])
 As a result, other severely impacted counties and states that may not fit FEMA
@@
 exposure assignments of counties, both within and between storms, when different
-metrics of exposure are used [@Grabich2016; @Grabich2015]. Such
+metrics of exposure are used [@grabich2015measuring; @grabich2015]. Such
 missclassification can have important consequences in assessments of the public
@@
 One study used FEMA disaster declarations as a binary indicator of tropical
-storm exposure [@Grabich2016]. They found that this metric tended to assess a
+storm exposure [@grabich2015measuring]. They found that this metric tended to assess a
 lot more counties as "exposed" to a storm than distance- or wind-based metrics
-[@Grabich2016], although they only compared the two exposure metrics as applied
+[@grabich2015measuring], although they only compared the two exposure metrics as applied
 to four storms (Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in
@@
 changes in development in at-risk areas, to be comparable over extended time
-periods [@Pielke1998].
+periods [@pielke1998].
 
@@
 
-1. If exposure to the different hazards strongly agrees (i.e., when a county is
-exposed to one hazard, they are typically also exposed to a second), this would
-have several implications for epidemiological research, especially for
-multi-year studies:
-  + From a statistical modeling point of view, if you tried to include each
-  hazard as a separate independent variable, you would run into problems with
-  collinearity if fitting the study data with a generalized linear model or
-  other regression model. One implication of this is that the confidence
-  intervals on effect estimates for the two variables would be very inflated,
+The reviewer makes a good point that it is important for us to clarify our
+motivation for measuring agreement between the hazard-based metrics, as well as
+to show readers how our results from these measures can be used to help in
+designing studies and planning statistical analysis for tropical cyclone
+epidemiology.
+
+A tropical cyclone brings a number of physical hazards, including the four 
+included in our study---wind, rain, flooding, and tornadoes. These separate 
+hazards may each cause health impacts and in some cases might act together 
+in synergy in elevating health risks. For example, severe wind and rain 
+together might create a higher probability of power outages, since the rain 
+could soften the ground and make it more likely that the wind can uproot 
+trees. These physical hazards tend to follow different patterns, and this is
+something that's intuitively understood by atmospheric scientists who study 
+the storms. Storm winds tend to decrease rapidly once a storm's center is 
+over land, for example, because the storm loses its main energy source---warm
+water. Rainfall can persist, and might be particularly severe when the storm 
+hits landscape that is not smooth, like the Appalachian Mountains. Tornadoes
+can be generated in outer bands of the storm, far from the storm's center, 
+but this tends to only happen in the southern half of the US. Rain and flooding
+will sometimes align spatially, but you can also have severe flooding downstream
+from the area that receives the most severe rain, particularly in coastal areas
+with large networks of waterways.
+
+None of these patterns would be particularly surprising to an atmospheric scientist, 
+but many epidemiologists might not be aware of them, or of how they play into
+exposure assessment for tropical cyclone epidemiology. Many environmental 
+epidemiologists have a training that was focused more on epidemiology and 
+biostatistics and rely on collaborations with experts in engineering,
+atmospheric science, or environmental health to ensure that they understand
+and are appropriately applying exposure assessment. In the case of tropical 
+cyclones, it is clear that they generate a mixture of physical hazards. 
+There has been a lot of methodological development and thought recently in 
+environmental epidemiology about how to study mixtures of chemicals---for example, 
+the mixture of air pollutants that a person might be exposed to, rather than 
+studying the individual effects of single air pollutants. 
+These mixture methods require an understanding of how the different components
+are associated with each other. In the case of binary, rather than continuous,
+exposure, this will be expressed as agreement among the different exposures.
+
+First, if exposure to a pair of different hazards strongly agrees (i.e., when a
+county is exposed to one hazard, they are typically also exposed to a second),
+this would have several implications for epidemiological research, especially
+for multi-year studies:
+
+- From a statistical modeling point of view, if you tried to include each hazard
+as a separate independent variable, you would run into problems with
+collinearity if fitting the study data with a generalized linear model or other
+regression model. One implication of this is that the confidence intervals on
+effect estimates for the two variables would be inflated, potentially a lot,
 because the model would struggle to figure out which of the two variables the
@@
 size of the associated confidence intervals.
-  + If you see an effect, it would be very hard to try to determine if the
-  causal pathway goes through the first hazard or the second. To mitigate
-  effects, it is important to understand the pathway through which they happen,
-  and it might be easier to get clues towards pathways if the hazards are better
-  separated. (Analogy: if you were able to conduct a controlled experiment to
-  look at the effects of each hazard, jointly and in combination, you would
-  design separation in the assignment of these treatments, so you could
-  distinguish. We can't do that because we are limited to an observational
-  analysis, but the area of causal inference is doing a lot to explore how we
-  can design observational studies and analyze data to come closer to what we
-  might be able to see with a randomized control study, and which of those
-  methods we might be able to use for tropical cyclone epidemiology will depend
-  on the interplay of these patterns of exposures.)
+- If you see an effect, it would be very hard to try to determine if the causal
+pathway goes through the first hazard or the second. To mitigate effects, it is
+important to understand the pathway through which they happen, and it might be
+easier to get clues towards pathways if the hazards are better separated. By
+analogy, if you were able to conduct a controlled experiment to look at the
+effects of each hazard, jointly and in combination, you would design separation
+in the assignment of these treatments, so you could distinguish between the two
+factors. For tropical cyclone epidemiology, we are limited to an observational
+analysis, but the area of causal inference is doing a lot to explore how we can
+design observational studies and analyze data to come closer to what we might be
+able to see with a randomized control study. To determine which of those methods
+could be used for tropical cyclone epidemiology, it is important to understand
+the interplay of these patterns of exposures.
+- If this is the case, and a study is conducted on a single hazard of the storm, 
+then there is the risk that results have been confounded by a different hazard
+of the storm. 
+- On the positive side, in this case the researcher may be able to fit a simpler
+statistical model to estimate the joint effects of exposure to the storm as a
+whole. Statistical models of mixtures can be complex to lay out, and their 
+results can be hard to interpret and explain to others. If all hazards of a 
+tropical cyclone follow very similar spatial patterns, then it would be 
+reasonable to model tropical cyclone exposure based on one of the hazards
+(and it agreement is strong enough, it wouldn't make much difference which 
+of the hazards was used for this). This would allow for the use of very 
+straightforward statistical modeling, drawing on well-developed models currently
+used in environmental epidemiology to study things like the health risk associated
+with exposure to a single pollutant or a single physical hazard like outdoor 
+temperature.
+
+Second, if a pair of hazards tend to have little spatial overlap among 
+counties, this also brings some implications for epidemiological study design
+and statistical analysis:
+
+- In this case, a single-hazard analysis could miss a lot of the risk created
+by the storm as a whole. This is a concern because there have been some cases
+of studies that have assessed exposure based on a single storm, typically wind.
+[Grabich example, Darren Sun example]. We are aware of one other paper 
+currently under review that takes this same approach of focusing on the wind
+hazard of these storms. .... In interpreting these papers, it is helpful to
+know if this hazard can stand as a reasonable proxy to the collection of 
+physical hazards of the storm. If not, the paper must be interpreted as 
+investigating
+
+In the past, much of tropical cyclone epidemiology focused on cases studies of
+single storms. Multi-year studies are becoming more common. Both types of
+studies are important in exploring both typical and unusual health risks that
+can be caused by these storms. Therefore, it's important to understand the
+patterns between these different hazards both for single storms and across many
+storms. In our analysis, we find that these hazards typically are not in strong
+agreement across most storms, suggesting that mixture methods could be
+meaningfully applied in most cases, and that there may often be the chance to
+separate the role of different hazards through such models. However, we also
+find that there are a few storms for which the hazards follow very similar
+spatial patterns (Hurricane Floyd in 1999, for example), and in these cases
+there may be problems of multi-collinearity and non-separability of different
+hazard effects if a researcher tries to fit a model that incorporates each
+separate hazard.
+
+Issues:
+
+1. **Interpretation** Effect of one hazard, or storm as a whole?
+
+2. **Confounding**
+
+3. **Multi-collinearity**
+
+4. **Disentangling role of different hazards**
+
+5. **Statistical power** (to separate effects of different hazards)
+
+------------------------------------------------------------------------------
+
 2. These results will not be surprising to an atmospheric scientist / disaster
@@
 
-The largest-scale county-level exposure study is likely that of @Zandbergen2009,
+The largest-scale county-level exposure study is likely that of @zandbergen2009,
 which estimated exposure in US counties to all US landfalling Atlantic basin
@@
 exposure to tropical storms and its location, distance from the coast, size, and
-shape [@Zandbergen2009]. @Kruk2010 explored exposure to hurricane-related winds
+shape [@zandbergen2009]. @kruk2010 explored exposure to hurricane-related winds
 in the United States, including inland areas, for 1900-2008.
@@
 this study and also other reasonable thresholds that could be considered.
+This is Table S1 of the revised manuscript and has been reproduced as 
+Table [x] at the end of this response document.
+
+\input{tables/thresholds_table.tex}
+
 Ultimately, this choice will depend on what a researcher considers as plausible
@@
 be more reasonable. Some studies have done sensitivity analyses to consider how
-results changed using different exposure metrics. For example, @Czajkowski2011
+results changed using different exposure metrics. For example, @czajkowski2011
 performed a sensitivity analysis using different assumptions about wind decay
-away from the center of a storm, while @Currie2013, who used a distance-based
+away from the center of a storm, while @currie2013, who used a distance-based
 exposure metric, considered different buffer distances as a sensitivity
@@
 to provide storm rainfall totals from the slow-moving Hurricane Danny in 1997
-[@Medlin2007]. Single-storm studies often measure rainfall as the total rainfall
-over the full "event" (e.g., Medlin2007); however, the timing of the event is
+[@medlin2007]. Single-storm studies often measure rainfall as the total rainfall
+over the full "event" (e.g., @medlin2007); however, the timing of the event is
 determine in a customized way for each storm, making it difficult to extend to
@@
 metrics have defined "exposure" to occur only when a storm's center track passes
-through a county's boundary (e.g., @Zandbergen2009, who describe this metric of
-exposure as a "hit" on the county; @Kinney2008 is another example that
+through a county's boundary (e.g., @zandbergen2009, who describe this metric of
+exposure as a "hit" on the county; @kinney2008 is another example that
 considered highest exposures for counties directly on the storm's path). This
@@
 conditions from the storm but were not directly on the hurricane's track
-[@Kruk2010], especially since hurricanes can have a width ([typical storm width;
+[@kruk2010], especially since hurricanes can have a width ([typical storm width;
 citation])[JMF-- Weatherford and Gray paper referenced in Grabich 2016 may be a
@@
 winds can extend well away from the storm's center. Further, exposure assessed
-using this method is associated with the size [@Zandbergen2009; @Kruk2010] and
-shape [@Zandbergen2009] of a county. (While some studies have avoided this
-problem by determining exposure for equal-area grid cells (e.g., [@Kruk2010]),
+using this method is associated with the size [@zandbergen2009; @kruk2010] and
+shape [@zandbergen2009] of a county. (While some studies have avoided this
+problem by determining exposure for equal-area grid cells (e.g., [@kruk2010]),
 in studies where exposure needs to be matched with a county-level outcome, some
@@
 distance, recent studies have defined the distance from the storm track affected
-by a given storm differently. @Czajkowski2011 assessed county-level risk and
-exposure based on a three-tierd definition, with primary counties being those
+by a given storm differently. @czajkowski2011 assessed county-level risk and
+exposure based on a three-tiered definition, with primary counties being those
 closest to the storm track on either side, secondary counties being adjacent to
-primary coutnies, and tertiary counties adjacent to secondary counties. Such a
+primary counties, and tertiary counties adjacent to secondary counties. Such a
 definition resulted in an exposure defenition based on an average distance
-radius of 120 km [75 miles] on either side of the storm track [@Czajkowski2011].
+radius of 120 km [75 miles] on either side of the storm track [@czajkowski2011].
 Such a distance is slightly larger than that commonly used by public health
-departments (i.e., 100 km) [@Czajkowski2011]. Other studies used distance
-buffers of 30 kilometers [@Currie2013] or of 30, 60, and 100 kilometers as
-thresholds for identifying distance-based exposure [@Grabich2016] [check the
+departments (i.e., 100 km) [@czajkowski2011]. Other studies used distance
+buffers of 30 kilometers [@currie2013] or of 30, 60, and 100 kilometers as
+thresholds for identifying distance-based exposure [@grabich2015measuring] [check the
 Grabich paper-- was this distance to any part of a county, or to the center of
@@
 counties for which the 60 km buffer fully or partially passed through the
-county's boundaries as exposed [@Grabich2015] [again, check to see if this was
+county's boundaries as exposed [@grabich2015] [again, check to see if this was
 to any point in the county or to some center point of the county][JMF-- it is
@@
 to the omission of residences 100 km from the storm path, and that results
-varied insignificantly from 30-75 km [@Currie2013]. One study used distance to
+varied insignificantly from 30-75 km [@currie2013]. One study used distance to
 determine exposure to landfalling US hurricanes by assigning counties along the
@@
 (120 kilometers) from the storm track for a county to qualify as "exposed" to a
-storm [@Czajkowski2011]. The study found that about 30 direct hurricane
+storm [@czajkowski2011]. The study found that about 30 direct hurricane
 fatalities over their study period (1970--2007) occurred outside of this
-distance range from the storm center [@Czajkowski2011], suggesting that this
+distance range from the storm center [@czajkowski2011], suggesting that this
 distance choice can cause some counties with dangerous exposures to a storm to

### 2020-09-09 ehp_revision/tables/thresholds_table.tex
Work some more on revision

\begin{table}
\caption{Caption. These are provided for the three exposure metrics for which our database includes continuous data, and so a threshold is selected to determine binary exposure based on the metric. This table provides reasoning for the choice of threshold used in this paper as well as guidance on other thresholds that could be considered, depending on the hypothesized pathways for an epidemiological study.}
\centering
\begin{tabular}{>{\bfseries\leavevmode\color{black}}l>{\raggedright\arraybackslash}p{40em}}
\toprule
Metric & Threshold choice\\
\midrule
Distance & In the manuscript, distance-based exposure was assessed if the storm track came within 100 km of county population mean center. This threshold for distance from the storm center has been used in prior research as a proxy for exposure to hazards from the storm \cite{grabich2015measuring}. Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km \cite{mallin2006, quiring2011variations}, and dangerous winds can extend beyond these maximum winds. One study \cite{czajkowski2011} assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties. Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track \cite{czajkowski2011}. Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research \cite{grabich2015measuring, grabich2015, currie2013}.\\
Rain & In the manuscript, rain-based exposure was assessed if the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center. This threshold... Other reasonable definitions could vary both the threshold of rainfall (in mm) and also the choice of days surrounding the storm to include when calculating the cumulative rainfall. Tropical cyclone precipitation arrives before the center of the storm \cite{zhou2017spatial}. A variety of definitions have been used to identify both extreme or heavy rain (whether associated with a tropical cyclone or not) and tropical cyclone--associated rain. Some definitions of extreme rain events are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours \cite{schumacher2006characteristics}. Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]). \cite{schumacher2006characteristics, schumacher2005organization, stevenson201410}. A study focused on the risks that result from extreme rainfall may therefore reasonably choose a higher threshold than the 75 mm threshold we use in this analysis. In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall \cite{zhou2017spatial} and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location \cite{feldmann2019estimation}. \\
Wind & In the manuscript, wind-based exposure was assessed if modeled storm-associated peak sustained surface wind of $\ge34$ kts at the county’s population mean center. This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale. This limit is used as the outer limit in measuring storm size through the US National Hurricane Center's wind radii for tropical cyclone forecasts \cite{cangialosi2016examination}. Other thresholds could be selected based on other points on the Beaufort scale---for example, $\ge48$ kts for capturing storm-force winds or $\ge64$ kts for capturing hurricane-force winds. As a note, hurricane-force winds will be rarely experienced for counties, as it will likely only be observed for very severe storms and even for those for counties near the storm's landfall. Many presentations of the Beaufort wind scale include descriptions of the conditions that winds in each category would produce both over land and at sea.\\
\bottomrule
\end{tabular}
\end{table}

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 where not everyone in a county would have been exposed and the difficulties of
-going from country-level to individual -level risks. Further the comparison of
+going from country-level to individual-level risks. Further the comparison of
 different storm hazards could be clarified. The reviewers identified several
@@
 
+For figures and tables that were changed in the revision, we have reproduced them 
+at the end of this response, as well as some figures showing how figures will be
+perceived for those who are colorblind and when printed in grayscale.
+
+\medskip
+
 # Reviewer 1: {-#reviewer-1:}
@@
 each in detail below.
+
+\medskip
 
@@
 aggregated to the county level. **This spatial scale allows for easy integration
-with health outcome data aggregated at the county level. Such aggregated or
-ecological data is often used for disaster epidemiology, as aggregated data may
-be easier to access than individual-level data, especially at a scale that
-covers many locations and a long time period and so allows higher statistical
-power and a broader range of exposure levels within the study
-[@wakefield2008overcoming].**
+with health outcome data aggregated at the county level. Such data is often used
+for disaster epidemiology, as aggregated data may be easier to access than
+individual-level data, especially at a scale that covers many locations and a
+long time period and so allows higher statistical power and a broader range of
+exposure levels within the study [@wakefield2008overcoming].**
 
@@
 
+\medskip
+
 \begin{shaded}
@@
 municipalities in Mexico, districts in India. There would be a few challenges,
-however, in developing the exposure dataset for the hazards that we cover in the
-US in the data presented in this paper. We have added text to the paper's
-Discussion on opportunities and barriers to expanding this approach
-internationally:
+however, in extending the exposure dataset for the hazards that we present in
+this paper. We have added text to the paper's Discussion on opportunities and
+barriers to expanding this approach internationally:
 
@@
 separate validation and calibration to ensure it performs appropriately in those
-settings. Other wind field modeling software is also available that already
+settings. Alternatively, other wind field modeling software is available that
 provides a global coverage, like Geoscience Australia's Tropical Cyclone Risk
@@
 municipalities in Mexico, districts in India).**"
+
+\medskip
 
@@
 
-This is a great question, and we are aware of a number of studies on the
-societal impacts of tropical cyclones that have used FEMA disaster declarations
-to assess county-level exposure to tropical cyclones.
-
-There are two elements that are important related to this question. First, many
-of the health impacts of tropical cyclones will likely come through indirect
-pathways, including through damage to property and infrastructure. Ultimately,
-it would be useful to have large-scale databases of these secondary exposure
-pathways to use for more complex studies of the pathways between the physical
-hazards of these storms and how they affect human health. In this sense,
-something like the FEMA declarations could be considered not as a proxy for
-assessing exposure to physical hazards, but rather as a proxy of damage caused
-by the storm, and so a measurement informative for one of these indirect
-pathways from physical hazards of the storm to health impacts, through something
-like a mediation analysis. We have added text to the manuscript on this point:
-
-...
-
-In terms of using FEMA declarations as a source for this, however, there are 
-some limitations. These declarations are on average tied to the amount of 
-damage and other impacts experienced following a storm, but there are a number
-of other considerations that drive whether a disaster is declared or not. 
-These contributing factors could vary by location (especially from one state
-to another) and also over time, especially over the long period we've included
-for our other data. Given that, these may not be the best data source to use
-to try to capture estimates of the damage or impacts to a community in a multi-year, 
-multi-site dataset like the one we present.
-
+This is a great question. For tropical cyclones, some health impacts will come
+directly from physical exposures like wind, rain, and flooding. However, many of
+the impacts might come through indirect pathways, which start from physical
+hazards but move through pathways related to property damage, infrastructure
+damage, access to typical medical care, mold growth, and a number of other
+intermediates. Ultimately, it would be useful to have large-scale databases of
+these secondary exposure pathways to use for more complex studies of the
+pathways between the physical hazards of these storms and how they affect human
+health. In this sense, something like the FEMA declarations could be considered
+not as a proxy for assessing exposure to physical hazards, but rather as a proxy
+of damage caused by the storm, and so a measurement informative for one of these
+indirect pathways from physical hazards of the storm to health impacts, through
+something like a mediation analysis.
+
+However, there are some limitations to the FEMA disaster declarations in terms
+of using them as a proxy for the amount of damage a storm caused in a county.
 FEMA's disaster declarations are issued to provide assistance, based on damage
@@
 nature of disaster declaration process, these declarations are often subject to
-many political and economic factors [@mccarthy2014fema; @logue1981research]. One
-epidemiological study specifically focused on exposure assessment for tropical
-cyclone epidemiology and compared use of FEMA declarations with other methods of
-exposure assessment, including one based directly on the storm hazard of wind
-[@grabich2015measuring]. They found that exposure assessment based on FEMA
-declarations tended to overassign counties to be "exposed," resulting in false
-positives [@grabich2015measuring]. A further concern is that, given the
-political and economic nature of these declarations, there is likely variation
-across time and geography in the likelihood of a given exposure resulting in a
-disaster declaration. This is particularly worrisome for a study that seeks to
-explore risk across multiple years and affected communities.
-
-Other sources of data may be better suited to use as a marker of damage to
-assess that pathway of indirect health impacts from a tropical cyclone. For
-example, data on storm impacts, including property damage or insurance claims
-data, are sometimes available through storm databases or publications (e.g.,
-NOAA's Storm Events data, *Monthly Weather Review*). While such data could
-potentially be used to create an exposure metric, tropical storm damage data
-requires some normalization, to incorporate both changes in dollar value over
-time and also changes in development in at-risk areas, to be comparable over
-extended time periods [@pielke1998].
-
----------------------------------------------------------------------------
-
-However, evidence from prior tropical cyclone epidemiology that doing so is
-problematic and should be avoided [@grabich2015measuring].
-
-We have added some discussion of this point...
-
-- Future research could expand the work by [@grabich2015measuring] to determine
-the extent of these issues with use of FEMA disaster declarations for
-epidemiological exposure assessment. This would need to go beyond the comparison
-with exposure assessment based on storm hazards, because sometimes the
-characteristics of the affected community modifies the amount of disaster that
-results from hazards of a certain severity. For example, a community that it
-exposed often might be more "hardened" against damage from the storm (any work
-from Seth on this) and so might experience less infrastructure damage from a
-tropical cyclone with certain characteristics compared to a community that is
-rarely exposed to tropical cyclones. One interesting direction would be to
-investigate if there is evidence of differences in probablity of disaster
-declarations at geopolitical boundaries, like when comparing counties on either
-side of a state border.
-
-There are some cases where damage reports are directly used to infer the
-intensity of the weather event. For example, damage surveying is used following
-a tornado to assign the tornado's class on the [enhanced] Fujita scale. However,
-this is not the intent of FEMA damage reports, which instead are used to
-determine post-disaster aid [?] and are influenced by a number of political
-factors.
-
-> "Because of the historic lack of direct measurements and remotely sensed
-tornado wind speeds at or near group level, damage surveying has remained the
-most common form for indicating tornado strength. ... The occurrence of a direct
-tornado strict upon a fixed, sufficiently sturdy, and well-calibrated wind
-measuring station is quite rare. Only 31 direct in situ tornado observations are
-evident between 1894 and 2011." [@edwards2013tornado]
-
-This question could be explored to some degree through a study that uses the 
-physical hazard data provided in this dataset as the exposure and FEMA disaster
-declarations as the outcome. Such a study could clarify the degree to which 
-each hazard---and its intensity---helps in explaining that probablity that 
-a disaster declaration is issued. It could also clarify the degree to which 
-other factors, beyond the intensity of these hazards, drive the probability of
-a county receiving a disaster declaration. For example, if two counties 
-are exposed to a storm-associated hazard of the same intensity, they may 
-still have different probabilities of associated societal impacts, as a 
-community that is regularly exposed to that hazard may be more "hardened" 
-against it. Other differences in infrastructure and factors associated with 
-vulnerability could similarly introduce variability.
-
-At a more complex level, studies could explore the causal pathways from the 
-initial physical hazard to resulting damage/property loss through to associations
-with health outcomes, to provide a clearer idea of the direct and indirect 
-pathways leading to health impacts from tropical cyclones.
-
-For both of these studies, however, there may be better sources of damage
-estimates to use than FEMA disaster declarations. FEMA disaster declarations 
-can be prone to political and other influences, and since their primary function
-is not to record damage but rather to be a step in getting help to communities, 
-they can be problematic to use as an unbiased estimate of damage levels in 
-epidmiological research. One study, for example, found ... . There are other
-estimates of property damage that might be less problematic, including the 
-estimates recorded in NOAA's Storm Events database and data from private 
-insurers, although the quality of disaster loss data is well known to be
-less than ideal for secondary research [can be prone to bias through a number
-of mechanisms] [ref to Gall paper?].
-
-FEMA declaration is largely subjective, requiring state governors and/or tribal
-chiefs to submit a proposal for government aid on the basis that local resources
-are overwhelmed [@FEMA]. As a result, using FEMA disaster declarations as
-exposure classification to hurricanes restricts exposure to 1) those states
-whose resources are overwhelmed and 2) those states who apply for FEMA aid.
-(FEMA is likely less objective given societal effects (politics) [@kunkel1999])
-As a result, other severely impacted counties and states that may not fit FEMA
-criteria are excluded and may therefore be incorrectly identified as unexposed,
-preventing researchers from assessing the full impact of a given hurricane
-[JMF-- the Grabich 2016 study actually found that FEMA assigned the highest
-exposure, and may overestimate the health effects of a hurricane; GBA-- let's
-think about why that might have had "false positives". I think that would be an
-interesting point to include here.]. Researchers have found heterogeneity in
-exposure assignments of counties, both within and between storms, when different
-metrics of exposure are used [@grabich2015measuring; @grabich2015]. Such
-missclassification can have important consequences in assessments of the public
-health and economic impacts of tropical storms.
-
-One study used FEMA disaster declarations as a binary indicator of tropical
-storm exposure [@grabich2015measuring]. They found that this metric tended to assess a
-lot more counties as "exposed" to a storm than distance- or wind-based metrics
-[@grabich2015measuring], although they only compared the two exposure metrics as applied
-to four storms (Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in
-2004).
-
-Data on storm impacts, including property damage or insurance claims data, are
-sometimes available through storm databases or publications (e.g., NOAA's Storm
-Events data, *Monthly Weather Review*). While such data could potentially be
-used to create an exposure metric, tropical storm damage data requires some
+many political and economic factors [@mccarthy2014fema; @logue1981research].
+Given the political and economic nature of these declarations, there is likely
+variation across time and geography in the likelihood of a given exposure
+resulting in a disaster declaration. This would be particularly worrisome for a
+study that seeks to explore risk across multiple years and affected communities.
+
+Other sources of data may therefore be better suited to use as a marker of
+damage to assess that pathway of indirect health impacts from a tropical
+cyclone. For example, data on storm impacts, including property damage or
+insurance claims data, are sometimes available through storm databases or
+publications (e.g., NOAA's Storm Events data, tropical cyclone reports published
+in the *Monthly Weather Review*). While such data could potentially be used to
+create an exposure metric, tropical storm damage data requires some
 normalization, to incorporate both changes in dollar value over time and also
@@
 periods [@pielke1998].
+
+We have added text discussing this point to the Discussion:
+
+> **Text in revised manuscript (added text in bold):**
+"**The dataset we present focuses on the physical hazards of a tropical cyclone.
+However, health impacts will often some through indirect pathways, including
+through damage to property and infrastructure. In future work, it would be
+useful to expand this dataset to add data related to these pathways, to allow
+research exploring the role of indirect pathways from tropical cyclone phyiscal
+hazards to health risk. This expansion could include adding data on normalized
+storm-associated damages, or proxy measurements of such damages, from sources
+like the US NOAA's Storm Events database and the US Federal Emergency Management
+Agency's county-level disaster declarations.**"
+
+\medskip
 
@@
 them to estimate average associations over many disasters and to explore how a
-disaster's characteristics, or the characteristics ofthe **affected
+disaster's characteristics, or the characteristics of the **affected
 communities**, modify associated health risks (e.g., Anderson and Bell 2010; Son
@@
 We also used the word "Hits per county per decade" in the legend for Figure 5 in
-the main text and Figure S3 of the Supplemental Material. We have revised the
-wording in these figures to "Exposures per county per decade".
+the main text (reproduced in this response document as Figure \@ref(fig:fig5))
+and Figure S3 of the Supplemental Material. We have revised the wording in these
+figures to "Exposures per county per decade".
 
@@
 ```
+
+\medskip
 
@@
 
-It turns out that it should already be accessible to use on the DesignSafe
-platform, because we have shared the data as an R package.
+It turns out that the data we present should already be accessible to use on the
+DesignSafe platform, because we have shared the data as an R package.
 
@@
 drawing on data stored in the Data Depot and while using scientific programming
-tools.
-
-As part of its platform, DesignSafe allows users to work in a number of
+tools. As part of its platform, DesignSafe allows users to work in a number of
 programming languages, including R (through Jupyter Notebooks)
@@
 
+\medskip
+
 # Reviewer 2: {-#reviewer-2}
@@
 
-We have moved other suggestions that you sent in "Track Changes" of the Word 
-document for the original manuscript and address each in detail below. When 
-there were several comments on a related topic, we have grouped them together
-to provide a more cohesive response to the point.
+We have moved other suggestions that the reviewer sent in "Track Changes" of the
+Word document for the original manuscript and address each in detail below. When
+there were several comments on a related topic, we have grouped them together to
+provide a more cohesive response to the point.
+
+\medskip
 
@@
 correlation at higher precipitation values through the use of examples from
-specific storms, and discuss the potential reasons for it, but measuring the
+specific storms, and discuss the potential reasons for it. Measuring the
 correlation specifically within these higher-precipitation events would add more
@@
 We have added estimates of these correlations specifically for
-high-precipitation events in a new table in the Supplemental Material (Table S1,
-reproduced as Table \@ref(tab:highprecipcorr) of this response document).
+high-precipitation events in a new table in the Supplemental Material (Table S1
+of the manuscript, reproduced as Table \@ref(tab:highprecipcorr) of this
+response document).
 
@@
 
-The original and revised versions of Figure 4 of the paper are reproduced in
+The original and revised versions of Figure 4 of the manuscript are reproduced in
 Figure \@ref(fig:fig4) of this response.
@@
 
-Figure 7 of the main text is meant to help illustrate the potential for 
-exposure misclassification, including when using a distance-based metric as
-a proxy. We selected to use the Jaccard index because we are using data for 
-counties at a large spatial scale. For any storm, there will be large parts of
-our study area that are clearly unaffected, because they are so far from the 
-storm. If we measured misclassification using a metric that includes these
-counties in the denominator, it would be overly optimistic, and wouldn't help
-to pinpoint the potential for disagreement in areas nearer the storm path.
+Figure 7 of the main text is meant to help illustrate the potential for exposure
+misclassification, including when using a distance-based metric as a proxy. We
+used the Jaccard index because we are using data for all eastern US counties.
+For any storm, however, there will be large parts of this area that are
+unquestionably unaffected, because they are so far from the storm. For example,
+during Hurricane Floyd in 1999, the storm made landfall in North Carolina and
+moved northward along the Eastern Seaboard. While there will be questions in
+some case about whether counties in the Mid-Atlantic and New England were
+exposed to this storm, counties in Texas, Louisiana, and other states in the
+west of the study area were unquestionably unexposed to this storm. If we
+measured misclassification using a metric that includes these counties in the
+denominator, it would be overly optimistic because of the strong agreement in
+counties far from the storm, like Texas counties during Hurricane Floyd, and
+wouldn't help to pinpoint the potential for disagreement in areas nearer the
+storm path.
 
@@
 
+\input{tables/wind_misclass.tex}
+
+\input{tables/rain_misclass.tex}
+
+\input{tables/flood_misclass.tex}
+
+\input{tables/tornado_misclass.tex}
+
 We have added four new tables in the Supplemental Material, one for each of the
-hazard-based metrics we considered. We have reproduced these tables in this response
-document as Tables [x]--[x] Each breaks down agreement and disagreement
-between the distance-based proxy metric and that hazard-based metric within several
-storms in our study period, focusing on storms with at least 250 counties exposed
-by at least one of the metrics considered. All of these storms are included in 
-Figure 7.
-
-\input{tables/wind_misclass.tex}
-
-\input{tables/rain_misclass.tex}
-
-\input{tables/flood_misclass.tex}
-
-\input{tables/tornado_misclass.tex}
-
-These tables allow readers both to investigate general patterns in disagreement
-and exposure misclassification across these storms and also to see cases of unusual 
-storms. For example, for most storms, disagreement between the distance proxy 
-metric and the wind-based metric resulted mainly from mis-classification of 
-counties as exposed based on distance while they did not experience gale-force winds.
+hazard-based metrics we considered. We have reproduced these tables in this
+response document as Tables [x]--[x] Each breaks down agreement and disagreement
+between the distance-based proxy metric and that hazard-based metric within
+several storms in our study period. These tables focus on storms with at least
+250 counties exposed by at least one of the metrics considered. These tables
+allow readers both to investigate general patterns in disagreement and exposure
+misclassification across these storms and also to see cases of unusual storms.
+For example, for most storms, disagreement between the distance proxy metric and
+the wind-based metric resulted mainly from mis-classification of counties as
+exposed based on distance while they did not experience gale-force winds.
 However, there were a few storms, like Hurricane Bertha in 1996, where the
-inverse was the case, and a distance-based metric would have missed a large number
-of counties that experienced gale-force winds from the storm (misclassified them
-as unexposed), but not misclassified many counties as exposed when they in fact
-did not experience winds at this level.
+inverse was the case, and a distance-based metric would have missed a large
+number of counties that experienced gale-force winds from the storm
+(misclassified them as unexposed), but not misclassified many counties as
+exposed when they in fact did not experience winds at this level.
 
@@
 misclassification **(Tables S[x]--S[x])**."
+
+\medskip
 
@@
 
+\medskip
+
 # Reviewer 3: {-#reviewer-3}
@@
 concert with the tropical cyclone's forward motion."
+
+\medskip
 
@@
 Census** (US Census Bureau 2020)."
+
+\medskip
 
@@
 data."
+
+\medskip
 
@@
 
+\medskip
+
 # Requests from EHP: {-#requests-from-ehp}
@@
 paper are reproduced in this response document as Figure \@ref(fig:fig3).
+
+\medskip
 
@@
 overlap each other.
+
+\medskip
 
@@
 
+\medskip
+
 # References for the response {-#references-for-the-response}

### 2020-09-09 ehp_revision/writing/ehp_discussion.tex
Work some more on response

@@
 include separate validation and calibration to ensure it performs appropriately
-in those settings. Other wind field modeling software is also available that
-already provides a global coverage, like Geoscience Australia's Tropical
+in those settings. Alternatively, other wind field modeling software is
+available that provides a global coverage, like Geoscience Australia's Tropical
 Cyclone Risk Model (http://geoscienceaustralia.github.io/tcrm/). For tornado
@@
 (e.g., municipalities in Mexico, districts in India).}
+
+\textbf{The dataset we present focuses on the physical hazards of a tropical
+cyclone.  However, health impacts will often some through indirect pathways,
+including through damage to property and infrastructure. In future work, it
+would be useful to expand this dataset to add data related to these pathways,
+to allow research exploring the role of indirect pathways from tropical cyclone
+phyiscal hazards to health risk. This expansion could include adding data on
+normalized storm-associated damages, or proxy measurements of such damages,
+from sources like the US NOAA's Storm Events database and the US Federal
+Emergency Management Agency's county-level disaster declarations.}
 
@@
 scale allows for easy integration with health outcome data aggregated at the
-county level. Such aggregated or ecological data is often used for disaster
-epidemiology, as aggregated data may be easier to access than individual-level
-data, especially at a scale that covers many locations and a long time period
-and so allows higher statistical power and a broader range of exposure levels
-within the study \parencite{wakefield2008overcoming}.}
+county level. Such data is often used for disaster epidemiology, as aggregated
+data may be easier to access than individual-level data, especially at a scale
+that covers many locations and a long time period and so allows higher
+statistical power and a broader range of exposure levels within the study
+\parencite{wakefield2008overcoming}.}
 

### 2020-09-09 ehp_revision/writing/ehp_response_darlings.tex
Work some more on response

@@
 
+# R1 Comment 3
+
+In this paper, we present a dataset that focuses on exposure to the primary
+physical hazards of wind, rain, tornadoes, and flooding. We do not include data
+related to intermediate pathways, but such data could be a very interesting
+future expansion of the data.
+
+There are two elements that are important related to this question. First, many
+of the health impacts of tropical cyclones will likely come through indirect
+pathways, including through damage to property and infrastructure. Ultimately,
+it would be useful to have large-scale databases of these secondary exposure
+pathways to use for more complex studies of the pathways between the physical
+hazards of these storms and how they affect human health. In this sense,
+something like the FEMA declarations could be considered not as a proxy for
+assessing exposure to physical hazards, but rather as a proxy of damage caused
+by the storm, and so a measurement informative for one of these indirect
+pathways from physical hazards of the storm to health impacts, through something
+like a mediation analysis. We have added text to the manuscript on this point:
+
+In terms of using FEMA declarations as a source for this, however, there are 
+some limitations. These declarations are on average tied to the amount of 
+damage and other impacts experienced following a storm, but there are a number
+of other considerations that drive whether a disaster is declared or not. 
+These contributing factors could vary by location (especially from one state
+to another) and also over time, especially over the long period we've included
+for our other data. Given that, these may not be the best data source to use
+to try to capture estimates of the damage or impacts to a community in a multi-year, 
+multi-site dataset like the one we present.
+
+FEMA's disaster declarations are issued to provide assistance, based on damage
+assessment, to individuals or the entire public in certain locations where
+catastrophes overwhelm the state or local government [@mccarthy2014fema], and
+any use of them for exposure assessment is secondary. Due to the political
+nature of disaster declaration process, these declarations are often subject to
+many political and economic factors [@mccarthy2014fema; @logue1981research]. One
+epidemiological study specifically focused on exposure assessment for tropical
+cyclone epidemiology and compared use of FEMA declarations with other methods of
+exposure assessment, including one based directly on the storm hazard of wind
+[@grabich2015measuring]. They found that exposure assessment based on FEMA
+declarations tended to overassign counties to be "exposed," resulting in false
+positives [@grabich2015measuring]. A further concern is that, given the
+political and economic nature of these declarations, there is likely variation
+across time and geography in the likelihood of a given exposure resulting in a
+disaster declaration. This is particularly worrisome for a study that seeks to
+explore risk across multiple years and affected communities.
+
+Other sources of data may be better suited to use as a marker of damage to
+assess that pathway of indirect health impacts from a tropical cyclone. For
+example, data on storm impacts, including property damage or insurance claims
+data, are sometimes available through storm databases or publications (e.g.,
+NOAA's Storm Events data, *Monthly Weather Review*). While such data could
+potentially be used to create an exposure metric, tropical storm damage data
+requires some normalization, to incorporate both changes in dollar value over
+time and also changes in development in at-risk areas, to be comparable over
+extended time periods [@pielke1998].
+
+However, evidence from prior tropical cyclone epidemiology that doing so is
+problematic and should be avoided [@grabich2015measuring].
+
+We have added some discussion of this point...
+
+- Future research could expand the work by [@grabich2015measuring] to determine
+the extent of these issues with use of FEMA disaster declarations for
+epidemiological exposure assessment. This would need to go beyond the comparison
+with exposure assessment based on storm hazards, because sometimes the
+characteristics of the affected community modifies the amount of disaster that
+results from hazards of a certain severity. For example, a community that it
+exposed often might be more "hardened" against damage from the storm (any work
+from Seth on this) and so might experience less infrastructure damage from a
+tropical cyclone with certain characteristics compared to a community that is
+rarely exposed to tropical cyclones. One interesting direction would be to
+investigate if there is evidence of differences in probablity of disaster
+declarations at geopolitical boundaries, like when comparing counties on either
+side of a state border.
+
+There are some cases where damage reports are directly used to infer the
+intensity of the weather event. For example, damage surveying is used following
+a tornado to assign the tornado's class on the [enhanced] Fujita scale. However,
+this is not the intent of FEMA damage reports, which instead are used to
+determine post-disaster aid [?] and are influenced by a number of political
+factors.
+
+> "Because of the historic lack of direct measurements and remotely sensed
+tornado wind speeds at or near group level, damage surveying has remained the
+most common form for indicating tornado strength. ... The occurrence of a direct
+tornado strict upon a fixed, sufficiently sturdy, and well-calibrated wind
+measuring station is quite rare. Only 31 direct in situ tornado observations are
+evident between 1894 and 2011." [@edwards2013tornado]
+
+This question could be explored to some degree through a study that uses the 
+physical hazard data provided in this dataset as the exposure and FEMA disaster
+declarations as the outcome. Such a study could clarify the degree to which 
+each hazard---and its intensity---helps in explaining that probablity that 
+a disaster declaration is issued. It could also clarify the degree to which 
+other factors, beyond the intensity of these hazards, drive the probability of
+a county receiving a disaster declaration. For example, if two counties 
+are exposed to a storm-associated hazard of the same intensity, they may 
+still have different probabilities of associated societal impacts, as a 
+community that is regularly exposed to that hazard may be more "hardened" 
+against it. Other differences in infrastructure and factors associated with 
+vulnerability could similarly introduce variability.
+
+At a more complex level, studies could explore the causal pathways from the 
+initial physical hazard to resulting damage/property loss through to associations
+with health outcomes, to provide a clearer idea of the direct and indirect 
+pathways leading to health impacts from tropical cyclones.
+
+For both of these studies, however, there may be better sources of damage
+estimates to use than FEMA disaster declarations. FEMA disaster declarations 
+can be prone to political and other influences, and since their primary function
+is not to record damage but rather to be a step in getting help to communities, 
+they can be problematic to use as an unbiased estimate of damage levels in 
+epidmiological research. One study, for example, found ... . There are other
+estimates of property damage that might be less problematic, including the 
+estimates recorded in NOAA's Storm Events database and data from private 
+insurers, although the quality of disaster loss data is well known to be
+less than ideal for secondary research [can be prone to bias through a number
+of mechanisms] [ref to Gall paper?].
+
+FEMA declaration is largely subjective, requiring state governors and/or tribal
+chiefs to submit a proposal for government aid on the basis that local resources
+are overwhelmed [@FEMA]. As a result, using FEMA disaster declarations as
+exposure classification to hurricanes restricts exposure to 1) those states
+whose resources are overwhelmed and 2) those states who apply for FEMA aid.
+(FEMA is likely less objective given societal effects (politics) [@kunkel1999])
+As a result, other severely impacted counties and states that may not fit FEMA
+criteria are excluded and may therefore be incorrectly identified as unexposed,
+preventing researchers from assessing the full impact of a given hurricane
+[JMF-- the Grabich 2016 study actually found that FEMA assigned the highest
+exposure, and may overestimate the health effects of a hurricane; GBA-- let's
+think about why that might have had "false positives". I think that would be an
+interesting point to include here.]. Researchers have found heterogeneity in
+exposure assignments of counties, both within and between storms, when different
+metrics of exposure are used [@grabich2015measuring; @grabich2015]. Such
+missclassification can have important consequences in assessments of the public
+health and economic impacts of tropical storms.
+
+One study used FEMA disaster declarations as a binary indicator of tropical
+storm exposure [@grabich2015measuring]. They found that this metric tended to assess a
+lot more counties as "exposed" to a storm than distance- or wind-based metrics
+[@grabich2015measuring], although they only compared the two exposure metrics as applied
+to four storms (Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in
+2004).
+
+Data on storm impacts, including property damage or insurance claims data, are
+sometimes available through storm databases or publications (e.g., NOAA's Storm
+Events data, *Monthly Weather Review*). While such data could potentially be
+used to create an exposure metric, tropical storm damage data requires some
+normalization, to incorporate both changes in dollar value over time and also
+changes in development in at-risk areas, to be comparable over extended time
+periods [@pielke1998].
+
 # R1 Comment 6

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd
Work some more on revisions

@@
 hazard-based metrics we considered. We have reproduced these tables in this
-response document as Tables [x]--[x] Each breaks down agreement and disagreement
-between the distance-based proxy metric and that hazard-based metric within
-several storms in our study period. These tables focus on storms with at least
-250 counties exposed by at least one of the metrics considered. These tables
-allow readers both to investigate general patterns in disagreement and exposure
-misclassification across these storms and also to see cases of unusual storms.
-For example, for most storms, disagreement between the distance proxy metric and
-the wind-based metric resulted mainly from mis-classification of counties as
-exposed based on distance while they did not experience gale-force winds.
-However, there were a few storms, like Hurricane Bertha in 1996, where the
-inverse was the case, and a distance-based metric would have missed a large
-number of counties that experienced gale-force winds from the storm
-(misclassified them as unexposed), but not misclassified many counties as
-exposed when they in fact did not experience winds at this level.
+response document as Tables \@ref(tab:misclasswind)--\@ref(tab:misclassflood)
+Each breaks down agreement and disagreement between the distance-based proxy
+metric and that hazard-based metric within several storms in our study period.
+These tables focus on storms with at least 250 counties exposed by at least one
+of the metrics considered. These tables allow readers both to investigate
+general patterns in disagreement and exposure misclassification across these
+storms and also to see cases of unusual storms. For example, for most storms,
+disagreement between the distance proxy metric and the wind-based metric
+resulted mainly from mis-classification of counties as exposed based on distance
+while they did not experience gale-force winds. However, there were a few
+storms, like Hurricane Bertha in 1996, where the inverse was the case, and a
+distance-based metric would have missed a large number of counties that
+experienced gale-force winds from the storm (misclassified them as unexposed),
+but not misclassified many counties as exposed when they in fact did not
+experience winds at this level.
 

### 2020-09-09 ehp_revision/R/misclassification_tables.R

@@ -145,7 +145,8 @@ misclass_rain %>%
          Storm = paste0(Storm, ")")) %>% 
   mutate_at(2:5, prettyNum, big.mark = ",") %>% 
   xtable(align = "ccp{3cm}p{3cm}p{3cm}p{3cm}",
-          caption = "Caption. Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.  Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.).") %>% 
+         caption = "Caption. Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.  Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.).", 
+         label = "tab:misclassrain") %>% 
   print(include.rownames = FALSE, booktabs = TRUE, 
          only.contents = FALSE,
          file = "ehp_revision/tables/rain_misclass.tex", 
@@ -177,7 +178,8 @@ misclass_wind %>%
          Storm = paste0(Storm, ")")) %>% 
   mutate_at(2:5, prettyNum, big.mark = ",") %>% 
   xtable(align = "ccp{3cm}p{3cm}p{3cm}p{3cm}",
-         caption = "Caption. Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns). Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.") %>% 
[…]

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 day into a central office, which used this information to create synoptic
-weather maps that captured major weather systems in the county and from this
+weather maps that captured major weather systems in the county and then
 sent back forecasts to local weather offices, also at standardized times
@@
 This is Table S1 of the revised manuscript and has been reproduced as 
-Table [x] at the end of this response document.
+Table \@ref(tab:thresholds) at the end of this response document.
 
@@
 
-- **Wind:** This threshold is the threshold for gale-force and higher winds
-based on the Beaufort wind scale. This value is used by the National Hurricane
-Center [?] as the outer limit of winds for its wind radii product in storm 
-tracking data. Other reasonable choices for a threshold of wind exposure 
-could include the two more intense thresholds also used in the NHC [?] wind
-radii estimates of storm extent (50 knots and 64 knots, the later of which 
-is the threshold for hurricane-force winds on the Beaufort Scale) or other
-thresholds on the Beaufort Scale (e.g., "strong gale", which starts at 
-41 knots; "storm", which starts at 48 knots; "violent storm", which starts
-at 56 knots).
-
-- **Rain:** This threshold is ... Other reasonable thresholds might be ... or
-might incorporate different ranges of days around the storm's closest approach
-to a county in calculating the cumulative precipitation. For example, ... For
-the most part, using our window (two days before to one day after a storm's
-closest approach) captured most of the storm-associated precipitation that would
-have been counted using a wider window from five days before to two days after a
-storm's closest approach, even for a large, slow storm. However, for a storm
-that loops and so passes some counties twice, such as Allison (2001), this
-four-day window may not be adequate, and lead to an underestimation of total
-rainfall associated with the storm relative to estimates for faster-moving
-storms. Other storms that looped over or near land and so passed close to US
-counties twice include Allison (1989), Alberto (1994), Gordon (1994), Dennis
-(1999), Edouard (2002), Fay (2002), Alex (2004), Ivan (2004), Dennis (2005), and
-Ophelia (2005). One difficulty in measuring rain exposure consistently across
-many different storms is how many days to include in the rain total. Some
-storms, especially ones associated with dangerous rainfalls, can move very
-slowly, meaning a window of several days needs to be used to fully capture
-storm-related rainfall. For example, one study used a period of over four days
-to provide storm rainfall totals from the slow-moving Hurricane Danny in 1997
-[@medlin2007]. Single-storm studies often measure rainfall as the total rainfall
-over the full "event" (e.g., @medlin2007); however, the timing of the event is
-determine in a customized way for each storm, making it difficult to extend to
-multi-storm analyses. We included more days prior to the storm because
-storm-related rains often precede the passing of the storm's center [ref].
-
-- **Distance:** This threshold is one that has been commonly used in studies in
-which distance from the storm's track is used as a proxy for tropical cyclone
-exposure (for example, [refs]). Further, [anything in relation to storm size?].
-Other reasonable thresholds for distance include ..., all of which have also
-been used in previous research in assessing tropical cyclone exposure using
-distance from the storm's track as a proxy. Other distance-based exposure
-metrics have defined "exposure" to occur only when a storm's center track passes
-through a county's boundary (e.g., @zandbergen2009, who describe this metric of
-exposure as a "hit" on the county; @kinney2008 is another example that
-considered highest exposures for counties directly on the storm's path). This
-way of measuring exposure can exclude nearby counties that suffered extreme
-conditions from the storm but were not directly on the hurricane's track
-[@kruk2010], especially since hurricanes can have a width ([typical storm width;
-citation])[JMF-- Weatherford and Gray paper referenced in Grabich 2016 may be a
-good place to start. Suggest 30-60km width of the eye of the storm itself] that
-is much larger than a typical county's width and since a hurricane's maximum
-winds can extend well away from the storm's center. Further, exposure assessed
-using this method is associated with the size [@zandbergen2009; @kruk2010] and
-shape [@zandbergen2009] of a county. (While some studies have avoided this
-problem by determining exposure for equal-area grid cells (e.g., [@kruk2010]),
-in studies where exposure needs to be matched with a county-level outcome, some
-process of aggregation from grid cells to counties would still need to be
-performed.) By contrast, assessing distance-based exposure based on measuring
-distance between the storm track and each county's center would be less
-susceptible to differences in county size and shape and so might be a more
-reliable metric of exposure than determining if a storm track ever crossed a
-county boundary. Further, assessing the minimum distance between county center
-and the storm track, as we do here, allows for a continuous, rather than binary,
-metric of distance-based exposure to a storm. To date, distance parameters
-involved with assessing risk of a particular storm have been rather arbitrary,
-contributing to the necessity in understanding how such parameters influence a
-county's exposure status. In assessing the risk of a given storm based on
-distance, recent studies have defined the distance from the storm track affected
-by a given storm differently. @czajkowski2011 assessed county-level risk and
-exposure based on a three-tiered definition, with primary counties being those
-closest to the storm track on either side, secondary counties being adjacent to
-primary counties, and tertiary counties adjacent to secondary counties. Such a
-definition resulted in an exposure defenition based on an average distance
-radius of 120 km [75 miles] on either side of the storm track [@czajkowski2011].
-Such a distance is slightly larger than that commonly used by public health
-departments (i.e., 100 km) [@czajkowski2011]. Other studies used distance
-buffers of 30 kilometers [@currie2013] or of 30, 60, and 100 kilometers as
-thresholds for identifying distance-based exposure [@grabich2015measuring] [check the
-Grabich paper-- was this distance to any part of a county, or to the center of
-the county?; JMF-- I think any part of the county]. Another study used a
-distance criterion of 60 kilometers from the storm track, classifying all
-counties for which the 60 km buffer fully or partially passed through the
-county's boundaries as exposed [@grabich2015] [again, check to see if this was
-to any point in the county or to some center point of the county][JMF-- it is
-not abundantly clear in the Grabich 2016 article; based on their graphs it seems
-like they considered a county exposed if the definition they used crossed any
-county boundary. The Grabich 2015 paper specifies both full and partial counties
-being included]. In a study assessing the association between hurricanes and
-undesireable birth outcomes, researchers found that results were not sensitive
-to the omission of residences 100 km from the storm path, and that results
-varied insignificantly from 30-75 km [@currie2013]. One study used distance to
-determine exposure to landfalling US hurricanes by assigning counties along the
-storm's tracks as "primary" exposures, counties neighboring primary counties as
-"secondary" counties, and counties neighboring secondary counties as "tertiary"
-counties; this method correspond to a distance buffer of approximately 75 miles
-(120 kilometers) from the storm track for a county to qualify as "exposed" to a
-storm [@czajkowski2011]. The study found that about 30 direct hurricane
-fatalities over their study period (1970--2007) occurred outside of this
-distance range from the storm center [@czajkowski2011], suggesting that this
-distance choice can cause some counties with dangerous exposures to a storm to
-be mis-identified as "unexposed". Tropical cyclones vary dramatically in size:
-US tropical cyclones have been observed with radii to maximum winds as small as
-20-km and as large as 200-km [@mallin2006; @quiring2011variations].
-
 \medskip
@@
 distinguished when the figure is printed in black and white [@viridis;
-@van2015mpl].
-
-For these figures, we selected colormaps that were created to meet both these
-criteria. These colormaps are all multi-hue sequential maps that avoid red-green
-contrasts, while maximizing dynamic range in comparison to something
+@van2015mpl]. The colormaps we used are all multi-hue sequential maps that avoid
+red-green contrasts, while maximizing dynamic range in comparison to something
 single-hued, like grayscale [@nunez2018optimizing; @viridis]. They are all
@@
 
-For example, in Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2) (left)
+To help illustrate, in Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2)
 of this response document, we have reproduced the current version of Figure 5 as
@@
 protanopia (defective red cone cells) and tritanopia (defective blue cone
-cells). We have also shown how the Figure would appear in grayscale in
+cells). We have also shown how the Figure would appear in grayscale in Figure
 \@ref(fig:fig5check2) (right). We have included similar versions of other
@@
 
-We have reproduced both the original and revised versions of Figure 4 as Figure
-\@ref(fig:fig4) of this response. We have made several changes to make this
+We have made several changes to make this
 figure easier for readers to see and interpret. These changes include:
@@
 overlap each other.
+
+We have reproduced both the original and revised versions of Figure 4 as Figure
+\@ref(fig:fig4) of this response.
 

### 2020-09-09 ehp_revision/writing/ehp_response_darlings.tex
Work some more on response

@@
 
-R3 Comment 5
+# R3 Comment 5
 
@@
 [@Ashley2007]
+
+# R3 Comment 6
+
+- **Wind:** This threshold is the threshold for gale-force and higher winds
+based on the Beaufort wind scale. This value is used by the National Hurricane
+Center [?] as the outer limit of winds for its wind radii product in storm 
+tracking data. Other reasonable choices for a threshold of wind exposure 
+could include the two more intense thresholds also used in the NHC [?] wind
+radii estimates of storm extent (50 knots and 64 knots, the later of which 
+is the threshold for hurricane-force winds on the Beaufort Scale) or other
+thresholds on the Beaufort Scale (e.g., "strong gale", which starts at 
+41 knots; "storm", which starts at 48 knots; "violent storm", which starts
+at 56 knots).
+
+- **Rain:** This threshold is ... Other reasonable thresholds might be ... or
+might incorporate different ranges of days around the storm's closest approach
+to a county in calculating the cumulative precipitation. For example, ... For
+the most part, using our window (two days before to one day after a storm's
+closest approach) captured most of the storm-associated precipitation that would
+have been counted using a wider window from five days before to two days after a
+storm's closest approach, even for a large, slow storm. However, for a storm
+that loops and so passes some counties twice, such as Allison (2001), this
+four-day window may not be adequate, and lead to an underestimation of total
+rainfall associated with the storm relative to estimates for faster-moving
+storms. Other storms that looped over or near land and so passed close to US
+counties twice include Allison (1989), Alberto (1994), Gordon (1994), Dennis
+(1999), Edouard (2002), Fay (2002), Alex (2004), Ivan (2004), Dennis (2005), and
+Ophelia (2005). One difficulty in measuring rain exposure consistently across
+many different storms is how many days to include in the rain total. Some
+storms, especially ones associated with dangerous rainfalls, can move very
+slowly, meaning a window of several days needs to be used to fully capture
+storm-related rainfall. For example, one study used a period of over four days
+to provide storm rainfall totals from the slow-moving Hurricane Danny in 1997
+[@medlin2007]. Single-storm studies often measure rainfall as the total rainfall
+over the full "event" (e.g., @medlin2007); however, the timing of the event is
+determine in a customized way for each storm, making it difficult to extend to
+multi-storm analyses. We included more days prior to the storm because
+storm-related rains often precede the passing of the storm's center [ref].
+
+- **Distance:** This threshold is one that has been commonly used in studies in
+which distance from the storm's track is used as a proxy for tropical cyclone
+exposure (for example, [refs]). Further, [anything in relation to storm size?].
+Other reasonable thresholds for distance include ..., all of which have also
+been used in previous research in assessing tropical cyclone exposure using
+distance from the storm's track as a proxy. Other distance-based exposure
+metrics have defined "exposure" to occur only when a storm's center track passes
+through a county's boundary (e.g., @zandbergen2009, who describe this metric of
+exposure as a "hit" on the county; @kinney2008 is another example that
+considered highest exposures for counties directly on the storm's path). This
+way of measuring exposure can exclude nearby counties that suffered extreme
+conditions from the storm but were not directly on the hurricane's track
+[@kruk2010], especially since hurricanes can have a width ([typical storm width;
+citation])[JMF-- Weatherford and Gray paper referenced in Grabich 2016 may be a
+good place to start. Suggest 30-60km width of the eye of the storm itself] that
+is much larger than a typical county's width and since a hurricane's maximum
+winds can extend well away from the storm's center. Further, exposure assessed
+using this method is associated with the size [@zandbergen2009; @kruk2010] and
+shape [@zandbergen2009] of a county. (While some studies have avoided this
+problem by determining exposure for equal-area grid cells (e.g., [@kruk2010]),
+in studies where exposure needs to be matched with a county-level outcome, some
+process of aggregation from grid cells to counties would still need to be
+performed.) By contrast, assessing distance-based exposure based on measuring
+distance between the storm track and each county's center would be less
+susceptible to differences in county size and shape and so might be a more
+reliable metric of exposure than determining if a storm track ever crossed a
+county boundary. Further, assessing the minimum distance between county center
+and the storm track, as we do here, allows for a continuous, rather than binary,
+metric of distance-based exposure to a storm. To date, distance parameters
+involved with assessing risk of a particular storm have been rather arbitrary,
+contributing to the necessity in understanding how such parameters influence a
+county's exposure status. In assessing the risk of a given storm based on
+distance, recent studies have defined the distance from the storm track affected
+by a given storm differently. @czajkowski2011 assessed county-level risk and
+exposure based on a three-tiered definition, with primary counties being those
+closest to the storm track on either side, secondary counties being adjacent to
+primary counties, and tertiary counties adjacent to secondary counties. Such a
+definition resulted in an exposure defenition based on an average distance
+radius of 120 km [75 miles] on either side of the storm track [@czajkowski2011].
+Such a distance is slightly larger than that commonly used by public health
+departments (i.e., 100 km) [@czajkowski2011]. Other studies used distance
+buffers of 30 kilometers [@currie2013] or of 30, 60, and 100 kilometers as
+thresholds for identifying distance-based exposure [@grabich2015measuring] [check the
+Grabich paper-- was this distance to any part of a county, or to the center of
+the county?; JMF-- I think any part of the county]. Another study used a
+distance criterion of 60 kilometers from the storm track, classifying all
+counties for which the 60 km buffer fully or partially passed through the
+county's boundaries as exposed [@grabich2015] [again, check to see if this was
+to any point in the county or to some center point of the county][JMF-- it is
+not abundantly clear in the Grabich 2016 article; based on their graphs it seems
+like they considered a county exposed if the definition they used crossed any
+county boundary. The Grabich 2015 paper specifies both full and partial counties
+being included]. In a study assessing the association between hurricanes and
+undesireable birth outcomes, researchers found that results were not sensitive
+to the omission of residences 100 km from the storm path, and that results
+varied insignificantly from 30-75 km [@currie2013]. One study used distance to
+determine exposure to landfalling US hurricanes by assigning counties along the
+storm's tracks as "primary" exposures, counties neighboring primary counties as
+"secondary" counties, and counties neighboring secondary counties as "tertiary"
+counties; this method correspond to a distance buffer of approximately 75 miles
+(120 kilometers) from the storm track for a county to qualify as "exposed" to a
+storm [@czajkowski2011]. The study found that about 30 direct hurricane
+fatalities over their study period (1970--2007) occurred outside of this
+distance range from the storm center [@czajkowski2011], suggesting that this
+distance choice can cause some counties with dangerous exposures to a storm to
+be mis-identified as "unexposed". Tropical cyclones vary dramatically in size:
+US tropical cyclones have been observed with radii to maximum winds as small as
+20-km and as large as 200-km [@mallin2006; @quiring2011variations].

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd
Work more on revision

@@
 - \pagenumbering{gobble}
+- \newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
 ---
@@
 
-We have made several changes to make this
-figure easier for readers to see and interpret. These changes include:
+We have made several changes to make this figure easier for readers to see and
+interpret. These changes include:
 
@@
 vision deficiency and when printed in grayscale Figures \@ref(fig:fig5check1)
-and \@ref(fig:fig5check1) show what Figure 5 would look like under three common
-types of color vision deficiency and when printed in black and white. Figures
-\@ref(fig:fig6check1) and \@ref(fig:fig6check2) show the same thing for Figure
-6. Therefore, these figures should not require any exemption from accessibility
-requirements, as they should satisfy these requirements.
+and \@ref(fig:fig5check2) show what Figure 5 of the main text would look like
+under three common types of color vision deficiency and when printed in black
+and white. Figures \@ref(fig:fig6check1) and \@ref(fig:fig6check2) show the same
+thing for Figure 6 of the main text. Therefore, these figures should not require
+any exemption from accessibility requirements, as they should satisfy these
+requirements.
 

### 2020-09-09 ehp_revision/tables/flood_misclass.tex
Work some more on revisions
Work more on revision

@@
 % latex table generated in R 4.0.2 by xtable 1.8-4 package
-% Thu Sep  3 12:19:54 2020
+% Wed Sep  9 17:22:05 2020
 \begin{table}[ht]
@@
 \caption{Caption.
-Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study.
+Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
 Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
 Exposure assessment is based on the thresholds given in Table 1 of the main text.
-The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
+The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
 Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.}
-\begin{tabular}{cp{3cm}p{3cm}p{3cm}p{3cm}}
+\label{tab:misclassflood}
+\begin{tabular}{cP{3cm}P{3cm}P{3cm}P{3cm}}
   \toprule
@@
   Ernesto (2006) & 44 & 148 & 27 & 2,177 \\ 
+  Hanna (2008) & 52 & 140 & 26 & 2,178 \\ 
+  Helene (2000) & 13 & 179 & 20 & 2,184 \\ 
   Bertha (1996) & 27 & 156 & 17 & 2,196 \\ 
+  Irene (2011) & 108 & 12 & 79 & 2,197 \\ 
   Rita (2005) & 4 & 161 & 26 & 2,205 \\ 
+  Earl (1998) & 4 & 165 & 11 & 2,216 \\ 
   Lee (2011) & 18 & 126 & 36 & 2,216 \\ 
+  Hermine (2010) & 16 & 107 & 40 & 2,233 \\ 
+  Isabel (2003) & 39 & 108 & 9 & 2,240 \\ 
+  Josephine (1996) & 27 & 96 & 32 & 2,241 \\ 
+  Bonnie (2004) & 10 & 99 & 29 & 2,258 \\ 
+  Frances (1998) & 8 & 54 & 23 & 2,311 \\ 
    \bottomrule

### 2020-09-09 ehp_revision/tables/rain_misclass.tex
Work some more on revisions
Work more on revision

@@
 \caption{Caption.
-Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study.
+Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
 Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
 Exposure assessment is based on the thresholds given in Table 1 of the main text.
-The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.
+The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.
 Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.).}
-\begin{tabular}{cp{3cm}p{3cm}p{3cm}p{3cm}}
+\label{tab:misclassrain}
+\begin{tabular}{cP{3cm}P{3cm}P{3cm}P{3cm}}
   \toprule
@@
   Ernesto (2006) & 124 & 68 & 63 & 2,141 \\ 
+  Irene (2011) & 119 & 1 & 106 & 2,170 \\ 
+  Hanna (2008) & 100 & 92 & 33 & 2,171 \\ 
+  Hermine (2010) & 44 & 79 & 102 & 2,171 \\ 
   Fran (1996) & 89 & 62 & 73 & 2,172 \\ 
+  Earl (1998) & 72 & 97 & 38 & 2,189 \\ 
   Bertha (1996) & 103 & 80 & 20 & 2,193 \\ 
+  Helene (2000) & 28 & 164 & 11 & 2,193 \\ 
+  Frances (1998) & 27 & 35 & 139 & 2,195 \\ 
+  Josephine (1996) & 94 & 29 & 77 & 2,196 \\ 
+  Bonnie (2004) & 56 & 53 & 56 & 2,231 \\ 
+  Isabel (2003) & 88 & 59 & 17 & 2,232 \\ 
    \bottomrule

### 2020-09-09 ehp_revision/tables/tornado_misclass.tex
Work some more on revisions
Work more on revision

@@
 \caption{Caption.
-Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study.
+Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
 Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
 Exposure assessment is based on the thresholds given in Table 1 of the main text.
-The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
+The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
 Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.}
-\begin{tabular}{cp{3cm}p{3cm}p{3cm}p{3cm}}
+\label{tab:misclasstornado}
+\begin{tabular}{cP{3cm}P{3cm}P{3cm}P{3cm}}
   \toprule
@@
   Dennis (1999) & 0 & 200 & 1 & 2,195 \\ 
+  Helene (2000) & 5 & 187 & 6 & 2,198 \\ 
   Ernesto (2006) & 2 & 190 & 1 & 2,203 \\ 
+  Hanna (2008) & 0 & 192 & 1 & 2,203 \\ 
   Bertha (1996) & 13 & 170 & 4 & 2,209 \\ 
+  Earl (1998) & 8 & 161 & 7 & 2,220 \\ 
   Fran (1996) & 1 & 150 & 4 & 2,241 \\ 
   Lee (2011) & 14 & 130 & 11 & 2,241 \\ 
+  Isabel (2003) & 0 & 147 & 1 & 2,248 \\ 
   Floyd (1999) & 11 & 128 & 1 & 2,256 \\ 
+  Hermine (2010) & 1 & 122 & 12 & 2,261 \\ 
+  Josephine (1996) & 7 & 116 & 12 & 2,261 \\ 
+  Bonnie (2004) & 13 & 96 & 22 & 2,265 \\ 
+  Irene (2011) & 9 & 111 & 0 & 2,276 \\ 
+  Frances (1998) & 1 & 61 & 10 & 2,324 \\ 
    \bottomrule

### 2020-09-09 ehp_revision/tables/wind_misclass.tex
Work some more on revisions
Work more on revision

@@
 \caption{Caption.
-Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study.
+Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
 Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
 Exposure assessment is based on the thresholds given in Table 1 of the main text.
-The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
+The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
 Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.}
-\begin{tabular}{cp{3cm}p{3cm}p{3cm}p{3cm}}
+\label{tab:misclasswind}
+\begin{tabular}{cP{3cm}P{3cm}P{3cm}P{3cm}}
   \toprule
@@
   Dennis (1999) & 17 & 183 & 13 & 2,183 \\ 
+  Hanna (2008) & 135 & 57 & 12 & 2,192 \\ 
+  Isabel (2003) & 105 & 42 & 56 & 2,193 \\ 
   Ernesto (2006) & 93 & 99 & 4 & 2,200 \\ 
+  Helene (2000) & 18 & 174 & 0 & 2,204 \\ 
   Rita (2005) & 26 & 139 & 20 & 2,211 \\ 
   Fran (1996) & 46 & 105 & 30 & 2,215 \\ 
+  Earl (1998) & 133 & 36 & 11 & 2,216 \\ 
   Floyd (1999) & 124 & 15 & 40 & 2,217 \\ 
+  Irene (2011) & 114 & 6 & 46 & 2,230 \\ 
   Lee (2011) & 34 & 110 & 0 & 2,252 \\ 
+  Josephine (1996) & 62 & 61 & 3 & 2,270 \\ 
+  Hermine (2010) & 22 & 101 & 1 & 2,272 \\ 
+  Bonnie (2004) & 0 & 109 & 0 & 2,287 \\ 
+  Frances (1998) & 15 & 47 & 0 & 2,334 \\ 
    \bottomrule

### 2020-09-09 ehp_revision/R/misclassification_tables.R

@@ -117,7 +117,7 @@ misclass <- distance_data %>%
   dplyr::group_by(storm_id) %>%
   dplyr::mutate(counties_exposed = sum(any_exposure)) %>%
   ungroup() %>%
-  dplyr::filter(counties_exposed >= 250)
+  dplyr::filter(counties_exposed >= 200)
 
 misclass_rain <- misclass %>% 
   mutate(dist_rain = distance & rain, 
@@ -144,8 +144,8 @@ misclass_rain %>%
   mutate(Storm = str_replace(Storm, "-", " ("), 
          Storm = paste0(Storm, ")")) %>% 
   mutate_at(2:5, prettyNum, big.mark = ",") %>% 
-  xtable(align = "ccp{3cm}p{3cm}p{3cm}p{3cm}",
-         caption = "Caption. Limited to storms with at least 250 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.  Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.).", 
+  xtable(align = "ccP{3cm}P{3cm}P{3cm}P{3cm}",
+         caption = "Caption. Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.  Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.).", 
          label = "tab:misclassrain") %>% 
   print(include.rownames = FALSE, booktabs = TRUE, 
          only.contents = FALSE,
@@ -177,8 +177,8 @@ misclass_wind %>%
   mutate(Storm = str_replace(Storm, "-", " ("), 
          Storm = paste0(Storm, ")")) %>% 
   mutate_at(2:5, prettyNum, big.mark = ",") %>% 
-  xtable(align = "ccp{3cm}p{3cm}p{3cm}p{3cm}",
[…]

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd
Work some more on revisions

@@
 We have added estimates of these correlations specifically for
-high-precipitation events in a new table in the Supplemental Material (Table S1
+high-precipitation events in a new table in the Supplemental Material (Table S2
 of the manuscript, reproduced as Table \@ref(tab:highprecipcorr) of this
@@
 100 or more counties were exposed based on at least one metric (Figure 7;
-**Tables S[x]--S[x] provide the underlying numbers for the most extensive of
+**Tables S3--S6 provide underlying numbers for the most extensive of
 these, the storms for which 250 of more counties were exposed based on at least
@@
 exposure to any of the other hazards considered, would often introduce exposure
-misclassification **(Tables S[x]--S[x])**."
+misclassification **(Tables S3--S6)**."
 
@@
 analysis.
+
+We have added text to introduce this supplemental table in the revised manuscript, 
+both in the Methods and in the Discussion:
+
+> **Text in the revised Methods (added text in bold)**:
+"For other exposure metrics, each county was
+classified as exposed to a tropical cyclone based on whether the exposure
+metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
+picked reasonable thresholds (e.g., the threshold for gale-force wind for wind
+exposure; **Table S1**), but others could be used with the open-source data and its
+associated software."
+
+> **Text in the revised Discussion (added text in bold)**:
+"Finally, to assess patterns and agreement for binary exposure classifications,
+we have chosen one set of sensible thresholds for binary classifications based
+on continuous metrics (rainfall, maximum sustained surface wind, and distance
+from the storm's track), **but other thresholds would be reasonable
+depending on hypothesized pathways for a given epidemiological study (Table
+S1)**."
 
@@
 vision deficiency and when printed in grayscale Figures \@ref(fig:fig5check1)
-and \@ref(fig:fig5check2) show what Figure 5 of the main text would look like
-under three common types of color vision deficiency and when printed in black
-and white. Figures \@ref(fig:fig6check1) and \@ref(fig:fig6check2) show the same
-thing for Figure 6 of the main text. Therefore, these figures should not require
-any exemption from accessibility requirements, as they should satisfy these
-requirements.
+and \@ref(fig:fig5check2) at the end of this response show what Figure 5 of the
+main text would look like under three common types of color vision deficiency
+and when printed in black and white. Figures \@ref(fig:fig6check1) and
+\@ref(fig:fig6check2) at the end of this response show the same thing for Figure
+6 of the main text.
+
+For Figure 6 of the main text, for some types of color balance, it could be
+unclear which of the two color legends pairs with which part of the graph. We
+have therefore added a note to the figure legend to clarify:
+
+> **Figure legend for Figure 6 in revised manuscript (added text shown in bold)**: 
+"Agreement between exposure classifications based on different single-hazard
+exposure metrics for all storms between~1996 and~2011 for which at least~100
+counties were exposed based on at least one metric. Each row shows one storm,
+and the color of each cell shows the measured Jaccard index for each pair of
+exposure metrics (proportion of counties classified as exposed by both metrics
+out of counties classified as exposed by either metric). For Grace in 2003 and
+Ida in 2009, there were no county exposures for either the tornado-based metric
+or the wind-based metric (indicated by gray squares). Colors to the right of the
+heatmap show the number of exposed counties based on any of the metrics, **and
+this panel is linked with the color scale labeled '# of counties exposed by
+any metric'.** Storms are displayed within clusters that have similar patterns
+based on hierarchical clustering."
+
+These figures should therefore not require any exemption from accessibility
+requirements, as they should satisfy these requirements.
 

### 2020-09-09 ehp_revision/writing/ehp_discussion.tex
Work some more on revisions

@@
 exposure to any of the other hazards considered, would often introduce exposure
-misclassification \textbf{(Tables S[x]--S[x])}. This conclusion reinforces
+misclassification \textbf{(Tables S3--S6)}. This conclusion reinforces
 similar findings from a study of Florida's 2004 storm
@@
 on continuous metrics (rainfall, maximum sustained surface wind, and distance
-from the storm's track).  Results and conclusions would differ somewhat with
-other threshold choices. We have published code for this analysis online
+from the storm's track), \textbf{but other thresholds would be reasonable
+depending on hypothesized pathways for a given epidemiological study (Table
+S1)}.  Results and conclusions would differ somewhat with other threshold
+choices. We have published code for this analysis online
 (\url{https://github.com/geanders/county_hurricane_exposure}), allowing other

### 2020-09-09 ehp_revision/writing/ehp_methods.tex
Work some more on revisions

@@
 near-land Atlantic basin tropical cyclones while excluding storms that never
-neared the \ac{US}.  \subsection*{Distance-based exposure metric}
+neared the \ac{US}.
+
+\subsection*{Distance-based exposure metric}
 
@@
 we measured the distance between the storm's center and each county's
-population mean center, **as of the 2010 US Decennial Census**~\parencite{countycenters}. We took the minimum distance
-for each county as a measure of how close the storm came to the county over its
-lifetime.  We also recorded the time when this occurred for each county so we
-could later link observed data on precipitation, flooding, and tornadoes.  To
-allow matching with data recorded in local time (e.g., health data), we
-converted these times from \ac{UTC} to local time~\parencite{countytimezones}.
+population mean center, \textbf{as of the 2010 US Decennial
+Census}~\parencite{countycenters}. We took the minimum distance for each county
+as a measure of how close the storm came to the county over its lifetime.  We
+also recorded the time when this occurred for each county so we could later
+link observed data on precipitation, flooding, and tornadoes.  To allow
+matching with data recorded in local time (e.g., health data), we converted
+these times from \ac{UTC} to local time~\parencite{countytimezones}.
 
@@
 metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
-picked reasonable thresholds (e.g., the threshold for gale-force wind for 
-wind exposure), but others could be used with the open-source data and its
+picked reasonable thresholds (e.g., the threshold for gale-force wind for wind
+exposure; \textbf{Table S1}), but others could be used with the open-source data and its
 associated software.

### 2020-09-09 ehp_revision/writing/ehp_results.tex
Work some more on revisions

--- 2020-02-24 writing/ehp_results.tex
@@
 which~100 or more counties were exposed based on at least one metric
-(Figure~\ref{fig:jaccard}). In this figure, each row provides results for one
-tropical cyclone, and each box in that row shows the Jaccard coefficient for a
-pair of metrics. For all pairs of metrics, agreement in exposure assessment
-was, at best, moderate for all but a few storms. When comparing distance- and
-wind-based exposure assessment, only about 10\% of storms had Jaccard indices
-higher than 0.6 (i.e., out of the counties assessed as exposed by at least one
-of the two metrics in the pair,~60\si{\percent} or more were assessed the same
-under both metrics). For comparisons of assessments based on other combinations
-of distance-, wind-, rain-, and flood-based metrics, fewer than 5\% of storms
-had Jaccard indices above 0.6.  The tornado-based metric had universally poor
-agreement with other metrics in county-level classification across the tropical
-cyclones considered.
+(Figure~\ref{fig:jaccard}; \textbf{Tables S3--S6 provide underlying
+numbers for the most extensive of these, the storms for which 250 of more
+counties were exposed based on at least one metric}). In this figure, each row
+provides results for one tropical cyclone, and each box in that row shows the
+Jaccard coefficient for a pair of metrics. For all pairs of metrics, agreement
+in exposure assessment was, at best, moderate for all but a few storms. When
+comparing distance- and wind-based exposure assessment, only about 10\% of
+storms had Jaccard indices higher than 0.6 (i.e., out of the counties assessed
+as exposed by at least one of the two metrics in the pair,~60\si{\percent} or
+more were assessed the same under both metrics). For comparisons of assessments
+based on other combinations of distance-, wind-, rain-, and flood-based
+metrics, fewer than 5\% of storms had Jaccard indices above 0.6.  The
+tornado-based metric had universally poor agreement with other metrics in
+county-level classification across the tropical cyclones considered.
 

### 2020-09-10 ehp_revision/response_to_reviewers.Rmd
Work some more on revisions

@@
 
-> **Text from Results of original manuscript:**
-"Across the four storm hazards considered, there was wide variation in the
-average number of county exposures per year (Table 2). For tropical cyclone
+**Response:**
+
+We hope that these results can help epidemiologists who are planning new
+single-storm or multi-year studies of tropical cyclones and associated health
+risks. For planning a multi-year study, it is helpful to know how many exposures
+the study might incorporate, as this can help in giving a general idea of
+whether the study might have adequate statistical power to investigate a certain
+research question. Similarly, for single-storm studies, it is helpful to know
+how widespread exposure can be for a specific hazard during a very large storm.
+This table provides some basic summaries of the extent of exposure to different
+tropical cyclone hazards, both on average for the storms in the study and also
+for those storms that were most extensive.
+
+We have added text to help the readers interpret this table:
+
+> **Text from Results of revised manuscript (added text in bold):**
+"**Table 2 provides summary statistics describing the extent of counties
+assessed as exposed based on specific storm hazards, to help epidemiologists in
+planning studies, including understanding the potential statistical power to
+study specific exposures, both for multi-storm and single-storm studies.**
+Across the four storm hazards considered, there was wide variation in the
+average number of county exposures per year. For tropical cyclone
 tornadoes, there were on average about 40 county exposures per year within our
@@
 in 2004 based on flooding and tornadoes (Table 2)."
-
-**Response:**
-
-We hope that these results can help epidemiologists who are planning new
-multi-year studies of tropical cyclones and associated health risks. For 
-planning a multi-year study, it is helpful to know how many exposures the 
-study might incorporate, as this can help in giving a general idea of whether
-the study might have adequate statistical power to investigate a certain 
-research question. This table provides some basic summaries of the extent
-of exposure to different tropical cyclone hazards, both on average for the
-storms in the study and also for those storms that were most extensive. 
-Further, as with other results in the study, it provides evidence that 
-patterns in different hazards can vary substantially, both because some 
-hazards are much less likely to expose counties (e.g., torndaoes are a fairly 
-rare exposure, based on this table) and also because there is variation in 
-the most extensive storms across the hazards considered.
-
-We have added text to help the readers interpret this table ...
 

### 2020-09-10 ehp_revision/writing/ehp_results.tex
Work some more on revisions

@@
 to four different hazards for each tropical cyclone and then explored patterns
-over years with available data. Across the four storm hazards considered, there
-was wide variation in the average number of county exposures per year
-(Table~\ref{tab:exposuresummaries}). For tropical cyclone tornadoes,
-there were on average about 40 county exposures per year within our study.
-County exposures were more frequent for tropical cyclone wind exposures ($>$160
-/ yr on average), even more frequent for tropical cyclone flood exposures
-($>$190 / yr on average), and most frequent for tropical cyclone rain exposure
-($>$290 / yr on average). For every hazard except tornadoes, we identified at
-least one tropical cyclone that exposed over~250 counties
+over years with available data. \textbf{Table~\ref{tab:exposuresummaries}
+provides summary statistics describing the extent of counties assessed as
+exposed based on specific storm hazards, to help epidemiologists in planning
+studies, including understanding the potential statistical power to study
+specific exposures, both for multi-storm and single-storm studies.} Across the
+four storm hazards considered, there was wide variation in the average number
+of county exposures per year. For tropical cyclone tornadoes, there were on
+average about 40 county exposures per year within our study.  County exposures
+were more frequent for tropical cyclone wind exposures ($>$160 / yr on
+average), even more frequent for tropical cyclone flood exposures ($>$190 / yr
+on average), and most frequent for tropical cyclone rain exposure ($>$290 / yr
+on average). For every hazard except tornadoes, we identified at least one
+tropical cyclone that exposed over~250 counties
 (Table~\ref{tab:exposuresummaries}).  However, the largest-extent tropical

### 2020-09-10 ehp_revision/response_to_reviewers.Rmd
Work some more on revisions

@@
 ```{r fig5, fig.cap="Reproduction of Figure 5 from the original manuscript (left) and revised version in the resubmitted manuscript (right).
-In the revised version, the figure legend was changed from '\\textbf{Hits} per county per decade' to '\\textbf{Exposures} per county per decade' (bold used to highlight change).
-Figure caption in the revised manuscript: \\textit{Figure caption.}", fig.show = "hold", out.width="40%", fig.align = "center"}
+In the revised version, the figure legend was changed from '\\textbf{Hits} per county per decade' to '\\textbf{Exposures} per county per decade' (bold used to highlight change).}", fig.show = "hold", out.width="40%", fig.align = "center"}
 include_graphics("../figures/averageexposureonly.pdf")
@@
 
+We have included a measurement of agreement between different hazard-based
+exposure metrics for two reasons. First, while we agree with the reviewer that a
+direct measurement is always preferable to a proxy, there are still some studies
+that, when assessing tropical cyclone exposure based on measurements of physical
+hazards, use a single hazard for this exposure assessment. We are most familar
+with cases where a wind-based metric has been the focus in these cases (e.g.,
+@sun2020tropical, @grabich2015measuring). Therefore, sometimes might not be
+ideal, it is done in practice. To interpret these studies, it is helpful to
+understand the degree to which this single-hazard metric agrees with exposure
+patterns for other hazards. If the exposure patterns are in strong agreement for
+all hazards, then these studies could be interpreted as estimating risk
+associated with exposure to tropical cyclones as a whole. If the exposure
+patterns are in low to moderate agreement, then these studies should be
+interpreted instead as estimating the risk associated with that specific storm
+hazard, but not for all storm hazards. In this case, if the agreement is
+moderate, though, there is a chance of confounding from other storm hazards
+(i.e., an association that is attributed to one storm hazard was actually the
+result of another that was not included in the model). In this paper, we present
+measurements of agreement across four hazard-based exposure metrics that can
+provide insight on these points.
+
+Second (and related), epidemiologists need to understand how these exposure
+measurements agree to help in planning studies and statistical approaches to
+interpreting the resulting data. In this sense, "agreement" is providing a
+measure similar to correlation for continuous variables; our choice of agreement
+with the binary exposure classifications reflects the more common use of binary
+classifications for most tropical cyclone epidemiology to date.
+
+There are several things that an epidemiologist could derive from these results
+to help in designing a study and the associated statistical analysis. First, the
+results we present can help epidemiologists determine if they can use a single
+metric of tropical cyclone exposure in their studies, or if they need to instead
+consider tropical cyclones as a mixture of hazards. There are a number of
+practical reasons to use a single exposure metric, if doing so is reasonable
+(even if it's not conceptually ideal). While there have been advances in methods
+for mixtures research in environmental epidemiology, these are often more
+complex than the long-used and well-developed methods for studying single
+exposures, and many epidemiologists are much more comfortable with methods for
+single exposures compared to methods for mixtures. Further, it can be easier to
+communicate results from single-exposure studies clearly and succinctly.
+Therefore, if there is a reasonable way to measure exposure with a single
+metric, there are practical reasons that recommend that approach, even if
+conceptually it might be preferable to incorporate separate hazards.
+
+If, at the county level, there were strong agreement among exposure patterns for
+the different tropical cyclone hazards we considered, it might have been
+reasonable to recommend the use of a single hazard-based metric as a proxy for
+exposure to all storm hazards in epidemiological studies for practical reasons.
+We found that this is almost never the case, although there are a few select
+storms (like Hurricane Floyd in 1999) where such an approach might be
+justifiable. Most atmospheric scientists would not be surprised by our findings
+about this disagreement in exposure patterns for different tropical cyclone
+hazards, but this finding is not as self-evident for environmental
+epidemiologists. Based on a recent conference of the Interational Society for
+Environmental Epidemiology, for example, there are a number of environmental
+epidemiologists with backgrounds focused on methods and other ambient exposures
+(e.g., air pollution, heat) who are beginning to include tropical cyclones as an
+exposure of interest in their studies, and we believe these results would be
+novel and useful to many of them.
+
+Second, in cases where it is clear that a mixture approach is necessary, these
+results can help in understanding what challenges might arise in statistical
+analysis. While one approach to including several hazards could be to create an
+index of tropical cyclone exposure by taking a weighted sum of all hazards, for
+most health outcomes it is unclear *a priori* what the weight for each hazard
+should be in such an index, and so it will often be preferable to discover those
+weights from the data by fitting a statistical model that includes each hazard
+as an independent variable. In these cases, strong correlation (or agreement, in
+the case of binary exposures) among separate independent variables can create
+both numerical issues (e.g., variance inflation from multicollinearity) and
+difficulties in interpretation (e.g., difficulty infering the role of specific
+hazards and separating the influece of each). If an epidemiologist has a good
+understanding of these patterns in their data, he or she can adopt methods that
+address potential issues, including using matching-based study designs to create
+a subset of storms to study in which separate hazards are well-distinguished
+from each other, incorporation of dimension-reduction methods like principal
+components analysis, or shrinkage techniques that allow for a small amount of
+bias in estimation of associations in return for reducing problems from
+collinearity (e.g., ridge regression) [@leal2012multicollinearity;
+@schisterman2017collinearity].
+
 The reviewer makes a good point that it is important for us to clarify our
@@
 designing studies and planning statistical analysis for tropical cyclone
-epidemiology.
-
-A tropical cyclone brings a number of physical hazards, including the four 
-included in our study---wind, rain, flooding, and tornadoes. These separate 
-hazards may each cause health impacts and in some cases might act together 
-in synergy in elevating health risks. For example, severe wind and rain 
-together might create a higher probability of power outages, since the rain 
-could soften the ground and make it more likely that the wind can uproot 
-trees. These physical hazards tend to follow different patterns, and this is
-something that's intuitively understood by atmospheric scientists who study 
-the storms. Storm winds tend to decrease rapidly once a storm's center is 
-over land, for example, because the storm loses its main energy source---warm
-water. Rainfall can persist, and might be particularly severe when the storm 
-hits landscape that is not smooth, like the Appalachian Mountains. Tornadoes
-can be generated in outer bands of the storm, far from the storm's center, 
-but this tends to only happen in the southern half of the US. Rain and flooding
-will sometimes align spatially, but you can also have severe flooding downstream
-from the area that receives the most severe rain, particularly in coastal areas
-with large networks of waterways.
-
-None of these patterns would be particularly surprising to an atmospheric scientist, 
-but many epidemiologists might not be aware of them, or of how they play into
-exposure assessment for tropical cyclone epidemiology. Many environmental 
-epidemiologists have a training that was focused more on epidemiology and 
-biostatistics and rely on collaborations with experts in engineering,
-atmospheric science, or environmental health to ensure that they understand
-and are appropriately applying exposure assessment. In the case of tropical 
-cyclones, it is clear that they generate a mixture of physical hazards. 
-There has been a lot of methodological development and thought recently in 
-environmental epidemiology about how to study mixtures of chemicals---for example, 
-the mixture of air pollutants that a person might be exposed to, rather than 
-studying the individual effects of single air pollutants. 
-These mixture methods require an understanding of how the different components
-are associated with each other. In the case of binary, rather than continuous,
-exposure, this will be expressed as agreement among the different exposures.
-
-First, if exposure to a pair of different hazards strongly agrees (i.e., when a
-county is exposed to one hazard, they are typically also exposed to a second),
-this would have several implications for epidemiological research, especially
-for multi-year studies:
-
-- From a statistical modeling point of view, if you tried to include each hazard
-as a separate independent variable, you would run into problems with
-collinearity if fitting the study data with a generalized linear model or other
-regression model. One implication of this is that the confidence intervals on
-effect estimates for the two variables would be inflated, potentially a lot,
-because the model would struggle to figure out which of the two variables the
-weight should be placed on, and so this uncertainty would be reflected in the
-size of the associated confidence intervals.
-- If you see an effect, it would be very hard to try to determine if the causal
-pathway goes through the first hazard or the second. To mitigate effects, it is
-important to understand the pathway through which they happen, and it might be
-easier to get clues towards pathways if the hazards are better separated. By
-analogy, if you were able to conduct a controlled experiment to look at the
-effects of each hazard, jointly and in combination, you would design separation
-in the assignment of these treatments, so you could distinguish between the two
-factors. For tropical cyclone epidemiology, we are limited to an observational
-analysis, but the area of causal inference is doing a lot to explore how we can
-design observational studies and analyze data to come closer to what we might be
-able to see with a randomized control study. To determine which of those methods
-could be used for tropical cyclone epidemiology, it is important to understand
-the interplay of these patterns of exposures.
-- If this is the case, and a study is conducted on a single hazard of the storm, 
-then there is the risk that results have been confounded by a different hazard
-of the storm. 
-- On the positive side, in this case the researcher may be able to fit a simpler
-statistical model to estimate the joint effects of exposure to the storm as a
-whole. Statistical models of mixtures can be complex to lay out, and their 
-results can be hard to interpret and explain to others. If all hazards of a 
-tropical cyclone follow very similar spatial patterns, then it would be 
-reasonable to model tropical cyclone exposure based on one of the hazards
-(and it agreement is strong enough, it wouldn't make much difference which 
-of the hazards was used for this). This would allow for the use of very 
-straightforward statistical modeling, drawing on well-developed models currently
-used in environmental epidemiology to study things like the health risk associated
-with exposure to a single pollutant or a single physical hazard like outdoor 
-temperature.
-
-Second, if a pair of hazards tend to have little spatial overlap among 
-counties, this also brings some implications for epidemiological study design
-and statistical analysis:
-
-- In this case, a single-hazard analysis could miss a lot of the risk created
-by the storm as a whole. This is a concern because there have been some cases
-of studies that have assessed exposure based on a single storm, typically wind.
-[Grabich example, Darren Sun example]. We are aware of one other paper 
-currently under review that takes this same approach of focusing on the wind
-hazard of these storms. .... In interpreting these papers, it is helpful to
-know if this hazard can stand as a reasonable proxy to the collection of 
-physical hazards of the storm. If not, the paper must be interpreted as 
-investigating
-
-In the past, much of tropical cyclone epidemiology focused on cases studies of
-single storms. Multi-year studies are becoming more common. Both types of
-studies are important in exploring both typical and unusual health risks that
-can be caused by these storms. Therefore, it's important to understand the
-patterns between these different hazards both for single storms and across many
-storms. In our analysis, we find that these hazards typically are not in strong
-agreement across most storms, suggesting that mixture methods could be
-meaningfully applied in most cases, and that there may often be the chance to
-separate the role of different hazards through such models. However, we also
-find that there are a few storms for which the hazards follow very similar
-spatial patterns (Hurricane Floyd in 1999, for example), and in these cases
-there may be problems of multi-collinearity and non-separability of different
-hazard effects if a researcher tries to fit a model that incorporates each
-separate hazard.
-
-Issues:
-
-1. **Interpretation** Effect of one hazard, or storm as a whole?
-
-2. **Confounding**
-
-3. **Multi-collinearity**
-
-4. **Disentangling role of different hazards**
-
-5. **Statistical power** (to separate effects of different hazards)
-
-------------------------------------------------------------------------------
-
-2. These results will not be surprising to an atmospheric scientist / disaster
-researcher, but there are environmental epidemiologists who have done research
-on other exposures and are now starting to study tropical cyclones (session at
-ISEE, for example). It is useful for epidemiologists entering into tropical
-cyclone research to understand these facets of tropical cyclone exposure,
-including the distinction between the very coastal wind exposures and other
-hazards that tend to reach further inland.
-3. There are epidemiological studies that have used only one hazard for exposure
-assessment. This has most often been the case for wind. If this exposure
-assessment is meant to capture "tropical cyclone exposure" in general, with all
-the hazards that are related to tropical cyclones, then there is the chance for
-exposure misclassification if this single hazard is used.
-4. In the past, it was a very complex issue to try to pull together data on all
-these hazards. We agree that now that we have provided a dataset with all of
-them, epidemiologists should be thinking about multi-hazard analyses. However,
-there still will be some barriers to that. [Recent work on multi-pollutant
-exposures and the complexities associated with that.] If one of the hazards
-dominates in explaining associated health risk, then in some cases it might be
-reasonable to focus on that hazard, but in that case it is critical, based on
-our results, to remember that the exposure assessment is not capturing all
-hazards of the storms.
-
-The largest-scale county-level exposure study is likely that of @zandbergen2009,
-which estimated exposure in US counties to all US landfalling Atlantic basin
-tropical storms between 1851 and 2003, using both distance and an exposure
-metric that incorporated distance and windspeed. They used this exposure
-evaluation to create maps of total exposure to tropical storms within US
-counties, as well as to explore associations between a county's long-term
-exposure to tropical storms and its location, distance from the coast, size, and
-shape [@zandbergen2009]. @kruk2010 explored exposure to hurricane-related winds
-in the United States, including inland areas, for 1900-2008.
-
-Geographical patterns can differ for these different
-storm hazards, and so exposure assessment based on one hazard could create
-exposure misclassification if the pathway for health risks is through a
-different storm hazard.
-
-A tropical cyclone's high winds can bring health risks and property damage
-through structural damage of houses and other buildings, falling trees, and
-wind-borne debris [@rappaport2000] and cause power
-outages [@liu2005; han2009], which introduce a number of threats to
-human and ecological health, including water quality risks if the outage affects
-wastewater treatment plants [@mallin2006].  Other risks can exist
-without severe wind; for example, one study found that most of the direct
-hurricane-related deaths in the US between 1970 and 1999 occurred in cases
-when wind was below hurricane strength, including for Tropical Storms Charley
-in 1998 and Alberto in 1994 [@rappaport2000].  Tropical cyclones can
-produce excessive rain, especially in certain topographies (e.g., near
-mountains), so counties well inland sometimes experience more extreme rain than
-coastal counties. Flood risks from tropical cyclones can result from this rain,
-although the two risks are not perfectly correlated [@chen2015]. In the
-US, over half of hurricane-related direct deaths from 1970 to 1999 from
-Atlantic basin storms were a result of freshwater
-flooding [@rappaport2000]. Flooding can also degrade water
-quality [@mallin2006], which can threaten both human and ecological
-health.
-
-If a study misclassifies exposure to the hazard or hazards that cause the
-health risk being studied, the study will generate biased estimates of tropical
-cyclone risks and impacts. Further, when different studies use different
-methods to assess exposure to tropical cyclones, their results are difficult to
-meaningfully compare and aggregate. If different methods identify
-similar sets of communities as ``exposed'',  these concerns are less serious.
-However, if different methods differ substantially in which communities they
-identify as ``exposed'', it makes it very important that epidemiological
-studies are thoughtful in how they assess exposure.
-
-These analyses can also help researchers compare and interpret results from
-previous epidemiological studies that have assessed tropical cyclone exposure in
-different ways, helping to [synthesize] previous epidemiologic results to form a
-better picture of how these storms can affect human health.
-
-However, these frequency maps, together with evidence from
-specific tropical cyclones (Figures \ref{fig:ivanexposure}
-and \ref{fig:jaccard}),  do illustrate the potential for strong differences in
-spatial patterns in tropical cyclone exposures, depending on which tropical
-cyclone hazards are considered.  A few previous studies have sought to
-determine county-level exposure to tropical cyclones over multi-year periods,
-including @zandbergen2009, which estimated exposure in US
-counties to all US landfalling Atlantic-basin tropical cyclones
-between 1851 and 2003, using both a distance-based metric and a metric that
-combined distance and windspeed, and [@kruk2010], which explored
-exposure to hurricane-related winds in the US, including inland areas,
-for 1900--2008.  Our results suggest that such exposure assessments may
-perform well in capturing some tropical cyclone hazards (e.g., wind), but
-likely miss other potentially dangerous tropical cyclone exposures, especially
-for hazards that repeatedly threaten northern or inland counties (e.g., rain,
-flooding).
-
-Therefore, use of distance to assess tropical cyclone exposure for impact
-studies could result in problematic exposure misclassification, which could mask
-true associations, even strong associations, between tropical cyclone exposure
-and outcomes of interest in impact studies [@savitz2016interpreting;
-armstrong1998effect]. The use of a single hazard-based metric (e.g., wind) could
-cause similar problems if the impact is driven, at least in part, by a different
-hazard or by multiple hazards of the tropical cyclone.
-
-If measurements of one storm hazard are used as a proxy to capture exposure to
-a different hazard, or a non-hazard proxy like distance to the storm track is
-used to assess exposure to storm hazards, the resulting exposure
-misclassification could be differential (i.e., associated with the outcome of
-interest or with factors associated with risk of the outcome of interest).
-
-Therefore for most tropical cyclones, since exposure assessments often differ
-across storm hazards, as well as between each storm hazard and a distance-based
-proxy measurement, exposure misclassification is a potential risk.
-
-Further, we investigated agreement in exposure assessment across these metrics
-when applied individually. We found large differences in which counties are
-exposed to different hazards of tropical cyclones and that distance is, at
-best, a moderate, and often a very poor, surrogate for exposure to the specific
-tropical cyclone hazards of high wind, extreme rainfall, flooding, and
-tornadoes. Use of distance as a surrogate for any of these hazards, or exposure
-assessment based on one hazard when the pathway for health impacts is in part
-or full through another storm hazard, could lead to exposure misclassification.
-In the case of tropical cyclone risk and impact studies, including
-epidemiological studies, this would result in biased estimates.  Our findings
-highlight the importance of clarifying the potential pathway from tropical
-cyclone hazards to health impacts when conducting tropical cyclone
-epidemiological studies, and then basing exposure assessment on these hazards.
-
----------------------------------------------------------------------------
-
-It is obvious that it is typically better, but in cases where the proxy is much
-more practical to measure, it may need to be more than just a bit better, but
-instead substantially better. Also, there could be some cases where the proxy
-might actually be better---for example, if the proxy can be measured with very
-low level, while the direct exposure measure might only be measurable with high
-error. In that case, the proxy could be the better metric and might provide a
-more precise estimate of the association with health risk.
-
-[Env. Epi. book---advantages and disadvantages of different types of measures]
-
-We agree that it is helpful to clarify for readers that these independent
-hazards can come together in a storm event at different intensities and at
-different time windows. We have added some text on this:
-
-[Added text]
-
-However, we disagree that this should be down as an alternative to calculating
-agreement. Epidemiologists need to understand how these exposure measurements
-agree to help in planning studies and statistical approaches to interpreting the
-resulting data. In this sense, "agreement" is providing a measure similar to
-correlation for continuous variables; our choice of agreement with the binary
-exposure classifications reflects the more common use of binary classifications
-for most tropical cyclone epidemiology to date.
-
-Epidemiological studies might focus on tropical cyclone exposures at two levels.
-First, they might wish to focus on a specific hazard. For example, a study might
-hypothesize that mold growth following flooding from a tropical storm might
-increase risk of respiratory health outcomes in the months after a storm. In
-this case, the clear causal pathway recommends a focus on one of the storm
-hazards, rather than all storm hazards. However, other storm hazards could
-confound the association between flood exposure and respiratory health risk if
-the other hazards are ignored *and* if they are associated with both the
-exposure of interest (storm-associated flooding) and the health outcome. For
-example, if wind exposure assessments are in high agreement with flood exposure
-assessments for storms, and if wind affects risk of respiratory outcomes by
-causing extensive property damage and creating extreme psychological stress,
-which can have repercussions on respiratory health, then the wind hazard could
-confound the estimated association between storm-related flood exposure and the
-health outcome.
-
-This bias would be higher when the confounder and exposure of interest are more
-strongly correlated (for continuous variables) or in higher agreement (for
-categorical variables). Conversely, if the exposure and the potential confounder
-are, in practice, only weakly associated, then the potential for confounding is
-low. The Jaccard index is one measure of similarities between categorical
-(including binary) coefficients.
-
-Another consideration is collinearity. If one hazard is the exposure of interest
-and another hazard is a potential confounder, then the model should adjust the
-confounder or it will be misspecified and be prone to bias. However, if the
-exposure of interest and the potential confounder are very strongly
-correlated/in agreement, then including both as independent variables in a
-regression model often causes numerical problems in fitting the regression
-model, including instability in the estimated coefficients and inflated
-variance. In very extreme cases, the statistical software may not be able to
-invert the matrix [? only for linear regression?].
-
-Therefore, it is helpful for epidemiologists to understand if these independent
-storm-associated hazards tend to be strongly associated (in high agreement),
-both within specific storms (for single-storm studies) and on average across
-most large storms (for multi-storm studies). This knowledge will help them in
-identifying potential confounders if they are interested in studying the effects
-of a single hazard of the storm and also in determining if control for these
-potential confounders among other storm hazards might cause numerical problems
-or other collinearity-associated problems when fitting a regression model.
-
-If a researcher wishes to study the storm as a whole, then the researcher 
-could either seek a single measure to capture exposure to the storm as a whole
-or investigate more complex models that model the influence of the storm through
-pathways of several different hazards.
-
-If a researcher wants a single measure for the storm as a whole, then it is
-helpful to know how well exposure assessments based on the hazard components of
-the storm agree with each other. If there is strong agreement, then any of those
-hazards could be used as a primary marker of exposure assessment. This is
-conceptually similar to checking for observer / rater agreement.
-
-> "When the row and column variable srepresent different observers' rating the
-same subjects or objects, interest is focused on *observer agreement* rather
-than mere association. In this case, measures and tests of agreement provide a
-method of assessing the reliability of a subjective classification or assessment
-procedure." [@friendly2015discrete]
-
-> "In assessing the strength of *agreement* we usually have a more stringent
-criterion than in measuring the strength of *association*, because observers
-ratings can be strongly associated without strong agreement. For example, one
-rater could use a more stringent criterion and thus consistently rate subjects
-one category lower (on an ordinal scale) than another rater. More generally,
-measures of agreement must take account of the marginal frequencies with which
-two raters use the categories. If observers tend to use the categories with
-different frequency, this will affect measures of agreement."
-[@friendly2015discrete]
-
-When you have an array of two or more exposures that are strongly correlated,
-including them together in a regression model can cause problems from
-collinearity [@schisterman2017collinearity].
-
-> "Environmental exposures ... often consist of an array of related individual
-factors that may be highly correlated with each other, causing concern about the
-impact of collinearity when attempting to identify individual effects."
-[@schisterman2017collinearity]
-
-**Commonly have multiple factors of interest, correlated**
-
-In some areas of epidemiology, researchers have moved toward the idea of a 
-"web of causality":
-
-> "'Multiple causation' is the canon of contemporary epidemiology, and its
-metaphor and model is the 'web of causation'. Expressed through the notion
-of 'multifactorial etiology' and embedded in the statistical techniques
-of 'multivariate analysis', the belief that population patterns of health 
-and disease can be explained by a complex web of numerous interconnected 
-risk and protective factors has become one of this discipline's central 
-concepts. Equally entrenched is the corrolary that epidemiology's power
-to improve the public's health rests upon its ability to identify---and predict
-the results of breaking---selected strands of this causal web." [@krieger1994epidemiology]
-
-This is in contrast to focusing on single factors:
-
-> "Expressly challenging the still-pervasive tendency of epidemiologists to
-think in terms of single 'agents' causing discrete diseases, the provocative
-metaphor and model of the 'web' invited epidemiologists to embrace a more
-sophisticated view of causality." [@krieger1994epidemiology]
-
-In this framework, epidemiology helps by identifying where it would help to
-break strands of the web:
-
-> "Using this model, MacMahon et al. drew several important inferences about
-prevention and research that remain part of epidemiologic thinking to this day.
-Arguing that 'to effect preventative measures, it is not necessary to understand
-causal mechanisms in their entirity', they stated that 'even knowledge of one
-small component may allow some degree of prevention' since 'wherever the chain
-is broken the disease will be prevented'." [@krieger1994epidemiology]
-
-In other words, if you can identify *necessary* causes, you can craft 
-*practical* interventions.
-
-In some fields of epidemiology (e.g., nutritional epidemiology), it is common 
-to study cases where there are multiple potential risk factors, and these 
-factors are often strongly correlated [@schisterman2017collinearity].
-
-> "The dynamics of many important ecological and evolutionary processes are 
-often influenced by multiple interacting factors, and attempting to understand 
-the dynamics of such systems presents difficult technical and philosophical 
-problems." [@petraitis1996inferring]
-
-> "Often in practice, investigators are faced with decisions of how to handle
-highly correlated variables." [@schisterman2017collinearity]
-
-**Methods to tackle multiple causality**
-
-When the aim is to explore the influences of multiple associated factors, there
-are some different strategies that can be used. These include multiple
-regression and path analysis:
-
-> "Evolutionary ecologists have recently turned to path analysis, multiple
-regression, and related techniques to analyse systems of multiple causality.
-While these techniques are often seen as competitors, they are, in fact,
-related, and as a result the problems of these techniques are similar."
-[@petraitis1996inferring]
-
-**Implications of collinearity**
-
-> "Two variables are defined as collinear if one can be expressed as an exact
-or near linear combination of the other. ... In general, collinearity between
-variables is described by the magnitude of their correlation." 
-[@schisterman2017collinearity]
-
-It can complicate inference of model coefficients. This is always a concern for 
-inferential models. For predictive models, it can be a problem if you later use
-the model to predict in a case where the correlation structure is different than
-in the training data.
-
-> "The dependence of regression and path coefficients on the correlation
-structure is often glossed over. Yet this dependence means that inferences about
-the strengths of path coefficients are conditional upon the correlational
-structure among the predictor variables. If the correlation structure of the
-sample does not match the correlational structure of the population then
-regression and path coefficients will not estimate relative strengths
-accurately. This can occur if sampling is not random or if analysis is based on
-data from experiments with factorial designs, which break the correlational
-structure amnog the predictor variables." [@petraitis1996inferring]
-
-These implications can depend on whether the modeling goal is predictive or 
-inferential.
-
-> "When the goal is to choose covariates based on their predictive value, 
-concerns about collinearity stem from the desire to improve parameter estimates'
-precision, which can be inflated when highly correlated covariates are included
-in the model." [@schisterman2017collinearity]
-
-> "In a causal framework, on the other hand, interest is often centered on a
-specific effect from a correctly specified causal model, even in the presence of
-highly correlated data, despite a potential loss of statistical efficiency.
-Under such a viewpoint, bias due to misspecification of the causal associations
-is a more devastating result than reduced precision of parameter estimates
-correctly estimated. Furthermore, a precisely but incorrectly estimated
-parameter may lead to invalid inferences. ... Relying on conventional
-variable-selection algorithms and standard analyses can ignore important
-confounders and even produce extremely imprecise confidence intervals."
-[@schisterman2017collinearity]
-
-The amount of bias in the unadjusted model (in terms of bias from an unadjusted
-confounder) depends on the strength of association between the exposure of
-interest and the potential confounder [@schisterman2017collinearity].
-
-> "This is especially important for the confounding variable scenario, where 
-correctly specifying the model may result in adjusting for a nearly 
-collinear variable." [@schisterman2017collinearity]
-
-The degree of agreement will influence the degree of confounding in an 
-unadjusted model, for either one or two confounders:
-
-> "In this confounding scenario, increasing correlation exacerbates the bias
-due to misspecification, similar to results for a single confounder."
-[@schisterman2017collinearity]
-
-Extreme collinearity can create numerical problems
-[@schisterman2017collinearity].
-
-> "Under the extremely unlikely scenario of extreme collinearity [(abs. rho of
-0.999 or higher)], the numerical matrix inversion required for estimation fails,
-resulting in unstable and biased predictors." [@schisterman2017collinearity]
-
-> "Only for extreme values of rho cloes to 1 or -1 is beta-squareed
-nonidentifiable due to numerical instability; otherwise, the correctly specified
-model produces the most desirable (e.g., unbiased) estimates of the total effect
-of the exposure on the outcome." [@schisterman2017collinearity]
-
-Collinearity can cause imprecision in coefficient estimates, even if those
-estimates are unbiased [@schisterman2017collinearity].
-
-> "Multicollinearity: A term that refers to correlation among the independent
-variables in a multiple regression model; it is usually invoked when some 
-correlations are 'large', but an actual magnitude test is not well defined.[1]"
-[@gunasekara2008glossary]
-
-> "Collinearity occurs in multiple regression models when two (or
-more---multicollinearity) exposure variables are included in the model but are
-so similar to each other that they are essentially measuring at least part of
-the same thing. Collinearity creates problems with interpreting the analysis by
-affecting the standard errors of the variables (as it increases variance)[24]
-and by causing biased estimates for one or both of the collinear terms
-(sometimes dramatically so) [25]. If there is perfect collinearity between two
-variables then one should be dropped from the model. However, dropping exposure
-variables to reduce multicollinearity may lead to bias in the model from
-dropping useful information." [@gunasekara2008glossary]
-
-> "It might be argued in some cases that it is appropriate to include
-(collinear) variables in a model even when they are highly correlated, because
-the theoretical basis for including the variables is strong and the results of
-the model are consistent with expectations [10]. In such cases, models with and
-without the variables would be tested and compared. Such consideraion of
-alternative models tends to be described as an issue of (mis)-specification in
-econometrics. However, in epidemiology this tends to be described in terms of
-mediating and *intermediary variables*. Thus, an epidemiologist might argue for
-including an intermediate variable in a model (and testing it against a model
-without the variable) because including that variable might elucidate direct and
-indirect pathways between the exposure and outcome variables. An econometrician
-might do the same thing but describe this in terms of finding the correct model
-specification." [@gunasekara2008glossary]
-
-**Techniques to address collinearity**
-
-There are study designs that might help to separate different effects. For
-example, some matching designs might help in creating a subset of data in which
-the association that exists in the full dataset is broken (e.g.,
-@leal2012multicollinearity).
-
-> "Matching is typically employed to reduce model dependence and estimate 
-associations in a more empirical way than would be necessary without 
-matching. In line with this practice, matching was employed as a diagnostic
-tool to verify whether the effects of environmental variables that are highly 
-correlated with each other could be disentangled." [@leal2012multicollinearity]
-
-You could also explore techniques of dimension reduction, to create a smaller
-subset of independent variables for which the correlation is broken. For
-example, you could use PCA [@leal2012multicollinearity].
-
-There are other techniques that allow some bias in estimates in an attempt to
-reduce problems from collinearity, like ridge regression
-[@petraitis1996inferring].
-
-> "Several biased estimation techniquest ahve been proposed as a way to combat
-collinearity (Myers 1990). These are controversial methods. Ridge regression,
-for example, biases the regression coefficients in order to improve the model's
-ability to predict the criterion variable. The estimates of the individual
-regression coefficients may be very biased even though when taken together they
-provide a very good prediction of the criterion variable."
-[@petraitis1996inferring]
-
-> "When correct specification of the models still leads to 
-problems for conventional regression methods, one can apply modern techniques
-such as Bayesian hierarchical models with shrinkage estimators."
-[@schisterman2017collinearity]
-
-**Are effects of different factors 'separable'?**
-
-Epidemiology often aims to inform policy decisions. In this case, it can be
-important to characterize the separate effects of different associated factors
-that contribute to the health risk [@leal2012multicollinearity].
-
-> "To develop efficient public health interventions addressing the obesity 
-epidemic, it is important to identify exactly which aspects of the
-environment influence obesity risk. For example, demonstrating that the density 
-of fast-food restaurants was associated with obesity risk and demnstrating that
-the density of sports facilities was associated with obesity risk would lead to 
-different interventions." [@leal2012multicollinearity]
-
-However, it can be very difficult to separate the effects of different factors
-if those factors are strongly correlated [@leal2012multicollinearity]:
-
-> "However, many neighborhood characteristics, especially those related to the
-densities of physical features and services, are strongly correlated with each
-other. Therefore, it is not clear whether it is even possible to disentangle the
-effects of these different environmental dimensions, even if they are
-hypothesized to influence obesity risk through distinct causal pathways."
-[@leal2012multicollinearity]
-
------------------------------------------------------------------------------
-
-As we note in our general response to your comments, we agree in general with
-your instinct here. However, there are some subtle points to note.
-
-First, we understand the potential appeal of creating a single exposure summary
-score based on a summation of exposures. This simplifies some of the statistical
-modeling, for example, and also provides a single exposure estimate to provide
-as the key study conclusion. However, in many areas of tropical cyclone
-epidemiology, it is unlikely that exposure to each type of hazard should be
-given equal weight. For example, an analysis of respiratory health in the months
-after the storm might hypothesize a strong connection with flooding, and smaller
-connections with rainfall and winds (through the pathway of property damage).
-Use of a pre-defined summation of exposure to each hazard would mask some of the
-potential association.
-
-While a summary statistic could be defined that adds a weight to each exposure,
-it would often be unclear *a priori* what weight each hazard exposure should be
-given. A better method might be to use a regression framework, where each hazard
-exposure is included as a separate explanatory variable, and the model uses the
-observed data to determine the appropriate weight for each hazard in terms of
-modifying risk to the outcome of interest. To explore the influence of multiple
-concurrent hazards, interaction terms could be added.
-
-However, if our analysis had found that hazard exposures follow very similar
-patterns, in terms of the counties exposed during a storm, then this strong
-agreement would result in a number of problems in fitting and interpreting these
-types of models. First, strong agreement in two or more of the explanatory
-variables in a regression model leads to collinearity, which among other
-implications often leads to a large inflation of confidence intervals for the
-effect estimates of the correlated explanatory variables. Further, if the hazard
-exposures were in strong agreement, then it would be difficult or impossible to
-disentagle the effects of the separate hazards, as a strong effect for one would
-be impossible to interpret with confidence, since the strong agreement in the
-data (and resulting low co-variability in the variables in the data) could cause
-the model to mistakenly attribute the effect of one hazard to the other.
-Further, it may become impossible in this case to estimate synergistic effects
-between the two, as an indicator of an interaction between the two exposures
-would be almost collinear with the two main effects.
-
-In our analysis, we find that this is usually not the case for the four tropical
-cyclone hazards that we consider. Instead, we find that for most storms, the
-hazards are fairly well-separated in terms of the counties they effect. There is
-a particularly strong pattern where inland counties can be exposed to extreme
-rain and flooding, but not wind. This separation opens the door for some
-interesting study designs and statistical modeling. For example, these patterns
-suggest that a study could investigate whether storm-associated rains alone
-create a health risk, as there are many exposures with only rain in inland
-counties, and so health associations could be explored here and contrasted with
-coastal areas that experienced both wind and extreme rain.
-
--------------------------------------------------------------------------------
-
-> "A job exposure matrix (JEM) may be defined as a cross classification of jobs
-and occupational exposures. Some matrices are based on tasks, instead of jobs. 
-... Usually there is a group of experts who build the matrix based on their 
-own knowledge and experience on occupational exopsures. The experts should 
-assess exposure to an open or fixed list of occupational exposures (that can 
-include chemical, physical, biological, and/or psychosocial agents) for an
-open or fixed list of jobs (or tasks) usually coded according to some established
-national or international classification system. Some JEMs also include 
-information for different calendar time periods." [@garcia2003glossary]
-
-[New advances in multi-pollutant analysis.]
-
-[Some of the hazards could create similar risks, if they converge on a common
-pathway of risk. For example, if pathway is through stress from property damage,
-individually and in the county, then several hazards could indeed lead to the
-same health outcomes (e.g., PTSD, other psychological markers, risk of acute
-cardiovascular outcomes, etc.). One example is extreme level of destruction both
-from wind (Hurricane Andrew, Hurricane Hugo), from river flooding associated
-with among of precipitation (Hurricane Floyd, Hurricane Florence), and from
-flash flooding associated with the rate of precipitation (Virginia storms).]
-
-[A better future direction might be through regression models with multiple
-explanatory pathways. This allows the model to fit different weights to each.
-Other models can also be explored, like regression trees, which might have an
-easier time capturing non-linearity (if exposure is included as a continuous
-value, as with wind and rain) and interactions when there are co-exposures to
-two or more hazards. Both of these are able (and will) either exclude or give a
-very low weight to an explanatory variable that does not help much in explaining
-variation in the health outcome.]
+epidemiology. We have added text to the manuscript on these points:
+
+> "Finally, we investigated how well exposure assessment agreed across these
+metrics. **These results can help epidemiologists answer several key questions as
+they interpret previous research and design new studies. For example, if a
+single storm hazard has been used to measure exposure in an epidemiological
+study, can the result be interpreted as an association with tropical cyclone
+exposure in general, or should it be limited to represent an association with
+that specific storm hazard? If a study investigates the association between a
+single storm hazard and a health outcome, could this estimate be confounded by
+other storm hazards? When planning a new study that will incorporate several
+storm hazards in a mixtures approach, will there likely be numerical problems in
+fitting the model from multicollinearity, or difficulties in disentangling the
+role of separate hazards?**"
 
@@
 
-We hope that these results can help epidemiologists who are planning new
-single-storm or multi-year studies of tropical cyclones and associated health
-risks. For planning a multi-year study, it is helpful to know how many exposures
-the study might incorporate, as this can help in giving a general idea of
-whether the study might have adequate statistical power to investigate a certain
-research question. Similarly, for single-storm studies, it is helpful to know
-how widespread exposure can be for a specific hazard during a very large storm.
-This table provides some basic summaries of the extent of exposure to different
-tropical cyclone hazards, both on average for the storms in the study and also
-for those storms that were most extensive.
+We included this table with the intent that these results can help
+epidemiologists who are planning new single-storm or multi-year studies of
+tropical cyclones and associated health risks. For planning a multi-year study,
+it is helpful to know how many exposures the study might incorporate, as this
+can help in giving a general idea of whether the study might have adequate
+statistical power to investigate a certain research question. Similarly, for
+single-storm studies, it is helpful to know how widespread exposure can be for a
+specific hazard during a very large storm. This table provides some basic
+summaries of the extent of exposure to different tropical cyclone hazards, both
+on average for the storms in the study and also for those storms that were most
+extensive.
 
@@
 In the revised version, the x-axis has been labelled to clarify that this axis is showing results based on the wind radii dataset described in the text, in response to Reviewer 3's Comment 7.
-The revised version also uses a more focused range for the x axis, as requested in EHP's Request 1.
-Figure caption in the revised manuscript: \\textit{Comparison of two sources of wind exposure estimates: (1) modeled peak sustained surface wind and (2) estimates based on HURDAT2's wind speed radii.
-Each point represents a storm, with the x-axis giving the percent of counties classified in the same category of peak sustained surface wind (< 34 kt; 34--49.9 kt; 50--63.9 kt; $\\ge$64 kt) by both sources of data.
-The color of each point gives the number of study counties that were exposed to peak sustained surface wind of at least 34 kt (based on modeled wind).
-Estimates are shown for study storms since 2004, the earliest year for which post-storm reanalysis wind speed radii are routinely available in HURDAT2, and for which at least one study county had a peak sustained wind of $\\ge$34 kt based on the post-storm wind radii.}", fig.show = "hold", out.width="50%"}
+The revised version also uses a more focused range for the x axis, as requested in EHP's Request 1.", fig.show = "hold", out.width="50%"}
 include_graphics("../figures/windcomparison.pdf")

### 2020-09-10 ehp_revision/writing/ehp_methods.tex
Work some more on revisions

@@
 
-Finally, we investigated how well exposure assessment agreed across 
-these metrics. For each storm we investigated the degree to which the 
-set of counties assessed as exposed based on one metric overlapped with the set
-assessed as exposed based on each other metric, including a metric that used
-distance of a county from the storm track as a proxy for exposure to storm hazards. 
-We calculated the within-storm
-Jaccard index ($J_s$)~\parencite{jaccard1901distribution, jaccard1908nouvelles} between
-each pair of exposure metrics. This measures similarity
-between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical cyclone~$s$ as the
-proportion of counties in which both of the metrics classify the county as
-exposed out of all counties classified as exposed by at least one of the
-metrics:
+Finally, we investigated how well exposure assessment agreed across these
+metrics. \textbf{These results can help epidemiologists answer several key
+questions as they interpret previous research and design new studies. For
+example, if a single storm hazard has been used to measure exposure in an
+epidemiological study, can the result be interpreted as an association with
+tropical cyclone exposure in general, or should it be limited to represent an
+association with that specific storm hazard? If a study investigates the
+association between a single storm hazard and a health outcome, could this
+estimate be confounded by other storm hazards? When planning a new study that
+will incorporate several storm hazards in a mixtures approach, will there
+likely be numerical problems in fitting the model from multicollinearity, or
+difficulties in disentangling the role of separate hazards?}
+
+For each storm we investigated the degree to which the set of counties assessed
+as exposed based on one metric overlapped with the set assessed as exposed
+based on each other metric, including a metric that used distance of a county
+from the storm track as a proxy for exposure to storm hazards.  We calculated
+the within-storm Jaccard index ($J_s$)~\parencite{jaccard1901distribution,
+jaccard1908nouvelles} between each pair of exposure metrics. This measures
+similarity between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical
+cyclone~$s$ as the proportion of counties in which both of the metrics classify
+the county as exposed out of all counties classified as exposed by at least one
+of the metrics:
 

### 2020-09-10 ehp_revision/writing/ehp_response_darlings.tex
Work some more on revisions

@@
 curation and publication needs." [@rathje2017designsafe]
+
+# R2 Comment 1
+
+A tropical cyclone brings a number of physical hazards, including the four 
+included in our study---wind, rain, flooding, and tornadoes. These separate 
+hazards may each cause health impacts and in some cases might act together 
+in synergy in elevating health risks. For example, severe wind and rain 
+together might create a higher probability of power outages, since the rain 
+could soften the ground and make it more likely that the wind can uproot 
+trees. These physical hazards tend to follow different patterns, and this is
+something that's intuitively understood by atmospheric scientists who study 
+the storms. Storm winds tend to decrease rapidly once a storm's center is 
+over land, for example, because the storm loses its main energy source---warm
+water. Rainfall can persist, and might be particularly severe when the storm 
+hits landscape that is not smooth, like the Appalachian Mountains. Tornadoes
+can be generated in outer bands of the storm, far from the storm's center, 
+but this tends to only happen in the southern half of the US. Rain and flooding
+will sometimes align spatially, but you can also have severe flooding downstream
+from the area that receives the most severe rain, particularly in coastal areas
+with large networks of waterways.
+
+None of these patterns would be particularly surprising to an atmospheric scientist, 
+but many epidemiologists might not be aware of them, or of how they play into
+exposure assessment for tropical cyclone epidemiology. Many environmental 
+epidemiologists have a training that was focused more on epidemiology and 
+biostatistics and rely on collaborations with experts in engineering,
+atmospheric science, or environmental health to ensure that they understand
+and are appropriately applying exposure assessment. In the case of tropical 
+cyclones, it is clear that they generate a mixture of physical hazards. 
+There has been a lot of methodological development and thought recently in 
+environmental epidemiology about how to study mixtures of chemicals---for example, 
+the mixture of air pollutants that a person might be exposed to, rather than 
+studying the individual effects of single air pollutants. 
+These mixture methods require an understanding of how the different components
+are associated with each other. In the case of binary, rather than continuous,
+exposure, this will be expressed as agreement among the different exposures.
+
+First, if exposure to a pair of different hazards strongly agrees (i.e., when a
+county is exposed to one hazard, they are typically also exposed to a second),
+this would have several implications for epidemiological research, especially
+for multi-year studies:
+
+- From a statistical modeling point of view, if you tried to include each hazard
+as a separate independent variable, you would run into problems with
+collinearity if fitting the study data with a generalized linear model or other
+regression model. One implication of this is that the confidence intervals on
+effect estimates for the two variables would be inflated, potentially a lot,
+because the model would struggle to figure out which of the two variables the
+weight should be placed on, and so this uncertainty would be reflected in the
+size of the associated confidence intervals.
+- If you see an effect, it would be very hard to try to determine if the causal
+pathway goes through the first hazard or the second. To mitigate effects, it is
+important to understand the pathway through which they happen, and it might be
+easier to get clues towards pathways if the hazards are better separated. By
+analogy, if you were able to conduct a controlled experiment to look at the
+effects of each hazard, jointly and in combination, you would design separation
+in the assignment of these treatments, so you could distinguish between the two
+factors. For tropical cyclone epidemiology, we are limited to an observational
+analysis, but the area of causal inference is doing a lot to explore how we can
+design observational studies and analyze data to come closer to what we might be
+able to see with a randomized control study. To determine which of those methods
+could be used for tropical cyclone epidemiology, it is important to understand
+the interplay of these patterns of exposures.
+- If this is the case, and a study is conducted on a single hazard of the storm, 
+then there is the risk that results have been confounded by a different hazard
+of the storm. 
+- On the positive side, in this case the researcher may be able to fit a simpler
+statistical model to estimate the joint effects of exposure to the storm as a
+whole. Statistical models of mixtures can be complex to lay out, and their 
+results can be hard to interpret and explain to others. If all hazards of a 
+tropical cyclone follow very similar spatial patterns, then it would be 
+reasonable to model tropical cyclone exposure based on one of the hazards
+(and it agreement is strong enough, it wouldn't make much difference which 
+of the hazards was used for this). This would allow for the use of very 
+straightforward statistical modeling, drawing on well-developed models currently
+used in environmental epidemiology to study things like the health risk associated
+with exposure to a single pollutant or a single physical hazard like outdoor 
+temperature.
+
+Second, if a pair of hazards tend to have little spatial overlap among 
+counties, this also brings some implications for epidemiological study design
+and statistical analysis:
+
+- In this case, a single-hazard analysis could miss a lot of the risk created
+by the storm as a whole. This is a concern because there have been some cases
+of studies that have assessed exposure based on a single storm, typically wind.
+[Grabich example, Darren Sun example]. We are aware of one other paper 
+currently under review that takes this same approach of focusing on the wind
+hazard of these storms. .... In interpreting these papers, it is helpful to
+know if this hazard can stand as a reasonable proxy to the collection of 
+physical hazards of the storm. If not, the paper must be interpreted as 
+investigating
+
+In the past, much of tropical cyclone epidemiology focused on cases studies of
+single storms. Multi-year studies are becoming more common. Both types of
+studies are important in exploring both typical and unusual health risks that
+can be caused by these storms. Therefore, it's important to understand the
+patterns between these different hazards both for single storms and across many
+storms. In our analysis, we find that these hazards typically are not in strong
+agreement across most storms, suggesting that mixture methods could be
+meaningfully applied in most cases, and that there may often be the chance to
+separate the role of different hazards through such models. However, we also
+find that there are a few storms for which the hazards follow very similar
+spatial patterns (Hurricane Floyd in 1999, for example), and in these cases
+there may be problems of multi-collinearity and non-separability of different
+hazard effects if a researcher tries to fit a model that incorporates each
+separate hazard.
+
+Issues:
+
+1. **Interpretation** Effect of one hazard, or storm as a whole?
+
+2. **Confounding**
+
+3. **Multi-collinearity**
+
+4. **Disentangling role of different hazards**
+
+5. **Statistical power** (to separate effects of different hazards)
+
+------------------------------------------------------------------------------
+
+2. These results will not be surprising to an atmospheric scientist / disaster
+researcher, but there are environmental epidemiologists who have done research
+on other exposures and are now starting to study tropical cyclones (session at
+ISEE, for example). It is useful for epidemiologists entering into tropical
+cyclone research to understand these facets of tropical cyclone exposure,
+including the distinction between the very coastal wind exposures and other
+hazards that tend to reach further inland.
+3. There are epidemiological studies that have used only one hazard for exposure
+assessment. This has most often been the case for wind. If this exposure
+assessment is meant to capture "tropical cyclone exposure" in general, with all
+the hazards that are related to tropical cyclones, then there is the chance for
+exposure misclassification if this single hazard is used.
+4. In the past, it was a very complex issue to try to pull together data on all
+these hazards. We agree that now that we have provided a dataset with all of
+them, epidemiologists should be thinking about multi-hazard analyses. However,
+there still will be some barriers to that. [Recent work on multi-pollutant
+exposures and the complexities associated with that.] If one of the hazards
+dominates in explaining associated health risk, then in some cases it might be
+reasonable to focus on that hazard, but in that case it is critical, based on
+our results, to remember that the exposure assessment is not capturing all
+hazards of the storms.
+
+The largest-scale county-level exposure study is likely that of @zandbergen2009,
+which estimated exposure in US counties to all US landfalling Atlantic basin
+tropical storms between 1851 and 2003, using both distance and an exposure
+metric that incorporated distance and windspeed. They used this exposure
+evaluation to create maps of total exposure to tropical storms within US
+counties, as well as to explore associations between a county's long-term
+exposure to tropical storms and its location, distance from the coast, size, and
+shape [@zandbergen2009]. @kruk2010 explored exposure to hurricane-related winds
+in the United States, including inland areas, for 1900-2008.
+
+Geographical patterns can differ for these different
+storm hazards, and so exposure assessment based on one hazard could create
+exposure misclassification if the pathway for health risks is through a
+different storm hazard.
+
+A tropical cyclone's high winds can bring health risks and property damage
+through structural damage of houses and other buildings, falling trees, and
+wind-borne debris [@rappaport2000] and cause power
+outages [@liu2005; han2009], which introduce a number of threats to
+human and ecological health, including water quality risks if the outage affects
+wastewater treatment plants [@mallin2006].  Other risks can exist
+without severe wind; for example, one study found that most of the direct
+hurricane-related deaths in the US between 1970 and 1999 occurred in cases
+when wind was below hurricane strength, including for Tropical Storms Charley
+in 1998 and Alberto in 1994 [@rappaport2000].  Tropical cyclones can
+produce excessive rain, especially in certain topographies (e.g., near
+mountains), so counties well inland sometimes experience more extreme rain than
+coastal counties. Flood risks from tropical cyclones can result from this rain,
+although the two risks are not perfectly correlated [@chen2015]. In the
+US, over half of hurricane-related direct deaths from 1970 to 1999 from
+Atlantic basin storms were a result of freshwater
+flooding [@rappaport2000]. Flooding can also degrade water
+quality [@mallin2006], which can threaten both human and ecological
+health.
+
+If a study misclassifies exposure to the hazard or hazards that cause the
+health risk being studied, the study will generate biased estimates of tropical
+cyclone risks and impacts. Further, when different studies use different
+methods to assess exposure to tropical cyclones, their results are difficult to
+meaningfully compare and aggregate. If different methods identify
+similar sets of communities as ``exposed'',  these concerns are less serious.
+However, if different methods differ substantially in which communities they
+identify as ``exposed'', it makes it very important that epidemiological
+studies are thoughtful in how they assess exposure.
+
+These analyses can also help researchers compare and interpret results from
+previous epidemiological studies that have assessed tropical cyclone exposure in
+different ways, helping to [synthesize] previous epidemiologic results to form a
+better picture of how these storms can affect human health.
+
+However, these frequency maps, together with evidence from
+specific tropical cyclones (Figures \ref{fig:ivanexposure}
+and \ref{fig:jaccard}),  do illustrate the potential for strong differences in
+spatial patterns in tropical cyclone exposures, depending on which tropical
+cyclone hazards are considered.  A few previous studies have sought to
+determine county-level exposure to tropical cyclones over multi-year periods,
+including @zandbergen2009, which estimated exposure in US
+counties to all US landfalling Atlantic-basin tropical cyclones
+between 1851 and 2003, using both a distance-based metric and a metric that
+combined distance and windspeed, and [@kruk2010], which explored
+exposure to hurricane-related winds in the US, including inland areas,
+for 1900--2008.  Our results suggest that such exposure assessments may
+perform well in capturing some tropical cyclone hazards (e.g., wind), but
+likely miss other potentially dangerous tropical cyclone exposures, especially
+for hazards that repeatedly threaten northern or inland counties (e.g., rain,
+flooding).
+
+Therefore, use of distance to assess tropical cyclone exposure for impact
+studies could result in problematic exposure misclassification, which could mask
+true associations, even strong associations, between tropical cyclone exposure
+and outcomes of interest in impact studies [@savitz2016interpreting;
+armstrong1998effect]. The use of a single hazard-based metric (e.g., wind) could
+cause similar problems if the impact is driven, at least in part, by a different
+hazard or by multiple hazards of the tropical cyclone.
+
+If measurements of one storm hazard are used as a proxy to capture exposure to
+a different hazard, or a non-hazard proxy like distance to the storm track is
+used to assess exposure to storm hazards, the resulting exposure
+misclassification could be differential (i.e., associated with the outcome of
+interest or with factors associated with risk of the outcome of interest).
+
+Therefore for most tropical cyclones, since exposure assessments often differ
+across storm hazards, as well as between each storm hazard and a distance-based
+proxy measurement, exposure misclassification is a potential risk.
+
+Further, we investigated agreement in exposure assessment across these metrics
+when applied individually. We found large differences in which counties are
+exposed to different hazards of tropical cyclones and that distance is, at
+best, a moderate, and often a very poor, surrogate for exposure to the specific
+tropical cyclone hazards of high wind, extreme rainfall, flooding, and
+tornadoes. Use of distance as a surrogate for any of these hazards, or exposure
+assessment based on one hazard when the pathway for health impacts is in part
+or full through another storm hazard, could lead to exposure misclassification.
+In the case of tropical cyclone risk and impact studies, including
+epidemiological studies, this would result in biased estimates.  Our findings
+highlight the importance of clarifying the potential pathway from tropical
+cyclone hazards to health impacts when conducting tropical cyclone
+epidemiological studies, and then basing exposure assessment on these hazards.
+
+---------------------------------------------------------------------------
+
+It is obvious that it is typically better, but in cases where the proxy is much
+more practical to measure, it may need to be more than just a bit better, but
+instead substantially better. Also, there could be some cases where the proxy
+might actually be better---for example, if the proxy can be measured with very
+low level, while the direct exposure measure might only be measurable with high
+error. In that case, the proxy could be the better metric and might provide a
+more precise estimate of the association with health risk.
+
+[Env. Epi. book---advantages and disadvantages of different types of measures]
+
+We agree that it is helpful to clarify for readers that these independent
+hazards can come together in a storm event at different intensities and at
+different time windows. We have added some text on this:
+
+[Added text]
+
+However, we disagree that this should be down as an alternative to calculating
+agreement. Epidemiologists need to understand how these exposure measurements
+agree to help in planning studies and statistical approaches to interpreting the
+resulting data. In this sense, "agreement" is providing a measure similar to
+correlation for continuous variables; our choice of agreement with the binary
+exposure classifications reflects the more common use of binary classifications
+for most tropical cyclone epidemiology to date.
+
+Epidemiological studies might focus on tropical cyclone exposures at two levels.
+First, they might wish to focus on a specific hazard. For example, a study might
+hypothesize that mold growth following flooding from a tropical storm might
+increase risk of respiratory health outcomes in the months after a storm. In
+this case, the clear causal pathway recommends a focus on one of the storm
+hazards, rather than all storm hazards. However, other storm hazards could
+confound the association between flood exposure and respiratory health risk if
+the other hazards are ignored *and* if they are associated with both the
+exposure of interest (storm-associated flooding) and the health outcome. For
+example, if wind exposure assessments are in high agreement with flood exposure
+assessments for storms, and if wind affects risk of respiratory outcomes by
+causing extensive property damage and creating extreme psychological stress,
+which can have repercussions on respiratory health, then the wind hazard could
+confound the estimated association between storm-related flood exposure and the
+health outcome.
+
+This bias would be higher when the confounder and exposure of interest are more
+strongly correlated (for continuous variables) or in higher agreement (for
+categorical variables). Conversely, if the exposure and the potential confounder
+are, in practice, only weakly associated, then the potential for confounding is
+low. The Jaccard index is one measure of similarities between categorical
+(including binary) coefficients.
+
+Another consideration is collinearity. If one hazard is the exposure of interest
+and another hazard is a potential confounder, then the model should adjust the
+confounder or it will be misspecified and be prone to bias. However, if the
+exposure of interest and the potential confounder are very strongly
+correlated/in agreement, then including both as independent variables in a
+regression model often causes numerical problems in fitting the regression
+model, including instability in the estimated coefficients and inflated
+variance. In very extreme cases, the statistical software may not be able to
+invert the matrix [? only for linear regression?].
+
+Therefore, it is helpful for epidemiologists to understand if these independent
+storm-associated hazards tend to be strongly associated (in high agreement),
+both within specific storms (for single-storm studies) and on average across
+most large storms (for multi-storm studies). This knowledge will help them in
+identifying potential confounders if they are interested in studying the effects
+of a single hazard of the storm and also in determining if control for these
+potential confounders among other storm hazards might cause numerical problems
+or other collinearity-associated problems when fitting a regression model.
+
+If a researcher wishes to study the storm as a whole, then the researcher 
+could either seek a single measure to capture exposure to the storm as a whole
+or investigate more complex models that model the influence of the storm through
+pathways of several different hazards.
+
+If a researcher wants a single measure for the storm as a whole, then it is
+helpful to know how well exposure assessments based on the hazard components of
+the storm agree with each other. If there is strong agreement, then any of those
+hazards could be used as a primary marker of exposure assessment. This is
+conceptually similar to checking for observer / rater agreement.
+
+> "When the row and column variable srepresent different observers' rating the
+same subjects or objects, interest is focused on *observer agreement* rather
+than mere association. In this case, measures and tests of agreement provide a
+method of assessing the reliability of a subjective classification or assessment
+procedure." [@friendly2015discrete]
+
+> "In assessing the strength of *agreement* we usually have a more stringent
+criterion than in measuring the strength of *association*, because observers
+ratings can be strongly associated without strong agreement. For example, one
+rater could use a more stringent criterion and thus consistently rate subjects
+one category lower (on an ordinal scale) than another rater. More generally,
+measures of agreement must take account of the marginal frequencies with which
+two raters use the categories. If observers tend to use the categories with
+different frequency, this will affect measures of agreement."
+[@friendly2015discrete]
+
+When you have an array of two or more exposures that are strongly correlated,
+including them together in a regression model can cause problems from
+collinearity [@schisterman2017collinearity].
+
+> "Environmental exposures ... often consist of an array of related individual
+factors that may be highly correlated with each other, causing concern about the
+impact of collinearity when attempting to identify individual effects."
+[@schisterman2017collinearity]
+
+**Commonly have multiple factors of interest, correlated**
+
+In some areas of epidemiology, researchers have moved toward the idea of a 
+"web of causality":
+
+> "'Multiple causation' is the canon of contemporary epidemiology, and its
+metaphor and model is the 'web of causation'. Expressed through the notion
+of 'multifactorial etiology' and embedded in the statistical techniques
+of 'multivariate analysis', the belief that population patterns of health 
+and disease can be explained by a complex web of numerous interconnected 
+risk and protective factors has become one of this discipline's central 
+concepts. Equally entrenched is the corrolary that epidemiology's power
+to improve the public's health rests upon its ability to identify---and predict
+the results of breaking---selected strands of this causal web." [@krieger1994epidemiology]
+
+This is in contrast to focusing on single factors:
+
+> "Expressly challenging the still-pervasive tendency of epidemiologists to
+think in terms of single 'agents' causing discrete diseases, the provocative
+metaphor and model of the 'web' invited epidemiologists to embrace a more
+sophisticated view of causality." [@krieger1994epidemiology]
+
+In this framework, epidemiology helps by identifying where it would help to
+break strands of the web:
+
+> "Using this model, MacMahon et al. drew several important inferences about
+prevention and research that remain part of epidemiologic thinking to this day.
+Arguing that 'to effect preventative measures, it is not necessary to understand
+causal mechanisms in their entirity', they stated that 'even knowledge of one
+small component may allow some degree of prevention' since 'wherever the chain
+is broken the disease will be prevented'." [@krieger1994epidemiology]
+
+In other words, if you can identify *necessary* causes, you can craft 
+*practical* interventions.
+
+In some fields of epidemiology (e.g., nutritional epidemiology), it is common 
+to study cases where there are multiple potential risk factors, and these 
+factors are often strongly correlated [@schisterman2017collinearity].
+
+> "The dynamics of many important ecological and evolutionary processes are 
+often influenced by multiple interacting factors, and attempting to understand 
+the dynamics of such systems presents difficult technical and philosophical 
+problems." [@petraitis1996inferring]
+
+> "Often in practice, investigators are faced with decisions of how to handle
+highly correlated variables." [@schisterman2017collinearity]
+
+**Methods to tackle multiple causality**
+
+When the aim is to explore the influences of multiple associated factors, there
+are some different strategies that can be used. These include multiple
+regression and path analysis:
+
+> "Evolutionary ecologists have recently turned to path analysis, multiple
+regression, and related techniques to analyse systems of multiple causality.
+While these techniques are often seen as competitors, they are, in fact,
+related, and as a result the problems of these techniques are similar."
+[@petraitis1996inferring]
+
+**Implications of collinearity**
+
+> "Two variables are defined as collinear if one can be expressed as an exact
+or near linear combination of the other. ... In general, collinearity between
+variables is described by the magnitude of their correlation." 
+[@schisterman2017collinearity]
+
+It can complicate inference of model coefficients. This is always a concern for 
+inferential models. For predictive models, it can be a problem if you later use
+the model to predict in a case where the correlation structure is different than
+in the training data.
+
+> "The dependence of regression and path coefficients on the correlation
+structure is often glossed over. Yet this dependence means that inferences about
+the strengths of path coefficients are conditional upon the correlational
+structure among the predictor variables. If the correlation structure of the
+sample does not match the correlational structure of the population then
+regression and path coefficients will not estimate relative strengths
+accurately. This can occur if sampling is not random or if analysis is based on
+data from experiments with factorial designs, which break the correlational
+structure amnog the predictor variables." [@petraitis1996inferring]
+
+These implications can depend on whether the modeling goal is predictive or 
+inferential.
+
+> "When the goal is to choose covariates based on their predictive value, 
+concerns about collinearity stem from the desire to improve parameter estimates'
+precision, which can be inflated when highly correlated covariates are included
+in the model." [@schisterman2017collinearity]
+
+> "In a causal framework, on the other hand, interest is often centered on a
+specific effect from a correctly specified causal model, even in the presence of
+highly correlated data, despite a potential loss of statistical efficiency.
+Under such a viewpoint, bias due to misspecification of the causal associations
+is a more devastating result than reduced precision of parameter estimates
+correctly estimated. Furthermore, a precisely but incorrectly estimated
+parameter may lead to invalid inferences. ... Relying on conventional
+variable-selection algorithms and standard analyses can ignore important
+confounders and even produce extremely imprecise confidence intervals."
+[@schisterman2017collinearity]
+
+The amount of bias in the unadjusted model (in terms of bias from an unadjusted
+confounder) depends on the strength of association between the exposure of
+interest and the potential confounder [@schisterman2017collinearity].
+
+> "This is especially important for the confounding variable scenario, where 
+correctly specifying the model may result in adjusting for a nearly 
+collinear variable." [@schisterman2017collinearity]
+
+The degree of agreement will influence the degree of confounding in an 
+unadjusted model, for either one or two confounders:
+
+> "In this confounding scenario, increasing correlation exacerbates the bias
+due to misspecification, similar to results for a single confounder."
+[@schisterman2017collinearity]
+
+Extreme collinearity can create numerical problems
+[@schisterman2017collinearity].
+
+> "Under the extremely unlikely scenario of extreme collinearity [(abs. rho of
+0.999 or higher)], the numerical matrix inversion required for estimation fails,
+resulting in unstable and biased predictors." [@schisterman2017collinearity]
+
+> "Only for extreme values of rho cloes to 1 or -1 is beta-squareed
+nonidentifiable due to numerical instability; otherwise, the correctly specified
+model produces the most desirable (e.g., unbiased) estimates of the total effect
+of the exposure on the outcome." [@schisterman2017collinearity]
+
+Collinearity can cause imprecision in coefficient estimates, even if those
+estimates are unbiased [@schisterman2017collinearity].
+
+> "Multicollinearity: A term that refers to correlation among the independent
+variables in a multiple regression model; it is usually invoked when some 
+correlations are 'large', but an actual magnitude test is not well defined.[1]"
+[@gunasekara2008glossary]
+
+> "Collinearity occurs in multiple regression models when two (or
+more---multicollinearity) exposure variables are included in the model but are
+so similar to each other that they are essentially measuring at least part of
+the same thing. Collinearity creates problems with interpreting the analysis by
+affecting the standard errors of the variables (as it increases variance)[24]
+and by causing biased estimates for one or both of the collinear terms
+(sometimes dramatically so) [25]. If there is perfect collinearity between two
+variables then one should be dropped from the model. However, dropping exposure
+variables to reduce multicollinearity may lead to bias in the model from
+dropping useful information." [@gunasekara2008glossary]
+
+> "It might be argued in some cases that it is appropriate to include
+(collinear) variables in a model even when they are highly correlated, because
+the theoretical basis for including the variables is strong and the results of
+the model are consistent with expectations [10]. In such cases, models with and
+without the variables would be tested and compared. Such consideraion of
+alternative models tends to be described as an issue of (mis)-specification in
+econometrics. However, in epidemiology this tends to be described in terms of
+mediating and *intermediary variables*. Thus, an epidemiologist might argue for
+including an intermediate variable in a model (and testing it against a model
+without the variable) because including that variable might elucidate direct and
+indirect pathways between the exposure and outcome variables. An econometrician
+might do the same thing but describe this in terms of finding the correct model
+specification." [@gunasekara2008glossary]
+
+**Techniques to address collinearity**
+
+There are study designs that might help to separate different effects. For
+example, some matching designs might help in creating a subset of data in which
+the association that exists in the full dataset is broken (e.g.,
+@leal2012multicollinearity).
+
+> "Matching is typically employed to reduce model dependence and estimate 
+associations in a more empirical way than would be necessary without 
+matching. In line with this practice, matching was employed as a diagnostic
+tool to verify whether the effects of environmental variables that are highly 
+correlated with each other could be disentangled." [@leal2012multicollinearity]
+
+You could also explore techniques of dimension reduction, to create a smaller
+subset of independent variables for which the correlation is broken. For
+example, you could use PCA [@leal2012multicollinearity].
+
+There are other techniques that allow some bias in estimates in an attempt to
+reduce problems from collinearity, like ridge regression
+[@petraitis1996inferring].
+
+> "Several biased estimation techniquest ahve been proposed as a way to combat
+collinearity (Myers 1990). These are controversial methods. Ridge regression,
+for example, biases the regression coefficients in order to improve the model's
+ability to predict the criterion variable. The estimates of the individual
+regression coefficients may be very biased even though when taken together they
+provide a very good prediction of the criterion variable."
+[@petraitis1996inferring]
+
+> "When correct specification of the models still leads to 
+problems for conventional regression methods, one can apply modern techniques
+such as Bayesian hierarchical models with shrinkage estimators."
+[@schisterman2017collinearity]
+
+**Are effects of different factors 'separable'?**
+
+Epidemiology often aims to inform policy decisions. In this case, it can be
+important to characterize the separate effects of different associated factors
+that contribute to the health risk [@leal2012multicollinearity].
+
+> "To develop efficient public health interventions addressing the obesity 
+epidemic, it is important to identify exactly which aspects of the
+environment influence obesity risk. For example, demonstrating that the density 
+of fast-food restaurants was associated with obesity risk and demnstrating that
+the density of sports facilities was associated with obesity risk would lead to 
+different interventions." [@leal2012multicollinearity]
+
+However, it can be very difficult to separate the effects of different factors
+if those factors are strongly correlated [@leal2012multicollinearity]:
+
+> "However, many neighborhood characteristics, especially those related to the
+densities of physical features and services, are strongly correlated with each
+other. Therefore, it is not clear whether it is even possible to disentangle the
+effects of these different environmental dimensions, even if they are
+hypothesized to influence obesity risk through distinct causal pathways."
+[@leal2012multicollinearity]
+
+-----------------------------------------------------------------------------
+
+As we note in our general response to your comments, we agree in general with
+your instinct here. However, there are some subtle points to note.
+
+First, we understand the potential appeal of creating a single exposure summary
+score based on a summation of exposures. This simplifies some of the statistical
+modeling, for example, and also provides a single exposure estimate to provide
+as the key study conclusion. However, in many areas of tropical cyclone
+epidemiology, it is unlikely that exposure to each type of hazard should be
+given equal weight. For example, an analysis of respiratory health in the months
+after the storm might hypothesize a strong connection with flooding, and smaller
+connections with rainfall and winds (through the pathway of property damage).
+Use of a pre-defined summation of exposure to each hazard would mask some of the
+potential association.
+
+While a summary statistic could be defined that adds a weight to each exposure,
+it would often be unclear *a priori* what weight each hazard exposure should be
+given. A better method might be to use a regression framework, where each hazard
+exposure is included as a separate explanatory variable, and the model uses the
+observed data to determine the appropriate weight for each hazard in terms of
+modifying risk to the outcome of interest. To explore the influence of multiple
+concurrent hazards, interaction terms could be added.
+
+However, if our analysis had found that hazard exposures follow very similar
+patterns, in terms of the counties exposed during a storm, then this strong
+agreement would result in a number of problems in fitting and interpreting these
+types of models. First, strong agreement in two or more of the explanatory
+variables in a regression model leads to collinearity, which among other
+implications often leads to a large inflation of confidence intervals for the
+effect estimates of the correlated explanatory variables. Further, if the hazard
+exposures were in strong agreement, then it would be difficult or impossible to
+disentagle the effects of the separate hazards, as a strong effect for one would
+be impossible to interpret with confidence, since the strong agreement in the
+data (and resulting low co-variability in the variables in the data) could cause
+the model to mistakenly attribute the effect of one hazard to the other.
+Further, it may become impossible in this case to estimate synergistic effects
+between the two, as an indicator of an interaction between the two exposures
+would be almost collinear with the two main effects.
+
+In our analysis, we find that this is usually not the case for the four tropical
+cyclone hazards that we consider. Instead, we find that for most storms, the
+hazards are fairly well-separated in terms of the counties they effect. There is
+a particularly strong pattern where inland counties can be exposed to extreme
+rain and flooding, but not wind. This separation opens the door for some
+interesting study designs and statistical modeling. For example, these patterns
+suggest that a study could investigate whether storm-associated rains alone
+create a health risk, as there are many exposures with only rain in inland
+counties, and so health associations could be explored here and contrasted with
+coastal areas that experienced both wind and extreme rain.
+
+-------------------------------------------------------------------------------
+
+> "A job exposure matrix (JEM) may be defined as a cross classification of jobs
+and occupational exposures. Some matrices are based on tasks, instead of jobs. 
+... Usually there is a group of experts who build the matrix based on their 
+own knowledge and experience on occupational exopsures. The experts should 
+assess exposure to an open or fixed list of occupational exposures (that can 
+include chemical, physical, biological, and/or psychosocial agents) for an
+open or fixed list of jobs (or tasks) usually coded according to some established
+national or international classification system. Some JEMs also include 
+information for different calendar time periods." [@garcia2003glossary]
+
+[New advances in multi-pollutant analysis.]
+
+[Some of the hazards could create similar risks, if they converge on a common
+pathway of risk. For example, if pathway is through stress from property damage,
+individually and in the county, then several hazards could indeed lead to the
+same health outcomes (e.g., PTSD, other psychological markers, risk of acute
+cardiovascular outcomes, etc.). One example is extreme level of destruction both
+from wind (Hurricane Andrew, Hurricane Hugo), from river flooding associated
+with among of precipitation (Hurricane Floyd, Hurricane Florence), and from
+flash flooding associated with the rate of precipitation (Virginia storms).]
+
+[A better future direction might be through regression models with multiple
+explanatory pathways. This allows the model to fit different weights to each.
+Other models can also be explored, like regression trees, which might have an
+easier time capturing non-linearity (if exposure is included as a continuous
+value, as with wind and rain) and interactions when there are co-exposures to
+two or more hazards. Both of these are able (and will) either exclude or give a
+very low weight to an explanatory variable that does not help much in explaining
+variation in the health outcome.]
 

### 2020-09-11 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 ```{r fig5, fig.cap="Reproduction of Figure 5 from the original manuscript (left) and revised version in the resubmitted manuscript (right).
-In the revised version, the figure legend was changed from '\\textbf{Hits} per county per decade' to '\\textbf{Exposures} per county per decade' (bold used to highlight change).}", fig.show = "hold", out.width="40%", fig.align = "center"}
+In the revised version, the figure legend was changed from '\\textbf{Hits} per county per decade' to '\\textbf{Exposures} per county per decade' (bold used to highlight change).
+Figure caption in the revised manuscript: \\textit{Figure caption.}", fig.show = "hold", out.width="40%", fig.align = "center"}
 include_graphics("../figures/averageexposureonly.pdf")
@@
 In the revised version, the x-axis has been labelled to clarify that this axis is showing results based on the wind radii dataset described in the text, in response to Reviewer 3's Comment 7.
-The revised version also uses a more focused range for the x axis, as requested in EHP's Request 1.", fig.show = "hold", out.width="50%"}
+The revised version also uses a more focused range for the x axis, as requested in EHP's Request 1.
+Figure caption in the revised manuscript: \\textit{Comparison of two sources of wind exposure estimates: (1) modeled peak sustained surface wind and (2) estimates based on HURDAT2's wind speed radii.
+Each point represents a storm, with the x-axis giving the percent of counties classified in the same category of peak sustained surface wind (< 34 kt; 34--49.9 kt; 50--63.9 kt; $\\ge$64 kt) by both sources of data.
+The color of each point gives the number of study counties that were exposed to peak sustained surface wind of at least 34 kt (based on modeled wind).
+Estimates are shown for study storms since 2004, the earliest year for which post-storm reanalysis wind speed radii are routinely available in HURDAT2, and for which at least one study county had a peak sustained wind of $\\ge$34 kt based on the post-storm wind radii.}", fig.show = "hold", out.width="50%"}
 include_graphics("../figures/windcomparison.pdf")

### 2020-09-13 ehp_revision/response_to_reviewers.Rmd
Continue work on revisions

@@
 - \newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
+- \usepackage{longtable}
+- \DeclareDelayedFloatFlavor{longtable}{table}
 ---
@@
 
-> **Text added to Discussion (added text in bold):** "Second, these data are
-aggregated to the county level. **This spatial scale allows for easy integration
-with health outcome data aggregated at the county level. Such data is often used
-for disaster epidemiology, as aggregated data may be easier to access than
-individual-level data, especially at a scale that covers many locations and a
-long time period and so allows higher statistical power and a broader range of
-exposure levels within the study [@wakefield2008overcoming].**
-
-> "**These ecological exposure estimates, however, set a common exposure level
+> **Text added to Discussion (added text in bold):** 
+"Second, these data are aggregated to the county level. **This spatial scale
+allows for easy integration with health outcome data aggregated at the county
+level. Such data is often used for disaster epidemiology, as aggregated data may
+be easier to access than individual-level data, especially at a scale that
+covers many locations and years and so allows higher statistical power and
+includes a broader range of exposure levels [@wakefield2008overcoming].**
+
+> "**Such ecological exposure assessment, however, sets a common exposure level
 throughout the county, ignoring within-county variability, even though such
-variability often exists. For some hazards, this within-county variation in
-exposure level could be stark. For example, tornadoes cause very localized
-damage, directly along the tornado's path---a tornado can destroy homes on one
-side of a street while leaving those on the opposite side untouched. Levels of
-other hazards, like storm-associated winds and rain, will also vary within a
-county, but typically with smoother variation. In particular, it will be
-unlikely that a county will have one area that is exposed to extremes of these
-hazards from a tropical cyclone while other parts of the county are completely
-unexposed, as both the wind fields and rain fields of tropical cyclones tend to
-be large in comparison to the size of a county.**
-
-> "**Aggregated data can be used to infer contextual level associations---for
+variability exists. For some hazards, this within-county variation could be
+stark. For example, tornadoes cause very localized damage, directly along the
+tornado's path---a tornado can destroy homes on one side of a street while
+leaving those on the opposite side untouched. Levels of other hazards, like
+storm-associated winds and rain, will also vary within a county, but typically
+with smoother variation. In particular, it will be unlikely that a county will
+have one area that is exposed to extremes of these hazards while other parts of
+the county are completely unexposed, as both the wind fields and rain fields of
+tropical cyclones tend to be large in comparison to the size of a county.**
+
+> "**Aggregated data can be used to infer contextual-level associations---for
 example, the association between county-level exposure to a storm hazard and
@@
 > Text added to the Discussion (added text in bold): 
-"**This dataset is limited to the contiguous US. Some exposure estimates would
-be fairly straightforward to extend internationally. The precipitation data is
-extracted and aggregated to the county level from a re-analysis data product.
-While the re-analysis product used here (NLDAS-2) only covers the contiguous US,
-other similar re-analysis products, as well as other types of precipitation
-datasets, have global coverage [@sun2018review]. The wind data are generated
-based on a model that is currently US-focused [@stormwindmodel], but could be
-extended to other areas, although this would require adding new land/sea masks
-within the associated software, as well as accounting for differences across
-storm basins in wind averaging periods [@harper2010guidelines] and in the
-direction of cyclonic winds in the Northern versus Southern Hemisphere. Further,
-since the core of the wind model was developed based on data from Atlantic-basin
-storms [@willoughby2006parametric], an extension to other areas should include
-separate validation and calibration to ensure it performs appropriately in those
-settings. Alternatively, other wind field modeling software is available that
-provides a global coverage, like Geoscience Australia's Tropical Cyclone Risk
-Model (http://geoscienceaustralia.github.io/tcrm/). For tornado and flood
-events, the data described in this paper drew on a US-focused storm events
-database, and so international extension of data on these hazards would require
-access to similar databases covering other countries. Finally, the relevant
-geopolitical boundaries to use for aggregation would vary by country (e.g.,
-municipalities in Mexico, districts in India).**"
+"**This dataset is limited to the contiguous US, and while expansion to global
+coverage would be useful and feasible, there would be some challenges. For
+precipitation data, the re-analysis product used here (NLDAS-2) only covers the
+contiguous US, although other re-analysis products, as well as other types of
+precipitation datasets, have global coverage [@sun2018review]. The wind data are
+generated based on a model that is currently US-focused [@stormwindmodel], but
+could be extended to other areas, although this would require adding new
+land/sea masks within the associated software, as well as accounting for
+differences across storm basins in wind averaging periods
+[@harper2010guidelines] and in the direction of cyclonic winds in the Northern
+versus Southern Hemisphere. Further, since the core of the wind model was
+developed based on data from Atlantic-basin storms [@willoughby2006parametric],
+an extension to other areas should include separate validation and calibration
+to ensure it performs appropriately in those settings. Alternatively, other wind
+field modeling software is available that provides a global coverage, like
+Geoscience Australia's Tropical Cyclone Risk Model
+(http://geoscienceaustralia.github.io/tcrm/). For tornado and flood events, the
+data described in this paper drew on a US-focused storm events database, and so
+international extension of data on these hazards would require access to similar
+databases covering other countries. Finally, the relevant geopolitical
+boundaries to use for aggregation would vary by country (e.g., municipalities in
+Mexico, districts in India).**"
 
@@
 "**The dataset we present focuses on the physical hazards of a tropical cyclone.
-However, health impacts will often some through indirect pathways, including
+However, health impacts will often come through indirect pathways, including
 through damage to property and infrastructure. In future work, it would be
@@
 > **Text added to the Methods of the revised manuscript (added text in bold):**
-"**By sharing this data as an R package, it is also accessible through 
+"We have posted most study data as an open-source R package (Anderson et al.
+2020). **By sharing this data as an R package, it is also accessible through
 cloud-based computing platforms that incorporate Jupyter notebooks, including
@@
 > "Finally, we investigated how well exposure assessment agreed across these
-metrics. **These results can help epidemiologists answer several key questions as
-they interpret previous research and design new studies. For example, if a
-single storm hazard has been used to measure exposure in an epidemiological
-study, can the result be interpreted as an association with tropical cyclone
-exposure in general, or should it be limited to represent an association with
-that specific storm hazard? If a study investigates the association between a
-single storm hazard and a health outcome, could this estimate be confounded by
-other storm hazards? When planning a new study that will incorporate several
-storm hazards in a mixtures approach, will there likely be numerical problems in
-fitting the model from multicollinearity, or difficulties in disentangling the
-role of separate hazards?**"
+metrics. **These results can help epidemiologists answer several key
+questions as they interpret previous research and design new studies. For
+example, if a single storm hazard has been used to measure exposure in an
+epidemiological study, can the result be interpreted as an association with
+tropical cyclone exposure in general, or should it be limited to represent an
+association with that specific storm hazard? If a study investigates the
+association between a single storm hazard and a health outcome, could this
+estimate be confounded by other storm hazards? When planning a new study that
+will incorporate several storm hazards, might there be variance inflation from
+multicollinearity or difficulties in disentangling the roles of separate
+hazards?**"
 
@@
 
-> **Text in revised manuscript (relevant addition in bold):**
+> **Text in revised Results (relevant addition in bold):**
 "The rainfall data are generally well-correlated with ground-based
 observations, but may sometimes underestimate very high rainfall values (Figure
-2), **and in some counties the correlation was substantially lower when 
+2). **Further, in some counties, the correlation was substantially lower when 
 considering only tropical cyclones with cumulative local rainfall of 75 mm
 or more (Supplemental Table 1)**."
+
+> **Text in revised Discussion (relevant addition in bold):**
+"The rainfall data are generally well-correlated with ground-based observations,
+but may sometimes underestimate very high rainfall values (Figure 2), **and in
+some counties the correlation was substantially lower when considering only
+tropical cyclones with cumulative local rainfall of 75 mm or more (Supplemental
+Table 1).**"
 
@@
 "We drew similar conclusions when we investigated all 46 tropical cyclones
-between 1996 and 2011 (when data for all five metrics were available) for which
-100 or more counties were exposed based on at least one metric (Figure 7;
-**Tables S3--S6 provide underlying numbers for the most extensive of
-these, the storms for which 250 of more counties were exposed based on at least
-one metric)**."
+between 1996 and 2011 (when data for all five metrics were available) for
+which~100 or more counties were exposed based on at least one metric (Figure7;
+**for the most extensive of these, storms for which 250 of more counties were
+exposed based on at least one metric, Tables S3--S6 provide underlying numbers
+comparing exposure assessments for the distance-based proxy to each hazard-based
+metric**)."
 
@@
 exposure to any of the other hazards considered, would often introduce exposure
-misclassification **(Tables S3--S6)**."
+misclassification **(Figure 7, Tables S3--S6)**."
 
@@
 
-\input{tables/thresholds_table.tex}
+\begin{longtable}{lp{40em}}
+\caption{Caption.
+These are provided for the three exposure metrics for which our database includes continuous data, and so a threshold is selected to determine binary exposure based on the metric.
+This table provides reasoning for the choice of threshold used in this paper as well as guidance on other thresholds that could be considered, depending on the hypothesized pathways for an epidemiological study.}
+\label{tab:thresholds} \\
+\hline
+Metric & Threshold choice\\
+\hline
+Distance & In the manuscript, distance-based exposure was assessed if the storm track came within 100 km of county population mean center.
+This threshold for distance from the storm center has been used in prior research as a proxy for exposure to hazards from the storm \cite{grabich2015measuring}.
+Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km \cite{mallin2006, quiring2011variations}, and dangerous winds can extend beyond these maximum winds.
+One study \cite{czajkowski2011} assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties.
+Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track \cite{czajkowski2011}.
+Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research \cite{grabich2015measuring, grabich2015, currie2013}.\\
+Rain & In the manuscript, rain-based exposure was assessed if the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
+This threshold...
+Other reasonable definitions could vary both the threshold of rainfall (in mm) and also the choice of days surrounding the storm to include when calculating the cumulative rainfall.
+Tropical cyclone precipitation arrives before the center of the storm \cite{zhou2017spatial}.
+A variety of definitions have been used to identify both extreme or heavy rain (whether associated with a tropical cyclone or not) and tropical cyclone--associated rain.
+Some definitions of extreme rain events are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours \cite{schumacher2006characteristics}.
+Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]).
+\cite{schumacher2006characteristics, schumacher2005organization, stevenson201410}.
+A study focused on the risks that result from extreme rainfall may therefore reasonably choose a higher threshold than the 75 mm threshold we use in this analysis.
+In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall \cite{zhou2017spatial} and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location \cite{feldmann2019estimation}.
+\\
+Wind & In the manuscript, wind-based exposure was assessed if modeled storm-associated peak sustained surface wind of $\ge34$ kts at the county’s population mean center.
+This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale.
+This limit is used as the outer limit in measuring storm size through the US National Hurricane Center's wind radii for tropical cyclone forecasts \cite{cangialosi2016examination}.
+Other thresholds could be selected based on other points on the Beaufort scale---for example, $\ge48$ kts for capturing storm-force winds or $\ge64$ kts for capturing hurricane-force winds.
+As a note, hurricane-force winds will be rarely experienced for counties, as it will likely only be observed for very severe storms and even for those for counties near the storm's landfall.
+Many presentations of the Beaufort wind scale include descriptions of the conditions that winds in each category would produce both over land and at sea.\\
+\hline
+\end{longtable}
 
@@
 
+\clearpage
+
 # References for the response {-#references-for-the-response}

### 2020-09-13 ehp_revision/tables/thresholds_table.tex
Work some more on response
Continue work on revisions

@@
-\begin{table}
+\begin{longtable}{lp{40em}}
 \caption{Caption.
@@
 This table provides reasoning for the choice of threshold used in this paper as well as guidance on other thresholds that could be considered, depending on the hypothesized pathways for an epidemiological study.}
-\centering
-\begin{tabular}{>{\bfseries\leavevmode\color{black}}l>{\raggedright\arraybackslash}p{40em}}
-\toprule
+\label{tab:thresholds} \\
+\hline
 Metric & Threshold choice\\
-\midrule
+\hline
 Distance & In the manuscript, distance-based exposure was assessed if the storm track came within 100 km of county population mean center.
-This threshold for distance from the storm center has been used in prior research as a proxy for exposure to hazards from the storm \cite{grabich2015measuring}.
-Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km \cite{mallin2006, quiring2011variations}, and dangerous winds can extend beyond these maximum winds.
-One study \cite{czajkowski2011} assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties.
-Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track \cite{czajkowski2011}.
-Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research \cite{grabich2015measuring, grabich2015, currie2013}.\\
+This threshold for distance from the storm center has been used in prior research as a proxy for exposure to hazards from the storm \parencite{grabich2015measuring}.
+Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km \parencite{mallin2006, quiring2011variations}, and dangerous winds can extend beyond these maximum winds.
+One study \parencite{czajkowski2011} assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties.
+Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track \parencite{czajkowski2011}.
+Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research \parencite{grabich2015measuring, grabich2015, currie2013}.\\
 Rain & In the manuscript, rain-based exposure was assessed if the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
@@
 Other reasonable definitions could vary both the threshold of rainfall (in mm) and also the choice of days surrounding the storm to include when calculating the cumulative rainfall.
-Tropical cyclone precipitation arrives before the center of the storm \cite{zhou2017spatial}.
+Tropical cyclone precipitation arrives before the center of the storm \parencite{zhou2017spatial}.
 A variety of definitions have been used to identify both extreme or heavy rain (whether associated with a tropical cyclone or not) and tropical cyclone--associated rain.
-Some definitions of extreme rain events are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours \cite{schumacher2006characteristics}.
+Some definitions of extreme rain events are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours \parencite{schumacher2006characteristics}.
 Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]).
-\cite{schumacher2006characteristics, schumacher2005organization, stevenson201410}.
+\parencite{schumacher2006characteristics, schumacher2005organization, stevenson201410}.
 A study focused on the risks that result from extreme rainfall may therefore reasonably choose a higher threshold than the 75 mm threshold we use in this analysis.
-In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall \cite{zhou2017spatial} and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location \cite{feldmann2019estimation}.
+In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall \parencite{zhou2017spatial} and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location \parencite{feldmann2019estimation}.
 \\
@@
 This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale.
-This limit is used as the outer limit in measuring storm size through the US National Hurricane Center's wind radii for tropical cyclone forecasts \cite{cangialosi2016examination}.
+This limit is used as the outer limit in measuring storm size through the US National Hurricane Center's wind radii for tropical cyclone forecasts \parencite{cangialosi2016examination}.
 Other thresholds could be selected based on other points on the Beaufort scale---for example, $\ge48$ kts for capturing storm-force winds or $\ge64$ kts for capturing hurricane-force winds.
@@
 Many presentations of the Beaufort wind scale include descriptions of the conditions that winds in each category would produce both over land and at sea.\\
-\bottomrule
-\end{tabular}
-\end{table}
+\hline
+\end{longtable}

### 2020-09-13 ehp_revision/writing/ehp_discussion.tex
Continue work on revisions

@@
 ground-based observations, but may sometimes underestimate very high rainfall
-values (Figure~\ref{fig:raincomparison}), and in some counties the correlation
-was substantially lower when considering only tropical cyclones with cumulative
-local rainfall of 75 mm or more (Supplemental Table 1). When rainfall data is
-used to create binary exposure classifications, this disagreement is unlikely
-to influence results, as both data sources agree in identifying these as storms
-with high rainfall, but would be important to consider for cases that include
-rainfall as a continuous measurement.
+values (Figure~\ref{fig:raincomparison}), \textbf{and in some counties the
+correlation was substantially lower when considering only tropical cyclones
+with cumulative local rainfall of 75 mm or more (Supplemental Table 1).} When
+rainfall data is used to create binary exposure classifications, this
+disagreement is unlikely to influence results, as both data sources agree in
+identifying these as storms with high rainfall, but would be important to
+consider for cases that include rainfall as a continuous measurement.
 
@@
 
-\textbf{This dataset is limited to the contiguous US. The precipitation data is
-extracted and aggregated to the county level from a re-analysis data product.
-While the re-analysis product used here (NLDAS-2) only covers the contiguous
-US, other similar re-analysis products, as well as other types of precipitation
-datasets, have global coverage \parencite{sun2018review}. The wind data are
-generated based on a model that is currently US-focused
+\textbf{This dataset is limited to the contiguous US, and while expansion to
+global coverage would be useful and feasible, there would be some challenges.
+For precipitation data, the re-analysis product used here (NLDAS-2) only covers
+the contiguous US, although other re-analysis products, as well as other types
+of precipitation datasets, have global coverage \parencite{sun2018review}. The
+wind data are generated based on a model that is currently US-focused
 \parencite{stormwindmodel}, but could be extended to other areas, although this
@@
 \textbf{The dataset we present focuses on the physical hazards of a tropical
-cyclone.  However, health impacts will often some through indirect pathways,
+cyclone. However, health impacts will often come through indirect pathways,
 including through damage to property and infrastructure. In future work, it
@@
 exposure to any of the other hazards considered, would often introduce exposure
-misclassification \textbf{(Tables S3--S6)}. This conclusion reinforces
-similar findings from a study of Florida's 2004 storm
+misclassification \textbf{(Figure \ref{fig:jaccard}, Tables S3--S6)}. This
+conclusion reinforces similar findings from a study of Florida's 2004 storm
 season~\parencite{grabich2015measuring}.  For some studies, such exposure
@@
 data may be easier to access than individual-level data, especially at a scale
-that covers many locations and a long time period and so allows higher
-statistical power and a broader range of exposure levels within the study
+that covers many locations and years and so allows higher statistical power and
+includes a broader range of exposure levels
 \parencite{wakefield2008overcoming}.}
 
-\textbf{These ecological exposure estimates, however, set a common exposure
+\textbf{Such ecological exposure assessment, however, sets a common exposure
 level throughout the county, ignoring within-county variability, even though
-such variability often exists. For some hazards, this within-county variation
-in exposure level could be stark. For example, tornadoes cause very localized
-damage, directly along the tornado's path---a tornado can destroy homes on one
-side of a street while leaving those on the opposite side untouched. Levels of
-other hazards, like storm-associated winds and rain, will also vary within a
-county, but typically with smoother variation. In particular, it will be
-unlikely that a county will have one area that is exposed to extremes of these
-hazards from a tropical cyclone while other parts of the county are completely
-unexposed, as both the wind fields and rain fields of tropical cyclones tend to
-be large in comparison to the size of a county.}
+such variability exists. For some hazards, this within-county variation could
+be stark. For example, tornadoes cause very localized damage, directly along
+the tornado's path---a tornado can destroy homes on one side of a street while
+leaving those on the opposite side untouched. Levels of other hazards, like
+storm-associated winds and rain, will also vary within a county, but typically
+with smoother variation. In particular, it will be unlikely that a county will
+have one area that is exposed to extremes of these hazards while other parts of
+the county are completely unexposed, as both the wind fields and rain fields of
+tropical cyclones tend to be large in comparison to the size of a county.}
 
-\textbf{Aggregated data can be used to infer contextual level
+\textbf{Aggregated data can be used to infer contextual-level
 associations---for example, the association between county-level exposure to a

### 2020-09-13 ehp_revision/writing/ehp_figures.tex
Work some more on revisions
Continue work on revisions

--- 2020-02-24 writing/ehp_figures.tex
@@
 	radii are routinely available in \ac{HURDAT2}, and for which at
-	least one study county had a peak sustained wind of $\ge$34~\si{\knot}
+	least one study county had a peak sustained wind of $\ge$34~kt
 	based on the post-storm wind radii.
@@
 	\includegraphics[width = 0.8\linewidth]{jaccard_heatmap.pdf}
-	\caption{Agreement between exposure
-	classifications based on different single-hazard exposure metrics for all storms
-	between~1996 and~2011 for which at least~100 counties were exposed
-	based on at least one metric.  Each row shows one storm, and the color
-	of each cell shows the measured Jaccard index for each pair of exposure
-	metrics (proportion of counties classified as exposed by both metrics
-	out of counties classified as exposed by either metric). For 
-	Grace in 2003 and Ida in 2009, there were no county exposures for
-	either the tornado-based metric or the wind-based metric (indicated 
-	by gray squares). Colors to the right of the
-	heatmap show the number of exposed counties
-	based on any of the metrics. Storms are displayed
-	within clusters that have similar patterns based on hierarchical 
-	clustering.}
+	\caption{Agreement between exposure classifications based on different
+	single-hazard exposure metrics for all storms between~1996 and~2011 for
+	which at least~100 counties were exposed based on at least one metric.
+	Each row shows one storm, and the color of each cell shows the measured
+	Jaccard index for each pair of exposure metrics (proportion of counties
+	classified as exposed by both metrics out of counties classified as
+	exposed by either metric). For Grace in 2003 and Ida in 2009, there
+	were no county exposures for either the tornado-based metric or the
+	wind-based metric (indicated by gray squares). Colors to the right of
+	the heatmap show the number of exposed counties based on any of the
+	metrics, \textbf{and this panel is linked with the color scale labeled ``\# of
+	counties exposed by any metric''.} Storms are displayed within clusters
+	that have similar patterns based on hierarchical clustering.}
 \label{fig:jaccard} 

### 2020-09-13 ehp_revision/writing/ehp_methods.tex
Continue work on revisions

@@
 used tracking data from \ac{HURDAT2}, which records the storm center's position
-at four standardized times for weather data collection (synoptic times),
-6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We interpolated this
-position to~15-\si{\minute} intervals using natural cubic spline
+at \textbf{four standardized times for weather data collection (synoptic
+times)}, 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We interpolated
+this position to~15-\si{\minute} intervals using natural cubic spline
 interpolation~\parencite{hurricaneexposure}. At each~15-\si{\minute} interval,
@@
 data} for the \ac{US}. \textbf{There are no other collections of tornado data
-comparable in emporal or geographic scope,} and so we did not further validate
+comparable in temporal or geographic scope,} and so we did not further validate
 the tornado event data. It is difficult to characterize flooding at the county
@@
 estimate be confounded by other storm hazards? When planning a new study that
-will incorporate several storm hazards in a mixtures approach, will there
-likely be numerical problems in fitting the model from multicollinearity, or
-difficulties in disentangling the role of separate hazards?}
+will incorporate several storm hazards, might there be variance inflation from
+multicollinearity or difficulties in disentangling the roles of separate
+hazards?}
 

### 2020-09-13 ehp_revision/writing/ehp_results.tex
Continue work on revisions

@@
 from other available sources.  For estimates of storm-associated rainfall, we
-compared data in the open-source package with ground-based
-observations in nine sample counties (Figure~\ref{fig:raincomparison}). Within
-these counties, storm-related rainfall measurements were well-correlated
-between the two data sources, with rank correlations (bottom right of each
-graph in Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. 
-There was some evidence that our primary rainfall metric may tend to
-underestimate rainfall totals in storms with extremely high rainfall, based on
-a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
-County, SC, and Wake County, NC (Figure~\ref{fig:raincomparison}). However,
-even in these cases, it was rare for a storm to be classified differently
-(under the precipitation threshold of 75~\si{\milli\metre} we used for binary
-exposure classification for later analysis) based on the source of
-precipitation data. Horizontal and vertical lines in each small plot in
-Figure~\ref{fig:raincomparison} show the threshold of 75~\si{\milli\metre}, so
-storms in the lower left and upper right quadrants would be classified the same
-(``exposed'' or ``unexposed'') regardless of the precipitation data source,
-while storms in the upper left and lower right quadrants would be classified
-differently. Such cases were rare.
+compared data in the open-source package with ground-based observations in nine
+sample counties (Figure~\ref{fig:raincomparison}). Within these counties,
+storm-related rainfall measurements were well-correlated between the two data
+sources, with rank correlations (bottom right of each graph in
+Figure~\ref{fig:raincomparison}) between 0.87 and 0.98.  There was some
+evidence that our primary rainfall metric may tend to underestimate rainfall
+totals in storms with extremely high rainfall, based on a few heavy-rainfall
+storms in Harris County, TX, Mobile County, AL, Charleston County, SC, and Wake
+County, NC (Figure~\ref{fig:raincomparison}). \textbf{Further, in some
+counties, the correlation was substantially lower when considering only
+tropical cyclones with cumulative local rainfall of 75 mm or more (Supplemental
+Table 1).} However, even in these cases, it was rare for a storm to be
+classified differently (under the precipitation threshold of
+75~\si{\milli\metre} we used for binary exposure classification for later
+analysis) based on the source of precipitation data. Horizontal and vertical
+lines in each small plot in Figure~\ref{fig:raincomparison} show the threshold
+of 75~\si{\milli\metre}, so storms in the lower left and upper right quadrants
+would be classified the same (``exposed'' or ``unexposed'') regardless of the
+precipitation data source, while storms in the upper left and lower right
+quadrants would be classified differently. Such cases were rare.
 
@@
 which~100 or more counties were exposed based on at least one metric
-(Figure~\ref{fig:jaccard}; \textbf{Tables S3--S6 provide underlying
-numbers for the most extensive of these, the storms for which 250 of more
-counties were exposed based on at least one metric}). In this figure, each row
+(Figure~\ref{fig:jaccard}; \textbf{for the most extensive of these, storms for
+which 250 of more counties were exposed based on at least one metric, Tables
+S3--S6 provide underlying numbers comparing exposure assessments for the
+distance-based proxy to each hazard-based metric}). In \textbf{Figure 7}, each row
 provides results for one tropical cyclone, and each box in that row shows the

### 2020-09-16 ehp_revision/tables/precip_high_corr.tex
Work some more on response

@@
                      The same sample of counties is shown as in Figure 2 of the main text.
-                     Events are cases where a tropical cyclone came within ... km of each of the 
+                     Events are cases where a tropical cyclone came within 500 km of each of the 
                      listed counties. The number of total events gives the sum of all points
@@
                      precipitation was 75 mm or higher based on at least one of the two measures
-                     considered in this comparison (NLDAS and monitors). The Spearman correlation
-                     for these high-precipitation events is given in the last column of the 
+                     considered in this comparison (NLDAS-2 reanalysis data and ground-based stations). The Spearman correlation between these two precipitation data sources is given
+                     for these high-precipitation events in the last column of the 
                      table.}

### 2020-09-16 ehp_revision/tables/thresholds_table.tex
Work some more on response

@@
 \begin{longtable}{lp{40em}}
-\caption{Caption.
+\caption{Reasons behind the choices of thresholds for binary exposure classifications, as well as discussion of some other reasonable choices.
 These are provided for the three exposure metrics for which our database includes continuous data, and so a threshold is selected to determine binary exposure based on the metric.
-This table provides reasoning for the choice of threshold used in this paper as well as guidance on other thresholds that could be considered, depending on the hypothesized pathways for an epidemiological study.}
+This table provides reasoning for the choice of threshold used in this paper as well as guidance on other thresholds that could be considered, depending on the hypothesized pathways for an epidemiological study.} \\
 \label{tab:thresholds} \\
@@
 \hline
-Distance & In the manuscript, distance-based exposure was assessed if the storm track came within 100 km of county population mean center.
-This threshold for distance from the storm center has been used in prior research as a proxy for exposure to hazards from the storm \parencite{grabich2015measuring}.
+Distance & Distance-based exposure was determined based on whether the storm track came within 100 km of county population mean center.
+This threshold has been used in prior research as a proxy for exposure to hazards from the storm (e.g., \cite{grabich2015measuring}).
 Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km \parencite{mallin2006, quiring2011variations}, and dangerous winds can extend beyond these maximum winds.
-One study \parencite{czajkowski2011} assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties.
-Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track \parencite{czajkowski2011}.
-Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research \parencite{grabich2015measuring, grabich2015, currie2013}.\\
-Rain & In the manuscript, rain-based exposure was assessed if the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
+One study assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties, which resulted in an average distance radius of 120 km on either side of the storm track \parencite{czajkowski2011}.
+Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research \parencite{grabich2015measuring, grabich2015, currie2013}.
+However, based on the results presented in the main manuscript, hazard-based metrics should often be used directly rather than a distance-based proxy.\\
+Rain & Rain-based exposure was determined based on whether the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
 This threshold...
@@
 \\
-Wind & In the manuscript, wind-based exposure was assessed if modeled storm-associated peak sustained surface wind of $\ge34$ kts at the county’s population mean center.
+Wind & Wind-based exposure was determined based on whether modeled storm-associated peak sustained surface wind were $\ge34$ kts at the county’s population mean center.
 This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale.

### 2020-09-16 ehp_revision/R/rain_comparison_figure.R

@@ -385,16 +385,16 @@ supp_tab <- high_corr_table %>%
                      booktabs = TRUE, 
                      escape = FALSE, 
                      caption = "Precipitation correlation during all versus high-precipitation events.
-                     The same sample of counties is shown as in Figure ... of the main text.
-                     Events are cases where a tropical cyclone came within ... km of each of the 
+                     The same sample of counties is shown as in Figure 2 of the main text.
+                     Events are cases where a tropical cyclone came within 500 km of each of the 
                      listed counties. The number of total events gives the sum of all points
-                     shown on the main plot for the county in Figure ... of the main text. 
-                     The Spearman correlation for all events is the same as that shown in Figure ...
+                     shown on the main plot for the county in Figure 2 of the main text. 
+                     The Spearman correlation for all events is the same as that shown in Figure 2
                      of the main text. High-precipitation events are those for which storm-associated
                      precipitation was 75 mm or higher based on at least one of the two measures
-                     considered in this comparison (NLDAS and monitors). The Spearman correlation
-                     for these high-precipitation events is given in the last column of the 
-                     table.", 
+                     considered in this comparison (NLDAS-2 reanalysis data and ground-based stations). 
+                     The Spearman correlation between these two precipitation data sources is given
+                     for these high-precipitation events in the last column of the table.", 
                      align = "lcccc", 
                      label = "highprecipcorr") %>% 
         add_header_above(c(" ", "All events" = 2, "High-precipitation events" = 2)) 

### 2020-09-18 ehp_revision/response_to_reviewers.Rmd
Work some more on revision

@@
 - \DeclareDelayedFloatFlavor{longtable}{table}
+- \usepackage{pdflscape}
+- \setlength\LTcapwidth{\textwidth} 
 ---
@@
 variation across time and geography in the likelihood of a given exposure
-resulting in a disaster declaration. This would be particularly worrisome for a
+resulting in a disaster declaration. This could be problematic for a
 study that seeks to explore risk across multiple years and affected communities.
@@
 
-We have moved other suggestions that the reviewer sent in "Track Changes" of the
+We have moved suggestions that the reviewer sent in "Track Changes" of the
 Word document for the original manuscript and address each in detail below. When
@@
 exposure metrics for two reasons. First, while we agree with the reviewer that a
-direct measurement is always preferable to a proxy, there are still some studies
+direct measurement of each relevant hazard is preferable to a proxy, there are still some studies
 that, when assessing tropical cyclone exposure based on measurements of physical
 hazards, use a single hazard for this exposure assessment. We are most familar
-with cases where a wind-based metric has been the focus in these cases (e.g.,
-@sun2020tropical, @grabich2015measuring). Therefore, sometimes might not be
+with cases where a wind-based metric has been the focus (e.g.,
+@sun2020tropical, @grabich2015measuring). While this may not be 
 ideal, it is done in practice. To interpret these studies, it is helpful to
@@
 hazard, but not for all storm hazards. In this case, if the agreement is
-moderate, though, there is a chance of confounding from other storm hazards
+moderate, there is a chance of confounding from other storm hazards
 (i.e., an association that is attributed to one storm hazard was actually the
@@
 measurements of agreement across four hazard-based exposure metrics that can
-provide insight on these points.
+provide insights on these points.
 
@@
 both numerical issues (e.g., variance inflation from multicollinearity) and
-difficulties in interpretation (e.g., difficulty infering the role of specific
-hazards and separating the influece of each). If an epidemiologist has a good
-understanding of these patterns in their data, he or she can adopt methods that
+difficulties in interpretation (e.g., difficulty inferring the role of specific
+hazards and separating the influence of each). If an epidemiologist has a good
+understanding of these patterns in the data, he or she can adopt methods that
 address potential issues, including using matching-based study designs to create
-a subset of storms to study in which separate hazards are well-distinguished
-from each other, incorporation of dimension-reduction methods like principal
+a subset of storms in which separate hazards are well-distinguished,
+incorporation of dimension-reduction methods like principal
 components analysis, or shrinkage techniques that allow for a small amount of
-bias in estimation of associations in return for reducing problems from
+bias in estimation in return for reducing problems from
 collinearity (e.g., ridge regression) [@leal2012multicollinearity;
@@
 correlation specifically within these higher-precipitation events would add more
-quantitative evidence to this discussion.
-
-We have added estimates of these correlations specifically for
-high-precipitation events in a new table in the Supplemental Material (Table S2
-of the manuscript, reproduced as Table \@ref(tab:highprecipcorr) of this
-response document).
+quantitative evidence to this discussion. We have added estimates of these
+correlations specifically for high-precipitation events in a new table in the
+Supplemental Material (Table S2 of the manuscript, reproduced as Table
+\@ref(tab:highprecipcorr) of this response document).
 
@@
 remained high (e.g., Harris County, TX; Orleans Parish, LA), although in all
-cases there are small numbers of extreme precipitation events (5--20 across
+cases there are few extreme precipitation events (5--20 across
 these sample counties), and so there is likely substantial random variation in
@@
 2). **Further, in some counties, the correlation was substantially lower when 
-considering only tropical cyclones with cumulative local rainfall of 75 mm
-or more (Supplemental Table 1)**."
+considering only tropical cyclones with cumulative local rainfall of $\ge75$ mm
+(Supplemental Table 1)**."
 
@@
 some counties the correlation was substantially lower when considering only
-tropical cyclones with cumulative local rainfall of 75 mm or more (Supplemental
+tropical cyclones with cumulative local rainfall of $\ge75$ mm (Supplemental
 Table 1).**"
@@
 This is a great suggestion. We have made this change, as well as several changes
-in response to EHP's Request 4 later in this response.
-
-The original and revised versions of Figure 4 of the manuscript are reproduced in
-Figure \@ref(fig:fig4) of this response.
+in response to EHP's Request 4 later in this response. The original and revised
+versions of Figure 4 of the manuscript are reproduced in Figure \@ref(fig:fig4)
+of this response.
 
@@
 
-\begin{longtable}{lp{40em}}
-\caption{Caption.
+\begin{longtable}{lp{35em}}
+\caption{Reasons behind the choices of thresholds for binary exposure classifications, as well as discussion of some other reasonable choices.
 These are provided for the three exposure metrics for which our database includes continuous data, and so a threshold is selected to determine binary exposure based on the metric.
-This table provides reasoning for the choice of threshold used in this paper as well as guidance on other thresholds that could be considered, depending on the hypothesized pathways for an epidemiological study.}
+This table provides reasoning for the choice of threshold used in this paper as well as guidance on other thresholds that could be considered, depending on the hypothesized pathways for an epidemiological study.} \\
 \label{tab:thresholds} \\
@@
 \hline
-Distance & In the manuscript, distance-based exposure was assessed if the storm track came within 100 km of county population mean center.
-This threshold for distance from the storm center has been used in prior research as a proxy for exposure to hazards from the storm \cite{grabich2015measuring}.
-Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km \cite{mallin2006, quiring2011variations}, and dangerous winds can extend beyond these maximum winds.
-One study \cite{czajkowski2011} assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties.
-Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track \cite{czajkowski2011}.
-Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research \cite{grabich2015measuring, grabich2015, currie2013}.\\
-Rain & In the manuscript, rain-based exposure was assessed if the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
-This threshold...
-Other reasonable definitions could vary both the threshold of rainfall (in mm) and also the choice of days surrounding the storm to include when calculating the cumulative rainfall.
-Tropical cyclone precipitation arrives before the center of the storm \cite{zhou2017spatial}.
-A variety of definitions have been used to identify both extreme or heavy rain (whether associated with a tropical cyclone or not) and tropical cyclone--associated rain.
-Some definitions of extreme rain events are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours \cite{schumacher2006characteristics}.
+\textbf{Distance} & Distance-based exposure was determined based on whether the storm track came within 100 km of county population mean center.
+This threshold has been used in prior research as a proxy for exposure to hazards from the storm (e.g., @grabich2015measuring).
+Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km [@mallin2006; @quiring2011variations], and dangerous winds can extend beyond these maximum winds.
+One study assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties, which resulted in an average distance radius of 120 km on either side of the storm track [@czajkowski2011].
+Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research [@grabich2015measuring; @grabich2015; @currie2013].
+However, based on the results presented in the main manuscript, hazard-based metrics should often be used directly rather than a distance-based proxy.\\
+
+\textbf{Rain} & Rain-based exposure was determined based on whether the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
+One recent study has highlighted that a two-year rainfall value (which is the median annual maximum rainfall) for a location can provide a useful threshold in identifying rainfall events with potential for societal impacts [@bosma2020intuitive].
+For some of the more northern, inland communities included in our study area, the two-year rainfall value for two-, three- and four-day windows is in the 65--85 mm range.
+For example, Pittsburgh, PA, has two-year rainfall values of 69 mm for a two-day window, 74 mm for a three-day window, and 78 mm for a four-day window [@noaaatlas], consistent with the 75 mm threshold we selected for exposure classification in this paper.
+However, other thresholds, particularly higher thresholds, would be reasonable in some cases.
+For example, the two-year rainfall values tend to be much higher in counties of the study area that are further south and close to the coast.
+The two-year rainfall value for Miami for a three-day window, for example is 180 mm [@noaaatlas].
+A variety of definitions have been used previously to identify both extreme or heavy rain (whether associated with a tropical cyclone or not) and tropical cyclone--associated rain.
+In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall [@zhou2017spatial] and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---\parfillskip=0pt\tabularnewline
+
+&in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location [@feldmann2019estimation].
+Other definitions of extreme rain events---including but not limited to tropical cyclone--associated rainfall---are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours [@schumacher2006characteristics].
 Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]).
-\cite{schumacher2006characteristics, schumacher2005organization, stevenson201410}.
-A study focused on the risks that result from extreme rainfall may therefore reasonably choose a higher threshold than the 75 mm threshold we use in this analysis.
-In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall \cite{zhou2017spatial} and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location \cite{feldmann2019estimation}.
-\\
-Wind & In the manuscript, wind-based exposure was assessed if modeled storm-associated peak sustained surface wind of $\ge34$ kts at the county’s population mean center.
+[@schumacher2006characteristics; @schumacher2005organization; @stevenson201410].\\
+
+\textbf{Wind} & Wind-based exposure was determined based on whether modeled storm-associated peak sustained surface wind were $\ge34$ kts at the county’s population mean center.
 This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale.
-This limit is used as the outer limit in measuring storm size through the US National Hurricane Center's wind radii for tropical cyclone forecasts \cite{cangialosi2016examination}.
+This limit is used as the outer limit in measuring storm size through the US National Hurricane Center's wind radii for tropical cyclone forecasts [@cangialosi2016examination].
 Other thresholds could be selected based on other points on the Beaufort scale---for example, $\ge48$ kts for capturing storm-force winds or $\ge64$ kts for capturing hurricane-force winds.

### 2020-09-18 ehp_revision/tables/flood_misclass.tex
Work some more on revision

@@
 % latex table generated in R 4.0.2 by xtable 1.8-4 package
-% Wed Sep  9 17:22:05 2020
+% Fri Sep 18 10:32:13 2020
 \begin{table}[ht]
 \centering
-\caption{Caption.
-Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
-Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
-Exposure assessment is based on the thresholds given in Table 1 of the main text.
-The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
-Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.}
+\caption{Agreement between flood-based exposure assessment and a distance-based
+         proxy of storm exposure for tropical cyclones with at least 200 counties assessed 
+         as exposed based on at least one exposure metric considered in this study. Numbers 
+         are out of 2,396 counties in the study area (states in the eastern half of the US; 
+         Figure 1 of the main text). Exposure assessment is based on the thresholds given in 
+         Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
+         calculated as the value in the second column divided by the sum of numbers in the 
+         second through fourth columns.  Storms are ordered based on the number of counties 
+         assessed as exposed to at least one of these two exposure metrics.).} 
 \label{tab:misclassflood}

### 2020-09-18 ehp_revision/tables/rain_misclass.tex
Work some more on revision

@@
 % latex table generated in R 4.0.2 by xtable 1.8-4 package
-% Wed Sep  9 17:22:05 2020
+% Fri Sep 18 10:32:05 2020
 \begin{table}[ht]
 \centering
-\caption{Caption.
-Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
-Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
-Exposure assessment is based on the thresholds given in Table 1 of the main text.
-The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.
-Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.).}
+\caption{Agreement between rain-based exposure assessment and a distance-based
+         proxy of storm exposure for tropical cyclones with at least 200 counties assessed 
+         as exposed based on at least one exposure metric considered in this study. Numbers 
+         are out of 2,396 counties in the study area (states in the eastern half of the US; 
+         Figure 1 of the main text). Exposure assessment is based on the thresholds given in 
+         Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
+         calculated as the value in the second column divided by the sum of numbers in the 
+         second through fourth columns.  Storms are ordered based on the number of counties 
+         assessed as exposed to at least one of these two exposure metrics.).} 
 \label{tab:misclassrain}

### 2020-09-18 ehp_revision/tables/thresholds_table.tex
Work some more on revision

@@
-\begin{longtable}{lp{40em}}
+\begin{longtable}{lp{35em}}
 \caption{Reasons behind the choices of thresholds for binary exposure classifications, as well as discussion of some other reasonable choices.
@@
 \hline
-Distance & Distance-based exposure was determined based on whether the storm track came within 100 km of county population mean center.
+\textbf{Distance} & Distance-based exposure was determined based on whether the storm track came within 100 km of county population mean center.
 This threshold has been used in prior research as a proxy for exposure to hazards from the storm (e.g., \cite{grabich2015measuring}).
@@
 However, based on the results presented in the main manuscript, hazard-based metrics should often be used directly rather than a distance-based proxy.\\
-Rain & Rain-based exposure was determined based on whether the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
-This threshold...
-Other reasonable definitions could vary both the threshold of rainfall (in mm) and also the choice of days surrounding the storm to include when calculating the cumulative rainfall.
-Tropical cyclone precipitation arrives before the center of the storm \parencite{zhou2017spatial}.
-A variety of definitions have been used to identify both extreme or heavy rain (whether associated with a tropical cyclone or not) and tropical cyclone--associated rain.
-Some definitions of extreme rain events are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours \parencite{schumacher2006characteristics}.
+
+\textbf{Rain} & Rain-based exposure was determined based on whether the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
+One recent study has highlighted that a two-year rainfall value (which is the median annual maximum rainfall) for a location can provide a useful threshold in identifying rainfall events with potential for societal impacts \parencite{bosma2020intuitive}.
+For some of the more northern, inland communities included in our study area, the two-year rainfall value for two-, three- and four-day windows is in the 65--85 mm range.
+For example, Pittsburgh, PA, has two-year rainfall values of 69 mm for a two-day window, 74 mm for a three-day window, and 78 mm for a four-day window \parencite{noaaatlas}, consistent with the 75 mm threshold we selected for exposure classification in this paper.
+However, other thresholds, particularly higher thresholds, would be reasonable in some cases.
+For example, the two-year rainfall values tend to be much higher in counties of the study area that are further south and close to the coast.
+The two-year rainfall value for Miami for a three-day window, for example is 180 mm \parencite{noaaatlas}.
+A variety of definitions have been used previously to identify both extreme or heavy rain (whether associated with a tropical cyclone or not) and tropical cyclone--associated rain.
+In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall \parencite{zhou2017spatial} and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---\parfillskip=0pt\tabularnewline
+
+&in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location \parencite{feldmann2019estimation}.
+Other definitions of extreme rain events---including but not limited to tropical cyclone--associated rainfall---are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours \parencite{schumacher2006characteristics}.
 Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]).
-\parencite{schumacher2006characteristics, schumacher2005organization, stevenson201410}.
-A study focused on the risks that result from extreme rainfall may therefore reasonably choose a higher threshold than the 75 mm threshold we use in this analysis.
-In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall \parencite{zhou2017spatial} and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location \parencite{feldmann2019estimation}.
-\\
-Wind & Wind-based exposure was determined based on whether modeled storm-associated peak sustained surface wind were $\ge34$ kts at the county’s population mean center.
+\parencite{schumacher2006characteristics, schumacher2005organization, stevenson201410}.\\
+
+\textbf{Wind} & Wind-based exposure was determined based on whether modeled storm-associated peak sustained surface wind were $\ge34$ kts at the county’s population mean center.
 This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale.

### 2020-09-18 ehp_revision/tables/tornado_misclass.tex
Work some more on revision

@@
 % latex table generated in R 4.0.2 by xtable 1.8-4 package
-% Wed Sep  9 17:22:05 2020
+% Fri Sep 18 10:32:16 2020
 \begin{table}[ht]
 \centering
-\caption{Caption.
-Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
-Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
-Exposure assessment is based on the thresholds given in Table 1 of the main text.
-The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
-Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.}
+\caption{Agreement between tornado-based exposure assessment and a distance-based
+         proxy of storm exposure for tropical cyclones with at least 200 counties assessed 
+         as exposed based on at least one exposure metric considered in this study. Numbers 
+         are out of 2,396 counties in the study area (states in the eastern half of the US; 
+         Figure 1 of the main text). Exposure assessment is based on the thresholds given in 
+         Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
+         calculated as the value in the second column divided by the sum of numbers in the 
+         second through fourth columns.  Storms are ordered based on the number of counties 
+         assessed as exposed to at least one of these two exposure metrics.).} 
 \label{tab:misclasstornado}

### 2020-09-18 ehp_revision/tables/wind_misclass.tex
Work some more on revision

@@
 % latex table generated in R 4.0.2 by xtable 1.8-4 package
-% Wed Sep  9 17:22:05 2020
+% Fri Sep 18 10:32:09 2020
 \begin{table}[ht]
 \centering
-\caption{Caption.
-Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study.
-Numbers are out of 2,396 counties in the study area (states in the eastern half of the US).
-Exposure assessment is based on the thresholds given in Table 1 of the main text.
-The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns).
-Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.}
+\caption{Agreement between wind-based exposure assessment and a distance-based
+         proxy of storm exposure for tropical cyclones with at least 200 counties assessed 
+         as exposed based on at least one exposure metric considered in this study. Numbers 
+         are out of 2,396 counties in the study area (states in the eastern half of the US; 
+         Figure 1 of the main text). Exposure assessment is based on the thresholds given in 
+         Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
+         calculated as the value in the second column divided by the sum of numbers in the 
+         second through fourth columns.  Storms are ordered based on the number of counties 
+         assessed as exposed to at least one of these two exposure metrics.).} 
 \label{tab:misclasswind}

### 2020-09-18 ehp_revision/writing/ehp_discussion.tex
Work some more on revision

@@
 correlation was substantially lower when considering only tropical cyclones
-with cumulative local rainfall of 75 mm or more (Supplemental Table 1).} When
+with cumulative local rainfall of $\ge75$ mm (Supplemental Table 1).} When
 rainfall data is used to create binary exposure classifications, this

### 2020-09-18 ehp_revision/writing/ehp_results.tex
Work some more on revision

@@
 counties, the correlation was substantially lower when considering only
-tropical cyclones with cumulative local rainfall of 75 mm or more (Supplemental
+tropical cyclones with cumulative local rainfall of $\ge75$ mm (Supplemental
 Table 1).} However, even in these cases, it was rare for a storm to be

### 2020-09-18 ehp_revision/R/misclassification_tables.R

@@ -145,7 +145,15 @@ misclass_rain %>%
          Storm = paste0(Storm, ")")) %>% 
   mutate_at(2:5, prettyNum, big.mark = ",") %>% 
   xtable(align = "ccP{3cm}P{3cm}P{3cm}P{3cm}",
-         caption = "Caption. Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns.  Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.).", 
+         caption = "Agreement between rain-based exposure assessment and a distance-based
+         proxy of storm exposure for tropical cyclones with at least 200 counties assessed 
+         as exposed based on at least one exposure metric considered in this study. Numbers 
+         are out of 2,396 counties in the study area (states in the eastern half of the US; 
+         Figure 1 of the main text). Exposure assessment is based on the thresholds given in 
+         Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
+         calculated as the value in the second column divided by the sum of numbers in the 
+         second through fourth columns.  Storms are ordered based on the number of counties 
+         assessed as exposed to at least one of these two exposure metrics.).", 
          label = "tab:misclassrain") %>% 
   print(include.rownames = FALSE, booktabs = TRUE, 
          only.contents = FALSE,
@@ -178,7 +186,15 @@ misclass_wind %>%
          Storm = paste0(Storm, ")")) %>% 
   mutate_at(2:5, prettyNum, big.mark = ",") %>% 
   xtable(align = "ccP{3cm}P{3cm}P{3cm}P{3cm}",
-         caption = "Caption. Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns). Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.",
+         caption = "Agreement between wind-based exposure assessment and a distance-based
[…]

### 2020-09-18 ehp_revision/response_to_reviewers.Rmd
Work some more on revisions

@@
 
-DesignSafe is an online platform sponsered by the National Science Foundation as
+DesignSafe is an online platform sponsored by the National Science Foundation as
 part of the Natural Hazards Engineering Research Infrastructure
@@
 that, when assessing tropical cyclone exposure based on measurements of physical
-hazards, use a single hazard for this exposure assessment. We are most familar
+hazards, use a single hazard for this exposure assessment. We are most familiar
 with cases where a wind-based metric has been the focus (e.g.,
@@
 hazards, but this finding is not as self-evident for environmental
-epidemiologists. Based on a recent conference of the Interational Society for
+epidemiologists. Based on a recent conference of the International Society for
 Environmental Epidemiology, for example, there are a number of environmental
@@
 used the Jaccard index because we are using data for all eastern US counties.
-For any storm, however, there will be large parts of this area that are
+For any storm, there will be large parts of this area that are
 unquestionably unaffected, because they are so far from the storm. For example,
@@
 
+We therefore used the Jaccard index, which focuses on counties that were
+assessed as exposed based on at least one of the two metrics being compared.
 However, one downside of this metric is that it does not help readers understand
-the direction that disagreement is going. For example, it does not provide 
-information about how many counties were classified as "exposed" based on a 
-distance proxy but "unexposed" by a hazard-based proxy. This could be useful in 
-interpreting our results in Figure 7 of the main table.
+the direction of that disagreement. For example, it does not provide information
+about how many counties were classified as "exposed" based on a distance proxy
+but "unexposed" by a hazard-based proxy. This could be useful in interpreting
+our results in Figure 7 of the main table.
 
@@
 
-We have added four new tables in the Supplemental Material, one for each of the
-hazard-based metrics we considered. We have reproduced these tables in this
-response document as Tables \@ref(tab:misclasswind)--\@ref(tab:misclassflood)
-Each breaks down agreement and disagreement between the distance-based proxy
-metric and that hazard-based metric within several storms in our study period.
-These tables focus on storms with at least 250 counties exposed by at least one
-of the metrics considered. These tables allow readers both to investigate
-general patterns in disagreement and exposure misclassification across these
-storms and also to see cases of unusual storms. For example, for most storms,
-disagreement between the distance proxy metric and the wind-based metric
-resulted mainly from mis-classification of counties as exposed based on distance
-while they did not experience gale-force winds. However, there were a few
-storms, like Hurricane Bertha in 1996, where the inverse was the case, and a
-distance-based metric would have missed a large number of counties that
-experienced gale-force winds from the storm (misclassified them as unexposed),
-but not misclassified many counties as exposed when they in fact did not
-experience winds at this level.
+We have added four new tables in the Supplemental Material (Tables S3--S6), one
+for each of the hazard-based metrics we considered. We have reproduced these
+tables at the end of this response document as Tables
+\@ref(tab:misclasswind)--\@ref(tab:misclassflood). Each breaks down agreement
+and disagreement between the distance-based proxy metric and one hazard-based
+metric, focusing on storms with at least 200 counties exposed by at least one of
+the metrics considered. These tables allow readers both to investigate general
+patterns in disagreement and exposure misclassification across these storms and
+also to see cases of unusual storms. For example, for most storms, disagreement
+between the distance proxy metric and the wind-based metric resulted mainly from
+mis-classification of counties as exposed based on distance while they did not
+experience gale-force winds. However, there were a few storms, like Hurricane
+Bertha in 1996, where the inverse was the case, and a distance-based metric
+would have missed a large number of counties that experienced gale-force winds
+from the storm (misclassified them as unexposed), but not misclassified many
+counties as exposed when they in fact did not experience winds at this level.
 
@@
 [@mccarthy2003nws]. It is the original source for the only other national
-tornado database available for the country (the Storm Prediction Center [SPC]'s
-National Tornado Database [@center2020storm]), which provides a version of the
+tornado database available for the country, the Storm Prediction Center [SPC]'s
+National Tornado Database [@center2020storm], which provides a version of the
 same data in which tornado events are joined into a single record in cases where
@@
 
-For all outcomes, there are various reasonable choices that can be picked for
+Ultimately, this choice will depend on what a researcher considers as plausible
+causal pathways for the outcome he or she is studying. For example, if power
+outages form a plausible step in the pathway, then the threshold for gale-force
+or strong gale-force winds might be reasonable, while if the pathway is more
+likely to be through larger-scale building destruction, a higher threshold would
+be more reasonable.
+
+For all outcomes, there are therefore various reasonable choices that can be picked for
 the threshold in defining exposure. There is not room to go into these details
@@
 Table \@ref(tab:thresholds) at the end of this response document.
+
+We have added text to introduce this supplemental table in the revised manuscript, 
+both in the Methods and in the Discussion:
+
+> **Text in the revised Methods (added text in bold)**:
+"For other exposure metrics, each county was
+classified as exposed to a tropical cyclone based on whether the exposure
+metric exceeded a certain threshold (Table 1). We
+picked reasonable thresholds (e.g., the threshold for gale-force wind for wind
+exposure; **Table S1**), but others could be used with the open-source data and its
+associated software."
+
+> **Text in the revised Discussion (added text in bold)**:
+"Finally, to assess patterns and agreement for binary exposure classifications,
+we have chosen one set of sensible thresholds for binary classifications based
+on continuous metrics (rainfall, maximum sustained surface wind, and distance
+from the storm's track), **but other thresholds would be reasonable
+depending on hypothesized pathways for a given epidemiological study (Table
+S1)**."
 
@@
 
-Ultimately, this choice will depend on what a researcher considers as plausible
-causal pathways for the outcome he or she is studying. For example, if power
-outages form a plausible step in the pathway, then the threshold for gale-force
-or strong gale-force winds might be reasonable, while if the pathway is more
-likely to be through larger-scale building destruction, a higher threshold would
-be more reasonable. Some studies have done sensitivity analyses to consider how
-results changed using different exposure metrics. For example, @czajkowski2011
-performed a sensitivity analysis using different assumptions about wind decay
-away from the center of a storm, while @currie2013, who used a distance-based
-exposure metric, considered different buffer distances as a sensitivity
-analysis.
-
-We have added text to introduce this supplemental table in the revised manuscript, 
-both in the Methods and in the Discussion:
-
-> **Text in the revised Methods (added text in bold)**:
-"For other exposure metrics, each county was
-classified as exposed to a tropical cyclone based on whether the exposure
-metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
-picked reasonable thresholds (e.g., the threshold for gale-force wind for wind
-exposure; **Table S1**), but others could be used with the open-source data and its
-associated software."
-
-> **Text in the revised Discussion (added text in bold)**:
-"Finally, to assess patterns and agreement for binary exposure classifications,
-we have chosen one set of sensible thresholds for binary classifications based
-on continuous metrics (rainfall, maximum sustained surface wind, and distance
-from the storm's track), **but other thresholds would be reasonable
-depending on hypothesized pathways for a given epidemiological study (Table
-S1)**."
-
 \medskip
@@
 forms of color deficiency [@nunez2018optimizing]. They reproduce their original
-sequential gradient when printed in grayscale and so are perceptable when
+sequential gradient when printed in grayscale and so are perceptible when
 printed on a black-and-white printer [@van2015mpl; @nunez2018optimizing].

### 2020-09-19 ehp_revision/response_to_reviewers.Rmd
Work some more on revisions

@@
 considering only tropical cyclones with cumulative local rainfall of $\ge75$ mm
-(Supplemental Table 1)**."
+(Supplemental Table 2)**."
 
@@
 tropical cyclones with cumulative local rainfall of $\ge75$ mm (Supplemental
-Table 1).**"
+Table 2).**"
 
@@
 which~100 or more counties were exposed based on at least one metric (Figure7;
-**for the most extensive of these, storms for which 250 of more counties were
+**for the most extensive of these, storms for which 200 of more counties were
 exposed based on at least one metric, Tables S3--S6 provide underlying numbers
-comparing exposure assessments for the distance-based proxy to each hazard-based
-metric**)."
+comparing exposure assessments between the distance-based proxy and each
+hazard-based metric**)."
 
@@
 extreme winds are more common to the track's right where cyclonic winds move in
-concert with the tropical cyclone's forward motion."
+concert with the tropical cyclone's forward motion (Halverson 2015)."
 
@@
 Other definitions of extreme rain events---including but not limited to tropical cyclone--associated rainfall---are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours [@schumacher2006characteristics].
-Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]).
-[@schumacher2006characteristics; @schumacher2005organization; @stevenson201410].\\
-
-\textbf{Wind} & Wind-based exposure was determined based on whether modeled storm-associated peak sustained surface wind were $\ge34$ kts at the county’s population mean center.
+Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]) [@schumacher2006characteristics; @schumacher2005organization; @stevenson201410].\\
+
+\textbf{Wind} & Wind-based exposure was determined based on whether modeled storm-associated peak sustained surface wind was $\ge34$ kts at the county’s population mean center.
 This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale.
@@
 Other thresholds could be selected based on other points on the Beaufort scale---for example, $\ge48$ kts for capturing storm-force winds or $\ge64$ kts for capturing hurricane-force winds.
-As a note, hurricane-force winds will be rarely experienced for counties, as it will likely only be observed for very severe storms and even for those for counties near the storm's landfall.
+As a note, hurricane-force winds will be rarely experienced for counties, as it will likely only be observed for very severe storms and even for those, only for counties near the storm's landfall.
 Many presentations of the Beaufort wind scale include descriptions of the conditions that winds in each category would produce both over land and at sea.\\

### 2020-09-19 ehp_revision/tables/thresholds_table.tex
Work some more on revisions

@@
 Other definitions of extreme rain events---including but not limited to tropical cyclone--associated rainfall---are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours \parencite{schumacher2006characteristics}.
-Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]).
-\parencite{schumacher2006characteristics, schumacher2005organization, stevenson201410}.\\
+Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]) \parencite{schumacher2006characteristics, schumacher2005organization, stevenson201410}.\\
 
-\textbf{Wind} & Wind-based exposure was determined based on whether modeled storm-associated peak sustained surface wind were $\ge34$ kts at the county’s population mean center.
+\textbf{Wind} & Wind-based exposure was determined based on whether modeled storm-associated peak sustained surface wind was $\ge34$ kts at the county’s population mean center.
 This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale.

### 2020-09-19 ehp_revision/writing/ehp_discussion.tex
Work some more on revisions

@@
 correlation was substantially lower when considering only tropical cyclones
-with cumulative local rainfall of $\ge75$ mm (Supplemental Table 1).} When
+with cumulative local rainfall of $\ge75$ mm (Supplemental Table 2).} When
 rainfall data is used to create binary exposure classifications, this

### 2020-09-19 ehp_revision/writing/ehp_introduction.tex
Work more on response
Work some more on revisions

--- 2020-07-10 ehp_revision/writing/ehp_introduction.tex
@@
 cyclone tend to be distributed around its center non-symmetrically.  \textbf{In
-the Northern Hemisphere,} extreme winds are more common to the track's right,
-where counterclockwise cyclonic winds move in concert with the tropical
-cyclone's forward motion~\parencite{halverson2015}, and almost all of the fatal
-tornadoes associated with \ac{US} tropical cyclones between~1995 and~2009
-occurred to the right of the tropical cyclone's track~\parencite{moore2012}.
-Rain, conversely, is often heaviest to the left of the track, especially when
-the tropical cyclone interacts with other weather
+the Northern Hemisphere, cyclonic winds are counterclockwise,} so extreme winds
+are more common to the track's right where cyclonic winds move in concert with
+the tropical cyclone's forward motion~\parencite{halverson2015}, and almost all
+of the fatal tornadoes associated with \ac{US} tropical cyclones between~1995
+and~2009 occurred to the right of the tropical cyclone's
+track~\parencite{moore2012}.  Rain, conversely, is often heaviest to the left
+of the track, especially when the tropical cyclone interacts with other weather
 systems~\parencite{atallah2003, atallah2007, zhu2013variations} or undergoes an

### 2020-09-19 ehp_revision/writing/ehp_results.tex
Work some more on revisions

@@
 tropical cyclones with cumulative local rainfall of $\ge75$ mm (Supplemental
-Table 1).} However, even in these cases, it was rare for a storm to be
-classified differently (under the precipitation threshold of
-75~\si{\milli\metre} we used for binary exposure classification for later
-analysis) based on the source of precipitation data. Horizontal and vertical
-lines in each small plot in Figure~\ref{fig:raincomparison} show the threshold
-of 75~\si{\milli\metre}, so storms in the lower left and upper right quadrants
-would be classified the same (``exposed'' or ``unexposed'') regardless of the
-precipitation data source, while storms in the upper left and lower right
-quadrants would be classified differently. Such cases were rare.
+Table 2).} However, it was rare for a storm to be classified differently (under
+the precipitation threshold of 75~\si{\milli\metre} we used for binary exposure
+classification for later analysis) based on the source of precipitation data.
+Horizontal and vertical lines in each small plot in
+Figure~\ref{fig:raincomparison} show the threshold of 75~\si{\milli\metre}, so
+storms in the lower left and upper right quadrants would be classified the same
+(``exposed'' or ``unexposed'') regardless of the precipitation data source,
+while storms in the upper left and lower right quadrants would be classified
+differently. Such cases were rare.
 
@@
 (Figure~\ref{fig:jaccard}; \textbf{for the most extensive of these, storms for
-which 250 of more counties were exposed based on at least one metric, Tables
-S3--S6 provide underlying numbers comparing exposure assessments for the
-distance-based proxy to each hazard-based metric}). In \textbf{Figure 7}, each row
+which 200 of more counties were exposed based on at least one metric, Tables
+S3--S6 provide underlying numbers comparing exposure assessments between the
+distance-based proxy and each hazard-based metric}). In \textbf{Figure 7}, each row
 provides results for one tropical cyclone, and each box in that row shows the

### 2020-09-19 ehp_revision/R/misclassification_tables.R

@@ -153,7 +153,7 @@ misclass_rain %>%
          Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
          calculated as the value in the second column divided by the sum of numbers in the 
          second through fourth columns.  Storms are ordered based on the number of counties 
-         assessed as exposed to at least one of these two exposure metrics.).", 
+         assessed as exposed to at least one of these two exposure metrics.", 
          label = "tab:misclassrain") %>% 
   print(include.rownames = FALSE, booktabs = TRUE, 
          only.contents = FALSE,
@@ -194,7 +194,7 @@ misclass_wind %>%
          Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
          calculated as the value in the second column divided by the sum of numbers in the 
          second through fourth columns.  Storms are ordered based on the number of counties 
-         assessed as exposed to at least one of these two exposure metrics.).", 
+         assessed as exposed to at least one of these two exposure metrics.", 
          label = "tab:misclasswind") %>% 
   print(include.rownames = FALSE, booktabs = TRUE, 
         only.contents = FALSE,
@@ -236,7 +236,7 @@ misclass_flood %>%
          Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
          calculated as the value in the second column divided by the sum of numbers in the 
          second through fourth columns.  Storms are ordered based on the number of counties 
-         assessed as exposed to at least one of these two exposure metrics.).", 
+         assessed as exposed to at least one of these two exposure metrics.", 
          label = "tab:misclassflood") %>% 
   print(include.rownames = FALSE, booktabs = TRUE, 
         only.contents = FALSE,
@@ -277,7 +277,7 @@ misclass_tornado %>%
          Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
          calculated as the value in the second column divided by the sum of numbers in the 
          second through fourth columns.  Storms are ordered based on the number of counties 
-         assessed as exposed to at least one of these two exposure metrics.).", 
+         assessed as exposed to at least one of these two exposure metrics.", 
          label = "tab:misclasstornado") %>% 
   print(include.rownames = FALSE, booktabs = TRUE, 
         only.contents = FALSE,

### 2020-09-19 ehp_revision/response_to_reviewers.Rmd
Work some more on response

@@
 have geopolitical areas that are fairly comparable to US counties, like
-municipalities in Mexico, districts in India. There would be a few challenges,
+municipalities in Mexico and districts in India. There would be a few challenges,
 however, in extending the exposure dataset for the hazards that we present in
@@
 We also used the word "Hits per county per decade" in the legend for Figure 5 in
-the main text (reproduced in this response document as Figure \@ref(fig:fig5))
-and Figure S3 of the Supplemental Material. We have revised the wording in these
-figures to "Exposures per county per decade".
+the main text (reproduced at the end of this response document as Figure
+\@ref(fig:fig5)) and Figure S3 of the Supplemental Material. We have revised the
+wording in these figures to "Exposures per county per decade".
 
@@
 
-Second, in cases where it is clear that a mixture approach is necessary, these
+Further, in cases where it is clear that a mixture approach is necessary, these
 results can help in understanding what challenges might arise in statistical
@@
 
-> "Finally, we investigated how well exposure assessment agreed across these
+> **Text in revised manuscript (relevant added text in bold):** 
+"Finally, we investigated how well exposure assessment agreed across these
 metrics. **These results can help epidemiologists answer several key
@@
 Supplemental Material (Table S2 of the manuscript, reproduced as Table
-\@ref(tab:highprecipcorr) of this response document).
+\@ref(tab:highprecipcorr) at the end of this response document).
 
@@
 versions of Figure 4 of the manuscript are reproduced in Figure \@ref(fig:fig4)
-of this response.
+at the end of this response.
 
@@
 moved northward along the Eastern Seaboard. While there will be questions in
-some case about whether counties in the Mid-Atlantic and New England were
+some cases about whether counties in the Mid-Atlantic and New England were
 exposed to this storm, counties in Texas, Louisiana, and other states in the
@@
 \textbf{Distance} & Distance-based exposure was determined based on whether the storm track came within 100 km of county population mean center.
-This threshold has been used in prior research as a proxy for exposure to hazards from the storm (e.g., @grabich2015measuring).
-Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km [@mallin2006; @quiring2011variations], and dangerous winds can extend beyond these maximum winds.
-One study assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties, which resulted in an average distance radius of 120 km on either side of the storm track [@czajkowski2011].
-Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research [@grabich2015measuring; @grabich2015; @currie2013].
+This threshold has been used in prior research as a proxy for exposure to hazards from the storm (e.g., Grabich et al.
+2015a).
+Tropical cyclones vary dramatically in size: US tropical cyclones have been observed with radii to maximum winds as small as 20 km and as large as 200 km (Mallin and Corbett 2006; Quiring et al.
+2011), and dangerous winds can extend beyond these maximum winds.
+One study assessed county-level risk and exposure based on a three-tiered definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary counties, and tertiary counties adjacent to secondary counties, which resulted in an average distance radius of 120 km on either side of the storm track (Czajkowski et al.
+2011).
+Other distance thresholds that could be considered include 60 km and 30 km, both of which have been used in previous research (Grabich et al.
+2015a; Grabich et al.
+2015b; Currie and Rossin-Slater 2013).
 However, based on the results presented in the main manuscript, hazard-based metrics should often be used directly rather than a distance-based proxy.\\
@@
 \textbf{Rain} & Rain-based exposure was determined based on whether the county had cumulative rainfall of $\ge75$ mm over the period from two days before to one day after the storm’s closest approach and the storm track came within 500 km of the county population mean center.
-One recent study has highlighted that a two-year rainfall value (which is the median annual maximum rainfall) for a location can provide a useful threshold in identifying rainfall events with potential for societal impacts [@bosma2020intuitive].
+One recent study has highlighted that a two-year rainfall value (which is the median annual maximum rainfall) for a location can provide a useful threshold in identifying rainfall events with potential for societal impacts (Bosma et al.
+2020).
 For some of the more northern, inland communities included in our study area, the two-year rainfall value for two-, three- and four-day windows is in the 65--85 mm range.
-For example, Pittsburgh, PA, has two-year rainfall values of 69 mm for a two-day window, 74 mm for a three-day window, and 78 mm for a four-day window [@noaaatlas], consistent with the 75 mm threshold we selected for exposure classification in this paper.
+For example, Pittsburgh, PA, has two-year rainfall values of 69 mm for a two-day window, 74 mm for a three-day window, and 78 mm for a four-day window (US NOAA 2020), consistent with the 75 mm threshold we selected for exposure classification in this paper.
 However, other thresholds, particularly higher thresholds, would be reasonable in some cases.
 For example, the two-year rainfall values tend to be much higher in counties of the study area that are further south and close to the coast.
-The two-year rainfall value for Miami for a three-day window, for example is 180 mm [@noaaatlas].
+The two-year rainfall value for Miami for a three-day window, for example is 180 mm (US NOAA 2020).
 A variety of definitions have been used previously to identify both extreme or heavy rain (whether associated with a tropical cyclone or not) and tropical cyclone--associated rain.
-In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall [@zhou2017spatial] and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---\parfillskip=0pt\tabularnewline
-
-&in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location [@feldmann2019estimation].
-Other definitions of extreme rain events---including but not limited to tropical cyclone--associated rainfall---are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours [@schumacher2006characteristics].
-Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]) [@schumacher2006characteristics; @schumacher2005organization; @stevenson201410].\\
+In defining precipitation associated with tropical cyclones, studies have used thresholds of 12.5 mm per day as a metric of regions of ``moderately heavy'' rainfall (Zhou and Matyas 2017) and, as a lower threshold, a lower limit of 10 mm of total storm precipitation---\parfillskip=0pt\tabularnewline
+
+&in conjunction with proximity to the storm's center---in identifying tropical cyclone precipitation events at a location (Feldmann et al.
+2019).
+Other definitions of extreme rain events---including but not limited to tropical cyclone--associated rainfall---are higher than the threshold we use here---for example, one paper defined extreme rain events as cases in which a gauge reported 125 mm or more of rain in 24 hours (Schumacher and Johnson 2006).
+Studies have also used definitions that are relative to the norms for a given location (e.g., 24 hour rainfall totals over the 50-year return value for the location, which the part of the US east of the Rocky Mountains range from 3.5 in [89 mm] to 13 in [330 mm]) (Schumacher and Johnson 2006; Schumacher and Johnson 2005; Stevenson and Schumacher 2014).\\
 
@@
 This threshold is being applied to local winds for each county, and it represents the threshold for gale-force winds on the Beaufort wind scale.
-This limit is used as the outer limit in measuring storm size through the US National Hurricane Center's wind radii for tropical cyclone forecasts [@cangialosi2016examination].
+This limit is used as the outer limit in measuring storm size through the US National Hurricane Center's wind radii for tropical cyclone forecasts (Cangialosi and Landsea 2016).
 Other thresholds could be selected based on other points on the Beaufort scale---for example, $\ge48$ kts for capturing storm-force winds or $\ge64$ kts for capturing hurricane-force winds.
@@
 requested. The original and revised versions of Figure 3 of the manuscript are
-reproduced in this response document as Figure \@ref(fig:fig3).
+reproduced at the end of this response document as Figure \@ref(fig:fig3).
 
@@
 value ($\sim75\%$), as requested. The original and revised versions of this
-paper are reproduced in this response document as Figure \@ref(fig:fig3).
+paper are reproduced at the end of this response document as Figure \@ref(fig:fig3).
 
@@
 To help illustrate, in Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2)
-of this response document, we have reproduced the current version of Figure 5 as
-it would appear to someone with three common types of color vision deficiencies:
-deuteranomaly (defective green cone cells, the most common type), as well as
-protanopia (defective red cone cells) and tritanopia (defective blue cone
-cells). We have also shown how the Figure would appear in grayscale in Figure
-\@ref(fig:fig5check2) (right). We have included similar versions of other
+at the end of this response document, we have reproduced the current version of
+Figure 5 as it would appear to someone with three common types of color vision
+deficiencies: deuteranomaly (defective green cone cells, the most common type),
+as well as protanopia (defective red cone cells) and tritanopia (defective blue
+cone cells). We have also shown how the Figure would appear in grayscale in
+Figure \@ref(fig:fig5check2) (right). We have included similar versions of other
 figures from the manuscript in response to later requests from EHP in this
@@
 
-As we note in our response to EHP Request 2, we used colormaps for all
-figures that are designed to be perceptible both for those with common types of
-color vision deficiency and when printed in grayscale. We have increased the
-size of the circles in Figure 3 of the revised manuscript (reproduced in Figure
-\@ref(fig:fig3) of this response) so that the colors are clearer, as well as
-truncated the x-axis as requested in EHP Request 1. Also, we show in Figures
-\@ref(fig:fig3check1) and \@ref(fig:fig3check2) of this response how the revised
-Figure 3 would look with three common types of color vision deficiency as well
-as when printed in black and white. The gradient is clear when this figure is
-printed in black and white.
+As we note in our response to EHP Request 2, we used colormaps for all figures
+that are designed to be perceptible both for those with common types of color
+vision deficiency and when printed in grayscale. We have increased the size of
+the circles in Figure 3 of the revised manuscript (reproduced in Figure
+\@ref(fig:fig3) at the end of this response) so that the colors are clearer, as
+well as truncated the x-axis as requested in EHP Request 1. Also, we show in
+Figures \@ref(fig:fig3check1) and \@ref(fig:fig3check2) of this response how the
+revised Figure 3 would look with three common types of color vision deficiency
+as well as when printed in black and white. The gradient is clear when this
+figure is printed in black and white.
 
@@
 
-We have reproduced both the original and revised versions of Figure 4 as Figure
-\@ref(fig:fig4) of this response.
+We have reproduced both the original and revised versions of Figure 4 in the
+manuscript as Figure \@ref(fig:fig4) at the end of this response.
 
@@
 printed in black and white, and if not, use a “color-blind friendly” palette if
-possible. If this is not feasible, we can request and exemption from
+possible. If this is not feasible, we can request an exemption from
 accessibility requirements, but we try to do this only when alternatives  would
@@
 and \@ref(fig:fig5check2) at the end of this response show what Figure 5 of the
-main text would look like under three common types of color vision deficiency
+manuscript would look like under three common types of color vision deficiency
 and when printed in black and white. Figures \@ref(fig:fig6check1) and
 \@ref(fig:fig6check2) at the end of this response show the same thing for Figure
-6 of the main text.
-
-For Figure 6 of the main text, for some types of color balance, it could be
-unclear which of the two color legends pairs with which part of the graph. We
-have therefore added a note to the figure legend to clarify:
+6 of the manuscript.
+
+For Figure 6 of the manuscript, for some types of color vision deficiency, it
+could be unclear which of the two color legends pairs with which part of the
+graph. We have therefore added a note to the figure legend to clarify:
 

### 2020-09-19 ehp_revision/response_to_reviewers.Rmd
Continue working on revisions

@@
 
-Broadly, we have:
-
-1. Added discussion on the question of county-level versus individual-level
-data. 
+The larger changes we have made include:
+
+1. Added discussion on the challenges and limitations of using aggregated, 
+county-level exposure assessment.
 2. Added explanations for how our comparisons among different storm hazards
-can help epidemiologists in designing studies and statistical analysis.
+can help epidemiologists in designing studies and planning statistical analysis.
+3. Added a supplemental table explaining the choice of thresholds for binary 
+exposure classification.
 3. Added a small analysis of the correlation in rainfall metrics specifically
@@
 in continuous measures of rainfall during extreme conditions. 
+4. Added supplemental tables that provide more details on patterns of exposure
+misclassification when comparing each hazard-based metric to a distance-based
+proxy metric.
 4. Changed labeling, axis range, and other presentation details for some
@@
 **Response:**
+
+\hl{Roger and Meilin, could you take a look at this response and let me know if
+you have any comments of suggestions to add?}
 
@@
 **Response:**
+
+\hl{Andrea, Mohammad, Seth, Steven, Bill, could you take a look at this response 
+and let me know if you have any comments of suggestions to add?}
 
@@
 **Response:**
+
+\hl{Dirk, could you take a look at this response and let me know if you have
+any comments of suggestions to add? In particular, I don't have a lot of experience
+working with cloud-based platforms with Jupyter Notebooks. It sounds like that's 
+how the workspace for this DesignSafe platform works. My best guess is that you'd
+be able to download and use R packages in that type of workspace, including 
+a package hosted on a drat repository. Do you have any experience or insights on
+installing packages from different repositories while working on a Jupyter Notebook
+through an online platform?}
 
@@
 
+\hl{Roger and Meilin, could you take a look at this response and let me know if
+you have any comments of suggestions to add?}
+
 We have included a measurement of agreement between different hazard-based
@@
 
+\hl{Andrea, could you take a look at this response and let me know if you have
+any comments of suggestions to add?}
+
 You are right, the opposite would be true in the Southern hemisphere. This
@@
 **Response:**
+
+\hl{Andrea, could you take a look at this response and let me know if you have
+any comments of suggestions to add?}
 
@@
 
+\hl{Andrea, could you take a look at this response and let me know if you have
+any comments of suggestions to add?}
+
 This database is NOAA's official database on tornado events in the United States
@@
 **Response:**
+
+\hl{Andrea, Mohammad, Seth, Steven, and Bill, could you take a look at this
+response and let me know if you have any comments of suggestions to add?}
 

### 2020-09-22 ehp_revision/response_to_reviewers.Rmd
Finalize revisions for submission

--- 2020-09-19 ehp_revision/response_to_reviewers.Rmd
@@
 allows for easy integration with health outcome data aggregated at the county
+level, and disaster response decisions are often made at a county or state
 level. Such data is often used for disaster epidemiology, as aggregated data may
@@
 "**This dataset is limited to the contiguous US, and while expansion to global
-coverage would be useful and feasible, there would be some challenges. For
-precipitation data, the re-analysis product used here (NLDAS-2) only covers the
-contiguous US, although other re-analysis products, as well as other types of
-precipitation datasets, have global coverage [@sun2018review]. The wind data are
-generated based on a model that is currently US-focused [@stormwindmodel], but
-could be extended to other areas, although this would require adding new
-land/sea masks within the associated software, as well as accounting for
-differences across storm basins in wind averaging periods
-[@harper2010guidelines] and in the direction of cyclonic winds in the Northern
-versus Southern Hemisphere. Further, since the core of the wind model was
-developed based on data from Atlantic-basin storms [@willoughby2006parametric],
-an extension to other areas should include separate validation and calibration
-to ensure it performs appropriately in those settings. Alternatively, other wind
+coverage would be useful and feasible, there would be some challenges. The key
+challenge would be for tornado and flood events. For these events, the data
+described in this paper drew on a US-focused storm events database, and so
+international extension of data on these hazards would require access to similar
+databases covering other countries. For precipitation data, while the
+re-analysis product used here (NLDAS-2) only covers the contiguous US, other
+re-analysis products, as well as other types of precipitation datasets, have
+global coverage [@sun2018review], including the NASA Global Land Data
+Assimilation System Version 2 (GLDAS-2; @rodell2004global), the Global
+Precipitation Measurement Mission (GPM), and the Tropical Rainfall Measuring
+Mission (TRMM). The wind data are generated based on a model that is currently
+US-focused [@stormwindmodel], but could be extended to other areas, although
+this would require adding new land/sea masks within the associated software, as
+well as accounting for differences across storm basins in wind averaging periods
+[@harper2010guidelines]. Further, since the core of the wind model was developed
+based on data from Atlantic-basin storms [@willoughby2006parametric], an
+extension to other areas should include separate validation and calibration to
+ensure it performs appropriately in those settings. Alternatively, other wind
 field modeling software is available that provides a global coverage, like
 Geoscience Australia's Tropical Cyclone Risk Model
-(http://geoscienceaustralia.github.io/tcrm/). For tornado and flood events, the
-data described in this paper drew on a US-focused storm events database, and so
-international extension of data on these hazards would require access to similar
-databases covering other countries. Finally, the relevant geopolitical
+(http://geoscienceaustralia.github.io/tcrm/). Finally, the relevant geopolitical
 boundaries to use for aggregation would vary by country (e.g., municipalities in
@@
 
-It turns out that the data we present should already be accessible to use on the
-DesignSafe platform, because we have shared the data as an R package.
+This is a very good, and very forward-looking suggestion. And it turns out that
+the data we present should already be accessible to use on the DesignSafe
+platform because we have shared the data as an R package.
 
@@
 programming languages, including R (through Jupyter Notebooks)
-[@rathje2017designsafe]. Therefore, DesignSafe users already have access to any
-data shared through an R package, including the data we present in this
-manuscript, without the data needing to be separately loaded and published
-through DesignSafe's Data Depot.
-
-This is great, because it might otherwise be difficult to try to maintain copies
-of the dataset in different repositories (i.e., one copy as an R package and
-another as a DesignSafe dataset), and there could be confusion about things like
+[@rathje2017designsafe]. Therefore, and thanks to the R package mechanism and
+its integration into the R environment, DesignSafe users already have access to
+any data shared through an R package.
+
+This is a very useful feature because it might otherwise be difficult to try to
+maintain copies of the dataset in different repositories (i.e., one copy as an R
+package and another as a DesignSafe dataset), and there could be confusion about
 the equivalence of the two copies. R packages are citable through a standard
@@
 each update to the R package (we have been updating several hazards as new data
-becomes available). DesignSafe's interface with R avoids this problem by
+becomes available). DesignSafe’s interface with R avoids this problem by
 allowing its users direct access to our package and its data through that path,
@@
 We have added a note in the manuscript to let readers know that, because the
-data is shared through an R package, and because DesignSafe's Discovery
+data is shared through an R package, and because DesignSafe’s Discovery
 Workspace allows users to work in R, the data can be fully accessed and used for
@@
 considering only tropical cyclones with cumulative local rainfall of $\ge75$ mm
-(Supplemental Table 2)**."
+(Table S2)**."
 
@@
 some counties the correlation was substantially lower when considering only
-tropical cyclones with cumulative local rainfall of $\ge75$ mm (Supplemental
-Table 2).**"
+tropical cyclones with cumulative local rainfall of $\ge75$ mm (Table S2).**"
 

### 2020-09-22 ehp_revision/writing/ehp_discussion.tex
Finalize revisions for submission

@@
 global coverage would be useful and feasible, there would be some challenges.
-For precipitation data, the re-analysis product used here (NLDAS-2) only covers
-the contiguous US, although other re-analysis products, as well as other types
-of precipitation datasets, have global coverage \parencite{sun2018review}. The
-wind data are generated based on a model that is currently US-focused
-\parencite{stormwindmodel}, but could be extended to other areas, although this
-would require adding new land/sea masks within the associated software, as well
-as accounting for differences across storm basins in wind averaging periods
-\parencite{harper2010guidelines} and in the direction of cyclonic winds in the
-Northern versus Southern Hemisphere. Further, since the core of the wind model
-was developed based on data from Atlantic-basin storms
-\parencite{willoughby2006parametric}, an extension to other areas should
-include separate validation and calibration to ensure it performs appropriately
-in those settings. Alternatively, other wind field modeling software is
-available that provides a global coverage, like Geoscience Australia's Tropical
-Cyclone Risk Model (http://geoscienceaustralia.github.io/tcrm/). For tornado
-and flood events, the data described in this paper drew on a US-focused storm
-events database, and so international extension of data on these hazards would
-require access to similar databases covering other countries. Finally, the
-relevant geopolitical boundaries to use for aggregation would vary by country
-(e.g., municipalities in Mexico, districts in India).}
+The key challenge would be for tornado and flood events. For these events, the
+data described in this paper drew on a US-focused storm events database, and so
+international extension of data on these hazards would require access to
+similar databases covering other countries.  For precipitation data, while the
+re-analysis product used here (NLDAS-2) only covers the contiguous US, other
+re-analysis products, as well as other types of precipitation datasets, have
+global coverage \parencite{sun2018review}. The wind data are generated based on
+a model that is currently US-focused \parencite{stormwindmodel}, but could be
+extended to other areas, although this would require adding new land/sea masks
+within the associated software, as well as accounting for differences across
+storm basins in wind averaging periods \parencite{harper2010guidelines}.
+Further, since the core of the wind model was developed based on data from
+Atlantic-basin storms \parencite{willoughby2006parametric}, an extension to
+other areas should include separate validation and calibration to ensure it
+performs appropriately in those settings. Alternatively, other wind field
+modeling software is available that provides a global coverage, like Geoscience
+Australia's Tropical Cyclone Risk Model
+(http://geoscienceaustralia.github.io/tcrm/). For tornado and flood events, the
+data described in this paper drew on a US-focused storm events database, and so
+international extension of data on these hazards would require access to
+similar databases covering other countries. Finally, the relevant geopolitical
+boundaries to use for aggregation would vary by country (e.g., municipalities
+in Mexico, districts in India).}
 
@@
 scale allows for easy integration with health outcome data aggregated at the
-county level. Such data is often used for disaster epidemiology, as aggregated
+county level, and disaster response decisions are often made at a county or
+state level. Such data is often used for disaster epidemiology, as aggregated
 data may be easier to access than individual-level data, especially at a scale
